### 2023-09-16 manuscript.Rmd
the first draft.

---
title: "Quantifying Stability and Change in Personal Culture Using Panel Data"
author:
- 
output:
  pdf_document:
    keep_tex: true
bibliography: kileybib.bib
csl: asa.csl
abstract: "Recent work has produced conflicting interpretations on the question of whether adults undergo durable changes in their personal culture over time. This paper asserts that this divergence is due more to the approach adopted by researchers –-a tournament of models-– than their findings, which are highly consonant. We reanalyze the data used previously in this debate with an approach that quantifies the processes of cultural difference, rather than one that detects the presence or absence of change. In doing so, we compare how much cultural difference is explained by intrapersonal change and interpersonal differences. Our results harmonize recent findings by showing that, while most measures of personal culture show change over time, the variance explained by intrapersonal change is typically small compared to interpersonal differences at baseline. We then exemplify how quantifying these processes offer new ways to address theoretical questions of when and how the relative importance of intrapersonal change and interpersonal differences shifts."
fontsize: 11pt
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F, message = F, warning = F, fig.align = "center")
```

# Introduction

A recurring focus in the sociology of culture is the question of whether people, as they move through their lives and transition between social roles, change their personal culture –- the declarative and nondeclarative attitudes, beliefs, values, and dispositions that manifest at the individual level [@lizardo2017; @kiley2020; @lersch2023].

This seemingly simple question underlies a number of central debates. For example, pragmatist theories of action claim that changes in social environments cause people to adapt their views and make new meanings [@swidler2001; @gross2009], while Bourdiesian practice theories argue that the “past conditions of production” leave a lasting mark on peoples’ personal culture throughout their lives [@bourdieu1990]. Similarly, models of social influence require people to adapt their culture in the face of new information [@vaisey2010a], while models of aggregate social and cultural change that place emphasis on cohort effects require them to be open to change while young but fairly resistant to change as they age [@ryder1965]. Finally, life course theories posit slower but still important changes over time as people advance through important transitions in their lives [@elder2016; @bardi2011].

Because all these theoretical perspectives have some empirical support, adjudicating when and how they apply is important for further understanding how early experiences and contemporary environments shape personal culture over the life course. That being said, scholars have struggled to reach a consensus on the relative importance of change during adulthood for explaining cultural differences. Over the span of a few years, most apparent changes in people’s responses to personal culture items appear to be transitory, with little evidence that adults make persistent changes of personal culture [@kiley2020; @vaisey2021]. When looked at over longer time horizons, however, there is at least some evidence that adults make persistent cultural changes, especially as they move through different social roles across the life course [@lersch2023].

These different conclusions are partly due to the fact that researchers decided to adjudicate the different perspectives by taking a “tournament of models” approach [@lersch2023: p. 228]. In this approach, researchers have formalized models of responses over time as demonstrating either “change” or “no change” and compared which model fits observed data better, functionally asking whether we ever observe within-person change in personal culture.

We argue, however, that asking only *whether* people change obscures what are actually highly compatible results. Whether change “really happened” in a given data set depends not only on sample populations, statistical methods and power, and item measurement, but crucially also on definitions of what counts as “change,” thus running the risk of getting caught up in unresolvable debates about related decisions. Moreover, it narrows the focus on particular inquiries such as, “do we see change in this measure?” Such inquiries are ill-suited to address what’s at the core of the theoretical debate. There is enough evidence to assume that neither socialization nor the environment alone can explain cultural differences. What’s left to answer is under what conditions, across which groups, and in what domains do we observe differences in the relative importance of intrapersonal change and interpersonal differences. This focus is necessarily comparative and requires an approach that quantifies forms of cultural change rather than calling victory for the one or the other perspective.

In this paper, we try to overcome previous debates and move away from the “tournament of models” approach toward a quantification of the processes that lead to differences in personal culture. Drawing on the seven panel data sets analyzed in these debates, we use established methods to quantify the amount of variance in responses that is due to between-people differences in baselines and the amount of variance in responses that is due to linear within-person change over time. This measure gives us a sense of the importance of intrapersonal change over the course of the panel relative to the differences that exist between people when they start. We also explore features of questions and populations that explain differences in within-person change over time.

The pattern of results help clarify and, hopefully, transcend previous debates. First, we observe similar patterns across the data sets that have produced conflicting interpretations. In all datasets we see similar rates of individual-level change over time across a broad range of measures of personal culture. But quantifying these processes also reveals that intrapersonal change plays a substantially smaller role in explaining the variance in peoples’ personal culture than between-person differences in their baselines. Whether this amount of intrapersonal change is “meaningful” ultimately remains a theoretical question reflected in prior expectations concerning the question, population, and time. Nevertheless, the results suggest that the intrapersonal change observed (in some cases over more than 10 years) does not seem to play a substantial role in explaining personal cultural differences between adults. While adding to existing debates, the quantification we propose extends them in new ways that enables scholars to more clearly adjudicate theoretical mechanisms at play. We provide a series of illustrative examples of the utility of this approach for addressing different theoretical questions.

# Background

Recent debates about  whether adults undergo intrapersonal cultural change emerged in part because sociological models of cultural change at the aggregate level tend to implicitly invoke one of two models of individual behavior. The first, what Kiley and Vaisey [-@kiley2020] call a “Settled Dispositions Model” (SDM), assumes that peoples’ personal culture is relatively fixed by the time they are adults. While they might make temporary changes in their declarative culture as their social environments provide new considerations, this model assumes that they return to a settled baseline over a (relatively) short time frame. This model of behavior is frequently invoked in models of cultural change that suggest people are strongly imprinted by the “past conditions of production” of socialization, such as Bourdieusian practice theories [@bourdieu1990], cohort replacement theories of aggregate change [@mannheim1952; @ryder1965], and control theories in social psychology [@smith-lovin1988; @robinson2007].

The second model summarized by Kiley and Vaisey, an “Active Updating Model” (AUM), posits that people continually update their personal culture as they move through life. This model suggests people make new cultural meanings, and therefore durably change their personal culture, as they encounter new social environments, discourses, and information [@gross2009; @swidler2001]. This model underlies models of cultural diffusion [@christakis2010], attitude alignment [@dellaposta2020], and polarization [@bail2018], among others. It is also implicit in most studies of whether specific experiences, changes in social roles, or political events, affect personal culture [@gelman2021; @visser2004; @slothuus2021].

There is no reason to think that only one of these models is “correct” in all times and places. A population observed over some time frame contains a mix of people who are changing and people who are not (or some people making very small or imperceptible changes in culture and others making perceptible changes). Instead, different perspectives argue that each of these ideal-typical models is more operative at different times, for different people, and for different elements of personal culture. For example, adolescence and early adulthood, typically viewed  as a “formative period” for personal cultural development, is likely characterized by higher rates of “active updating,” while middle age and later life are potentially characterized more by “settled dispositions” [@alwin1991; @krosnick1989; @eaton2009]. Similarly, highly salient issues, such as views around gay rights in the 2010s; issues that see substantial elite realignments, such as views around the Vietnam conflict in the 1970s [@zaller1992]; and novel issues of public opinion, such as views around vaccines during the Covid-19 pandemic, might be characterized by active updating, while low-salience, established issues are characterized by stability.

Comparing which of these models better fit a broad range of questions from the General Social Survey’s rotating panels, Kiley and Vaisey [-@kiley2020] found limited evidence of durable change. While the AUM was preferred for the majority of items, the total amount of durable change detected on these items was small. A substantial minority of questions (39 percent) favored the SDM, meaning they showed no evidence of durable change. Questions with more evidence of durable change included highly salient issues like gay marriage and “public” issues like partisan identification and religious service attendance. There was also more evidence of durable change among early adults (people ages 18-30) than the rest of the population. Overall, the researchers concluded that “results ultimately suggest that real, persistent attitude change is an uncommon phenomenon among adults'' [@kiley2020: p. 500; see also @vaisey2021]. This lack of durable change is consistent with other findings that cohort replacement plays a larger role than period effects in explaining differences in personal culture [@vaisey2016].

However, the claim that change is a relatively infrequent phenomenon has been difficult to square with research identifying individual change across a number of cultural dimensions, such as morality [@brocic2021], trust [@mewes2021], and concerns about immigration [@kratz2021]. Similarly, there are longitudinal studies showing that cues from political elites can change a person’s position on specific policy issues [@slothuus2021; @zaller1992] and that changes in a person’s close contacts and acquaintances can change their attitudes on group-related politics [@gelman2021; @dellaposta2018]. Given how often we observe people change their personal culture, it is hard to accept that adults do not change.

Drawing on these findings, Lersch challenged the SDM and AUM as a “needless dichotomy” [-@lersch2023] and proposed a “Life Course Adaption Model” (LCAM), which draws on the life course perspective and models personal culture over the course of a panel as a linear trajectory that varies for each respondent. Lersch’s work  addresses two shortcomings of the AUM and SDM comparison. First, the AUM as outlined by Kiley and Vaisey assumes that change can only occur as a Markovian process, in which responses to personal culture items at time $t$ are a function only of responses at time $t-1$. But early life experiences might indirectly shape peoples’ personal culture as they move into new social roles or environments, even if they do not affect personal culture right away. For example, experiences in childhood might not affect views about family structure until much later in life when people create their own families. The LCAM allows responses at earlier points than $t-1$ to influence responses at time $t$.

Second, Kiley and Vaisey only adjudicated the AUM and SDM on three-wave panel data covering four years, which might not be a long enough time span to adjudicate the two models they outline. In contrast, Lersch tests the LCAM against the AUM and SDM on panels covering a much longer time frame (between 3 and 36 years) with more waves (between 3 and 18 waves), allowing for a greater possibility of detecting durable change.

In his tournament of models, Lersch finds broad support for the LCAM in data from five countries, with the LCAM favored over the AUM and SDM on 297 of 428 questions he analyzed and the SDM favored over the other two models on 112 items, with the remaining questions producing inconclusive results. He concludes that “new experiences over the life course, which expose individuals to new social environments, including novel interactions, and different sets of resources, can persistently move individuals’ personal culture in novel directions” [@lersch2023: p. 243-244].

However, despite differences in how these researchers interpret their findings, the empirical results are not far apart. Kiley and Vaisey [-@kiley2020] found that the majority of questions they explored favored the AUM, meaning they showed some evidence of durable change among adults, a finding broadly compatible with Lersch’s results. At the same time, while Lersch [-@lersch2023] found that the LCAM was preferred in the majority of cases he examined, durable change does not account for a large proportion of variance explained on these items. On average, people changed only about .07 standard deviations over 10 years. Again, this is consistent with Kiley and Vaisey’s assertion that durable change is often small. And a quarter of the items Lersch studied still preferred the SDM, suggesting that even with looser assumptions about the process of change, more cases, and more waves, many questions show no evidence of durable change. Relatedly, in most studies purporting to find evidence of durable change in peoples’ personal culture, the change is often quite small. For example, Brocic and Miles [-@brocic2021] find that completing graduate degrees in humanities, arts, and social sciences shift peoples’ moral relativism only about .2 standard deviations on average compared to people with no college degree, and this is the largest effect they identify. And studies of aggregate change show that, even on items where cohort effects dominate, some people do change over time [@vaisey2016]. In other words, despite different interpretations, the results of previous work are in many ways highly consonant.

## Quantification, Not Adjudication

The shortcoming of current debates is that they focus on whether people *ever* change, a question that is not theoretically interesting since the answer is almost always “yes.”  As noted previously, if we observe enough people for long enough, we will likely observe some of them making durable changes. A lack of evidence for change in one question might simply indicate that the question was not asked at a clear enough level of resolution, for a long enough time frame, or to enough people to find evidence of change. Conversely, finding evidence of change tells us little about how much change is present in a population and how meaningful that change is for social differences.

A more theoretically productive approach is to start with a model that assumes that, in the window of a panel survey, some people might remain stable, some people might change a little, and some people might undergo substantial changes in personal culture. Lersch’s LCAM formally does this in modeling each person as following their own trajectory over time. But rather than adjudicating whether this model fits the data better than one in which people never change –-asking the question of whether change ever exists-– researchers should quantify the amount that people change over time and compare that change to the differences that exist between people at the beginning of the panel. In other words, rather than ask, “do people change?” we should ask, “what are the relative contributions of change and initial differences toward explaining differences in personal culture?”

In this approach, there are two numbers that each reflect a combination of relevant theoretical processes. First, interpersonal differences at the start of the survey, what Lersch calls “early imprinting,” reflects the accumulated experiences of people prior to entering the panel survey that shape how they answer questions. While this is typically associated with a “formative period” or a “settled disposition” (hence the name early imprinting), it could also reflect transformative experiences that happen at any age, so long as they occurred before the start of the panel and leave a lasting effect on how people answer questions. For example, among people already retired upon entering the panel, their “early imprinting” might reflect this life course transition. To some extent, then, this number also reflects variation across people in relevant social roles or statuses at the start of the panel that are important in shaping dispositions.

The second, the amount of intrapersonal linear change over the course of the survey, what Lersch calls “persistent change” or “adaption” and Kiley and Vaisey call “active updating,” reflects durable changes in personal culture over time. Lersch suggests that these changes are brought on by social triggers such as moving into a new environment, adopting new social roles, or broader changes in the social structure.

A third number, often called “residual variance” or “fluctuation,” necessarily accounts for the remainder of variance in peoples’ responses over time. This non-durable change potentially emerges for a variety of reasons. For example, people might not have clear personal cultural dispositions as measured by surveys. Instead, they might internalize a broad set of considerations and construct an opinion or attitude in the context of the survey interview, with different considerations coming to the forefront of their cognition each interview [@tourangeau2000; @zaller1992; @feldman1992]. This variance in considerations can produce variance in responses even if peoples’ “dispositions” --the full set of considerations they could draw on-- remains stable over time. This variance could also include true errors of measurement, such as respondents misunderstanding the question, respondents or administrators selecting the wrong response, or responses getting coded incorrectly. While this third process reflects important aspects of personal culture, it does not directly touch on the ongoing debate outlined here. Because of this we quantify this component but focus principally on the other two.

We assert that, at the theoretical level, the existence of these two principal processes --interpersonal differences and intrapersonal change-- is indisputable, and that they are, in many ways, related. If intrapersonal change occurs at any point in the life course, it will inevitably produce interpersonal differences by the time people enter a panel survey. Similarly, unless people are fully socialized early in the life course and never deviate from these dispositions (a proposition against which there is substantial evidence), then we expect to observe intrapersonal change in some proportion of the population whenever we survey them over time.

Furthermore, these numbers are not necessarily useful in isolation or without a sense of time. What is important for the theoretical debate is the relative contribution of these two components in explaining cultural differences in a population. In other words, when we look at a population of adults over some time frame, is there enough intrapersonal change relative to interpersonal difference to view it as an important explanation for why people demonstrate cultural differences? That is a question that can only be answered by quantifying the two processes.

However, the real value in quantification lies not in adjudicating past debates, but in opening up avenues of exploration that were limited under previous approaches. For example, classifying questions as either displaying durable change or not obscures differences among questions where there is evidence of durable change, but the relative importance of durable change is different. In quantifying the relative contributions of interpersonal differences and intrapersonal change, researchers can compare the relative importance of these processes in explaining differences across societies, across groups within the same population, across items in the same survey, and across time.

## Expectations

By shifting the question and approach, we expect to see broad similarities across data sets previously used in this analysis.  First, we expect to observe some level of intrapersonal change in personal culture for most measures of personal culture in each data set. However, previous studies found a substantial number of questions (likely between 25 and 35 percent) where there was no evidence of durable change. Consistent with this, we expect to observe significant variation in the proportion of variance explained by intrapersonal change, with some questions showing very little variance explained by intrapersonal change.

Second, we anticipate that the amount of variance explained by intrapersonal change will generally be small relative to differences that exist between people when they start the survey. This assertion should not be controversial, regardless of which theoretical process we expect to be operative for a particular question. Panel surveys do not cover the full span of peoples’ lives, so even if intrapersonal change plays a large role in explaining cultural differences, all the intrapersonal change observed in a panel survey covering even 20 years is likely not enough to account for the full variance of interpersonal differences.

To this point, our expectations are driven principally by what was observed in previous studies. However, theoretical perspectives differ in the amount of variance they expect to be attributable to intrapersonal change, and it is here that quantification can provide a new perspective on existing questions. “Formative period” and cohort theories that expect early-life socialization to be a strong force in adults’ personal culture would require that the amount of variance attributable to intrapersonal change would be insufficient to explain interpersonal differences even when it is extrapolated to a longer time frame. In contrast, other perspectives that attribute more explanatory power to the roles that people currently occupy would expect intrapersonal change to explain a rather large component of variance when extrapolated. While the assumption that the amount of change we observe during the panel could be extrapolated beyond the panel window is not necessarily valid, we believe it is the most valid starting point.

Third, we expect the amount of variance to be explained by each process –interpersonal differences and intrapersonal change– to depend on features of the question and panel. Most obviously, we anticipate that the longer we observe respondents, the greater the degree of resolution at which we observe them (i.e., the number of response options for a particular question), and the larger the sample is, the greater share of variance will be explained by intrapersonal change. Because the Life Course Adaption model (and the various theoretical perspectives that underlie it) suggests that social experiences drive changes of personal culture, and because social experiences are a function of time, observing people for a longer period increases the chance that each of them will undergo some experience that changes their personal culture.

# Analytical Strategy

## Data

We use data from 7 nationally representative panel surveys from Australia, Germany, Great Britain, Switzerland, and the United States (summarized in Table 1), combining all the data files used in previous work [@kiley2020; @lersch2023]. These studies cover a long period of time (the range of surveys spans from 1968 to 2021), with 610 personal culture items that capture attitudes, beliefs, values, self-assessments, self-descriptions, and behaviors [@alwin2007]. We restricted the sample such that individuals between the ages of 18 and 79 are included without further elimination, and in all surveys, we used all possible cases for which respondents provided responses. In the end, the analyses that follow rely on a cross-cultural sample with a cross-domain set of items to capture the breadth of individual personal culture. Appendix A documents the list of all variables used in the upcoming analyses.

```{r, message = FALSE}

library(tidyverse)
library(knitr)
library(kableExtra)

table <- tibble(
  Country = c("Australia", "Germany", "Great Britian", "Great Britian", "Switzerland", "United States", "United States"),
  Survey = c("Household, Income and Labor Dynamics in Australia (HILDA)",
             "Socio-Economic Panel Study (SOEP)",
             "British Household Panel Survey (BHPS)",
             "Understanding Society/UK Household Longitudinal Study (UKHLS)",
             "Swiss Household Panel (SHP)",
             "General Social Survey (GSS)",
             "Panel Study of Income Dynamics (PSID)"),
  Period = c("2001-2021", "1984-2020", "1991-2008", "2009-2020", "1999-2019", "2006-2012", "1968-2019"),
  Outcomes = c(30, 122, 83, 66, 77, 183, 49)
)

kableExtra::kable(table, 
                  format = "latex", 
                  caption = "The Description of the Data Sources") |> 
  kable_styling(latex_options = c("scale_down", "HOLD_position"))

```

## Life Course Adaption Model

We start with Lersch’s [-@lersch2023] Life Course Adaption Model, which formalizes survey responses at time $t$ as a function of individual-level random intercepts and slopes for survey age, what is often called a mixed-effects growth curve model. This model formally assumes a set of propositions about change that reflect the theoretical debate to this point. First, consistent with the settled disposition model, it assumes that people start the survey with cultural differences, modeled as random intercepts for each respondent. Second, it assumes that people change over time, taking the form of random slopes for each respondent as a function of age. Third, it assumes that people deviate around this baseline randomly with age, reflecting “fluctuation” or short-term non-persistent change. Finally, it assumes that people draw on earlier time points in reacting to new events. Formally, this can be written as

$$
y_{i, t} = \beta_0 + u_{0, i} + \beta_1 x_t + u_{1, i} x_{i, t} + \epsilon_{i, t}
$$

where $\beta_0$ is the average intercept, $u_{0, i}$ is the random intercept for individuals, $\beta_1 x_t$ is the average slope, $u_{1, i} x_{i, t}$ is the random slope for individuals, and $\epsilon_{i, t}$ is the random error term that captures transitory disturbances.

As a first step in our analysis, we fit LCAMs to each of the 610 measures of personal culture outlined above. We then use these models to derive two measures of residual variance that correspond to our theoretical processes. First, to measure the importance of interpersonal differences, we calculate the residuals derived from measuring each respondent at the midpoint of their age trajectory, denoted as $\tau$. By dividing this value by the total variance in responses and subtracting it from 1, we calculate the proportion of variance explained by interpersonal differences. Second, we calculate residuals for additionally accounting for the age trajectory, denoted as $\delta$. Subtracting the variance explained by interpersonal differences and dividing this new value by the total variance give us the additional proportion of variance explained by intrapersonal change. These two components reflect total “explained” variance of the model. Our principal outcome of interest is the proportion of explained variance accounted for by intrapersonal change, which we calculate as

$$
\frac{\delta}{\tau + \delta}
$$

As a second step in our analysis, we use ordinary least squares regression to model the proportion of variance explained by intrapersonal change as a function of features of the question and panel. This allows us to explore how these features affect the prevalence of intrapersonal change. These covariates include the number of response options (measured as 2, 3-5, and more than 5), the panel, the log of the number of participants, the number of waves observed, the time since the question was first asked, and the total span of time the question was asked.

# Results

Our results proceed in three parts. We first explore the overall pattern of variance explained by interpersonal differences at baseline and intrapersonal change over time across the seven panels. Next, we analyze the association between the relative proportion of variance explained by intrapersonal change as a function of question, panel, and sample features. Finally, we highlight a few specific comparisons to illustrate the value of quantification relative to adjudication for further exploring theoretical questions.

## Variance Decomposition

Figure 1 plots the proportion of variance explained by interpersonal differences (“between”) and intrapersonal change (“within”) for each of the 610 questions analyzed, plotted separately by panel. We highlight two top-line findings. First, despite the differences in interpretation present in previous works, results from the GSS appear remarkably consistent with results from the other panels. All panels show a wide range of the proportion of explained variance attributable to intrapersonal change, with all panels including questions that show essentially no variance explained by intrapersonal change and all panels including a few questions where more than 20 percent of explained variance is attributable to intrapersonal change. Across all the questions, interpersonal differences explain a much larger share of the variance in responses. Again, this is to be expected. Interpersonal differences capture not just pre-adult socialization, but all accumulated experiences up to the start of the panel.

The substantive importance of intrapersonal change depends on a range of factors, including how long the panel runs, whether assumptions about linear change hold, and whether the panel is capturing a distinctly turbulent period or a distinctly stable period for the item under question. However, we observe an average of about `PERCENT` percent of explained variance being attributed to intrapersonal change across all the questions, with an average duration of `YEARS`. If we assume that the period under observation is “typical” for a question – not a time of extremely heightened (or lowered) sensitivity or change – then it does not seem realistic for this process of intrapersonal change to explain a large share of interpersonal differences.

## Meta-Analysis

Figure 2 plots the results from an ordinary least squares regression of proportion of explained variance attributable to intrapersonal change as a function of question, panel, and sample features. We find that the more response options respondents are given and the larger the sample, the larger the proportion of explained variance attributable to intrapersonal change. We interpret these coefficients as suggesting that greater resolution on a question makes it easier to detect and model change. The earlier a question was asked, even net of how long it has been asked, is associated with lower proportion of intrapersonal change.

The most notable result in the meta analysis is the negative coefficient attached to the span of years covered by the question. Theories of personal cultural change that link changes in personal culture to social experiences, including the LCAM, suggest that the longer we observe respondents, the more likely people are to undergo such experiences and therefore the more variance would be explained by intrapersonal change. The fact that we find a negative coefficient attached to this variable –-the longer we observe people, the less variance is explained by intrapersonal change-– is a challenge to that assumption.

As a further test of this coefficient, we performed a within-question analysis, comparing the proportion of variance explained by intrapersonal change by the same question when using the full panel as when we dropped the final wave, and therefore reduced the total duration of observation for the question. Results from this analysis are presented in Appendix B, which contradict the coefficient from the regression model. The longer we observe a question, the higher proportion of explained variance is attributable to intrapersonal change. We interpret this combination of findings as suggesting that the kinds of questions asked for a longer time period tend to demonstrate less intrapersonal change than questions asked over shorter periods.

## A Case Study: College Education

# Discussion

This paper set out to clarify previous findings about the prevalence of durable change in personal culture, the attitudes, beliefs, values, and preferences that manifest at the individual level. Relying on an approach that quantified the amount of response variance attributable to either intrapersonal change or interpersonal differences at baseline, we found a consistent pattern across data sets previously used in this debate. We see evidence of intrapersonal change over time, with some measures of personal culture displaying quite high levels of intrapersonal change relative to the amount of variance between people. On some questions (such as factors that affect one’s life satisfaction or attitudes about government benefits), it is highly plausible that changes in personal culture as a result of peoples’ experiences as adults are the primary explanation for differences in personal culture.

However, the amount of intrapersonal change is often small (`NUMBER?`) relative to interpersonal differences. On many questions, the amount of explained variance attributable to intrapersonal change is essentially zero. These include `EXAMPLES`. On these questions, assuming the window of observation is not uniquely stable, the amount of variance attributable to intrapersonal change does not seem sufficient to explain how and why adults end up with different personal cultures. For these items, the explanation for personal cultural differences seems to reside principally in early-life experiences.

It is difficult to draw a single conclusion about the relative importance of intrapersonal change in explaining cultural differences across a wide range of questions measured on different scales and over different durations. A key assertion of this work is that blanket statements of “change” versus “stability” as the single pattern among adults is likely to be insufficient. Change and stability are always present among adults, and no viable theoretical model expects just one. But the overall picture, with important caveats, suggests that intrapersonal change during adulthood is not the dominant explanation for cultural differences among adults for most items we measure in surveys. There simply is not enough evidence of intrapersonal change on items relative to the amount of difference we observe across people.

To be clear, this does not mean that adults do not change or are incapable of changing. On basically all questions, some members of the sample change their personal culture. And our results should not be taken to imply that intrapersonal change is not important or consequential. Small numbers of changers can have important consequences in the aggregate. A shift from 49 percent supporting gay marriage to 51 percent supporting it can have significnat political ramifications, even as most respondents remain fixed in their preferences.

## Implications for Cultural Sociology

While we find that people continually reshape their personal culture over time, our results continue to direct attention to the “past conditions of production” that shape peoples’ personal culture early in life. To the extent that we wish to understand what makes adults different, focusing attention on peoples’ transition between social roles, selection into and out of organizational environments, and restructuring of social networks, as well as changes in the cultural resources people have at their disposal, seems to explain a much smaller proportion of their differences than the formative experiences that happen early in life.

The findings we present here are consonant with a range of recent causal inference work that suggests selection effects, rather than treatment effects, seem to explain the majority of differences in personal culture between people in different social roles or positions [@campbell2016; @wodtke2018]. While some people clearly change over time as they move into new roles, this amount of change does not appear to be sufficient to explain what are often substantial differences between people in different roles. Our results seem to suggest that, to the extent we observe differences in personal culture between people in different social roles, such as parenthood, having children, education, or working in positions of authority [@longest2013], or between people in different occupations [@weeden2005], our prior expectation should be that this difference is selection, rather than treatment.

This stability in adulthood could be driven by a range of theoretical processes. It could be the case that adults are exposed to new information at comparable rates as young people but are resistant to changing their personal culture. It could also mean that they are just given fewer opportunities to change their personal culture. For example, adult social environments like workplaces and voluntary associations might expose people to less new information than adolescent social environments such as schools and colleges.

## Implications for Social Change

Because intrapersonal change in personal culture does not appear to be a driving force in cultural differences among adults, our results suggest that processes of social and cultural change at the aggregate level are likely driven by cohort replacement – one group of people with cultural dispositions existing the population and a different set of people with a potentially different set of cultural dispositions entering the population. This is likely true for both large-scale change in societies over time as well as changes within organizations, political parties, occupations, and other social aggregates.

## Implications for Survey Measurement

A major takeaway from the meta-analysis is that greater response resolution – the number of response options survey respondents are given – is associated with a greater share of variance being explained by within-individual change. Again, it is possible that this reflects different kinds of issues being measured with different kinds of scales. But it could also suggest that within-person change, when it happens, is relatively small – moving from “agree” to “strongly agree”, rather than “strongly agree” to “strongly disagree,” for example – and therefore easier to detect with greater resolution.

Second, our results continue to highlight the value of panel surveys for adjudicating different theories of culture that cannot be evaluated with cross-sectional data alone. The process of how individual people store and report personal culture is key to understand the aggregate processes of social change, and . Additionally, we echo previous calls for studies – including panel studies – of the socialization processes of young people. Our results suggest that much of the work of social differentiation takes place before people turn 18. In order to understand why adults are different, we need to understand them before they are adults.

## Limitations

The approach outlined here generates estimates of how much variance in a population is due to each of the two mechanisms outlined here. It does not quantify the proportion of people who “change.” When we observe intrapersonal change in a question, we cannot be sure whether this is driven by many people making small changes or a few people making large changes. Adjudicating this question will require different models that distinguish different kinds of change.

Similarly, the models presented here, even as they allow different slopes for different respondents, still assume that change is a linear function of time. This assumption likely does not hold true in practice. Change over time likely takes non-linear forms. People might jump from one “stable” disposition to another, for example. This assumption is in part a limitation of the data, as most questions are only observed for three to six waves with a large number of missing cases.

Similarly, our approach, in assuming a linear trajectory for change, assumes that durable change is unidirectional. To the extent that people make durable change as a result of a social transition and then return to a previous position later in life as a result of a second transition, this is modeled as residual within-person variance, rather than one of our two theoretical processes.

## Implications for Future Research

We believe our results push past the “needless dichotomy” of asking whether people change their personal culture established by previous work. Our central hope is that, in quantifying the level of intrapersonal change compared to the amount of interpersonal difference at baseline, researchers can start focusing on the individual and social conditions that affect the amount of change we observe.

For example, our results do not speak to the specific mechanisms that limit change in personal culture during adulthood. It is highly plausible that we observe limited amounts of intrapersonal change because the structuring of adulthood rarely provides the social circumstances necessary for intrapersonal change to happen, and that a different social structuring of adulthood would produce higher rates of intrapersonal change. Rather than adults being resistant to change, it could be the case that they are shielded from opportunities for change.

Researchers with strong prior expectations about the social factors that matter for the relative importance or intrapersonal change can partition samples and test hypotheses about the relative importance of intrapersonal change across these partitions.

# Conclusion

In this paper we sought to move past the “needless dichotomy” implicit in the question of whether people change or not. Accepting that some amount of change in personal culture will appear when explored over a long enough time scale, at a high enough degree of resolution, or across a large number of questions, we sought to quantify the relative contribution of intrapersonal change and interpersonal difference across a large range of personal culture, to make broader claims about the importance of change to cultural differences.

Our findings presented here should not be construed as suggesting that social experiences during adulthood are irrelevant for personal culture or can never have meaningful effects. Clearly identifying causal effects of experiences, social transitions, or interventions on personal culture is an important undertaking, even if those treatments do not explain a large variance in personal culture across people. It is only through such studies that researchers will be able to better clarify the mechanisms that produce personal cultural change.

Our hope is that the findings presented here clarify what are actually highly compatible findings from recent debates. People do change. Just not very much.

# References

### 2023-09-17 manuscript.Rmd
data citations added.

@@
 
-We use data from 7 nationally representative panel surveys from Australia, Germany, Great Britain, Switzerland, and the United States (summarized in Table 1), combining all the data files used in previous work [@kiley2020; @lersch2023].
-These studies cover a long period of time (the range of surveys spans from 1968 to 2021), with 610 personal culture items that capture attitudes, beliefs, values, self-assessments, self-descriptions, and behaviors [@alwin2007].
+We use data from 7 nationally representative panel surveys from Australia, Germany, Great Britain, Switzerland, and the United States (summarized in Table 1), combining all the data files used in previous work [@kiley2020; @lersch2023].[^1] These studies cover a long period of time (the range of surveys spans from 1968 to 2021), with 610 personal culture items that capture attitudes, beliefs, values, self-assessments, self-descriptions, and behaviors [@alwin2007].
 We restricted the sample such that individuals between the ages of 18 and 79 are included without further elimination, and in all surveys, we used all possible cases for which respondents provided responses.
@@
 Appendix A documents the list of all variables used in the upcoming analyses.
+
+[^1]: For more information on these data sources, see [@data_gss; @data_hilda; @data_bhps; @data_ukhls; @data_soep; @data_shp; @data_psid].
 

### 2023-09-17 manuscript.Rmd
very small tweaks.

@@
 
-We use data from 7 nationally representative panel surveys from Australia, Germany, Great Britain, Switzerland, and the United States (summarized in Table 1), combining all the data files used in previous work [@kiley2020; @lersch2023].[^1] These studies cover a long period of time (the range of surveys spans from 1968 to 2021), with 610 personal culture items that capture attitudes, beliefs, values, self-assessments, self-descriptions, and behaviors [@alwin2007].
+We use data from 7 nationally representative panel surveys from Australia, Germany, Great Britain, Switzerland, and the United States (summarized in Table 1), combining all the data files used in previous work [@kiley2020; @lersch2023].[^1] These studies cover a long period of time (the range of surveys spans from 1968 to 2021), with 609 personal culture items that capture attitudes, beliefs, values, self-assessments, self-descriptions, and behaviors [@alwin2007].
 We restricted the sample such that individuals between the ages of 18 and 79 are included without further elimination, and in all surveys, we used all possible cases for which respondents provided responses.
@@
   Period = c("2001-2021", "1984-2020", "1991-2008", "2009-2020", "1999-2019", "2006-2012", "1968-2019"),
-  Outcomes = c(30, 122, 83, 66, 77, 183, 49)
+  Outcomes = c(30, 122, 83, 66, 77, 183, 48)
 )
@@
 
-As a first step in our analysis, we fit LCAMs to each of the 610 measures of personal culture outlined above.
+As a first step in our analysis, we fit LCAMs to each of the 609 measures of personal culture outlined above.
 We then use these models to derive two measures of residual variance that correspond to our theoretical processes.
@@
 
-Figure 1 plots the proportion of variance explained by interpersonal differences (“between”) and intrapersonal change (“within”) for each of the 610 questions analyzed, plotted separately by panel.
+Figure 1 plots the proportion of variance explained by interpersonal differences (“between”) and intrapersonal change (“within”) for each of the 609 questions analyzed, plotted separately by panel.
 We highlight two top-line findings.
@@
 The substantive importance of intrapersonal change depends on a range of factors, including how long the panel runs, whether assumptions about linear change hold, and whether the panel is capturing a distinctly turbulent period or a distinctly stable period for the item under question.
-However, we observe an average of about `PERCENT` percent of explained variance being attributed to intrapersonal change across all the questions, with an average duration of `YEARS`.
-If we assume that the period under observation is “typical” for a question – not a time of extremely heightened (or lowered) sensitivity or change – then it does not seem realistic for this process of intrapersonal change to explain a large share of interpersonal differences.
+However, we observe an average of about 9.7 percent of explained variance being attributed to intrapersonal change across all the questions.
+If we assume that the period under observation is “typical” for a question –-not a time of extremely heightened (or lowered) sensitivity or change-– then it does not seem realistic for this process of intrapersonal change to explain a large share of interpersonal differences.
 
@@
 
-This paper set out to clarify previous findings about the prevalence of durable change in personal culture, the attitudes, beliefs, values, and preferences that manifest at the individual level.
+This paper set out to clarify previous findings about the prevalence of durable change in personal culture -- the attitudes, beliefs, values, and preferences that manifest at the individual level.
 Relying on an approach that quantified the amount of response variance attributable to either intrapersonal change or interpersonal differences at baseline, we found a consistent pattern across data sets previously used in this debate.
@@
 
-However, the amount of intrapersonal change is often small (`NUMBER?`) relative to interpersonal differences.
+However, the amount of intrapersonal change is often small (approximately 10% across all surveys) relative to interpersonal differences.
 On many questions, the amount of explained variance attributable to intrapersonal change is essentially zero.
-These include `EXAMPLES`.
+These include questions about civil liberties, abortion, interindividual trust, or sense of civic duty.
 On these questions, assuming the window of observation is not uniquely stable, the amount of variance attributable to intrapersonal change does not seem sufficient to explain how and why adults end up with different personal cultures.
@@
 Small numbers of changers can have important consequences in the aggregate.
-A shift from 49 percent supporting gay marriage to 51 percent supporting it can have significnat political ramifications, even as most respondents remain fixed in their preferences.
+A shift from 49 percent supporting gay marriage to 51 percent supporting it can have significant political ramifications, even as most respondents remain fixed in their preferences.
 
@@
 While we find that people continually reshape their personal culture over time, our results continue to direct attention to the “past conditions of production” that shape peoples’ personal culture early in life.
-To the extent that we wish to understand what makes adults different, focusing attention on peoples’ transition between social roles, selection into and out of organizational environments, and restructuring of social networks, as well as changes in the cultural resources people have at their disposal, seems to explain a much smaller proportion of their differences than the formative experiences that happen early in life.
+To the extent that we wish to understand what makes adults different, focusing attention on peoples’ transition between social roles, selection into and out of organizational environments, and restructuring of social networks, as well as changes in the cultural resources people have at their disposal, seem to explain a much smaller proportion of their differences than the formative experiences that happen early in life.
 
@@
 While some people clearly change over time as they move into new roles, this amount of change does not appear to be sufficient to explain what are often substantial differences between people in different roles.
-Our results seem to suggest that, to the extent we observe differences in personal culture between people in different social roles, such as parenthood, having children, education, or working in positions of authority [@longest2013], or between people in different occupations [@weeden2005], our prior expectation should be that this difference is selection, rather than treatment.
+Our results suggest that, to the extent we observe differences in personal culture between people in different social roles, such as parenthood, having children, education, or working in positions of authority [@longest2013], or between people in different occupations [@weeden2005], our prior expectation should be that this difference is selection, rather than treatment.
 
@@
 
-Because intrapersonal change in personal culture does not appear to be a driving force in cultural differences among adults, our results suggest that processes of social and cultural change at the aggregate level are likely driven by cohort replacement – one group of people with cultural dispositions existing the population and a different set of people with a potentially different set of cultural dispositions entering the population.
+Because intrapersonal change in personal culture does not appear to be a driving force in cultural differences among adults, our results suggest that processes of social and cultural change at the aggregate level are likely driven by cohort replacement –- one group of people with certain cultural dispositions exiting the population and a different set of people with a potentially different set of cultural dispositions entering the population [@vaisey2016].
 This is likely true for both large-scale change in societies over time as well as changes within organizations, political parties, occupations, and other social aggregates.
@@
 
-A major takeaway from the meta-analysis is that greater response resolution – the number of response options survey respondents are given – is associated with a greater share of variance being explained by within-individual change.
+A major takeaway from the meta-analysis is that greater response resolution –-the number of response options survey respondents are given-– is associated with a greater share of variance being explained by within-individual change.
 Again, it is possible that this reflects different kinds of issues being measured with different kinds of scales.
-But it could also suggest that within-person change, when it happens, is relatively small – moving from “agree” to “strongly agree”, rather than “strongly agree” to “strongly disagree,” for example – and therefore easier to detect with greater resolution.
+But it could also suggest that within-person change, when it happens, is relatively small –-moving from “agree” to “strongly agree”, rather than “strongly agree” to “strongly disagree,” for example-– and therefore easier to detect with greater resolution.
 
 Second, our results continue to highlight the value of panel surveys for adjudicating different theories of culture that cannot be evaluated with cross-sectional data alone.
-The process of how individual people store and report personal culture is key to understand the aggregate processes of social change, and .
-Additionally, we echo previous calls for studies – including panel studies – of the socialization processes of young people.
+The process of how individual people store and report personal culture is key to understand the aggregate processes of social change.
+
+Additionally, we echo previous calls for studies –-including panel studies-– of the socialization processes of young people.
 Our results suggest that much of the work of social differentiation takes place before people turn 18.

### 2023-09-19 manuscript.Rmd
draft updates: case, results, measure

@@
 Drawing on the seven panel data sets analyzed in these debates, we use established methods to quantify the amount of variance in responses that is due to between-people differences in baselines and the amount of variance in responses that is due to linear within-person change over time.
-This measure gives us a sense of the importance of intrapersonal change over the course of the panel relative to the differences that exist between people when they start.
+This measure gives us a sense of the importance of intrapersonal change over the course of the panel relative to the differences that exist between people when they start, a quantity with more diverse applications than a simple judgment about whether change happened.
 We also explore features of questions and populations that explain differences in within-person change over time.
@@
 The pattern of results help clarify and, hopefully, transcend previous debates.
-First, we observe similar patterns across the data sets that have produced conflicting interpretations.
+We observe similar patterns across the data sets that have produced conflicting interpretations.
 In all datasets we see similar rates of individual-level change over time across a broad range of measures of personal culture.
@@
 Nevertheless, the results suggest that the intrapersonal change observed (in some cases over more than 10 years) does not seem to play a substantial role in explaining personal cultural differences between adults.
+
 While adding to existing debates, the quantification we propose extends them in new ways that enables scholars to more clearly adjudicate theoretical mechanisms at play.
-We provide a series of illustrative examples of the utility of this approach for addressing different theoretical questions.
+We provide an example by exploring differences in the relative importance of intrapersonal change for people with and without college degrees, comparing both within and across questions in ways that the tournament of models cannot.
 
@@
 Instead, different perspectives argue that each of these ideal-typical models is more operative at different times, for different people, and for different elements of personal culture.
-For example, adolescence and early adulthood, typically viewed  as a “formative period” for personal cultural development, is likely characterized by higher rates of “active updating,” while middle age and later life are potentially characterized more by “settled dispositions” [@alwin1991; @krosnick1989; @eaton2009].
+For example, adolescence and early adulthood, typically viewed  as a “formative period” for personal cultural development, is likely characterized by higher rates of active updating, while middle age and later life are potentially characterized more by settled dispositions [@alwin1991; @krosnick1989; @eaton2009].
 Similarly, highly salient issues, such as views around gay rights in the 2010s; issues that see substantial elite realignments, such as views around the Vietnam conflict in the 1970s [@zaller1992]; and novel issues of public opinion, such as views around vaccines during the Covid-19 pandemic, might be characterized by active updating, while low-salience, established issues are characterized by stability.
@@
 
-However, the claim that change is a relatively infrequent phenomenon has been difficult to square with research identifying individual change across a number of cultural dimensions, such as morality [@brocic2021], trust [@mewes2021], and concerns about immigration [@kratz2021].
+However, the claim that change is a relatively infrequent phenomenon has been difficult to square with research identifying durable individual change across a number of cultural dimensions, such as morality [@brocic2021], trust [@mewes2021], and concerns about immigration [@kratz2021], as a result of social experiences.
 Similarly, there are longitudinal studies showing that cues from political elites can change a person’s position on specific policy issues [@slothuus2021; @zaller1992] and that changes in a person’s close contacts and acquaintances can change their attitudes on group-related politics [@gelman2021; @dellaposta2018].
@@
 
-In his tournament of models, Lersch finds broad support for the LCAM in data from five countries, with the LCAM favored over the AUM and SDM on 297 of 428 questions he analyzed and the SDM favored over the other two models on 112 items, with the remaining questions producing inconclusive results.
+In his tournament of models, Lersch finds broad support for the LCAM in data from five countries, with the LCAM favored over the AUM and SDM on 297 of 428 questions he analyzed and the SDM favored over the other two models on 112 items, with the remaining questions producing inconclusive results (no questions preferred the AUM).
 He concludes that “new experiences over the life course, which expose individuals to new social environments, including novel interactions, and different sets of resources, can persistently move individuals’ personal culture in novel directions” [@lersch2023: p.
@@
 
-However, despite differences in how these researchers interpret their findings, the empirical results are not far apart.
+Despite differences in how these researchers interpret their findings, the empirical results are not far apart.
 Kiley and Vaisey [-@kiley2020] found that the majority of questions they explored favored the AUM, meaning they showed some evidence of durable change among adults, a finding broadly compatible with Lersch’s results.
@@
 For example, Brocic and Miles [-@brocic2021] find that completing graduate degrees in humanities, arts, and social sciences shift peoples’ moral relativism only about .2 standard deviations on average compared to people with no college degree, and this is the largest effect they identify.
-And studies of aggregate change show that, even on items where cohort effects dominate, some people do change over time [@vaisey2016].
+And studies of aggregate change show that, even on items where cohort effects dominate, there is always evidence that some people change over time [@vaisey2016].
 In other words, despite different interpretations, the results of previous work are in many ways highly consonant.
@@
 
-The shortcoming of current debates is that they focus on whether people *ever* change, a question that is not theoretically interesting since the answer is almost always “yes.”  As noted previously, if we observe enough people for long enough, we will likely observe some of them making durable changes.
+The shortcoming of current debates is that they focus on whether people *ever* change, a question that is not theoretically interesting since the answer is almost always “yes.”  As noted previously, if researchers observe enough people for long enough, they will likely observe some of them making durable changes.
 A lack of evidence for change in one question might simply indicate that the question was not asked at a clear enough level of resolution, for a long enough time frame, or to enough people to find evidence of change.
@@
 First, interpersonal differences at the start of the survey, what Lersch calls “early imprinting,” reflects the accumulated experiences of people prior to entering the panel survey that shape how they answer questions.
-While this is typically associated with a “formative period” or a “settled disposition” (hence the name early imprinting), it could also reflect transformative experiences that happen at any age, so long as they occurred before the start of the panel and leave a lasting effect on how people answer questions.
+While this is typically associated with a formative period or a settled disposition (hence the name early imprinting), it could also reflect transformative experiences that happen at any age, so long as they occurred before the start of the panel and leave a lasting effect on how people answer questions.
 For example, among people already retired upon entering the panel, their “early imprinting” might reflect this life course transition.
@@
 
-The second, the amount of intrapersonal linear change over the course of the survey, what Lersch calls “persistent change” or “adaption” and Kiley and Vaisey call “active updating,” reflects durable changes in personal culture over time.
-Lersch suggests that these changes are brought on by social triggers such as moving into a new environment, adopting new social roles, or broader changes in the social structure.
+The second set of processes, what Lersch calls “persistent change” or “adaption” and Kiley and Vaisey call “active updating,” reflects durable changes in personal culture over time and is captured by the amount of intrapersonal linear change over the course of the survey.
+Lersch suggests that these changes are brought on by social triggers such as moving into a new environment or adopting new social roles.
+They could also reflect the diffusion of new culture across social networks, changes in social network composition (or knowledge about who is in a person's network), new cues from political elites or other cultural leaders, the emergence of new issues in politics or culture, or other large-scale social shifts.
 
@@
 First, consistent with the settled disposition model, it assumes that people start the survey with cultural differences, modeled as random intercepts for each respondent.
-Second, it assumes that people change over time, taking the form of random slopes for each respondent as a function of age.
-Third, it assumes that people deviate around this baseline randomly with age, reflecting “fluctuation” or short-term non-persistent change.
+Second, it assumes that people change over time, taking the form of random slopes for each respondent as a linear function of time.
+Third, it assumes that people deviate around this baseline randomly over time, reflecting “fluctuation” or short-term non-persistent change.
 Finally, it assumes that people draw on earlier time points in reacting to new events.
@@
 As a first step in our analysis, we fit LCAMs to each of the 609 measures of personal culture outlined above.
-We then use these models to derive two measures of residual variance that correspond to our theoretical processes.
-First, to measure the importance of interpersonal differences, we calculate the residuals derived from measuring each respondent at the midpoint of their age trajectory, denoted as $\tau$.
-By dividing this value by the total variance in responses and subtracting it from 1, we calculate the proportion of variance explained by interpersonal differences.
-Second, we calculate residuals for additionally accounting for the age trajectory, denoted as $\delta$.
-Subtracting the variance explained by interpersonal differences and dividing this new value by the total variance give us the additional proportion of variance explained by intrapersonal change.
-These two components reflect total “explained” variance of the model.
-Our principal outcome of interest is the proportion of explained variance accounted for by intrapersonal change, which we calculate as
+We then use these models to derive two measures of systematic variance that correspond to the two collections of theoretical processes outlined above: interpersonal differences at baseline and intrapersonal change over time.
+First, to measure the importance of interpersonal differences, we calculate the sum of the squared residuals derived from measuring each respondent at the midpoint of their age trajectory.[^midpoint] We then divide this value by the total sum of squares and subtract it from from 1, producing a value we denote as $\tau$, or the proportion of variance explained by interpersonal differences.
+This number is equivalent to an $R^2$ value generated by using random intercepts for all respondents.
+Second, we calculate the sum of squared residuals generated by additionally accounting for the age trajectory.
+We divide this number by the total sum of squares and subtract it from 1.
+We then subtract $\tau$ from this value.
+This gives us a value we denote as $\delta$, the additional proportion of variance explained by intrapersonal change.
+
+These two numbers reflect the total systematic or "explained" variance of the model that can be attributed to either differences across people or linear change over the course of the panel.[^explained] Our principal outcome of interest is the proportion of systematic variance accounted for by intrapersonal change, which we calculate as
 
@@
 $$
+
+[^midpoint]: We use the scale midpoint, rather than the actual difference at baseline, to better capture the differences between people independent of their age trajectories.
+
+[^explained]: This does not mean that other factors might not explain additional variance if they were included in the model.
+For example, peoples' responses about religious beliefs and participation might systematically vary by the day of the week or month in which they are surveyed.
+This variation is moved into the "random" or non-systematic component of the model for us.
+
+We can think of this number as akin to an Intraclass Correlation Coefficient focused only on the systematic components of the model while ignoring residual variance that might be attributable to measurement error or non-persisting change [see also the variance decomposition of period and cohort by @vaisey2016].
+The higher the number is, the more variance in responses is attributable to intrapersonal change during the panel.
+The lower the number, the less intrapersonal change seems to play a large role in explaining cultural differences.
+
+Because this quotient summarizes variance and therefore has no natural referent, it is hard to say at which values we observe "a lot" of intrapersonal change.
+But summarizing questions in this way allows us to compare the relative importance of our two broad processes across a range of questions that might have very different levels of non-systematic variance, giving us a baseline sense of the relative prevalence of these two processes for each question (something not achievable under the tournament of models approach), and allowing us to compare which kinds of questions show more or less intrapersonal change relative to interpersonal differences.
+While the tournament of models approach allowed researchers to say whether questions in general showed evidence of updating, this approach allows us to make more specific claims about the prevalence of these theoretical processes for particular questions relative to each other.
 
@@
 Our results proceed in three parts.
-We first explore the overall pattern of variance explained by interpersonal differences at baseline and intrapersonal change over time across the seven panels.
-Next, we analyze the association between the relative proportion of variance explained by intrapersonal change as a function of question, panel, and sample features.
+We first explore the overall pattern of systematic variance explained by interpersonal differences at baseline and intrapersonal change over time across the seven panels.
+Next, we analyze the association between the relative proportion of systematic variance explained by intrapersonal change as a function of question, panel, and sample features.
 Finally, we highlight a few specific comparisons to illustrate the value of quantification relative to adjudication for further exploring theoretical questions.
@@
 
-Figure 1 plots the proportion of variance explained by interpersonal differences (“between”) and intrapersonal change (“within”) for each of the 609 questions analyzed, plotted separately by panel.
-We highlight two top-line findings.
-First, despite the differences in interpretation present in previous works, results from the GSS appear remarkably consistent with results from the other panels.
-All panels show a wide range of the proportion of explained variance attributable to intrapersonal change, with all panels including questions that show essentially no variance explained by intrapersonal change and all panels including a few questions where more than 20 percent of explained variance is attributable to intrapersonal change.
-Across all the questions, interpersonal differences explain a much larger share of the variance in responses.
+Figure 1 plots the proportion of systematic variance attributable to interpersonal differences (“between”) and intrapersonal change (“within”) for each of the 609 questions analyzed, plotted separately by panel.
+All panels show a similar range of the proportion of systematic variance attributable to intrapersonal change, with all panels including questions that show essentially no systematic variance attributable to intrapersonal change and all panels including a few questions where more than 20 percent of systematic variance is attributable to intrapersonal change.
+Across all questions, interpersonal differences explain a much larger share of the systematic variance in responses.
 Again, this is to be expected.
-Interpersonal differences capture not just pre-adult socialization, but all accumulated experiences up to the start of the panel.
+Interpersonal differences capture not just pre-adult socialization, but all accumulated experiences up to the start of the panel that might have a bearing on personal culture.
+Despite the differences in interpretation present in previous works, results from the GSS are highly consistent with results from the other panels.
 
-The substantive importance of intrapersonal change depends on a range of factors, including how long the panel runs, whether assumptions about linear change hold, and whether the panel is capturing a distinctly turbulent period or a distinctly stable period for the item under question.
-However, we observe an average of about 9.7 percent of explained variance being attributed to intrapersonal change across all the questions.
+PARAGRAPH HERE ABOUT RESULTS: MOST/LEAST SYSTEMATIC VARIANCE. DISTRIBUTION. NEED THAT INFO.
+
+Appendix A presents the estimated proportion of variance attributable to interpersonal differences and intrapersonal change, as well as the proportion of residual variance, for each question analyzed.
+In general, interpersonal differences are almost always the largest component of the total variance and tend to explain between 50 and 70 percent of total variance, while intrapersonal change is always the smallest, typically explaining between 0 and 15 percent of total variance.
+Residual variance tends to account for between 20 to 40 percent of variance (though on several questions residual variance is greater than 50 percent).
+
+The substantive importance of intrapersonal change for understanding cultural differences depends on a range of factors, including how long the panel runs, whether assumptions about linear change hold, and whether the panel is capturing a distinctly turbulent period or a distinctly stable period for the item under question.
+However, we observe a mean of about 9.7 percent of explained variance being attributed to intrapersonal change across all the questions.
 If we assume that the period under observation is “typical” for a question –-not a time of extremely heightened (or lowered) sensitivity or change-– then it does not seem realistic for this process of intrapersonal change to explain a large share of interpersonal differences.
@@
 
-Figure 2 plots the results from an ordinary least squares regression of proportion of explained variance attributable to intrapersonal change as a function of question, panel, and sample features.
-We find that the more response options respondents are given and the larger the sample, the larger the proportion of explained variance attributable to intrapersonal change.
-We interpret these coefficients as suggesting that greater resolution on a question makes it easier to detect and model change.
-The earlier a question was asked, even net of how long it has been asked, is associated with lower proportion of intrapersonal change.
+Figure 2 plots the results from an ordinary least squares regression of proportion of systematic variance attributable to intrapersonal change as a function of question, panel, and sample features.
+We find that the more response options respondents are given and the larger the sample, the larger the proportion of systematic variance is attributable to intrapersonal change.
+We interpret these coefficients as suggesting greater resolution on a question makes it easier to detect and model change.
+The earlier a question was asked, even net of how long it has been asked, the lower the proportion of systematic variance is attributable to intrapersonal change.
 
 The most notable result in the meta analysis is the negative coefficient attached to the span of years covered by the question.
-Theories of personal cultural change that link changes in personal culture to social experiences, including the LCAM, suggest that the longer we observe respondents, the more likely people are to undergo such experiences and therefore the more variance would be explained by intrapersonal change.
+Theories of personal cultural change that link changes in personal culture to social experiences, including the LCAM, suggest that the longer we observe respondents, the more likely people are to undergo potentially transformative experiences and therefore the more variance would be explained by intrapersonal change.
 The fact that we find a negative coefficient attached to this variable –-the longer we observe people, the less variance is explained by intrapersonal change-– is a challenge to that assumption.
 
-As a further test of this coefficient, we performed a within-question analysis, comparing the proportion of variance explained by intrapersonal change by the same question when using the full panel as when we dropped the final wave, and therefore reduced the total duration of observation for the question.
-Results from this analysis are presented in Appendix B, which contradict the coefficient from the regression model.
-The longer we observe a question, the higher proportion of explained variance is attributable to intrapersonal change.
-We interpret this combination of findings as suggesting that the kinds of questions asked for a longer time period tend to demonstrate less intrapersonal change than questions asked over shorter periods.
+As a further test of this coefficient, we compared the proportion of systematic variance attributable to intrapersonal change when using the full duration of a panel compared to when when we dropped the final wave and therefore reduced the total duration of observation for the question.
+If the coefficient reflects a true negative "causal" effect of duration on systematic variance attributable to intrapersonal change, we should see that same effect within questions.
+Results from this analysis are presented in Appendix B and contradict the coefficient from the regression model.
+The longer we observe a question, the higher proportion of systematic variance is attributable to intrapersonal change.
+We interpret this combination of findings as suggesting that the kinds of questions asked for a longer time period tend to demonstrate less intrapersonal change than questions asked over shorter periods, rather than a true function of time.
 
 ## A Case Study: College Education
+
+While these broad results help clarify previous debates, they mostly reframe previous empirical results.
+To better illustrate the value of our approach, we turn to a specific empirical example: the relative importance of intrapersonal change and interpersonal difference for explaining cultural differences by education levels.
+
+Previous work has established a positive relationship between education and attitude stability, especially on issues related to American politics.
+This stability is often attributed to education facilitating "chronic knowledge" -- a general understanding of the domain of American politics, including the positions held by major parties and political figures and how issues relate to one another at a logical or socio-logical level [@zaller1992; @alvarez2002; @boutyline2017].
+Because they have more knowledge of American politics, people with a college education are better able to consistently connect the considerations in their cognition with the answer choices they are presented with in a survey.
+
+This work has tended to focus on the fact that college-educated Americans give responses that are less likely to be affected by measurement error or short-term influences [@zaller1992; @alwin2003].
+However, this focus on the difference in the non-systematic (residual) component of variance across groups obscures the fact that the relative importance of different components of systematic variance might differ across these two groups as well.
+
+It could be the case that education, chronic knowledge, and attention to the political sphere makes intrapersonal change more important in explaining differences among people with higher levels of education.
+For example, these people might be more likely to make systematic changes in response to the emergence of new information, new issues, or political realignments, while those without chronic knowledge display more variance around their baselines without their baselines changing much [@zaller1992].
+In contrast, it could be the case that the stability demonstrated by educated respondents reflects the fact that they have already formed durable opinions and are relatively closed off to new information.
+College could serve as a formative experience that solidifies attitudes for people who attend.
+But among people who do not attend college, later life experiences might prove more important in forming or changing attitudes, as these experiences potentially provide information that college-educated peers received earlier.
+
+To test these competing propositions, we use the models presented above to calculate the total systematic variance for people with at least a bachelor's degree and people with less than a bachelor's degree at wave 1[^wave] of the General Social Survey's panels.[^missing] We focus on the GSS because it contains the largest number of questions tapping general political dispositions, which is the domain where education has proven particularly relevant for understanding attitude stability.
+The GSS also covers a particularly turbulent window of American politics from 2006 to 2014.
+This window covers the start of the Great Recession, debates about federal intervention in and regulation of Wall Street, the election of Barack Obama as the first black president in U.S.
+history, debates about the role of the federal government in the health care sector, the emergence of the Tea Party, and political realignment and clarification on the issue of gay marriage, among other topics.
+
+[^wave]: A small number of respondents report different highest degrees at wave 1, wave 2, and wave 3.
+Some of this is due to measurement error, and some of it is due to a small number of people obtaining a higher degree during the four years of the panel.
+Estimating the panel with highest at wave 3 or highest degree reported across the panel produces functionally identical results.
+
+[^missing]: We estimate the models presented in [earlier figure] using the full sample, but only use respondents who have information on their highest degree to generate the results presented in figures [these figures].
+In total, ### respondents (%%% percent) reported no degree information.
+
+Figure XXX plots the distribution of differences between people with less than a bachelor's degree and people with at least a bachelor's degree at wave 1 of the panel.
+Values greater than 0 would indicate that intrapersonal change is more important in explaining the responses of college graduates than those of people without a college degree, while values less than 0 would indicate that intrapersonal change is less important for explaining variation among people without a college degree than those with a college degree.
+
+There is a clear pattern in [figure above]: for more than 80 percent of GSS items explored here, intrapersonal change is a larger component of systematic variance for people without a college degree than for people with a college degree (they have negative values in the figure).
+While most of these differences are small in absolute terms (less than 2 percentage points), several are greater than 5 percentage points.
+Given the distribution observed in Figure [the earlier one] showing that the systematic variance attributable to intrapersonal change averages [average across all questions], a 5 percentage point difference between groups is quite substantial.
+
+To more clearly illustrate some of these differences, we highlight eight questions designed to tap general political dispositions: partisan identification (Democrat vs.
+Republican) and ideological identification (liberal vs.
+conservative) on seven-point scales; four questions about the government's role in improving the condition of the poor, paying people's medical bills, giving special treatment to blacks, and doing things that private businesses could do, measured on five-point scales; a question about whether the government should do more to reduce income differences, measured on a seven-point scale; and one question about whether blacks should be given preferences in hiring, measured on a five-point scale.
+Estimates for the proportion of systematic variance attributable to intrapersonal change on these eight questions, for both education groups, are presented in Figure XXX.
+
+On all eight questions presented in Figure XXX, intrapersonal change explains less systematic variance for people with a college degree than for people without a college degree.
+There are also large differences across questions on the amount of systematic variance attributable to intrapersonal change for both groups.
+For example, on the question of whether the government should try to solve more problems or leave those problems to be solved by private businesses ("government do more or less"), less than one percent of the systematic variance is attributable to intrapersonal change for both groups.
+In other words, while people might vacillate on this question at random (37 percent of variance is residual for this question), there is functionally no evidence that people make systematic changes of opinion on this issue during the GSS panel.
+
+In contrast, partisan identification and political ideology both show a larger proportion of systematic variance attributable to intrapersonal change, with quite large differences between people with a college degree and those without.
+Compared to the other questions explored here, change as adults (and, potentially, their experiences as adults) plays a much larger role in explaining why people report the partisan identification and political ideology they do.
+And this is particularly true for respondents who do not have a college degree.
+The amount of systematic variance in ideological identification that is attributable to intrapersonal change among non-college educated respondents is almost four times that of college-educated respondents.
+
+We note here that this meaningful difference in the proportion of systematic variance attributable to intrapersonal change across education groups and across questions would not be detectable using previous methods.
+For partisan identification and political ideology, both college-educated respondents and people with less than college degree would likely favor the AUM or LCAM over the SDM because they show evidence of some members of the population making intrapersonal change.
+In other words, using the tournament of models approach obscures the fact that the relative importance of intrapersonal change in explaining differences on these two items is very different across these two groups, that intrapersonal change explains more systematic variance for ideological identification than partisan identification for people without a college degree, or that there appears to be more durable change on questions of affirmative action than on questions of government aid to black Americans.
+
+It is important to note here that these differences should not be interpreted as causal effects of attending college on the importance of social experiences for cultural change.
+These results are potentially confounded by age, social class, race, gender, and other factors that explain selection into higher education.
+We simply use these estimates as a starting point for a broader discussion.
 
@@
 This paper set out to clarify previous findings about the prevalence of durable change in personal culture -- the attitudes, beliefs, values, and preferences that manifest at the individual level.
-Relying on an approach that quantified the amount of response variance attributable to either intrapersonal change or interpersonal differences at baseline, we found a consistent pattern across data sets previously used in this debate.
-We see evidence of intrapersonal change over time, with some measures of personal culture displaying quite high levels of intrapersonal change relative to the amount of variance between people.
+Relying on an approach that quantified the amount of systematic variance attributable to either intrapersonal change or interpersonal differences at baseline, we found a consistent pattern across data sets previously used in this debate.
+We see evidence of intrapersonal change over time on almost all questions, with some measures of personal culture displaying quite high levels of intrapersonal change relative to the amount of variance between people.
 On some questions (such as factors that affect one’s life satisfaction or attitudes about government benefits), it is highly plausible that changes in personal culture as a result of peoples’ experiences as adults are the primary explanation for differences in personal culture.
 
-However, the amount of intrapersonal change is often small (approximately 10% across all surveys) relative to interpersonal differences.
-On many questions, the amount of explained variance attributable to intrapersonal change is essentially zero.
-These include questions about civil liberties, abortion, interindividual trust, or sense of civic duty.
+However, the amount of intrapersonal change is often small relative to interpersonal differences (approximately 10 percent across all surveys).
+On many questions, the amount of systematic variance attributable to intrapersonal change is essentially zero.
+These include questions about civil liberties, abortion, interindividual (kk note: generalized?) trust, or sense of civic duty.
 On these questions, assuming the window of observation is not uniquely stable, the amount of variance attributable to intrapersonal change does not seem sufficient to explain how and why adults end up with different personal cultures.
@@
 
-It is difficult to draw a single conclusion about the relative importance of intrapersonal change in explaining cultural differences across a wide range of questions measured on different scales and over different durations.
+It is difficult to draw a single conclusion about the relative importance of intrapersonal change in explaining cultural differences across a wide range of questions measured on different scales and over different durations of time.
 A key assertion of this work is that blanket statements of “change” versus “stability” as the single pattern among adults is likely to be insufficient.
-Change and stability are always present among adults, and no viable theoretical model expects just one.
+Change and stability are always present among adults, and no viable theoretical model expects just one process to explain all difference.
 But the overall picture, with important caveats, suggests that intrapersonal change during adulthood is not the dominant explanation for cultural differences among adults for most items we measure in surveys.
-There simply is not enough evidence of intrapersonal change on items relative to the amount of difference we observe across people.
+There simply is not enough evidence of intrapersonal change on these items relative to the amount of difference we observe across people.
 
@@
 A shift from 49 percent supporting gay marriage to 51 percent supporting it can have significant political ramifications, even as most respondents remain fixed in their preferences.
+Our results could also be interpreted as suggesting that opportunities for durable change or the social conditions that facilitate durable change are just extremely uncommon.
+Changes in social structures, the life course, or cultural expectations could make intrapersonal change a much more consequential explanation for cultural differences.
+
+Furthermore, the results we present related to education and political views suggest that the relative importance of different theoretical processes might differ by group.
+In line with life course perspectives, the results presented here suggest that importance of different experiences for attitude change are potentially contingent on previous experiences.
+For example, to the extent that we want to explain why people have different personal culture, the answer is likely different for people who have a college degree and those who do not.
+Mid-life experiences appear to have more significant implications for the personal culture for people without a college degree than for people with a college degree.
 

### 2023-09-20 manuscript.Rmd
results and discussion

@@
 
-A recurring focus in the sociology of culture is the question of whether people, as they move through their lives and transition between social roles, change their personal culture –- the declarative and nondeclarative attitudes, beliefs, values, and dispositions that manifest at the individual level [@lizardo2017; @kiley2020; @lersch2023].
+A recurring focus in the sociology of culture is the question of whether people, as they move through their lives and transition between social roles, change their personal culture –- the declarative and non-declarative attitudes, beliefs, values, and dispositions that manifest at the individual level [@lizardo2017; @kiley2020; @lersch2023].
 
@@
 We observe similar patterns across the data sets that have produced conflicting interpretations.
-In all datasets we see similar rates of individual-level change over time across a broad range of measures of personal culture.
+In all data sets we see similar rates of individual-level change over time across a broad range of measures of personal culture.
 But quantifying these processes also reveals that intrapersonal change plays a substantially smaller role in explaining the variance in peoples’ personal culture than between-person differences in their baselines.
@@
 The higher the number is, the more variance in responses is attributable to intrapersonal change during the panel.
-The lower the number, the less intrapersonal change seems to play a large role in explaining cultural differences.
+The lower the number, the less intrapersonal change seems to play a large role in explaining cultural differences during the panel.
 
@@
 Figure 1 plots the proportion of systematic variance attributable to interpersonal differences (“between”) and intrapersonal change (“within”) for each of the 609 questions analyzed, plotted separately by panel.
-All panels show a similar range of the proportion of systematic variance attributable to intrapersonal change, with all panels including questions that show essentially no systematic variance attributable to intrapersonal change and all panels including a few questions where more than 20 percent of systematic variance is attributable to intrapersonal change.
-Across all questions, interpersonal differences explain a much larger share of the systematic variance in responses.
+To broadly summarize the results, all panels show a similar range of the proportion of systematic variance attributable to intrapersonal change.
+Across panels, the median proportion of systematic variance attributable to intrapersonal change is 0.094 (mean of 0.098), with an interquartile range of 0.053 to 0.139.
+All panels include questions that show essentially no systematic variance attributable to intrapersonal change and all panels include questions where more than 20 percent of systematic variance is attributable to intrapersonal change.
+
+Across all questions, interpersonal differences explain a much larger share of the systematic variance in responses than intrapersonal change.
 Again, this is to be expected.
 Interpersonal differences capture not just pre-adult socialization, but all accumulated experiences up to the start of the panel that might have a bearing on personal culture.
-Despite the differences in interpretation present in previous works, results from the GSS are highly consistent with results from the other panels.
 
-PARAGRAPH HERE ABOUT RESULTS: MOST/LEAST SYSTEMATIC VARIANCE. DISTRIBUTION. NEED THAT INFO.
+To the extent that there are differences across the panels, the PSID has the highest mean (.120) and median (.135) proportion of systematic variance attributable to intrapersonal change.
+While we cannot disentangle features of the sample from features of the questions asked to each sample, the specific samples for many PSID questions have lower average ages than those from other panels.
+To the extent that younger respondents might be more likely to make durable changes of opinion, this higher explanatory power of intrapersonal change might reflect the distinct age profile of respondents in this sample.
+At the other end, the GSS has the lowest mean (.073) and median (.069) proportion of variance attributable to intrapersonal change.
+This potentially reflects the fact that the GSS observes people for a shorter duration, on average, than the other panels.
+If, consistent with life course adaption theories, people are more likely to make significant cultural changes the longer we observe them, then duration likely affects the amount of variance attributable to intrapersonal change (a point we explore more below).
+However, the GSS results are still highly consistent with results from the other panels.
+
+While there are some differences between panels, these differences are small compared to the differences within panels.
+For about 6 percent of items explored here, more than 20 percent of systematic variance is attributable to intrapersonal change.
+These questions tend to ask about external referents (e.g., confidence in specific government leaders or political parties), life satisfaction, or current financial position.
+At the other end, questions about religious identification, views on gender roles, and support for civil liberties tend to have very low proportion of variance attributable to intrapersonal change.
+
+In contrast to the tournament of models approach, quantifying change this way allows us to explore variation in the relative importance of intrapersonal change across questions that all both show evidence of change.
+For example, Kiley and Vaisey [-@kileyvaisey2020] found that confidence in the press and confidence in religious leaders were both characterized by active updating, our results show that intrapersonal change is much more important for explaining variance in confidence in the press (0.164) than confidence in religion (0.049).
 
 Appendix A presents the estimated proportion of variance attributable to interpersonal differences and intrapersonal change, as well as the proportion of residual variance, for each question analyzed.
-In general, interpersonal differences are almost always the largest component of the total variance and tend to explain between 50 and 70 percent of total variance, while intrapersonal change is always the smallest, typically explaining between 0 and 15 percent of total variance.
-Residual variance tends to account for between 20 to 40 percent of variance (though on several questions residual variance is greater than 50 percent).
+In general, interpersonal differences are almost always the largest component of the total variance and tend to explain between 55 and 70 percent of total variance, while intrapersonal change is always the smallest, typically explaining between 3 and 8 percent of total variance.
+Residual variance tends to account for between 22 and 37 percent of variance (though on several questions residual variance is greater than 50 percent).
 
-The substantive importance of intrapersonal change for understanding cultural differences depends on a range of factors, including how long the panel runs, whether assumptions about linear change hold, and whether the panel is capturing a distinctly turbulent period or a distinctly stable period for the item under question.
-However, we observe a mean of about 9.7 percent of explained variance being attributed to intrapersonal change across all the questions.
-If we assume that the period under observation is “typical” for a question –-not a time of extremely heightened (or lowered) sensitivity or change-– then it does not seem realistic for this process of intrapersonal change to explain a large share of interpersonal differences.
+As noted above, the substantive importance of intrapersonal change for understanding cultural differences depends on a range of factors, including how long the panel runs, whether assumptions about linear change hold, and whether the panel is capturing a distinctly turbulent period or a distinctly stable period for the item under question.
+However, if we assume that the period under observation is “typical” for a question –-not a time of extremely heightened (or lowered) sensitivity or change-– then it does not seem realistic for this process of intrapersonal change to explain a large share of interpersonal differences for many of the questions we observe.
 
@@
 Figure 2 plots the results from an ordinary least squares regression of proportion of systematic variance attributable to intrapersonal change as a function of question, panel, and sample features.
-We find that the more response options respondents are given and the larger the sample, the larger the proportion of systematic variance is attributable to intrapersonal change.
+These models also include fixed effects for panels (not shown), so coefficients reflect the association between a particular sample or question characteristic within a panel.[^paneleffects]
+
+[^paneleffects]: KK NOTE: Can we summarize these panel fixed effects at all?
+
+Figure 2 shows that the more response options respondents are given and the larger the sample, the larger the proportion of systematic variance is attributable to intrapersonal change.
 We interpret these coefficients as suggesting greater resolution on a question makes it easier to detect and model change.
-The earlier a question was asked, even net of how long it has been asked, the lower the proportion of systematic variance is attributable to intrapersonal change.
+The earlier a question was asked, even net of how long it was been asked, the lower the proportion of systematic variance is attributable to intrapersonal change.
 
 The most notable result in the meta analysis is the negative coefficient attached to the span of years covered by the question.
+The longer a question is observed, the less systematic variance is attributable to intrapersonal change.
 Theories of personal cultural change that link changes in personal culture to social experiences, including the LCAM, suggest that the longer we observe respondents, the more likely people are to undergo potentially transformative experiences and therefore the more variance would be explained by intrapersonal change.
-The fact that we find a negative coefficient attached to this variable –-the longer we observe people, the less variance is explained by intrapersonal change-– is a challenge to that assumption.
+The fact that we find a negative coefficient attached to this variable is a challenge to that assumption.
 
 As a further test of this coefficient, we compared the proportion of systematic variance attributable to intrapersonal change when using the full duration of a panel compared to when when we dropped the final wave and therefore reduced the total duration of observation for the question.
-If the coefficient reflects a true negative "causal" effect of duration on systematic variance attributable to intrapersonal change, we should see that same effect within questions.
+If the coefficient reflects a true negative effect of duration on systematic variance attributable to intrapersonal change, we should see that same effect within questions.
 Results from this analysis are presented in Appendix B and contradict the coefficient from the regression model.
-The longer we observe a question, the higher proportion of systematic variance is attributable to intrapersonal change.
+The longer we observe the same question, the higher proportion of systematic variance is attributable to intrapersonal change.
 We interpret this combination of findings as suggesting that the kinds of questions asked for a longer time period tend to demonstrate less intrapersonal change than questions asked over shorter periods, rather than a true function of time.
+
+This reinforces the explanation for why the GSS displays the lowest average proportion of systematic variance attributable to intrapersonal change: its shorter duration.
+[KK NOTE: Does this inform any other panel differences?
+Does the PSID, BHPS or UKHLS have longer durations on average?]
 
@@
 
-While these broad results help clarify previous debates, they mostly reframe previous empirical results.
-To better illustrate the value of our approach, we turn to a specific empirical example: the relative importance of intrapersonal change and interpersonal difference for explaining cultural differences by education levels.
+While these broad results help speak to previous debates, they mostly re-frame previous findings without extending the theoretical debate.
+To better illustrate the value of our approach for addressing different theoretical questions, we turn to a specific empirical example: the relative importance of intrapersonal change and interpersonal difference for explaining cultural differences by education levels.
 
@@
 
-This paper set out to clarify previous findings about the prevalence of durable change in personal culture -- the attitudes, beliefs, values, and preferences that manifest at the individual level.
+This paper set out to clarify previous findings about the prevalence of durable change in personal culture -- the attitudes, beliefs, values, behaviors, and preferences that manifest at the individual level.
 Relying on an approach that quantified the amount of systematic variance attributable to either intrapersonal change or interpersonal differences at baseline, we found a consistent pattern across data sets previously used in this debate.
 We see evidence of intrapersonal change over time on almost all questions, with some measures of personal culture displaying quite high levels of intrapersonal change relative to the amount of variance between people.
-On some questions (such as factors that affect one’s life satisfaction or attitudes about government benefits), it is highly plausible that changes in personal culture as a result of peoples’ experiences as adults are the primary explanation for differences in personal culture.
+On some questions, such as the factors that affect one’s life satisfaction or attitudes about government benefits, it is highly plausible that differences in peoples’ experiences as adults are the primary explanation for differences in personal culture.
 
-However, the amount of intrapersonal change is often small relative to interpersonal differences (approximately 10 percent across all surveys).
+However, the amount of intrapersonal change is often small relative to interpersonal differences (less than 10 percent on average across all surveys).
 On many questions, the amount of systematic variance attributable to intrapersonal change is essentially zero.
-These include questions about civil liberties, abortion, interindividual (kk note: generalized?) trust, or sense of civic duty.
-On these questions, assuming the window of observation is not uniquely stable, the amount of variance attributable to intrapersonal change does not seem sufficient to explain how and why adults end up with different personal cultures.
+These include questions about civil liberties, abortion, inter-individual (kk note: generalized?) trust, or sense of civic duty.
+On these questions and the large number of others with a low proportion of systematic variance attributable to intrapersonal change, there does not appear to be sufficient cultural change during adulthood for experiences and transitions between social locations to be the primary explanation for why adults end up with different personal cultures.
 For these items, the explanation for personal cultural differences seems to reside principally in early-life experiences.
 
-It is difficult to draw a single conclusion about the relative importance of intrapersonal change in explaining cultural differences across a wide range of questions measured on different scales and over different durations of time.
-A key assertion of this work is that blanket statements of “change” versus “stability” as the single pattern among adults is likely to be insufficient.
-Change and stability are always present among adults, and no viable theoretical model expects just one process to explain all difference.
-But the overall picture, with important caveats, suggests that intrapersonal change during adulthood is not the dominant explanation for cultural differences among adults for most items we measure in surveys.
+More important than adjudicating past debates, however, we show that the approach developed in this paper can adjudicate novel theoretical questions about the conditions in which change or stability is more prevalent and more important for explaining cultural differences within a population.
+As an example, we showed that the role of intrapersonal change in explaining cultural differences varies significantly by question and by education.
+Intrapersonal change plays almost no role in explaining general sentiments about the government's role in the economy, while it plays a much larger role in explaining ideological and partisan identification.
+At the same time, across issues, intrapersonal change is much more important for explaining cultural variation among people without a college degree than for people with a college degree.
+While this pattern might be confounded by age and other factors that affect college completion, it potentially suggests that college-going locks in personal culture in a way that other experiences do not, making adult experiences less influential for political dispositions for college graduates compared to the rest of the population.
+
+## Implications for Cultural Sociology
+
+To this point, the theoretical debate in cultural sociology has focused on the existence of personal cultural change among adults and its relative importance.
+We believe the "existence" debate is theoretically misguided.
+Personal cultural change will always exist to some degree in an adult population, and no viable theoretical model expects just one process --stability or change-- to characterize a population.
+Our results reinforce that assertion, finding a mixture of interpersonal differences and intrapersonal change for all items.
+Even when we observe very low levels of systematic variance attributable to intrapersonal change for particular questions, this is potentially attributable to insufficient data, insufficient resolution on the underlying belief, or insufficient duration of observation.
+
+A more reasonable theoretical debate pushed by previous work is the relative importance of these two processes for explaining cultural variation.
+While change might always be prevalent, its impact on social differences might be small or large, and our approach has been designed to clarify the contribution of intrapersonal change to such differences.
+It is difficult to draw a single conclusion on this question across a wide range of questions measured on different scales and over different durations of time, but the overall picture, with important caveats, suggests that intrapersonal change during adulthood is not the dominant explanation for cultural differences among adults for most items we measure in surveys.
 There simply is not enough evidence of intrapersonal change on these items relative to the amount of difference we observe across people.
 
-To be clear, this does not mean that adults do not change or are incapable of changing.
-On basically all questions, some members of the sample change their personal culture.
-And our results should not be taken to imply that intrapersonal change is not important or consequential.
+To be clear, none of our results should be interpreted as suggesting that adults do not change or are incapable of changing.
+On all questions, some members of the sample change their personal culture at least a little bit.
+Our results also should not be taken to imply that intrapersonal change is not important or consequential.
 Small numbers of changers can have important consequences in the aggregate.
 A shift from 49 percent supporting gay marriage to 51 percent supporting it can have significant political ramifications, even as most respondents remain fixed in their preferences.
-Our results could also be interpreted as suggesting that opportunities for durable change or the social conditions that facilitate durable change are just extremely uncommon.
+And our results should not detract from work trying to estimate the mechanisms through which personal cultural differences --however small-- emerge among adults.
+Just because change is small does not mean the mechanisms that produce is are not important to understand.
+
+Our results should be read principally as suggesting that if researchers want to explain why people differ in their personal culture, it is imperative to look at the conditions in which they were raised and the experiences that happened before they reached adulthood.
+To the extent that we want to explain what makes adults different, focusing attention on peoples’ transition between social roles, their selection into and out of organizational environments, and changes in their social networks, as much sociological work does, appears to explain a much smaller proportion of their differences than the formative experiences that happen early in life.
+
+In this way, our findings echo a range of recent causal inference work that suggests selection effects, rather than treatment effects, seem to explain the majority of differences in personal culture between people in different social roles or positions [@campbell2016; @wodtke2018].
+While some people clearly change over time as they move into new social roles or into new social environments, this amount of change does not appear to be sufficient to explain what are often substantial differences between people in different roles.
+Our results suggest that, to the extent we observe differences in personal culture between people in different social roles, such as parenthood, having children, education, or working in positions of authority [@longest2013], or between people in different occupations [@weeden2005], our prior expectation should be that this difference is selection, rather than treatment.
+
+Our results do not speak to exactly why intrapersonal change lacks significant explanatory power in understanding culture.
+First, it could be the case that adults are exposed to opportunities for change at the same rate, but become less willing to change.
+Or it could be that opportunities for durable change or the social conditions that facilitate durable change are just extremely uncommon.
 Changes in social structures, the life course, or cultural expectations could make intrapersonal change a much more consequential explanation for cultural differences.
@@
 
-## Implications for Cultural Sociology
-
-While we find that people continually reshape their personal culture over time, our results continue to direct attention to the “past conditions of production” that shape peoples’ personal culture early in life.
-To the extent that we wish to understand what makes adults different, focusing attention on peoples’ transition between social roles, selection into and out of organizational environments, and restructuring of social networks, as well as changes in the cultural resources people have at their disposal, seem to explain a much smaller proportion of their differences than the formative experiences that happen early in life.
-
-The findings we present here are consonant with a range of recent causal inference work that suggests selection effects, rather than treatment effects, seem to explain the majority of differences in personal culture between people in different social roles or positions [@campbell2016; @wodtke2018].
-While some people clearly change over time as they move into new roles, this amount of change does not appear to be sufficient to explain what are often substantial differences between people in different roles.
-Our results suggest that, to the extent we observe differences in personal culture between people in different social roles, such as parenthood, having children, education, or working in positions of authority [@longest2013], or between people in different occupations [@weeden2005], our prior expectation should be that this difference is selection, rather than treatment.
-
-This stability in adulthood could be driven by a range of theoretical processes.
-It could be the case that adults are exposed to new information at comparable rates as young people but are resistant to changing their personal culture.
-It could also mean that they are just given fewer opportunities to change their personal culture.
-For example, adult social environments like workplaces and voluntary associations might expose people to less new information than adolescent social environments such as schools and colleges.
-
 ## Implications for Social Change
 
-Because intrapersonal change in personal culture does not appear to be a driving force in cultural differences among adults, our results suggest that processes of social and cultural change at the aggregate level are likely driven by cohort replacement –- one group of people with certain cultural dispositions exiting the population and a different set of people with a potentially different set of cultural dispositions entering the population [@vaisey2016].
+Because intrapersonal change in personal culture does not appear to be a driving force in cultural differences among adults, our results suggest that processes of social and cultural change at the aggregate level are likely driven more by cohort replacement, one group of people with certain cultural dispositions exiting the population and a different set of people with a potentially different set of cultural dispositions entering the population, than by contemporaneous social conditions [@vaisey2016; @underwood2022].
 This is likely true for both large-scale change in societies over time as well as changes within organizations, political parties, occupations, and other social aggregates.
+
+However, because our results suggest that the role of formative experiences compared to contemporaneous environments differs by a person's level of education, changes in educational attainment over time, as well as changes in the social structuring of the life course might be shifting the relative importance of these two factors over time.
+If we observe further differences across socio-demographic groups like education, social class, and race/ethnicity in the relative importance of these processes, this would argue for a greater incorporation the interaction of demographic processes with processes of individual cultural change into our explanations of aggregate cultural change.
 
@@
 
-Second, our results continue to highlight the value of panel surveys for adjudicating different theories of culture that cannot be evaluated with cross-sectional data alone.
-The process of how individual people store and report personal culture is key to understand the aggregate processes of social change.
+Our results continue to highlight the value of long-running panel surveys for adjudicating different theories of culture that cannot be evaluated with cross-sectional data alone.
+Across our results, residual fluctuations in responses account for more than four times as much variance, on average, as linear persistent change, meaning that most change in responses is not systematic.
+Under these circumstances, extending panels beyond two waves provides significantly more leverage to understand how and why personal culture changes.
 
-Additionally, we echo previous calls for studies –-including panel studies-– of the socialization processes of young people.
-Our results suggest that much of the work of social differentiation takes place before people turn 18.
-In order to understand why adults are different, we need to understand them before they are adults.
+Additionally, we echo previous calls for more studies –-including panel studies-– of the socialization processes of young people.
+Our results suggest that most of the variance between adults emerges before people turn 18 or relatively soon after.
+In limiting our empirical perspective to experiences that happen after that point, social scientists potentially focus on questions that, while important, lack the capacity to explain major cultural differences.
 
@@
 
-The approach outlined here generates estimates of how much variance in a population is due to each of the two mechanisms outlined here.
+The approach outlined here generates estimates of how much variance in a population is due to each of two broad processes: intrapersonal change and interpersonal differences at baseline.
 It does not quantify the proportion of people who “change.” When we observe intrapersonal change in a question, we cannot be sure whether this is driven by many people making small changes or a few people making large changes.
@@
 
-Similarly, the models presented here, even as they allow different slopes for different respondents, still assume that change is a linear function of time.
+Our results and interpretations also hinge on how we have defined change and the assumptions built into the Life Course Adaption Model.
+This model treats cultural trajectories as varying linear slopes for each respondent, assuming that change is a linear function of time.
 This assumption likely does not hold true in practice.
-Change over time likely takes non-linear forms.
-People might jump from one “stable” disposition to another, for example.
+Change over time likely takes non-linear and discontinuous forms.
+People might jump from one “stable” disposition to another, for example, or experience a discontinuous "turning point" that upsets a stable trajectory.
 This assumption is in part a limitation of the data, as most questions are only observed for three to six waves with a large number of missing cases.
+Panels with more waves might allow for more flexible models of change that do not require this assumption of linearity.
 
-Similarly, our approach, in assuming a linear trajectory for change, assumes that durable change is unidirectional.
-To the extent that people make durable change as a result of a social transition and then return to a previous position later in life as a result of a second transition, this is modeled as residual within-person variance, rather than one of our two theoretical processes.
+Similarly, in assuming a linear trajectory for change, our approach assumes that durable change is unidirectional.
+To the extent that people make durable change as a result of a social transition and then return to a previous position later in life as a result of a second transition (marriage and divorce, for example), this is captured as residual within-person variance, rather than one of our two broad theoretical processes.
+This is a potentially reasonable assumption on short panels -- change that does not last two years is hard to classify as durable -- but it might not be true in longer panels.
+Other researchers with more flexible definitions of "durable change" might potentially identify more durable change even in the same population.
 
-## Implications for Future Research
+Finally, while we examine an incredible broad array of measures of personal culture, from assessments about the relative importance of different features when buying a new car and policy preferences to religious beliefs and core values, we are limited by the kinds of questions that are asked in panel surveys and the contexts in which these surveys were administered (Australia, Germany, Switzerland, the United States, and the United Kingdom).
+This means our findings tend to reflect questions about general (national) politics, gender roles, immigration and race relations, and general well being.
+Our data does not cover some dimensions of culture that might be relevant for other aspects of sociological exploration (artistic tastes, leisure activities, and time use, for example).
 
-We believe our results push past the “needless dichotomy” of asking whether people change their personal culture established by previous work.
-Our central hope is that, in quantifying the level of intrapersonal change compared to the amount of interpersonal difference at baseline, researchers can start focusing on the individual and social conditions that affect the amount of change we observe.
+# Implications for Future Research
 
-For example, our results do not speak to the specific mechanisms that limit change in personal culture during adulthood.
-It is highly plausible that we observe limited amounts of intrapersonal change because the structuring of adulthood rarely provides the social circumstances necessary for intrapersonal change to happen, and that a different social structuring of adulthood would produce higher rates of intrapersonal change.
-Rather than adults being resistant to change, it could be the case that they are shielded from opportunities for change.
+We believe the approach outlined here can push past the “needless dichotomy” implicit in the question of whether people change or not and address more theoretically informed questions about the conditions under which the relative importance of intrapersonal change and interpersonal differences matter for explaining cultural differences.
+Characterizing questions as displaying change or not can only take researchers so far, but the question of why some questions demonstrate more change while others do not, or why some groups are characterized by more stability than others, has the potential to weigh in on a broader range of theoretical debates.
+We hope researchers move toward those questions.
 
-Researchers with strong prior expectations about the social factors that matter for the relative importance or intrapersonal change can partition samples and test hypotheses about the relative importance of intrapersonal change across these partitions.
+While we focused on education as a determinant of the explanatory power of intrapersonal change, our approach can readily be extended to other socio-demographic categories that might affect intrapersonal change -- race/ethnicity, religiosity, social class, gender, and political affiliation, for example -- or their intersections.
+Researchers with different data might be able to evaluate the relative importance of intrapersonal change and interpersonal differences across contexts as well, either national contexts if the same questions are administered across countries or organizational contexts if data traces students within schools, workers within workplaces, or congregation members within congregations.
 
-# Conclusion
-
-In this paper we sought to move past the “needless dichotomy” implicit in the question of whether people change or not.
-Accepting that some amount of change in personal culture will appear when explored over a long enough time scale, at a high enough degree of resolution, or across a large number of questions, we sought to quantify the relative contribution of intrapersonal change and interpersonal difference across a large range of personal culture, to make broader claims about the importance of change to cultural differences.
-
-Our findings presented here should not be construed as suggesting that social experiences during adulthood are irrelevant for personal culture or can never have meaningful effects.
-Clearly identifying causal effects of experiences, social transitions, or interventions on personal culture is an important undertaking, even if those treatments do not explain a large variance in personal culture across people.
-It is only through such studies that researchers will be able to better clarify the mechanisms that produce personal cultural change.
-
-Our hope is that the findings presented here clarify what are actually highly compatible findings from recent debates. People do change. Just not very much.
+The wide variation across questions in the proportion of systematic variation attributable to intrapersonal change also calls for a more systematic exploration for why some forms of culture are characterized by stability while others are characterized by change, and what the consequences of these differences for the organization of beliefs in the public might be.
 

### 2023-09-20 manuscript.Rmd
small changes in wording

@@
 
+## Change and Stability in Personal Culture
+
 Recent debates about whether adults undergo intrapersonal cultural change emerged in part because sociological models of cultural change at the aggregate level tend to implicitly invoke one of two models of individual behavior.
@@
 
-The second model summarized by Kiley and Vaisey, an “Active Updating Model” (AUM), posits that people continually update their personal culture as they move through life.
+The second model summarized by Kiley and Vaisey [-@kiley2020], an “Active Updating Model” (AUM), posits that people continually update their personal culture as they move through life.
 This model suggests people make new cultural meanings, and therefore durably change their personal culture, as they encounter new social environments, discourses, and information [@gross2009; @swidler2001].
@@
 In contrast to the tournament of models approach, quantifying change this way allows us to explore variation in the relative importance of intrapersonal change across questions that all both show evidence of change.
-For example, Kiley and Vaisey [-@kileyvaisey2020] found that confidence in the press and confidence in religious leaders were both characterized by active updating, our results show that intrapersonal change is much more important for explaining variance in confidence in the press (0.164) than confidence in religion (0.049).
+For example, Kiley and Vaisey [-@kiley2020] found that confidence in the press and confidence in religious leaders were both characterized by active updating, our results show that intrapersonal change is much more important for explaining variance in confidence in the press (0.164) than confidence in religion (0.049).
 
@@
 
-This work has tended to focus on the fact that college-educated Americans give responses that are less likely to be affected by measurement error or short-term influences [@zaller1992; @alwin2003].
+This work has tended to focus on the fact that college-educated Americans give responses that are less likely to be affected by measurement error or short-term influences [@zaller1992; @alwin2007].
 However, this focus on the difference in the non-systematic (residual) component of variance across groups obscures the fact that the relative importance of different components of systematic variance might differ across these two groups as well.
@@
 In contrast, it could be the case that the stability demonstrated by educated respondents reflects the fact that they have already formed durable opinions and are relatively closed off to new information.
-College could serve as a formative experience that solidifies attitudes for people who attend.
-But among people who do not attend college, later life experiences might prove more important in forming or changing attitudes, as these experiences potentially provide information that college-educated peers received earlier.
+College could serve as a formative experience that solidifies personal culture for people who attend.
+But among people who do not attend college, later life experiences might prove more important in forming or changing personal culture, as these experiences potentially provide information that college-educated peers received earlier.
 
@@
 Some of this is due to measurement error, and some of it is due to a small number of people obtaining a higher degree during the four years of the panel.
-Estimating the panel with highest at wave 3 or highest degree reported across the panel produces functionally identical results.
+Estimating the panel with highest at wave 1 or highest degree reported across the panel produces functionally identical results.
 
@@
 Figure XXX plots the distribution of differences between people with less than a bachelor's degree and people with at least a bachelor's degree at wave 1 of the panel.
-Values greater than 0 would indicate that intrapersonal change is more important in explaining the responses of college graduates than those of people without a college degree, while values less than 0 would indicate that intrapersonal change is less important for explaining variation among people without a college degree than those with a college degree.
+Values greater than 0 would indicate that intrapersonal change is more important in explaining the responses of college graduates than those of people without a college degree, while values less than 0 would indicate that intrapersonal change is less important for explaining variation among people with a college degree than those without a college degree.
 
@@
 
-This paper set out to clarify previous findings about the prevalence of durable change in personal culture -- the attitudes, beliefs, values, behaviors, and preferences that manifest at the individual level.
+This paper set out to clarify previous findings about the prevalence of durable change in personal culture -- the attitudes, beliefs, values, preferences, and behaviors that manifest at the individual level.
 Relying on an approach that quantified the amount of systematic variance attributable to either intrapersonal change or interpersonal differences at baseline, we found a consistent pattern across data sets previously used in this debate.
@@
 On many questions, the amount of systematic variance attributable to intrapersonal change is essentially zero.
-These include questions about civil liberties, abortion, inter-individual (kk note: generalized?) trust, or sense of civic duty.
+These include questions about civil liberties, abortion, generalized trust, or sense of civic duty.
 On these questions and the large number of others with a low proportion of systematic variance attributable to intrapersonal change, there does not appear to be sufficient cultural change during adulthood for experiences and transitions between social locations to be the primary explanation for why adults end up with different personal cultures.
@@
 And our results should not detract from work trying to estimate the mechanisms through which personal cultural differences --however small-- emerge among adults.
-Just because change is small does not mean the mechanisms that produce is are not important to understand.
+Just because change is small does not mean the mechanisms that produce it are not important to understand.
 
@@
 Furthermore, the results we present related to education and political views suggest that the relative importance of different theoretical processes might differ by group.
-In line with life course perspectives, the results presented here suggest that importance of different experiences for attitude change are potentially contingent on previous experiences.
+In line with life course perspectives, the results presented here suggest that importance of different experiences for cultural change are potentially contingent on previous experiences.
 For example, to the extent that we want to explain why people have different personal culture, the answer is likely different for people who have a college degree and those who do not.
@@
 However, because our results suggest that the role of formative experiences compared to contemporaneous environments differs by a person's level of education, changes in educational attainment over time, as well as changes in the social structuring of the life course might be shifting the relative importance of these two factors over time.
-If we observe further differences across socio-demographic groups like education, social class, and race/ethnicity in the relative importance of these processes, this would argue for a greater incorporation the interaction of demographic processes with processes of individual cultural change into our explanations of aggregate cultural change.
+If we observe further differences across socio-demographic groups like education, social class, and race/ethnicity in the relative importance of these processes, this would argue for a greater incorporation of the interaction of demographic processes with processes of individual cultural change into our explanations of aggregate cultural change.
 
@@
 Again, it is possible that this reflects different kinds of issues being measured with different kinds of scales.
-But it could also suggest that within-person change, when it happens, is relatively small –-moving from “agree” to “strongly agree”, rather than “strongly agree” to “strongly disagree,” for example-– and therefore easier to detect with greater resolution.
+But it could also suggest that within-person change, when it happens, is relatively small –-moving from “agree” to “strongly agree,” rather than “strongly agree” to “strongly disagree,” for example-– and therefore easier to detect with greater resolution.
 
@@
 To the extent that people make durable change as a result of a social transition and then return to a previous position later in life as a result of a second transition (marriage and divorce, for example), this is captured as residual within-person variance, rather than one of our two broad theoretical processes.
-This is a potentially reasonable assumption on short panels -- change that does not last two years is hard to classify as durable -- but it might not be true in longer panels.
+This is a potentially reasonable assumption on short panels --change that does not last two years is hard to classify as durable-- but it might not be true in longer panels.
 Other researchers with more flexible definitions of "durable change" might potentially identify more durable change even in the same population.
 
-Finally, while we examine an incredible broad array of measures of personal culture, from assessments about the relative importance of different features when buying a new car and policy preferences to religious beliefs and core values, we are limited by the kinds of questions that are asked in panel surveys and the contexts in which these surveys were administered (Australia, Germany, Switzerland, the United States, and the United Kingdom).
+Finally, while we examine an incredible broad array of measures of personal culture, from assessments about the relative importance of different features when buying a new car and policy preferences to religious beliefs and core values, we are limited by the kinds of questions that are asked in panel surveys and the contexts in which these surveys were administered (Australia, Germany, Switzerland, the United Kingdom and the United States).
 This means our findings tend to reflect questions about general (national) politics, gender roles, immigration and race relations, and general well being.
@@
 
-# Implications for Future Research
+## Implications for Future Research
 
@@
 
-While we focused on education as a determinant of the explanatory power of intrapersonal change, our approach can readily be extended to other socio-demographic categories that might affect intrapersonal change -- race/ethnicity, religiosity, social class, gender, and political affiliation, for example -- or their intersections.
-Researchers with different data might be able to evaluate the relative importance of intrapersonal change and interpersonal differences across contexts as well, either national contexts if the same questions are administered across countries or organizational contexts if data traces students within schools, workers within workplaces, or congregation members within congregations.
+While we focused on education as a determinant of the explanatory power of intrapersonal change, our approach can readily be extended to other socio-demographic categories that might affect intrapersonal change --race/ethnicity, religiosity, social class, gender, and political affiliation, for example-- or their intersections.
+Researchers with different data might be able to evaluate the relative importance of intrapersonal change and interpersonal differences across contexts as well, either national contexts if the same questions are administered across countries, or organizational contexts if data traces students within schools, workers within workplaces, or congregation members within congregations.
 

### 2023-09-21 manuscript.Rmd

@@
 abstract: "Recent work has produced conflicting interpretations on the question of whether adults undergo durable changes in their personal culture over time.
-This paper asserts that this divergence is due more to the approach adopted by researchers –-a tournament of models-– than their findings, which are highly consonant.
-We reanalyze the data used previously in this debate with an approach that quantifies the processes of cultural difference, rather than one that detects the presence or absence of change.
+This paper asserts that this divergence is less due to their findings than the approach taken -- a tournament of models focusing on the presence or absence of change.
+To move beyond this, we reanalyze the data used previously in this debate with an approach that quantifies the processes of cultural difference.
 In doing so, we compare how much cultural difference is explained by intrapersonal change and interpersonal differences.
 Our results harmonize recent findings by showing that, while most measures of personal culture show change over time, the variance explained by intrapersonal change is typically small compared to interpersonal differences at baseline.
-We then exemplify how quantifying these processes offer new ways to address theoretical questions of when and how the relative importance of intrapersonal change and interpersonal differences shifts."
+Quantifying these processes offers new ways to address theoretical questions of when and how their relative importance shifts.
+We examplify this with the case of college experience."
 fontsize: 11pt
@@
 
-A recurring focus in the sociology of culture is the question of whether people, as they move through their lives and transition between social roles, change their personal culture –- the declarative and non-declarative attitudes, beliefs, values, and dispositions that manifest at the individual level [@lizardo2017; @kiley2020; @lersch2023].
+A recurring focus in the sociology of culture is the question of whether people, as they move through their lives and transition between social roles, change their personal culture –- the declarative and non-declarative attitudes, beliefs, values, practices, and dispositions [@lizardo2017; @kiley2020; @lersch2023].
 
 This seemingly simple question underlies a number of central debates.
-For example, pragmatist theories of action claim that changes in social environments cause people to adapt their views and make new meanings [@swidler2001; @gross2009], while Bourdiesian practice theories argue that the “past conditions of production” leave a lasting mark on peoples’ personal culture throughout their lives [@bourdieu1990].
-Similarly, models of social influence require people to adapt their culture in the face of new information [@vaisey2010a], while models of aggregate social and cultural change that place emphasis on cohort effects require them to be open to change while young but fairly resistant to change as they age [@ryder1965].
+For example, pragmatist theories of action claim that changes in social environments cause people to adapt their views and make new meanings [@swidler2001; @gross2009], while Bourdieusian practice theories argue that the “past conditions of production” leave a mark on people's personal culture that lasts throughout their lives [@bourdieu1990].
+Similarly, models of social influence assume that people adapt their culture in the face of new information [@vaisey2010a], while the emphasis on cohort effects in models of aggregate social and cultural change requires them to be open to change while young but become fairly resistant to it as they age [@ryder1965].
 Finally, life course theories posit slower but still important changes over time as people advance through important transitions in their lives [@elder2016; @bardi2011].
 
-Because all these theoretical perspectives have some empirical support, adjudicating when and how they apply is important for further understanding how early experiences and contemporary environments shape personal culture over the life course.
+All of the above theoretical perspectives have some empirical support.
+All the more it is important to adjudicate when and how they apply to advance an understanding the differences in people's personal culture that we observe.
 That being said, scholars have struggled to reach a consensus on the relative importance of change during adulthood for explaining cultural differences.
-Over the span of a few years, most apparent changes in people’s responses to personal culture items appear to be transitory, with little evidence that adults make persistent changes of personal culture [@kiley2020; @vaisey2021].
-When looked at over longer time horizons, however, there is at least some evidence that adults make persistent cultural changes, especially as they move through different social roles across the life course [@lersch2023].
+Over the span of a few years, most apparent changes in people’s responses to personal culture items appear to be transitory, with little evidence of persistent change among adults [@kiley2020; @vaisey2021].
+When looked at over longer time horizons, however, there is at least some evidence that adults make persistent cultural changes, especially as they move through distinct social roles across the life course [@lersch2023].
 
-These different conclusions are partly due to the fact that researchers decided to adjudicate the different perspectives by taking a “tournament of models” approach [@lersch2023: p.
-228].
-In this approach, researchers have formalized models of responses over time as demonstrating either “change” or “no change” and compared which model fits observed data better, functionally asking whether we ever observe within-person change in personal culture.
+These seemingly inconsistent findings are partly due to the fact that researchers have taken a “tournament of models” approach [@lersch2023: p.
+228] in adjudicating the different perspectives.
+Accordingly, researchers have formalized models of responses to cultural items over time as demonstrating either “change” or “no change” to then compare which model fits observed data better, thus functionally asking whether we ever observe within-person change in personal culture.
 
 We argue, however, that asking only *whether* people change obscures what are actually highly compatible results.
-Whether change “really happened” in a given data set depends not only on sample populations, statistical methods and power, and item measurement, but crucially also on definitions of what counts as “change,” thus running the risk of getting caught up in unresolvable debates about related decisions.
-Moreover, it narrows the focus on particular inquiries such as, “do we see change in this measure?” Such inquiries are ill-suited to address what’s at the core of the theoretical debate.
-There is enough evidence to assume that neither socialization nor the environment alone can explain cultural differences.
-What’s left to answer is under what conditions, across which groups, and in what domains do we observe differences in the relative importance of intrapersonal change and interpersonal differences.
+Whether change “really happened” in a given data set depends not only on sample populations, item measurement, statistical methods and power, but crucially also on definitions of what counts as “change,” thus running the risk of getting caught up in unresolvable debates about related decisions.
+[MAYBE BETTER TO SAY HERE: WHICH QUICKLY RELEGATES THE POSSIBILITY OF AN ANSWER BACK TO AN ULTIMATELY UNRESOLVABLE THEORETICAL DEBATE.] Moreover, it narrows the focus on particular questions such as, “Do we see change in this measure?” Such inquiries are ill-suited to address what’s at the core of the theoretical debate.
+There is enough evidence to assume that neither socialization at an early age nor the influence of the contemporary environment alone can explain cultural differences.
+The question that remains is under what conditions, across which groups, and in what domains do we observe differences in the relative importance of intrapersonal change and interpersonal differences.
 This focus is necessarily comparative and requires an approach that quantifies forms of cultural change rather than calling victory for the one or the other perspective.
 
-In this paper, we try to overcome previous debates and move away from the “tournament of models” approach toward a quantification of the processes that lead to differences in personal culture.
-Drawing on the seven panel data sets analyzed in these debates, we use established methods to quantify the amount of variance in responses that is due to between-people differences in baselines and the amount of variance in responses that is due to linear within-person change over time.
-This measure gives us a sense of the importance of intrapersonal change over the course of the panel relative to the differences that exist between people when they start, a quantity with more diverse applications than a simple judgment about whether change happened.
-We also explore features of questions and populations that explain differences in within-person change over time.
+In this paper, we try to move beyond previous debates by moving away from the “tournament of models” approach toward a quantification of the processes that lead to differences in personal culture.
+Drawing on the seven panel data sets analyzed in these debates, we use established methods to quantify the amount of variance in responses that is due to between-people differences in baselines versus the amount of variance in responses that is due to linear within-person change over time.
+This measure gives us a sense of the importance of intrapersonal change over the course of the panel relative to the differences that exist between people when they start.
+We then explore features of questions and populations that explain differences in within-person change over time.
 
-The pattern of results help clarify and, hopefully, transcend previous debates.
 We observe similar patterns across the data sets that have produced conflicting interpretations.
@@
 But quantifying these processes also reveals that intrapersonal change plays a substantially smaller role in explaining the variance in peoples’ personal culture than between-person differences in their baselines.
-Whether this amount of intrapersonal change is “meaningful” ultimately remains a theoretical question reflected in prior expectations concerning the question, population, and time.
+Of course, whether this amount of intrapersonal change is “meaningful” ultimately remains a theoretical question reflected in prior expectations concerning the question, population, and time.
 Nevertheless, the results suggest that the intrapersonal change observed (in some cases over more than 10 years) does not seem to play a substantial role in explaining personal cultural differences between adults.
 
-While adding to existing debates, the quantification we propose extends them in new ways that enables scholars to more clearly adjudicate theoretical mechanisms at play.
-We provide an example by exploring differences in the relative importance of intrapersonal change for people with and without college degrees, comparing both within and across questions in ways that the tournament of models cannot.
+This pattern of results helps clarify and, hopefully, transcend previous debates.
+What is more, the measure we propose offers a quantity with more diverse applications than a simple judgment about whether change happened.
+To examplify this, we apply the measure to explore differences in the relative importance of intrapersonal change for people with and without college degrees, comparing both within and across questions in ways that the tournament of models cannot.
+
+While adding to existing debates, we hope the quantification we propose extends them in new ways and enables scholars to more clearly adjudicate theoretical mechanisms at play in shaping people's personal culture.
 

### 2023-09-21 manuscript.Rmd

@@
 abstract: "Recent work has produced conflicting interpretations on the question of whether adults undergo durable changes in their personal culture over time.
-This paper asserts that this divergence is less due to their findings than the approach taken -- a tournament of models focusing on the presence or absence of change.
+This paper asserts that this divergence is less due to researchers' findings than the approach taken -- a tournament of models focusing on the presence or absence of change.
 To move beyond this, we reanalyze the data used previously in this debate with an approach that quantifies the processes of cultural difference.
@@
 
-A recurring focus in the sociology of culture is the question of whether people, as they move through their lives and transition between social roles, change their personal culture –- the declarative and non-declarative attitudes, beliefs, values, practices, and dispositions [@lizardo2017; @kiley2020; @lersch2023].
+A recurring focus in the sociology of culture is the question of whether people, as they move through their lives and transition between social roles, change their personal culture –- the declarative and non-declarative attitudes, beliefs, values, practices, and dispositions that exist at the individual level [@lizardo2017; @kiley2020; @lersch2023].
 
-This seemingly simple question underlies a number of central debates.
+This seemingly simple question underlies a number of central debates across sociological subfields.
 For example, pragmatist theories of action claim that changes in social environments cause people to adapt their views and make new meanings [@swidler2001; @gross2009], while Bourdieusian practice theories argue that the “past conditions of production” leave a mark on people's personal culture that lasts throughout their lives [@bourdieu1990].
@@
 
-All of the above theoretical perspectives have some empirical support.
-All the more it is important to adjudicate when and how they apply to advance an understanding the differences in people's personal culture that we observe.
-That being said, scholars have struggled to reach a consensus on the relative importance of change during adulthood for explaining cultural differences.
+Because all of the above theoretical perspectives have some empirical support, the current theoretical debate revolves around not whether they exist, but when and how they each apply and the relative contribution of each to explaining cultural differences.
+That being said, scholars have struggled to reach a consensus on the relative importance of change during adulthood.
 Over the span of a few years, most apparent changes in people’s responses to personal culture items appear to be transitory, with little evidence of persistent change among adults [@kiley2020; @vaisey2021].
+These results suggest change is not a strong force in explaining personal cultural differences.
 When looked at over longer time horizons, however, there is at least some evidence that adults make persistent cultural changes, especially as they move through distinct social roles across the life course [@lersch2023].
 
-These seemingly inconsistent findings are partly due to the fact that researchers have taken a “tournament of models” approach [@lersch2023: p.
-228] in adjudicating the different perspectives.
-Accordingly, researchers have formalized models of responses to cultural items over time as demonstrating either “change” or “no change” to then compare which model fits observed data better, thus functionally asking whether we ever observe within-person change in personal culture.
+These seemingly inconsistent findings are partly due to the fact that researchers have taken a “tournament of models” approach in adjudicating the different perspectives [@lersch2023: p.
+228].
+Accordingly, researchers have formalized models of responses to cultural items over time as demonstrating either “change” or “no change” to then compared which model fits observed data better, thus functionally asking whether we ever observe within-person change in personal culture.
 
-We argue, however, that asking only *whether* people change obscures what are actually highly compatible results.
-Whether change “really happened” in a given data set depends not only on sample populations, item measurement, statistical methods and power, but crucially also on definitions of what counts as “change,” thus running the risk of getting caught up in unresolvable debates about related decisions.
-[MAYBE BETTER TO SAY HERE: WHICH QUICKLY RELEGATES THE POSSIBILITY OF AN ANSWER BACK TO AN ULTIMATELY UNRESOLVABLE THEORETICAL DEBATE.] Moreover, it narrows the focus on particular questions such as, “Do we see change in this measure?” Such inquiries are ill-suited to address what’s at the core of the theoretical debate.
+We argue, however, that asking only *whether* people change obscures what are actually highly compatible empirical patterns.
+Whether researchers are able to detect "change" in a given data set depends not only on sample populations, item measurement, statistical methods and power, but crucially also on definitions of what counts as “change.” These methodological differences therefore limit the possibility of convergence on the theoretical debates.
+Moreover, asking whether these is evidence of change limits the kind of questions that can be asked to, “Do we see change in this measure?” But such inquiries are ill-suited to address what’s at the core of the theoretical debate.
 There is enough evidence to assume that neither socialization at an early age nor the influence of the contemporary environment alone can explain cultural differences.
@@
 In this paper, we try to move beyond previous debates by moving away from the “tournament of models” approach toward a quantification of the processes that lead to differences in personal culture.
-Drawing on the seven panel data sets analyzed in these debates, we use established methods to quantify the amount of variance in responses that is due to between-people differences in baselines versus the amount of variance in responses that is due to linear within-person change over time.
-This measure gives us a sense of the importance of intrapersonal change over the course of the panel relative to the differences that exist between people when they start.
-We then explore features of questions and populations that explain differences in within-person change over time.
+Drawing on the seven panel data sets analyzed in these debates, we use established methods to quantify the amount of variance in responses that is due to between-people differences at a single point during the survey versus the amount of variance in responses that is due to linear within-person change over time.
+This measure gives us a sense of the importance of intrapersonal change over the course of the panel relative to the differences that exist between people assuming they never changed.
 
@@
 Of course, whether this amount of intrapersonal change is “meaningful” ultimately remains a theoretical question reflected in prior expectations concerning the question, population, and time.
-Nevertheless, the results suggest that the intrapersonal change observed (in some cases over more than 10 years) does not seem to play a substantial role in explaining personal cultural differences between adults.
+Nevertheless, with important exceptions, the results suggest that intrapersonal change (in some cases over more than 10 years) does not seem to play a substantial role in explaining a broad range of personal cultural differences we observe between adults.
 
@@
 What is more, the measure we propose offers a quantity with more diverse applications than a simple judgment about whether change happened.
-To examplify this, we apply the measure to explore differences in the relative importance of intrapersonal change for people with and without college degrees, comparing both within and across questions in ways that the tournament of models cannot.
-
-While adding to existing debates, we hope the quantification we propose extends them in new ways and enables scholars to more clearly adjudicate theoretical mechanisms at play in shaping people's personal culture.
+To exemplify this, we apply the measure to explore differences in the relative importance of intrapersonal change for people with and without college degrees, comparing both within and across questions in ways that the tournament of models cannot.
+The results suggest new directions for further clarifying the theoretical mechanisms that shape personal cultural change over the life course.
 

### 2023-09-23 manuscript.Rmd
background 1

@@
 
-Recent debates about whether adults undergo intrapersonal cultural change emerged in part because sociological models of cultural change at the aggregate level tend to implicitly invoke one of two models of individual behavior.
+Recent debates about whether adults undergo intrapersonal cultural change emerged in part because theories of cultural change at the aggregate level tend to implicitly invoke one of two models of individual behavior.
 The first, what Kiley and Vaisey [-@kiley2020] call a “Settled Dispositions Model” (SDM), assumes that peoples’ personal culture is relatively fixed by the time they are adults.
-While they might make temporary changes in their declarative culture as their social environments provide new considerations, this model assumes that they return to a settled baseline over a (relatively) short time frame.
-This model of behavior is frequently invoked in models of cultural change that suggest people are strongly imprinted by the “past conditions of production” of socialization, such as Bourdieusian practice theories [@bourdieu1990], cohort replacement theories of aggregate change [@mannheim1952; @ryder1965], and control theories in social psychology [@smith-lovin1988; @robinson2007].
+While they might make temporary changes in their declarative culture in reaction to their environments, this model assumes that people return to a settled baseline over a (relatively) short period of time.
+This model is frequently invoked in theories of cultural change that suggest people are strongly imprinted by early socialization experiences such as the “past conditions of production” a Bourdieusian practice theory [@bourdieu1990], cohort replacement theories of aggregate change [@mannheim1952; @ryder1965], and control theories in social psychology [@smith-lovin1988; @robinson2007].
 
 The second model summarized by Kiley and Vaisey [-@kiley2020], an “Active Updating Model” (AUM), posits that people continually update their personal culture as they move through life.
-This model suggests people make new cultural meanings, and therefore durably change their personal culture, as they encounter new social environments, discourses, and information [@gross2009; @swidler2001].
-This model underlies models of cultural diffusion [@christakis2010], attitude alignment [@dellaposta2020], and polarization [@bail2018], among others.
-It is also implicit in most studies of whether specific experiences, changes in social roles, or political events, affect personal culture [@gelman2021; @visser2004; @slothuus2021].
+This model suggests people change their personal culture as they adapt and make new meanings when encountering new social environments, discourses, and information [@gross2009; @swidler2001].
+This model underlies, among others, theories of cultural diffusion [@christakis2010], attitude alignment [@dellaposta2020], and polarization [@bail2018] It is also implicit in most studies that ask whether specific experiences, changes in social roles, or political events, affect personal culture [@gelman2021; @visser2004; @slothuus2021].
 
-There is no reason to think that only one of these models is “correct” in all times and places.
-A population observed over some time frame contains a mix of people who are changing and people who are not (or some people making very small or imperceptible changes in culture and others making perceptible changes).
+There is no reason to believe that only one of these two models is “correct” at all times and in all places.
+A population observed over some period of time contains a mix of people who are noticeably changing and people who are not.
 Instead, different perspectives argue that each of these ideal-typical models is more operative at different times, for different people, and for different elements of personal culture.
-For example, adolescence and early adulthood, typically viewed as a “formative period” for personal cultural development, is likely characterized by higher rates of active updating, while middle age and later life are potentially characterized more by settled dispositions [@alwin1991; @krosnick1989; @eaton2009].
-Similarly, highly salient issues, such as views around gay rights in the 2010s; issues that see substantial elite realignments, such as views around the Vietnam conflict in the 1970s [@zaller1992]; and novel issues of public opinion, such as views around vaccines during the Covid-19 pandemic, might be characterized by active updating, while low-salience, established issues are characterized by stability.
+For example, adolescence and early adulthood is typically viewed as a “formative period” for personal cultural development and thus characterized by higher rates of active updating, while middle age and later life are potentially characterized more by settled dispositions [@alwin1991; @krosnick1989; @eaton2009].
+Similarly, highly salient issues, such as views around gay rights in the 2010s; issues that see substantial elite realignments, such as views around the Vietnam conflict in the 1970s [@zaller1992]; and novel issues of public opinion, such as views around vaccines during the Covid-19 pandemic, might be characterized by active updating, while established issues of low salience are characterized by stability.
 
-Comparing which of these models better fit a broad range of questions from the General Social Survey’s rotating panels, Kiley and Vaisey [-@kiley2020] found limited evidence of durable change.
+Empirically comparing which of these models better fit a broad range of questions from the General Social Survey’s rotating panels, Kiley and Vaisey [-@kiley2020] found limited evidence of durable change.
 While the AUM was preferred for the majority of items, the total amount of durable change detected on these items was small.
@@
 
-However, the claim that change is a relatively infrequent phenomenon has been difficult to square with research identifying durable individual change across a number of cultural dimensions, such as morality [@brocic2021], trust [@mewes2021], and concerns about immigration [@kratz2021], as a result of social experiences.
-Similarly, there are longitudinal studies showing that cues from political elites can change a person’s position on specific policy issues [@slothuus2021; @zaller1992] and that changes in a person’s close contacts and acquaintances can change their attitudes on group-related politics [@gelman2021; @dellaposta2018].
+On the other hand, the claim that change is a relatively infrequent phenomenon has been difficult to square with research identifying durable change as a result of social experiences across a number of cultural dimensions, such as morality [@brocic2021], trust [@mewes2021], and concerns about immigration [@kratz2021].
+Similarly, there are longitudinal studies showing that cues from political elites can change an individual’s position on specific policy issues [@slothuus2021; @zaller1992] and that changes in their close contacts and acquaintances can change individuals' attitudes on group-related politics [@gelman2021; @dellaposta2018].
 Given how often we observe people change their personal culture, it is hard to accept that adults do not change.
 
-Drawing on these findings, Lersch challenged the SDM and AUM as a “needless dichotomy” [-@lersch2023] and proposed a “Life Course Adaption Model” (LCAM), which draws on the life course perspective and models personal culture over the course of a panel as a linear trajectory that varies for each respondent.
+Drawing on these findings, Lersch challenged the SDM and AUM as a “needless dichotomy” [-@lersch2023] and proposed a “Life Course Adaption Model” (LCAM) instead, which draws on the life course perspective and models personal culture over the course of a panel as a linear trajectory that varies for each respondent.
 Lersch’s work addresses two shortcomings of the AUM and SDM comparison.

### 2023-09-23 manuscript.Rmd
Background 2

@@
 
-Drawing on these findings, Lersch challenged the SDM and AUM as a “needless dichotomy” [-@lersch2023] and proposed a “Life Course Adaption Model” (LCAM) instead, which draws on the life course perspective and models personal culture over the course of a panel as a linear trajectory that varies for each respondent.
-Lersch’s work addresses two shortcomings of the AUM and SDM comparison.
-First, the AUM as outlined by Kiley and Vaisey assumes that change can only occur as a Markovian process, in which responses to personal culture items at time $t$ are a function only of responses at time $t-1$.
-But early life experiences might indirectly shape peoples’ personal culture as they move into new social roles or environments, even if they do not affect personal culture right away.
-For example, experiences in childhood might not affect views about family structure until much later in life when people create their own families.
-The LCAM allows responses at earlier points than $t-1$ to influence responses at time $t$.
+Drawing on these findings, Lersch challenged the SDM and AUM as a “needless dichotomy” [-@lersch2023].
+Instead he proposed the “Life Course Adaption Model” (LCAM).
+This model draws on the life course perspective to model personal culture as a linear trajectory over the course of a panel that varies for each respondent.
+Lersch rectifies two shortcomings of the AUM and SDM.
+First, the AUM, as described by Kiley and Vaisey, posits that changes follow a Markovian process where personal culture responses at time $t$ rely solely on responses at time $t-1$.
+Howevre, earlier life experiences can (directly and indirectly) mold personal culture when transitioning to new social roles or into new environments, even if their initial impact is delayed.
+For example, childhood events might influence views on family structures later when individuals form their own families.
+The LCAM considers therefore considers influences from earlier than just $t-1$ on responses at time $t$.
 
-Second, Kiley and Vaisey only adjudicated the AUM and SDM on three-wave panel data covering four years, which might not be a long enough time span to adjudicate the two models they outline.
-In contrast, Lersch tests the LCAM against the AUM and SDM on panels covering a much longer time frame (between 3 and 36 years) with more waves (between 3 and 18 waves), allowing for a greater possibility of detecting durable change.
+Second, Kiley and Vaisey's analysis of the AUM and SDM is based on three-wave panel data over four years, which might not be extensive enough to properly adjudicate [COMPARE?
+WE ARE USING ADJUDICATING A LOT] the two models.
+Lersch, however, evaluates the LCAM against the AUM and SDM using panel data spanning a wider duration (from 3 to 36 years) and more waves (3 to 18), offering a better change to observe durable change.
 
-In his tournament of models, Lersch finds broad support for the LCAM in data from five countries, with the LCAM favored over the AUM and SDM on 297 of 428 questions he analyzed and the SDM favored over the other two models on 112 items, with the remaining questions producing inconclusive results (no questions preferred the AUM).
-He concludes that “new experiences over the life course, which expose individuals to new social environments, including novel interactions, and different sets of resources, can persistently move individuals’ personal culture in novel directions” [@lersch2023: p.
+In Lersch's tournament of models, the LCAM gains broad [EXTENSIVE?] support from data across five countries.
+Out of 428 questions, the LCAM was favored in 297 cases while the SDM was preferred in only 112.
+The rest did not yield inconclusive results, and notably, no questions favoured the AUM.
+He concludes that “new experiences over the life course [...] can persistently move individuals’ personal culture in novel directions” [@lersch2023: p.
 243-244].
 
-Despite differences in how these researchers interpret their findings, the empirical results are not far apart.
-Kiley and Vaisey [-@kiley2020] found that the majority of questions they explored favored the AUM, meaning they showed some evidence of durable change among adults, a finding broadly compatible with Lersch’s results.
-At the same time, while Lersch [-@lersch2023] found that the LCAM was preferred in the majority of cases he examined, durable change does not account for a large proportion of variance explained on these items.
-On average, people changed only about .07 standard deviations over 10 years.
-Again, this is consistent with Kiley and Vaisey’s assertion that durable change is often small.
-And a quarter of the items Lersch studied still preferred the SDM, suggesting that even with looser assumptions about the process of change, more cases, and more waves, many questions show no evidence of durable change.
-Relatedly, in most studies purporting to find evidence of durable change in peoples’ personal culture, the change is often quite small.
-For example, Brocic and Miles [-@brocic2021] find that completing graduate degrees in humanities, arts, and social sciences shift peoples’ moral relativism only about .2 standard deviations on average compared to people with no college degree, and this is the largest effect they identify.
-And studies of aggregate change show that, even on items where cohort effects dominate, there is always evidence that some people change over time [@vaisey2016].
+Despite differing interpretations of these scholars, their empirical results align closely.
+Kiley and Vaisey [-@kiley2020] discovered that most questions they investigated favored the AUM, indicating evidence of durable change among adults -- a finding largely in line with Lersch’s.
+However, while Lersch [-@lersch2023] found a preference for the LCAM in many cases, durable change only accounted for a minor proportion of variance explained.
+On average, individuals changed [CAN WE SAY SHIFTED TO INDICATE THE LINEAR DRIFT] about .07 standard deviations over 10 years, echoing Kiley's and Vaisey’s view that such changes are typically modest.
+Furthermore, a quarter of the items in Lersch's study favoured the SDM, suggesting that despite broader assumptions about change and more extensive data, many questions reveal no evidence of durable change.
+Likewise, in studies claiming evidence of durable change in personal culture, the magnitude is often minimal.
+For instance, Brocic and Miles [-@brocic2021] note that obtaining graduate degrees in humanities, arts, and social sciences only shift [NUDGES?] peoples’ moral relativism by about .2 standard deviations compared to non-degree holders [DELETED THE LAST SENTENCE PART HERE; FELT UNNECESSARY].
+Moreover, studies of aggregate change demonstrate that even on questions where cohort effects prevail, there is always evidence that some individuals change over time [@vaisey2016].
 In other words, despite different interpretations, the results of previous work are in many ways highly consonant.

### 2023-09-23 manuscript.Rmd
improve prominance of edu case in the intro

@@
 
-This pattern of results helps clarify and, hopefully, transcend previous debates.
-What is more, the measure we propose offers a quantity with more diverse applications than a simple judgment about whether change happened.
-To exemplify this, we apply the measure to explore differences in the relative importance of intrapersonal change for people with and without college degrees, comparing both within and across questions in ways that the tournament of models cannot.
-The results suggest new directions for further clarifying the theoretical mechanisms that shape personal cultural change over the life course.
+These results help clarify and, hopefully, transcend previous debates.
+What is more, the measure we propose offers a tool with wider utility than merely discerning durable change that can help advance theoretical debates.
+To exemplify this, we turn to the role of education for cultural differences.
+Specifically, we use our measure to better understand what mechanisms might underlie differences observed between people with and those without college degrees.
+Focusing on eight questions that tab general political dispositions, our measure enables us to compare the relative importance of intrapersonal change for both groups in ways that the tournament of models cannot.
+We find less evidence of durable change among those with college degree.
+This may imply that the college experience predominantly serves as a formative period that solidifies political dispositions, rather than fostering an openness [ADAPTABILITY?] to new information and influences later in the life.
+[MISSING AN ENDING SENTENCE NOW.]
 

### 2023-09-23 manuscript.Rmd
make edu case more prominent in the intro
Background III

@@
 
-The shortcoming of current debates is that they focus on whether people *ever* change, a question that is not theoretically interesting since the answer is almost always “yes.”  As noted previously, if researchers observe enough people for long enough, they will likely observe some of them making durable changes.
-A lack of evidence for change in one question might simply indicate that the question was not asked at a clear enough level of resolution, for a long enough time frame, or to enough people to find evidence of change.
+Current debates often narrow down to ask whether people *ever* change, which is rarely theoretically interesting since the answer is almost always “yes.”  Over sufficient time and in a large enough sample, researchers will likely observe some evidence for durable changes in individuals.
+A lack thereof may simply due to a lack of clarity in the question asked or a low level of resolution in the possible answers, or that the question simply was not asked long enough or to enough people.
 Conversely, finding evidence of change tells us little about how much change is present in a population and how meaningful that change is for social differences.
 
-A more theoretically productive approach is to start with a model that assumes that, in the window of a panel survey, some people might remain stable, some people might change a little, and some people might undergo substantial changes in personal culture.
-Lersch’s LCAM formally does this in modeling each person as following their own trajectory over time.
-But rather than adjudicating whether this model fits the data better than one in which people never change –-asking the question of whether change ever exists-– researchers should quantify the amount that people change over time and compare that change to the differences that exist between people at the beginning of the panel.
-In other words, rather than ask, “do people change?” we should ask, “what are the relative contributions of change and initial differences toward explaining differences in personal culture?”
+A more theoretically productive approach begins with a model assuming that during an observed time period some individuals might remain stable, might change a little, or might undergo significant shifts in their personal culture.
+Lersch’s LCAM formally does this by modeling each individual as following their own trajectory over time.
+But rather than debating whether this model is superior to a model that assumes that people never change, researchers should quantify individual changes over time and compare those to the differences that exist between people at the panel's start.
+In other words, the focus should shift from asking, “do people change?” to asking, “how do change and initial differences collectively explain variation in personal culture?” [I CHANGED THAT QUESTION WORDING, PLEASE CHECK!]
 
-In this approach, there are two numbers that each reflect a combination of relevant theoretical processes.
-First, interpersonal differences at the start of the survey, what Lersch calls “early imprinting,” reflects the accumulated experiences of people prior to entering the panel survey that shape how they answer questions.
-While this is typically associated with a formative period or a settled disposition (hence the name early imprinting), it could also reflect transformative experiences that happen at any age, so long as they occurred before the start of the panel and leave a lasting effect on how people answer questions.
-For example, among people already retired upon entering the panel, their “early imprinting” might reflect this life course transition.
-To some extent, then, this number also reflects variation across people in relevant social roles or statuses at the start of the panel that are important in shaping dispositions.
+In this approach, two metrics reflect distinct theoretical processes.
+First, "early imprinting," as Lersch calls it, denotes interpersonal differences at the start of the panel, capturing the accumulated experiences of people before participating in the panel that influence their answers.
+While commonly associated with experiences during an early formative period that result in settled disposition, this could also reflect transformative experiences [-@paul2014transformative] that happen at any stage in life, as long as they predate the panel and consistently impact subsequent responses.
+For instance, for those people entering the panel post-retirement, this “imprinting” might reflect this pivotal life transition.
+Consequently, this metric also reflects variation in individuals' social roles or statuses at the start of the panel that were important in shaping their dispositions.
 
-The second set of processes, what Lersch calls “persistent change” or “adaption” and Kiley and Vaisey call “active updating,” reflects durable changes in personal culture over time and is captured by the amount of intrapersonal linear change over the course of the survey.
-Lersch suggests that these changes are brought on by social triggers such as moving into a new environment or adopting new social roles.
-They could also reflect the diffusion of new culture across social networks, changes in social network composition (or knowledge about who is in a person's network), new cues from political elites or other cultural leaders, the emergence of new issues in politics or culture, or other large-scale social shifts.
+The second process, termed “persistent change” or “adaption” by Lersch and "active updating" by Kiley and Vaisey, denotes durable changes in personal culture over time, captured by the amount of intrapersonal linear change during the panel.
+Lersch attributes these changes to social triggers such as moving into a new environment or adopting new social roles.
+Additionally, they might reflect the diffusion of new cultural forms across social networks, cues from political elites or otherwise culturally influential leaders, the emergence of issues in politics or culture, or large-scale social shifts.
 
-A third number, often called “residual variance” or “fluctuation,” necessarily accounts for the remainder of variance in peoples’ responses over time.
-This non-durable change potentially emerges for a variety of reasons.
-For example, people might not have clear personal cultural dispositions as measured by surveys.
-Instead, they might internalize a broad set of considerations and construct an opinion or attitude in the context of the survey interview, with different considerations coming to the forefront of their cognition each interview [@tourangeau2000; @zaller1992; @feldman1992].
-This variance in considerations can produce variance in responses even if peoples’ “dispositions” --the full set of considerations they could draw on-- remains stable over time.
-This variance could also include true errors of measurement, such as respondents misunderstanding the question, respondents or administrators selecting the wrong response, or responses getting coded incorrectly.
-While this third process reflects important aspects of personal culture, it does not directly touch on the ongoing debate outlined here.
-Because of this we quantify this component but focus principally on the other two.
+A third metric, often termed “residual variance” or “fluctuation,” represents the remainder of variance in peoples’ responses over time.
+This non-durable change may arise for various reasons.
+Some might lack clear cultural dispositions, leading them to construct an opinion or attitude during the survey, with varying thoughts and considerations dominating their thinking in each interview [@tourangeau2000; @zaller1992; @feldman1992].
+Even though the full set of considerations they could draw on remains stable over time, this could yield to such variance in responses.
+Additionally, this variance can also include measurement errors, such as misinterpreted questions, erroneous response selections, or responses getting coded incorrectly.
+While this third process thus might reflect potentially important aspects of personal culture [I FELT WE NEED TO PLAY IT DOWN A BIT], it is less central to the debate outlined here.
+Hence, we focus on the first two metrics.
 
-We assert that, at the theoretical level, the existence of these two principal processes --interpersonal differences and intrapersonal change-- is indisputable, and that they are, in many ways, related.
-If intrapersonal change occurs at any point in the life course, it will inevitably produce interpersonal differences by the time people enter a panel survey.
-Similarly, unless people are fully socialized early in the life course and never deviate from these dispositions (a proposition against which there is substantial evidence), then we expect to observe intrapersonal change in some proportion of the population whenever we survey them over time.
+At a theoretical level, the coexistence of two main processes -- interpersonal differences and intrapersonal change -- is undeniable, and they are linked in many ways [THOUGHT IF intrinsically linked].
+Any intrapersonal change during one's life will inevitably manifest as interpersonal differences by the time people enter a panel survey.
+Furthermore, unless people are entirely socialized early on and never deviate from these dispositions -- an understanding challenged by evidence -- we expect observing intrapersonal change in a segment of the population when surveyed over time.
 
-Furthermore, these numbers are not necessarily useful in isolation or without a sense of time.
-What is important for the theoretical debate is the relative contribution of these two components in explaining cultural differences in a population.
-In other words, when we look at a population of adults over some time frame, is there enough intrapersonal change relative to interpersonal difference to view it as an important explanation for why people demonstrate cultural differences?
-That is a question that can only be answered by quantifying the two processes.
+Furthermore, these metrics are not necessarily useful in isolation or without context.
+Central to the theoretical debate is how the two processes they capture collectively explain cultural differences in a population.
+In other words, when examining adults across a period of time, does intrapersonal change, compared to interpersonal differences, sufficiently account for the observed cultural differences?
+Only through quantifying the two processes we can get closer to a true answer to this question.
 
-However, the real value in quantification lies not in adjudicating past debates, but in opening up avenues of exploration that were limited under previous approaches.
-For example, classifying questions as either displaying durable change or not obscures differences among questions where there is evidence of durable change, but the relative importance of durable change is different.
-In quantifying the relative contributions of interpersonal differences and intrapersonal change, researchers can compare the relative importance of these processes in explaining differences across societies, across groups within the same population, across items in the same survey, and across time.
+Yet, the true merit of the proposed quantification goes beyond settling past debates; it opens up new avenues for investigations that remained closed under previous approaches.
+For example, merely classifying questions based on whether they show durable change [DO WE NEED "or not"?
+NEVER KNOW THAT AS A NON-NATIVE] overlooks possibly important differences among those that do.
+By quantifying the contributions of interpersonal differences and intrapersonal change, researchers can gauge their relative importance in understanding [INSTEAD explaining] differences among survey items, between groups within a given population, across time, and even across societies.
 

### 2023-09-23 manuscript.Rmd
Background IV -- moving on

@@
 We argue, however, that asking only *whether* people change obscures what are actually highly compatible empirical patterns.
-Whether researchers are able to detect "change" in a given data set depends not only on sample populations, item measurement, statistical methods and power, but crucially also on definitions of what counts as “change.” These methodological differences therefore limit the possibility of convergence on the theoretical debates.
+Whether researchers are able to detect "change" in a given dataset depends not only on sample populations, item measurement, statistical methods and power, but crucially also on definitions of what counts as “change.” These methodological differences therefore limit the possibility of convergence on the theoretical debates.
 Moreover, asking whether these is evidence of change limits the kind of questions that can be asked to, “Do we see change in this measure?” But such inquiries are ill-suited to address what’s at the core of the theoretical debate.
@@
 In this paper, we try to move beyond previous debates by moving away from the “tournament of models” approach toward a quantification of the processes that lead to differences in personal culture.
-Drawing on the seven panel data sets analyzed in these debates, we use established methods to quantify the amount of variance in responses that is due to between-people differences at a single point during the survey versus the amount of variance in responses that is due to linear within-person change over time.
+Drawing on the seven panel datasets analyzed in these debates, we use established methods to quantify the amount of variance in responses that is due to between-people differences at a single point during the survey versus the amount of variance in responses that is due to linear within-person change over time.
 This measure gives us a sense of the importance of intrapersonal change over the course of the panel relative to the differences that exist between people assuming they never changed.
 
-We observe similar patterns across the data sets that have produced conflicting interpretations.
-In all data sets we see similar rates of individual-level change over time across a broad range of measures of personal culture.
+We observe similar patterns across the datasets that have produced conflicting interpretations.
+In all datasets we see similar rates of individual-level change over time across a broad range of measures of personal culture.
 But quantifying these processes also reveals that intrapersonal change plays a substantially smaller role in explaining the variance in peoples’ personal culture than between-person differences in their baselines.
@@
 
-By shifting the question and approach, we expect to see broad similarities across data sets previously used in this analysis.
-First, we expect to observe some level of intrapersonal change in personal culture for most measures of personal culture in each data set.
-However, previous studies found a substantial number of questions (likely between 25 and 35 percent) where there was no evidence of durable change.
-Consistent with this, we expect to observe significant variation in the proportion of variance explained by intrapersonal change, with some questions showing very little variance explained by intrapersonal change.
+By reframing the question and adjusting the approach, we expect consistent patterns across datasets previously analysed in the debate.
+First, we anticipate evidence for some intrapersonal change in most measures of personal culture across datasets.
+Yet, earlier studies found many questions (between 25 and 35 percent) lacking evidence of durable change.
+Aligning wiht this, we expect considerable variation in the variance explained by intrapersonal change, with some questions indicating minimal variance attributable to such change.
+[I KEPT THE "VARIANCE EXPLAINED" PHRASING THOUGH THE "VARIANCE ATTRIBUTABLE" MIGHT BE LESS OF A HARD CLAIM WITHOUT LOSING MUCH?]
 
-Second, we anticipate that the amount of variance explained by intrapersonal change will generally be small relative to differences that exist between people when they start the survey.
-This assertion should not be controversial, regardless of which theoretical process we expect to be operative for a particular question.
-Panel surveys do not cover the full span of peoples’ lives, so even if intrapersonal change plays a large role in explaining cultural differences, all the intrapersonal change observed in a panel survey covering even 20 years is likely not enough to account for the full variance of interpersonal differences.
+Second, we expect the variance explained by intrapersonal to be modest compared to interpersonal differences at the start of the survey.
+This assertion [EXPECTATION?
+PREMISE?] is fairly uncontroversial, regardless of the theoretical process assumed to be most relevant for a particular question.
+Panel surveys do not encompase individuals’ entire lifetimes.
+Thus, even if intrapersonal change explains a large part of cultural differences, the change detected in a panel covering 20 years probably will not fully account for the entire variance of interpersonal differences.
 
-To this point, our expectations are driven principally by what was observed in previous studies.
-However, theoretical perspectives differ in the amount of variance they expect to be attributable to intrapersonal change, and it is here that quantification can provide a new perspective on existing questions.
-“Formative period” and cohort theories that expect early-life socialization to be a strong force in adults’ personal culture would require that the amount of variance attributable to intrapersonal change would be insufficient to explain interpersonal differences even when it is extrapolated to a longer time frame.
-In contrast, other perspectives that attribute more explanatory power to the roles that people currently occupy would expect intrapersonal change to explain a rather large component of variance when extrapolated.
-While the assumption that the amount of change we observe during the panel could be extrapolated beyond the panel window is not necessarily valid, we believe it is the most valid starting point.
+Our expectations so far largely echo what was observed in previous studies.
+However, theoretical perspectives differ in how much variance they attribute to intrapersonal change, enabling quantification to provide a new perspective on existing questions.
+“Formative period” and cohort theories, which emphasize early-life socialization as a strong force in shaping adult personal culture, suggest that the variance from intrapersonal change would still fall short of explaining interpersonal differences, even when extrapolated to a longer time frame [WE COULD DROP HERE A REFERENCE TO THE SPAN..., possibly beyond the panel window].
+Conversely, other theoretical perspectives emphasizing the importance of people's current social roles expect intrapersonal change could account for a large portion of the variance when extrapolated [CAN WE FIND AN ALTERNATIVE TO "EXTRAPOLATE"?].
+While projecting observed panel change beyond the covered time period may have its critics, we see it as a reasonable starting point.
+[DO WE NEED THIS LAST SENTENCE?
+IT FEELS WE ARE MAKING OUR SELVES VULNERABLE TO A CRITIQUE TO NO BENEFIT IN THE ARGUMENT.]
 
-Third, we expect the amount of variance to be explained by each process –interpersonal differences and intrapersonal change– to depend on features of the question and panel.
-Most obviously, we anticipate that the longer we observe respondents, the greater the degree of resolution at which we observe them (i.e., the number of response options for a particular question), and the larger the sample is, the greater share of variance will be explained by intrapersonal change.
-Because the Life Course Adaption model (and the various theoretical perspectives that underlie it) suggests that social experiences drive changes of personal culture, and because social experiences are a function of time, observing people for a longer period increases the chance that each of them will undergo some experience that changes their personal culture.
+Third, we expect the variance explained by each process –interpersonal differences and intrapersonal change– to depend on specific attributes of the question and panel.
+Noteably, the longer we observe respondents, the higher the response resolution for a given question, and the larger the sample, the greater the variance explained by intrapersonal change should be.
+Given that the LCAM (supported by the various underlying theoretical perspectives) posits that social experiences drive changes in personal culture and that these experiences correlate with time, observing people for longer durations [or TIME SPANS?] increases the chance that they undergo experiences that change their personal culture.
+[WE WOULDN'T EVEN NEED THE "they undergo experiences that" IN THAT SENTENCE.]
 
@@
 This paper set out to clarify previous findings about the prevalence of durable change in personal culture -- the attitudes, beliefs, values, preferences, and behaviors that manifest at the individual level.
-Relying on an approach that quantified the amount of systematic variance attributable to either intrapersonal change or interpersonal differences at baseline, we found a consistent pattern across data sets previously used in this debate.
+Relying on an approach that quantified the amount of systematic variance attributable to either intrapersonal change or interpersonal differences at baseline, we found a consistent pattern across datasets previously used in this debate.
 We see evidence of intrapersonal change over time on almost all questions, with some measures of personal culture displaying quite high levels of intrapersonal change relative to the amount of variance between people.

### 2023-09-23 manuscript.Rmd
Discussion I

@@
 
-A recurring focus in the sociology of culture is the question of whether people, as they move through their lives and transition between social roles, change their personal culture –- the declarative and non-declarative attitudes, beliefs, values, practices, and dispositions that exist at the individual level [@lizardo2017; @kiley2020; @lersch2023].
+A recurring focus in the sociology of culture is the question of whether people, as they move through their lives, change their personal culture –- the declarative and non-declarative attitudes, beliefs, values, practices, and dispositions [@lizardo2017; @kiley2020; @lersch2023].
 
@@
 228].
-Accordingly, researchers have formalized models of responses to cultural items over time as demonstrating either “change” or “no change” to then compared which model fits observed data better, thus functionally asking whether we ever observe within-person change in personal culture.
+Accordingly, researchers have formalized models of responses to cultural items over time as demonstrating either “change” or “no change” to then compare which model fits observed data better, thus functionally asking whether we ever observe within-person change in personal culture.
 
@@
 
-This paper set out to clarify previous findings about the prevalence of durable change in personal culture -- the attitudes, beliefs, values, preferences, and behaviors that manifest at the individual level.
-Relying on an approach that quantified the amount of systematic variance attributable to either intrapersonal change or interpersonal differences at baseline, we found a consistent pattern across datasets previously used in this debate.
-We see evidence of intrapersonal change over time on almost all questions, with some measures of personal culture displaying quite high levels of intrapersonal change relative to the amount of variance between people.
-On some questions, such as the factors that affect one’s life satisfaction or attitudes about government benefits, it is highly plausible that differences in peoples’ experiences as adults are the primary explanation for differences in personal culture.
+This paper set out intervene in the debate on whether people change their personal culture -- their attitudes, beliefs, values, practices, and dispositions -- as they move through life.
+Instead of falling back on a "tournament of models" approach [@lersch2023: p.
+228] to call victory for one or the other possible answer to this question, this paper seeks to refocus the debate to the theoretically more productive question of _how_ initial differences between and durable changes within people collectively explain the observed variation in personal culture.
+To do so, it proposes a new measure that quantifies the amount of systematic variance that is attributable to either interpersonal differences at baseline or intrapersonal change.
 
-However, the amount of intrapersonal change is often small relative to interpersonal differences (less than 10 percent on average across all surveys).
-On many questions, the amount of systematic variance attributable to intrapersonal change is essentially zero.
-These include questions about civil liberties, abortion, generalized trust, or sense of civic duty.
-On these questions and the large number of others with a low proportion of systematic variance attributable to intrapersonal change, there does not appear to be sufficient cultural change during adulthood for experiences and transitions between social locations to be the primary explanation for why adults end up with different personal cultures.
-For these items, the explanation for personal cultural differences seems to reside principally in early-life experiences.
+Applying this measure to all datasets previously discussed in this debate revealed a consistent pattern.
+Nearly all questions on personal culture show evidence of intrapersonal change over time, with some items showing noteably high intrapersonal change compared to interpersonal variance.
+For questions about factors that affect one’s life satisfaction or views on government benefits, it seems plausible that differences in adult experiences predominantly account for the observed differences between people.
 
-More important than adjudicating past debates, however, we show that the approach developed in this paper can adjudicate novel theoretical questions about the conditions in which change or stability is more prevalent and more important for explaining cultural differences within a population.
-As an example, we showed that the role of intrapersonal change in explaining cultural differences varies significantly by question and by education.
-Intrapersonal change plays almost no role in explaining general sentiments about the government's role in the economy, while it plays a much larger role in explaining ideological and partisan identification.
-At the same time, across issues, intrapersonal change is much more important for explaining cultural variation among people without a college degree than for people with a college degree.
-While this pattern might be confounded by age and other factors that affect college completion, it potentially suggests that college-going locks in personal culture in a way that other experiences do not, making adult experiences less influential for political dispositions for college graduates compared to the rest of the population.
+Intrapersonal change is often substantially less pronounced than interpersonal differences, accounting for less than 10 percent on average across all surveys.
+On numerous questions, like those on civil liberties, abortion, generalized trust, and civic duty, the systematic variance attributable to intrapersonal change is essentially zero.
+For these and other questions with minimal systematic variance attributable from intrapersonal change, it seems there is not enough cultural change during adulthood to attribute differences in personal culture mainly to experiences and social transitions.
+For these topics, the primary source of persona cultural differences appeasr to stem from early-life experiences.
+
+More important than adjudicating past debates, however, the approach presented in this paper offers insights into new theoretical inquiries regarding when change or stability in personal culture predominates and its significance in explaining cultural variances within a population.
+
+For instance, we detailed how the contribution that intrapersonal change can make in understanding cultural differences varies substantially by question and by individuals' education.
+It plays almost no part in explaining sentiments regarding the government's economic role but is more defining for ideological and partisan identities.
+Intriguingly, intrapersonal change is more important in explaining cultural variation among people without a college compared to college graduates in the U.S..
+While age and other factors related to college completion might confound this pattern, it suggests that college may solidify personal culture to an extent that renders the effect of other adult experiences on political dispositions for graduates less influential relative to individuals without such a college experience.
 

### 2023-09-23 manuscript.Rmd
- Updated figures, added them to repository, and included them into the text (should move them to end of doc); - Went through first part of the discussion.

@@
 
+```{r fig1, out.width = "400px", fig.cap="Proportions of explained variance in items of personal culture; based on predictions at participants' wave mid-point."}
+knitr::include_graphics("fig/variation_share_lm_mdpnt.png")
+```
+
 Across all questions, interpersonal differences explain a much larger share of the systematic variance in responses than intrapersonal change.
@@
 [^paneleffects]: KK NOTE: Can we summarize these panel fixed effects at all?
+
+```{r fig2, out.width = "300px", fig.cap="Proportion of explained variance attributable to intrapersonal change; based on predictions at participants' wave mid-point; supressed intercept model; survey indicators not shown."}
+knitr::include_graphics("fig/meta_no_int.png")
+```
 
@@
 
-Figure XXX plots the distribution of differences between people with less than a bachelor's degree and people with at least a bachelor's degree at wave 1 of the panel.
+Figure 3 plots the distribution of differences between people with less than a bachelor's degree and people with at least a bachelor's degree at wave 1 of the panel.
 Values greater than 0 would indicate that intrapersonal change is more important in explaining the responses of college graduates than those of people without a college degree, while values less than 0 would indicate that intrapersonal change is less important for explaining variation among people with a college degree than those without a college degree.
+
+```{r img, out.width = "450px", fig.cap="Proportions of intrapersonal change of explained variance by college degree; based on predictions at participants' wave mid-point."}
+knitr::include_graphics("fig/FIG_college2.png")
+```
 
@@
 Applying this measure to all datasets previously discussed in this debate revealed a consistent pattern.
-Nearly all questions on personal culture show evidence of intrapersonal change over time, with some items showing noteably high intrapersonal change compared to interpersonal variance.
+Nearly all questions on personal culture show evidence of intrapersonal change over time, with some items showing noteably high intrapersonal change compared to interpersonal differences.
 For questions about factors that affect one’s life satisfaction or views on government benefits, it seems plausible that differences in adult experiences predominantly account for the observed differences between people.
@@
 Intrapersonal change is often substantially less pronounced than interpersonal differences, accounting for less than 10 percent on average across all surveys.
-On numerous questions, like those on civil liberties, abortion, generalized trust, and civic duty, the systematic variance attributable to intrapersonal change is essentially zero.
-For these and other questions with minimal systematic variance attributable from intrapersonal change, it seems there is not enough cultural change during adulthood to attribute differences in personal culture mainly to experiences and social transitions.
-For these topics, the primary source of persona cultural differences appeasr to stem from early-life experiences.
+On numerous questions, like those on civil liberties, abortion, generalized trust, and civic duty, the systematic variance explained by intrapersonal change is essentially zero.
+For these questions, it seems there is not enough cultural change during adulthood to warrant attributing differences in personal culture mainly to experiences and social transitions; instead, the primary source of the observed differences appears to stem from early-life experiences.
 
-More important than adjudicating past debates, however, the approach presented in this paper offers insights into new theoretical inquiries regarding when change or stability in personal culture predominates and its significance in explaining cultural variances within a population.
-
-For instance, we detailed how the contribution that intrapersonal change can make in understanding cultural differences varies substantially by question and by individuals' education.
-It plays almost no part in explaining sentiments regarding the government's economic role but is more defining for ideological and partisan identities.
-Intriguingly, intrapersonal change is more important in explaining cultural variation among people without a college compared to college graduates in the U.S..
+More important than adjudicating past debates, however, the approach presented in this paper offers offers the possibility to pose new theoretical questions regarding when change or stability in personal culture predominates and its significance in explaining cultural variances within a population.
+To examplify this potential, we detailed how the proportion in cultural differences that intrapersonal change can explain not only varies substantially by survey item but also by individual characteristics such as their education.
+Intrapersonal change plays almost no part in explaining sentiments regarding the government's economic role but is more defining for ideological and partisan identities.
+Intriguingly, it is more important in explaining cultural variation among people without a college compared to college graduates in the U.S..
 While age and other factors related to college completion might confound this pattern, it suggests that college may solidify personal culture to an extent that renders the effect of other adult experiences on political dispositions for graduates less influential relative to individuals without such a college experience.
@@
 
-To this point, the theoretical debate in cultural sociology has focused on the existence of personal cultural change among adults and its relative importance.
-We believe the "existence" debate is theoretically misguided.
-Personal cultural change will always exist to some degree in an adult population, and no viable theoretical model expects just one process --stability or change-- to characterize a population.
-Our results reinforce that assertion, finding a mixture of interpersonal differences and intrapersonal change for all items.
-Even when we observe very low levels of systematic variance attributable to intrapersonal change for particular questions, this is potentially attributable to insufficient data, insufficient resolution on the underlying belief, or insufficient duration of observation.
+So far, the debate in cultural sociology has centered around the presence and significance of cultural change in adults.
+Yet, arguing solely over its existence sets the wrong theoretical focus.
+In any adult population, some degree of personal cultural change is inevitable.
+Likewise, no viable theoretical perspective would expect just one process --stable differences or internal change between people-- to characterize the differences in personal cultural observed between them.
+Our results underscore this perspective, showing a combination of interpersonal differences and intrapersonal change across all items.
+[DELETED SENTECE, MISSPLACED]
 
 A more reasonable theoretical debate pushed by previous work is the relative importance of these two processes for explaining cultural variation.
-While change might always be prevalent, its impact on social differences might be small or large, and our approach has been designed to clarify the contribution of intrapersonal change to such differences.
 It is difficult to draw a single conclusion on this question across a wide range of questions measured on different scales and over different durations of time, but the overall picture, with important caveats, suggests that intrapersonal change during adulthood is not the dominant explanation for cultural differences among adults for most items we measure in surveys.

### 2023-09-23 manuscript.Rmd
Added Apendix; updated figures; adjusted positioning of sections

@@
 
-```{r, message = FALSE}
-
-library(tidyverse)
-library(knitr)
-library(kableExtra)
-
-table <- tibble(
-  Country = c("Australia", "Germany", "Great Britian", "Great Britian", "Switzerland", "United States", "United States"),
-  Survey = c("Household, Income and Labor Dynamics in Australia (HILDA)",
-             "Socio-Economic Panel Study (SOEP)",
-             "British Household Panel Survey (BHPS)",
-             "Understanding Society/UK Household Longitudinal Study (UKHLS)",
-             "Swiss Household Panel (SHP)",
-             "General Social Survey (GSS)",
-             "Panel Study of Income Dynamics (PSID)"),
-  Period = c("2001-2021", "1984-2020", "1991-2008", "2009-2020", "1999-2019", "2006-2012", "1968-2019"),
-  Outcomes = c(30, 122, 83, 66, 77, 183, 48)
-)
-
-kableExtra::kable(table, 
-                  format = "latex", 
-                  caption = "The Description of the Data Sources") |> 
-  kable_styling(latex_options = c("scale_down", "HOLD_position"))
-
-```
+\begin{center}
+<about here> 
+Table 1: The Description of the Data Sources
+\end{center}
 
@@
 
-```{r fig1, out.width = "400px", fig.cap="Proportions of explained variance in items of personal culture; based on predictions at participants' wave mid-point."}
-knitr::include_graphics("fig/variation_share_lm_mdpnt.png")
-```
+\begin{center}
+<about here>
+Figure 1: Proportions of explained variance in items of personal culture; based on predictions at participants' wave mid-point.
+\end{center}
 
@@
 
-```{r fig2, out.width = "300px", fig.cap="Proportion of explained variance attributable to intrapersonal change; based on predictions at participants' wave mid-point; supressed intercept model; survey indicators not shown."}
-knitr::include_graphics("fig/meta_no_int.png")
-```
+\begin{center}
+<about here>
+Figure 2: Proportion of explained variance attributable to intrapersonal change; based on predictions at participants' wave mid-point; supressed intercept model; survey indicators not shown.
+\end{center}
 
@@
 
-```{r img, out.width = "450px", fig.cap="Proportions of intrapersonal change of explained variance by college degree; based on predictions at participants' wave mid-point."}
-knitr::include_graphics("fig/FIG_college2.png")
-```
+\begin{center}
+<about here>
+Figure 3: Proportions of intrapersonal change of explained variance by college degree; based on predictions at participants' wave mid-point.
+\end{center}
 
@@
 # References
+
+<div id="refs"></div>
+
+\newpage
+
+# Appendix
+
+## Tables
+
+```{r, message = FALSE}
+
+library(tidyverse)
+library(knitr)
+library(kableExtra)
+
+table <- tibble(
+  Country = c("Australia", "Germany", "Great Britian", "Great Britian", "Switzerland", "United States", "United States"),
+  Survey = c("Household, Income and Labor Dynamics in Australia (HILDA)",
+             "Socio-Economic Panel Study (SOEP)",
+             "British Household Panel Survey (BHPS)",
+             "Understanding Society/UK Household Longitudinal Study (UKHLS)",
+             "Swiss Household Panel (SHP)",
+             "General Social Survey (GSS)",
+             "Panel Study of Income Dynamics (PSID)"),
+  Period = c("2001-2021", "1984-2020", "1991-2008", "2009-2020", "1999-2019", "2006-2012", "1968-2019"),
+  Outcomes = c(30, 122, 83, 66, 77, 183, 48)
+)
+
+kableExtra::kable(table, 
+                  format = "latex", 
+                  caption = "The Description of the Data Sources") |> 
+  kable_styling(latex_options = c("scale_down", "HOLD_position"))
+
+```
+
+\newpage
+
+## Figures
+
+```{r fig1, fig.height=2, fig.cap="Proportions of explained variance in items of personal culture; based on predictions at participants' wave mid-point."}
+knitr::include_graphics("fig/variation_share_lm_mdpnt.png")
+```
+
+\newpage
+```{r fig2, out.width = "300px", fig.cap="Proportion of explained variance attributable to intrapersonal change; based on predictions at participants' wave mid-point; supressed intercept model; survey indicators not shown."}
+knitr::include_graphics("fig/meta_no_int.png")
+```
+
+\newpage
+```{r fig3, out.width = "450px", fig.cap="Proportions of intrapersonal change of explained variance by college degree; based on predictions at participants' wave mid-point."}
+knitr::include_graphics("fig/FIG_college2.png")
+```

### 2023-09-24 manuscript.Rmd

@@
 
-A recurring focus in the sociology of culture is the question of whether people, as they move through their lives, change their personal culture –- the declarative and non-declarative attitudes, beliefs, values, practices, and dispositions [@lizardo2017; @kiley2020; @lersch2023].
+A recurring focus in the sociology of culture is the question of whether people, as they move through their lives and transition between social roles, change their personal culture –- the declarative and non-declarative attitudes, beliefs, values, practices, and dispositions that exist at the individual level [@lizardo2017; @kiley2020; @lersch2023].
 

### 2023-09-24 manuscript.tex

--- 2023-09-23 manuscript.Rmd
@@
----
-title: "Quantifying Stability and Change in Personal Culture Using Panel Data"
-author:
-- 
-output:
-  pdf_document:
-    keep_tex: true
-bibliography: kileybib.bib
-csl: asa.csl
-abstract: "Recent work has produced conflicting interpretations on the question of whether adults undergo durable changes in their personal culture over time.
-This paper asserts that this divergence is less due to researchers' findings than the approach taken -- a tournament of models focusing on the presence or absence of change.
-To move beyond this, we reanalyze the data used previously in this debate with an approach that quantifies the processes of cultural difference.
-In doing so, we compare how much cultural difference is explained by intrapersonal change and interpersonal differences.
-Our results harmonize recent findings by showing that, while most measures of personal culture show change over time, the variance explained by intrapersonal change is typically small compared to interpersonal differences at baseline.
-Quantifying these processes offers new ways to address theoretical questions of when and how their relative importance shifts.
-We examplify this with the case of college experience."
-fontsize: 11pt
----
+% Options for packages loaded elsewhere
+\PassOptionsToPackage{unicode}{hyperref}
+\PassOptionsToPackage{hyphens}{url}
+%
+\documentclass[
+  11pt,
+]{article}
+\usepackage{amsmath,amssymb}
+\usepackage{lmodern}
+\usepackage{iftex}
+\ifPDFTeX
+  \usepackage[T1]{fontenc}
+  \usepackage[utf8]{inputenc}
+  \usepackage{textcomp} % provide euro and other symbols
+\else % if luatex or xetex
+  \usepackage{unicode-math}
+  \defaultfontfeatures{Scale=MatchLowercase}
+  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
+\fi
+% Use upquote if available, for straight quotes in verbatim environments
+\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
+\IfFileExists{microtype.sty}{% use microtype if available
+  \usepackage[]{microtype}
+  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
+}{}
+\makeatletter
+\@ifundefined{KOMAClassName}{% if non-KOMA class
+  \IfFileExists{parskip.sty}{%
+    \usepackage{parskip}
+  }{% else
+    \setlength{\parindent}{0pt}
+    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
+}{% if KOMA class
+  \KOMAoptions{parskip=half}}
+\makeatother
+\usepackage{xcolor}
+\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
+\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
+\hypersetup{
+  pdftitle={Quantifying Stability and Change in Personal Culture Using Panel Data},
+  hidelinks,
+  pdfcreator={LaTeX via pandoc}}
+\urlstyle{same} % disable monospaced font for URLs
+\usepackage[margin=1in]{geometry}
+\usepackage{graphicx}
+\makeatletter
+\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
+\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
+\makeatother
+% Scale images if necessary, so that they will not overflow the page
+% margins by default, and it is still possible to overwrite the defaults
+% using explicit options in \includegraphics[width, height, ...]{}
+\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
+% Set default figure placement to htbp
+\makeatletter
+\def\fps@figure{htbp}
+\makeatother
+\setlength{\emergencystretch}{3em} % prevent overfull lines
+\providecommand{\tightlist}{%
+  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
+\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
+\newlength{\cslhangindent}
+\setlength{\cslhangindent}{1.5em}
+\newlength{\csllabelwidth}
+\setlength{\csllabelwidth}{3em}
+\newlength{\cslentryspacingunit} % times entry-spacing
+\setlength{\cslentryspacingunit}{\parskip}
+\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
+ {% don't indent paragraphs
+  \setlength{\parindent}{0pt}
+  % turn on hanging indent if param 1 is 1
+  \ifodd #1
+  \let\oldpar\par
+  \def\par{\hangindent=\cslhangindent\oldpar}
+  \fi
+  % set entry spacing
+  \setlength{\parskip}{#2\cslentryspacingunit}
+ }%
+ {}
+\usepackage{calc}
+\newcommand{\CSLBlock}[1]{#1\hfill\break}
+\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
+\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}
+\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}
+\usepackage{booktabs}
+\usepackage{longtable}
+\usepackage{array}
+\usepackage{multirow}
+\usepackage{wrapfig}
+\usepackage{float}
+\usepackage{colortbl}
+\usepackage{pdflscape}
+\usepackage{tabu}
+\usepackage{threeparttable}
+\usepackage{threeparttablex}
+\usepackage[normalem]{ulem}
+\usepackage{makecell}
+\usepackage{xcolor}
+\ifLuaTeX
+  \usepackage{selnolig}  % disable illegal ligatures
+\fi
 
-```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = F, message = F, warning = F, fig.align = "center")
-```
+\title{Quantifying Stability and Change in Personal Culture Using Panel
+Data}
+\author{}
+\date{\vspace{-2.5em}}
 
-# Introduction
+\begin{document}
+\maketitle
+\begin{abstract}
+Recent work has produced conflicting interpretations on the question of
+whether adults undergo durable changes in their personal culture over
+time. This paper asserts that this divergence is less due to
+researchers' findings than the approach taken -- a tournament of models
+focusing on the presence or absence of change. To move beyond this, we
+reanalyze the data used previously in this debate with an approach that
+quantifies the processes of cultural difference. In doing so, we compare
+how much cultural difference is explained by intrapersonal change and
+interpersonal differences. Our results harmonize recent findings by
+showing that, while most measures of personal culture show change over
+time, the variance explained by intrapersonal change is typically small
+compared to interpersonal differences at baseline. Quantifying these
+processes offers new ways to address theoretical questions of when and
+how their relative importance shifts. We examplify this with the case of
+college experience.
+\end{abstract}
 
-A recurring focus in the sociology of culture is the question of whether people, as they move through their lives, change their personal culture –- the declarative and non-declarative attitudes, beliefs, values, practices, and dispositions [@lizardo2017; @kiley2020; @lersch2023].
+\hypertarget{introduction}{%
+\section{Introduction}\label{introduction}}
 
-This seemingly simple question underlies a number of central debates across sociological subfields.
-For example, pragmatist theories of action claim that changes in social environments cause people to adapt their views and make new meanings [@swidler2001; @gross2009], while Bourdieusian practice theories argue that the “past conditions of production” leave a mark on people's personal culture that lasts throughout their lives [@bourdieu1990].
-Similarly, models of social influence assume that people adapt their culture in the face of new information [@vaisey2010a], while the emphasis on cohort effects in models of aggregate social and cultural change requires them to be open to change while young but become fairly resistant to it as they age [@ryder1965].
-Finally, life course theories posit slower but still important changes over time as people advance through important transitions in their lives [@elder2016; @bardi2011].
+A recurring focus in the sociology of culture is the question of whether
+people, as they move through their lives, change their personal culture
+--- the declarative and non-declarative attitudes, beliefs, values,
+practices, and dispositions (Kiley and Vaisey 2020; Lersch 2023; Lizardo
+2017).
 
-Because all of the above theoretical perspectives have some empirical support, the current theoretical debate revolves around not whether they exist, but when and how they each apply and the relative contribution of each to explaining cultural differences.
-That being said, scholars have struggled to reach a consensus on the relative importance of change during adulthood.
-Over the span of a few years, most apparent changes in people’s responses to personal culture items appear to be transitory, with little evidence of persistent change among adults [@kiley2020; @vaisey2021].
-These results suggest change is not a strong force in explaining personal cultural differences.
-When looked at over longer time horizons, however, there is at least some evidence that adults make persistent cultural changes, especially as they move through distinct social roles across the life course [@lersch2023].
+This seemingly simple question underlies a number of central debates
+across sociological subfields. For example, pragmatist theories of
+action claim that changes in social environments cause people to adapt
+their views and make new meanings (Gross 2009; Swidler 2001), while
+Bourdieusian practice theories argue that the ``past conditions of
+production'' leave a mark on people's personal culture that lasts
+throughout their lives (Bourdieu 1990). Similarly, models of social
+influence assume that people adapt their culture in the face of new
+information (Vaisey and Lizardo 2010), while the emphasis on cohort
+effects in models of aggregate social and cultural change requires them
+to be open to change while young but become fairly resistant to it as
+they age (Ryder 1965). Finally, life course theories posit slower but
+still important changes over time as people advance through important
+transitions in their lives (Bardi and Goodwin 2011; Elder and George
+2016).
 
-These seemingly inconsistent findings are partly due to the fact that researchers have taken a “tournament of models” approach in adjudicating the different perspectives [@lersch2023: p.
-228].
-Accordingly, researchers have formalized models of responses to cultural items over time as demonstrating either “change” or “no change” to then compare which model fits observed data better, thus functionally asking whether we ever observe within-person change in personal culture.
+Because all of the above theoretical perspectives have some empirical
+support, the current theoretical debate revolves around not whether they
+exist, but when and how they each apply and the relative contribution of
+each to explaining cultural differences. That being said, scholars have
+struggled to reach a consensus on the relative importance of change
+during adulthood. Over the span of a few years, most apparent changes in
+people's responses to personal culture items appear to be transitory,
+with little evidence of persistent change among adults (Kiley and Vaisey
+2020; Vaisey and Kiley 2021). These results suggest change is not a
+strong force in explaining personal cultural differences. When looked at
+over longer time horizons, however, there is at least some evidence that
+adults make persistent cultural changes, especially as they move through
+distinct social roles across the life course (Lersch 2023).
 
-We argue, however, that asking only *whether* people change obscures what are actually highly compatible empirical patterns.
-Whether researchers are able to detect "change" in a given dataset depends not only on sample populations, item measurement, statistical methods and power, but crucially also on definitions of what counts as “change.” These methodological differences therefore limit the possibility of convergence on the theoretical debates.
-Moreover, asking whether these is evidence of change limits the kind of questions that can be asked to, “Do we see change in this measure?” But such inquiries are ill-suited to address what’s at the core of the theoretical debate.
-There is enough evidence to assume that neither socialization at an early age nor the influence of the contemporary environment alone can explain cultural differences.
-The question that remains is under what conditions, across which groups, and in what domains do we observe differences in the relative importance of intrapersonal change and interpersonal differences.
-This focus is necessarily comparative and requires an approach that quantifies forms of cultural change rather than calling victory for the one or the other perspective.
+These seemingly inconsistent findings are partly due to the fact that
+researchers have taken a ``tournament of models'' approach in
+adjudicating the different perspectives (Lersch 2023: p.~228).
+Accordingly, researchers have formalized models of responses to cultural
+items over time as demonstrating either ``change'' or ``no change'' to
+then compare which model fits observed data better, thus functionally
+asking whether we ever observe within-person change in personal culture.
 
-In this paper, we try to move beyond previous debates by moving away from the “tournament of models” approach toward a quantification of the processes that lead to differences in personal culture.
-Drawing on the seven panel datasets analyzed in these debates, we use established methods to quantify the amount of variance in responses that is due to between-people differences at a single point during the survey versus the amount of variance in responses that is due to linear within-person change over time.
-This measure gives us a sense of the importance of intrapersonal change over the course of the panel relative to the differences that exist between people assuming they never changed.
+We argue, however, that asking only \emph{whether} people change
+obscures what are actually highly compatible empirical patterns. Whether
+researchers are able to detect ``change'' in a given dataset depends not
+only on sample populations, item measurement, statistical methods and
+power, but crucially also on definitions of what counts as ``change.''
+These methodological differences therefore limit the possibility of
+convergence on the theoretical debates. Moreover, asking whether these
+is evidence of change limits the kind of questions that can be asked to,
+``Do we see change in this measure?'' But such inquiries are ill-suited
+to address what's at the core of the theoretical debate. There is enough
+evidence to assume that neither socialization at an early age nor the
+influence of the contemporary environment alone can explain cultural
+differences. The question that remains is under what conditions, across
+which groups, and in what domains do we observe differences in the
+relative importance of intrapersonal change and interpersonal
+differences. This focus is necessarily comparative and requires an
+approach that quantifies forms of cultural change rather than calling
+victory for the one or the other perspective.
 
-We observe similar patterns across the datasets that have produced conflicting interpretations.
-In all datasets we see similar rates of individual-level change over time across a broad range of measures of personal culture.
-But quantifying these processes also reveals that intrapersonal change plays a substantially smaller role in explaining the variance in peoples’ personal culture than between-person differences in their baselines.
-Of course, whether this amount of intrapersonal change is “meaningful” ultimately remains a theoretical question reflected in prior expectations concerning the question, population, and time.
-Nevertheless, with important exceptions, the results suggest that intrapersonal change (in some cases over more than 10 years) does not seem to play a substantial role in explaining a broad range of personal cultural differences we observe between adults.
+In this paper, we try to move beyond previous debates by moving away
+from the ``tournament of models'' approach toward a quantification of
+the processes that lead to differences in personal culture. Drawing on
+the seven panel datasets analyzed in these debates, we use established
+methods to quantify the amount of variance in responses that is due to
+between-people differences at a single point during the survey versus
+the amount of variance in responses that is due to linear within-person
+change over time. This measure gives us a sense of the importance of
+intrapersonal change over the course of the panel relative to the
+differences that exist between people assuming they never changed.
+
+We observe similar patterns across the datasets that have produced
+conflicting interpretations. In all datasets we see similar rates of
+individual-level change over time across a broad range of measures of
+personal culture. But quantifying these processes also reveals that
+intrapersonal change plays a substantially smaller role in explaining
+the variance in peoples' personal culture than between-person
+differences in their baselines. Of course, whether this amount of
+intrapersonal change is ``meaningful'' ultimately remains a theoretical
+question reflected in prior expectations concerning the question,
+population, and time. Nevertheless, with important exceptions, the
+results suggest that intrapersonal change (in some cases over more than
+10 years) does not seem to play a substantial role in explaining a broad
+range of personal cultural differences we observe between adults.
 
 These results help clarify and, hopefully, transcend previous debates.
-What is more, the measure we propose offers a tool with wider utility than merely discerning durable change that can help advance theoretical debates.
-To exemplify this, we turn to the role of education for cultural differences.
-Specifically, we use our measure to better understand what mechanisms might underlie differences observed between people with and those without college degrees.
-Focusing on eight questions that tab general political dispositions, our measure enables us to compare the relative importance of intrapersonal change for both groups in ways that the tournament of models cannot.
-We find less evidence of durable change among those with college degree.
-This may imply that the college experience predominantly serves as a formative period that solidifies political dispositions, rather than fostering an openness [ADAPTABILITY?] to new information and influences later in the life.
-[MISSING AN ENDING SENTENCE NOW.]
+What is more, the measure we propose offers a tool with wider utility
+than merely discerning durable change that can help advance theoretical
+debates. To exemplify this, we turn to the role of education for
+cultural differences. Specifically, we use our measure to better
+understand what mechanisms might underlie differences observed between
+people with and those without college degrees. Focusing on eight
+questions that tab general political dispositions, our measure enables
+us to compare the relative importance of intrapersonal change for both
+groups in ways that the tournament of models cannot. We find less
+evidence of durable change among those with college degree. This may
+imply that the college experience predominantly serves as a formative
+period that solidifies political dispositions, rather than fostering an
+openness {[}ADAPTABILITY?{]} to new information and influences later in
+the life. {[}MISSING AN ENDING SENTENCE NOW.{]}
 
-# Background
+\hypertarget{background}{%
+\section{Background}\label{background}}
 
-## Change and Stability in Personal Culture
+\hypertarget{change-and-stability-in-personal-culture}{%
+\subsection{Change and Stability in Personal
+Culture}\label{change-and-stability-in-personal-culture}}
 
-Recent debates about whether adults undergo intrapersonal cultural change emerged in part because theories of cultural change at the aggregate level tend to implicitly invoke one of two models of individual behavior.
-The first, what Kiley and Vaisey [-@kiley2020] call a “Settled Dispositions Model” (SDM), assumes that peoples’ personal culture is relatively fixed by the time they are adults.
-While they might make temporary changes in their declarative culture in reaction to their environments, this model assumes that people return to a settled baseline over a (relatively) short period of time.
-This model is frequently invoked in theories of cultural change that suggest people are strongly imprinted by early socialization experiences such as the “past conditions of production” a Bourdieusian practice theory [@bourdieu1990], cohort replacement theories of aggregate change [@mannheim1952; @ryder1965], and control theories in social psychology [@smith-lovin1988; @robinson2007].
+Recent debates about whether adults undergo intrapersonal cultural
+change emerged in part because theories of cultural change at the
+aggregate level tend to implicitly invoke one of two models of
+individual behavior. The first, what Kiley and Vaisey (2020) call a
+``Settled Dispositions Model'' (SDM), assumes that peoples' personal
+culture is relatively fixed by the time they are adults. While they
+might make temporary changes in their declarative culture in reaction to
+their environments, this model assumes that people return to a settled
+baseline over a (relatively) short period of time. This model is
+frequently invoked in theories of cultural change that suggest people
+are strongly imprinted by early socialization experiences such as the
+``past conditions of production'' a Bourdieusian practice theory
+(Bourdieu 1990), cohort replacement theories of aggregate change
+(Mannheim 1952; Ryder 1965), and control theories in social psychology
+(Robinson 2007; Smith-Lovin and Heise 1988).
 
-The second model summarized by Kiley and Vaisey [-@kiley2020], an “Active Updating Model” (AUM), posits that people continually update their personal culture as they move through life.
-This model suggests people change their personal culture as they adapt and make new meanings when encountering new social environments, discourses, and information [@gross2009; @swidler2001].
-This model underlies, among others, theories of cultural diffusion [@christakis2010], attitude alignment [@dellaposta2020], and polarization [@bail2018] It is also implicit in most studies that ask whether specific experiences, changes in social roles, or political events, affect personal culture [@gelman2021; @visser2004; @slothuus2021].
+The second model summarized by Kiley and Vaisey (2020), an ``Active
+Updating Model'' (AUM), posits that people continually update their
+personal culture as they move through life. This model suggests people
+change their personal culture as they adapt and make new meanings when
+encountering new social environments, discourses, and information (Gross
+2009; Swidler 2001). This model underlies, among others, theories of
+cultural diffusion (Christakis and Fowler 2010), attitude alignment
+(DellaPosta 2020), and polarization (Bail et al. 2018) It is also
+implicit in most studies that ask whether specific experiences, changes
+in social roles, or political events, affect personal culture (Gelman
+and Margalit 2021; Slothuus and Bisgaard 2021; Visser and Mirabile
+2004).
 
-There is no reason to believe that only one of these two models is “correct” at all times and in all places.
-A population observed over some period of time contains a mix of people who are noticeably changing and people who are not.
-Instead, different perspectives argue that each of these ideal-typical models is more operative at different times, for different people, and for different elements of personal culture.
-For example, adolescence and early adulthood is typically viewed as a “formative period” for personal cultural development and thus characterized by higher rates of active updating, while middle age and later life are potentially characterized more by settled dispositions [@alwin1991; @krosnick1989; @eaton2009].
-Similarly, highly salient issues, such as views around gay rights in the 2010s; issues that see substantial elite realignments, such as views around the Vietnam conflict in the 1970s [@zaller1992]; and novel issues of public opinion, such as views around vaccines during the Covid-19 pandemic, might be characterized by active updating, while established issues of low salience are characterized by stability.
+There is no reason to believe that only one of these two models is
+``correct'' at all times and in all places. A population observed over
+some period of time contains a mix of people who are noticeably changing
+and people who are not. Instead, different perspectives argue that each
+of these ideal-typical models is more operative at different times, for
+different people, and for different elements of personal culture. For
+example, adolescence and early adulthood is typically viewed as a
+``formative period'' for personal cultural development and thus
+characterized by higher rates of active updating, while middle age and
+later life are potentially characterized more by settled dispositions
+(Alwin and Krosnick 1991; Eaton et al. 2009; Krosnick and Alwin 1989).
+Similarly, highly salient issues, such as views around gay rights in the
+2010s; issues that see substantial elite realignments, such as views
+around the Vietnam conflict in the 1970s (Zaller 1992); and novel issues
+of public opinion, such as views around vaccines during the Covid-19
+pandemic, might be characterized by active updating, while established
+issues of low salience are characterized by stability.
 
-Empirically comparing which of these models better fit a broad range of questions from the General Social Survey’s rotating panels, Kiley and Vaisey [-@kiley2020] found limited evidence of durable change.
-While the AUM was preferred for the majority of items, the total amount of durable change detected on these items was small.
-A substantial minority of questions (39 percent) favored the SDM, meaning they showed no evidence of durable change.
-Questions with more evidence of durable change included highly salient issues like gay marriage and “public” issues like partisan identification and religious service attendance.
-There was also more evidence of durable change among early adults (people ages 18-30) than the rest of the population.
-Overall, the researchers concluded that “results ultimately suggest that real, persistent attitude change is an uncommon phenomenon among adults'' [@kiley2020: p.
-500; see also @vaisey2021].
-This lack of durable change is consistent with other findings that cohort replacement plays a larger role than period effects in explaining differences in personal culture [@vaisey2016].
+Empirically comparing which of these models better fit a broad range of
+questions from the General Social Survey's rotating panels, Kiley and
+Vaisey (2020) found limited evidence of durable change. While the AUM
+was preferred for the majority of items, the total amount of durable
+change detected on these items was small. A substantial minority of
+questions (39 percent) favored the SDM, meaning they showed no evidence
+of durable change. Questions with more evidence of durable change
+included highly salient issues like gay marriage and ``public'' issues
+like partisan identification and religious service attendance. There was
+also more evidence of durable change among early adults (people ages
+18-30) than the rest of the population. Overall, the researchers
+concluded that ``results ultimately suggest that real, persistent
+attitude change is an uncommon phenomenon among adults'\,' (Kiley and
+Vaisey 2020: p.~500; see also Vaisey and Kiley 2021). This lack of
+durable change is consistent with other findings that cohort replacement
+plays a larger role than period effects in explaining differences in
+personal culture (Vaisey and Lizardo 2016).
 
-On the other hand, the claim that change is a relatively infrequent phenomenon has been difficult to square with research identifying durable change as a result of social experiences across a number of cultural dimensions, such as morality [@brocic2021], trust [@mewes2021], and concerns about immigration [@kratz2021].
-Similarly, there are longitudinal studies showing that cues from political elites can change an individual’s position on specific policy issues [@slothuus2021; @zaller1992] and that changes in their close contacts and acquaintances can change individuals' attitudes on group-related politics [@gelman2021; @dellaposta2018].
-Given how often we observe people change their personal culture, it is hard to accept that adults do not change.
+On the other hand, the claim that change is a relatively infrequent
+phenomenon has been difficult to square with research identifying
+durable change as a result of social experiences across a number of
+cultural dimensions, such as morality (Broćić and Miles 2021), trust
+(Mewes et al. 2021), and concerns about immigration (Kratz 2021).
+Similarly, there are longitudinal studies showing that cues from
+political elites can change an individual's position on specific policy
+issues (Slothuus and Bisgaard 2021; Zaller 1992) and that changes in
+their close contacts and acquaintances can change individuals' attitudes
+on group-related politics (DellaPosta 2018; Gelman and Margalit 2021).
+Given how often we observe people change their personal culture, it is
+hard to accept that adults do not change.
 
-Drawing on these findings, Lersch challenged the SDM and AUM as a “needless dichotomy” [-@lersch2023].
-Instead he proposed the “Life Course Adaption Model” (LCAM).
-This model draws on the life course perspective to model personal culture as a linear trajectory over the course of a panel that varies for each respondent.
-Lersch rectifies two shortcomings of the AUM and SDM.
-First, the AUM, as described by Kiley and Vaisey, posits that changes follow a Markovian process where personal culture responses at time $t$ rely solely on responses at time $t-1$.
-Howevre, earlier life experiences can (directly and indirectly) mold personal culture when transitioning to new social roles or into new environments, even if their initial impact is delayed.
-For example, childhood events might influence views on family structures later when individuals form their own families.
-The LCAM considers therefore considers influences from earlier than just $t-1$ on responses at time $t$.
+Drawing on these findings, Lersch challenged the SDM and AUM as a
+``needless dichotomy'' (2023). Instead he proposed the ``Life Course
+Adaption Model'' (LCAM). This model draws on the life course perspective
+to model personal culture as a linear trajectory over the course of a
+panel that varies for each respondent. Lersch rectifies two shortcomings
+of the AUM and SDM. First, the AUM, as described by Kiley and Vaisey,
+posits that changes follow a Markovian process where personal culture
+responses at time \(t\) rely solely on responses at time \(t-1\).
+Howevre, earlier life experiences can (directly and indirectly) mold
+personal culture when transitioning to new social roles or into new
+environments, even if their initial impact is delayed. For example,
+childhood events might influence views on family structures later when
+individuals form their own families. The LCAM considers therefore
+considers influences from earlier than just \(t-1\) on responses at time
+\(t\).
 
-Second, Kiley and Vaisey's analysis of the AUM and SDM is based on three-wave panel data over four years, which might not be extensive enough to properly adjudicate [COMPARE?
-WE ARE USING ADJUDICATING A LOT] the two models.
-Lersch, however, evaluates the LCAM against the AUM and SDM using panel data spanning a wider duration (from 3 to 36 years) and more waves (3 to 18), offering a better change to observe durable change.
+Second, Kiley and Vaisey's analysis of the AUM and SDM is based on
+three-wave panel data over four years, which might not be extensive
+enough to properly adjudicate {[}COMPARE? WE ARE USING ADJUDICATING A
+LOT{]} the two models. Lersch, however, evaluates the LCAM against the
+AUM and SDM using panel data spanning a wider duration (from 3 to 36
+years) and more waves (3 to 18), offering a better change to observe
+durable change.
 
-In Lersch's tournament of models, the LCAM gains broad [EXTENSIVE?] support from data across five countries.
-Out of 428 questions, the LCAM was favored in 297 cases while the SDM was preferred in only 112.
-The rest did not yield inconclusive results, and notably, no questions favoured the AUM.
-He concludes that “new experiences over the life course [...] can persistently move individuals’ personal culture in novel directions” [@lersch2023: p.
-243-244].
+In Lersch's tournament of models, the LCAM gains broad {[}EXTENSIVE?{]}
+support from data across five countries. Out of 428 questions, the LCAM
+was favored in 297 cases while the SDM was preferred in only 112. The
+rest did not yield inconclusive results, and notably, no questions
+favoured the AUM. He concludes that ``new experiences over the life
+course {[}\ldots{]} can persistently move individuals' personal culture
+in novel directions'' (Lersch 2023: p.~243-244).
 
-Despite differing interpretations of these scholars, their empirical results align closely.
-Kiley and Vaisey [-@kiley2020] discovered that most questions they investigated favored the AUM, indicating evidence of durable change among adults -- a finding largely in line with Lersch’s.
-However, while Lersch [-@lersch2023] found a preference for the LCAM in many cases, durable change only accounted for a minor proportion of variance explained.
-On average, individuals changed [CAN WE SAY SHIFTED TO INDICATE THE LINEAR DRIFT] about .07 standard deviations over 10 years, echoing Kiley's and Vaisey’s view that such changes are typically modest.
-Furthermore, a quarter of the items in Lersch's study favoured the SDM, suggesting that despite broader assumptions about change and more extensive data, many questions reveal no evidence of durable change.
-Likewise, in studies claiming evidence of durable change in personal culture, the magnitude is often minimal.
-For instance, Brocic and Miles [-@brocic2021] note that obtaining graduate degrees in humanities, arts, and social sciences only shift [NUDGES?] peoples’ moral relativism by about .2 standard deviations compared to non-degree holders [DELETED THE LAST SENTENCE PART HERE; FELT UNNECESSARY].
-Moreover, studies of aggregate change demonstrate that even on questions where cohort effects prevail, there is always evidence that some individuals change over time [@vaisey2016].
-In other words, despite different interpretations, the results of previous work are in many ways highly consonant.
+Despite differing interpretations of these scholars, their empirical
+results align closely. Kiley and Vaisey (2020) discovered that most
+questions they investigated favored the AUM, indicating evidence of
+durable change among adults -- a finding largely in line with Lersch's.
+However, while Lersch (2023) found a preference for the LCAM in many
+cases, durable change only accounted for a minor proportion of variance
+explained. On average, individuals changed {[}CAN WE SAY SHIFTED TO
+INDICATE THE LINEAR DRIFT{]} about .07 standard deviations over 10
+years, echoing Kiley's and Vaisey's view that such changes are typically
+modest. Furthermore, a quarter of the items in Lersch's study favoured
+the SDM, suggesting that despite broader assumptions about change and
+more extensive data, many questions reveal no evidence of durable
+change. Likewise, in studies claiming evidence of durable change in
+personal culture, the magnitude is often minimal. For instance, Brocic
+and Miles (2021) note that obtaining graduate degrees in humanities,
+arts, and social sciences only shift {[}NUDGES?{]} peoples' moral
+relativism by about .2 standard deviations compared to non-degree
+holders {[}DELETED THE LAST SENTENCE PART HERE; FELT UNNECESSARY{]}.
+Moreover, studies of aggregate change demonstrate that even on questions
+where cohort effects prevail, there is always evidence that some
+individuals change over time (Vaisey and Lizardo 2016). In other words,
+despite different interpretations, the results of previous work are in
+many ways highly consonant.
 
-## Quantification, Not Adjudication
+\hypertarget{quantification-not-adjudication}{%
+\subsection{Quantification, Not
+Adjudication}\label{quantification-not-adjudication}}
 
-Current debates often narrow down to ask whether people *ever* change, which is rarely theoretically interesting since the answer is almost always “yes.”  Over sufficient time and in a large enough sample, researchers will likely observe some evidence for durable changes in individuals.
-A lack thereof may simply due to a lack of clarity in the question asked or a low level of resolution in the possible answers, or that the question simply was not asked long enough or to enough people.
-Conversely, finding evidence of change tells us little about how much change is present in a population and how meaningful that change is for social differences.
+Current debates often narrow down to ask whether people \emph{ever}
+change, which is rarely theoretically interesting since the answer is
+almost always ``yes.'' Over sufficient time and in a large enough
+sample, researchers will likely observe some evidence for durable
+changes in individuals. A lack thereof may simply due to a lack of
+clarity in the question asked or a low level of resolution in the
+possible answers, or that the question simply was not asked long enough
+or to enough people. Conversely, finding evidence of change tells us
+little about how much change is present in a population and how
+meaningful that change is for social differences.
 
-A more theoretically productive approach begins with a model assuming that during an observed time period some individuals might remain stable, might change a little, or might undergo significant shifts in their personal culture.
-Lersch’s LCAM formally does this by modeling each individual as following their own trajectory over time.
-But rather than debating whether this model is superior to a model that assumes that people never change, researchers should quantify individual changes over time and compare those to the differences that exist between people at the panel's start.
-In other words, the focus should shift from asking, “do people change?” to asking, “how do change and initial differences collectively explain variation in personal culture?” [I CHANGED THAT QUESTION WORDING, PLEASE CHECK!]
+A more theoretically productive approach begins with a model assuming
+that during an observed time period some individuals might remain
+stable, might change a little, or might undergo significant shifts in
+their personal culture. Lersch's LCAM formally does this by modeling
+each individual as following their own trajectory over time. But rather
+than debating whether this model is superior to a model that assumes
+that people never change, researchers should quantify individual changes
+over time and compare those to the differences that exist between people
+at the panel's start. In other words, the focus should shift from
+asking, ``do people change?'' to asking, ``how do change and initial
+differences collectively explain variation in personal culture?'' {[}I
+CHANGED THAT QUESTION WORDING, PLEASE CHECK!{]}
 
 In this approach, two metrics reflect distinct theoretical processes.
-First, "early imprinting," as Lersch calls it, denotes interpersonal differences at the start of the panel, capturing the accumulated experiences of people before participating in the panel that influence their answers.
-While commonly associated with experiences during an early formative period that result in settled disposition, this could also reflect transformative experiences [-@paul2014transformative] that happen at any stage in life, as long as they predate the panel and consistently impact subsequent responses.
-For instance, for those people entering the panel post-retirement, this “imprinting” might reflect this pivotal life transition.
-Consequently, this metric also reflects variation in individuals' social roles or statuses at the start of the panel that were important in shaping their dispositions.
+First, ``early imprinting,'' as Lersch calls it, denotes interpersonal
+differences at the start of the panel, capturing the accumulated
+experiences of people before participating in the panel that influence
+their answers. While commonly associated with experiences during an
+early formative period that result in settled disposition, this could
+also reflect transformative experiences (2014) that happen at any stage
+in life, as long as they predate the panel and consistently impact
+subsequent responses. For instance, for those people entering the panel
+post-retirement, this ``imprinting'' might reflect this pivotal life
+transition. Consequently, this metric also reflects variation in
+individuals' social roles or statuses at the start of the panel that
+were important in shaping their dispositions.
 
-The second process, termed “persistent change” or “adaption” by Lersch and "active updating" by Kiley and Vaisey, denotes durable changes in personal culture over time, captured by the amount of intrapersonal linear change during the panel.
-Lersch attributes these changes to social triggers such as moving into a new environment or adopting new social roles.
-Additionally, they might reflect the diffusion of new cultural forms across social networks, cues from political elites or otherwise culturally influential leaders, the emergence of issues in politics or culture, or large-scale social shifts.
+The second process, termed ``persistent change'' or ``adaption'' by
+Lersch and ``active updating'' by Kiley and Vaisey, denotes durable
+changes in personal culture over time, captured by the amount of
+intrapersonal linear change during the panel. Lersch attributes these
+changes to social triggers such as moving into a new environment or
+adopting new social roles. Additionally, they might reflect the
+diffusion of new cultural forms across social networks, cues from
+political elites or otherwise culturally influential leaders, the
+emergence of issues in politics or culture, or large-scale social
+shifts.
 
-A third metric, often termed “residual variance” or “fluctuation,” represents the remainder of variance in peoples’ responses over time.
-This non-durable change may arise for various reasons.
-Some might lack clear cultural dispositions, leading them to construct an opinion or attitude during the survey, with varying thoughts and considerations dominating their thinking in each interview [@tourangeau2000; @zaller1992; @feldman1992].
-Even though the full set of considerations they could draw on remains stable over time, this could yield to such variance in responses.
-Additionally, this variance can also include measurement errors, such as misinterpreted questions, erroneous response selections, or responses getting coded incorrectly.
-While this third process thus might reflect potentially important aspects of personal culture [I FELT WE NEED TO PLAY IT DOWN A BIT], it is less central to the debate outlined here.
-Hence, we focus on the first two metrics.
+A third metric, often termed ``residual variance'' or ``fluctuation,''
+represents the remainder of variance in peoples' responses over time.
+This non-durable change may arise for various reasons. Some might lack
+clear cultural dispositions, leading them to construct an opinion or
+attitude during the survey, with varying thoughts and considerations
+dominating their thinking in each interview (Feldman and Zaller 1992;
+Tourangeau, Rips, and Rasinski 2000; Zaller 1992). Even though the full
+set of considerations they could draw on remains stable over time, this
+could yield to such variance in responses. Additionally, this variance
+can also include measurement errors, such as misinterpreted questions,
+erroneous response selections, or responses getting coded incorrectly.
+While this third process thus might reflect potentially important
+aspects of personal culture {[}I FELT WE NEED TO PLAY IT DOWN A BIT{]},
+it is less central to the debate outlined here. Hence, we focus on the
+first two metrics.
 
-At a theoretical level, the coexistence of two main processes -- interpersonal differences and intrapersonal change -- is undeniable, and they are linked in many ways [THOUGHT IF intrinsically linked].
-Any intrapersonal change during one's life will inevitably manifest as interpersonal differences by the time people enter a panel survey.
-Furthermore, unless people are entirely socialized early on and never deviate from these dispositions -- an understanding challenged by evidence -- we expect observing intrapersonal change in a segment of the population when surveyed over time.
+At a theoretical level, the coexistence of two main processes --
+interpersonal differences and intrapersonal change -- is undeniable, and
+they are linked in many ways {[}THOUGHT IF intrinsically linked{]}. Any
+intrapersonal change during one's life will inevitably manifest as
+interpersonal differences by the time people enter a panel survey.
+Furthermore, unless people are entirely socialized early on and never
+deviate from these dispositions -- an understanding challenged by
+evidence -- we expect observing intrapersonal change in a segment of the
+population when surveyed over time.
 
-Furthermore, these metrics are not necessarily useful in isolation or without context.
-Central to the theoretical debate is how the two processes they capture collectively explain cultural differences in a population.
-In other words, when examining adults across a period of time, does intrapersonal change, compared to interpersonal differences, sufficiently account for the observed cultural differences?
-Only through quantifying the two processes we can get closer to a true answer to this question.
+Furthermore, these metrics are not necessarily useful in isolation or
+without context. Central to the theoretical debate is how the two
+processes they capture collectively explain cultural differences in a
+population. In other words, when examining adults across a period of
+time, does intrapersonal change, compared to interpersonal differences,
+sufficiently account for the observed cultural differences? Only through
+quantifying the two processes we can get closer to a true answer to this
+question.
 
-Yet, the true merit of the proposed quantification goes beyond settling past debates; it opens up new avenues for investigations that remained closed under previous approaches.
-For example, merely classifying questions based on whether they show durable change [DO WE NEED "or not"?
-NEVER KNOW THAT AS A NON-NATIVE] overlooks possibly important differences among those that do.
-By quantifying the contributions of interpersonal differences and intrapersonal change, researchers can gauge their relative importance in understanding [INSTEAD explaining] differences among survey items, between groups within a given population, across time, and even across societies.
+Yet, the true merit of the proposed quantification goes beyond settling
+past debates; it opens up new avenues for investigations that remained
+closed under previous approaches. For example, merely classifying
+questions based on whether they show durable change {[}DO WE NEED ``or
+not''? NEVER KNOW THAT AS A NON-NATIVE{]} overlooks possibly important
+differences among those that do. By quantifying the contributions of
+interpersonal differences and intrapersonal change, researchers can
+gauge their relative importance in understanding {[}INSTEAD
+explaining{]} differences among survey items, between groups within a
+given population, across time, and even across societies.
 
-## Expectations
+\hypertarget{expectations}{%
+\subsection{Expectations}\label{expectations}}
 
-By reframing the question and adjusting the approach, we expect consistent patterns across datasets previously analysed in the debate.
-First, we anticipate evidence for some intrapersonal change in most measures of personal culture across datasets.
-Yet, earlier studies found many questions (between 25 and 35 percent) lacking evidence of durable change.
-Aligning wiht this, we expect considerable variation in the variance explained by intrapersonal change, with some questions indicating minimal variance attributable to such change.
-[I KEPT THE "VARIANCE EXPLAINED" PHRASING THOUGH THE "VARIANCE ATTRIBUTABLE" MIGHT BE LESS OF A HARD CLAIM WITHOUT LOSING MUCH?]
+By reframing the question and adjusting the approach, we expect
+consistent patterns across datasets previously analysed in the debate.
+First, we anticipate evidence for some intrapersonal change in most
+measures of personal culture across datasets. Yet, earlier studies found
+many questions (between 25 and 35 percent) lacking evidence of durable
+change. Aligning wiht this, we expect considerable variation in the
+variance explained by intrapersonal change, with some questions
+indicating minimal variance attributable to such change. {[}I KEPT THE
+``VARIANCE EXPLAINED'' PHRASING THOUGH THE ``VARIANCE ATTRIBUTABLE''
+MIGHT BE LESS OF A HARD CLAIM WITHOUT LOSING MUCH?{]}
 
-Second, we expect the variance explained by intrapersonal to be modest compared to interpersonal differences at the start of the survey.
-This assertion [EXPECTATION?
-PREMISE?] is fairly uncontroversial, regardless of the theoretical process assumed to be most relevant for a particular question.
-Panel surveys do not encompase individuals’ entire lifetimes.
-Thus, even if intrapersonal change explains a large part of cultural differences, the change detected in a panel covering 20 years probably will not fully account for the entire variance of interpersonal differences.
+Second, we expect the variance explained by intrapersonal to be modest
+compared to interpersonal differences at the start of the survey. This
+assertion {[}EXPECTATION? PREMISE?{]} is fairly uncontroversial,
+regardless of the theoretical process assumed to be most relevant for a
+particular question. Panel surveys do not encompase individuals' entire
+lifetimes. Thus, even if intrapersonal change explains a large part of
+cultural differences, the change detected in a panel covering 20 years
+probably will not fully account for the entire variance of interpersonal
+differences.
 
-Our expectations so far largely echo what was observed in previous studies.
-However, theoretical perspectives differ in how much variance they attribute to intrapersonal change, enabling quantification to provide a new perspective on existing questions.
-“Formative period” and cohort theories, which emphasize early-life socialization as a strong force in shaping adult personal culture, suggest that the variance from intrapersonal change would still fall short of explaining interpersonal differences, even when extrapolated to a longer time frame [WE COULD DROP HERE A REFERENCE TO THE SPAN..., possibly beyond the panel window].
-Conversely, other theoretical perspectives emphasizing the importance of people's current social roles expect intrapersonal change could account for a large portion of the variance when extrapolated [CAN WE FIND AN ALTERNATIVE TO "EXTRAPOLATE"?].
-While projecting observed panel change beyond the covered time period may have its critics, we see it as a reasonable starting point.
-[DO WE NEED THIS LAST SENTENCE?
-IT FEELS WE ARE MAKING OUR SELVES VULNERABLE TO A CRITIQUE TO NO BENEFIT IN THE ARGUMENT.]
+Our expectations so far largely echo what was observed in previous
+studies. However, theoretical perspectives differ in how much variance
+they attribute to intrapersonal change, enabling quantification to
+provide a new perspective on existing questions. ``Formative period''
+and cohort theories, which emphasize early-life socialization as a
+strong force in shaping adult personal culture, suggest that the
+variance from intrapersonal change would still fall short of explaining
+interpersonal differences, even when extrapolated to a longer time frame
+{[}WE COULD DROP HERE A REFERENCE TO THE SPAN\ldots, possibly beyond the
+panel window{]}. Conversely, other theoretical perspectives emphasizing
+the importance of people's current social roles expect intrapersonal
+change could account for a large portion of the variance when
+extrapolated {[}CAN WE FIND AN ALTERNATIVE TO ``EXTRAPOLATE''?{]}. While
+projecting observed panel change beyond the covered time period may have
+its critics, we see it as a reasonable starting point. {[}DO WE NEED
+THIS LAST SENTENCE? IT FEELS WE ARE MAKING OUR SELVES VULNERABLE TO A
+CRITIQUE TO NO BENEFIT IN THE ARGUMENT.{]}
 
-Third, we expect the variance explained by each process –interpersonal differences and intrapersonal change– to depend on specific attributes of the question and panel.
-Noteably, the longer we observe respondents, the higher the response resolution for a given question, and the larger the sample, the greater the variance explained by intrapersonal change should be.
-Given that the LCAM (supported by the various underlying theoretical perspectives) posits that social experiences drive changes in personal culture and that these experiences correlate with time, observing people for longer durations [or TIME SPANS?] increases the chance that they undergo experiences that change their personal culture.
-[WE WOULDN'T EVEN NEED THE "they undergo experiences that" IN THAT SENTENCE.]
+Third, we expect the variance explained by each process --interpersonal
+differences and intrapersonal change-- to depend on specific attributes
+of the question and panel. Noteably, the longer we observe respondents,
+the higher the response resolution for a given question, and the larger
+the sample, the greater the variance explained by intrapersonal change
+should be. Given that the LCAM (supported by the various underlying
+theoretical perspectives) posits that social experiences drive changes
+in personal culture and that these experiences correlate with time,
+observing people for longer durations {[}or TIME SPANS?{]} increases the
+chance that they undergo experiences that change their personal culture.
+{[}WE WOULDN'T EVEN NEED THE ``they undergo experiences that'' IN THAT
+SENTENCE.{]}
 
-# Analytical Strategy
+\hypertarget{analytical-strategy}{%
+\section{Analytical Strategy}\label{analytical-strategy}}
 
-## Data
+\hypertarget{data}{%
+\subsection{Data}\label{data}}
 
-We use data from 7 nationally representative panel surveys from Australia, Germany, Great Britain, Switzerland, and the United States (summarized in Table 1), combining all the data files used in previous work [@kiley2020; @lersch2023].[^1] These studies cover a long period of time (the range of surveys spans from 1968 to 2021), with 609 personal culture items that capture attitudes, beliefs, values, self-assessments, self-descriptions, and behaviors [@alwin2007].
-We restricted the sample such that individuals between the ages of 18 and 79 are included without further elimination, and in all surveys, we used all possible cases for which respondents provided responses.
-In the end, the analyses that follow rely on a cross-cultural sample with a cross-domain set of items to capture the breadth of individual personal culture.
-Appendix A documents the list of all variables used in the upcoming analyses.
-
-[^1]: For more information on these data sources, see [@data_gss; @data_hilda; @data_bhps; @data_ukhls; @data_soep; @data_shp; @data_psid].
+We use data from 7 nationally representative panel surveys from
+Australia, Germany, Great Britain, Switzerland, and the United States
+(summarized in Table 1), combining all the data files used in previous
+work (Kiley and Vaisey 2020; Lersch 2023).\footnote{For more information
+  on these data sources, see (Goebel et al. 2019; Income Dynamics 2013;
+  Smith et al. 2022; Summerfield et al. 2011; Taylor 1996; University of
+  Essex and Research 2019; Voorpostel et al. 2016).} These studies cover
+a long period of time (the range of surveys spans from 1968 to 2021),
+with 609 personal culture items that capture attitudes, beliefs, values,
+self-assessments, self-descriptions, and behaviors (Alwin 2007). We
+restricted the sample such that individuals between the ages of 18 and
+79 are included without further elimination, and in all surveys, we used
+all possible cases for which respondents provided responses. In the end,
+the analyses that follow rely on a cross-cultural sample with a
+cross-domain set of items to capture the breadth of individual personal
+culture. Appendix A documents the list of all variables used in the
+upcoming analyses.
 
@@
 
-## Life Course Adaption Model
+\hypertarget{life-course-adaption-model}{%
+\subsection{Life Course Adaption
+Model}\label{life-course-adaption-model}}
 
-We start with Lersch’s [-@lersch2023] Life Course Adaption Model, which formalizes survey responses at time $t$ as a function of individual-level random intercepts and slopes for survey age, what is often called a mixed-effects growth curve model.
-This model formally assumes a set of propositions about change that reflect the theoretical debate to this point.
-First, consistent with the settled disposition model, it assumes that people start the survey with cultural differences, modeled as random intercepts for each respondent.
-Second, it assumes that people change over time, taking the form of random slopes for each respondent as a linear function of time.
-Third, it assumes that people deviate around this baseline randomly over time, reflecting “fluctuation” or short-term non-persistent change.
-Finally, it assumes that people draw on earlier time points in reacting to new events.
-Formally, this can be written as
+We start with Lersch's (2023) Life Course Adaption Model, which
+formalizes survey responses at time \(t\) as a function of
+individual-level random intercepts and slopes for survey age, what is
+often called a mixed-effects growth curve model. This model formally
+assumes a set of propositions about change that reflect the theoretical
+debate to this point. First, consistent with the settled disposition
+model, it assumes that people start the survey with cultural
+differences, modeled as random intercepts for each respondent. Second,
+it assumes that people change over time, taking the form of random
+slopes for each respondent as a linear function of time. Third, it
+assumes that people deviate around this baseline randomly over time,
+reflecting ``fluctuation'' or short-term non-persistent change. Finally,
+it assumes that people draw on earlier time points in reacting to new
+events. Formally, this can be written as
 
-$$
+\[
 y_{i, t} = \beta_0 + u_{0, i} + \beta_1 x_t + u_{1, i} x_{i, t} + \epsilon_{i, t}
-$$
+\]
 
-where $\beta_0$ is the average intercept, $u_{0, i}$ is the random intercept for individuals, $\beta_1 x_t$ is the average slope, $u_{1, i} x_{i, t}$ is the random slope for individuals, and $\epsilon_{i, t}$ is the random error term that captures transitory disturbances.
+where \(\beta_0\) is the average intercept, \(u_{0, i}\) is the random
+intercept for individuals, \(\beta_1 x_t\) is the average slope,
+\(u_{1, i} x_{i, t}\) is the random slope for individuals, and
+\(\epsilon_{i, t}\) is the random error term that captures transitory
+disturbances.
 
-As a first step in our analysis, we fit LCAMs to each of the 609 measures of personal culture outlined above.
-We then use these models to derive two measures of systematic variance that correspond to the two collections of theoretical processes outlined above: interpersonal differences at baseline and intrapersonal change over time.
-First, to measure the importance of interpersonal differences, we calculate the sum of the squared residuals derived from measuring each respondent at the midpoint of their age trajectory.[^midpoint] We then divide this value by the total sum of squares and subtract it from from 1, producing a value we denote as $\tau$, or the proportion of variance explained by interpersonal differences.
-This number is equivalent to an $R^2$ value generated by using random intercepts for all respondents.
-Second, we calculate the sum of squared residuals generated by additionally accounting for the age trajectory.
-We divide this number by the total sum of squares and subtract it from 1.
-We then subtract $\tau$ from this value.
-This gives us a value we denote as $\delta$, the additional proportion of variance explained by intrapersonal change.
+As a first step in our analysis, we fit LCAMs to each of the 609
+measures of personal culture outlined above. We then use these models to
+derive two measures of systematic variance that correspond to the two
+collections of theoretical processes outlined above: interpersonal
+differences at baseline and intrapersonal change over time. First, to
+measure the importance of interpersonal differences, we calculate the
+sum of the squared residuals derived from measuring each respondent at
+the midpoint of their age trajectory.\footnote{We use the scale
+  midpoint, rather than the actual difference at baseline, to better
+  capture the differences between people independent of their age
+  trajectories.} We then divide this value by the total sum of squares
+and subtract it from from 1, producing a value we denote as \(\tau\), or
+the proportion of variance explained by interpersonal differences. This
+number is equivalent to an \(R^2\) value generated by using random
+intercepts for all respondents. Second, we calculate the sum of squared
+residuals generated by additionally accounting for the age trajectory.
+We divide this number by the total sum of squares and subtract it from
+1. We then subtract \(\tau\) from this value. This gives us a value we
+denote as \(\delta\), the additional proportion of variance explained by
+intrapersonal change.
 
-These two numbers reflect the total systematic or "explained" variance of the model that can be attributed to either differences across people or linear change over the course of the panel.[^explained] Our principal outcome of interest is the proportion of systematic variance accounted for by intrapersonal change, which we calculate as
+These two numbers reflect the total systematic or ``explained'' variance
+of the model that can be attributed to either differences across people
+or linear change over the course of the panel.\footnote{This does not
+  mean that other factors might not explain additional variance if they
+  were included in the model. For example, peoples' responses about
+  religious beliefs and participation might systematically vary by the
+  day of the week or month in which they are surveyed. This variation is
+  moved into the ``random'' or non-systematic component of the model for
+  us.} Our principal outcome of interest is the proportion of systematic
+variance accounted for by intrapersonal change, which we calculate as
 
-$$
+\[
 \frac{\delta}{\tau + \delta}
-$$
+\]
 
-[^midpoint]: We use the scale midpoint, rather than the actual difference at baseline, to better capture the differences between people independent of their age trajectories.
+We can think of this number as akin to an Intraclass Correlation
+Coefficient focused only on the systematic components of the model while
+ignoring residual variance that might be attributable to measurement
+error or non-persisting change (see also the variance decomposition of
+period and cohort by Vaisey and Lizardo 2016). The higher the number is,
+the more variance in responses is attributable to intrapersonal change
+during the panel. The lower the number, the less intrapersonal change
+seems to play a large role in explaining cultural differences during the
+panel.
 
-[^explained]: This does not mean that other factors might not explain additional variance if they were included in the model.
-For example, peoples' responses about religious beliefs and participation might systematically vary by the day of the week or month in which they are surveyed.
-This variation is moved into the "random" or non-systematic component of the model for us.
+Because this quotient summarizes variance and therefore has no natural
+referent, it is hard to say at which values we observe ``a lot'' of
+intrapersonal change. But summarizing questions in this way allows us to
+compare the relative importance of our two broad processes across a
+range of questions that might have very different levels of
+non-systematic variance, giving us a baseline sense of the relative
+prevalence of these two processes for each question (something not
+achievable under the tournament of models approach), and allowing us to
+compare which kinds of questions show more or less intrapersonal change
+relative to interpersonal differences. While the tournament of models
+approach allowed researchers to say whether questions in general showed
+evidence of updating, this approach allows us to make more specific
+claims about the prevalence of these theoretical processes for
+particular questions relative to each other.
 
-We can think of this number as akin to an Intraclass Correlation Coefficient focused only on the systematic components of the model while ignoring residual variance that might be attributable to measurement error or non-persisting change [see also the variance decomposition of period and cohort by @vaisey2016].
-The higher the number is, the more variance in responses is attributable to intrapersonal change during the panel.
-The lower the number, the less intrapersonal change seems to play a large role in explaining cultural differences during the panel.
+As a second step in our analysis, we use ordinary least squares
+regression to model the proportion of variance explained by
+intrapersonal change as a function of features of the question and
+panel. This allows us to explore how these features affect the
+prevalence of intrapersonal change. These covariates include the number
+of response options (measured as 2, 3-5, and more than 5), the panel,
+the log of the number of participants, the number of waves observed, the
+time since the question was first asked, and the total span of time the
+question was asked.
 
-Because this quotient summarizes variance and therefore has no natural referent, it is hard to say at which values we observe "a lot" of intrapersonal change.
-But summarizing questions in this way allows us to compare the relative importance of our two broad processes across a range of questions that might have very different levels of non-systematic variance, giving us a baseline sense of the relative prevalence of these two processes for each question (something not achievable under the tournament of models approach), and allowing us to compare which kinds of questions show more or less intrapersonal change relative to interpersonal differences.
-While the tournament of models approach allowed researchers to say whether questions in general showed evidence of updating, this approach allows us to make more specific claims about the prevalence of these theoretical processes for particular questions relative to each other.
+\hypertarget{results}{%
+\section{Results}\label{results}}
 
-As a second step in our analysis, we use ordinary least squares regression to model the proportion of variance explained by intrapersonal change as a function of features of the question and panel.
-This allows us to explore how these features affect the prevalence of intrapersonal change.
-These covariates include the number of response options (measured as 2, 3-5, and more than 5), the panel, the log of the number of participants, the number of waves observed, the time since the question was first asked, and the total span of time the question was asked.
+Our results proceed in three parts. We first explore the overall pattern
+of systematic variance explained by interpersonal differences at
+baseline and intrapersonal change over time across the seven panels.
+Next, we analyze the association between the relative proportion of
+systematic variance explained by intrapersonal change as a function of
+question, panel, and sample features. Finally, we highlight a few
+specific comparisons to illustrate the value of quantification relative
+to adjudication for further exploring theoretical questions.
 
-# Results
+\hypertarget{variance-decomposition}{%
+\subsection{Variance Decomposition}\label{variance-decomposition}}
 
-Our results proceed in three parts.
-We first explore the overall pattern of systematic variance explained by interpersonal differences at baseline and intrapersonal change over time across the seven panels.
-Next, we analyze the association between the relative proportion of systematic variance explained by intrapersonal change as a function of question, panel, and sample features.
-Finally, we highlight a few specific comparisons to illustrate the value of quantification relative to adjudication for further exploring theoretical questions.
-
-## Variance Decomposition
-
-Figure 1 plots the proportion of systematic variance attributable to interpersonal differences (“between”) and intrapersonal change (“within”) for each of the 609 questions analyzed, plotted separately by panel.
-To broadly summarize the results, all panels show a similar range of the proportion of systematic variance attributable to intrapersonal change.
-Across panels, the median proportion of systematic variance attributable to intrapersonal change is 0.094 (mean of 0.098), with an interquartile range of 0.053 to 0.139.
-All panels include questions that show essentially no systematic variance attributable to intrapersonal change and all panels include questions where more than 20 percent of systematic variance is attributable to intrapersonal change.
+Figure 1 plots the proportion of systematic variance attributable to
+interpersonal differences (``between'') and intrapersonal change
+(``within'') for each of the 609 questions analyzed, plotted separately
+by panel. To broadly summarize the results, all panels show a similar
+range of the proportion of systematic variance attributable to
+intrapersonal change. Across panels, the median proportion of systematic
+variance attributable to intrapersonal change is 0.094 (mean of 0.098),
+with an interquartile range of 0.053 to 0.139. All panels include
+questions that show essentially no systematic variance attributable to
+intrapersonal change and all panels include questions where more than 20
+percent of systematic variance is attributable to intrapersonal change.
 
@@
 
-Across all questions, interpersonal differences explain a much larger share of the systematic variance in responses than intrapersonal change.
-Again, this is to be expected.
-Interpersonal differences capture not just pre-adult socialization, but all accumulated experiences up to the start of the panel that might have a bearing on personal culture.
+Across all questions, interpersonal differences explain a much larger
+share of the systematic variance in responses than intrapersonal change.
+Again, this is to be expected. Interpersonal differences capture not
+just pre-adult socialization, but all accumulated experiences up to the
+start of the panel that might have a bearing on personal culture.
 
-To the extent that there are differences across the panels, the PSID has the highest mean (.120) and median (.135) proportion of systematic variance attributable to intrapersonal change.
-While we cannot disentangle features of the sample from features of the questions asked to each sample, the specific samples for many PSID questions have lower average ages than those from other panels.
-To the extent that younger respondents might be more likely to make durable changes of opinion, this higher explanatory power of intrapersonal change might reflect the distinct age profile of respondents in this sample.
-At the other end, the GSS has the lowest mean (.073) and median (.069) proportion of variance attributable to intrapersonal change.
-This potentially reflects the fact that the GSS observes people for a shorter duration, on average, than the other panels.
-If, consistent with life course adaption theories, people are more likely to make significant cultural changes the longer we observe them, then duration likely affects the amount of variance attributable to intrapersonal change (a point we explore more below).
-However, the GSS results are still highly consistent with results from the other panels.
+To the extent that there are differences across the panels, the PSID has
+the highest mean (.120) and median (.135) proportion of systematic
+variance attributable to intrapersonal change. While we cannot
+disentangle features of the sample from features of the questions asked
+to each sample, the specific samples for many PSID questions have lower
+average ages than those from other panels. To the extent that younger
+respondents might be more likely to make durable changes of opinion,
+this higher explanatory power of intrapersonal change might reflect the
+distinct age profile of respondents in this sample. At the other end,
+the GSS has the lowest mean (.073) and median (.069) proportion of
+variance attributable to intrapersonal change. This potentially reflects
+the fact that the GSS observes people for a shorter duration, on
+average, than the other panels. If, consistent with life course adaption
+theories, people are more likely to make significant cultural changes
+the longer we observe them, then duration likely affects the amount of
+variance attributable to intrapersonal change (a point we explore more
+below). However, the GSS results are still highly consistent with
+results from the other panels.
 
-While there are some differences between panels, these differences are small compared to the differences within panels.
-For about 6 percent of items explored here, more than 20 percent of systematic variance is attributable to intrapersonal change.
-These questions tend to ask about external referents (e.g., confidence in specific government leaders or political parties), life satisfaction, or current financial position.
-At the other end, questions about religious identification, views on gender roles, and support for civil liberties tend to have very low proportion of variance attributable to intrapersonal change.
+While there are some differences between panels, these differences are
+small compared to the differences within panels. For about 6 percent of
+items explored here, more than 20 percent of systematic variance is
+attributable to intrapersonal change. These questions tend to ask about
+external referents (e.g., confidence in specific government leaders or
+political parties), life satisfaction, or current financial position. At
+the other end, questions about religious identification, views on gender
+roles, and support for civil liberties tend to have very low proportion
+of variance attributable to intrapersonal change.
 
-In contrast to the tournament of models approach, quantifying change this way allows us to explore variation in the relative importance of intrapersonal change across questions that all both show evidence of change.
-For example, Kiley and Vaisey [-@kiley2020] found that confidence in the press and confidence in religious leaders were both characterized by active updating, our results show that intrapersonal change is much more important for explaining variance in confidence in the press (0.164) than confidence in religion (0.049).
+In contrast to the tournament of models approach, quantifying change
+this way allows us to explore variation in the relative importance of
+intrapersonal change across questions that all both show evidence of
+change. For example, Kiley and Vaisey (2020) found that confidence in
+the press and confidence in religious leaders were both characterized by
+active updating, our results show that intrapersonal change is much more
+important for explaining variance in confidence in the press (0.164)
+than confidence in religion (0.049).
 
-Appendix A presents the estimated proportion of variance attributable to interpersonal differences and intrapersonal change, as well as the proportion of residual variance, for each question analyzed.
-In general, interpersonal differences are almost always the largest component of the total variance and tend to explain between 55 and 70 percent of total variance, while intrapersonal change is always the smallest, typically explaining between 3 and 8 percent of total variance.
-Residual variance tends to account for between 22 and 37 percent of variance (though on several questions residual variance is greater than 50 percent).
+Appendix A presents the estimated proportion of variance attributable to
+interpersonal differences and intrapersonal change, as well as the
+proportion of residual variance, for each question analyzed. In general,
+interpersonal differences are almost always the largest component of the
+total variance and tend to explain between 55 and 70 percent of total
+variance, while intrapersonal change is always the smallest, typically
+explaining between 3 and 8 percent of total variance. Residual variance
+tends to account for between 22 and 37 percent of variance (though on
+several questions residual variance is greater than 50 percent).
 
-As noted above, the substantive importance of intrapersonal change for understanding cultural differences depends on a range of factors, including how long the panel runs, whether assumptions about linear change hold, and whether the panel is capturing a distinctly turbulent period or a distinctly stable period for the item under question.
-However, if we assume that the period under observation is “typical” for a question –-not a time of extremely heightened (or lowered) sensitivity or change-– then it does not seem realistic for this process of intrapersonal change to explain a large share of interpersonal differences for many of the questions we observe.
+As noted above, the substantive importance of intrapersonal change for
+understanding cultural differences depends on a range of factors,
+including how long the panel runs, whether assumptions about linear
+change hold, and whether the panel is capturing a distinctly turbulent
+period or a distinctly stable period for the item under question.
+However, if we assume that the period under observation is ``typical''
+for a question ---not a time of extremely heightened (or lowered)
+sensitivity or change-\/-- then it does not seem realistic for this
+process of intrapersonal change to explain a large share of
+interpersonal differences for many of the questions we observe.
 
-## Meta-Analysis
+\hypertarget{meta-analysis}{%
+\subsection{Meta-Analysis}\label{meta-analysis}}
 
-Figure 2 plots the results from an ordinary least squares regression of proportion of systematic variance attributable to intrapersonal change as a function of question, panel, and sample features.
-These models also include fixed effects for panels (not shown), so coefficients reflect the association between a particular sample or question characteristic within a panel.[^paneleffects]
-
-[^paneleffects]: KK NOTE: Can we summarize these panel fixed effects at all?
+Figure 2 plots the results from an ordinary least squares regression of
+proportion of systematic variance attributable to intrapersonal change
+as a function of question, panel, and sample features. These models also
+include fixed effects for panels (not shown), so coefficients reflect
+the association between a particular sample or question characteristic
+within a panel.\footnote{KK NOTE: Can we summarize these panel fixed
+  effects at all?}
 
@@
 
-Figure 2 shows that the more response options respondents are given and the larger the sample, the larger the proportion of systematic variance is attributable to intrapersonal change.
-We interpret these coefficients as suggesting greater resolution on a question makes it easier to detect and model change.
-The earlier a question was asked, even net of how long it was been asked, the lower the proportion of systematic variance is attributable to intrapersonal change.
+Figure 2 shows that the more response options respondents are given and
+the larger the sample, the larger the proportion of systematic variance
+is attributable to intrapersonal change. We interpret these coefficients
+as suggesting greater resolution on a question makes it easier to detect
+and model change. The earlier a question was asked, even net of how long
+it was been asked, the lower the proportion of systematic variance is
+attributable to intrapersonal change.
 
-The most notable result in the meta analysis is the negative coefficient attached to the span of years covered by the question.
-The longer a question is observed, the less systematic variance is attributable to intrapersonal change.
-Theories of personal cultural change that link changes in personal culture to social experiences, including the LCAM, suggest that the longer we observe respondents, the more likely people are to undergo potentially transformative experiences and therefore the more variance would be explained by intrapersonal change.
-The fact that we find a negative coefficient attached to this variable is a challenge to that assumption.
+The most notable result in the meta analysis is the negative coefficient
+attached to the span of years covered by the question. The longer a
+question is observed, the less systematic variance is attributable to
+intrapersonal change. Theories of personal cultural change that link
+changes in personal culture to social experiences, including the LCAM,
+suggest that the longer we observe respondents, the more likely people
+are to undergo potentially transformative experiences and therefore the
+more variance would be explained by intrapersonal change. The fact that
+we find a negative coefficient attached to this variable is a challenge
+to that assumption.
 
-As a further test of this coefficient, we compared the proportion of systematic variance attributable to intrapersonal change when using the full duration of a panel compared to when when we dropped the final wave and therefore reduced the total duration of observation for the question.
-If the coefficient reflects a true negative effect of duration on systematic variance attributable to intrapersonal change, we should see that same effect within questions.
-Results from this analysis are presented in Appendix B and contradict the coefficient from the regression model.
-The longer we observe the same question, the higher proportion of systematic variance is attributable to intrapersonal change.
-We interpret this combination of findings as suggesting that the kinds of questions asked for a longer time period tend to demonstrate less intrapersonal change than questions asked over shorter periods, rather than a true function of time.
+As a further test of this coefficient, we compared the proportion of
+systematic variance attributable to intrapersonal change when using the
+full duration of a panel compared to when when we dropped the final wave
+and therefore reduced the total duration of observation for the
+question. If the coefficient reflects a true negative effect of duration
+on systematic variance attributable to intrapersonal change, we should
+see that same effect within questions. Results from this analysis are
+presented in Appendix B and contradict the coefficient from the
+regression model. The longer we observe the same question, the higher
+proportion of systematic variance is attributable to intrapersonal
+change. We interpret this combination of findings as suggesting that the
+kinds of questions asked for a longer time period tend to demonstrate
+less intrapersonal change than questions asked over shorter periods,
+rather than a true function of time.
 
-This reinforces the explanation for why the GSS displays the lowest average proportion of systematic variance attributable to intrapersonal change: its shorter duration.
-[KK NOTE: Does this inform any other panel differences?
-Does the PSID, BHPS or UKHLS have longer durations on average?]
+This reinforces the explanation for why the GSS displays the lowest
+average proportion of systematic variance attributable to intrapersonal
+change: its shorter duration. {[}KK NOTE: Does this inform any other
+panel differences? Does the PSID, BHPS or UKHLS have longer durations on
+average?{]}
 
-## A Case Study: College Education
+\hypertarget{a-case-study-college-education}{%
+\subsection{A Case Study: College
+Education}\label{a-case-study-college-education}}
 
-While these broad results help speak to previous debates, they mostly re-frame previous findings without extending the theoretical debate.
-To better illustrate the value of our approach for addressing different theoretical questions, we turn to a specific empirical example: the relative importance of intrapersonal change and interpersonal difference for explaining cultural differences by education levels.
+While these broad results help speak to previous debates, they mostly
+re-frame previous findings without extending the theoretical debate. To
+better illustrate the value of our approach for addressing different
+theoretical questions, we turn to a specific empirical example: the
+relative importance of intrapersonal change and interpersonal difference
+for explaining cultural differences by education levels.
 
-Previous work has established a positive relationship between education and attitude stability, especially on issues related to American politics.
-This stability is often attributed to education facilitating "chronic knowledge" -- a general understanding of the domain of American politics, including the positions held by major parties and political figures and how issues relate to one another at a logical or socio-logical level [@zaller1992; @alvarez2002; @boutyline2017].
-Because they have more knowledge of American politics, people with a college education are better able to consistently connect the considerations in their cognition with the answer choices they are presented with in a survey.
+Previous work has established a positive relationship between education
+and attitude stability, especially on issues related to American
+politics. This stability is often attributed to education facilitating
+``chronic knowledge'' -- a general understanding of the domain of
+American politics, including the positions held by major parties and
+political figures and how issues relate to one another at a logical or
+socio-logical level (Alvarez and Brehm 2002; Boutyline and Vaisey 2017;
+Zaller 1992). Because they have more knowledge of American politics,
+people with a college education are better able to consistently connect
+the considerations in their cognition with the answer choices they are
+presented with in a survey.
 
-This work has tended to focus on the fact that college-educated Americans give responses that are less likely to be affected by measurement error or short-term influences [@zaller1992; @alwin2007].
-However, this focus on the difference in the non-systematic (residual) component of variance across groups obscures the fact that the relative importance of different components of systematic variance might differ across these two groups as well.
+This work has tended to focus on the fact that college-educated
+Americans give responses that are less likely to be affected by
+measurement error or short-term influences (Alwin 2007; Zaller 1992).
+However, this focus on the difference in the non-systematic (residual)
+component of variance across groups obscures the fact that the relative
+importance of different components of systematic variance might differ
+across these two groups as well.
 
-It could be the case that education, chronic knowledge, and attention to the political sphere makes intrapersonal change more important in explaining differences among people with higher levels of education.
-For example, these people might be more likely to make systematic changes in response to the emergence of new information, new issues, or political realignments, while those without chronic knowledge display more variance around their baselines without their baselines changing much [@zaller1992].
-In contrast, it could be the case that the stability demonstrated by educated respondents reflects the fact that they have already formed durable opinions and are relatively closed off to new information.
-College could serve as a formative experience that solidifies personal culture for people who attend.
-But among people who do not attend college, later life experiences might prove more important in forming or changing personal culture, as these experiences potentially provide information that college-educated peers received earlier.
+It could be the case that education, chronic knowledge, and attention to
+the political sphere makes intrapersonal change more important in
+explaining differences among people with higher levels of education. For
+example, these people might be more likely to make systematic changes in
+response to the emergence of new information, new issues, or political
+realignments, while those without chronic knowledge display more
+variance around their baselines without their baselines changing much
+(Zaller 1992). In contrast, it could be the case that the stability
+demonstrated by educated respondents reflects the fact that they have
+already formed durable opinions and are relatively closed off to new
+information. College could serve as a formative experience that
+solidifies personal culture for people who attend. But among people who
+do not attend college, later life experiences might prove more important
+in forming or changing personal culture, as these experiences
+potentially provide information that college-educated peers received
+earlier.
 
-To test these competing propositions, we use the models presented above to calculate the total systematic variance for people with at least a bachelor's degree and people with less than a bachelor's degree at wave 1[^wave] of the General Social Survey's panels.[^missing] We focus on the GSS because it contains the largest number of questions tapping general political dispositions, which is the domain where education has proven particularly relevant for understanding attitude stability.
-The GSS also covers a particularly turbulent window of American politics from 2006 to 2014.
-This window covers the start of the Great Recession, debates about federal intervention in and regulation of Wall Street, the election of Barack Obama as the first black president in U.S.
-history, debates about the role of the federal government in the health care sector, the emergence of the Tea Party, and political realignment and clarification on the issue of gay marriage, among other topics.
+To test these competing propositions, we use the models presented above
+to calculate the total systematic variance for people with at least a
+bachelor's degree and people with less than a bachelor's degree at wave
+1\footnote{A small number of respondents report different highest
+  degrees at wave 1, wave 2, and wave 3. Some of this is due to
+  measurement error, and some of it is due to a small number of people
+  obtaining a higher degree during the four years of the panel.
+  Estimating the panel with highest at wave 1 or highest degree reported
+  across the panel produces functionally identical results.} of the
+General Social Survey's panels.\footnote{We estimate the models
+  presented in {[}earlier figure{]} using the full sample, but only use
+  respondents who have information on their highest degree to generate
+  the results presented in figures {[}these figures{]}. In total, \#\#\#
+  respondents (\%\%\% percent) reported no degree information.} We focus
+on the GSS because it contains the largest number of questions tapping
+general political dispositions, which is the domain where education has
+proven particularly relevant for understanding attitude stability. The
+GSS also covers a particularly turbulent window of American politics
+from 2006 to 2014. This window covers the start of the Great Recession,
+debates about federal intervention in and regulation of Wall Street, the
+election of Barack Obama as the first black president in U.S. history,
+debates about the role of the federal government in the health care
+sector, the emergence of the Tea Party, and political realignment and
+clarification on the issue of gay marriage, among other topics.
 
-[^wave]: A small number of respondents report different highest degrees at wave 1, wave 2, and wave 3.
-Some of this is due to measurement error, and some of it is due to a small number of people obtaining a higher degree during the four years of the panel.
-Estimating the panel with highest at wave 1 or highest degree reported across the panel produces functionally identical results.
-
-[^missing]: We estimate the models presented in [earlier figure] using the full sample, but only use respondents who have information on their highest degree to generate the results presented in figures [these figures].
-In total, ### respondents (%%% percent) reported no degree information.
-
-Figure 3 plots the distribution of differences between people with less than a bachelor's degree and people with at least a bachelor's degree at wave 1 of the panel.
-Values greater than 0 would indicate that intrapersonal change is more important in explaining the responses of college graduates than those of people without a college degree, while values less than 0 would indicate that intrapersonal change is less important for explaining variation among people with a college degree than those without a college degree.
+Figure 3 plots the distribution of differences between people with less
+than a bachelor's degree and people with at least a bachelor's degree at
+wave 1 of the panel. Values greater than 0 would indicate that
+intrapersonal change is more important in explaining the responses of
+college graduates than those of people without a college degree, while
+values less than 0 would indicate that intrapersonal change is less
+important for explaining variation among people with a college degree
+than those without a college degree.
 
@@
 <about here>
-Figure 3: Proportions of intrapersonal change of explained variance by college degree; based on predictions at participants' wave mid-point.
+Figure 3: Distribution of differences share of variance explained by intrapersonal change between respondents without and with college degree.
 \end{center}
 
-There is a clear pattern in [figure above]: for more than 80 percent of GSS items explored here, intrapersonal change is a larger component of systematic variance for people without a college degree than for people with a college degree (they have negative values in the figure).
-While most of these differences are small in absolute terms (less than 2 percentage points), several are greater than 5 percentage points.
-Given the distribution observed in Figure [the earlier one] showing that the systematic variance attributable to intrapersonal change averages [average across all questions], a 5 percentage point difference between groups is quite substantial.
+There is a clear pattern in Figure 3: for more than 80 percent of GSS
+items explored here, intrapersonal change is a larger component of
+systematic variance for people without a college degree than for people
+with a college degree (they have negative values in the figure). While
+most of these differences are small in absolute terms (less than 2
+percentage points), several are greater than 5 percentage points. Given
+the distribution observed in Figure {[}the earlier one{]} showing that
+the systematic variance attributable to intrapersonal change averages
+{[}average across all questions{]}, a 5 percentage point difference
+between groups is quite substantial.
 
-To more clearly illustrate some of these differences, we highlight eight questions designed to tap general political dispositions: partisan identification (Democrat vs.
-Republican) and ideological identification (liberal vs.
-conservative) on seven-point scales; four questions about the government's role in improving the condition of the poor, paying people's medical bills, giving special treatment to blacks, and doing things that private businesses could do, measured on five-point scales; a question about whether the government should do more to reduce income differences, measured on a seven-point scale; and one question about whether blacks should be given preferences in hiring, measured on a five-point scale.
-Estimates for the proportion of systematic variance attributable to intrapersonal change on these eight questions, for both education groups, are presented in Figure XXX.
+To more clearly illustrate some of these differences, we highlight eight
+questions designed to tap general political dispositions: partisan
+identification (Democrat vs.~Republican) and ideological identification
+(liberal vs.~conservative) on seven-point scales; four questions about
+the government's role in improving the condition of the poor, paying
+people's medical bills, giving special treatment to blacks, and doing
+things that private businesses could do, measured on five-point scales;
+a question about whether the government should do more to reduce income
+differences, measured on a seven-point scale; and one question about
+whether blacks should be given preferences in hiring, measured on a
+five-point scale. Estimates for the proportion of systematic variance
+attributable to intrapersonal change on these eight questions, for both
+education groups, are presented in Figure 4.
 
-On all eight questions presented in Figure XXX, intrapersonal change explains less systematic variance for people with a college degree than for people without a college degree.
-There are also large differences across questions on the amount of systematic variance attributable to intrapersonal change for both groups.
-For example, on the question of whether the government should try to solve more problems or leave those problems to be solved by private businesses ("government do more or less"), less than one percent of the systematic variance is attributable to intrapersonal change for both groups.
-In other words, while people might vacillate on this question at random (37 percent of variance is residual for this question), there is functionally no evidence that people make systematic changes of opinion on this issue during the GSS panel.
+On all eight questions presented in Figure 4, intrapersonal change
+explains less systematic variance for people with a college degree than
+for people without a college degree. There are also large differences
+across questions on the amount of systematic variance attributable to
+intrapersonal change for both groups. For example, on the question of
+whether the government should try to solve more problems or leave those
+problems to be solved by private businesses (``government do more or
+less''), less than one percent of the systematic variance is
+attributable to intrapersonal change for both groups. In other words,
+while people might vacillate on this question at random (37 percent of
+variance is residual for this question), there is functionally no
+evidence that people make systematic changes of opinion on this issue
+during the GSS panel.
 
-In contrast, partisan identification and political ideology both show a larger proportion of systematic variance attributable to intrapersonal change, with quite large differences between people with a college degree and those without.
-Compared to the other questions explored here, change as adults (and, potentially, their experiences as adults) plays a much larger role in explaining why people report the partisan identification and political ideology they do.
-And this is particularly true for respondents who do not have a college degree.
-The amount of systematic variance in ideological identification that is attributable to intrapersonal change among non-college educated respondents is almost four times that of college-educated respondents.
+\begin{center}
+<about here>
+Figure 4: Proportions of intrapersonal change of explained variance by college degree; based on predictions at participants' wave mid-point.
+\end{center}
 
-We note here that this meaningful difference in the proportion of systematic variance attributable to intrapersonal change across education groups and across questions would not be detectable using previous methods.
-For partisan identification and political ideology, both college-educated respondents and people with less than college degree would likely favor the AUM or LCAM over the SDM because they show evidence of some members of the population making intrapersonal change.
-In other words, using the tournament of models approach obscures the fact that the relative importance of intrapersonal change in explaining differences on these two items is very different across these two groups, that intrapersonal change explains more systematic variance for ideological identification than partisan identification for people without a college degree, or that there appears to be more durable change on questions of affirmative action than on questions of government aid to black Americans.
+In contrast, partisan identification and political ideology both show a
+larger proportion of systematic variance attributable to intrapersonal
+change, with quite large differences between people with a college
+degree and those without. Compared to the other questions explored here,
+change as adults (and, potentially, their experiences as adults) plays a
+much larger role in explaining why people report the partisan
+identification and political ideology they do. And this is particularly
+true for respondents who do not have a college degree. The amount of
+systematic variance in ideological identification that is attributable
+to intrapersonal change among non-college educated respondents is almost
+four times that of college-educated respondents.
 
-It is important to note here that these differences should not be interpreted as causal effects of attending college on the importance of social experiences for cultural change.
-These results are potentially confounded by age, social class, race, gender, and other factors that explain selection into higher education.
-We simply use these estimates as a starting point for a broader discussion.
+We note here that this meaningful difference in the proportion of
+systematic variance attributable to intrapersonal change across
+education groups and across questions would not be detectable using
+previous methods. For partisan identification and political ideology,
+both college-educated respondents and people with less than college
+degree would likely favor the AUM or LCAM over the SDM because they show
+evidence of some members of the population making intrapersonal change.
+In other words, using the tournament of models approach obscures the
+fact that the relative importance of intrapersonal change in explaining
+differences on these two items is very different across these two
+groups, that intrapersonal change explains more systematic variance for
+ideological identification than partisan identification for people
+without a college degree, or that there appears to be more durable
+change on questions of affirmative action than on questions of
+government aid to black Americans.
 
-# Discussion
+It is important to note here that these differences should not be
+interpreted as causal effects of attending college on the importance of
+social experiences for cultural change. These results are potentially
+confounded by age, social class, race, gender, and other factors that
+explain selection into higher education. We simply use these estimates
+as a starting point for a broader discussion.
 
-This paper set out intervene in the debate on whether people change their personal culture -- their attitudes, beliefs, values, practices, and dispositions -- as they move through life.
-Instead of falling back on a "tournament of models" approach [@lersch2023: p.
-228] to call victory for one or the other possible answer to this question, this paper seeks to refocus the debate to the theoretically more productive question of _how_ initial differences between and durable changes within people collectively explain the observed variation in personal culture.
-To do so, it proposes a new measure that quantifies the amount of systematic variance that is attributable to either interpersonal differences at baseline or intrapersonal change.
+\hypertarget{discussion}{%
+\section{Discussion}\label{discussion}}
 
-Applying this measure to all datasets previously discussed in this debate revealed a consistent pattern.
-Nearly all questions on personal culture show evidence of intrapersonal change over time, with some items showing noteably high intrapersonal change compared to interpersonal differences.
-For questions about factors that affect one’s life satisfaction or views on government benefits, it seems plausible that differences in adult experiences predominantly account for the observed differences between people.
+This paper set out intervene in the debate on whether people change
+their personal culture -- their attitudes, beliefs, values, practices,
+and dispositions -- as they move through life. Instead of falling back
+on a ``tournament of models'' approach (Lersch 2023: p.~228) to call
+victory for one or the other possible answer to this question, this
+paper seeks to refocus the debate to the theoretically more productive
+question of \emph{how} initial differences between and durable changes
+within people collectively explain the observed variation in personal
+culture. To do so, it proposes a new measure that quantifies the amount
+of systematic variance that is attributable to either interpersonal
+differences at baseline or intrapersonal change.
 
-Intrapersonal change is often substantially less pronounced than interpersonal differences, accounting for less than 10 percent on average across all surveys.
-On numerous questions, like those on civil liberties, abortion, generalized trust, and civic duty, the systematic variance explained by intrapersonal change is essentially zero.
-For these questions, it seems there is not enough cultural change during adulthood to warrant attributing differences in personal culture mainly to experiences and social transitions; instead, the primary source of the observed differences appears to stem from early-life experiences.
+Applying this measure to all datasets previously discussed in this
+debate revealed a consistent pattern. Nearly all questions on personal
+culture show evidence of intrapersonal change over time, with some items
+showing noteably high intrapersonal change compared to interpersonal
+differences. For questions about factors that affect one's life
+satisfaction or views on government benefits, it seems plausible that
+differences in adult experiences predominantly account for the observed
+differences between people.
 
-More important than adjudicating past debates, however, the approach presented in this paper offers offers the possibility to pose new theoretical questions regarding when change or stability in personal culture predominates and its significance in explaining cultural variances within a population.
-To examplify this potential, we detailed how the proportion in cultural differences that intrapersonal change can explain not only varies substantially by survey item but also by individual characteristics such as their education.
-Intrapersonal change plays almost no part in explaining sentiments regarding the government's economic role but is more defining for ideological and partisan identities.
-Intriguingly, it is more important in explaining cultural variation among people without a college compared to college graduates in the U.S..
-While age and other factors related to college completion might confound this pattern, it suggests that college may solidify personal culture to an extent that renders the effect of other adult experiences on political dispositions for graduates less influential relative to individuals without such a college experience.
+Intrapersonal change is often substantially less pronounced than
+interpersonal differences, accounting for less than 10 percent on
+average across all surveys. On numerous questions, like those on civil
+liberties, abortion, generalized trust, and civic duty, the systematic
+variance explained by intrapersonal change is essentially zero. For
+these questions, it seems there is not enough cultural change during
+adulthood to warrant attributing differences in personal culture mainly
+to experiences and social transitions; instead, the primary source of
+the observed differences appears to stem from early-life experiences.
 
-## Implications for Cultural Sociology
+More important than adjudicating past debates, however, the approach
+presented in this paper offers offers the possibility to pose new
+theoretical questions regarding when change or stability in personal
+culture predominates and its significance in explaining cultural
+variances within a population. To examplify this potential, we detailed
+how the proportion in cultural differences that intrapersonal change can
+explain not only varies substantially by survey item but also by
+individual characteristics such as their education. Intrapersonal change
+plays almost no part in explaining sentiments regarding the government's
+economic role but is more defining for ideological and partisan
+identities. Intriguingly, it is more important in explaining cultural
+variation among people without a college compared to college graduates
+in the U.S.. While age and other factors related to college completion
+might confound this pattern, it suggests that college may solidify
+personal culture to an extent that renders the effect of other adult
+experiences on political dispositions for graduates less influential
+relative to individuals without such a college experience.
 
-So far, the debate in cultural sociology has centered around the presence and significance of cultural change in adults.
-Yet, arguing solely over its existence sets the wrong theoretical focus.
-In any adult population, some degree of personal cultural change is inevitable.
-Likewise, no viable theoretical perspective would expect just one process --stable differences or internal change between people-- to characterize the differences in personal cultural observed between them.
-Our results underscore this perspective, showing a combination of interpersonal differences and intrapersonal change across all items.
-[DELETED SENTECE, MISSPLACED]
+\hypertarget{implications-for-cultural-sociology}{%
+\subsection{Implications for Cultural
+Sociology}\label{implications-for-cultural-sociology}}
 
-A more reasonable theoretical debate pushed by previous work is the relative importance of these two processes for explaining cultural variation.
-It is difficult to draw a single conclusion on this question across a wide range of questions measured on different scales and over different durations of time, but the overall picture, with important caveats, suggests that intrapersonal change during adulthood is not the dominant explanation for cultural differences among adults for most items we measure in surveys.
-There simply is not enough evidence of intrapersonal change on these items relative to the amount of difference we observe across people.
+So far, the debate in cultural sociology has centered around the
+presence and significance of cultural change in adults. Yet, arguing
+solely over its existence sets the wrong theoretical focus. In any adult
+population, some degree of personal cultural change is inevitable.
+Likewise, no viable theoretical perspective would expect just one
+process --stable differences or internal change between people-- to
+characterize the differences in personal cultural observed between them.
+Our results underscore this perspective, showing a combination of
+interpersonal differences and intrapersonal change across all items.
+{[}DELETED SENTECE, MISSPLACED{]}
 
-To be clear, none of our results should be interpreted as suggesting that adults do not change or are incapable of changing.
-On all questions, some members of the sample change their personal culture at least a little bit.
-Our results also should not be taken to imply that intrapersonal change is not important or consequential.
-Small numbers of changers can have important consequences in the aggregate.
-A shift from 49 percent supporting gay marriage to 51 percent supporting it can have significant political ramifications, even as most respondents remain fixed in their preferences.
-And our results should not detract from work trying to estimate the mechanisms through which personal cultural differences --however small-- emerge among adults.
-Just because change is small does not mean the mechanisms that produce it are not important to understand.
+A more reasonable theoretical debate pushed by previous work is the
+relative importance of these two processes for explaining cultural
+variation. It is difficult to draw a single conclusion on this question
+across a wide range of questions measured on different scales and over
+different durations of time, but the overall picture, with important
+caveats, suggests that intrapersonal change during adulthood is not the
+dominant explanation for cultural differences among adults for most
+items we measure in surveys. There simply is not enough evidence of
+intrapersonal change on these items relative to the amount of difference
+we observe across people.
 
-Our results should be read principally as suggesting that if researchers want to explain why people differ in their personal culture, it is imperative to look at the conditions in which they were raised and the experiences that happened before they reached adulthood.
-To the extent that we want to explain what makes adults different, focusing attention on peoples’ transition between social roles, their selection into and out of organizational environments, and changes in their social networks, as much sociological work does, appears to explain a much smaller proportion of their differences than the formative experiences that happen early in life.
+To be clear, none of our results should be interpreted as suggesting
+that adults do not change or are incapable of changing. On all
+questions, some members of the sample change their personal culture at
+least a little bit. Our results also should not be taken to imply that
+intrapersonal change is not important or consequential. Small numbers of
+changers can have important consequences in the aggregate. A shift from
+49 percent supporting gay marriage to 51 percent supporting it can have
+significant political ramifications, even as most respondents remain
+fixed in their preferences. And our results should not detract from work
+trying to estimate the mechanisms through which personal cultural
+differences --however small-- emerge among adults. Just because change
+is small does not mean the mechanisms that produce it are not important
+to understand.
 
-In this way, our findings echo a range of recent causal inference work that suggests selection effects, rather than treatment effects, seem to explain the majority of differences in personal culture between people in different social roles or positions [@campbell2016; @wodtke2018].
-While some people clearly change over time as they move into new social roles or into new social environments, this amount of change does not appear to be sufficient to explain what are often substantial differences between people in different roles.
-Our results suggest that, to the extent we observe differences in personal culture between people in different social roles, such as parenthood, having children, education, or working in positions of authority [@longest2013], or between people in different occupations [@weeden2005], our prior expectation should be that this difference is selection, rather than treatment.
+Our results should be read principally as suggesting that if researchers
+want to explain why people differ in their personal culture, it is
+imperative to look at the conditions in which they were raised and the
+experiences that happened before they reached adulthood. To the extent
+that we want to explain what makes adults different, focusing attention
+on peoples' transition between social roles, their selection into and
+out of organizational environments, and changes in their social
+networks, as much sociological work does, appears to explain a much
+smaller proportion of their differences than the formative experiences
+that happen early in life.
 
-Our results do not speak to exactly why intrapersonal change lacks significant explanatory power in understanding culture.
-First, it could be the case that adults are exposed to opportunities for change at the same rate, but become less willing to change.
-Or it could be that opportunities for durable change or the social conditions that facilitate durable change are just extremely uncommon.
-Changes in social structures, the life course, or cultural expectations could make intrapersonal change a much more consequential explanation for cultural differences.
+In this way, our findings echo a range of recent causal inference work
+that suggests selection effects, rather than treatment effects, seem to
+explain the majority of differences in personal culture between people
+in different social roles or positions (Campbell and Horowitz 2016;
+Wodtke 2018). While some people clearly change over time as they move
+into new social roles or into new social environments, this amount of
+change does not appear to be sufficient to explain what are often
+substantial differences between people in different roles. Our results
+suggest that, to the extent we observe differences in personal culture
+between people in different social roles, such as parenthood, having
+children, education, or working in positions of authority (Longest,
+Hitlin, and Vaisey 2013), or between people in different occupations
+(Weeden and Grusky 2005), our prior expectation should be that this
+difference is selection, rather than treatment.
 
-Furthermore, the results we present related to education and political views suggest that the relative importance of different theoretical processes might differ by group.
-In line with life course perspectives, the results presented here suggest that importance of different experiences for cultural change are potentially contingent on previous experiences.
-For example, to the extent that we want to explain why people have different personal culture, the answer is likely different for people who have a college degree and those who do not.
-Mid-life experiences appear to have more significant implications for the personal culture for people without a college degree than for people with a college degree.
+Our results do not speak to exactly why intrapersonal change lacks
+significant explanatory power in understanding culture. First, it could
+be the case that adults are exposed to opportunities for change at the
+same rate, but become less willing to change. Or it could be that
+opportunities for durable change or the social conditions that
+facilitate durable change are just extremely uncommon. Changes in social
+structures, the life course, or cultural expectations could make
+intrapersonal change a much more consequential explanation for cultural
+differences.
 
-## Implications for Social Change
+Furthermore, the results we present related to education and political
+views suggest that the relative importance of different theoretical
+processes might differ by group. In line with life course perspectives,
+the results presented here suggest that importance of different
+experiences for cultural change are potentially contingent on previous
+experiences. For example, to the extent that we want to explain why
+people have different personal culture, the answer is likely different
+for people who have a college degree and those who do not. Mid-life
+experiences appear to have more significant implications for the
+personal culture for people without a college degree than for people
+with a college degree.
 
-Because intrapersonal change in personal culture does not appear to be a driving force in cultural differences among adults, our results suggest that processes of social and cultural change at the aggregate level are likely driven more by cohort replacement, one group of people with certain cultural dispositions exiting the population and a different set of people with a potentially different set of cultural dispositions entering the population, than by contemporaneous social conditions [@vaisey2016; @underwood2022].
-This is likely true for both large-scale change in societies over time as well as changes within organizations, political parties, occupations, and other social aggregates.
+\hypertarget{implications-for-social-change}{%
+\subsection{Implications for Social
+Change}\label{implications-for-social-change}}
 
-However, because our results suggest that the role of formative experiences compared to contemporaneous environments differs by a person's level of education, changes in educational attainment over time, as well as changes in the social structuring of the life course might be shifting the relative importance of these two factors over time.
-If we observe further differences across socio-demographic groups like education, social class, and race/ethnicity in the relative importance of these processes, this would argue for a greater incorporation of the interaction of demographic processes with processes of individual cultural change into our explanations of aggregate cultural change.
+Because intrapersonal change in personal culture does not appear to be a
+driving force in cultural differences among adults, our results suggest
+that processes of social and cultural change at the aggregate level are
+likely driven more by cohort replacement, one group of people with
+certain cultural dispositions exiting the population and a different set
+of people with a potentially different set of cultural dispositions
+entering the population, than by contemporaneous social conditions
+(Underwoood et al. 2022; Vaisey and Lizardo 2016). This is likely true
+for both large-scale change in societies over time as well as changes
+within organizations, political parties, occupations, and other social
+aggregates.
 
-## Implications for Survey Measurement
+However, because our results suggest that the role of formative
+experiences compared to contemporaneous environments differs by a
+person's level of education, changes in educational attainment over
+time, as well as changes in the social structuring of the life course
+might be shifting the relative importance of these two factors over
+time. If we observe further differences across socio-demographic groups
+like education, social class, and race/ethnicity in the relative
+importance of these processes, this would argue for a greater
+incorporation of the interaction of demographic processes with processes
+of individual cultural change into our explanations of aggregate
+cultural change.
 
-A major takeaway from the meta-analysis is that greater response resolution –-the number of response options survey respondents are given-– is associated with a greater share of variance being explained by within-individual change.
-Again, it is possible that this reflects different kinds of issues being measured with different kinds of scales.
-But it could also suggest that within-person change, when it happens, is relatively small –-moving from “agree” to “strongly agree,” rather than “strongly agree” to “strongly disagree,” for example-– and therefore easier to detect with greater resolution.
+\hypertarget{implications-for-survey-measurement}{%
+\subsection{Implications for Survey
+Measurement}\label{implications-for-survey-measurement}}
 
-Our results continue to highlight the value of long-running panel surveys for adjudicating different theories of culture that cannot be evaluated with cross-sectional data alone.
-Across our results, residual fluctuations in responses account for more than four times as much variance, on average, as linear persistent change, meaning that most change in responses is not systematic.
-Under these circumstances, extending panels beyond two waves provides significantly more leverage to understand how and why personal culture changes.
+A major takeaway from the meta-analysis is that greater response
+resolution ---the number of response options survey respondents are
+given-\/-- is associated with a greater share of variance being
+explained by within-individual change. Again, it is possible that this
+reflects different kinds of issues being measured with different kinds
+of scales. But it could also suggest that within-person change, when it
+happens, is relatively small ---moving from ``agree'' to ``strongly
+agree,'' rather than ``strongly agree'' to ``strongly disagree,'' for
+example-\/-- and therefore easier to detect with greater resolution.
 
-Additionally, we echo previous calls for more studies –-including panel studies-– of the socialization processes of young people.
-Our results suggest that most of the variance between adults emerges before people turn 18 or relatively soon after.
-In limiting our empirical perspective to experiences that happen after that point, social scientists potentially focus on questions that, while important, lack the capacity to explain major cultural differences.
+Our results continue to highlight the value of long-running panel
+surveys for adjudicating different theories of culture that cannot be
+evaluated with cross-sectional data alone. Across our results, residual
+fluctuations in responses account for more than four times as much
+variance, on average, as linear persistent change, meaning that most
+change in responses is not systematic. Under these circumstances,
+extending panels beyond two waves provides significantly more leverage
+to understand how and why personal culture changes.
 
-## Limitations
+Additionally, we echo previous calls for more studies ---including panel
+studies-\/-- of the socialization processes of young people. Our results
+suggest that most of the variance between adults emerges before people
+turn 18 or relatively soon after. In limiting our empirical perspective
+to experiences that happen after that point, social scientists
+potentially focus on questions that, while important, lack the capacity
+to explain major cultural differences.
 
-The approach outlined here generates estimates of how much variance in a population is due to each of two broad processes: intrapersonal change and interpersonal differences at baseline.
-It does not quantify the proportion of people who “change.” When we observe intrapersonal change in a question, we cannot be sure whether this is driven by many people making small changes or a few people making large changes.
-Adjudicating this question will require different models that distinguish different kinds of change.
+\hypertarget{limitations}{%
+\subsection{Limitations}\label{limitations}}
 
-Our results and interpretations also hinge on how we have defined change and the assumptions built into the Life Course Adaption Model.
-This model treats cultural trajectories as varying linear slopes for each respondent, assuming that change is a linear function of time.
-This assumption likely does not hold true in practice.
-Change over time likely takes non-linear and discontinuous forms.
-People might jump from one “stable” disposition to another, for example, or experience a discontinuous "turning point" that upsets a stable trajectory.
-This assumption is in part a limitation of the data, as most questions are only observed for three to six waves with a large number of missing cases.
-Panels with more waves might allow for more flexible models of change that do not require this assumption of linearity.
+The approach outlined here generates estimates of how much variance in a
+population is due to each of two broad processes: intrapersonal change
+and interpersonal differences at baseline. It does not quantify the
+proportion of people who ``change.'' When we observe intrapersonal
+change in a question, we cannot be sure whether this is driven by many
+people making small changes or a few people making large changes.
+Adjudicating this question will require different models that
+distinguish different kinds of change.
 
-Similarly, in assuming a linear trajectory for change, our approach assumes that durable change is unidirectional.
-To the extent that people make durable change as a result of a social transition and then return to a previous position later in life as a result of a second transition (marriage and divorce, for example), this is captured as residual within-person variance, rather than one of our two broad theoretical processes.
-This is a potentially reasonable assumption on short panels --change that does not last two years is hard to classify as durable-- but it might not be true in longer panels.
-Other researchers with more flexible definitions of "durable change" might potentially identify more durable change even in the same population.
+Our results and interpretations also hinge on how we have defined change
+and the assumptions built into the Life Course Adaption Model. This
+model treats cultural trajectories as varying linear slopes for each
+respondent, assuming that change is a linear function of time. This
+assumption likely does not hold true in practice. Change over time
+likely takes non-linear and discontinuous forms. People might jump from
+one ``stable'' disposition to another, for example, or experience a
+discontinuous ``turning point'' that upsets a stable trajectory. This
+assumption is in part a limitation of the data, as most questions are
+only observed for three to six waves with a large number of missing
+cases. Panels with more waves might allow for more flexible models of
+change that do not require this assumption of linearity.
 
-Finally, while we examine an incredible broad array of measures of personal culture, from assessments about the relative importance of different features when buying a new car and policy preferences to religious beliefs and core values, we are limited by the kinds of questions that are asked in panel surveys and the contexts in which these surveys were administered (Australia, Germany, Switzerland, the United Kingdom and the United States).
-This means our findings tend to reflect questions about general (national) politics, gender roles, immigration and race relations, and general well being.
-Our data does not cover some dimensions of culture that might be relevant for other aspects of sociological exploration (artistic tastes, leisure activities, and time use, for example).
+Similarly, in assuming a linear trajectory for change, our approach
+assumes that durable change is unidirectional. To the extent that people
+make durable change as a result of a social transition and then return
+to a previous position later in life as a result of a second transition
+(marriage and divorce, for example), this is captured as residual
+within-person variance, rather than one of our two broad theoretical
+processes. This is a potentially reasonable assumption on short panels
+--change that does not last two years is hard to classify as durable--
+but it might not be true in longer panels. Other researchers with more
+flexible definitions of ``durable change'' might potentially identify
+more durable change even in the same population.
 
-## Implications for Future Research
+Finally, while we examine an incredible broad array of measures of
+personal culture, from assessments about the relative importance of
+different features when buying a new car and policy preferences to
+religious beliefs and core values, we are limited by the kinds of
+questions that are asked in panel surveys and the contexts in which
+these surveys were administered (Australia, Germany, Switzerland, the
+United Kingdom and the United States). This means our findings tend to
+reflect questions about general (national) politics, gender roles,
+immigration and race relations, and general well being. Our data does
+not cover some dimensions of culture that might be relevant for other
+aspects of sociological exploration (artistic tastes, leisure
+activities, and time use, for example).
 
-We believe the approach outlined here can push past the “needless dichotomy” implicit in the question of whether people change or not and address more theoretically informed questions about the conditions under which the relative importance of intrapersonal change and interpersonal differences matter for explaining cultural differences.
-Characterizing questions as displaying change or not can only take researchers so far, but the question of why some questions demonstrate more change while others do not, or why some groups are characterized by more stability than others, has the potential to weigh in on a broader range of theoretical debates.
-We hope researchers move toward those questions.
+\hypertarget{implications-for-future-research}{%
+\subsection{Implications for Future
+Research}\label{implications-for-future-research}}
 
-While we focused on education as a determinant of the explanatory power of intrapersonal change, our approach can readily be extended to other socio-demographic categories that might affect intrapersonal change --race/ethnicity, religiosity, social class, gender, and political affiliation, for example-- or their intersections.
-Researchers with different data might be able to evaluate the relative importance of intrapersonal change and interpersonal differences across contexts as well, either national contexts if the same questions are administered across countries, or organizational contexts if data traces students within schools, workers within workplaces, or congregation members within congregations.
+We believe the approach outlined here can push past the ``needless
+dichotomy'' implicit in the question of whether people change or not and
+address more theoretically informed questions about the conditions under
+which the relative importance of intrapersonal change and interpersonal
+differences matter for explaining cultural differences. Characterizing
+questions as displaying change or not can only take researchers so far,
+but the question of why some questions demonstrate more change while
+others do not, or why some groups are characterized by more stability
+than others, has the potential to weigh in on a broader range of
+theoretical debates. We hope researchers move toward those questions.
 
-The wide variation across questions in the proportion of systematic variation attributable to intrapersonal change also calls for a more systematic exploration for why some forms of culture are characterized by stability while others are characterized by change, and what the consequences of these differences for the organization of beliefs in the public might be.
+While we focused on education as a determinant of the explanatory power
+of intrapersonal change, our approach can readily be extended to other
+socio-demographic categories that might affect intrapersonal change
+--race/ethnicity, religiosity, social class, gender, and political
+affiliation, for example-- or their intersections. Researchers with
+different data might be able to evaluate the relative importance of
+intrapersonal change and interpersonal differences across contexts as
+well, either national contexts if the same questions are administered
+across countries, or organizational contexts if data traces students
+within schools, workers within workplaces, or congregation members
+within congregations.
 
-# References
+The wide variation across questions in the proportion of systematic
+variation attributable to intrapersonal change also calls for a more
+systematic exploration for why some forms of culture are characterized
+by stability while others are characterized by change, and what the
+consequences of these differences for the organization of beliefs in the
+public might be.
 
-<div id="refs"></div>
+\hypertarget{references}{%
+\section{References}\label{references}}
+
+\hypertarget{refs}{}
+\begin{CSLReferences}{1}{0}
+\leavevmode\vadjust pre{\hypertarget{ref-alvarez2002}{}}%
+Alvarez, R. Michael, and John Brehm. 2002. \emph{Hard Choices, Easy
+Answers: Values, Information, and {American} Public Opinion}. Princeton,
+N.J: Princeton University Press.
+
+\leavevmode\vadjust pre{\hypertarget{ref-alwin2007}{}}%
+Alwin, Duane F. 2007. \emph{Margins of {Error}: {A} {Study} of
+{Reliability} in {Survey} {Measurement}}. Hoboken, N.J.: John Wiley \&
+Sons.
+
+\leavevmode\vadjust pre{\hypertarget{ref-alwin1991}{}}%
+Alwin, Duane F., and Jon A. Krosnick. 1991.
+{``\href{https://www.jstor.org/stable/2781642}{Aging, {Cohorts}, and the
+{Stability} of {Sociopolitical} {Orientations} {Over} the {Life}
+{Span}}.''} \emph{American Journal of Sociology} 97(1):169--95.
+
+\leavevmode\vadjust pre{\hypertarget{ref-bail2018}{}}%
+Bail, Christopher A., Lisa P. Argyle, Taylor W. Brown, John P. Bumpus,
+Haohan Chen, M. B. Fallin Hunzaker, Jaemin Lee, Marcus Mann, Friedolin
+Merhout, and Alexander Volfovsky. 2018. {``Exposure to Opposing Views on
+Social Media Can Increase Political Polarization.''} \emph{Proceedings
+of the National Academy of Sciences} 115(37):9216--21. doi:
+\href{https://doi.org/10.1073/pnas.1804840115}{10.1073/pnas.1804840115}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-bardi2011}{}}%
+Bardi, Anat, and Robin Goodwin. 2011. {``The {Dual} {Route} to {Value}
+{Change}: {Individual} {Processes} and {Cultural} {Moderators}.''}
+\emph{Journal of Cross-Cultural Psychology} 42(2):271--87. doi:
+\href{https://doi.org/10.1177/0022022110396916}{10.1177/0022022110396916}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-bourdieu1990}{}}%
+Bourdieu, Pierre. 1990. \emph{The Logic of Practice}. Stanford, Calif. :
+Stanford University Press, 1990.
+
+\leavevmode\vadjust pre{\hypertarget{ref-boutyline2017}{}}%
+Boutyline, Andrei, and Stephen Vaisey. 2017. {``Belief {Network}
+{Analysis}: {A} {Relational} {Approach} to {Understanding} the
+{Structure} of {Attitudes}.''} \emph{American Journal of Sociology}
+122(5):1371--1447. doi:
+\href{https://doi.org/10.1086/691274}{10.1086/691274}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-brocic2021}{}}%
+Broćić, Miloš, and Andrew Miles. 2021. {``College and the {`{Culture}
+{War}'}: {Assessing} {Higher} {Education}'s {Influence} on {Moral}
+{Attitudes}.''} \emph{American Sociological Review} 00031224211041094.
+doi:
+\href{https://doi.org/10.1177/00031224211041094}{10.1177/00031224211041094}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-campbell2016}{}}%
+Campbell, Colin, and Jonathan Horowitz. 2016. {``Does {College}
+{Influence} {Sociopolitical} {Attitudes}?''} \emph{Sociology of
+Education} 89(1):40--58. doi:
+\href{https://doi.org/10.1177/0038040715617224}{10.1177/0038040715617224}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-christakis2010}{}}%
+Christakis, Nicholas, and James Fowler. 2010. \emph{Connected: {The}
+{Amazing} {Power} of {Social} {Networks} and {How} {They} {Shape} {Our}
+{Lives}}. HarperCollins Publishers.
+
+\leavevmode\vadjust pre{\hypertarget{ref-dellaposta2018}{}}%
+DellaPosta, Daniel. 2018. {``Gay {Acquaintanceship} and {Attitudes}
+Toward {Homosexuality}: {A} {Conservative} {Test}.''} \emph{Socius}
+4:2378023118798959. doi:
+\href{https://doi.org/10.1177/2378023118798959}{10.1177/2378023118798959}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-dellaposta2020}{}}%
+DellaPosta, Daniel. 2020. {``Pluralistic {Collapse}: {The} {`{Oil}
+{Spill}'} {Model} of {Mass} {Opinion} {Polarization}.''} \emph{American
+Sociological Review} 85(3):507--36. doi:
+\href{https://doi.org/10.1177/0003122420922989}{10.1177/0003122420922989}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-eaton2009}{}}%
+Eaton, Asia A., Penny S. Visser, Jon A. Krosnick, and Sowmya Anand.
+2009. {``Social {Power} and {Attitude} {Strength} {Over} the {Life}
+{Course}.''} \emph{Personality and Social Psychology Bulletin}
+35(12):1646--60. doi:
+\href{https://doi.org/10.1177/0146167209349114}{10.1177/0146167209349114}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-elder2016}{}}%
+Elder, Glen H., and Linda K. George. 2016.
+{``\href{https://doi.org/10.1007/978-3-319-20880-0_3}{Age, {Cohorts},
+and the {Life} {Course}}.''} Pp. 59--85 in \emph{Handbook of the {Life}
+{Course}: {Volume} {II}}, \emph{Handbooks of {Sociology} and {Social}
+{Research}}, edited by M. J. Shanahan, J. T. Mortimer, and M.
+Kirkpatrick Johnson. Cham: Springer International Publishing.
+
+\leavevmode\vadjust pre{\hypertarget{ref-feldman1992}{}}%
+Feldman, Stanley, and John Zaller. 1992. {``The {Political} {Culture} of
+{Ambivalence}: {Ideological} {Responses} to the {Welfare} {State}.''}
+\emph{American Journal of Political Science} 36(1):268--307. doi:
+\href{https://doi.org/10.2307/2111433}{10.2307/2111433}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-gelman2021}{}}%
+Gelman, Andrew, and Yotam Margalit. 2021. {``Social Penumbras Predict
+Political Attitudes.''} \emph{Proceedings of the National Academy of
+Sciences} 118(6):e2019375118. doi:
+\href{https://doi.org/10.1073/pnas.2019375118}{10.1073/pnas.2019375118}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-data_soep}{}}%
+Goebel, Jan, Markus M. Grabka, Stefan Liebig, Martin Kroh, David
+Richter, Carsten Schröder, and Jürgen Schupp. 2019. {``The German
+Socio-Economic Panel (SOEP).''} \emph{Jahrb{ü}cher f{ü}r
+National{ö}konomie Und Statistik} 239(2):345--60.
+
+\leavevmode\vadjust pre{\hypertarget{ref-gross2009}{}}%
+Gross, Neil. 2009. {``A {Pragmatist} {Theory} of {Social}
+{Mechanisms}.''} \emph{American Sociological Review} 74(3):358--79. doi:
+\href{https://doi.org/10.1177/000312240907400302}{10.1177/000312240907400302}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-data_psid}{}}%
+Income Dynamics, Panel Study of. 2013. {``Public Use Dataset.''}
+\emph{The Survey Research Center Institute for Social Research,
+University of Michgan}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-kiley2020}{}}%
+Kiley, Kevin, and Stephen Vaisey. 2020. {``Measuring {Stability} and
+{Change} in {Personal} {Culture} {Using} {Panel} {Data}.''}
+\emph{American Sociological Review} 85(3):477--506. doi:
+\href{https://doi.org/10.1177/0003122420921538}{10.1177/0003122420921538}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-kratz2021}{}}%
+Kratz, Fabian. 2021. {``Do Concerns about Immigration Change After
+Adolescence? How Education and Critical Life Events Affect Concerns
+about Immigration.''} \emph{European Sociological Review}
+37(6):987--1003.
+
+\leavevmode\vadjust pre{\hypertarget{ref-krosnick1989}{}}%
+Krosnick, Jon, and Duane F. Alwin. 1989. {``Aging and {Susceptibility}
+to {Attitude} {Change}.''} \emph{Journal of Personality and Social
+Psychology} 57:416--25. doi:
+\href{https://doi.org/10.1037//0022-3514.57.3.416}{10.1037//0022-3514.57.3.416}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-lersch2023}{}}%
+Lersch, Philipp M. 2023. {``Change in {Personal} {Culture} over the
+{Life} {Course}.''} \emph{American Sociological Review} 88(2):220--51.
+doi:
+\href{https://doi.org/10.1177/00031224231156456}{10.1177/00031224231156456}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-lizardo2017}{}}%
+Lizardo, Omar. 2017. {``Improving {Cultural} {Analysis}: {Considering}
+{Personal} {Culture} in Its {Declarative} and {Nondeclarative}
+{Modes}.''} \emph{American Sociological Review} 82(1):88--115. doi:
+\href{https://doi.org/10.1177/0003122416675175}{10.1177/0003122416675175}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-longest2013}{}}%
+Longest, K. C., S. Hitlin, and S. Vaisey. 2013. {``Position and
+{Disposition}: {The} {Contextual} {Development} of {Human} {Values}.''}
+\emph{Social Forces} 91(4):1499--1528. doi:
+\href{https://doi.org/10.1093/sf/sot045}{10.1093/sf/sot045}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-mannheim1952}{}}%
+Mannheim, Karl. 1952. {``The {Problem} of {Generations}.''} Pp. 276--322
+in \emph{Essays on the {Sociology} of {Knowledge}}. New York: Oxford
+University Press.
+
+\leavevmode\vadjust pre{\hypertarget{ref-mewes2021}{}}%
+Mewes, Jan, Malcolm Fairbrother, Giuseppe Nicola Giordano, Cary Wu, and
+Rima Wilkes. 2021. {``Experiences Matter: A Longitudinal Study of
+Individual-Level Sources of Declining Social Trust in the United
+States.''} \emph{Social Science Research} 95:102537.
+
+\leavevmode\vadjust pre{\hypertarget{ref-paul2014transformative}{}}%
+Paul, Laurie Ann. 2014. \emph{Transformative Experience}. OUP Oxford.
+
+\leavevmode\vadjust pre{\hypertarget{ref-robinson2007}{}}%
+Robinson, Dawn T. 2007. {``Control {Theories} in {Sociology}.''}
+\emph{Annual Review of Sociology} 33(1):157--74. doi:
+\href{https://doi.org/10.1146/annurev.soc.32.061604.123110}{10.1146/annurev.soc.32.061604.123110}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-ryder1965}{}}%
+Ryder, Norman B. 1965. {``The {Cohort} as a {Concept} in the {Study} of
+{Social} {Change}.''} \emph{American Sociological Review} 30(6):843--61.
+doi: \href{https://doi.org/10.2307/2090964}{10.2307/2090964}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-slothuus2021}{}}%
+Slothuus, Rune, and Martin Bisgaard. 2021. {``How {Political} {Parties}
+{Shape} {Public} {Opinion} in the {Real} {World}.''} \emph{American
+Journal of Political Science} 65(4):896--911. doi:
+\href{https://doi.org/10.1111/AJPS.12550}{10.1111/AJPS.12550}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-data_gss}{}}%
+Smith, Tom W., Michael Davern, Jeremy Freese, and Stephen L. Morgan.
+2022. {``General Social Surveys, 1972-2022: Cumulative Codebook /
+Principal Investigator, Tom w. Smith; Co-Principal Investigators,
+Michael Davern, Jeremy Freese and Stephen l. Morgan.''} \emph{Chicago:
+NORC}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-smith-lovin1988}{}}%
+Smith-Lovin, Lynn, and David R. Heise. 1988. \emph{Analyzing {Social}
+{Interaction}: {Advances} in {Affect} {Control} {Theory}}. New York:
+Gordon; Breach Sciences Publishers.
+
+\leavevmode\vadjust pre{\hypertarget{ref-data_hilda}{}}%
+Summerfield, Michelle, Ross Dunn, Simon Freidin, Markus Hahn, Peter
+Ittak, Milica Kecmanovic, Ning Li, Ninette Macalalad, Nicole Watson,
+Roger Wilkins, and others. 2011. {``HILDA User Manual -- Release 10.''}
+\emph{Melbourne Institute of Applied Economic and Social Research,
+University of Melbourne}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-swidler2001}{}}%
+Swidler, Ann. 2001. \emph{Talk of {Love}: {How} {Culture} {Matters}}.
+Chicago: University of Chicago Press.
+
+\leavevmode\vadjust pre{\hypertarget{ref-data_bhps}{}}%
+Taylor, Marcia Freed. 1996. {``British Household Panel Survey User
+Manual: Volume a-Introduction, Technical Report and Appendices.''}
+
+\leavevmode\vadjust pre{\hypertarget{ref-tourangeau2000}{}}%
+Tourangeau, Roger, Lance J. Rips, and Kenneth A. Rasinski. 2000.
+\emph{The Psychology of Survey Response}. Cambridge, U.K. ; New York:
+Cambridge University Press.
+
+\leavevmode\vadjust pre{\hypertarget{ref-underwood2022}{}}%
+Underwoood, Ted, Kevin Kiley, Wenyi Shang, and Stephen Vaisey. 2022.
+{``Cohort Succession Explains Most Change in Literary Culture.''}
+\emph{Sociological Science} 9:184--205.
+
+\leavevmode\vadjust pre{\hypertarget{ref-data_ukhls}{}}%
+University of Essex, Institute for Social, and Economic Research. 2019.
+{``Understanding Society: Waves 1--9, 2009--2018 and Harmonised BHPS:
+Waves 1--18, 1991--2009.''} \emph{UK Data Service}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-vaisey2021}{}}%
+Vaisey, Stephen, and Kevin Kiley. 2021. {``A {Model}-{Based} {Method}
+for {Detecting} {Persistent} {Cultural} {Change} {Using} {Panel}
+{Data}.''} \emph{Sociological Science} 8:83--95. doi:
+\href{https://doi.org/10.15195/v8.a5}{10.15195/v8.a5}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-vaisey2010a}{}}%
+Vaisey, Stephen, and Omar Lizardo. 2010. {``Can {Cultural} {Worldviews}
+{Influence} {Network} {Composition}?''} \emph{Social Forces}
+88(4):1595--1618. doi:
+\href{https://doi.org/10.1353/sof.2010.0009}{10.1353/sof.2010.0009}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-vaisey2016}{}}%
+Vaisey, Stephen, and Omar Lizardo. 2016. {``Cultural {Fragmentation} or
+{Acquired} {Dispositions}? {A} {New} {Approach} to {Accounting} for
+{Patterns} of {Cultural} {Change}.''} \emph{Socius} 2:2378023116669726.
+doi:
+\href{https://doi.org/10.1177/2378023116669726}{10.1177/2378023116669726}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-visser2004}{}}%
+Visser, Penny S., and Robert R. Mirabile. 2004. {``Attitudes in the
+Social Context: {The} Impact of Social Network Composition on
+Individual-Level Attitude Strength.''} \emph{Journal of Personality and
+Social Psychology} 779--95.
+
+\leavevmode\vadjust pre{\hypertarget{ref-data_shp}{}}%
+Voorpostel, Marieke, Robin Tillmann, Florence Lebert, Ursina Kuhn,
+Oliver Lipps, Valérie-Anne Ryser, and B. Wernli. 2016. {``Swiss
+Household Panel User Guide (1999--2015).''} \emph{Lausanne: FORS}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-weeden2005}{}}%
+Weeden, Kim A., and David B. Grusky. 2005. {``The {Case} for a {New}
+{Class} {Map}.''} \emph{American Journal of Sociology} 111(1):141--212.
+doi: \href{https://doi.org/10.1086/428815}{10.1086/428815}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-wodtke2018}{}}%
+Wodtke, Geoffrey T. 2018. {``The {Effects} of {Education} on {Beliefs}
+about {Racial} {Inequality}.''} \emph{Social Psychology Quarterly}
+81(4):273--94. doi:
+\href{https://doi.org/10.1177/0190272518804145}{10.1177/0190272518804145}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-zaller1992}{}}%
+Zaller, John. 1992.
+\emph{\href{https://doi.org/10.1017/CBO9780511818691}{The {Nature} and
+{Origins} of {Mass} {Opinion}}}. Cambridge: Cambridge University Press.
+
+\end{CSLReferences}
 
@@
 
-# Appendix
+\hypertarget{appendix}{%
+\section{Appendix}\label{appendix}}
 
-## Tables
+\hypertarget{tables}{%
+\subsection{Tables}\label{tables}}
 
-```{r, message = FALSE}
+\begin{table}[H]
 
-library(tidyverse)
-library(knitr)
-library(kableExtra)
-
-table <- tibble(
-  Country = c("Australia", "Germany", "Great Britian", "Great Britian", "Switzerland", "United States", "United States"),
-  Survey = c("Household, Income and Labor Dynamics in Australia (HILDA)",
-             "Socio-Economic Panel Study (SOEP)",
-             "British Household Panel Survey (BHPS)",
-             "Understanding Society/UK Household Longitudinal Study (UKHLS)",
-             "Swiss Household Panel (SHP)",
-             "General Social Survey (GSS)",
-             "Panel Study of Income Dynamics (PSID)"),
-  Period = c("2001-2021", "1984-2020", "1991-2008", "2009-2020", "1999-2019", "2006-2012", "1968-2019"),
-  Outcomes = c(30, 122, 83, 66, 77, 183, 48)
-)
-
-kableExtra::kable(table, 
-                  format = "latex", 
-                  caption = "The Description of the Data Sources") |> 
-  kable_styling(latex_options = c("scale_down", "HOLD_position"))
-
-```
+\caption{\label{tab:unnamed-chunk-1}The Description of the Data Sources}
+\centering
+\resizebox{\linewidth}{!}{
+\begin{tabular}[t]{l|l|l|r}
+\hline
+Country & Survey & Period & Outcomes\\
+\hline
+Australia & Household, Income and Labor Dynamics in Australia (HILDA) & 2001-2021 & 30\\
+\hline
+Germany & Socio-Economic Panel Study (SOEP) & 1984-2020 & 122\\
+\hline
+Great Britian & British Household Panel Survey (BHPS) & 1991-2008 & 83\\
+\hline
+Great Britian & Understanding Society/UK Household Longitudinal Study (UKHLS) & 2009-2020 & 66\\
+\hline
+Switzerland & Swiss Household Panel (SHP) & 1999-2019 & 77\\
+\hline
+United States & General Social Survey (GSS) & 2006-2012 & 183\\
+\hline
+United States & Panel Study of Income Dynamics (PSID) & 1968-2019 & 48\\
+\hline
+\end{tabular}}
+\end{table}
 
@@
 
-## Figures
+\hypertarget{figures}{%
+\subsection{Figures}\label{figures}}
 
-```{r fig1, fig.height=2, fig.cap="Proportions of explained variance in items of personal culture; based on predictions at participants' wave mid-point."}
-knitr::include_graphics("fig/variation_share_lm_mdpnt.png")
-```
+\begin{figure}
+
+{\centering \includegraphics[width=20.71in]{fig/variation_share_lm_mdpnt}
+
+}
+
+\caption{Proportions of explained variance in items of personal culture; based on predictions at participants' wave mid-point.}\label{fig:fig1}
+\end{figure}
 
 \newpage
-```{r fig2, out.width = "300px", fig.cap="Proportion of explained variance attributable to intrapersonal change; based on predictions at participants' wave mid-point; supressed intercept model; survey indicators not shown."}
-knitr::include_graphics("fig/meta_no_int.png")
-```
+\begin{figure}
+
+{\centering \includegraphics[width=300px]{fig/meta_no_int}
+
+}
+
+\caption{Proportion of explained variance attributable to intrapersonal change; based on predictions at participants' wave mid-point; supressed intercept model; survey indicators not shown.}\label{fig:fig2}
+\end{figure}
 
 \newpage
-```{r fig3, out.width = "450px", fig.cap="Proportions of intrapersonal change of explained variance by college degree; based on predictions at participants' wave mid-point."}
-knitr::include_graphics("fig/FIG_college2.png")
-```
+\begin{figure}
+
+{\centering \includegraphics[width=450px]{fig/FIG_college1}
+
+}
+
+\caption{Figure 3: Distribution of differences share of variance explained by intrapersonal change between respondents without and with college degree.}\label{fig:fig3}
+\end{figure}
+
+\newpage
+\begin{figure}
+
+{\centering \includegraphics[width=450px]{fig/FIG_college2}
+
+}
+
+\caption{Proportions of intrapersonal change of explained variance by college degree; based on predictions at participants' wave mid-point.}\label{fig:fig4}
+\end{figure}
+
+\end{document}

### 2023-09-24 manuscript.Rmd
intro changes; adjusted formatting to fit ASR requirements

@@
     keep_tex: true
+geometry: margin=1in 
+mainfont: Times New Roman
+fontsize: 12pt
+header-includes: \usepackage{setspace}\doublespacing
 bibliography: kileybib.bib
@@
 We examplify this with the case of college experience."
-fontsize: 11pt
 ---
@@
 
-Because all of the above theoretical perspectives have some empirical support, the current theoretical debate revolves around not whether they exist, but when and how they each apply and the relative contribution of each to explaining cultural differences.
+Because all of the above theoretical perspectives have some empirical support, the current debate revolves around not whether they exist, but when and how they each apply and the relative contribution of each to explaining cultural differences.
 That being said, scholars have struggled to reach a consensus on the relative importance of change during adulthood.
 Over the span of a few years, most apparent changes in people’s responses to personal culture items appear to be transitory, with little evidence of persistent change among adults [@kiley2020; @vaisey2021].
-These results suggest change is not a strong force in explaining personal cultural differences.
-When looked at over longer time horizons, however, there is at least some evidence that adults make persistent cultural changes, especially as they move through distinct social roles across the life course [@lersch2023].
+This suggests that change during adulthood is not a strong force in explaining personal cultural differences.
+When looked at over longer time horizons, however, there is at least some evidence that adults make persistent cultural changes, especially as they move through distinct social roles [@lersch2023].
 
-These seemingly inconsistent findings are partly due to the fact that researchers have taken a “tournament of models” approach in adjudicating the different perspectives [@lersch2023: p.
+These seemingly inconsistent findings are partly due to the fact that researchers have taken a “tournament of models” approach in adjudicating different theoretical perspectives [@lersch2023: p.
 228].
-Accordingly, researchers have formalized models of responses to cultural items over time as demonstrating either “change” or “no change” to then compare which model fits observed data better, thus functionally asking whether we ever observe within-person change in personal culture.
+Accordingly, researchers have formalized models of responses to cultural items over time as demonstrating either “change” or “no change” to then compare their empirical fit, which essentially asks whether we ever observe within-person change in personal culture [WE COULD ALREADY HERE SWTICH TO "INTRAPERSONAL CHANGE" OR INTRODUCE THE TERM AS AN ALTERNATIVE TO WITHIN-PERSON CHANGE].
 
-We argue, however, that asking only *whether* people change obscures what are actually highly compatible empirical patterns.
-Whether researchers are able to detect "change" in a given dataset depends not only on sample populations, item measurement, statistical methods and power, but crucially also on definitions of what counts as “change.” These methodological differences therefore limit the possibility of convergence on the theoretical debates.
-Moreover, asking whether these is evidence of change limits the kind of questions that can be asked to, “Do we see change in this measure?” But such inquiries are ill-suited to address what’s at the core of the theoretical debate.
+We argue that asking only *whether* people change obscures what are actually highly compatible empirical patterns.
+Being able to detect "change" in a given dataset depends not only on sample populations, item measurement, statistical methods and power, but crucially also on definitions of what counts as “change.” These methodological differences therefore limit the possibility of convergence on the theoretical debates.
+Moreover, asking whether there is evidence of change limits the kind of questions that can be asked to, “Do we see change in this measure?” But such inquiries are ill-suited to address what’s at the core of the theoretical debate.
 There is enough evidence to assume that neither socialization at an early age nor the influence of the contemporary environment alone can explain cultural differences.
@@
 
-In this paper, we try to move beyond previous debates by moving away from the “tournament of models” approach toward a quantification of the processes that lead to differences in personal culture.
-Drawing on the seven panel datasets analyzed in these debates, we use established methods to quantify the amount of variance in responses that is due to between-people differences at a single point during the survey versus the amount of variance in responses that is due to linear within-person change over time.
+In this paper, we try to intervene in this debate by moving beyond the “tournament of models” approach toward a quantification of the processes that lead to differences in personal culture.
+Drawing on the seven panel datasets analyzed in the debate, we use established methods to quantify the amount of variance in responses that is due to between-people differences at a single point during the survey versus the amount of variance in responses that is due to linear within-person change over time.
 This measure gives us a sense of the importance of intrapersonal change over the course of the panel relative to the differences that exist between people assuming they never changed.
 
-We observe similar patterns across the datasets that have produced conflicting interpretations.
-In all datasets we see similar rates of individual-level change over time across a broad range of measures of personal culture.
-But quantifying these processes also reveals that intrapersonal change plays a substantially smaller role in explaining the variance in peoples’ personal culture than between-person differences in their baselines.
-Of course, whether this amount of intrapersonal change is “meaningful” ultimately remains a theoretical question reflected in prior expectations concerning the question, population, and time.
+We observe similar patterns across the datasets that yielded previously conflicting interpretations.
+In all datasets we find similar rates of personal change over time across a broad range of measures of personal culture.
+But quantifying these processes also reveals that intrapersonal change plays a substantially smaller role in explaining the variance in peoples’ personal culture than between-person differences at their baselines.
+Whether the found contribution of intrapersonal change is large enough to be “meaningful” ultimately remains a theoretical question reflected in prior expectations about particular cultural item, population, and duration.
 Nevertheless, with important exceptions, the results suggest that intrapersonal change (in some cases over more than 10 years) does not seem to play a substantial role in explaining a broad range of personal cultural differences we observe between adults.
@@
 These results help clarify and, hopefully, transcend previous debates.
-What is more, the measure we propose offers a tool with wider utility than merely discerning durable change that can help advance theoretical debates.
+What is more, the measure we propose offers a tool with wider utility than merely discerning durable change that can help advance theoretical perspectives.
 To exemplify this, we turn to the role of education for cultural differences.
 Specifically, we use our measure to better understand what mechanisms might underlie differences observed between people with and those without college degrees.
-Focusing on eight questions that tab general political dispositions, our measure enables us to compare the relative importance of intrapersonal change for both groups in ways that the tournament of models cannot.
+Focusing on eight questions that tab general political dispositions, our measure enables us to compare the relative importance of intrapersonal change for both groups.
 We find less evidence of durable change among those with college degree.
-This may imply that the college experience predominantly serves as a formative period that solidifies political dispositions, rather than fostering an openness [ADAPTABILITY?] to new information and influences later in the life.
-[MISSING AN ENDING SENTENCE NOW.]
+This may imply that the college experience predominantly serves as a formative period that solidifies political dispositions, rather than fosters an openness and adaptability to new information and influences later in life.
+This offers a novel, empirical basis to theorize about the role of college in shaping personal culture.
+In the same way, we believe that our approach has the potential to advance the theoretical debate on personal culture beyond what a tournament of models approach can do.
+[TRIED TO GET A CATCHY END SENTENCE IN.]
 

### 2023-09-24 manuscript.Rmd
included Fig 3 again; moved endnotes to bottom (still need to tweak that); tweaks on intro and beginning of discussion.

@@
 
-[^1]: For more information on these data sources, see [@data_gss; @data_hilda; @data_bhps; @data_ukhls; @data_soep; @data_shp; @data_psid].
-
 \begin{center}
@@
 
-[^midpoint]: We use the scale midpoint, rather than the actual difference at baseline, to better capture the differences between people independent of their age trajectories.
-
-[^explained]: This does not mean that other factors might not explain additional variance if they were included in the model.
-For example, peoples' responses about religious beliefs and participation might systematically vary by the day of the week or month in which they are surveyed.
-This variation is moved into the "random" or non-systematic component of the model for us.
-
 We can think of this number as akin to an Intraclass Correlation Coefficient focused only on the systematic components of the model while ignoring residual variance that might be attributable to measurement error or non-persisting change [see also the variance decomposition of period and cohort by @vaisey2016].
@@
 <about here>
+\end{center}
+
 Figure 1: Proportions of explained variance in items of personal culture; based on predictions at participants' wave mid-point.
-\end{center}
 
@@
 Figure 2 plots the results from an ordinary least squares regression of proportion of systematic variance attributable to intrapersonal change as a function of question, panel, and sample features.
-These models also include fixed effects for panels (not shown), so coefficients reflect the association between a particular sample or question characteristic within a panel.[^paneleffects]
-
-[^paneleffects]: KK NOTE: Can we summarize these panel fixed effects at all?
+These models also include fixed effects for panels (not shown), so coefficients reflect the association between a particular sample or question characteristic within a panel.[KK NOTE: CAN WE SUMMARIZE THESE PANEL FIXED EFFECTS AT ALL?
+AE: WE COULD SHOW THAT THEY DON'T CAPTURE TOPIC EFFECTS BEYOND OUR CLASSIFICATION IF WE INCLUDE THOSE / THAT THEY POSSIBLY CAPTURE UNMEASURED COUNTRY/CONTEXT DIFFERENCES OR DIFFERENCES IN SURVEY METHODOLOGY?]
 
@@
 <about here>
+\end{center}
 Figure 2: Proportion of explained variance attributable to intrapersonal change; based on predictions at participants' wave mid-point; supressed intercept model; survey indicators not shown.
-\end{center}
 
@@
 
-[^wave]: A small number of respondents report different highest degrees at wave 1, wave 2, and wave 3.
-Some of this is due to measurement error, and some of it is due to a small number of people obtaining a higher degree during the four years of the panel.
-Estimating the panel with highest at wave 1 or highest degree reported across the panel produces functionally identical results.
-
-[^missing]: We estimate the models presented in [earlier figure] using the full sample, but only use respondents who have information on their highest degree to generate the results presented in figures [these figures].
-In total, ### respondents (%%% percent) reported no degree information.
-
 Figure 3 plots the distribution of differences between people with less than a bachelor's degree and people with at least a bachelor's degree at wave 1 of the panel.
@@
 <about here>
-Figure 3: Proportions of intrapersonal change of explained variance by college degree; based on predictions at participants' wave mid-point.
 \end{center}
+Figure 3: Difference between share of intrapersonal change and share of inter-personal differences in expalined variance; based on predictions at participants' wave mid-point
 
-There is a clear pattern in [figure above]: for more than 80 percent of GSS items explored here, intrapersonal change is a larger component of systematic variance for people without a college degree than for people with a college degree (they have negative values in the figure).
+There is a clear pattern in Figure 4: for more than 80 percent of GSS items explored here, intrapersonal change is a larger component of systematic variance for people without a college degree than for people with a college degree (they have negative values in the figure).
 While most of these differences are small in absolute terms (less than 2 percentage points), several are greater than 5 percentage points.
@@
 conservative) on seven-point scales; four questions about the government's role in improving the condition of the poor, paying people's medical bills, giving special treatment to blacks, and doing things that private businesses could do, measured on five-point scales; a question about whether the government should do more to reduce income differences, measured on a seven-point scale; and one question about whether blacks should be given preferences in hiring, measured on a five-point scale.
-Estimates for the proportion of systematic variance attributable to intrapersonal change on these eight questions, for both education groups, are presented in Figure XXX.
+Estimates for the proportion of systematic variance attributable to intrapersonal change on these eight questions, for both education groups, are presented in Figure 4.
 
-On all eight questions presented in Figure XXX, intrapersonal change explains less systematic variance for people with a college degree than for people without a college degree.
+\begin{center}
+<about here>
+\end{center}
+Figure 4: Proportions of intrapersonal change of explained variance by college degree; based on predictions at participants' wave mid-point.
+
+On all eight questions presented in Figure 4, intrapersonal change explains less systematic variance for people with a college degree than for people without a college degree.
 There are also large differences across questions on the amount of systematic variance attributable to intrapersonal change for both groups.
@@
 Instead of falling back on a "tournament of models" approach [@lersch2023: p.
-228] to call victory for one or the other possible answer to this question, this paper seeks to refocus the debate to the theoretically more productive question of _how_ initial differences between and durable changes within people collectively explain the observed variation in personal culture.
+228] to call victory for one or the other possible answer to this question, it seeks to refocus the debate to the theoretically more productive issue of _how_ initial differences between and durable changes within people collectively explain the variation in personal culture that is observed.
 To do so, it proposes a new measure that quantifies the amount of systematic variance that is attributable to either interpersonal differences at baseline or intrapersonal change.
@@
 Applying this measure to all datasets previously discussed in this debate revealed a consistent pattern.
-Nearly all questions on personal culture show evidence of intrapersonal change over time, with some items showing noteably high intrapersonal change compared to interpersonal differences.
+Nearly all questions on personal culture show evidence of durable, intrapersonal change over time, with some items showing noteably high intrapersonal change compared to interpersonal differences.
 For questions about factors that affect one’s life satisfaction or views on government benefits, it seems plausible that differences in adult experiences predominantly account for the observed differences between people.
@@
 
-More important than adjudicating past debates, however, the approach presented in this paper offers offers the possibility to pose new theoretical questions regarding when change or stability in personal culture predominates and its significance in explaining cultural variances within a population.
-To examplify this potential, we detailed how the proportion in cultural differences that intrapersonal change can explain not only varies substantially by survey item but also by individual characteristics such as their education.
+More important than adjudicating past debates, however, the approach presented in this paper unlocks new opportunities to ask when change or stability in personal culture predominates and about their role in explaining cultural variation within a population.
+To exemplify this, we detailed how the proportion in cultural differences that intrapersonal change can explain not only varies substantially by survey item but also by individual characteristics such as their education.
 Intrapersonal change plays almost no part in explaining sentiments regarding the government's economic role but is more defining for ideological and partisan identities.
-Intriguingly, it is more important in explaining cultural variation among people without a college compared to college graduates in the U.S..
-While age and other factors related to college completion might confound this pattern, it suggests that college may solidify personal culture to an extent that renders the effect of other adult experiences on political dispositions for graduates less influential relative to individuals without such a college experience.
+Intriguingly, it is more important in explaining cultural variation among non-college graduates compared to college graduates in the U.S..
+While age and other factors related to college completion might confound this pattern, this suggests that experiencing college may solidify personal culture to an extent that renders the effect of other adult experiences on political dispositions for graduates less influential relative to individuals without a college experience.
 
@@
 
+# Endnotes
+
+[^1]: For more information on these data sources, see [@data_gss; @data_hilda; @data_bhps; @data_ukhls; @data_soep; @data_shp; @data_psid].
+
+[^midpoint]: We use the scale midpoint, rather than the actual difference at baseline, to better capture the differences between people independent of their age trajectories.
+
+[^explained]: This does not mean that other factors might not explain additional variance if they were included in the model.
+For example, peoples' responses about religious beliefs and participation might systematically vary by the day of the week or month in which they are surveyed.
+This variation is moved into the "random" or non-systematic component of the model for us.
+
+[^wave]: A small number of respondents report different highest degrees at wave 1, wave 2, and wave 3.
+Some of this is due to measurement error, and some of it is due to a small number of people obtaining a higher degree during the four years of the panel.
+Estimating the panel with highest at wave 1 or highest degree reported across the panel produces functionally identical results.
+
+[^missing]: We estimate the models presented in Figures 1 and 2 using the full sample, but only use respondents who have information on their highest degree to generate the results presented in Figures 3 and 4.
+In total, ### respondents (%%% percent) reported no degree information.
+
 # References
@@
 \newpage
-```{r fig3, out.width = "450px", fig.cap="Proportions of intrapersonal change of explained variance by college degree; based on predictions at participants' wave mid-point."}
+```{r fig3, out.width = "450px", fig.cap="Difference between share of intrapersonal change and share of inter-personal differences in expalined variance; based on predictions at participants' wave mid-point."}
+knitr::include_graphics("fig/FIG_college1.png")
+```
+
+\newpage
+```{r fig4, out.width = "450px", fig.cap="Proportions of intrapersonal change of explained variance by college degree; based on predictions at participants' wave mid-point."}
 knitr::include_graphics("fig/FIG_college2.png")

### 2023-09-24 manuscript.Rmd
Implications Cultural Sociology I

@@
 This model suggests people change their personal culture as they adapt and make new meanings when encountering new social environments, discourses, and information [@gross2009; @swidler2001].
-This model underlies, among others, theories of cultural diffusion [@christakis2010], attitude alignment [@dellaposta2020], and polarization [@bail2018] It is also implicit in most studies that ask whether specific experiences, changes in social roles, or political events, affect personal culture [@gelman2021; @visser2004; @slothuus2021].
+This model underlies, among others, theories of cultural diffusion [@christakis2010], attitude alignment [@dellaposta2020], and polarization [@bail2018].
+It is also implicit in most studies that ask whether specific experiences, changes in social roles, or political events, affect personal culture [@gelman2021; @visser2004; @slothuus2021].
 
@@
 
-## A Case Study: College Education
+## A Case Study — The Role of College for Change in Political Dispositions Among Adults
 
@@
 
-So far, the debate in cultural sociology has centered around the presence and significance of cultural change in adults.
-Yet, arguing solely over its existence sets the wrong theoretical focus.
-In any adult population, some degree of personal cultural change is inevitable.
-Likewise, no viable theoretical perspective would expect just one process --stable differences or internal change between people-- to characterize the differences in personal cultural observed between them.
-Our results underscore this perspective, showing a combination of interpersonal differences and intrapersonal change across all items.
-[DELETED SENTECE, MISSPLACED]
+So far, sociologists interested in understanding cultural differences have largely asked about the presence and significance of cultural change in adults [e.g., @kiley2020; @vaisey2021; @lersch2023].
+Yet doing so, might have unnecessarily limited and even set the wrong theoretical focus.
+In any adult population, some degree of cultural change in people is inevitable.
+Likewise, no viable theoretical perspective would expect just one process -- stable differences or internal change between people -- to underlie the differences in personal culture observed between them.
+Our results reinforce this point, showing the relevance of both interpersonal differences and intrapersonal change in understanding differences in personal culture across all items analyzed.
+[TRIED TO BETTER TUCK IT IN, BUT COULD LET GO OF THE LAST SENTENCE]
 
-A more reasonable theoretical debate pushed by previous work is the relative importance of these two processes for explaining cultural variation.
-It is difficult to draw a single conclusion on this question across a wide range of questions measured on different scales and over different durations of time, but the overall picture, with important caveats, suggests that intrapersonal change during adulthood is not the dominant explanation for cultural differences among adults for most items we measure in surveys.
-There simply is not enough evidence of intrapersonal change on these items relative to the amount of difference we observe across people.
+Recent findings suggest that it is theoretically more insightful to ask about the relative importance of these two processes in understanding cultural variation.
+Drawing a unified conclusion from data on various cultural forms, measured on different scales and across different time frames is challenging.
+However, the general pattern suggests that, for most survey items, intrapersonal change in adulthood is not the primary reason for cultural differences among people.
 
-To be clear, none of our results should be interpreted as suggesting that adults do not change or are incapable of changing.
-On all questions, some members of the sample change their personal culture at least a little bit.
-Our results also should not be taken to imply that intrapersonal change is not important or consequential.
-Small numbers of changers can have important consequences in the aggregate.
-A shift from 49 percent supporting gay marriage to 51 percent supporting it can have significant political ramifications, even as most respondents remain fixed in their preferences.
-And our results should not detract from work trying to estimate the mechanisms through which personal cultural differences --however small-- emerge among adults.
-Just because change is small does not mean the mechanisms that produce it are not important to understand.
+To clarify, our findings do not imply that adults remain static or that changes are inconsequential.
+Even minor shifts in personal culture, such as a 2% change in support for gay marriage, can have significant ramifications.
+While the majority may remain consistent in their views, understanding the underlying mechanisms of even such minor shifts remains a crucial task for the sociology of culture [COULD SHORTEN TO "EVEN SUCH MINOR CULTURAL SHIFTS REMAIN CRUCIAL" BUT WANTED TO APPEAL TO THE SCHOLARS IN THE FIELD...].
 
-Our results should be read principally as suggesting that if researchers want to explain why people differ in their personal culture, it is imperative to look at the conditions in which they were raised and the experiences that happened before they reached adulthood.
-To the extent that we want to explain what makes adults different, focusing attention on peoples’ transition between social roles, their selection into and out of organizational environments, and changes in their social networks, as much sociological work does, appears to explain a much smaller proportion of their differences than the formative experiences that happen early in life.
+Our findings primarily suggest that understanding variations in personal culture requires scholars to examine the conditions and experiences of early life.
+While transitions between social roles, the experience of organizational environments [CAN WE SAY SOCIAL INSTITUTIONS?], and changes in social networks are often emphasized in sociological research, these factors seem to account for a smaller proportion of differences among adults than the formative experiences early in life.
 
-In this way, our findings echo a range of recent causal inference work that suggests selection effects, rather than treatment effects, seem to explain the majority of differences in personal culture between people in different social roles or positions [@campbell2016; @wodtke2018].
-While some people clearly change over time as they move into new social roles or into new social environments, this amount of change does not appear to be sufficient to explain what are often substantial differences between people in different roles.
-Our results suggest that, to the extent we observe differences in personal culture between people in different social roles, such as parenthood, having children, education, or working in positions of authority [@longest2013], or between people in different occupations [@weeden2005], our prior expectation should be that this difference is selection, rather than treatment.
+This IMPLICATION/CONCLUSION aligns with a range of recent causal inference work suggesting that selection effects, rather than treatment effects, predominantly account for personal cultural differences among individuals in varied social roles and positions [@campbell2016; @wodtke2018].
+While some individuals clearly change as they transition into new roles or environments, this magnitude of this change seems to be insufficient to explain the pronounced differences among people in diverse roles.
+When observing differences in personal culture across social roles, such as parenthood, education, or professional authority [@longest2013], or across occupations [@weeden2005], our initial presumption should lean towards selection as their primary cause, rather than treatment.
 

### 2023-09-24 manuscript.Rmd
Implications Cultural Sociology II

@@
 
-Our results do not speak to exactly why intrapersonal change lacks significant explanatory power in understanding culture.
-First, it could be the case that adults are exposed to opportunities for change at the same rate, but become less willing to change.
-Or it could be that opportunities for durable change or the social conditions that facilitate durable change are just extremely uncommon.
-Changes in social structures, the life course, or cultural expectations could make intrapersonal change a much more consequential explanation for cultural differences.
+Our analysis does not provide an answer as to why intrapersonal change seems to have limited impact on understanding cultural differences.
+The situations that promote durable change in personal culture might simply be rare.
+Alternatively, it is possible that adults do encounter opportunities, necessities, and incentives for change, but their ability or willingness to change decreases.
+All the more it is important to research when and how social situations can have encourage durable change in adults.
 
-Furthermore, the results we present related to education and political views suggest that the relative importance of different theoretical processes might differ by group.
-In line with life course perspectives, the results presented here suggest that importance of different experiences for cultural change are potentially contingent on previous experiences.
-For example, to the extent that we want to explain why people have different personal culture, the answer is likely different for people who have a college degree and those who do not.
-Mid-life experiences appear to have more significant implications for the personal culture for people without a college degree than for people with a college degree.
+In this regard, our results regarding education and political views suggest that the significance of theoretical processes that can lead to such change can vary by group.
+Aligning with life course theories, it seems the importance [IMPACT ...
+ON ?
+IF IT'S NOT TOO CAUSAL] of experiences for cultural change may be contingent on other, prior experiences.
+For example, factors that shape personal cultural differences may likely differ for college and non-college graduates; the latter might be more profoundly influenced by mid-life experiences than the former in this regard.
+In trying to understand differences in personal culture, sociologists  should therefore pay more attention to the heterogeneous effects that different factors including social events, encounters, and situations can have.
+[NOT HAPPY WITH THIS LAST SENTENCE, BUT NEED SOME "MESSAGE".
+IF WE ARE TOO LONG, WE SHOULD LET GO OR EVEN COMPRESS THE LAST TWO PARAGRAPHS INTO ONE MUCH SHORTER ONE.]
 

### 2023-09-24 manuscript.Rmd
integrated aggregate change into implications for cultural sociology.

@@
 
-## Implications for Social Change
-
-Because intrapersonal change in personal culture does not appear to be a driving force in cultural differences among adults, our results suggest that processes of social and cultural change at the aggregate level are likely driven more by cohort replacement, one group of people with certain cultural dispositions exiting the population and a different set of people with a potentially different set of cultural dispositions entering the population, than by contemporaneous social conditions [@vaisey2016; @underwood2022].
-This is likely true for both large-scale change in societies over time as well as changes within organizations, political parties, occupations, and other social aggregates.
-
-However, because our results suggest that the role of formative experiences compared to contemporaneous environments differs by a person's level of education, changes in educational attainment over time, as well as changes in the social structuring of the life course might be shifting the relative importance of these two factors over time.
-If we observe further differences across socio-demographic groups like education, social class, and race/ethnicity in the relative importance of these processes, this would argue for a greater incorporation of the interaction of demographic processes with processes of individual cultural change into our explanations of aggregate cultural change.
+Our findings also have implications for understanding cultural change at the aggregate level.
+Given that intrapersonal change does not appear as the primary driver for cultural differences among adults, our results suggest that cultural change at the aggregate level is more likely driven by processes of cohort replacement than by contemporaneous social conditions influencing a given cohort's cultural dispositions [@vaisey2016; @underwood2022].
+This likely holds true for cultural change at the macro societal level as well as the micro level such as within organizations, political parties, and professions.
+At the same time our findings point to the fact that the importance of formative experiences versus contemporaneous social influence for cultural change hinges on individuals' level of education.
+The relative importance of these two factors might thus shift as education becomes more or less prominent at different points in the life course.
+Should scholars find similar differences across other social categorizations such as class or race, it would suggest to more strongly incorporate demographic processes into the analysis of change at the aggregate level.
 

### 2023-09-25 manuscript.Rmd
Implications for Survey Measurement/Research

@@
 
-This paper set out intervene in the debate on whether people change their personal culture -- their attitudes, beliefs, values, practices, and dispositions -- as they move through life.
+This paper set out intervene in the debate on whether people change their personal culture -- their attitudes, beliefs, values, practices, and dispositions -- as they move through their adult life.
 Instead of falling back on a "tournament of models" approach [@lersch2023: p.
-228] to call victory for one or the other possible answer to this question, it seeks to refocus the debate to the theoretically more productive issue of _how_ initial differences between and durable changes within people collectively explain the variation in personal culture that is observed.
-To do so, it proposes a new measure that quantifies the amount of systematic variance that is attributable to either interpersonal differences at baseline or intrapersonal change.
+228] to call victory for one or the other possible answer to this question, it seeks to refocus the debate to the theoretically more productive question of _how_ initial differences between and durable changes within people collectively explain differences in the personal culture they exhibit.
+To enable exploring this and related questions, it proposes a new approach that quantifies the amount of observed systematic variance that is attributable to either interpersonal differences at baseline or intrapersonal change.
 
-Applying this measure to all datasets previously discussed in this debate revealed a consistent pattern.
-Nearly all questions on personal culture show evidence of durable, intrapersonal change over time, with some items showing noteably high intrapersonal change compared to interpersonal differences.
+Applying this measure to all panel datasets previously discussed in this debate revealed a consistent pattern.
+Nearly all questions on personal culture show evidence of durable, intrapersonal change over time, with some showing noteably high intrapersonal change compared to interpersonal differences.
 For questions about factors that affect one’s life satisfaction or views on government benefits, it seems plausible that differences in adult experiences predominantly account for the observed differences between people.
 
-Intrapersonal change is often substantially less pronounced than interpersonal differences, accounting for less than 10 percent on average across all surveys.
+Intrapersonal change is often substantially less pronounced than interpersonal differences, accounting for less than 10 percent on average across all datasets.
 On numerous questions, like those on civil liberties, abortion, generalized trust, and civic duty, the systematic variance explained by intrapersonal change is essentially zero.
-For these questions, it seems there is not enough cultural change during adulthood to warrant attributing differences in personal culture mainly to experiences and social transitions; instead, the primary source of the observed differences appears to stem from early-life experiences.
+For these questions, it seems there is not enough cultural change during adulthood to warrant attributing differences in personal culture mainly to experiences and social transitions; instead, the primary source of the observed differences appears to stem from early-life experiences [OR "EXPERIENCES PREDATING ADULT LIFE"].
 
-More important than adjudicating past debates, however, the approach presented in this paper unlocks new opportunities to ask when change or stability in personal culture predominates and about their role in explaining cultural variation within a population.
-To exemplify this, we detailed how the proportion in cultural differences that intrapersonal change can explain not only varies substantially by survey item but also by individual characteristics such as their education.
+More than adjudicating past positions in this debate, the approach presented in this paper enables scholars to ask when change or stability in personal culture predominates within a given population and examine their role in explaining the cultural variation within it.
+To exemplify this, we detailed how the proportion in cultural differences that intrapersonal change can explain not only varies substantially by survey item but also by individuals' characteristics such as their education.
 Intrapersonal change plays almost no part in explaining sentiments regarding the government's economic role but is more defining for ideological and partisan identities.
 Intriguingly, it is more important in explaining cultural variation among non-college graduates compared to college graduates in the U.S..
-While age and other factors related to college completion might confound this pattern, this suggests that experiencing college may solidify personal culture to an extent that renders the effect of other adult experiences on political dispositions for graduates less influential relative to individuals without a college experience.
+While age and other factors related to college completion might confound this pattern, this suggests that experiencing college may solidify personal culture to an extent that renders the effect of other adult experiences on political dispositions less influential for college graduates relative to non-college graduates.
 
@@
 So far, sociologists interested in understanding cultural differences have largely asked about the presence and significance of cultural change in adults [e.g., @kiley2020; @vaisey2021; @lersch2023].
-Yet doing so, might have unnecessarily limited and even set the wrong theoretical focus.
-In any adult population, some degree of cultural change in people is inevitable.
-Likewise, no viable theoretical perspective would expect just one process -- stable differences or internal change between people -- to underlie the differences in personal culture observed between them.
-Our results reinforce this point, showing the relevance of both interpersonal differences and intrapersonal change in understanding differences in personal culture across all items analyzed.
+Yet doing so, has unnecessarily limited and might even set the wrong focus.
+In any population, some degree of cultural change in adults is inevitable.
+Likewise, no viable theoretical perspective would expect just one process -- stable differences or internal change -- to shape their personal culture.
+Our results reinforce this point, showing the relevance of both interpersonal differences and intrapersonal change in understanding the differences in personal culture among adults across all items analyzed.
 [TRIED TO BETTER TUCK IT IN, BUT COULD LET GO OF THE LAST SENTENCE]
 
-Recent findings suggest that it is theoretically more insightful to ask about the relative importance of these two processes in understanding cultural variation.
-Drawing a unified conclusion from data on various cultural forms, measured on different scales and across different time frames is challenging.
-However, the general pattern suggests that, for most survey items, intrapersonal change in adulthood is not the primary reason for cultural differences among people.
+This and other recent findings [CITES] suggest that it is theoretically more productive to ask about the relative importance of these two processes in understanding cultural variation.
+Drawing a unified conclusion from survey items measuring various cultural forms on different scales and across different time frames is challenging.
+However, the general pattern suggests that, for most items, intrapersonal change in adulthood is not the primary reason for cultural differences among people.
 
-To clarify, our findings do not imply that adults remain static or that changes are inconsequential.
+To clarify, our findings do not imply that adults remain static or that their changes are inconsequential.
 Even minor shifts in personal culture, such as a 2% change in support for gay marriage, can have significant ramifications.
-While the majority may remain consistent in their views, understanding the underlying mechanisms of even such minor shifts remains a crucial task for the sociology of culture [COULD SHORTEN TO "EVEN SUCH MINOR CULTURAL SHIFTS REMAIN CRUCIAL" BUT WANTED TO APPEAL TO THE SCHOLARS IN THE FIELD...].
+While the majority may remain consistent in their views, understanding the underlying mechanisms of even such minor shifts remains a crucial task for the sociology of culture.
 
 Our findings primarily suggest that understanding variations in personal culture requires scholars to examine the conditions and experiences of early life.
-While transitions between social roles, the experience of organizational environments [CAN WE SAY SOCIAL INSTITUTIONS?], and changes in social networks are often emphasized in sociological research, these factors seem to account for a smaller proportion of differences among adults than the formative experiences early in life.
+While transitions between social roles, changes in social networks, and the experience of organizational environments are often emphasized in sociological research, these factors seem to account for a smaller proportion of differences among adults than the formative experiences early in life.
 
-This IMPLICATION/CONCLUSION aligns with a range of recent causal inference work suggesting that selection effects, rather than treatment effects, predominantly account for personal cultural differences among individuals in varied social roles and positions [@campbell2016; @wodtke2018].
-While some individuals clearly change as they transition into new roles or environments, this magnitude of this change seems to be insufficient to explain the pronounced differences among people in diverse roles.
-When observing differences in personal culture across social roles, such as parenthood, education, or professional authority [@longest2013], or across occupations [@weeden2005], our initial presumption should lean towards selection as their primary cause, rather than treatment.
+This suggestion aligns with a range of recent causal inference work suggesting ["ADVOCATING" TO AVOID REPETITION] that selection effects, rather than treatment effects, predominantly account for personal cultural differences among individuals in varied social roles and positions [@campbell2016; @wodtke2018].
+While some individuals clearly change as they transition into new roles or environments, this change seems, on average, to be insufficient in magnitude and duration to possibly explain the pronounced differences among people in diverse roles.
+[I'VE TRIED TO FIND A WAY TO GET BOTH IMPACT AND DURATION IN THERE; COULDN'T FIND A BETTER SOLUTION THOUGH--IDEAS?
+ALSO WE NEED TO BE CAREFUL WITH THIS SENTENCE, IT SEEMS WE ARE OVERSTATING HERE; LET'S FIND A WAY TO RELATIVIZE THE CLAIM HERE?].
+When observing differences in personal culture across social roles, such as parenthood, education, or professional authority [@longest2013], or across occupations [@weeden2005], scholar are well advised to expect selection as their primary cause, rather than treatment [ALTERNATIVE: "OUR INITIAL PRESUMPTION SHOULD LEAN TOWARDS..."].
 
@@
 Alternatively, it is possible that adults do encounter opportunities, necessities, and incentives for change, but their ability or willingness to change decreases.
-All the more it is important to research when and how social situations can have encourage durable change in adults.
+All the more it is important to research when and how social situations can provoke durable change in adults.
 
-In this regard, our results regarding education and political views suggest that the significance of theoretical processes that can lead to such change can vary by group.
-Aligning with life course theories, it seems the importance [IMPACT ...
-ON ?
-IF IT'S NOT TOO CAUSAL] of experiences for cultural change may be contingent on other, prior experiences.
-For example, factors that shape personal cultural differences may likely differ for college and non-college graduates; the latter might be more profoundly influenced by mid-life experiences than the former in this regard.
-In trying to understand differences in personal culture, sociologists  should therefore pay more attention to the heterogeneous effects that different factors including social events, encounters, and situations can have.
-[NOT HAPPY WITH THIS LAST SENTENCE, BUT NEED SOME "MESSAGE".
-IF WE ARE TOO LONG, WE SHOULD LET GO OR EVEN COMPRESS THE LAST TWO PARAGRAPHS INTO ONE MUCH SHORTER ONE.]
+Related to this, our results regarding education and political views suggest that the importance of processes that lead to such change can vary by group.
+Aligning with life course theories, it seems the significance of experiences for cultural change may be contingent on other, prior experiences.
+For example, factors that shape cultural dispositions may likely differ for college and non-college graduates; the latter might be more profoundly influenced by mid-life experiences than the former in this regard.
+In trying to understand differences in personal culture, sociologists  should therefore pay more attention to the heterogeneous effects that various factors including social events, encounters, and situations can have.
+[IF WE ARE TOO LONG, CONSIDER COMPRESSING THE LAST TWO PARAGRAPHS INTO ONE OR EVEN DELETING ENTIRELY.]
 
 Our findings also have implications for understanding cultural change at the aggregate level.
-Given that intrapersonal change does not appear as the primary driver for cultural differences among adults, our results suggest that cultural change at the aggregate level is more likely driven by processes of cohort replacement than by contemporaneous social conditions influencing a given cohort's cultural dispositions [@vaisey2016; @underwood2022].
-This likely holds true for cultural change at the macro societal level as well as the micro level such as within organizations, political parties, and professions.
-At the same time our findings point to the fact that the importance of formative experiences versus contemporaneous social influence for cultural change hinges on individuals' level of education.
-The relative importance of these two factors might thus shift as education becomes more or less prominent at different points in the life course.
-Should scholars find similar differences across other social categorizations such as class or race, it would suggest to more strongly incorporate demographic processes into the analysis of change at the aggregate level.
+Given that intrapersonal change does not appear as the primary driver for cultural differences among adults, our results suggest that cultural change at the aggregate level is more likely driven by processes of cohort replacement than the by contemporaneous social conditions changing people's cultural dispositions [@vaisey2016; @underwood2022].
+[COULD EVEN JUST SAY "THAN THE INFLUENCE OF CONTEMPORANEOUS SOCIAL CONDITIONS" BUT FELT A BIT TOO SHORT.] This likely holds true for cultural change at the macro societal level as well as the micro level such as within organizations, political parties, and professions.
+Simultaneously, our findings indicate that the impact of formative experiences, as opposed to contemporaneous social factors, on cultural change is contingent upon an individuals' level of education.
+The weight of these two factors might thus change as the prominence of education shifts in the life course.
+If scholars find similar differences across other social categorizations like class or race, it would advocate for a more thorough integration of demographic processes into the analysis of cultural change at the aggregate level.
 
@@
 
-A major takeaway from the meta-analysis is that greater response resolution –-the number of response options survey respondents are given-– is associated with a greater share of variance being explained by within-individual change.
-Again, it is possible that this reflects different kinds of issues being measured with different kinds of scales.
-But it could also suggest that within-person change, when it happens, is relatively small –-moving from “agree” to “strongly agree,” rather than “strongly agree” to “strongly disagree,” for example-– and therefore easier to detect with greater resolution.
+Our findings underscore the value of extended panel surveys to advance theories of culture.
+On average, non-systematic fluctuations in responses account for more than four times as much variance as intrapersonal change, measured as linear change.
+Differentiating between the two is therefore crucial for understanding cultural differences and cannot be done with cross-sectional data.
+[MAYBE THE COMPARISON BETWEEN NON-SYSTEMATIC VARIATION AND SYSTEMATIC (LINEAR+BASELINE DIFFERENCES) WOULD EVEN BE BETTER TO MAKE THAT POINT?].
+Instead we should extend panels beyond two waves to gain more leverage to understand when and in what form personal culture changes.
 
-Our results continue to highlight the value of long-running panel surveys for adjudicating different theories of culture that cannot be evaluated with cross-sectional data alone.
-Across our results, residual fluctuations in responses account for more than four times as much variance, on average, as linear persistent change, meaning that most change in responses is not systematic.
-Under these circumstances, extending panels beyond two waves provides significantly more leverage to understand how and why personal culture changes.
+Further, we align with others [CITES?] advocating for more studies -- including panel studies -- focusing on the socialization of youth.
+Our analysis points to the majority of variance among adults rooting in differences encountered before turning 18.
+By empirically constraining their analysis to experiences after that point, social scientists may inadvertently concentrate on topics and questions that, while important, might not be able to help explain major cultural differences in adult populations.
 
-Additionally, we echo previous calls for more studies –-including panel studies-– of the socialization processes of young people.
-Our results suggest that most of the variance between adults emerges before people turn 18 or relatively soon after.
-In limiting our empirical perspective to experiences that happen after that point, social scientists potentially focus on questions that, while important, lack the capacity to explain major cultural differences.
+Finally, a key insight from our meta-analysis is that an increase in response resolution -- the range of response options survey respondents are given -- correlates with an increased share of systematic variance explained by intrapersonal change.
+This may reflect specific issues being assessed by survey items using [COULD DELETE "SURVEY ITEMS USING"] different scales.
+Alternatively, it might imply that when intrapersonal changes occurs, they are often subtle with answers moving from “agree” to “strongly agree,” rather than from “strongly agree” to “strongly disagree.” Such subtle changes require finer response options to be detected.
 

### 2023-09-25 manuscript.Rmd
Limitations

@@
 
-## Implications for Survey Measurement
+## Implications for Survey Research
 
@@
 
-The approach outlined here generates estimates of how much variance in a population is due to each of two broad processes: intrapersonal change and interpersonal differences at baseline.
-It does not quantify the proportion of people who “change.” When we observe intrapersonal change in a question, we cannot be sure whether this is driven by many people making small changes or a few people making large changes.
-Adjudicating this question will require different models that distinguish different kinds of change.
+The approach outlined here generates estimates of how much of the observed variance in a population can be attributed to one of two theoretical processes: intrapersonal change and interpersonal differences at baseline.
+It does not quantify the proportion of people who “change,” nor can we be sure that the amount of interpersonal change we detect is driven by many people making small changes or a few people making large changes.
 
-Our results and interpretations also hinge on how we have defined change and the assumptions built into the Life Course Adaption Model.
-This model treats cultural trajectories as varying linear slopes for each respondent, assuming that change is a linear function of time.
-This assumption likely does not hold true in practice.
-Change over time likely takes non-linear and discontinuous forms.
-People might jump from one “stable” disposition to another, for example, or experience a discontinuous "turning point" that upsets a stable trajectory.
-This assumption is in part a limitation of the data, as most questions are only observed for three to six waves with a large number of missing cases.
-Panels with more waves might allow for more flexible models of change that do not require this assumption of linearity.
+Our results and interpretations also hinge on how we have defined change.
+Similarly to the LCAM, our approach treats cultural trajectories as varying linear slopes for each respondent, thus assuming that change is a linear function of time.
+There is no doubt, this assumption severely simplifies reality in which change likely also takes non-linear and discontinuous forms.
+People might jump from one “stable” disposition to another or experience a "turning point" in their lives that upsets an otherwise stable trajectory.
+In parts, our assumption of linearity is a limitation of the data, as most questions are only observed for three to six waves.
+Panels with more waves might allow to loosen this assumption to test alternative, more flexible models of change.
 
-Similarly, in assuming a linear trajectory for change, our approach assumes that durable change is unidirectional.
-To the extent that people make durable change as a result of a social transition and then return to a previous position later in life as a result of a second transition (marriage and divorce, for example), this is captured as residual within-person variance, rather than one of our two broad theoretical processes.
-This is a potentially reasonable assumption on short panels --change that does not last two years is hard to classify as durable-- but it might not be true in longer panels.
-Other researchers with more flexible definitions of "durable change" might potentially identify more durable change even in the same population.
+Similarly, our approach assumes that durable change is unidirectional.
+This is a sensible assumption on short panels where classifying change that lasts less than two years as durable seems unreasonable.
+Practically, it means that people making durable changes in their cultural dispositions to then return to a previous state later in life are classified as neither accounted for by the two theoretical processes.
+Longer panels and more flexible definitions of "change" would allow us to account for such trajectories and thus potentially also identify more durable change even in the same population.
 
-Finally, while we examine an incredible broad array of measures of personal culture, from assessments about the relative importance of different features when buying a new car and policy preferences to religious beliefs and core values, we are limited by the kinds of questions that are asked in panel surveys and the contexts in which these surveys were administered (Australia, Germany, Switzerland, the United Kingdom and the United States).
-This means our findings tend to reflect questions about general (national) politics, gender roles, immigration and race relations, and general well being.
-Our data does not cover some dimensions of culture that might be relevant for other aspects of sociological exploration (artistic tastes, leisure activities, and time use, for example).
+Finally, we had chosen to examine the broadest array of measures of personal culture available to us, ranging from religious beliefs and core values, to policy preferences, and even to the importance of different features when buying a new car across five countries (Australia, Germany, Switzerland, the United Kingdom and the United States).
+Nevertheless, our findings remain limited to the kinds of questions that are asked in panel surveys and in the contexts they were administered, reflecting issues of general (national) politics, gender roles, immigration and race relations, and general well being.
+Although we have no reason to believe results to be different, our findings do not directly speak to other dimensions of culture such as artistic tastes, leisure activities, and time use.
 

### 2023-09-25 manuscript.Rmd
greek letters, suggestions, and rants you can ignore

@@
 228].
-Accordingly, researchers have formalized models of responses to cultural items over time as demonstrating either “change” or “no change” to then compare their empirical fit, which essentially asks whether we ever observe within-person change in personal culture [WE COULD ALREADY HERE SWTICH TO "INTRAPERSONAL CHANGE" OR INTRODUCE THE TERM AS AN ALTERNATIVE TO WITHIN-PERSON CHANGE].
+Accordingly, researchers have formalized models of responses to cultural items over time as demonstrating either “change” or “no change,” which essentially asks whether we *ever* observe intrapersonal change in personal culture.
 
@@
 In this paper, we try to intervene in this debate by moving beyond the “tournament of models” approach toward a quantification of the processes that lead to differences in personal culture.
-Drawing on the seven panel datasets analyzed in the debate, we use established methods to quantify the amount of variance in responses that is due to between-people differences at a single point during the survey versus the amount of variance in responses that is due to linear within-person change over time.
+Drawing on the seven panel datasets analyzed in the debate, we use established methods to quantify the amount of variance in responses that is due to between-people differences at a single point during the survey versus the amount of variance in responses that is due to linear intraindividual change over time.
 This measure gives us a sense of the importance of intrapersonal change over the course of the panel relative to the differences that exist between people assuming they never changed.
@@
 But quantifying these processes also reveals that intrapersonal change plays a substantially smaller role in explaining the variance in peoples’ personal culture than between-person differences at their baselines.
-Whether the found contribution of intrapersonal change is large enough to be “meaningful” ultimately remains a theoretical question reflected in prior expectations about particular cultural item, population, and duration.
+Whether the contribution of intrapersonal change is large enough to be “meaningful” ultimately remains a theoretical question reflected in prior expectations about particular cultural item, population, and duration.
 Nevertheless, with important exceptions, the results suggest that intrapersonal change (in some cases over more than 10 years) does not seem to play a substantial role in explaining a broad range of personal cultural differences we observe between adults.
@@
 
-Drawing on these findings, Lersch challenged the SDM and AUM as a “needless dichotomy” [-@lersch2023].
-Instead he proposed the “Life Course Adaption Model” (LCAM).
+Drawing on these findings, Lersch [-@lersch2023] challenged the SDM and AUM as a “needless dichotomy,” proposing the “Life Course Adaption Model” (LCAM) as an alternative.
 This model draws on the life course perspective to model personal culture as a linear trajectory over the course of a panel that varies for each respondent.
-Lersch rectifies two shortcomings of the AUM and SDM.
+In doing this, Lersch rectified two shortcomings of the AUM and SDM.
 First, the AUM, as described by Kiley and Vaisey, posits that changes follow a Markovian process where personal culture responses at time $t$ rely solely on responses at time $t-1$.
-Howevre, earlier life experiences can (directly and indirectly) mold personal culture when transitioning to new social roles or into new environments, even if their initial impact is delayed.
+However, earlier life experiences can (directly and indirectly) mold personal culture when transitioning to new social roles or into new environments, even if their initial impact is delayed.
 For example, childhood events might influence views on family structures later when individuals form their own families.
-The LCAM considers therefore considers influences from earlier than just $t-1$ on responses at time $t$.
+Therefore, the LCAM considers influences from earlier than just $t-1$ on responses at time $t$.
 
 Second, Kiley and Vaisey's analysis of the AUM and SDM is based on three-wave panel data over four years, which might not be extensive enough to properly adjudicate [COMPARE?
-WE ARE USING ADJUDICATING A LOT] the two models.
-Lersch, however, evaluates the LCAM against the AUM and SDM using panel data spanning a wider duration (from 3 to 36 years) and more waves (3 to 18), offering a better change to observe durable change.
+WE ARE USING ADJUDICATING A LOT -- TK NOTE: I think adjudication is *the* thing they tried to do?] the two models.
+Lersch, however, evaluates the LCAM against the AUM and SDM using panel data spanning a wider duration (from 3 to 36 years) and more waves (3 to 18), offering a better chance to observe durable change.
 
-In Lersch's tournament of models, the LCAM gains broad [EXTENSIVE?] support from data across five countries.
+In Lersch's tournament of models, the LCAM gains broad [EXTENSIVE?
+-- TK Note: I wouldn't say 70% is extensive, tbh] support from data across five countries.
 Out of 428 questions, the LCAM was favored in 297 cases while the SDM was preferred in only 112.
-The rest did not yield inconclusive results, and notably, no questions favoured the AUM.
+The rest did not yield conclusive results, and notably, no questions favoured the AUM.
 He concludes that “new experiences over the life course [...] can persistently move individuals’ personal culture in novel directions” [@lersch2023: p.
@@
 However, while Lersch [-@lersch2023] found a preference for the LCAM in many cases, durable change only accounted for a minor proportion of variance explained.
-On average, individuals changed [CAN WE SAY SHIFTED TO INDICATE THE LINEAR DRIFT] about .07 standard deviations over 10 years, echoing Kiley's and Vaisey’s view that such changes are typically modest.
+On average, individuals changed [CAN WE SAY SHIFTED TO INDICATE THE LINEAR DRIFT -- TK Note: "shift" does not sound very layman to me, and I think we should stick to layman lingo as much as possible] about .07 standard deviations over 10 years, echoing Kiley's and Vaisey’s view that such changes are typically modest.
 Furthermore, a quarter of the items in Lersch's study favoured the SDM, suggesting that despite broader assumptions about change and more extensive data, many questions reveal no evidence of durable change.
 Likewise, in studies claiming evidence of durable change in personal culture, the magnitude is often minimal.
-For instance, Brocic and Miles [-@brocic2021] note that obtaining graduate degrees in humanities, arts, and social sciences only shift [NUDGES?] peoples’ moral relativism by about .2 standard deviations compared to non-degree holders [DELETED THE LAST SENTENCE PART HERE; FELT UNNECESSARY].
+For instance, Brocic and Miles [-@brocic2021] note that obtaining graduate degrees in humanities, arts, and social sciences only shift [NUDGES?
+-- TK Note: I have a *strong* dislike for the term nudge] peoples’ moral relativism by about .2 standard deviations compared to non-degree holders [DELETED THE LAST SENTENCE PART HERE; FELT UNNECESSARY].
 Moreover, studies of aggregate change demonstrate that even on questions where cohort effects prevail, there is always evidence that some individuals change over time [@vaisey2016].
@@
 Consequently, this metric also reflects variation in individuals' social roles or statuses at the start of the panel that were important in shaping their dispositions.
+[TK Note: I feel that this paragraph puts too much emphasis on pre-panel processes.
+It might be early imprinting or broad life-course, but why emphasize it this much?]
 
@@
 This non-durable change may arise for various reasons.
-Some might lack clear cultural dispositions, leading them to construct an opinion or attitude during the survey, with varying thoughts and considerations dominating their thinking in each interview [@tourangeau2000; @zaller1992; @feldman1992].
+Some might lack clear cultural dispositions, leading them to construct an opinion or attitude during the survey [TK Note: I am not sure saying "some" is true here.
+This might well be true for everyone, and I think Zaller's point is that it *is* true for everyone.
+And "lacking cultural dispositions" might just mean lacking them in the form of survey questions.
+So I think this part obscures some theoretical questions here.
+I would prefer the previous versions.], with varying thoughts and considerations dominating their thinking in each interview [@tourangeau2000; @zaller1992; @feldman1992].
 Even though the full set of considerations they could draw on remains stable over time, this could yield to such variance in responses.
@@
 Yet, earlier studies found many questions (between 25 and 35 percent) lacking evidence of durable change.
-Aligning wiht this, we expect considerable variation in the variance explained by intrapersonal change, with some questions indicating minimal variance attributable to such change.
+Aligning with this, we expect considerable variation in the variance explained by intrapersonal change, with some questions indicating minimal variance attributable to such change.
 [I KEPT THE "VARIANCE EXPLAINED" PHRASING THOUGH THE "VARIANCE ATTRIBUTABLE" MIGHT BE LESS OF A HARD CLAIM WITHOUT LOSING MUCH?]
@@
 PREMISE?] is fairly uncontroversial, regardless of the theoretical process assumed to be most relevant for a particular question.
-Panel surveys do not encompase individuals’ entire lifetimes.
+Panel surveys do not encompass individuals’ entire lifetimes.
 Thus, even if intrapersonal change explains a large part of cultural differences, the change detected in a panel covering 20 years probably will not fully account for the entire variance of interpersonal differences.
@@
 Third, we expect the variance explained by each process –interpersonal differences and intrapersonal change– to depend on specific attributes of the question and panel.
-Noteably, the longer we observe respondents, the higher the response resolution for a given question, and the larger the sample, the greater the variance explained by intrapersonal change should be.
+Notably, the longer we observe respondents, the higher the response resolution for a given question, and the larger the sample, the greater the variance explained by intrapersonal change should be.
 Given that the LCAM (supported by the various underlying theoretical perspectives) posits that social experiences drive changes in personal culture and that these experiences correlate with time, observing people for longer durations [or TIME SPANS?] increases the chance that they undergo experiences that change their personal culture.
@@
 
-We use data from 7 nationally representative panel surveys from Australia, Germany, Great Britain, Switzerland, and the United States (summarized in Table 1), combining all the data files used in previous work [@kiley2020; @lersch2023].[^1] These studies cover a long period of time (the range of surveys spans from 1968 to 2021), with 609 personal culture items that capture attitudes, beliefs, values, self-assessments, self-descriptions, and behaviors [@alwin2007].
+We use data from 7 nationally representative panel surveys from Australia, Germany, Great Britain, Switzerland, and the United States (summarized in Table 1), combining all the data files used in previous work [@kiley2020; @lersch2023].[^data] These studies cover a long period of time (the range of surveys spans from 1968 to 2021), with 609 personal culture items that capture attitudes, beliefs, values, self-assessments, self-descriptions, and behaviors [@alwin2007].
 We restricted the sample such that individuals between the ages of 18 and 79 are included without further elimination, and in all surveys, we used all possible cases for which respondents provided responses.
@@
 
+## Analysis Steps
+
 As a first step in our analysis, we fit LCAMs to each of the 609 measures of personal culture outlined above.
 We then use these models to derive two measures of systematic variance that correspond to the two collections of theoretical processes outlined above: interpersonal differences at baseline and intrapersonal change over time.
-First, to measure the importance of interpersonal differences, we calculate the sum of the squared residuals derived from measuring each respondent at the midpoint of their age trajectory.[^midpoint] We then divide this value by the total sum of squares and subtract it from from 1, producing a value we denote as $\tau$, or the proportion of variance explained by interpersonal differences.
+First, to measure the importance of interpersonal differences, we calculate the sum of the squared residuals derived from measuring each respondent at the midpoint of their age trajectory.[^midpoint] We then divide this value by the total sum of squares and subtract it from from 1, producing a value we denote as V(D), or the proportion of variance attributable to interpersonal differences.
 This number is equivalent to an $R^2$ value generated by using random intercepts for all respondents.
@@
 We divide this number by the total sum of squares and subtract it from 1.
-We then subtract $\tau$ from this value.
-This gives us a value we denote as $\delta$, the additional proportion of variance explained by intrapersonal change.
+We then subtract $V(D)$ from this value.
+This gives us a value we denote as $V(C)$, the additional proportion of variance explained by intrapersonal change [TK Note: I used Vaisey & Lizardo conventions of using a function term for $V$ here].
 
-These two numbers reflect the total systematic or "explained" variance of the model that can be attributed to either differences across people or linear change over the course of the panel.[^explained] Our principal outcome of interest is the proportion of systematic variance accounted for by intrapersonal change, which we calculate as
+These two numbers reflect the total systematic variance of the model that can be attributed to either differences across people or linear change over the course of the panel.[^explained] Our principal outcome of interest is the proportion of systematic variance accounted for by intrapersonal change, which we calculate as
 
 $$
-\frac{\delta}{\tau + \delta}
+\tau = \frac{V(C)}{V(C) + V(D)}
 $$
 
-We can think of this number as akin to an Intraclass Correlation Coefficient focused only on the systematic components of the model while ignoring residual variance that might be attributable to measurement error or non-persisting change [see also the variance decomposition of period and cohort by @vaisey2016].
+[TK Note: \phi taken in Kiley & Vaisey, \delta taken in Vaisey & Lizardo.
+I think using a Greek letter without strong associations (something like alpha, beta, or mu or whatever) is not good.
+So I settled on \tau.
+Feel free to change it.]
+
+We can think of this number, $\tau$ as akin to an Intraclass Correlation Coefficient focused only on the systematic components of the model while ignoring residual variance that might be attributable to measurement error or non-persisting change [see also the variance decomposition of period and cohort by @vaisey2016].
 The higher the number is, the more variance in responses is attributable to intrapersonal change during the panel.
 The lower the number, the less intrapersonal change seems to play a large role in explaining cultural differences during the panel.
+Moving forward, we use $\tau$ as a placeholder for *systematic variance attributable to intrapersonal change*.
 
@@
 
-As a second step in our analysis, we use ordinary least squares regression to model the proportion of variance explained by intrapersonal change as a function of features of the question and panel.
+As a second step in our analysis, we use ordinary least squares regression to model $\tau$ as a function of features of the question and panel.
 This allows us to explore how these features affect the prevalence of intrapersonal change.
@@
 
+In the third step... [TK Note: I guess we need to say something here about the college stuff?].
+
 # Results
 
+[TK Note: I think this paragraph should go].
 Our results proceed in three parts.
@@
 
-Figure 1 plots the proportion of systematic variance attributable to interpersonal differences (“between”) and intrapersonal change (“within”) for each of the 609 questions analyzed, plotted separately by panel.
+Figure 1 plots the proportion of systematic variance attributable to interpersonal differences and intrapersonal change for each of the 609 questions analyzed, plotted separately by panel.
 To broadly summarize the results, all panels show a similar range of the proportion of systematic variance attributable to intrapersonal change.
-Across panels, the median proportion of systematic variance attributable to intrapersonal change is 0.094 (mean of 0.098), with an interquartile range of 0.053 to 0.139.
-All panels include questions that show essentially no systematic variance attributable to intrapersonal change and all panels include questions where more than 20 percent of systematic variance is attributable to intrapersonal change.
+Across panels, the median proportion of $\tau$ is 0.094 (mean of 0.098), with an interquartile range of 0.053 to 0.139.
+All panels include questions that show essentially zero $\tau$ values and all panels include questions with more than 20 percent of $\tau$ values [TK Note: $\tau$ values?
+Sound kinda weird I guess.]
 
@@
 
-To the extent that there are differences across the panels, the PSID has the highest mean (.120) and median (.135) proportion of systematic variance attributable to intrapersonal change.
+To the extent that there are differences across the panels, the PSID has the highest $\tau$ values with mean .120 and median .135.
 While we cannot disentangle features of the sample from features of the questions asked to each sample, the specific samples for many PSID questions have lower average ages than those from other panels.
 To the extent that younger respondents might be more likely to make durable changes of opinion, this higher explanatory power of intrapersonal change might reflect the distinct age profile of respondents in this sample.
-At the other end, the GSS has the lowest mean (.073) and median (.069) proportion of variance attributable to intrapersonal change.
+At the other end, the GSS has the lowest range of $\tau$ with mean .073 and median .069.
 This potentially reflects the fact that the GSS observes people for a shorter duration, on average, than the other panels.
-If, consistent with life course adaption theories, people are more likely to make significant cultural changes the longer we observe them, then duration likely affects the amount of variance attributable to intrapersonal change (a point we explore more below).
+If, consistent with life course adaption theories, people are more likely to make significant cultural changes the longer we observe them, then duration likely affects the range of $\tau$ (a point we explore more below).
 However, the GSS results are still highly consistent with results from the other panels.
@@
 These questions tend to ask about external referents (e.g., confidence in specific government leaders or political parties), life satisfaction, or current financial position.
-At the other end, questions about religious identification, views on gender roles, and support for civil liberties tend to have very low proportion of variance attributable to intrapersonal change.
+At the other end, questions about religious identification, views on gender roles, and support for civil liberties tend to have very low proportion of variance attributable to intrapersonal change [TK Note: I sometimes keep the long format because I think it reads better in some places than others.
+Not sure if this is a principled decision, though.]
 
 In contrast to the tournament of models approach, quantifying change this way allows us to explore variation in the relative importance of intrapersonal change across questions that all both show evidence of change.
-For example, Kiley and Vaisey [-@kiley2020] found that confidence in the press and confidence in religious leaders were both characterized by active updating, our results show that intrapersonal change is much more important for explaining variance in confidence in the press (0.164) than confidence in religion (0.049).
+For example, Kiley and Vaisey [-@kiley2020] found that confidence in the press and confidence in religious leaders were both characterized by active updating.
+Our results show that intrapersonal change is much more important for explaining variance in confidence in the press (0.164) than confidence in religion (0.049).
 
@@
 
-Figure 2 plots the results from an ordinary least squares regression of proportion of systematic variance attributable to intrapersonal change as a function of question, panel, and sample features.
+Figure 2 plots the results from an ordinary least squares regression of $\tau$ as a function of question, panel, and sample features.
 These models also include fixed effects for panels (not shown), so coefficients reflect the association between a particular sample or question characteristic within a panel.[KK NOTE: CAN WE SUMMARIZE THESE PANEL FIXED EFFECTS AT ALL?
@@
 
-Figure 2 shows that the more response options respondents are given and the larger the sample, the larger the proportion of systematic variance is attributable to intrapersonal change.
+Figure 2 shows that the more response options respondents are given and the larger the sample, the larger the values of $\tau$.
 We interpret these coefficients as suggesting greater resolution on a question makes it easier to detect and model change.
-The earlier a question was asked, even net of how long it was been asked, the lower the proportion of systematic variance is attributable to intrapersonal change.
+The earlier a question was asked, even net of how long it was been asked, the lower the $\tau$.
 
-The most notable result in the meta analysis is the negative coefficient attached to the span of years covered by the question.
-The longer a question is observed, the less systematic variance is attributable to intrapersonal change.
+The most notable result in the meta analysis is the negative coefficient attached to the duration of years covered by the question.
+The longer a question is observed, the less is the value of $\tau$.
 Theories of personal cultural change that link changes in personal culture to social experiences, including the LCAM, suggest that the longer we observe respondents, the more likely people are to undergo potentially transformative experiences and therefore the more variance would be explained by intrapersonal change.
-The fact that we find a negative coefficient attached to this variable is a challenge to that assumption.
+The fact that we find a negative coefficient attached to this variable can be a challenge to that assumption [TK Note: since there is a complication of this, I changed is to can be, but feel free to ignore it.]
 
-As a further test of this coefficient, we compared the proportion of systematic variance attributable to intrapersonal change when using the full duration of a panel compared to when when we dropped the final wave and therefore reduced the total duration of observation for the question.
-If the coefficient reflects a true negative effect of duration on systematic variance attributable to intrapersonal change, we should see that same effect within questions.
-Results from this analysis are presented in Appendix B and contradict the coefficient from the regression model.
-The longer we observe the same question, the higher proportion of systematic variance is attributable to intrapersonal change.
+As a further test of this coefficient, we compared the $\tau$ values when using the full duration of a panel compared to when we dropped the final wave and therefore reduced the total duration of observation for the question.
+If the coefficient reflects a true negative effect of duration on $\tau$, we should see that same effect within questions.
+Results from this analysis are presented in Appendix B and contradict the coefficient from the regression model: we found that the longer we observe the same question, the higher the $\tau$ is.
 We interpret this combination of findings as suggesting that the kinds of questions asked for a longer time period tend to demonstrate less intrapersonal change than questions asked over shorter periods, rather than a true function of time.
@@
 ## A Case Study — The Role of College for Change in Political Dispositions Among Adults
+
+[TK Note: I feel like "dispositions" is a loaded term and a theoretical choice.
+Don't get me wrong, I like imputing misplaced concreteness as much as anyone (hehe), but can't we say something like political beliefs or political culture?]
 
@@
 
-Figure 3 plots the distribution of differences between people with less than a bachelor's degree and people with at least a bachelor's degree at wave 1 of the panel.
+Figure 3 plots the distribution of differences in $\tau$ between people with less than a bachelor's degree and people with at least a bachelor's degree at wave 1 of the panel.
 Values greater than 0 would indicate that intrapersonal change is more important in explaining the responses of college graduates than those of people without a college degree, while values less than 0 would indicate that intrapersonal change is less important for explaining variation among people with a college degree than those without a college degree.
@@
 While most of these differences are small in absolute terms (less than 2 percentage points), several are greater than 5 percentage points.
-Given the distribution observed in Figure [the earlier one] showing that the systematic variance attributable to intrapersonal change averages [average across all questions], a 5 percentage point difference between groups is quite substantial.
+Given the distribution observed in Figure 1 showing that the systematic variance attributable to intrapersonal change averages around 0.09, a 5 percentage point difference between groups is quite substantial.
 
@@
 conservative) on seven-point scales; four questions about the government's role in improving the condition of the poor, paying people's medical bills, giving special treatment to blacks, and doing things that private businesses could do, measured on five-point scales; a question about whether the government should do more to reduce income differences, measured on a seven-point scale; and one question about whether blacks should be given preferences in hiring, measured on a five-point scale.
-Estimates for the proportion of systematic variance attributable to intrapersonal change on these eight questions, for both education groups, are presented in Figure 4.
+Estimates for the $\tau$ on these eight questions, for both education groups, are presented in Figure 4.
 
@@
 On all eight questions presented in Figure 4, intrapersonal change explains less systematic variance for people with a college degree than for people without a college degree.
-There are also large differences across questions on the amount of systematic variance attributable to intrapersonal change for both groups.
+There are also large differences across questions on the amount of $\tau$ for both groups.
 For example, on the question of whether the government should try to solve more problems or leave those problems to be solved by private businesses ("government do more or less"), less than one percent of the systematic variance is attributable to intrapersonal change for both groups.
@@
 
-In contrast, partisan identification and political ideology both show a larger proportion of systematic variance attributable to intrapersonal change, with quite large differences between people with a college degree and those without.
+In contrast, partisan identification and political ideology both show a larger $\tau$, with quite large differences between people with a college degree and those without.
 Compared to the other questions explored here, change as adults (and, potentially, their experiences as adults) plays a much larger role in explaining why people report the partisan identification and political ideology they do.
 And this is particularly true for respondents who do not have a college degree.
-The amount of systematic variance in ideological identification that is attributable to intrapersonal change among non-college educated respondents is almost four times that of college-educated respondents.
+The $\tau$ value for ideological identification among non-college educated respondents is almost four times that of college-educated respondents.
 
-We note here that this meaningful difference in the proportion of systematic variance attributable to intrapersonal change across education groups and across questions would not be detectable using previous methods.
+We note here that this meaningful difference in $\tau$ across education groups and across questions would not be detectable using previous methods.
 For partisan identification and political ideology, both college-educated respondents and people with less than college degree would likely favor the AUM or LCAM over the SDM because they show evidence of some members of the population making intrapersonal change.
@@
 
-Applying this measure to all panel datasets previously discussed in this debate revealed a consistent pattern.
-Nearly all questions on personal culture show evidence of durable, intrapersonal change over time, with some showing noteably high intrapersonal change compared to interpersonal differences.
+Applying our proposed measure, $\tau$, to all panel datasets previously discussed in this debate revealed a consistent pattern.
+Nearly all questions on personal culture show evidence of durable, intrapersonal change over time, with some showing notably high intrapersonal change compared to interpersonal differences.
 For questions about factors that affect one’s life satisfaction or views on government benefits, it seems plausible that differences in adult experiences predominantly account for the observed differences between people.
 
+[TK Note: from now on I didn't change anything to $\tau$, but not sure if this is a correct way?].
+
 Intrapersonal change is often substantially less pronounced than interpersonal differences, accounting for less than 10 percent on average across all datasets.
-On numerous questions, like those on civil liberties, abortion, generalized trust, and civic duty, the systematic variance explained by intrapersonal change is essentially zero.
+On numerous questions, like those on civil liberties, abortion, generalized trust, and civic duty, the systematic variance attributable to intrapersonal change is essentially zero.
 For these questions, it seems there is not enough cultural change during adulthood to warrant attributing differences in personal culture mainly to experiences and social transitions; instead, the primary source of the observed differences appears to stem from early-life experiences [OR "EXPERIENCES PREDATING ADULT LIFE"].
@@
 
-[^1]: For more information on these data sources, see [@data_gss; @data_hilda; @data_bhps; @data_ukhls; @data_soep; @data_shp; @data_psid].
+[^data]: For more information on these data sources, see [@data_gss; @data_hilda; @data_bhps; @data_ukhls; @data_soep; @data_shp; @data_psid].
 

### 2023-09-26 manuscript.Rmd

@@
 Quantifying these processes offers new ways to address theoretical questions of when and how their relative importance shifts.
-We examplify this with the case of college experience."
+We examplify this approach by exploring how the importance of these processes differs between college graduates and the rest of the population."
 ---
@@
 
-This seemingly simple question underlies a number of central debates across sociological subfields.
-For example, pragmatist theories of action claim that changes in social environments cause people to adapt their views and make new meanings [@swidler2001; @gross2009], while Bourdieusian practice theories argue that the “past conditions of production” leave a mark on people's personal culture that lasts throughout their lives [@bourdieu1990].
-Similarly, models of social influence assume that people adapt their culture in the face of new information [@vaisey2010a], while the emphasis on cohort effects in models of aggregate social and cultural change requires them to be open to change while young but become fairly resistant to it as they age [@ryder1965].
-Finally, life course theories posit slower but still important changes over time as people advance through important transitions in their lives [@elder2016; @bardi2011].
+This seemingly simple question underlies a number of debates across sociological subfields.
+For example, pragmatist theories of action claim that changes in social environments cause people to adapt their views and make new cultural meanings [@swidler2001; @gross2009], while Bourdieusian practice theories argue that the “past conditions of production” leave a mark on people's personal culture that lasts throughout their lives [@bourdieu1990].
+Similarly, models of social influence assume that people adapt their culture in the face of new information [@vaisey2010a KK Note: I don't feel like this is a good cite for this.], while the emphasis on cohort effects in models of aggregate social and cultural change requires them to be open to change while young but become fairly resistant to it as they age [@ryder1965].
+Finally, life course theories posit slower but still important changes over time as people advance through important transitions in their lives [@elder2003; @bardi2009].
 
-Because all of the above theoretical perspectives have some empirical support, the current debate revolves around not whether they exist, but when and how they each apply and the relative contribution of each to explaining cultural differences.
+Because all of the above theoretical perspectives have some empirical support, the current debate revolves around not whether they exist, but when and how they each apply and the contribution of each to explaining cultural differences.
 That being said, scholars have struggled to reach a consensus on the relative importance of change during adulthood.
@@
 We argue that asking only *whether* people change obscures what are actually highly compatible empirical patterns.
-Being able to detect "change" in a given dataset depends not only on sample populations, item measurement, statistical methods and power, but crucially also on definitions of what counts as “change.” These methodological differences therefore limit the possibility of convergence on the theoretical debates.
+Being able to detect change in a given dataset depends not only on sample populations, item measurement, statistical methods, and power, but crucially also on definitions of what counts as “change.” These methodological decisions potentially limit the possibility of convergence on the theoretical debates.
 Moreover, asking whether there is evidence of change limits the kind of questions that can be asked to, “Do we see change in this measure?” But such inquiries are ill-suited to address what’s at the core of the theoretical debate.
@@
 In this paper, we try to intervene in this debate by moving beyond the “tournament of models” approach toward a quantification of the processes that lead to differences in personal culture.
-Drawing on the seven panel datasets analyzed in the debate, we use established methods to quantify the amount of variance in responses that is due to between-people differences at a single point during the survey versus the amount of variance in responses that is due to linear intraindividual change over time.
-This measure gives us a sense of the importance of intrapersonal change over the course of the panel relative to the differences that exist between people assuming they never changed.
+Drawing on the seven panel datasets analyzed in the debate, we use established methods to quantify the amount of variance in responses that is due to between-people differences at a single point during the survey versus the amount of variance in responses that is due to linear intrapersonal change over time.
+This measure gives us a sense of the importance of change over the duration of the panel relative to the differences that exist between people assuming they never changed.
 
 We observe similar patterns across the datasets that yielded previously conflicting interpretations.
-In all datasets we find similar rates of personal change over time across a broad range of measures of personal culture.
-But quantifying these processes also reveals that intrapersonal change plays a substantially smaller role in explaining the variance in peoples’ personal culture than between-person differences at their baselines.
-Whether the contribution of intrapersonal change is large enough to be “meaningful” ultimately remains a theoretical question reflected in prior expectations about particular cultural item, population, and duration.
-Nevertheless, with important exceptions, the results suggest that intrapersonal change (in some cases over more than 10 years) does not seem to play a substantial role in explaining a broad range of personal cultural differences we observe between adults.
+In all datasets we find a similar distribution of how much explanatory power can be attributed to intrapersonal change.
+But quantifying these processes also reveals that intrapersonal change plays a substantially smaller role in explaining the variance in peoples’ personal culture than interpersonal differences at their baselines.
+Whether the amount of change is large enough to be “meaningful” ultimately remains a theoretical question reflected in prior expectations about particular cultural item and population.
+It also crucially depends on how long people are observed in a panel.
+Nevertheless, on many measures of personal culture we do not observe enough intrapersonal change, in some cases over more than 10 years, for it to play a substantial role in explaining the differences we observe between adults, though notable exceptions exist.
 
-These results help clarify and, hopefully, transcend previous debates.
-What is more, the measure we propose offers a tool with wider utility than merely discerning durable change that can help advance theoretical perspectives.
-To exemplify this, we turn to the role of education for cultural differences.
-Specifically, we use our measure to better understand what mechanisms might underlie differences observed between people with and those without college degrees.
-Focusing on eight questions that tab general political dispositions, our measure enables us to compare the relative importance of intrapersonal change for both groups.
-We find less evidence of durable change among those with college degree.
-This may imply that the college experience predominantly serves as a formative period that solidifies political dispositions, rather than fosters an openness and adaptability to new information and influences later in life.
+These results help clarify and, hopefully, transcend previous debates about whether adults change.
+What is more, the approach we propose can be applied to a broader range of theoretical questions than approaches discerning whether durable change exists in a panel.
+To exemplify this, we explore differences in the explanatory power of intrapersonal change for people with college degrees and those without.
+We find that change has less explanatory power for people with college degrees than those without.
+This may imply that the college experience solidifies political dispositions rather than fosters an openness and adaptability to new information and influences later in life.
 This offers a novel, empirical basis to theorize about the role of college in shaping personal culture.
-In the same way, we believe that our approach has the potential to advance the theoretical debate on personal culture beyond what a tournament of models approach can do.
-[TRIED TO GET A CATCHY END SENTENCE IN.]
+We believe that this approach has the potential to advance the theoretical debate on the role of environments, social identities, and roles in facilitating personal cultural change and stability.
 
@@
 The first, what Kiley and Vaisey [-@kiley2020] call a “Settled Dispositions Model” (SDM), assumes that peoples’ personal culture is relatively fixed by the time they are adults.
-While they might make temporary changes in their declarative culture in reaction to their environments, this model assumes that people return to a settled baseline over a (relatively) short period of time.
-This model is frequently invoked in theories of cultural change that suggest people are strongly imprinted by early socialization experiences such as the “past conditions of production” a Bourdieusian practice theory [@bourdieu1990], cohort replacement theories of aggregate change [@mannheim1952; @ryder1965], and control theories in social psychology [@smith-lovin1988; @robinson2007].
+While they might make temporary changes in their declarative culture in reaction to their environments, this model assumes that people return to a settled baseline over a short period of time.
+This model underlies theories of cultural change that suggest people are imprinted by early socialization experiences such as the “past conditions of production” a Bourdieusian practice theory [@bourdieu1990], cohort replacement theories of aggregate change [@mannheim1952; @ryder1965], and control theories in social psychology [@smith-lovin1988; @robinson2007].
 
@@
 For example, adolescence and early adulthood is typically viewed as a “formative period” for personal cultural development and thus characterized by higher rates of active updating, while middle age and later life are potentially characterized more by settled dispositions [@alwin1991; @krosnick1989; @eaton2009].
-Similarly, highly salient issues, such as views around gay rights in the 2010s; issues that see substantial elite realignments, such as views around the Vietnam conflict in the 1970s [@zaller1992]; and novel issues of public opinion, such as views around vaccines during the Covid-19 pandemic, might be characterized by active updating, while established issues of low salience are characterized by stability.
+Similarly, salient issues, such as views around gay rights in the 2010s; issues that see substantial elite realignments, such as views around the Vietnam conflict in the 1970s [@zaller1992]; and novel issues of public opinion, such as views around vaccines during the Covid-19 pandemic [@scoville2022], might be characterized by active updating, while established issues of low salience are characterized by stability.
 
 Empirically comparing which of these models better fit a broad range of questions from the General Social Survey’s rotating panels, Kiley and Vaisey [-@kiley2020] found limited evidence of durable change.
-While the AUM was preferred for the majority of items, the total amount of durable change detected on these items was small.
+While the majority of items preferred the AUM, the amount of durable change detected on these items was small.
 A substantial minority of questions (39 percent) favored the SDM, meaning they showed no evidence of durable change.
-Questions with more evidence of durable change included highly salient issues like gay marriage and “public” issues like partisan identification and religious service attendance.
+Questions with more evidence of durable change included salient issues like gay marriage and questions tapping “public” statements or behaviors such as partisan identification and religious service attendance.
 There was also more evidence of durable change among early adults (people ages 18-30) than the rest of the population.
@@
 500; see also @vaisey2021].
-This lack of durable change is consistent with other findings that cohort replacement plays a larger role than period effects in explaining differences in personal culture [@vaisey2016].
+This lack of durable change is consistent with other recent findings that cohort replacement plays a larger role than period effects in explaining differences in personal culture [@vaisey2016].
 
@@
 Drawing on these findings, Lersch [-@lersch2023] challenged the SDM and AUM as a “needless dichotomy,” proposing the “Life Course Adaption Model” (LCAM) as an alternative.
-This model draws on the life course perspective to model personal culture as a linear trajectory over the course of a panel that varies for each respondent.
+This model draws on the life course perspective to model personal culture as a different linear trajectory over the duration of a panel for each respondent.
 In doing this, Lersch rectified two shortcomings of the AUM and SDM.
-First, the AUM, as described by Kiley and Vaisey, posits that changes follow a Markovian process where personal culture responses at time $t$ rely solely on responses at time $t-1$.
+First, the AUM, as described by Kiley and Vaisey, posits that changes follow a Markov process where responses at time $t$ are a function of responses at time $t-1$ but not earlier time points.
 However, earlier life experiences can (directly and indirectly) mold personal culture when transitioning to new social roles or into new environments, even if their initial impact is delayed.
@@
 
-Second, Kiley and Vaisey's analysis of the AUM and SDM is based on three-wave panel data over four years, which might not be extensive enough to properly adjudicate [COMPARE?
-WE ARE USING ADJUDICATING A LOT -- TK NOTE: I think adjudication is *the* thing they tried to do?] the two models.
-Lersch, however, evaluates the LCAM against the AUM and SDM using panel data spanning a wider duration (from 3 to 36 years) and more waves (3 to 18), offering a better chance to observe durable change.
+Second, Kiley and Vaisey's analysis of the AUM and SDM is based on three-wave panel data over four years, which might not be extensive enough to adjudicate the two models.
+Lersch evaluates the LCAM against the AUM and SDM using panel data spanning a wider duration (from 3 to 36 years) and more waves (3 to 18), offering a better chance to observe durable change.
 
-In Lersch's tournament of models, the LCAM gains broad [EXTENSIVE?
--- TK Note: I wouldn't say 70% is extensive, tbh] support from data across five countries.
-Out of 428 questions, the LCAM was favored in 297 cases while the SDM was preferred in only 112.
-The rest did not yield conclusive results, and notably, no questions favoured the AUM.
+When Lersch compared the LCAM to the AUM and SDM on data from five countries, 297 of the 428 questions he analyzed preferred the LCAM, suggesting that we observe some linear change for adults on most items.
+The SDM was preferred on 112 items, and the rest did not yield conclusive results.
+No questions favored the AUM.
 He concludes that “new experiences over the life course [...] can persistently move individuals’ personal culture in novel directions” [@lersch2023: p.
@@
 
-Despite differing interpretations of these scholars, their empirical results align closely.
-Kiley and Vaisey [-@kiley2020] discovered that most questions they investigated favored the AUM, indicating evidence of durable change among adults -- a finding largely in line with Lersch’s.
-However, while Lersch [-@lersch2023] found a preference for the LCAM in many cases, durable change only accounted for a minor proportion of variance explained.
-On average, individuals changed [CAN WE SAY SHIFTED TO INDICATE THE LINEAR DRIFT -- TK Note: "shift" does not sound very layman to me, and I think we should stick to layman lingo as much as possible] about .07 standard deviations over 10 years, echoing Kiley's and Vaisey’s view that such changes are typically modest.
-Furthermore, a quarter of the items in Lersch's study favoured the SDM, suggesting that despite broader assumptions about change and more extensive data, many questions reveal no evidence of durable change.
-Likewise, in studies claiming evidence of durable change in personal culture, the magnitude is often minimal.
-For instance, Brocic and Miles [-@brocic2021] note that obtaining graduate degrees in humanities, arts, and social sciences only shift [NUDGES?
--- TK Note: I have a *strong* dislike for the term nudge] peoples’ moral relativism by about .2 standard deviations compared to non-degree holders [DELETED THE LAST SENTENCE PART HERE; FELT UNNECESSARY].
-Moreover, studies of aggregate change demonstrate that even on questions where cohort effects prevail, there is always evidence that some individuals change over time [@vaisey2016].
+Despite differences in how these researchers interpret their findings, the empirical results are not far apart.
+Kiley and Vaisey [-@kiley2020] found that the majority of items they tested favored the AUM, meaning there was evidence of durable change among adults, a pattern consistent with Lersch’s results.
+While Lersch [-@lersch2023] found evidence of durable change on most items, this change only accounted for a small proportion of variance explained.
+On average, people changed only about .07 standard deviations over 10 years.
+This is consistent with Kiley and Vaisey’s finding that durable change is often small.
+And a quarter of the items Lersch studied still preferred the SDM, suggesting that even with different assumptions assumptions about change and more extensive data, many questions show no evidence of durable change.
+Likewise, in studies showing evidence of durable change in specific measures of personal culture, the change is often small.
+For example, Brocic and Miles [-@brocic2021] estimate that completing graduate degrees in humanities, arts, and social sciences shift peoples’ moral relativism only about .2 standard deviations on average compared to people with no college degree, and this is the largest effect they identify.
+And studies of aggregate change show that, even on items where cohort effects explain more variance than period effects, there is always evidence that some people change over time [@vaisey2016].
 In other words, despite different interpretations, the results of previous work are in many ways highly consonant.
+[I think the revisions that were done to this paragraph were confusing and mostly reverted it.]
 
@@
 Current debates often narrow down to ask whether people *ever* change, which is rarely theoretically interesting since the answer is almost always “yes.”  Over sufficient time and in a large enough sample, researchers will likely observe some evidence for durable changes in individuals.
-A lack thereof may simply due to a lack of clarity in the question asked or a low level of resolution in the possible answers, or that the question simply was not asked long enough or to enough people.
+A lack of evidence for change may simply due to a lack of clarity in the question asked or a low level of resolution in the possible answers, or that the question simply was not asked for long enough or to enough people.
 Conversely, finding evidence of change tells us little about how much change is present in a population and how meaningful that change is for social differences.
 
-A more theoretically productive approach begins with a model assuming that during an observed time period some individuals might remain stable, might change a little, or might undergo significant shifts in their personal culture.
+A more theoretically productive approach begins with a model assuming that during an observed time period people might remain stable, might change a little, or might undergo significant shifts in their personal culture.
 Lersch’s LCAM formally does this by modeling each individual as following their own trajectory over time.
-But rather than debating whether this model is superior to a model that assumes that people never change, researchers should quantify individual changes over time and compare those to the differences that exist between people at the panel's start.
-In other words, the focus should shift from asking, “do people change?” to asking, “how do change and initial differences collectively explain variation in personal culture?” [I CHANGED THAT QUESTION WORDING, PLEASE CHECK!]
+But rather than debating whether this model fits a particular data set better than a model that assumes that people never change, researchers should quantify individual changes over time and compare those to the differences that exist between people at the panel's start.
+In other words, the focus should shift from asking, “do people change?” to asking, “what are the relative contributions of change and initial differences toward explaining differences in personal culture?” [KK Note: I didn't like this change.
+It's not about collectively explaining, its about them relative to each other.]
 
-In this approach, two metrics reflect distinct theoretical processes.
-First, "early imprinting," as Lersch calls it, denotes interpersonal differences at the start of the panel, capturing the accumulated experiences of people before participating in the panel that influence their answers.
-While commonly associated with experiences during an early formative period that result in settled disposition, this could also reflect transformative experiences [-@paul2014transformative] that happen at any stage in life, as long as they predate the panel and consistently impact subsequent responses.
+In this approach, there are two metrics that each reflect a combination of relevant theoretical processes.
+The first metric, interpersonal differences at the start of the survey, reflect the accumulated experiences of people prior to entering the panel survey that shape how they answer questions, what Lersch calls "early imprinting." While commonly associated with experiences during an early formative period that result in settled dispositions, this variation could also reflect experiences that happen at any stage in life as long as they predate the panel and consistently impact subsequent responses.
 For instance, for those people entering the panel post-retirement, this “imprinting” might reflect this pivotal life transition.
@@
 [TK Note: I feel that this paragraph puts too much emphasis on pre-panel processes.
-It might be early imprinting or broad life-course, but why emphasize it this much?]
+It might be early imprinting or broad life-course, but why emphasize it this much?
+KK NOTE: Because it could be either, I don't want to treat it like it's just "imprinting."]
+[KK NOTE: You reversed what was being said here.
+You say "early imprinting" denotes intrapersonal change (process denotes number).
+We want to say the opposite -- what I had before -- number denotes process: "interpersonal differences" denotes "early imprinting." I changed it back.
+I also took out reference to "transformative" experiences because I don't think we're necessarily talking about those specifically, just any experience that changes you.
+Because we don't say a specific theoretical process ("socialization", etc.) I want to keep the language of each denoting combinations of processes.
+"Persistent change" isn't a theoretical process.
+It's a bunch of them.]
 
-The second process, termed “persistent change” or “adaption” by Lersch and "active updating" by Kiley and Vaisey, denotes durable changes in personal culture over time, captured by the amount of intrapersonal linear change during the panel.
+The second metric, the amount of intrapersonal linear change people make during the panel, captures durable changes in personal culture over time, a set of processes collectively called “persistent change” or “adaption” by Lersch and "active updating" by Kiley and Vaisey.
 Lersch attributes these changes to social triggers such as moving into a new environment or adopting new social roles.
@@
 
-A third metric, often termed “residual variance” or “fluctuation,” represents the remainder of variance in peoples’ responses over time.
-This non-durable change may arise for various reasons.
-Some might lack clear cultural dispositions, leading them to construct an opinion or attitude during the survey [TK Note: I am not sure saying "some" is true here.
-This might well be true for everyone, and I think Zaller's point is that it *is* true for everyone.
-And "lacking cultural dispositions" might just mean lacking them in the form of survey questions.
-So I think this part obscures some theoretical questions here.
-I would prefer the previous versions.], with varying thoughts and considerations dominating their thinking in each interview [@tourangeau2000; @zaller1992; @feldman1992].
-Even though the full set of considerations they could draw on remains stable over time, this could yield to such variance in responses.
+A third metric, often called “residual variance” or “fluctuation,” accounts for the remainder of variance in peoples’ responses over time.
+This non-durable change emerges for a variety of reasons.
+For example, people might not have clear personal cultural dispositions as measured by surveys.
+Instead, they might internalize a broad set of considerations and construct an opinion in the context of the survey interview, with different considerations coming to the forefront of their cognition each interview [@tourangeau2000; @zaller1992; @feldman1992].
+This variance in considerations can produce variance in responses even if peoples’ “dispositions” --the full set of considerations they could draw on-- remains stable over time.
 Additionally, this variance can also include measurement errors, such as misinterpreted questions, erroneous response selections, or responses getting coded incorrectly.
-While this third process thus might reflect potentially important aspects of personal culture [I FELT WE NEED TO PLAY IT DOWN A BIT], it is less central to the debate outlined here.
-Hence, we focus on the first two metrics.
+While this third process reflects important aspects of personal culture, it does not directly touch on the ongoing debate outlined here.
+Because of this we quantify this component but focus principally on the other two.
+[KK NOTE: I agreed with TK's note and mostly reverted this paragraph.]
 
-At a theoretical level, the coexistence of two main processes -- interpersonal differences and intrapersonal change -- is undeniable, and they are linked in many ways [THOUGHT IF intrinsically linked].
-Any intrapersonal change during one's life will inevitably manifest as interpersonal differences by the time people enter a panel survey.
-Furthermore, unless people are entirely socialized early on and never deviate from these dispositions -- an understanding challenged by evidence -- we expect observing intrapersonal change in a segment of the population when surveyed over time.
+At a theoretical level, the coexistence of the sets of processes captured in the two main metrics, interpersonal differences and intrapersonal change, is undeniable, and they are linked in many ways.
+Any intrapersonal change during one's life will likely manifest as interpersonal differences by the time people enter a panel survey.
+Furthermore, unless people are entirely socialized early on and never deviate from these dispositions, we expect observing intrapersonal change in a segment of the population when surveyed over time.
 
 Furthermore, these metrics are not necessarily useful in isolation or without context.
-Central to the theoretical debate is how the two processes they capture collectively explain cultural differences in a population.
-In other words, when examining adults across a period of time, does intrapersonal change, compared to interpersonal differences, sufficiently account for the observed cultural differences?
+What is important for the theoretical debate is the relative contribution of these two components in explaining cultural differences in a population over a specific duration of time.
+In other words, when we look at a population of adults over some time frame, is there enough intrapersonal change relative to interpersonal difference to view it as an important explanation for why people demonstrate cultural differences?
 Only through quantifying the two processes we can get closer to a true answer to this question.
 
-Yet, the true merit of the proposed quantification goes beyond settling past debates; it opens up new avenues for investigations that remained closed under previous approaches.
-For example, merely classifying questions based on whether they show durable change [DO WE NEED "or not"?
-NEVER KNOW THAT AS A NON-NATIVE] overlooks possibly important differences among those that do.
-By quantifying the contributions of interpersonal differences and intrapersonal change, researchers can gauge their relative importance in understanding [INSTEAD explaining] differences among survey items, between groups within a given population, across time, and even across societies.
+Yet the true utility of the proposed quantification goes beyond settling past debates; it allows researchers to ask questions that were difficult to explore with previous approaches.
+For example, classifying questions based on whether they show durable change overlooks possibly important differences among those that do.
+By quantifying the relative contributions of interpersonal differences and intrapersonal change, researchers can gauge their relative importance in explaining differences among survey items, between groups within a given population, across time, and across societies.
 
@@
 
-By reframing the question and adjusting the approach, we expect consistent patterns across datasets previously analysed in the debate.
-First, we anticipate evidence for some intrapersonal change in most measures of personal culture across datasets.
-Yet, earlier studies found many questions (between 25 and 35 percent) lacking evidence of durable change.
-Aligning with this, we expect considerable variation in the variance explained by intrapersonal change, with some questions indicating minimal variance attributable to such change.
-[I KEPT THE "VARIANCE EXPLAINED" PHRASING THOUGH THE "VARIANCE ATTRIBUTABLE" MIGHT BE LESS OF A HARD CLAIM WITHOUT LOSING MUCH?]
+By shifting the question and adjusting the approach, we expect consistent patterns across datasets previously analysed in the debate.
+First, because previous work found evidence of durable change on most measures of personal culture, we expect that intrapersonal change will have some explanatory power for most items.
+But because these earlier studies also found many questions (between 25 and 35 percent) lacking evidence of durable change, we expect considerable variation in the variance explained by intrapersonal change, with some questions indicating essentially no variance attributable to such change.
 
 Second, we expect the variance explained by intrapersonal to be modest compared to interpersonal differences at the start of the survey.
-This assertion [EXPECTATION?
-PREMISE?] is fairly uncontroversial, regardless of the theoretical process assumed to be most relevant for a particular question.
+This expectation should be uncontroversial, regardless of the theoretical process assumed to be most relevant for a particular question.
 Panel surveys do not encompass individuals’ entire lifetimes.
-Thus, even if intrapersonal change explains a large part of cultural differences, the change detected in a panel covering 20 years probably will not fully account for the entire variance of interpersonal differences.
+Thus, even if intrapersonal change during adulthood is the major driver of cultural differences, the change detected in a panel covering 20 years probably will not fully account for the entire variance of interpersonal differences.
 
 Our expectations so far largely echo what was observed in previous studies.
-However, theoretical perspectives differ in how much variance they attribute to intrapersonal change, enabling quantification to provide a new perspective on existing questions.
-“Formative period” and cohort theories, which emphasize early-life socialization as a strong force in shaping adult personal culture, suggest that the variance from intrapersonal change would still fall short of explaining interpersonal differences, even when extrapolated to a longer time frame [WE COULD DROP HERE A REFERENCE TO THE SPAN..., possibly beyond the panel window].
-Conversely, other theoretical perspectives emphasizing the importance of people's current social roles expect intrapersonal change could account for a large portion of the variance when extrapolated [CAN WE FIND AN ALTERNATIVE TO "EXTRAPOLATE"?].
-While projecting observed panel change beyond the covered time period may have its critics, we see it as a reasonable starting point.
-[DO WE NEED THIS LAST SENTENCE?
-IT FEELS WE ARE MAKING OUR SELVES VULNERABLE TO A CRITIQUE TO NO BENEFIT IN THE ARGUMENT.]
+However, theoretical perspectives differ in how much variance they would expect to be driven by intrapersonal change, enabling quantification to provide a new perspective on existing questions.
+“Formative period” and cohort theories, which emphasize early-life socialization as a strong force in shaping adult personal culture, suggest that the variance from intrapersonal change would still fall short of explaining interpersonal differences, even when we account for the duration of observation.
+Conversely, other theoretical perspectives emphasizing the importance of people's contemporaneous social environment for personal culture expect intrapersonal change to account for a large portion of the variance, regardless of the duration.
 
 Third, we expect the variance explained by each process –interpersonal differences and intrapersonal change– to depend on specific attributes of the question and panel.
-Notably, the longer we observe respondents, the higher the response resolution for a given question, and the larger the sample, the greater the variance explained by intrapersonal change should be.
-Given that the LCAM (supported by the various underlying theoretical perspectives) posits that social experiences drive changes in personal culture and that these experiences correlate with time, observing people for longer durations [or TIME SPANS?] increases the chance that they undergo experiences that change their personal culture.
-[WE WOULDN'T EVEN NEED THE "they undergo experiences that" IN THAT SENTENCE.]
+First, the higher the response resolution for a given question and the larger the sample, the greater the variance explained by intrapersonal change should be, as smaller changes in intrapersonal change will be easier to detect.
+Second, the LCAM and various theoretical perspectives that underlie it posit that social transitions drive changes in personal culture.
+Since the probability that any person experiences such a transition goes up the longer we observe them, we expect panel duration to be positively related to the proportion of variation explained by intrapersonal change.
 
@@
 We then subtract $V(D)$ from this value.
-This gives us a value we denote as $V(C)$, the additional proportion of variance explained by intrapersonal change [TK Note: I used Vaisey & Lizardo conventions of using a function term for $V$ here].
+This gives us a value we denote as $V(C)$, the additional proportion of variance explained by intrapersonal change.
 
@@
 
-In the third step... [TK Note: I guess we need to say something here about the college stuff?].
+[TK Note: I guess we need to say something here about the college stuff?
+KK Note: I didn't feel like that was necessary.
+It feels so distinct later on that I think it's OK to kind of "start over" there.
+But others should weigh in.].
 
 # Results
-
-[TK Note: I think this paragraph should go].
-Our results proceed in three parts.
-We first explore the overall pattern of systematic variance explained by interpersonal differences at baseline and intrapersonal change over time across the seven panels.
-Next, we analyze the association between the relative proportion of systematic variance explained by intrapersonal change as a function of question, panel, and sample features.
-Finally, we highlight a few specific comparisons to illustrate the value of quantification relative to adjudication for further exploring theoretical questions.
 
@@
 Across panels, the median proportion of $\tau$ is 0.094 (mean of 0.098), with an interquartile range of 0.053 to 0.139.
-All panels include questions that show essentially zero $\tau$ values and all panels include questions with more than 20 percent of $\tau$ values [TK Note: $\tau$ values?
-Sound kinda weird I guess.]
+All panels include questions where $\tau$ is essentially 0 and all panels include questions with $\tau$ values greater than 0.20.
 
@@
 While there are some differences between panels, these differences are small compared to the differences within panels.
-For about 6 percent of items explored here, more than 20 percent of systematic variance is attributable to intrapersonal change.
+For about 6 percent of items explored here,  $\tau$ is greater than 0.20.
 These questions tend to ask about external referents (e.g., confidence in specific government leaders or political parties), life satisfaction, or current financial position.
-At the other end, questions about religious identification, views on gender roles, and support for civil liberties tend to have very low proportion of variance attributable to intrapersonal change [TK Note: I sometimes keep the long format because I think it reads better in some places than others.
-Not sure if this is a principled decision, though.]
+At the other end, questions about religious identification, views on gender roles, and support for civil liberties tend to have very low proportion of variance attributable to intrapersonal change.
 
@@
 
-## A Case Study — The Role of College for Change in Political Dispositions Among Adults
-
-[TK Note: I feel like "dispositions" is a loaded term and a theoretical choice.
-Don't get me wrong, I like imputing misplaced concreteness as much as anyone (hehe), but can't we say something like political beliefs or political culture?]
+## College and Change in Political Culture
 
@@
 Previous work has established a positive relationship between education and attitude stability, especially on issues related to American politics.
-This stability is often attributed to education facilitating "chronic knowledge" -- a general understanding of the domain of American politics, including the positions held by major parties and political figures and how issues relate to one another at a logical or socio-logical level [@zaller1992; @alvarez2002; @boutyline2017].
+This stability is often attributed to education facilitating "chronic information" -- a general understanding of the domain of American politics, including the positions held by major parties and political figures and how issues relate to one another at a logical or socio-logical level [@zaller1992; @alvarez2002; @boutyline2017].
 Because they have more knowledge of American politics, people with a college education are better able to consistently connect the considerations in their cognition with the answer choices they are presented with in a survey.
 
-This work has tended to focus on the fact that college-educated Americans give responses that are less likely to be affected by measurement error or short-term influences [@zaller1992; @alwin2007].
+This work has tended to focus on the fact that college-educated Americans give responses that are less likely to be affected by measurement error or short-term influences than the rest of the population [@zaller1992; @alwin2007].
 However, this focus on the difference in the non-systematic (residual) component of variance across groups obscures the fact that the relative importance of different components of systematic variance might differ across these two groups as well.
 
-It could be the case that education, chronic knowledge, and attention to the political sphere makes intrapersonal change more important in explaining differences among people with higher levels of education.
-For example, these people might be more likely to make systematic changes in response to the emergence of new information, new issues, or political realignments, while those without chronic knowledge display more variance around their baselines without their baselines changing much [@zaller1992].
+It could be the case that education, chronic information, and attention to the political sphere makes intrapersonal change more important in explaining differences among people with higher levels of education.
+For example, these people might be more likely to make systematic changes in response to the emergence of new information, new issues, or political realignments, while those without chronic information display more variance around their baselines without their baselines changing much [@zaller1992].
 In contrast, it could be the case that the stability demonstrated by educated respondents reflects the fact that they have already formed durable opinions and are relatively closed off to new information.
@@
 
-Figure 3 plots the distribution of differences in $\tau$ between people with less than a bachelor's degree and people with at least a bachelor's degree at wave 1 of the panel.
+Figure 3 plots the distribution of differences in $\tau$ values between people with less than a bachelor's degree and people with at least a bachelor's degree at wave 1 of the panel for all 183 items int he GSS.
 Values greater than 0 would indicate that intrapersonal change is more important in explaining the responses of college graduates than those of people without a college degree, while values less than 0 would indicate that intrapersonal change is less important for explaining variation among people with a college degree than those without a college degree.
@@
 \end{center}
-Figure 3: Difference between share of intrapersonal change and share of inter-personal differences in expalined variance; based on predictions at participants' wave mid-point
+Figure 3: Difference between share of intrapersonal change and share of inter-personal differences in explained variance; based on predictions at participants' wave mid-point
 
-There is a clear pattern in Figure 4: for more than 80 percent of GSS items explored here, intrapersonal change is a larger component of systematic variance for people without a college degree than for people with a college degree (they have negative values in the figure).
+There is a clear pattern in Figure 3: for more than 80 percent of GSS items explored here, intrapersonal change is a larger component of systematic variance for people without a college degree than for people with a college degree (i.e., they have negative values).
 While most of these differences are small in absolute terms (less than 2 percentage points), several are greater than 5 percentage points.
@@
 conservative) on seven-point scales; four questions about the government's role in improving the condition of the poor, paying people's medical bills, giving special treatment to blacks, and doing things that private businesses could do, measured on five-point scales; a question about whether the government should do more to reduce income differences, measured on a seven-point scale; and one question about whether blacks should be given preferences in hiring, measured on a five-point scale.
-Estimates for the $\tau$ on these eight questions, for both education groups, are presented in Figure 4.
+Estimates of $\tau$ for these eight questions, for both education groups, are presented in Figure 4.
 
@@
 On all eight questions presented in Figure 4, intrapersonal change explains less systematic variance for people with a college degree than for people without a college degree.
-There are also large differences across questions on the amount of $\tau$ for both groups.
+There are also large differences in $\tau$ values across questions for both groups.
 For example, on the question of whether the government should try to solve more problems or leave those problems to be solved by private businesses ("government do more or less"), less than one percent of the systematic variance is attributable to intrapersonal change for both groups.
@@
 
-In contrast, partisan identification and political ideology both show a larger $\tau$, with quite large differences between people with a college degree and those without.
-Compared to the other questions explored here, change as adults (and, potentially, their experiences as adults) plays a much larger role in explaining why people report the partisan identification and political ideology they do.
+In contrast, partisan identification and political ideology both show larger values of $\tau$ than the other questions, as well as a large difference between people with a college degree and those without.
+Compared to the other questions explored here, change as adults plays a much larger role in explaining why people report the partisan identification and political ideology they do.
 And this is particularly true for respondents who do not have a college degree.
@@
 
-We note here that this meaningful difference in $\tau$ across education groups and across questions would not be detectable using previous methods.
+We note here that this meaningful difference in $\tau$ values across education groups and across questions would not be detectable using previous methods.
 For partisan identification and political ideology, both college-educated respondents and people with less than college degree would likely favor the AUM or LCAM over the SDM because they show evidence of some members of the population making intrapersonal change.
-In other words, using the tournament of models approach obscures the fact that the relative importance of intrapersonal change in explaining differences on these two items is very different across these two groups, that intrapersonal change explains more systematic variance for ideological identification than partisan identification for people without a college degree, or that there appears to be more durable change on questions of affirmative action than on questions of government aid to black Americans.
+In other words, using the tournament of models approach obscures the fact that the relative importance of intrapersonal change in explaining differences on these two items is different across these two groups, that intrapersonal change explains more systematic variance for ideological identification than partisan identification for people without a college degree, and that there appears to be more durable change on questions of affirmative action than on questions of government aid to black Americans.
 

### 2023-09-27 manuscript.Rmd
add updated descriptives table to repo
revisions to discussion, conclusion

@@
 In this approach, there are two metrics that each reflect a combination of relevant theoretical processes.
-The first metric, interpersonal differences at the start of the survey, reflect the accumulated experiences of people prior to entering the panel survey that shape how they answer questions, what Lersch calls "early imprinting." While commonly associated with experiences during an early formative period that result in settled dispositions, this variation could also reflect experiences that happen at any stage in life as long as they predate the panel and consistently impact subsequent responses.
+The first metric, interpersonal differences at the start of the survey, reflects the accumulated experiences of people prior to entering the panel survey that shape how they answer questions, what Lersch calls "early imprinting." While commonly associated with experiences during an early formative period that result in settled dispositions, this variation could also reflect experiences that happen at any stage in life as long as they predate the panel and consistently impact subsequent responses.
 For instance, for those people entering the panel post-retirement, this “imprinting” might reflect this pivotal life transition.
@@
 
-Second, we expect the variance explained by intrapersonal to be modest compared to interpersonal differences at the start of the survey.
+Second, we expect the variance explained by intrapersonal change to be modest compared to interpersonal differences at the start of the survey.
 This expectation should be uncontroversial, regardless of the theoretical process assumed to be most relevant for a particular question.
@@
 We restricted the sample such that individuals between the ages of 18 and 79 are included without further elimination, and in all surveys, we used all possible cases for which respondents provided responses.
-In the end, the analyses that follow rely on a cross-cultural sample with a cross-domain set of items to capture the breadth of individual personal culture.
+In the end, the analyses that follow rely on a cross-cultural sample with a cross-domain set of items to capture a broad range of individual personal culture.
 Appendix A documents the list of all variables used in the upcoming analyses.
@@
 $$
-y_{i, t} = \beta_0 + u_{0, i} + \beta_1 x_t + u_{1, i} x_{i, t} + \epsilon_{i, t}
+y_{i, t} = \beta_0 + u_{0, i} + \beta_{1} x_t + u_{1, i} x_{i, t} + \epsilon_{i, t}
 $$
+$$
+u_{0,i} \sim \mathcal{N}(0,\sigma_{\mu_0}^2)
+$$
+
+$$
+u_{1,i} \sim \mathcal{N}(0,\sigma_{u_1}^2)
+
+$$
+[KKNOTE: Is that right? I forget how to do that exactly.]
 
@@
 
-As a first step in our analysis, we fit LCAMs to each of the 609 measures of personal culture outlined above.
+As a first step in our analysis, we fit a LCAM to each of the 609 measures of personal culture outlined above.
 We then use these models to derive two measures of systematic variance that correspond to the two collections of theoretical processes outlined above: interpersonal differences at baseline and intrapersonal change over time.
@@
 We can think of this number, $\tau$ as akin to an Intraclass Correlation Coefficient focused only on the systematic components of the model while ignoring residual variance that might be attributable to measurement error or non-persisting change [see also the variance decomposition of period and cohort by @vaisey2016].
-The higher the number is, the more variance in responses is attributable to intrapersonal change during the panel.
-The lower the number, the less intrapersonal change seems to play a large role in explaining cultural differences during the panel.
+The higher this number is, the more systematic variance in responses is attributable to intrapersonal change during the panel.
+The lower the number, the less intrapersonal change seems to play a large role in explaining systematic cultural differences during the panel.
 Moving forward, we use $\tau$ as a placeholder for *systematic variance attributable to intrapersonal change*.
@@
 Because this quotient summarizes variance and therefore has no natural referent, it is hard to say at which values we observe "a lot" of intrapersonal change.
-But summarizing questions in this way allows us to compare the relative importance of our two broad processes across a range of questions that might have very different levels of non-systematic variance, giving us a baseline sense of the relative prevalence of these two processes for each question (something not achievable under the tournament of models approach), and allowing us to compare which kinds of questions show more or less intrapersonal change relative to interpersonal differences.
+But summarizing questions in this way allows us to compare the relative importance of our two broad processes across a range of questions that might have different levels of non-systematic variance, giving us a sense of the relative prevalence of these two processes for each question (something not achievable under the tournament of models approach), and allowing us to compare which kinds of questions show more or less intrapersonal change relative to interpersonal differences.
 While the tournament of models approach allowed researchers to say whether questions in general showed evidence of updating, this approach allows us to make more specific claims about the prevalence of these theoretical processes for particular questions relative to each other.
@@
 It feels so distinct later on that I think it's OK to kind of "start over" there.
+I'm worried that people will forget the plan by the time they get there, and we'll have to reiterate everything anyway.
 But others should weigh in.].
@@
 Across panels, the median proportion of $\tau$ is 0.094 (mean of 0.098), with an interquartile range of 0.053 to 0.139.
-All panels include questions where $\tau$ is essentially 0 and all panels include questions with $\tau$ values greater than 0.20.
+All panels include questions where $\tau$ is essentially 0, and all panels include questions with $\tau$ values greater than 0.20.
+The maximum $\tau$ value is 0.26.
 
@@
 If, consistent with life course adaption theories, people are more likely to make significant cultural changes the longer we observe them, then duration likely affects the range of $\tau$ (a point we explore more below).
-However, the GSS results are still highly consistent with results from the other panels.
+However, the GSS results are still consistent with results from the other panels.
 
@@
 
-Appendix A presents the estimated proportion of variance attributable to interpersonal differences and intrapersonal change, as well as the proportion of residual variance, for each question analyzed.
+Appendix A presents the estimated proportion of variance attributable to interpersonal differences and intrapersonal change, the estimated values of $\tau$, and the proportion of residual variance for each question analyzed.
 In general, interpersonal differences are almost always the largest component of the total variance and tend to explain between 55 and 70 percent of total variance, while intrapersonal change is always the smallest, typically explaining between 3 and 8 percent of total variance.
-Residual variance tends to account for between 22 and 37 percent of variance (though on several questions residual variance is greater than 50 percent).
+Residual variance tends to account for between 22 and 37 percent of variance, though on several questions residual variance is greater than 50 percent.
 
@@
 Figure 2 plots the results from an ordinary least squares regression of $\tau$ as a function of question, panel, and sample features.
-These models also include fixed effects for panels (not shown), so coefficients reflect the association between a particular sample or question characteristic within a panel.[KK NOTE: CAN WE SUMMARIZE THESE PANEL FIXED EFFECTS AT ALL?
+These models also include fixed effects for panels, so coefficients reflect the association between a particular sample or question characteristic within a panel.[KK NOTE: CAN WE SUMMARIZE THESE PANEL FIXED EFFECTS AT ALL?
 AE: WE COULD SHOW THAT THEY DON'T CAPTURE TOPIC EFFECTS BEYOND OUR CLASSIFICATION IF WE INCLUDE THOSE / THAT THEY POSSIBLY CAPTURE UNMEASURED COUNTRY/CONTEXT DIFFERENCES OR DIFFERENCES IN SURVEY METHODOLOGY?]
@@
 This reinforces the explanation for why the GSS displays the lowest average proportion of systematic variance attributable to intrapersonal change: its shorter duration.
-[KK NOTE: Does this inform any other panel differences?
-Does the PSID, BHPS or UKHLS have longer durations on average?]
+
+While we did not generate expectations for how question content relates to $\tau$, some of the associations between question structure, panel duration, and panels themselves might be driven by differences in the topics addressed by each panel.
+To address this, we coded each question as falling into one of nine different content categories and included these as indicator variables in the meta analysis.
+Appendix C presents the estimates for these content indicators.
+They show some associations with $\tau$, with questions about subjective SES; social life, social cohesion, and trust; environment and climate; and health and morale showing larger $\tau$ values than questions about religion and spirituality; politics, government and the economy; and gender and family.
+[KKNOTE: I FEEL LIKE WE WANTED TO SAY SOMETHING ELSE HERE BUT I CANT REMEMBER WHAT]
 
@@
 
-While these broad results help speak to previous debates, they mostly re-frame previous findings without extending the theoretical debate.
-To better illustrate the value of our approach for addressing different theoretical questions, we turn to a specific empirical example: the relative importance of intrapersonal change and interpersonal difference for explaining cultural differences by education levels.
+These results mostly re-frame and align previous findings, but the value of our approach lies in its ability to extend the debate to a broader set of theoretical questions.
+To show this potential, we turn to a specific empirical example: the relative importance of intrapersonal change and interpersonal difference for explaining cultural variation by education levels.
 
 Previous work has established a positive relationship between education and attitude stability, especially on issues related to American politics.
-This stability is often attributed to education facilitating "chronic information" -- a general understanding of the domain of American politics, including the positions held by major parties and political figures and how issues relate to one another at a logical or socio-logical level [@zaller1992; @alvarez2002; @boutyline2017].
-Because they have more knowledge of American politics, people with a college education are better able to consistently connect the considerations in their cognition with the answer choices they are presented with in a survey.
+This stability is often attributed to education facilitating "chronic information" -- a general understanding of and attention to the domain of American politics, including the positions held by major parties and political figures and how issues relate to one another at a logical or socio-logical level [@zaller1992; @alvarez2002; @boutyline2017].
+These perspectives argue that because college graduates have more knowledge of American politics, they are better able to consistently connect the considerations in their cognition with the answer choices they are presented with in a survey.
 
 This work has tended to focus on the fact that college-educated Americans give responses that are less likely to be affected by measurement error or short-term influences than the rest of the population [@zaller1992; @alwin2007].
-However, this focus on the difference in the non-systematic (residual) component of variance across groups obscures the fact that the relative importance of different components of systematic variance might differ across these two groups as well.
+But focus on the difference in the non-systematic or residual component of variance across groups obscures the fact that the relative importance of different components of systematic variance, intrapersonal change and interpersonal differences, might also differ across these groups.
 
-It could be the case that education, chronic information, and attention to the political sphere makes intrapersonal change more important in explaining differences among people with higher levels of education.
-For example, these people might be more likely to make systematic changes in response to the emergence of new information, new issues, or political realignments, while those without chronic information display more variance around their baselines without their baselines changing much [@zaller1992].
+There are theoretical reasons to believe that education might be associated with either more or less systematic variance attributable to intrapersonal change.
+On one hand, because college graduates are more connected to mainstream discourse and elite signals, they might be more likely to make durable changes in response to the emergence of new information, new issues, or political realignments, while those without chronic information display more variance around their baselines without their baselines changing much [@zaller1992].
 In contrast, it could be the case that the stability demonstrated by educated respondents reflects the fact that they have already formed durable opinions and are relatively closed off to new information.
-College could serve as a formative experience that solidifies personal culture for people who attend.
+In this perspective, college could serve as a formative experience that solidifies some dimensions of personal culture for people who attend.
 But among people who do not attend college, later life experiences might prove more important in forming or changing personal culture, as these experiences potentially provide information that college-educated peers received earlier.
 
-To test these competing propositions, we use the models presented above to calculate the total systematic variance for people with at least a bachelor's degree and people with less than a bachelor's degree at wave 1[^wave] of the General Social Survey's panels.[^missing] We focus on the GSS because it contains the largest number of questions tapping general political dispositions, which is the domain where education has proven particularly relevant for understanding attitude stability.
-The GSS also covers a particularly turbulent window of American politics from 2006 to 2014.
+To test these competing propositions, we use the models presented above to separately calculate $\tau$ values for people with at least a bachelor's degree and people with less than a bachelor's degree at wave 1[^wave] of the General Social Survey's panels.[^missing] We focus on the GSS because it contains the largest number of questions tapping general political dispositions, which is the domain where education has proven particularly relevant for understanding attitude stability.
+The GSS also covers a turbulent window of American politics from 2006 to 2014.
 This window covers the start of the Great Recession, debates about federal intervention in and regulation of Wall Street, the election of Barack Obama as the first black president in U.S.
@@
 
-Figure 3 plots the distribution of differences in $\tau$ values between people with less than a bachelor's degree and people with at least a bachelor's degree at wave 1 of the panel for all 183 items int he GSS.
-Values greater than 0 would indicate that intrapersonal change is more important in explaining the responses of college graduates than those of people without a college degree, while values less than 0 would indicate that intrapersonal change is less important for explaining variation among people with a college degree than those without a college degree.
+Figure 3 plots the distribution of differences in $\tau$ values between people with at least a bachelor's degree and people with less than a bachelor's degree at wave 1 of the panel for all 183 GSS items we explored.
+Values greater than 0 indicate that intrapersonal change is more important in explaining the responses of college graduates than those of people without a college degree, while values less than 0 indicate that intrapersonal change is less important for explaining variation among people with a college degree than those without a college degree.
 
@@
 
-There is a clear pattern in Figure 3: for more than 80 percent of GSS items explored here, intrapersonal change is a larger component of systematic variance for people without a college degree than for people with a college degree (i.e., they have negative values).
-While most of these differences are small in absolute terms (less than 2 percentage points), several are greater than 5 percentage points.
+There is a clear pattern in Figure 3: for more than 80 percent of GSS items explored here, intrapersonal change is a larger component of systematic variance for people without a college degree than for people with a college degree (i.e., the difference is negative).
+While most of these differences are small in absolute terms, less than 2 percentage points, several are greater than 5 percentage points.
 Given the distribution observed in Figure 1 showing that the systematic variance attributable to intrapersonal change averages around 0.09, a 5 percentage point difference between groups is quite substantial.
@@
 Republican) and ideological identification (liberal vs.
-conservative) on seven-point scales; four questions about the government's role in improving the condition of the poor, paying people's medical bills, giving special treatment to blacks, and doing things that private businesses could do, measured on five-point scales; a question about whether the government should do more to reduce income differences, measured on a seven-point scale; and one question about whether blacks should be given preferences in hiring, measured on a five-point scale.
+conservative) on seven-point scales; four questions about the government's role in improving the condition of the poor, paying people's medical bills, giving special treatment to black people, and doing things that private businesses could do, measured on five-point scales; a question about whether the government should do more to reduce income differences, measured on a seven-point scale; and one question about whether black people should be given preferences in hiring, measured on a five-point scale.
 Estimates of $\tau$ for these eight questions, for both education groups, are presented in Figure 4.
@@
 
-On all eight questions presented in Figure 4, intrapersonal change explains less systematic variance for people with a college degree than for people without a college degree.
+On all eight questions presented in Figure 4, $\tau$ values are greater for people without a college degree than for people with a college degree, meaning intrapersonal change explains less systematic variance for people with a college degree than for people without one.
 There are also large differences in $\tau$ values across questions for both groups.
@@
 
-In contrast, partisan identification and political ideology both show larger values of $\tau$ than the other questions, as well as a large difference between people with a college degree and those without.
+In contrast, partisan identification and political ideology both show larger values of $\tau$ than most other questions, as well as a larger absolute difference between people with a college degree and those without.
 Compared to the other questions explored here, change as adults plays a much larger role in explaining why people report the partisan identification and political ideology they do.
@@
 We note here that this meaningful difference in $\tau$ values across education groups and across questions would not be detectable using previous methods.
-For partisan identification and political ideology, both college-educated respondents and people with less than college degree would likely favor the AUM or LCAM over the SDM because they show evidence of some members of the population making intrapersonal change.
+For partisan identification and political ideology, both college-educated respondents and people with less than college degree would likely favor the AUM or LCAM over the SDM because they these questions both show evidence of some members of the population making intrapersonal change.
 In other words, using the tournament of models approach obscures the fact that the relative importance of intrapersonal change in explaining differences on these two items is different across these two groups, that intrapersonal change explains more systematic variance for ideological identification than partisan identification for people without a college degree, and that there appears to be more durable change on questions of affirmative action than on questions of government aid to black Americans.
 
-It is important to note here that these differences should not be interpreted as causal effects of attending college on the importance of social experiences for cultural change.
-These results are potentially confounded by age, social class, race, gender, and other factors that explain selection into higher education.
-We simply use these estimates as a starting point for a broader discussion.
+We believe these patterns shed new light on the mechanisms underlying differences in attitude behavior across groups.
+Something about college attendance limits the ability of experiences later in life to drive changes in personal culture.
+While these results should not be interpreted as causal effects of attending college -- they are potentially confounded by age, social class, race, gender, and other factors that explain selection into higher education -- they open up a set of new questions and dynamics to explore.
 
@@
 Instead of falling back on a "tournament of models" approach [@lersch2023: p.
-228] to call victory for one or the other possible answer to this question, it seeks to refocus the debate to the theoretically more productive question of _how_ initial differences between and durable changes within people collectively explain differences in the personal culture they exhibit.
-To enable exploring this and related questions, it proposes a new approach that quantifies the amount of observed systematic variance that is attributable to either interpersonal differences at baseline or intrapersonal change.
+228] to proclaim victory for only one answer to this question, it seeks to refocus the debate to the theoretically more productive question of the relative importance of intrapersonal change for explaining differences in the personal culture.
+It proposes a new approach that quantifies the amount of observed systematic variance that is attributable to either interpersonal differences at baseline or intrapersonal change over the duration of a panel.
 
-Applying our proposed measure, $\tau$, to all panel datasets previously discussed in this debate revealed a consistent pattern.
-Nearly all questions on personal culture show evidence of durable, intrapersonal change over time, with some showing notably high intrapersonal change compared to interpersonal differences.
-For questions about factors that affect one’s life satisfaction or views on government benefits, it seems plausible that differences in adult experiences predominantly account for the observed differences between people.
+Applying our proposed measure, $\tau$, to all questions about personal culture from the panel datasets previously discussed in this debate revealed a consistent pattern.
+Nearly all questions show evidence of people making durable, intrapersonal change over time, with some showing notably high intrapersonal change compared to interpersonal differences.
+For some questions about factors that affect one’s life satisfaction or views on government benefits, it seems plausible that differences in adult experiences predominantly account for the observed differences between people.
 
-[TK Note: from now on I didn't change anything to $\tau$, but not sure if this is a correct way?].
+However, intrapersonal change is often substantially less pronounced than interpersonal differences, accounting for less than 10 percent of systematic variance on average across all datasets.
+On numerous questions, like those on civil liberties, abortion, generalized trust, and civic duty, the systematic variance attributable to intrapersonal change is essentially zero.
+For these questions, it seems there is not enough cultural change during adulthood to warrant attributing differences in personal culture mainly to experiences and social transitions; instead, the primary source of the observed differences appears to stem from experiences in childhood, adolescence, and early adulthood.
 
-Intrapersonal change is often substantially less pronounced than interpersonal differences, accounting for less than 10 percent on average across all datasets.
-On numerous questions, like those on civil liberties, abortion, generalized trust, and civic duty, the systematic variance attributable to intrapersonal change is essentially zero.
-For these questions, it seems there is not enough cultural change during adulthood to warrant attributing differences in personal culture mainly to experiences and social transitions; instead, the primary source of the observed differences appears to stem from early-life experiences [OR "EXPERIENCES PREDATING ADULT LIFE"].
-
-More than adjudicating past positions in this debate, the approach presented in this paper enables scholars to ask when change or stability in personal culture predominates within a given population and examine their role in explaining the cultural variation within it.
-To exemplify this, we detailed how the proportion in cultural differences that intrapersonal change can explain not only varies substantially by survey item but also by individuals' characteristics such as their education.
-Intrapersonal change plays almost no part in explaining sentiments regarding the government's economic role but is more defining for ideological and partisan identities.
-Intriguingly, it is more important in explaining cultural variation among non-college graduates compared to college graduates in the U.S..
-While age and other factors related to college completion might confound this pattern, this suggests that experiencing college may solidify personal culture to an extent that renders the effect of other adult experiences on political dispositions less influential for college graduates relative to non-college graduates.
+More than adjudicating past positions in this debate, the approach presented in this paper enables scholars to more clearly explore the relative contributions of change and stability in explaining variation in personal culture, as well as how these contributions might differ across populations.
+To exemplify this, we detailed how amount of cultural difference explained by intrapersonal change varies substantially by survey item but also by individuals' characteristics such as their education.
+Intrapersonal change plays almost no part in explaining sentiments regarding the government's economic role, but it is more defining for ideological and partisan identities.
+Intriguingly, it is more important in explaining cultural variation among non-college graduates compared to college graduates in the United States.
+While age and other factors related to college completion might confound this pattern, it suggests that college completion may solidify personal culture to an extent that renders the effect of other adult experiences on political dispositions less influential for college graduates relative to non-college graduates.
 
@@
 
-So far, sociologists interested in understanding cultural differences have largely asked about the presence and significance of cultural change in adults [e.g., @kiley2020; @vaisey2021; @lersch2023].
-Yet doing so, has unnecessarily limited and might even set the wrong focus.
+Sociologists interested in understanding cultural differences have largely asked about the presence and significance of cultural change in adults [e.g., @kiley2020; @vaisey2021; @lersch2023].
+Yet doing so has unnecessarily limited and might even set the wrong focus.
 In any population, some degree of cultural change in adults is inevitable.
@@
 Our results reinforce this point, showing the relevance of both interpersonal differences and intrapersonal change in understanding the differences in personal culture among adults across all items analyzed.
-[TRIED TO BETTER TUCK IT IN, BUT COULD LET GO OF THE LAST SENTENCE]
 
-This and other recent findings [CITES] suggest that it is theoretically more productive to ask about the relative importance of these two processes in understanding cultural variation.
+This and other recent findings [@stewart2023; @quinn2023] suggest that it is theoretically more productive to ask about the relative importance of these two processes in understanding cultural variation.
 Drawing a unified conclusion from survey items measuring various cultural forms on different scales and across different time frames is challenging.
@@
 To clarify, our findings do not imply that adults remain static or that their changes are inconsequential.
-Even minor shifts in personal culture, such as a 2% change in support for gay marriage, can have significant ramifications.
+Even minor shifts in personal culture, such as a 2 percentage-point change in support for gay marriage, can have significant ramifications.
 While the majority may remain consistent in their views, understanding the underlying mechanisms of even such minor shifts remains a crucial task for the sociology of culture.
@@
 
-This suggestion aligns with a range of recent causal inference work suggesting ["ADVOCATING" TO AVOID REPETITION] that selection effects, rather than treatment effects, predominantly account for personal cultural differences among individuals in varied social roles and positions [@campbell2016; @wodtke2018].
-While some individuals clearly change as they transition into new roles or environments, this change seems, on average, to be insufficient in magnitude and duration to possibly explain the pronounced differences among people in diverse roles.
+This conclusion aligns with a range of recent causal inference work suggesting that selection effects, rather than treatment effects, predominantly account for personal cultural differences among individuals in varied social roles and positions [@campbell2016; @wodtke2018].
+While some people clearly change as they transition into new roles or environments, this change often seems to be insufficient in magnitude and duration to explain what are often pronounced differences among people in diverse roles.
 [I'VE TRIED TO FIND A WAY TO GET BOTH IMPACT AND DURATION IN THERE; COULDN'T FIND A BETTER SOLUTION THOUGH--IDEAS?
 ALSO WE NEED TO BE CAREFUL WITH THIS SENTENCE, IT SEEMS WE ARE OVERSTATING HERE; LET'S FIND A WAY TO RELATIVIZE THE CLAIM HERE?].
-When observing differences in personal culture across social roles, such as parenthood, education, or professional authority [@longest2013], or across occupations [@weeden2005], scholar are well advised to expect selection as their primary cause, rather than treatment [ALTERNATIVE: "OUR INITIAL PRESUMPTION SHOULD LEAN TOWARDS..."].
+These and our results suggest that when observing differences in personal culture across social roles, such as parenthood, education, or professional authority [@longest2013], or across occupations [@weeden2005], selection likely plays a large role in explaining these difference, though exceptions always exist.
 
-Our analysis does not provide an answer as to why intrapersonal change seems to have limited impact on understanding cultural differences.
-The situations that promote durable change in personal culture might simply be rare.
+Our analysis does not provide an answer as to why intrapersonal change seems to have limited impact on understanding cultural differences among adults.
+The situations that promote durable change in personal culture might simply be rare during adulthood.
 Alternatively, it is possible that adults do encounter opportunities, necessities, and incentives for change, but their ability or willingness to change decreases.
@@
 In trying to understand differences in personal culture, sociologists should therefore pay more attention to the heterogeneous effects that various factors including social events, encounters, and situations can have.
-[IF WE ARE TOO LONG, CONSIDER COMPRESSING THE LAST TWO PARAGRAPHS INTO ONE OR EVEN DELETING ENTIRELY.]
 
 Our findings also have implications for understanding cultural change at the aggregate level.
-Given that intrapersonal change does not appear as the primary driver for cultural differences among adults, our results suggest that cultural change at the aggregate level is more likely driven by processes of cohort replacement than the by contemporaneous social conditions changing people's cultural dispositions [@vaisey2016; @underwood2022].
-[COULD EVEN JUST SAY "THAN THE INFLUENCE OF CONTEMPORANEOUS SOCIAL CONDITIONS" BUT FELT A BIT TOO SHORT.] This likely holds true for cultural change at the macro societal level as well as the micro level such as within organizations, political parties, and professions.
-Simultaneously, our findings indicate that the impact of formative experiences, as opposed to contemporaneous social factors, on cultural change is contingent upon an individuals' level of education.
+Given that intrapersonal change does not appear as the primary driver for many cultural differences among adults, our results suggest that many forms of cultural change at the aggregate level are more likely driven by processes of cohort replacement than the by contemporaneous social conditions changing people's cultural dispositions [@vaisey2016; @underwood2022].
+This likely holds true for cultural change at the macro societal level as well as the micro level such as within organizations, political parties, and professions.
+Our findings also indicate that the impact of formative experiences, as opposed to contemporaneous social factors, on cultural change is contingent upon an individuals' level of education.
 The weight of these two factors might thus change as the prominence of education shifts in the life course.
@@
 Finally, a key insight from our meta-analysis is that an increase in response resolution -- the range of response options survey respondents are given -- correlates with an increased share of systematic variance explained by intrapersonal change.
-This may reflect specific issues being assessed by survey items using [COULD DELETE "SURVEY ITEMS USING"] different scales.
+This may reflect specific issues being assessed by survey items using different scales.
 Alternatively, it might imply that when intrapersonal changes occurs, they are often subtle with answers moving from “agree” to “strongly agree,” rather than from “strongly agree” to “strongly disagree.” Such subtle changes require finer response options to be detected.
@@
 
-The approach outlined here generates estimates of how much of the observed variance in a population can be attributed to one of two theoretical processes: intrapersonal change and interpersonal differences at baseline.
+The approach outlined here generates estimates of how much of the observed variance in a population can be attributed to one of two sets of theoretical processes: intrapersonal change and interpersonal differences at baseline.
 It does not quantify the proportion of people who “change,” nor can we be sure that the amount of interpersonal change we detect is driven by many people making small changes or a few people making large changes.
@@
 Similarly to the LCAM, our approach treats cultural trajectories as varying linear slopes for each respondent, thus assuming that change is a linear function of time.
-There is no doubt, this assumption severely simplifies reality in which change likely also takes non-linear and discontinuous forms.
+This assumption simplifies reality in which change likely also takes non-linear and discontinuous forms.
 People might jump from one “stable” disposition to another or experience a "turning point" in their lives that upsets an otherwise stable trajectory.
 In parts, our assumption of linearity is a limitation of the data, as most questions are only observed for three to six waves.
-Panels with more waves might allow to loosen this assumption to test alternative, more flexible models of change.
+Panels with more waves might allow researchers to loosen this assumption to test alternative, more flexible models of change.
 
@@
 This is a sensible assumption on short panels where classifying change that lasts less than two years as durable seems unreasonable.
-Practically, it means that people making durable changes in their cultural dispositions to then return to a previous state later in life are classified as neither accounted for by the two theoretical processes.
+Practically, it means that the variance produced by people making durable changes in their cultural dispositions to then returning to a previous state later in life is classified as residual variance, rather than intrapersonal change.
 Longer panels and more flexible definitions of "change" would allow us to account for such trajectories and thus potentially also identify more durable change even in the same population.
@@
 
-We believe the approach outlined here can push past the “needless dichotomy” implicit in the question of whether people change or not and address more theoretically informed questions about the conditions under which the relative importance of intrapersonal change and interpersonal differences matter for explaining cultural differences.
-Characterizing questions as displaying change or not can only take researchers so far, but the question of why some questions demonstrate more change while others do not, or why some groups are characterized by more stability than others, has the potential to weigh in on a broader range of theoretical debates.
+We believe the approach outlined here can push past the “needless dichotomy” implicit in the question of whether people change or not and address more theoretically informed questions about the conditions under which the relative importance of intrapersonal change and interpersonal differences in explaining cultural differences shifts.
+Characterizing questions as displaying change or not can only take researchers so far, but the question of whether some questions demonstrate more change than others, or whether some groups are characterized by more stability than others, has the potential to weigh in on a broader range of theoretical debates.
 We hope researchers move toward those questions.

### 2023-09-27 manuscript.Rmd
adding figure notes.

@@
 
-[TK Note: \phi taken in Kiley & Vaisey, \delta taken in Vaisey & Lizardo.
+[TK Note: \\phi taken in Kiley & Vaisey, \\delta taken in Vaisey & Lizardo.
 I think using a Greek letter without strong associations (something like alpha, beta, or mu or whatever) is not good.
-So I settled on \tau.
+So I settled on \\tau.
 Feel free to change it.]
@@
 
-```{r fig1, fig.height=2, fig.cap="Proportions of explained variance in items of personal culture; based on predictions at participants' wave mid-point."}
+```{=latex}
+\begin{figure}[ht]
+\caption{Proportions of Systematic Variance in Personal Culture}
+```
+```{r fig1, fig.height=2}
 knitr::include_graphics("fig/variation_share_lm_mdpnt.png")
+```
+```{=latex}
+\textit{Notes:} The figure shows $\tau$ and 1 - $\tau$ as the proportion of systematic variance attributable to intrapersonal change and interpersonal differences, respectively.
+See Appendix A for the full set of item values.
+\end{figure}
 ```
@@
 \newpage
-```{r fig2, out.width = "300px", fig.cap="Proportion of explained variance attributable to intrapersonal change; based on predictions at participants' wave mid-point; supressed intercept model; survey indicators not shown."}
+```{=latex}
+\begin{figure}[h]
+\caption{Regression Model Estimating $\tau$}
+
+```
+```{r fig2, out.width = "300px"}
 knitr::include_graphics("fig/meta_no_int.png")
+```
+```{=latex}
+\textit{Notes:} The model estimates the proportion of systematic variance attributable to intrapersonal change.
+It includes the log of the number of participants, the number of waves the variable was asked (ceiled at t = 10 for each item), the date of the first wave (measured as the year minus 1968 divided by 10), number of waves the question was asked averaged across participants, duration in years per item averaged across participants, and panel and topic fixed effects.
+Coefficients are estimated from suppressed intercept model based on predictions at participants' wave mid-point.
+Survey indicators and item topics not shown.
+See Appendix B for the full set of coefficients.
+\end{figure}
 ```
@@
 \newpage
-```{r fig3, out.width = "450px", fig.cap="Difference between share of intrapersonal change and share of inter-personal differences in expalined variance; based on predictions at participants' wave mid-point."}
+```{=latex}
+\begin{figure}[h]
+\caption{Difference in $\tau$ Across College Graduates and Non-Graduates}
+```
+```{r fig3, out.width = "450px"}
 knitr::include_graphics("fig/FIG_college1.png")
+```
+```{=latex}
+\textit{Notes:} The figure shows the difference in $\tau$ values across college graduates and college non-graduates.
+Values above (below) 0 means that those with college degree have higher (lower) variance of intrapersonal change.
+The dashed red line marks 0 difference.
+\end{figure}
 ```
@@
 \newpage
-```{r fig4, out.width = "450px", fig.cap="Proportions of intrapersonal change of explained variance by college degree; based on predictions at participants' wave mid-point."}
+```{=latex}
+\begin{figure}[h]
+\caption{Difference in $\tau$ Across College Graduates and Non-Graduates on Political Culture}
+```
+```{r fig4, out.width = "450px"}
 knitr::include_graphics("fig/FIG_college2.png")
 ```
+```{=latex}
+\textit{Notes:} $\tau$ values across college graduates and college non-graduates on 8 political culture items from the General Social Survey (2006-2014).
+\end{figure}
+```

### 2023-09-27 manuscript.tex
small changes and the PDF knitted

@@
 \documentclass[
-  11pt,
+  12pt,
 ]{article}
 \usepackage{amsmath,amssymb}
-\usepackage{lmodern}
 \usepackage{iftex}
@@
 \else % if luatex or xetex
-  \usepackage{unicode-math}
+  \usepackage{unicode-math} % this also loads fontspec
   \defaultfontfeatures{Scale=MatchLowercase}
   \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
+\fi
+\usepackage{lmodern}
+\ifPDFTeX\else
+  % xetex/luatex font selection
+  \setmainfont[]{Times New Roman}
 \fi
@@
 \usepackage{xcolor}
-\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
-\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
-\hypersetup{
-  pdftitle={Quantifying Stability and Change in Personal Culture Using Panel Data},
-  hidelinks,
-  pdfcreator={LaTeX via pandoc}}
-\urlstyle{same} % disable monospaced font for URLs
 \usepackage[margin=1in]{geometry}
@@
 \newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}
+\usepackage{setspace}\doublespacing
 \usepackage{booktabs}
@@
 \fi
+\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
+\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
+\urlstyle{same}
+\hypersetup{
+  pdftitle={Quantifying Stability and Change in Personal Culture Using Panel Data},
+  hidelinks,
+  pdfcreator={LaTeX via pandoc}}
 
@@
 processes offers new ways to address theoretical questions of when and
-how their relative importance shifts. We examplify this with the case of
-college experience.
+how their relative importance shifts. We examplify this approach by
+exploring how the importance of these processes differs between college
+graduates and the rest of the population.
 \end{abstract}
@@
 A recurring focus in the sociology of culture is the question of whether
-people, as they move through their lives, change their personal culture
---- the declarative and non-declarative attitudes, beliefs, values,
-practices, and dispositions (Kiley and Vaisey 2020; Lersch 2023; Lizardo
-2017).
-
-This seemingly simple question underlies a number of central debates
-across sociological subfields. For example, pragmatist theories of
-action claim that changes in social environments cause people to adapt
-their views and make new meanings (Gross 2009; Swidler 2001), while
+people, as they move through their lives and transition between social
+roles, change their personal culture --- the declarative and
+non-declarative attitudes, beliefs, values, practices, and dispositions
+that exist at the individual level (Kiley and Vaisey 2020; Lersch 2023;
+Lizardo 2017).
+
+This seemingly simple question underlies a number of debates across
+sociological subfields. For example, pragmatist theories of action claim
+that changes in social environments cause people to adapt their views
+and make new cultural meanings (Gross 2009; Swidler 2001), while
 Bourdieusian practice theories argue that the ``past conditions of
@@
 influence assume that people adapt their culture in the face of new
-information (Vaisey and Lizardo 2010), while the emphasis on cohort
-effects in models of aggregate social and cultural change requires them
-to be open to change while young but become fairly resistant to it as
-they age (Ryder 1965). Finally, life course theories posit slower but
-still important changes over time as people advance through important
-transitions in their lives (Bardi and Goodwin 2011; Elder and George
-2016).
+information (Vaisey and Lizardo 2010 KK Note: I don't feel like this is
+a good cite for this.), while the emphasis on cohort effects in models
+of aggregate social and cultural change requires them to be open to
+change while young but become fairly resistant to it as they age (Ryder
+1965). Finally, life course theories posit slower but still important
+changes over time as people advance through important transitions in
+their lives (Bardi et al. 2009; Elder, Johnson, and Crosnoe 2003).
 
 Because all of the above theoretical perspectives have some empirical
-support, the current theoretical debate revolves around not whether they
-exist, but when and how they each apply and the relative contribution of
-each to explaining cultural differences. That being said, scholars have
-struggled to reach a consensus on the relative importance of change
-during adulthood. Over the span of a few years, most apparent changes in
-people's responses to personal culture items appear to be transitory,
-with little evidence of persistent change among adults (Kiley and Vaisey
-2020; Vaisey and Kiley 2021). These results suggest change is not a
-strong force in explaining personal cultural differences. When looked at
-over longer time horizons, however, there is at least some evidence that
-adults make persistent cultural changes, especially as they move through
-distinct social roles across the life course (Lersch 2023).
+support, the current debate revolves around not whether they exist, but
+when and how they each apply and the contribution of each to explaining
+cultural differences. That being said, scholars have struggled to reach
+a consensus on the relative importance of change during adulthood. Over
+the span of a few years, most apparent changes in people's responses to
+personal culture items appear to be transitory, with little evidence of
+persistent change among adults (Kiley and Vaisey 2020; Vaisey and Kiley
+2021). This suggests that change during adulthood is not a strong force
+in explaining personal cultural differences. When looked at over longer
+time horizons, however, there is at least some evidence that adults make
+persistent cultural changes, especially as they move through distinct
+social roles (Lersch 2023).
 
@@
 researchers have taken a ``tournament of models'' approach in
-adjudicating the different perspectives (Lersch 2023: p.~228).
+adjudicating different theoretical perspectives (Lersch 2023: p.~228).
 Accordingly, researchers have formalized models of responses to cultural
-items over time as demonstrating either ``change'' or ``no change'' to
-then compare which model fits observed data better, thus functionally
-asking whether we ever observe within-person change in personal culture.
-
-We argue, however, that asking only \emph{whether} people change
-obscures what are actually highly compatible empirical patterns. Whether
-researchers are able to detect ``change'' in a given dataset depends not
-only on sample populations, item measurement, statistical methods and
-power, but crucially also on definitions of what counts as ``change.''
-These methodological differences therefore limit the possibility of
-convergence on the theoretical debates. Moreover, asking whether these
-is evidence of change limits the kind of questions that can be asked to,
-``Do we see change in this measure?'' But such inquiries are ill-suited
-to address what's at the core of the theoretical debate. There is enough
-evidence to assume that neither socialization at an early age nor the
-influence of the contemporary environment alone can explain cultural
-differences. The question that remains is under what conditions, across
-which groups, and in what domains do we observe differences in the
-relative importance of intrapersonal change and interpersonal
-differences. This focus is necessarily comparative and requires an
-approach that quantifies forms of cultural change rather than calling
-victory for the one or the other perspective.
-
-In this paper, we try to move beyond previous debates by moving away
-from the ``tournament of models'' approach toward a quantification of
-the processes that lead to differences in personal culture. Drawing on
-the seven panel datasets analyzed in these debates, we use established
-methods to quantify the amount of variance in responses that is due to
+items over time as demonstrating either ``change'' or ``no change,''
+which essentially asks whether we \emph{ever} observe intrapersonal
+change in personal culture.
+
+We argue that asking only \emph{whether} people change obscures what are
+actually highly compatible empirical patterns. Being able to detect
+change in a given dataset depends not only on sample populations, item
+measurement, statistical methods, and power, but crucially also on
+definitions of what counts as ``change.'' These methodological decisions
+potentially limit the possibility of convergence on the theoretical
+debates. Moreover, asking whether there is evidence of change limits the
+kind of questions that can be asked to, ``Do we see change in this
+measure?'' But such inquiries are ill-suited to address what's at the
+core of the theoretical debate. There is enough evidence to assume that
+neither socialization at an early age nor the influence of the
+contemporary environment alone can explain cultural differences. The
+question that remains is under what conditions, across which groups, and
+in what domains do we observe differences in the relative importance of
+intrapersonal change and interpersonal differences. This focus is
+necessarily comparative and requires an approach that quantifies forms
+of cultural change rather than calling victory for the one or the other
+perspective.
+
+In this paper, we try to intervene in this debate by moving beyond the
+``tournament of models'' approach toward a quantification of the
+processes that lead to differences in personal culture. Drawing on the
+seven panel datasets analyzed in the debate, we use established methods
+to quantify the amount of variance in responses that is due to
 between-people differences at a single point during the survey versus
-the amount of variance in responses that is due to linear within-person
+the amount of variance in responses that is due to linear intrapersonal
 change over time. This measure gives us a sense of the importance of
-intrapersonal change over the course of the panel relative to the
-differences that exist between people assuming they never changed.
-
-We observe similar patterns across the datasets that have produced
-conflicting interpretations. In all datasets we see similar rates of
-individual-level change over time across a broad range of measures of
-personal culture. But quantifying these processes also reveals that
+change over the duration of the panel relative to the differences that
+exist between people assuming they never changed.
+
+We observe similar patterns across the datasets that yielded previously
+conflicting interpretations. In all datasets we find a similar
+distribution of how much explanatory power can be attributed to
+intrapersonal change. But quantifying these processes also reveals that
 intrapersonal change plays a substantially smaller role in explaining
-the variance in peoples' personal culture than between-person
-differences in their baselines. Of course, whether this amount of
-intrapersonal change is ``meaningful'' ultimately remains a theoretical
-question reflected in prior expectations concerning the question,
-population, and time. Nevertheless, with important exceptions, the
-results suggest that intrapersonal change (in some cases over more than
-10 years) does not seem to play a substantial role in explaining a broad
-range of personal cultural differences we observe between adults.
-
-These results help clarify and, hopefully, transcend previous debates.
-What is more, the measure we propose offers a tool with wider utility
-than merely discerning durable change that can help advance theoretical
-debates. To exemplify this, we turn to the role of education for
-cultural differences. Specifically, we use our measure to better
-understand what mechanisms might underlie differences observed between
-people with and those without college degrees. Focusing on eight
-questions that tab general political dispositions, our measure enables
-us to compare the relative importance of intrapersonal change for both
-groups in ways that the tournament of models cannot. We find less
-evidence of durable change among those with college degree. This may
-imply that the college experience predominantly serves as a formative
-period that solidifies political dispositions, rather than fostering an
-openness {[}ADAPTABILITY?{]} to new information and influences later in
-the life. {[}MISSING AN ENDING SENTENCE NOW.{]}
+the variance in peoples' personal culture than interpersonal differences
+at their baselines. Whether the amount of change is large enough to be
+``meaningful'' ultimately remains a theoretical question reflected in
+prior expectations about particular cultural item and population. It
+also crucially depends on how long people are observed in a panel.
+Nevertheless, on many measures of personal culture we do not observe
+enough intrapersonal change, in some cases over more than 10 years, for
+it to play a substantial role in explaining the differences we observe
+between adults, though notable exceptions exist.
+
+These results help clarify and, hopefully, transcend previous debates
+about whether adults change. What is more, the approach we propose can
+be applied to a broader range of theoretical questions than approaches
+discerning whether durable change exists in a panel. To exemplify this,
+we explore differences in the explanatory power of intrapersonal change
+for people with college degrees and those without. We find that change
+has less explanatory power for people with college degrees than those
+without. This may imply that the college experience solidifies political
+dispositions rather than fosters an openness and adaptability to new
+information and influences later in life. This offers a novel, empirical
+basis to theorize about the role of college in shaping personal culture.
+We believe that this approach has the potential to advance the
+theoretical debate on the role of environments, social identities, and
+roles in facilitating personal cultural change and stability.
 
@@
 their environments, this model assumes that people return to a settled
-baseline over a (relatively) short period of time. This model is
-frequently invoked in theories of cultural change that suggest people
-are strongly imprinted by early socialization experiences such as the
-``past conditions of production'' a Bourdieusian practice theory
-(Bourdieu 1990), cohort replacement theories of aggregate change
-(Mannheim 1952; Ryder 1965), and control theories in social psychology
-(Robinson 2007; Smith-Lovin and Heise 1988).
+baseline over a short period of time. This model underlies theories of
+cultural change that suggest people are imprinted by early socialization
+experiences such as the ``past conditions of production'' a Bourdieusian
+practice theory (Bourdieu 1990), cohort replacement theories of
+aggregate change (Mannheim 1952; Ryder 1965), and control theories in
+social psychology (Robinson 2007; Smith-Lovin and Heise 1988).
 
@@
 cultural diffusion (Christakis and Fowler 2010), attitude alignment
-(DellaPosta 2020), and polarization (Bail et al. 2018) It is also
+(DellaPosta 2020), and polarization (Bail et al. 2018). It is also
 implicit in most studies that ask whether specific experiences, changes
@@
 (Alwin and Krosnick 1991; Eaton et al. 2009; Krosnick and Alwin 1989).
-Similarly, highly salient issues, such as views around gay rights in the
-2010s; issues that see substantial elite realignments, such as views
-around the Vietnam conflict in the 1970s (Zaller 1992); and novel issues
-of public opinion, such as views around vaccines during the Covid-19
-pandemic, might be characterized by active updating, while established
-issues of low salience are characterized by stability.
+Similarly, salient issues, such as views around gay rights in the 2010s;
+issues that see substantial elite realignments, such as views around the
+Vietnam conflict in the 1970s (Zaller 1992); and novel issues of public
+opinion, such as views around vaccines during the Covid-19 pandemic
+(Scoville et al. 2022), might be characterized by active updating, while
+established issues of low salience are characterized by stability.
 
@@
 questions from the General Social Survey's rotating panels, Kiley and
-Vaisey (2020) found limited evidence of durable change. While the AUM
-was preferred for the majority of items, the total amount of durable
-change detected on these items was small. A substantial minority of
-questions (39 percent) favored the SDM, meaning they showed no evidence
-of durable change. Questions with more evidence of durable change
-included highly salient issues like gay marriage and ``public'' issues
-like partisan identification and religious service attendance. There was
-also more evidence of durable change among early adults (people ages
-18-30) than the rest of the population. Overall, the researchers
-concluded that ``results ultimately suggest that real, persistent
-attitude change is an uncommon phenomenon among adults'\,' (Kiley and
-Vaisey 2020: p.~500; see also Vaisey and Kiley 2021). This lack of
-durable change is consistent with other findings that cohort replacement
-plays a larger role than period effects in explaining differences in
-personal culture (Vaisey and Lizardo 2016).
+Vaisey (2020) found limited evidence of durable change. While the
+majority of items preferred the AUM, the amount of durable change
+detected on these items was small. A substantial minority of questions
+(39 percent) favored the SDM, meaning they showed no evidence of durable
+change. Questions with more evidence of durable change included salient
+issues like gay marriage and questions tapping ``public'' statements or
+behaviors such as partisan identification and religious service
+attendance. There was also more evidence of durable change among early
+adults (people ages 18-30) than the rest of the population. Overall, the
+researchers concluded that ``results ultimately suggest that real,
+persistent attitude change is an uncommon phenomenon among adults'\,'
+(Kiley and Vaisey 2020: p.~500; see also Vaisey and Kiley 2021). This
+lack of durable change is consistent with other recent findings that
+cohort replacement plays a larger role than period effects in explaining
+differences in personal culture (Vaisey and Lizardo 2016).
 
@@
 
-Drawing on these findings, Lersch challenged the SDM and AUM as a
-``needless dichotomy'' (2023). Instead he proposed the ``Life Course
-Adaption Model'' (LCAM). This model draws on the life course perspective
-to model personal culture as a linear trajectory over the course of a
-panel that varies for each respondent. Lersch rectifies two shortcomings
-of the AUM and SDM. First, the AUM, as described by Kiley and Vaisey,
-posits that changes follow a Markovian process where personal culture
-responses at time \(t\) rely solely on responses at time \(t-1\).
-Howevre, earlier life experiences can (directly and indirectly) mold
-personal culture when transitioning to new social roles or into new
-environments, even if their initial impact is delayed. For example,
-childhood events might influence views on family structures later when
-individuals form their own families. The LCAM considers therefore
-considers influences from earlier than just \(t-1\) on responses at time
-\(t\).
+Drawing on these findings, Lersch (2023) challenged the SDM and AUM as a
+``needless dichotomy,'' proposing the ``Life Course Adaption Model''
+(LCAM) as an alternative. This model draws on the life course
+perspective to model personal culture as a different linear trajectory
+over the duration of a panel for each respondent. In doing this, Lersch
+rectified two shortcomings of the AUM and SDM. First, the AUM, as
+described by Kiley and Vaisey, posits that changes follow a Markov
+process where responses at time \(t\) are a function of responses at
+time \(t-1\) but not earlier time points. However, earlier life
+experiences can (directly and indirectly) mold personal culture when
+transitioning to new social roles or into new environments, even if
+their initial impact is delayed. For example, childhood events might
+influence views on family structures later when individuals form their
+own families. Therefore, the LCAM considers influences from earlier than
+just \(t-1\) on responses at time \(t\).
 
@@
 three-wave panel data over four years, which might not be extensive
-enough to properly adjudicate {[}COMPARE? WE ARE USING ADJUDICATING A
-LOT{]} the two models. Lersch, however, evaluates the LCAM against the
-AUM and SDM using panel data spanning a wider duration (from 3 to 36
-years) and more waves (3 to 18), offering a better change to observe
+enough to adjudicate the two models. Lersch evaluates the LCAM against
+the AUM and SDM using panel data spanning a wider duration (from 3 to 36
+years) and more waves (3 to 18), offering a better chance to observe
 durable change.
 
-In Lersch's tournament of models, the LCAM gains broad {[}EXTENSIVE?{]}
-support from data across five countries. Out of 428 questions, the LCAM
-was favored in 297 cases while the SDM was preferred in only 112. The
-rest did not yield inconclusive results, and notably, no questions
-favoured the AUM. He concludes that ``new experiences over the life
-course {[}\ldots{]} can persistently move individuals' personal culture
-in novel directions'' (Lersch 2023: p.~243-244).
-
-Despite differing interpretations of these scholars, their empirical
-results align closely. Kiley and Vaisey (2020) discovered that most
-questions they investigated favored the AUM, indicating evidence of
-durable change among adults -- a finding largely in line with Lersch's.
-However, while Lersch (2023) found a preference for the LCAM in many
-cases, durable change only accounted for a minor proportion of variance
-explained. On average, individuals changed {[}CAN WE SAY SHIFTED TO
-INDICATE THE LINEAR DRIFT{]} about .07 standard deviations over 10
-years, echoing Kiley's and Vaisey's view that such changes are typically
-modest. Furthermore, a quarter of the items in Lersch's study favoured
-the SDM, suggesting that despite broader assumptions about change and
-more extensive data, many questions reveal no evidence of durable
-change. Likewise, in studies claiming evidence of durable change in
-personal culture, the magnitude is often minimal. For instance, Brocic
-and Miles (2021) note that obtaining graduate degrees in humanities,
-arts, and social sciences only shift {[}NUDGES?{]} peoples' moral
-relativism by about .2 standard deviations compared to non-degree
-holders {[}DELETED THE LAST SENTENCE PART HERE; FELT UNNECESSARY{]}.
-Moreover, studies of aggregate change demonstrate that even on questions
-where cohort effects prevail, there is always evidence that some
-individuals change over time (Vaisey and Lizardo 2016). In other words,
-despite different interpretations, the results of previous work are in
-many ways highly consonant.
+When Lersch compared the LCAM to the AUM and SDM on data from five
+countries, 297 of the 428 questions he analyzed preferred the LCAM,
+suggesting that we observe some linear change for adults on most items.
+The SDM was preferred on 112 items, and the rest did not yield
+conclusive results. No questions favored the AUM. He concludes that
+``new experiences over the life course {[}\ldots{]} can persistently
+move individuals' personal culture in novel directions'' (Lersch 2023:
+p.~243-244).
+
+Despite differences in how these researchers interpret their findings,
+the empirical results are not far apart. Kiley and Vaisey (2020) found
+that the majority of items they tested favored the AUM, meaning there
+was evidence of durable change among adults, a pattern consistent with
+Lersch's results. While Lersch (2023) found evidence of durable change
+on most items, this change only accounted for a small proportion of
+variance explained. On average, people changed only about .07 standard
+deviations over 10 years. This is consistent with Kiley and Vaisey's
+finding that durable change is often small. And a quarter of the items
+Lersch studied still preferred the SDM, suggesting that even with
+different assumptions assumptions about change and more extensive data,
+many questions show no evidence of durable change. Likewise, in studies
+showing evidence of durable change in specific measures of personal
+culture, the change is often small. For example, Brocic and Miles (2021)
+estimate that completing graduate degrees in humanities, arts, and
+social sciences shift peoples' moral relativism only about .2 standard
+deviations on average compared to people with no college degree, and
+this is the largest effect they identify. And studies of aggregate
+change show that, even on items where cohort effects explain more
+variance than period effects, there is always evidence that some people
+change over time (Vaisey and Lizardo 2016). In other words, despite
+different interpretations, the results of previous work are in many ways
+highly consonant. {[}I think the revisions that were done to this
+paragraph were confusing and mostly reverted it.{]}
 
@@
 sample, researchers will likely observe some evidence for durable
-changes in individuals. A lack thereof may simply due to a lack of
-clarity in the question asked or a low level of resolution in the
-possible answers, or that the question simply was not asked long enough
-or to enough people. Conversely, finding evidence of change tells us
-little about how much change is present in a population and how
+changes in individuals. A lack of evidence for change may simply due to
+a lack of clarity in the question asked or a low level of resolution in
+the possible answers, or that the question simply was not asked for long
+enough or to enough people. Conversely, finding evidence of change tells
+us little about how much change is present in a population and how
 meaningful that change is for social differences.
@@
 A more theoretically productive approach begins with a model assuming
-that during an observed time period some individuals might remain
-stable, might change a little, or might undergo significant shifts in
-their personal culture. Lersch's LCAM formally does this by modeling
-each individual as following their own trajectory over time. But rather
-than debating whether this model is superior to a model that assumes
-that people never change, researchers should quantify individual changes
-over time and compare those to the differences that exist between people
-at the panel's start. In other words, the focus should shift from
-asking, ``do people change?'' to asking, ``how do change and initial
-differences collectively explain variation in personal culture?'' {[}I
-CHANGED THAT QUESTION WORDING, PLEASE CHECK!{]}
-
-In this approach, two metrics reflect distinct theoretical processes.
-First, ``early imprinting,'' as Lersch calls it, denotes interpersonal
-differences at the start of the panel, capturing the accumulated
-experiences of people before participating in the panel that influence
-their answers. While commonly associated with experiences during an
-early formative period that result in settled disposition, this could
-also reflect transformative experiences (2014) that happen at any stage
-in life, as long as they predate the panel and consistently impact
-subsequent responses. For instance, for those people entering the panel
-post-retirement, this ``imprinting'' might reflect this pivotal life
-transition. Consequently, this metric also reflects variation in
-individuals' social roles or statuses at the start of the panel that
-were important in shaping their dispositions.
-
-The second process, termed ``persistent change'' or ``adaption'' by
-Lersch and ``active updating'' by Kiley and Vaisey, denotes durable
-changes in personal culture over time, captured by the amount of
-intrapersonal linear change during the panel. Lersch attributes these
-changes to social triggers such as moving into a new environment or
-adopting new social roles. Additionally, they might reflect the
-diffusion of new cultural forms across social networks, cues from
-political elites or otherwise culturally influential leaders, the
+that during an observed time period people might remain stable, might
+change a little, or might undergo significant shifts in their personal
+culture. Lersch's LCAM formally does this by modeling each individual as
+following their own trajectory over time. But rather than debating
+whether this model fits a particular data set better than a model that
+assumes that people never change, researchers should quantify individual
+changes over time and compare those to the differences that exist
+between people at the panel's start. In other words, the focus should
+shift from asking, ``do people change?'' to asking, ``what are the
+relative contributions of change and initial differences toward
+explaining differences in personal culture?'' {[}KK Note: I didn't like
+this change. It's not about collectively explaining, its about them
+relative to each other.{]}
+
+In this approach, there are two metrics that each reflect a combination
+of relevant theoretical processes. The first metric, interpersonal
+differences at the start of the survey, reflects the accumulated
+experiences of people prior to entering the panel survey that shape how
+they answer questions, what Lersch calls ``early imprinting.'' While
+commonly associated with experiences during an early formative period
+that result in settled dispositions, this variation could also reflect
+experiences that happen at any stage in life as long as they predate the
+panel and consistently impact subsequent responses. For instance, for
+those people entering the panel post-retirement, this ``imprinting''
+might reflect this pivotal life transition. Consequently, this metric
+also reflects variation in individuals' social roles or statuses at the
+start of the panel that were important in shaping their dispositions.
+{[}TK Note: I feel that this paragraph puts too much emphasis on
+pre-panel processes. It might be early imprinting or broad life-course,
+but why emphasize it this much? KK NOTE: Because it could be either, I
+don't want to treat it like it's just ``imprinting.''{]} {[}KK NOTE: You
+reversed what was being said here. You say ``early imprinting'' denotes
+intrapersonal change (process denotes number). We want to say the
+opposite -- what I had before -- number denotes process: ``interpersonal
+differences'' denotes ``early imprinting.'' I changed it back. I also
+took out reference to ``transformative'' experiences because I don't
+think we're necessarily talking about those specifically, just any
+experience that changes you. Because we don't say a specific theoretical
+process (``socialization'', etc.) I want to keep the language of each
+denoting combinations of processes. ``Persistent change'' isn't a
+theoretical process. It's a bunch of them.{]}
+
+The second metric, the amount of intrapersonal linear change people make
+during the panel, captures durable changes in personal culture over
+time, a set of processes collectively called ``persistent change'' or
+``adaption'' by Lersch and ``active updating'' by Kiley and Vaisey.
+Lersch attributes these changes to social triggers such as moving into a
+new environment or adopting new social roles. Additionally, they might
+reflect the diffusion of new cultural forms across social networks, cues
+from political elites or otherwise culturally influential leaders, the
 emergence of issues in politics or culture, or large-scale social
@@
 
-A third metric, often termed ``residual variance'' or ``fluctuation,''
-represents the remainder of variance in peoples' responses over time.
-This non-durable change may arise for various reasons. Some might lack
-clear cultural dispositions, leading them to construct an opinion or
-attitude during the survey, with varying thoughts and considerations
-dominating their thinking in each interview (Feldman and Zaller 1992;
-Tourangeau, Rips, and Rasinski 2000; Zaller 1992). Even though the full
-set of considerations they could draw on remains stable over time, this
-could yield to such variance in responses. Additionally, this variance
-can also include measurement errors, such as misinterpreted questions,
-erroneous response selections, or responses getting coded incorrectly.
-While this third process thus might reflect potentially important
-aspects of personal culture {[}I FELT WE NEED TO PLAY IT DOWN A BIT{]},
-it is less central to the debate outlined here. Hence, we focus on the
-first two metrics.
-
-At a theoretical level, the coexistence of two main processes --
-interpersonal differences and intrapersonal change -- is undeniable, and
-they are linked in many ways {[}THOUGHT IF intrinsically linked{]}. Any
-intrapersonal change during one's life will inevitably manifest as
+A third metric, often called ``residual variance'' or ``fluctuation,''
+accounts for the remainder of variance in peoples' responses over time.
+This non-durable change emerges for a variety of reasons. For example,
+people might not have clear personal cultural dispositions as measured
+by surveys. Instead, they might internalize a broad set of
+considerations and construct an opinion in the context of the survey
+interview, with different considerations coming to the forefront of
+their cognition each interview (Feldman and Zaller 1992; Tourangeau,
+Rips, and Rasinski 2000; Zaller 1992). This variance in considerations
+can produce variance in responses even if peoples' ``dispositions''
+--the full set of considerations they could draw on-- remains stable
+over time. Additionally, this variance can also include measurement
+errors, such as misinterpreted questions, erroneous response selections,
+or responses getting coded incorrectly. While this third process
+reflects important aspects of personal culture, it does not directly
+touch on the ongoing debate outlined here. Because of this we quantify
+this component but focus principally on the other two. {[}KK NOTE: I
+agreed with TK's note and mostly reverted this paragraph.{]}
+
+At a theoretical level, the coexistence of the sets of processes
+captured in the two main metrics, interpersonal differences and
+intrapersonal change, is undeniable, and they are linked in many ways.
+Any intrapersonal change during one's life will likely manifest as
 interpersonal differences by the time people enter a panel survey.
 Furthermore, unless people are entirely socialized early on and never
-deviate from these dispositions -- an understanding challenged by
-evidence -- we expect observing intrapersonal change in a segment of the
-population when surveyed over time.
+deviate from these dispositions, we expect observing intrapersonal
+change in a segment of the population when surveyed over time.
 
 Furthermore, these metrics are not necessarily useful in isolation or
-without context. Central to the theoretical debate is how the two
-processes they capture collectively explain cultural differences in a
-population. In other words, when examining adults across a period of
-time, does intrapersonal change, compared to interpersonal differences,
-sufficiently account for the observed cultural differences? Only through
-quantifying the two processes we can get closer to a true answer to this
-question.
-
-Yet, the true merit of the proposed quantification goes beyond settling
-past debates; it opens up new avenues for investigations that remained
-closed under previous approaches. For example, merely classifying
-questions based on whether they show durable change {[}DO WE NEED ``or
-not''? NEVER KNOW THAT AS A NON-NATIVE{]} overlooks possibly important
-differences among those that do. By quantifying the contributions of
-interpersonal differences and intrapersonal change, researchers can
-gauge their relative importance in understanding {[}INSTEAD
-explaining{]} differences among survey items, between groups within a
-given population, across time, and even across societies.
+without context. What is important for the theoretical debate is the
+relative contribution of these two components in explaining cultural
+differences in a population over a specific duration of time. In other
+words, when we look at a population of adults over some time frame, is
+there enough intrapersonal change relative to interpersonal difference
+to view it as an important explanation for why people demonstrate
+cultural differences? Only through quantifying the two processes we can
+get closer to a true answer to this question.
+
+Yet the true utility of the proposed quantification goes beyond settling
+past debates; it allows researchers to ask questions that were difficult
+to explore with previous approaches. For example, classifying questions
+based on whether they show durable change overlooks possibly important
+differences among those that do. By quantifying the relative
+contributions of interpersonal differences and intrapersonal change,
+researchers can gauge their relative importance in explaining
+differences among survey items, between groups within a given
+population, across time, and across societies.
 
@@
 
-By reframing the question and adjusting the approach, we expect
+By shifting the question and adjusting the approach, we expect
 consistent patterns across datasets previously analysed in the debate.
-First, we anticipate evidence for some intrapersonal change in most
-measures of personal culture across datasets. Yet, earlier studies found
-many questions (between 25 and 35 percent) lacking evidence of durable
-change. Aligning wiht this, we expect considerable variation in the
+First, because previous work found evidence of durable change on most
+measures of personal culture, we expect that intrapersonal change will
+have some explanatory power for most items. But because these earlier
+studies also found many questions (between 25 and 35 percent) lacking
+evidence of durable change, we expect considerable variation in the
 variance explained by intrapersonal change, with some questions
-indicating minimal variance attributable to such change. {[}I KEPT THE
-``VARIANCE EXPLAINED'' PHRASING THOUGH THE ``VARIANCE ATTRIBUTABLE''
-MIGHT BE LESS OF A HARD CLAIM WITHOUT LOSING MUCH?{]}
-
-Second, we expect the variance explained by intrapersonal to be modest
-compared to interpersonal differences at the start of the survey. This
-assertion {[}EXPECTATION? PREMISE?{]} is fairly uncontroversial,
-regardless of the theoretical process assumed to be most relevant for a
-particular question. Panel surveys do not encompase individuals' entire
-lifetimes. Thus, even if intrapersonal change explains a large part of
-cultural differences, the change detected in a panel covering 20 years
-probably will not fully account for the entire variance of interpersonal
-differences.
+indicating essentially no variance attributable to such change.
+
+Second, we expect the variance explained by intrapersonal change to be
+modest compared to interpersonal differences at the start of the survey.
+This expectation should be uncontroversial, regardless of the
+theoretical process assumed to be most relevant for a particular
+question. Panel surveys do not encompass individuals' entire lifetimes.
+Thus, even if intrapersonal change during adulthood is the major driver
+of cultural differences, the change detected in a panel covering 20
+years probably will not fully account for the entire variance of
+interpersonal differences.
 
@@
 studies. However, theoretical perspectives differ in how much variance
-they attribute to intrapersonal change, enabling quantification to
-provide a new perspective on existing questions. ``Formative period''
-and cohort theories, which emphasize early-life socialization as a
-strong force in shaping adult personal culture, suggest that the
-variance from intrapersonal change would still fall short of explaining
-interpersonal differences, even when extrapolated to a longer time frame
-{[}WE COULD DROP HERE A REFERENCE TO THE SPAN\ldots, possibly beyond the
-panel window{]}. Conversely, other theoretical perspectives emphasizing
-the importance of people's current social roles expect intrapersonal
-change could account for a large portion of the variance when
-extrapolated {[}CAN WE FIND AN ALTERNATIVE TO ``EXTRAPOLATE''?{]}. While
-projecting observed panel change beyond the covered time period may have
-its critics, we see it as a reasonable starting point. {[}DO WE NEED
-THIS LAST SENTENCE? IT FEELS WE ARE MAKING OUR SELVES VULNERABLE TO A
-CRITIQUE TO NO BENEFIT IN THE ARGUMENT.{]}
+they would expect to be driven by intrapersonal change, enabling
+quantification to provide a new perspective on existing questions.
+``Formative period'' and cohort theories, which emphasize early-life
+socialization as a strong force in shaping adult personal culture,
+suggest that the variance from intrapersonal change would still fall
+short of explaining interpersonal differences, even when we account for
+the duration of observation. Conversely, other theoretical perspectives
+emphasizing the importance of people's contemporaneous social
+environment for personal culture expect intrapersonal change to account
+for a large portion of the variance, regardless of the duration.
 
@@
 differences and intrapersonal change-- to depend on specific attributes
-of the question and panel. Noteably, the longer we observe respondents,
-the higher the response resolution for a given question, and the larger
-the sample, the greater the variance explained by intrapersonal change
-should be. Given that the LCAM (supported by the various underlying
-theoretical perspectives) posits that social experiences drive changes
-in personal culture and that these experiences correlate with time,
-observing people for longer durations {[}or TIME SPANS?{]} increases the
-chance that they undergo experiences that change their personal culture.
-{[}WE WOULDN'T EVEN NEED THE ``they undergo experiences that'' IN THAT
-SENTENCE.{]}
+of the question and panel. First, the higher the response resolution for
+a given question and the larger the sample, the greater the variance
+explained by intrapersonal change should be, as smaller changes in
+intrapersonal change will be easier to detect. Second, the LCAM and
+various theoretical perspectives that underlie it posit that social
+transitions drive changes in personal culture. Since the probability
+that any person experiences such a transition goes up the longer we
+observe them, we expect panel duration to be positively related to the
+proportion of variation explained by intrapersonal change.
 
@@
 the analyses that follow rely on a cross-cultural sample with a
-cross-domain set of items to capture the breadth of individual personal
-culture. Appendix A documents the list of all variables used in the
-upcoming analyses.
+cross-domain set of items to capture a broad range of individual
+personal culture. Appendix A documents the list of all variables used in
+the upcoming analyses.
 
@@
 \[
-y_{i, t} = \beta_0 + u_{0, i} + \beta_1 x_t + u_{1, i} x_{i, t} + \epsilon_{i, t}
+y_{i, t} = \beta_0 + u_{0, i} + \beta_{1} x_t + u_{1, i} x_{i, t} + \epsilon_{i, t}
+\] \[
+u_{0,i} \sim \mathcal{N}(0,\sigma_{\mu_0}^2)
 \]
+
+\[
+u_{1,i} \sim \mathcal{N}(0,\sigma_{u_1}^2)
+\]
+
+{[}KKNOTE: Is that right? I forget how to do that exactly.{]}
 
@@
 
-As a first step in our analysis, we fit LCAMs to each of the 609
+\hypertarget{analysis-steps}{%
+\subsection{Analysis Steps}\label{analysis-steps}}
+
+As a first step in our analysis, we fit a LCAM to each of the 609
 measures of personal culture outlined above. We then use these models to
@@
   trajectories.} We then divide this value by the total sum of squares
-and subtract it from from 1, producing a value we denote as \(\tau\), or
-the proportion of variance explained by interpersonal differences. This
+and subtract it from from 1, producing a value we denote as V(D), or the
+proportion of variance attributable to interpersonal differences. This
 number is equivalent to an \(R^2\) value generated by using random
@@
 We divide this number by the total sum of squares and subtract it from
-1. We then subtract \(\tau\) from this value. This gives us a value we
-denote as \(\delta\), the additional proportion of variance explained by
+1. We then subtract \(V(D)\) from this value. This gives us a value we
+denote as \(V(C)\), the additional proportion of variance explained by
 intrapersonal change.
 
-These two numbers reflect the total systematic or ``explained'' variance
-of the model that can be attributed to either differences across people
-or linear change over the course of the panel.\footnote{This does not
-  mean that other factors might not explain additional variance if they
-  were included in the model. For example, peoples' responses about
-  religious beliefs and participation might systematically vary by the
-  day of the week or month in which they are surveyed. This variation is
-  moved into the ``random'' or non-systematic component of the model for
-  us.} Our principal outcome of interest is the proportion of systematic
-variance accounted for by intrapersonal change, which we calculate as
+These two numbers reflect the total systematic variance of the model
+that can be attributed to either differences across people or linear
+change over the course of the panel.\footnote{This does not mean that
+  other factors might not explain additional variance if they were
+  included in the model. For example, peoples' responses about religious
+  beliefs and participation might systematically vary by the day of the
+  week or month in which they are surveyed. This variation is moved into
+  the ``random'' or non-systematic component of the model for us.} Our
+principal outcome of interest is the proportion of systematic variance
+accounted for by intrapersonal change, which we calculate as
 
 \[
-\frac{\delta}{\tau + \delta}
+\tau = \frac{V(C)}{V(C) + V(D)}
 \]
 
-We can think of this number as akin to an Intraclass Correlation
-Coefficient focused only on the systematic components of the model while
-ignoring residual variance that might be attributable to measurement
-error or non-persisting change (see also the variance decomposition of
-period and cohort by Vaisey and Lizardo 2016). The higher the number is,
-the more variance in responses is attributable to intrapersonal change
-during the panel. The lower the number, the less intrapersonal change
-seems to play a large role in explaining cultural differences during the
-panel.
+{[}TK Note: \textbackslash phi taken in Kiley \& Vaisey,
+\textbackslash delta taken in Vaisey \& Lizardo. I think using a Greek
+letter without strong associations (something like alpha, beta, or mu or
+whatever) is not good. So I settled on \textbackslash tau. Feel free to
+change it.{]}
+
+We can think of this number, \(\tau\) as akin to an Intraclass
+Correlation Coefficient focused only on the systematic components of the
+model while ignoring residual variance that might be attributable to
+measurement error or non-persisting change (see also the variance
+decomposition of period and cohort by Vaisey and Lizardo 2016). The
+higher this number is, the more systematic variance in responses is
+attributable to intrapersonal change during the panel. The lower the
+number, the less intrapersonal change seems to play a large role in
+explaining systematic cultural differences during the panel. Moving
+forward, we use \(\tau\) as a placeholder for \emph{systematic variance
+attributable to intrapersonal change}.
 
@@
 compare the relative importance of our two broad processes across a
-range of questions that might have very different levels of
-non-systematic variance, giving us a baseline sense of the relative
-prevalence of these two processes for each question (something not
-achievable under the tournament of models approach), and allowing us to
-compare which kinds of questions show more or less intrapersonal change
-relative to interpersonal differences. While the tournament of models
-approach allowed researchers to say whether questions in general showed
-evidence of updating, this approach allows us to make more specific
-claims about the prevalence of these theoretical processes for
-particular questions relative to each other.
+range of questions that might have different levels of non-systematic
+variance, giving us a sense of the relative prevalence of these two
+processes for each question (something not achievable under the
+tournament of models approach), and allowing us to compare which kinds
+of questions show more or less intrapersonal change relative to
+interpersonal differences. While the tournament of models approach
+allowed researchers to say whether questions in general showed evidence
+of updating, this approach allows us to make more specific claims about
+the prevalence of these theoretical processes for particular questions
+relative to each other.
 
 As a second step in our analysis, we use ordinary least squares
-regression to model the proportion of variance explained by
-intrapersonal change as a function of features of the question and
-panel. This allows us to explore how these features affect the
+regression to model \(\tau\) as a function of features of the question
+and panel. This allows us to explore how these features affect the
 prevalence of intrapersonal change. These covariates include the number
@@
 
+{[}TK Note: I guess we need to say something here about the college
+stuff? KK Note: I didn't feel like that was necessary. It feels so
+distinct later on that I think it's OK to kind of ``start over'' there.
+I'm worried that people will forget the plan by the time they get there,
+and we'll have to reiterate everything anyway. But others should weigh
+in.{]}.
+
 \hypertarget{results}{%
@@
 
-Our results proceed in three parts. We first explore the overall pattern
-of systematic variance explained by interpersonal differences at
-baseline and intrapersonal change over time across the seven panels.
-Next, we analyze the association between the relative proportion of
-systematic variance explained by intrapersonal change as a function of
-question, panel, and sample features. Finally, we highlight a few
-specific comparisons to illustrate the value of quantification relative
-to adjudication for further exploring theoretical questions.
-
 \hypertarget{variance-decomposition}{%
@@
 Figure 1 plots the proportion of systematic variance attributable to
-interpersonal differences (``between'') and intrapersonal change
-(``within'') for each of the 609 questions analyzed, plotted separately
-by panel. To broadly summarize the results, all panels show a similar
-range of the proportion of systematic variance attributable to
-intrapersonal change. Across panels, the median proportion of systematic
-variance attributable to intrapersonal change is 0.094 (mean of 0.098),
-with an interquartile range of 0.053 to 0.139. All panels include
-questions that show essentially no systematic variance attributable to
-intrapersonal change and all panels include questions where more than 20
-percent of systematic variance is attributable to intrapersonal change.
+interpersonal differences and intrapersonal change for each of the 609
+questions analyzed, plotted separately by panel. To broadly summarize
+the results, all panels show a similar range of the proportion of
+systematic variance attributable to intrapersonal change. Across panels,
+the median proportion of \(\tau\) is 0.094 (mean of 0.098), with an
+interquartile range of 0.053 to 0.139. All panels include questions
+where \(\tau\) is essentially 0, and all panels include questions with
+\(\tau\) values greater than 0.20. The maximum \(\tau\) value is 0.26.
 
@@
 <about here>
-Figure 1: Proportions of explained variance in items of personal culture; based on predictions at participants' wave mid-point.
 \end{center}
+
+Figure 1: Proportions of explained variance in items of personal
+culture; based on predictions at participants' wave mid-point.
 
@@
 To the extent that there are differences across the panels, the PSID has
-the highest mean (.120) and median (.135) proportion of systematic
-variance attributable to intrapersonal change. While we cannot
-disentangle features of the sample from features of the questions asked
-to each sample, the specific samples for many PSID questions have lower
-average ages than those from other panels. To the extent that younger
-respondents might be more likely to make durable changes of opinion,
-this higher explanatory power of intrapersonal change might reflect the
-distinct age profile of respondents in this sample. At the other end,
-the GSS has the lowest mean (.073) and median (.069) proportion of
-variance attributable to intrapersonal change. This potentially reflects
-the fact that the GSS observes people for a shorter duration, on
-average, than the other panels. If, consistent with life course adaption
-theories, people are more likely to make significant cultural changes
-the longer we observe them, then duration likely affects the amount of
-variance attributable to intrapersonal change (a point we explore more
-below). However, the GSS results are still highly consistent with
-results from the other panels.
+the highest \(\tau\) values with mean .120 and median .135. While we
+cannot disentangle features of the sample from features of the questions
+asked to each sample, the specific samples for many PSID questions have
+lower average ages than those from other panels. To the extent that
+younger respondents might be more likely to make durable changes of
+opinion, this higher explanatory power of intrapersonal change might
+reflect the distinct age profile of respondents in this sample. At the
+other end, the GSS has the lowest range of \(\tau\) with mean .073 and
+median .069. This potentially reflects the fact that the GSS observes
+people for a shorter duration, on average, than the other panels. If,
+consistent with life course adaption theories, people are more likely to
+make significant cultural changes the longer we observe them, then
+duration likely affects the range of \(\tau\) (a point we explore more
+below). However, the GSS results are still consistent with results from
+the other panels.
 
@@
 small compared to the differences within panels. For about 6 percent of
-items explored here, more than 20 percent of systematic variance is
-attributable to intrapersonal change. These questions tend to ask about
-external referents (e.g., confidence in specific government leaders or
-political parties), life satisfaction, or current financial position. At
-the other end, questions about religious identification, views on gender
-roles, and support for civil liberties tend to have very low proportion
-of variance attributable to intrapersonal change.
+items explored here, \(\tau\) is greater than 0.20. These questions tend
+to ask about external referents (e.g., confidence in specific government
+leaders or political parties), life satisfaction, or current financial
+position. At the other end, questions about religious identification,
+views on gender roles, and support for civil liberties tend to have very
+low proportion of variance attributable to intrapersonal change.
 
@@
 the press and confidence in religious leaders were both characterized by
-active updating, our results show that intrapersonal change is much more
+active updating. Our results show that intrapersonal change is much more
 important for explaining variance in confidence in the press (0.164)
@@
 Appendix A presents the estimated proportion of variance attributable to
-interpersonal differences and intrapersonal change, as well as the
-proportion of residual variance, for each question analyzed. In general,
-interpersonal differences are almost always the largest component of the
-total variance and tend to explain between 55 and 70 percent of total
-variance, while intrapersonal change is always the smallest, typically
-explaining between 3 and 8 percent of total variance. Residual variance
-tends to account for between 22 and 37 percent of variance (though on
-several questions residual variance is greater than 50 percent).
+interpersonal differences and intrapersonal change, the estimated values
+of \(\tau\), and the proportion of residual variance for each question
+analyzed. In general, interpersonal differences are almost always the
+largest component of the total variance and tend to explain between 55
+and 70 percent of total variance, while intrapersonal change is always
+the smallest, typically explaining between 3 and 8 percent of total
+variance. Residual variance tends to account for between 22 and 37
+percent of variance, though on several questions residual variance is
+greater than 50 percent.
 
@@
 Figure 2 plots the results from an ordinary least squares regression of
-proportion of systematic variance attributable to intrapersonal change
-as a function of question, panel, and sample features. These models also
-include fixed effects for panels (not shown), so coefficients reflect
+\(\tau\) as a function of question, panel, and sample features. These
+models also include fixed effects for panels, so coefficients reflect
 the association between a particular sample or question characteristic
-within a panel.\footnote{KK NOTE: Can we summarize these panel fixed
-  effects at all?}
+within a panel.{[}KK NOTE: CAN WE SUMMARIZE THESE PANEL FIXED EFFECTS AT
+ALL? AE: WE COULD SHOW THAT THEY DON'T CAPTURE TOPIC EFFECTS BEYOND OUR
+CLASSIFICATION IF WE INCLUDE THOSE / THAT THEY POSSIBLY CAPTURE
+UNMEASURED COUNTRY/CONTEXT DIFFERENCES OR DIFFERENCES IN SURVEY
+METHODOLOGY?{]}
 
@@
 <about here>
-Figure 2: Proportion of explained variance attributable to intrapersonal change; based on predictions at participants' wave mid-point; supressed intercept model; survey indicators not shown.
 \end{center}
 
+Figure 2: Proportion of explained variance attributable to intrapersonal
+change; based on predictions at participants' wave mid-point; supressed
+intercept model; survey indicators not shown.
+
 Figure 2 shows that the more response options respondents are given and
-the larger the sample, the larger the proportion of systematic variance
-is attributable to intrapersonal change. We interpret these coefficients
-as suggesting greater resolution on a question makes it easier to detect
-and model change. The earlier a question was asked, even net of how long
-it was been asked, the lower the proportion of systematic variance is
-attributable to intrapersonal change.
+the larger the sample, the larger the values of \(\tau\). We interpret
+these coefficients as suggesting greater resolution on a question makes
+it easier to detect and model change. The earlier a question was asked,
+even net of how long it was been asked, the lower the \(\tau\).
 
 The most notable result in the meta analysis is the negative coefficient
-attached to the span of years covered by the question. The longer a
-question is observed, the less systematic variance is attributable to
-intrapersonal change. Theories of personal cultural change that link
-changes in personal culture to social experiences, including the LCAM,
-suggest that the longer we observe respondents, the more likely people
-are to undergo potentially transformative experiences and therefore the
-more variance would be explained by intrapersonal change. The fact that
-we find a negative coefficient attached to this variable is a challenge
-to that assumption.
-
-As a further test of this coefficient, we compared the proportion of
-systematic variance attributable to intrapersonal change when using the
-full duration of a panel compared to when when we dropped the final wave
-and therefore reduced the total duration of observation for the
-question. If the coefficient reflects a true negative effect of duration
-on systematic variance attributable to intrapersonal change, we should
-see that same effect within questions. Results from this analysis are
-presented in Appendix B and contradict the coefficient from the
-regression model. The longer we observe the same question, the higher
-proportion of systematic variance is attributable to intrapersonal
-change. We interpret this combination of findings as suggesting that the
-kinds of questions asked for a longer time period tend to demonstrate
-less intrapersonal change than questions asked over shorter periods,
-rather than a true function of time.
+attached to the duration of years covered by the question. The longer a
+question is observed, the less is the value of \(\tau\). Theories of
+personal cultural change that link changes in personal culture to social
+experiences, including the LCAM, suggest that the longer we observe
+respondents, the more likely people are to undergo potentially
+transformative experiences and therefore the more variance would be
+explained by intrapersonal change. The fact that we find a negative
+coefficient attached to this variable can be a challenge to that
+assumption {[}TK Note: since there is a complication of this, I changed
+is to can be, but feel free to ignore it.{]}
+
+As a further test of this coefficient, we compared the \(\tau\) values
+when using the full duration of a panel compared to when we dropped the
+final wave and therefore reduced the total duration of observation for
+the question. If the coefficient reflects a true negative effect of
+duration on \(\tau\), we should see that same effect within questions.
+Results from this analysis are presented in Appendix B and contradict
+the coefficient from the regression model: we found that the longer we
+observe the same question, the higher the \(\tau\) is. We interpret this
+combination of findings as suggesting that the kinds of questions asked
+for a longer time period tend to demonstrate less intrapersonal change
+than questions asked over shorter periods, rather than a true function
+of time.
 
@@
 average proportion of systematic variance attributable to intrapersonal
-change: its shorter duration. {[}KK NOTE: Does this inform any other
-panel differences? Does the PSID, BHPS or UKHLS have longer durations on
-average?{]}
-
-\hypertarget{a-case-study-college-education}{%
-\subsection{A Case Study: College
-Education}\label{a-case-study-college-education}}
-
-While these broad results help speak to previous debates, they mostly
-re-frame previous findings without extending the theoretical debate. To
-better illustrate the value of our approach for addressing different
-theoretical questions, we turn to a specific empirical example: the
-relative importance of intrapersonal change and interpersonal difference
-for explaining cultural differences by education levels.
+change: its shorter duration.
+
+While we did not generate expectations for how question content relates
+to \(\tau\), some of the associations between question structure, panel
+duration, and panels themselves might be driven by differences in the
+topics addressed by each panel. To address this, we coded each question
+as falling into one of nine different content categories and included
+these as indicator variables in the meta analysis. Appendix C presents
+the estimates for these content indicators. They show some associations
+with \(\tau\), with questions about subjective SES; social life, social
+cohesion, and trust; environment and climate; and health and morale
+showing larger \(\tau\) values than questions about religion and
+spirituality; politics, government and the economy; and gender and
+family. {[}KKNOTE: I FEEL LIKE WE WANTED TO SAY SOMETHING ELSE HERE BUT
+I CANT REMEMBER WHAT{]}
+
+\hypertarget{college-and-change-in-political-culture}{%
+\subsection{College and Change in Political
+Culture}\label{college-and-change-in-political-culture}}
+
+These results mostly re-frame and align previous findings, but the value
+of our approach lies in its ability to extend the debate to a broader
+set of theoretical questions. To show this potential, we turn to a
+specific empirical example: the relative importance of intrapersonal
+change and interpersonal difference for explaining cultural variation by
+education levels.
 
@@
 politics. This stability is often attributed to education facilitating
-``chronic knowledge'' -- a general understanding of the domain of
-American politics, including the positions held by major parties and
-political figures and how issues relate to one another at a logical or
-socio-logical level (Alvarez and Brehm 2002; Boutyline and Vaisey 2017;
-Zaller 1992). Because they have more knowledge of American politics,
-people with a college education are better able to consistently connect
-the considerations in their cognition with the answer choices they are
-presented with in a survey.
+``chronic information'' -- a general understanding of and attention to
+the domain of American politics, including the positions held by major
+parties and political figures and how issues relate to one another at a
+logical or socio-logical level (Alvarez and Brehm 2002; Boutyline and
+Vaisey 2017; Zaller 1992). These perspectives argue that because college
+graduates have more knowledge of American politics, they are better able
+to consistently connect the considerations in their cognition with the
+answer choices they are presented with in a survey.
 
@@
 Americans give responses that are less likely to be affected by
-measurement error or short-term influences (Alwin 2007; Zaller 1992).
-However, this focus on the difference in the non-systematic (residual)
-component of variance across groups obscures the fact that the relative
-importance of different components of systematic variance might differ
-across these two groups as well.
-
-It could be the case that education, chronic knowledge, and attention to
-the political sphere makes intrapersonal change more important in
-explaining differences among people with higher levels of education. For
-example, these people might be more likely to make systematic changes in
-response to the emergence of new information, new issues, or political
-realignments, while those without chronic knowledge display more
-variance around their baselines without their baselines changing much
-(Zaller 1992). In contrast, it could be the case that the stability
-demonstrated by educated respondents reflects the fact that they have
-already formed durable opinions and are relatively closed off to new
-information. College could serve as a formative experience that
-solidifies personal culture for people who attend. But among people who
-do not attend college, later life experiences might prove more important
-in forming or changing personal culture, as these experiences
-potentially provide information that college-educated peers received
-earlier.
+measurement error or short-term influences than the rest of the
+population (Alwin 2007; Zaller 1992). But focus on the difference in the
+non-systematic or residual component of variance across groups obscures
+the fact that the relative importance of different components of
+systematic variance, intrapersonal change and interpersonal differences,
+might also differ across these groups.
+
+There are theoretical reasons to believe that education might be
+associated with either more or less systematic variance attributable to
+intrapersonal change. On one hand, because college graduates are more
+connected to mainstream discourse and elite signals, they might be more
+likely to make durable changes in response to the emergence of new
+information, new issues, or political realignments, while those without
+chronic information display more variance around their baselines without
+their baselines changing much (Zaller 1992). In contrast, it could be
+the case that the stability demonstrated by educated respondents
+reflects the fact that they have already formed durable opinions and are
+relatively closed off to new information. In this perspective, college
+could serve as a formative experience that solidifies some dimensions of
+personal culture for people who attend. But among people who do not
+attend college, later life experiences might prove more important in
+forming or changing personal culture, as these experiences potentially
+provide information that college-educated peers received earlier.
 
 To test these competing propositions, we use the models presented above
-to calculate the total systematic variance for people with at least a
+to separately calculate \(\tau\) values for people with at least a
 bachelor's degree and people with less than a bachelor's degree at wave
@@
 General Social Survey's panels.\footnote{We estimate the models
-  presented in {[}earlier figure{]} using the full sample, but only use
+  presented in Figures 1 and 2 using the full sample, but only use
   respondents who have information on their highest degree to generate
-  the results presented in figures {[}these figures{]}. In total, \#\#\#
-  respondents (\%\%\% percent) reported no degree information.} We focus
-on the GSS because it contains the largest number of questions tapping
-general political dispositions, which is the domain where education has
-proven particularly relevant for understanding attitude stability. The
-GSS also covers a particularly turbulent window of American politics
-from 2006 to 2014. This window covers the start of the Great Recession,
-debates about federal intervention in and regulation of Wall Street, the
-election of Barack Obama as the first black president in U.S. history,
-debates about the role of the federal government in the health care
-sector, the emergence of the Tea Party, and political realignment and
-clarification on the issue of gay marriage, among other topics.
-
-Figure 3 plots the distribution of differences between people with less
-than a bachelor's degree and people with at least a bachelor's degree at
-wave 1 of the panel. Values greater than 0 would indicate that
-intrapersonal change is more important in explaining the responses of
-college graduates than those of people without a college degree, while
-values less than 0 would indicate that intrapersonal change is less
-important for explaining variation among people with a college degree
-than those without a college degree.
+  the results presented in Figures 3 and 4. In total, \#\#\# respondents
+  (\%\%\% percent) reported no degree information.} We focus on the GSS
+because it contains the largest number of questions tapping general
+political dispositions, which is the domain where education has proven
+particularly relevant for understanding attitude stability. The GSS also
+covers a turbulent window of American politics from 2006 to 2014. This
+window covers the start of the Great Recession, debates about federal
+intervention in and regulation of Wall Street, the election of Barack
+Obama as the first black president in U.S. history, debates about the
+role of the federal government in the health care sector, the emergence
+of the Tea Party, and political realignment and clarification on the
+issue of gay marriage, among other topics.
+
+Figure 3 plots the distribution of differences in \(\tau\) values
+between people with at least a bachelor's degree and people with less
+than a bachelor's degree at wave 1 of the panel for all 183 GSS items we
+explored. Values greater than 0 indicate that intrapersonal change is
+more important in explaining the responses of college graduates than
+those of people without a college degree, while values less than 0
+indicate that intrapersonal change is less important for explaining
+variation among people with a college degree than those without a
+college degree.
 
@@
 <about here>
-Figure 3: Distribution of differences share of variance explained by intrapersonal change between respondents without and with college degree.
 \end{center}
+
+Figure 3: Difference between share of intrapersonal change and share of
+inter-personal differences in explained variance; based on predictions
+at participants' wave mid-point
 
@@
 systematic variance for people without a college degree than for people
-with a college degree (they have negative values in the figure). While
-most of these differences are small in absolute terms (less than 2
-percentage points), several are greater than 5 percentage points. Given
-the distribution observed in Figure {[}the earlier one{]} showing that
-the systematic variance attributable to intrapersonal change averages
-{[}average across all questions{]}, a 5 percentage point difference
-between groups is quite substantial.
+with a college degree (i.e., the difference is negative). While most of
+these differences are small in absolute terms, less than 2 percentage
+points, several are greater than 5 percentage points. Given the
+distribution observed in Figure 1 showing that the systematic variance
+attributable to intrapersonal change averages around 0.09, a 5
+percentage point difference between groups is quite substantial.
 
@@
 the government's role in improving the condition of the poor, paying
-people's medical bills, giving special treatment to blacks, and doing
-things that private businesses could do, measured on five-point scales;
-a question about whether the government should do more to reduce income
-differences, measured on a seven-point scale; and one question about
-whether blacks should be given preferences in hiring, measured on a
-five-point scale. Estimates for the proportion of systematic variance
-attributable to intrapersonal change on these eight questions, for both
-education groups, are presented in Figure 4.
-
-On all eight questions presented in Figure 4, intrapersonal change
-explains less systematic variance for people with a college degree than
-for people without a college degree. There are also large differences
-across questions on the amount of systematic variance attributable to
-intrapersonal change for both groups. For example, on the question of
-whether the government should try to solve more problems or leave those
-problems to be solved by private businesses (``government do more or
-less''), less than one percent of the systematic variance is
-attributable to intrapersonal change for both groups. In other words,
-while people might vacillate on this question at random (37 percent of
-variance is residual for this question), there is functionally no
-evidence that people make systematic changes of opinion on this issue
-during the GSS panel.
+people's medical bills, giving special treatment to black people, and
+doing things that private businesses could do, measured on five-point
+scales; a question about whether the government should do more to reduce
+income differences, measured on a seven-point scale; and one question
+about whether black people should be given preferences in hiring,
+measured on a five-point scale. Estimates of \(\tau\) for these eight
+questions, for both education groups, are presented in Figure 4.
 
@@
 <about here>
-Figure 4: Proportions of intrapersonal change of explained variance by college degree; based on predictions at participants' wave mid-point.
 \end{center}
 
-In contrast, partisan identification and political ideology both show a
-larger proportion of systematic variance attributable to intrapersonal
-change, with quite large differences between people with a college
-degree and those without. Compared to the other questions explored here,
-change as adults (and, potentially, their experiences as adults) plays a
-much larger role in explaining why people report the partisan
+Figure 4: Proportions of intrapersonal change of explained variance by
+college degree; based on predictions at participants' wave mid-point.
+
+On all eight questions presented in Figure 4, \(\tau\) values are
+greater for people without a college degree than for people with a
+college degree, meaning intrapersonal change explains less systematic
+variance for people with a college degree than for people without one.
+There are also large differences in \(\tau\) values across questions for
+both groups. For example, on the question of whether the government
+should try to solve more problems or leave those problems to be solved
+by private businesses (``government do more or less''), less than one
+percent of the systematic variance is attributable to intrapersonal
+change for both groups. In other words, while people might vacillate on
+this question at random (37 percent of variance is residual for this
+question), there is functionally no evidence that people make systematic
+changes of opinion on this issue during the GSS panel.
+
+In contrast, partisan identification and political ideology both show
+larger values of \(\tau\) than most other questions, as well as a larger
+absolute difference between people with a college degree and those
+without. Compared to the other questions explored here, change as adults
+plays a much larger role in explaining why people report the partisan
 identification and political ideology they do. And this is particularly
-true for respondents who do not have a college degree. The amount of
-systematic variance in ideological identification that is attributable
-to intrapersonal change among non-college educated respondents is almost
-four times that of college-educated respondents.
-
-We note here that this meaningful difference in the proportion of
-systematic variance attributable to intrapersonal change across
+true for respondents who do not have a college degree. The \(\tau\)
+value for ideological identification among non-college educated
+respondents is almost four times that of college-educated respondents.
+
+We note here that this meaningful difference in \(\tau\) values across
 education groups and across questions would not be detectable using
@@
 both college-educated respondents and people with less than college
-degree would likely favor the AUM or LCAM over the SDM because they show
-evidence of some members of the population making intrapersonal change.
-In other words, using the tournament of models approach obscures the
-fact that the relative importance of intrapersonal change in explaining
-differences on these two items is very different across these two
-groups, that intrapersonal change explains more systematic variance for
-ideological identification than partisan identification for people
-without a college degree, or that there appears to be more durable
-change on questions of affirmative action than on questions of
-government aid to black Americans.
-
-It is important to note here that these differences should not be
-interpreted as causal effects of attending college on the importance of
-social experiences for cultural change. These results are potentially
-confounded by age, social class, race, gender, and other factors that
-explain selection into higher education. We simply use these estimates
-as a starting point for a broader discussion.
+degree would likely favor the AUM or LCAM over the SDM because they
+these questions both show evidence of some members of the population
+making intrapersonal change. In other words, using the tournament of
+models approach obscures the fact that the relative importance of
+intrapersonal change in explaining differences on these two items is
+different across these two groups, that intrapersonal change explains
+more systematic variance for ideological identification than partisan
+identification for people without a college degree, and that there
+appears to be more durable change on questions of affirmative action
+than on questions of government aid to black Americans.
+
+We believe these patterns shed new light on the mechanisms underlying
+differences in attitude behavior across groups. Something about college
+attendance limits the ability of experiences later in life to drive
+changes in personal culture. While these results should not be
+interpreted as causal effects of attending college -- they are
+potentially confounded by age, social class, race, gender, and other
+factors that explain selection into higher education -- they open up a
+set of new questions and dynamics to explore.
 
@@
 their personal culture -- their attitudes, beliefs, values, practices,
-and dispositions -- as they move through life. Instead of falling back
-on a ``tournament of models'' approach (Lersch 2023: p.~228) to call
-victory for one or the other possible answer to this question, this
-paper seeks to refocus the debate to the theoretically more productive
-question of \emph{how} initial differences between and durable changes
-within people collectively explain the observed variation in personal
-culture. To do so, it proposes a new measure that quantifies the amount
-of systematic variance that is attributable to either interpersonal
-differences at baseline or intrapersonal change.
-
-Applying this measure to all datasets previously discussed in this
-debate revealed a consistent pattern. Nearly all questions on personal
-culture show evidence of intrapersonal change over time, with some items
-showing noteably high intrapersonal change compared to interpersonal
-differences. For questions about factors that affect one's life
-satisfaction or views on government benefits, it seems plausible that
-differences in adult experiences predominantly account for the observed
-differences between people.
-
-Intrapersonal change is often substantially less pronounced than
-interpersonal differences, accounting for less than 10 percent on
-average across all surveys. On numerous questions, like those on civil
-liberties, abortion, generalized trust, and civic duty, the systematic
-variance explained by intrapersonal change is essentially zero. For
-these questions, it seems there is not enough cultural change during
-adulthood to warrant attributing differences in personal culture mainly
-to experiences and social transitions; instead, the primary source of
-the observed differences appears to stem from early-life experiences.
-
-More important than adjudicating past debates, however, the approach
-presented in this paper offers offers the possibility to pose new
-theoretical questions regarding when change or stability in personal
-culture predominates and its significance in explaining cultural
-variances within a population. To examplify this potential, we detailed
-how the proportion in cultural differences that intrapersonal change can
-explain not only varies substantially by survey item but also by
-individual characteristics such as their education. Intrapersonal change
-plays almost no part in explaining sentiments regarding the government's
-economic role but is more defining for ideological and partisan
-identities. Intriguingly, it is more important in explaining cultural
-variation among people without a college compared to college graduates
-in the U.S.. While age and other factors related to college completion
-might confound this pattern, it suggests that college may solidify
-personal culture to an extent that renders the effect of other adult
-experiences on political dispositions for graduates less influential
-relative to individuals without such a college experience.
+and dispositions -- as they move through their adult life. Instead of
+falling back on a ``tournament of models'' approach (Lersch 2023:
+p.~228) to proclaim victory for only one answer to this question, it
+seeks to refocus the debate to the theoretically more productive
+question of the relative importance of intrapersonal change for
+explaining differences in the personal culture. It proposes a new
+approach that quantifies the amount of observed systematic variance that
+is attributable to either interpersonal differences at baseline or
+intrapersonal change over the duration of a panel.
+
+Applying our proposed measure, \(\tau\), to all questions about personal
+culture from the panel datasets previously discussed in this debate
+revealed a consistent pattern. Nearly all questions show evidence of
+people making durable, intrapersonal change over time, with some showing
+notably high intrapersonal change compared to interpersonal differences.
+For some questions about factors that affect one's life satisfaction or
+views on government benefits, it seems plausible that differences in
+adult experiences predominantly account for the observed differences
+between people.
+
+However, intrapersonal change is often substantially less pronounced
+than interpersonal differences, accounting for less than 10 percent of
+systematic variance on average across all datasets. On numerous
+questions, like those on civil liberties, abortion, generalized trust,
+and civic duty, the systematic variance attributable to intrapersonal
+change is essentially zero. For these questions, it seems there is not
+enough cultural change during adulthood to warrant attributing
+differences in personal culture mainly to experiences and social
+transitions; instead, the primary source of the observed differences
+appears to stem from experiences in childhood, adolescence, and early
+adulthood.
+
+More than adjudicating past positions in this debate, the approach
+presented in this paper enables scholars to more clearly explore the
+relative contributions of change and stability in explaining variation
+in personal culture, as well as how these contributions might differ
+across populations. To exemplify this, we detailed how amount of
+cultural difference explained by intrapersonal change varies
+substantially by survey item but also by individuals' characteristics
+such as their education. Intrapersonal change plays almost no part in
+explaining sentiments regarding the government's economic role, but it
+is more defining for ideological and partisan identities. Intriguingly,
+it is more important in explaining cultural variation among non-college
+graduates compared to college graduates in the United States. While age
+and other factors related to college completion might confound this
+pattern, it suggests that college completion may solidify personal
+culture to an extent that renders the effect of other adult experiences
+on political dispositions less influential for college graduates
+relative to non-college graduates.
 
@@
 
-So far, the debate in cultural sociology has centered around the
-presence and significance of cultural change in adults. Yet, arguing
-solely over its existence sets the wrong theoretical focus. In any adult
-population, some degree of personal cultural change is inevitable.
-Likewise, no viable theoretical perspective would expect just one
-process --stable differences or internal change between people-- to
-characterize the differences in personal cultural observed between them.
-Our results underscore this perspective, showing a combination of
-interpersonal differences and intrapersonal change across all items.
-{[}DELETED SENTECE, MISSPLACED{]}
-
-A more reasonable theoretical debate pushed by previous work is the
-relative importance of these two processes for explaining cultural
-variation. It is difficult to draw a single conclusion on this question
-across a wide range of questions measured on different scales and over
-different durations of time, but the overall picture, with important
-caveats, suggests that intrapersonal change during adulthood is not the
-dominant explanation for cultural differences among adults for most
-items we measure in surveys. There simply is not enough evidence of
-intrapersonal change on these items relative to the amount of difference
-we observe across people.
-
-To be clear, none of our results should be interpreted as suggesting
-that adults do not change or are incapable of changing. On all
-questions, some members of the sample change their personal culture at
-least a little bit. Our results also should not be taken to imply that
-intrapersonal change is not important or consequential. Small numbers of
-changers can have important consequences in the aggregate. A shift from
-49 percent supporting gay marriage to 51 percent supporting it can have
-significant political ramifications, even as most respondents remain
-fixed in their preferences. And our results should not detract from work
-trying to estimate the mechanisms through which personal cultural
-differences --however small-- emerge among adults. Just because change
-is small does not mean the mechanisms that produce it are not important
-to understand.
-
-Our results should be read principally as suggesting that if researchers
-want to explain why people differ in their personal culture, it is
-imperative to look at the conditions in which they were raised and the
-experiences that happened before they reached adulthood. To the extent
-that we want to explain what makes adults different, focusing attention
-on peoples' transition between social roles, their selection into and
-out of organizational environments, and changes in their social
-networks, as much sociological work does, appears to explain a much
-smaller proportion of their differences than the formative experiences
-that happen early in life.
-
-In this way, our findings echo a range of recent causal inference work
-that suggests selection effects, rather than treatment effects, seem to
-explain the majority of differences in personal culture between people
-in different social roles or positions (Campbell and Horowitz 2016;
-Wodtke 2018). While some people clearly change over time as they move
-into new social roles or into new social environments, this amount of
-change does not appear to be sufficient to explain what are often
-substantial differences between people in different roles. Our results
-suggest that, to the extent we observe differences in personal culture
-between people in different social roles, such as parenthood, having
-children, education, or working in positions of authority (Longest,
-Hitlin, and Vaisey 2013), or between people in different occupations
-(Weeden and Grusky 2005), our prior expectation should be that this
-difference is selection, rather than treatment.
-
-Our results do not speak to exactly why intrapersonal change lacks
-significant explanatory power in understanding culture. First, it could
-be the case that adults are exposed to opportunities for change at the
-same rate, but become less willing to change. Or it could be that
-opportunities for durable change or the social conditions that
-facilitate durable change are just extremely uncommon. Changes in social
-structures, the life course, or cultural expectations could make
-intrapersonal change a much more consequential explanation for cultural
-differences.
-
-Furthermore, the results we present related to education and political
-views suggest that the relative importance of different theoretical
-processes might differ by group. In line with life course perspectives,
-the results presented here suggest that importance of different
-experiences for cultural change are potentially contingent on previous
-experiences. For example, to the extent that we want to explain why
-people have different personal culture, the answer is likely different
-for people who have a college degree and those who do not. Mid-life
-experiences appear to have more significant implications for the
-personal culture for people without a college degree than for people
-with a college degree.
-
-\hypertarget{implications-for-social-change}{%
-\subsection{Implications for Social
-Change}\label{implications-for-social-change}}
-
-Because intrapersonal change in personal culture does not appear to be a
-driving force in cultural differences among adults, our results suggest
-that processes of social and cultural change at the aggregate level are
-likely driven more by cohort replacement, one group of people with
-certain cultural dispositions exiting the population and a different set
-of people with a potentially different set of cultural dispositions
-entering the population, than by contemporaneous social conditions
-(Underwoood et al. 2022; Vaisey and Lizardo 2016). This is likely true
-for both large-scale change in societies over time as well as changes
-within organizations, political parties, occupations, and other social
-aggregates.
-
-However, because our results suggest that the role of formative
-experiences compared to contemporaneous environments differs by a
-person's level of education, changes in educational attainment over
-time, as well as changes in the social structuring of the life course
-might be shifting the relative importance of these two factors over
-time. If we observe further differences across socio-demographic groups
-like education, social class, and race/ethnicity in the relative
-importance of these processes, this would argue for a greater
-incorporation of the interaction of demographic processes with processes
-of individual cultural change into our explanations of aggregate
-cultural change.
-
-\hypertarget{implications-for-survey-measurement}{%
+Sociologists interested in understanding cultural differences have
+largely asked about the presence and significance of cultural change in
+adults (e.g., Kiley and Vaisey 2020; Lersch 2023; Vaisey and Kiley
+2021). Yet doing so has unnecessarily limited and might even set the
+wrong focus. In any population, some degree of cultural change in adults
+is inevitable. Likewise, no viable theoretical perspective would expect
+just one process -- stable differences or internal change -- to shape
+their personal culture. Our results reinforce this point, showing the
+relevance of both interpersonal differences and intrapersonal change in
+understanding the differences in personal culture among adults across
+all items analyzed.
+
+This and other recent findings (Quinn et al. 2023; Stewart and Berkman
+2023) suggest that it is theoretically more productive to ask about the
+relative importance of these two processes in understanding cultural
+variation. Drawing a unified conclusion from survey items measuring
+various cultural forms on different scales and across different time
+frames is challenging. However, the general pattern suggests that, for
+most items, intrapersonal change in adulthood is not the primary reason
+for cultural differences among people.
+
+To clarify, our findings do not imply that adults remain static or that
+their changes are inconsequential. Even minor shifts in personal
+culture, such as a 2 percentage-point change in support for gay
+marriage, can have significant ramifications. While the majority may
+remain consistent in their views, understanding the underlying
+mechanisms of even such minor shifts remains a crucial task for the
+sociology of culture.
+
+Our findings primarily suggest that understanding variations in personal
+culture requires scholars to examine the conditions and experiences of
+early life. While transitions between social roles, changes in social
+networks, and the experience of organizational environments are often
+emphasized in sociological research, these factors seem to account for a
+smaller proportion of differences among adults than the formative
+experiences early in life.
+
+This conclusion aligns with a range of recent causal inference work
+suggesting that selection effects, rather than treatment effects,
+predominantly account for personal cultural differences among
+individuals in varied social roles and positions (Campbell and Horowitz
+2016; Wodtke 2018). While some people clearly change as they transition
+into new roles or environments, this change often seems to be
+insufficient in magnitude and duration to explain what are often
+pronounced differences among people in diverse roles. {[}I'VE TRIED TO
+FIND A WAY TO GET BOTH IMPACT AND DURATION IN THERE; COULDN'T FIND A
+BETTER SOLUTION THOUGH--IDEAS? ALSO WE NEED TO BE CAREFUL WITH THIS
+SENTENCE, IT SEEMS WE ARE OVERSTATING HERE; LET'S FIND A WAY TO
+RELATIVIZE THE CLAIM HERE?{]}. These and our results suggest that when
+observing differences in personal culture across social roles, such as
+parenthood, education, or professional authority (Longest, Hitlin, and
+Vaisey 2013), or across occupations (Weeden and Grusky 2005), selection
+likely plays a large role in explaining these difference, though
+exceptions always exist.
+
+Our analysis does not provide an answer as to why intrapersonal change
+seems to have limited impact on understanding cultural differences among
+adults. The situations that promote durable change in personal culture
+might simply be rare during adulthood. Alternatively, it is possible
+that adults do encounter opportunities, necessities, and incentives for
+change, but their ability or willingness to change decreases. All the
+more it is important to research when and how social situations can
+provoke durable change in adults.
+
+Related to this, our results regarding education and political views
+suggest that the importance of processes that lead to such change can
+vary by group. Aligning with life course theories, it seems the
+significance of experiences for cultural change may be contingent on
+other, prior experiences. For example, factors that shape cultural
+dispositions may likely differ for college and non-college graduates;
+the latter might be more profoundly influenced by mid-life experiences
+than the former in this regard. In trying to understand differences in
+personal culture, sociologists should therefore pay more attention to
+the heterogeneous effects that various factors including social events,
+encounters, and situations can have.
+
+Our findings also have implications for understanding cultural change at
+the aggregate level. Given that intrapersonal change does not appear as
+the primary driver for many cultural differences among adults, our
+results suggest that many forms of cultural change at the aggregate
+level are more likely driven by processes of cohort replacement than the
+by contemporaneous social conditions changing people's cultural
+dispositions (Underwoood et al. 2022; Vaisey and Lizardo 2016). This
+likely holds true for cultural change at the macro societal level as
+well as the micro level such as within organizations, political parties,
+and professions. Our findings also indicate that the impact of formative
+experiences, as opposed to contemporaneous social factors, on cultural
+change is contingent upon an individuals' level of education. The weight
+of these two factors might thus change as the prominence of education
+shifts in the life course. If scholars find similar differences across
+other social categorizations like class or race, it would advocate for a
+more thorough integration of demographic processes into the analysis of
+cultural change at the aggregate level.
+
+\hypertarget{implications-for-survey-research}{%
 \subsection{Implications for Survey
-Measurement}\label{implications-for-survey-measurement}}
-
-A major takeaway from the meta-analysis is that greater response
-resolution ---the number of response options survey respondents are
-given-\/-- is associated with a greater share of variance being
-explained by within-individual change. Again, it is possible that this
-reflects different kinds of issues being measured with different kinds
-of scales. But it could also suggest that within-person change, when it
-happens, is relatively small ---moving from ``agree'' to ``strongly
-agree,'' rather than ``strongly agree'' to ``strongly disagree,'' for
-example-\/-- and therefore easier to detect with greater resolution.
-
-Our results continue to highlight the value of long-running panel
-surveys for adjudicating different theories of culture that cannot be
-evaluated with cross-sectional data alone. Across our results, residual
-fluctuations in responses account for more than four times as much
-variance, on average, as linear persistent change, meaning that most
-change in responses is not systematic. Under these circumstances,
-extending panels beyond two waves provides significantly more leverage
-to understand how and why personal culture changes.
-
-Additionally, we echo previous calls for more studies ---including panel
-studies-\/-- of the socialization processes of young people. Our results
-suggest that most of the variance between adults emerges before people
-turn 18 or relatively soon after. In limiting our empirical perspective
-to experiences that happen after that point, social scientists
-potentially focus on questions that, while important, lack the capacity
-to explain major cultural differences.
+Research}\label{implications-for-survey-research}}
+
+Our findings underscore the value of extended panel surveys to advance
+theories of culture. On average, non-systematic fluctuations in
+responses account for more than four times as much variance as
+intrapersonal change, measured as linear change. Differentiating between
+the two is therefore crucial for understanding cultural differences and
+cannot be done with cross-sectional data. {[}MAYBE THE COMPARISON
+BETWEEN NON-SYSTEMATIC VARIATION AND SYSTEMATIC (LINEAR+BASELINE
+DIFFERENCES) WOULD EVEN BE BETTER TO MAKE THAT POINT?{]}. Instead we
+should extend panels beyond two waves to gain more leverage to
+understand when and in what form personal culture changes.
+
+Further, we align with others {[}CITES?{]} advocating for more studies
+-- including panel studies -- focusing on the socialization of youth.
+Our analysis points to the majority of variance among adults rooting in
+differences encountered before turning 18. By empirically constraining
+their analysis to experiences after that point, social scientists may
+inadvertently concentrate on topics and questions that, while important,
+might not be able to help explain major cultural differences in adult
+populations.
+
+Finally, a key insight from our meta-analysis is that an increase in
+response resolution -- the range of response options survey respondents
+are given -- correlates with an increased share of systematic variance
+explained by intrapersonal change. This may reflect specific issues
+being assessed by survey items using different scales. Alternatively, it
+might imply that when intrapersonal changes occurs, they are often
+subtle with answers moving from ``agree'' to ``strongly agree,'' rather
+than from ``strongly agree'' to ``strongly disagree.'' Such subtle
+changes require finer response options to be detected.
 
@@
 
-The approach outlined here generates estimates of how much variance in a
-population is due to each of two broad processes: intrapersonal change
-and interpersonal differences at baseline. It does not quantify the
-proportion of people who ``change.'' When we observe intrapersonal
-change in a question, we cannot be sure whether this is driven by many
-people making small changes or a few people making large changes.
-Adjudicating this question will require different models that
-distinguish different kinds of change.
-
-Our results and interpretations also hinge on how we have defined change
-and the assumptions built into the Life Course Adaption Model. This
-model treats cultural trajectories as varying linear slopes for each
-respondent, assuming that change is a linear function of time. This
-assumption likely does not hold true in practice. Change over time
-likely takes non-linear and discontinuous forms. People might jump from
-one ``stable'' disposition to another, for example, or experience a
-discontinuous ``turning point'' that upsets a stable trajectory. This
-assumption is in part a limitation of the data, as most questions are
-only observed for three to six waves with a large number of missing
-cases. Panels with more waves might allow for more flexible models of
-change that do not require this assumption of linearity.
-
-Similarly, in assuming a linear trajectory for change, our approach
-assumes that durable change is unidirectional. To the extent that people
-make durable change as a result of a social transition and then return
-to a previous position later in life as a result of a second transition
-(marriage and divorce, for example), this is captured as residual
-within-person variance, rather than one of our two broad theoretical
-processes. This is a potentially reasonable assumption on short panels
---change that does not last two years is hard to classify as durable--
-but it might not be true in longer panels. Other researchers with more
-flexible definitions of ``durable change'' might potentially identify
-more durable change even in the same population.
-
-Finally, while we examine an incredible broad array of measures of
-personal culture, from assessments about the relative importance of
-different features when buying a new car and policy preferences to
-religious beliefs and core values, we are limited by the kinds of
-questions that are asked in panel surveys and the contexts in which
-these surveys were administered (Australia, Germany, Switzerland, the
-United Kingdom and the United States). This means our findings tend to
-reflect questions about general (national) politics, gender roles,
-immigration and race relations, and general well being. Our data does
-not cover some dimensions of culture that might be relevant for other
-aspects of sociological exploration (artistic tastes, leisure
-activities, and time use, for example).
+The approach outlined here generates estimates of how much of the
+observed variance in a population can be attributed to one of two sets
+of theoretical processes: intrapersonal change and interpersonal
+differences at baseline. It does not quantify the proportion of people
+who ``change,'' nor can we be sure that the amount of interpersonal
+change we detect is driven by many people making small changes or a few
+people making large changes.
+
+Our results and interpretations also hinge on how we have defined
+change. Similarly to the LCAM, our approach treats cultural trajectories
+as varying linear slopes for each respondent, thus assuming that change
+is a linear function of time. This assumption simplifies reality in
+which change likely also takes non-linear and discontinuous forms.
+People might jump from one ``stable'' disposition to another or
+experience a ``turning point'' in their lives that upsets an otherwise
+stable trajectory. In parts, our assumption of linearity is a limitation
+of the data, as most questions are only observed for three to six waves.
+Panels with more waves might allow researchers to loosen this assumption
+to test alternative, more flexible models of change.
+
+Similarly, our approach assumes that durable change is unidirectional.
+This is a sensible assumption on short panels where classifying change
+that lasts less than two years as durable seems unreasonable.
+Practically, it means that the variance produced by people making
+durable changes in their cultural dispositions to then returning to a
+previous state later in life is classified as residual variance, rather
+than intrapersonal change. Longer panels and more flexible definitions
+of ``change'' would allow us to account for such trajectories and thus
+potentially also identify more durable change even in the same
+population.
+
+Finally, we had chosen to examine the broadest array of measures of
+personal culture available to us, ranging from religious beliefs and
+core values, to policy preferences, and even to the importance of
+different features when buying a new car across five countries
+(Australia, Germany, Switzerland, the United Kingdom and the United
+States). Nevertheless, our findings remain limited to the kinds of
+questions that are asked in panel surveys and in the contexts they were
+administered, reflecting issues of general (national) politics, gender
+roles, immigration and race relations, and general well being. Although
+we have no reason to believe results to be different, our findings do
+not directly speak to other dimensions of culture such as artistic
+tastes, leisure activities, and time use.
 
@@
 which the relative importance of intrapersonal change and interpersonal
-differences matter for explaining cultural differences. Characterizing
+differences in explaining cultural differences shifts. Characterizing
 questions as displaying change or not can only take researchers so far,
-but the question of why some questions demonstrate more change while
-others do not, or why some groups are characterized by more stability
-than others, has the potential to weigh in on a broader range of
-theoretical debates. We hope researchers move toward those questions.
+but the question of whether some questions demonstrate more change than
+others, or whether some groups are characterized by more stability than
+others, has the potential to weigh in on a broader range of theoretical
+debates. We hope researchers move toward those questions.
 
@@
 
+\hypertarget{endnotes}{%
+\section{Endnotes}\label{endnotes}}
+
 \hypertarget{references}{%
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-bardi2011}{}}%
-Bardi, Anat, and Robin Goodwin. 2011. {``The {Dual} {Route} to {Value}
-{Change}: {Individual} {Processes} and {Cultural} {Moderators}.''}
-\emph{Journal of Cross-Cultural Psychology} 42(2):271--87. doi:
-\href{https://doi.org/10.1177/0022022110396916}{10.1177/0022022110396916}.
+\leavevmode\vadjust pre{\hypertarget{ref-bardi2009}{}}%
+Bardi, Anat, Julie Anne Lee, Nadi Hofmann-Towfigh, and Geoffrey Soutar.
+2009. {``The Structure of Intraindividual Value Change.''} \emph{Journal
+of Personality and Social Psychology} 97(5):913--29.
 
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-elder2016}{}}%
-Elder, Glen H., and Linda K. George. 2016.
-{``\href{https://doi.org/10.1007/978-3-319-20880-0_3}{Age, {Cohorts},
-and the {Life} {Course}}.''} Pp. 59--85 in \emph{Handbook of the {Life}
-{Course}: {Volume} {II}}, \emph{Handbooks of {Sociology} and {Social}
-{Research}}, edited by M. J. Shanahan, J. T. Mortimer, and M.
-Kirkpatrick Johnson. Cham: Springer International Publishing.
+\leavevmode\vadjust pre{\hypertarget{ref-elder2003}{}}%
+Elder, Glen H., Monica Kirkpatrick Johnson, and Robert Crosnoe. 2003.
+{``\href{https://doi.org/10.1007/978-0-306-48247-2_1}{The {Emergence}
+and {Development} of {Life} {Course} {Theory}}.''} Pp. 3--19 in
+\emph{Handbook of the {Life} {Course}}, \emph{Handbooks of {Sociology}
+and {Social} {Research}}, edited by J. T. Mortimer and M. J. Shanahan.
+Boston, MA: Springer US.
 
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-paul2014transformative}{}}%
-Paul, Laurie Ann. 2014. \emph{Transformative Experience}. OUP Oxford.
+\leavevmode\vadjust pre{\hypertarget{ref-quinn2023}{}}%
+Quinn, Joseph M., Robert E. Freeland, Kimberly B. Rogers, Jesse Hoey,
+and Lynn Smith-Lovin. 2023. {``How Cultural Meanings of Occupations in
+the u.s. Changed During the Covid-19 Pandemic.''} \emph{American
+Behavioral Scientist} 67(1):125--47.
 
@@
 doi: \href{https://doi.org/10.2307/2090964}{10.2307/2090964}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-scoville2022}{}}%
+Scoville, Caleb, Andrew McCumber, Razvan Amironesei, and June Jeon.
+2022. {``Mask {Refusal} {Backlash}: {The} {Politicization} of {Face}
+{Masks} in the {American} {Public} {Sphere} During the {Early} {Stages}
+of the {COVID}-19 {Pandemic}.''} \emph{Socius: Sociological Research for
+a Dynamic World} 8:22.
 
@@
 Gordon; Breach Sciences Publishers.
+
+\leavevmode\vadjust pre{\hypertarget{ref-stewart2023}{}}%
+Stewart, Evan, and Diane Berkman. 2023. {``Shifting or Settled? Tracking
+Racial Animus During COVID-19.''} \emph{Social Psychology Quarterly}
+86(3):312--33.
 
@@
 
-\begin{figure}
-
-{\centering \includegraphics[width=20.71in]{fig/variation_share_lm_mdpnt}
-
-}
-
-\caption{Proportions of explained variance in items of personal culture; based on predictions at participants' wave mid-point.}\label{fig:fig1}
+\begin{figure}[ht]
+\caption{Proportions of Systematic Variance in Personal Culture}
+
+\begin{center}\includegraphics[width=20.83in]{fig/variation_share_lm_mdpnt} \end{center}
+
+\textit{Notes:} The figure shows $\tau$ and 1 - $\tau$ as the proportion of systematic variance attributable to intrapersonal change and interpersonal differences, respectively.
+See Appendix A for the full set of item values.
 \end{figure}
@@
 \newpage
-\begin{figure}
-
-{\centering \includegraphics[width=300px]{fig/meta_no_int}
-
-}
-
-\caption{Proportion of explained variance attributable to intrapersonal change; based on predictions at participants' wave mid-point; supressed intercept model; survey indicators not shown.}\label{fig:fig2}
+
+\begin{figure}[ht]
+\caption{Regression Model Estimating $\tau$}
+
+\begin{center}\includegraphics[width=300px]{fig/meta_no_int} \end{center}
+
+\textit{Notes:} The model estimates the proportion of systematic variance attributable to intrapersonal change.
+It includes the log of the number of participants, the number of waves the variable was asked (ceiled at t = 10 for each item), the date of the first wave (measured as the year minus 1968 divided by 10), number of waves the question was asked averaged across participants, duration in years per item averaged across participants, and panel and topic fixed effects.
+Coefficients are estimated from suppressed intercept model based on predictions at participants' wave mid-point.
+Survey indicators and item topics not shown.
+See Appendix B for the full set of coefficients.
 \end{figure}
@@
 \newpage
-\begin{figure}
-
-{\centering \includegraphics[width=450px]{fig/FIG_college1}
-
-}
-
-\caption{Figure 3: Distribution of differences share of variance explained by intrapersonal change between respondents without and with college degree.}\label{fig:fig3}
+
+\begin{figure}[ht]
+\caption{Difference in $\tau$ Across College Graduates and Non-Graduates}
+
+\begin{center}\includegraphics[width=450px]{fig/FIG_college1} \end{center}
+
+\textit{Notes:} The figure shows the difference in $\tau$ values across college graduates and college non-graduates.
+Values above (below) 0 means that those with college degree have higher (lower) variance of intrapersonal change.
+The dashed red line marks 0 difference.
 \end{figure}
@@
 \newpage
-\begin{figure}
-
-{\centering \includegraphics[width=450px]{fig/FIG_college2}
-
-}
-
-\caption{Proportions of intrapersonal change of explained variance by college degree; based on predictions at participants' wave mid-point.}\label{fig:fig4}
+
+\begin{figure}[ht]
+\caption{Difference in $\tau$ Across College Graduates and Non-Graduates on Political Culture}
+
+\begin{center}\includegraphics[width=450px]{fig/FIG_college2} \end{center}
+
+\textit{Notes:} $\tau$ values across college graduates and college non-graduates on 8 political culture items from the General Social Survey (2006-2014).
 \end{figure}

### 2023-09-27 manuscript/manuscript.Rmd
cleaning up and updates

--- 2023-09-27 manuscript.Rmd
@@
 
-A recurring focus in the sociology of culture is the question of whether people, as they move through their lives and transition between social roles, change their personal culture –- the declarative and non-declarative attitudes, beliefs, values, practices, and dispositions that exist at the individual level [@lizardo2017; @kiley2020; @lersch2023].
+Does personal culture --an individual's attitudes, beliefs, values, and practices--- change over the life course, or is it largely fixed once a person reaches adulthood?
 
-This seemingly simple question underlies a number of debates across sociological subfields.
+This question underlies several contemporary debates in sociology [@lizardo2017; @kiley2020; @lersch2023].
 For example, pragmatist theories of action claim that changes in social environments cause people to adapt their views and make new cultural meanings [@swidler2001; @gross2009], while Bourdieusian practice theories argue that the “past conditions of production” leave a mark on people's personal culture that lasts throughout their lives [@bourdieu1990].
-Similarly, models of social influence assume that people adapt their culture in the face of new information [@vaisey2010a KK Note: I don't feel like this is a good cite for this.], while the emphasis on cohort effects in models of aggregate social and cultural change requires them to be open to change while young but become fairly resistant to it as they age [@ryder1965].
+Similarly, models of social influence assume that people adapt their culture in the face of new information, while the emphasis on cohort effects in models of aggregate social and cultural change requires them to be open to change while young but become fairly resistant to it as they age [@ryder1965].
 Finally, life course theories posit slower but still important changes over time as people advance through important transitions in their lives [@elder2003; @bardi2009].
 
-Because all of the above theoretical perspectives have some empirical support, the current debate revolves around not whether they exist, but when and how they each apply and the contribution of each to explaining cultural differences.
-That being said, scholars have struggled to reach a consensus on the relative importance of change during adulthood.
-Over the span of a few years, most apparent changes in people’s responses to personal culture items appear to be transitory, with little evidence of persistent change among adults [@kiley2020; @vaisey2021].
-This suggests that change during adulthood is not a strong force in explaining personal cultural differences.
-When looked at over longer time horizons, however, there is at least some evidence that adults make persistent cultural changes, especially as they move through distinct social roles [@lersch2023].
+Because all of the above theoretical perspectives have some empirical support, the current debate is not about whether they exist, but about their contribution to explaining cultural differences.
+That said, scholars have struggled to reach a consensus on the relative importance of change during adulthood.
+Over the span of a few years, most apparent changes in people’s responses to personal survey items appear to be transitory, with little evidence of persistent change among adults [@kiley2020; @vaisey2021].
+This suggests that change during adulthood is not a major factor in explaining personal cultural differences.
+When looked at over longer time horizons, however, there is at least some evidence that adults make persistent cultural changes [@lersch2023].
 
@@
 228].
-Accordingly, researchers have formalized models of responses to cultural items over time as demonstrating either “change” or “no change,” which essentially asks whether we *ever* observe intrapersonal change in personal culture.
+That is, researchers have compared models of responses to cultural items over time as consistent with either “change” or “no change,” which essentially asks whether we observe *any* intrapersonal change in personal culture.
 
 We argue that asking only *whether* people change obscures what are actually highly compatible empirical patterns.
-Being able to detect change in a given dataset depends not only on sample populations, item measurement, statistical methods, and power, but crucially also on definitions of what counts as “change.” These methodological decisions potentially limit the possibility of convergence on the theoretical debates.
-Moreover, asking whether there is evidence of change limits the kind of questions that can be asked to, “Do we see change in this measure?” But such inquiries are ill-suited to address what’s at the core of the theoretical debate.
-There is enough evidence to assume that neither socialization at an early age nor the influence of the contemporary environment alone can explain cultural differences.
-The question that remains is under what conditions, across which groups, and in what domains do we observe differences in the relative importance of intrapersonal change and interpersonal differences.
-This focus is necessarily comparative and requires an approach that quantifies forms of cultural change rather than calling victory for the one or the other perspective.
+Being able to detect change in a given dataset depends not only on sample populations, item measurement, statistical methods, and power, but also on definitions of what counts as “change.” These methodological decisions potentially limit the possibility of convergence on the theoretical debates.
+Moreover, asking whether there is evidence of change limits the kind of questions that can be asked to, “Do we see change in this measure?” But such inquiries are ill-suited to address the core of the theoretical debate: determining *how much* the process of intrapersonal change contributes to cultural differences.
 
-In this paper, we try to intervene in this debate by moving beyond the “tournament of models” approach toward a quantification of the processes that lead to differences in personal culture.
+There is enough evidence to assert that neither pre-adult socialization nor the contemporary environment alone can fully explain cultural differences.
+The question that remains is about the relative contributions of these processes.
+In the current debate, for a single construct (e.g., support for gay rights), existing approaches can only make claims about *whether* intrapersonal change is happening, not about how much.
+This limitation also prevents investigating under what conditions, across which groups, and in what domains do we see differences in the relative importance of intrapersonal change.
+These questions are about *degree*, not existence, and thus require precise measurement rather than declarations of victory for a particular perspective.
+
+In this paper, we intervene in this debate by moving beyond the “tournament of models” approach toward a quantification of the processes that lead to differences in personal culture, and introducing a method for quantifying the relative contributions of interpersonal differences and intrapersonal change to the patterns we see in empirical data.
 Drawing on the seven panel datasets analyzed in the debate, we use established methods to quantify the amount of variance in responses that is due to between-people differences at a single point during the survey versus the amount of variance in responses that is due to linear intrapersonal change over time.
@@
 In other words, despite different interpretations, the results of previous work are in many ways highly consonant.
-[I think the revisions that were done to this paragraph were confusing and mostly reverted it.]
 
@@
 But rather than debating whether this model fits a particular data set better than a model that assumes that people never change, researchers should quantify individual changes over time and compare those to the differences that exist between people at the panel's start.
-In other words, the focus should shift from asking, “do people change?” to asking, “what are the relative contributions of change and initial differences toward explaining differences in personal culture?” [KK Note: I didn't like this change.
-It's not about collectively explaining, its about them relative to each other.]
+In other words, the focus should shift from asking, “do people change?” to asking, “what are the relative contributions of change and initial differences toward explaining differences in personal culture?”
 
@@
 Consequently, this metric also reflects variation in individuals' social roles or statuses at the start of the panel that were important in shaping their dispositions.
-[TK Note: I feel that this paragraph puts too much emphasis on pre-panel processes.
-It might be early imprinting or broad life-course, but why emphasize it this much?
-KK NOTE: Because it could be either, I don't want to treat it like it's just "imprinting."]
-[KK NOTE: You reversed what was being said here.
-You say "early imprinting" denotes intrapersonal change (process denotes number).
-We want to say the opposite -- what I had before -- number denotes process: "interpersonal differences" denotes "early imprinting." I changed it back.
-I also took out reference to "transformative" experiences because I don't think we're necessarily talking about those specifically, just any experience that changes you.
-Because we don't say a specific theoretical process ("socialization", etc.) I want to keep the language of each denoting combinations of processes.
-"Persistent change" isn't a theoretical process.
-It's a bunch of them.]
 
@@
 Because of this we quantify this component but focus principally on the other two.
-[KK NOTE: I agreed with TK's note and mostly reverted this paragraph.]
 
@@
 
-Third, we expect the variance explained by each process –interpersonal differences and intrapersonal change– to depend on specific attributes of the question and panel.
+Third, we expect the variance explained by each process -–interpersonal differences and intrapersonal change-– to depend on specific attributes of the question and panel.
 First, the higher the response resolution for a given question and the larger the sample, the greater the variance explained by intrapersonal change should be, as smaller changes in intrapersonal change will be easier to detect.
@@
 In the end, the analyses that follow rely on a cross-cultural sample with a cross-domain set of items to capture a broad range of individual personal culture.
-Appendix A documents the list of all variables used in the upcoming analyses.
+Supplemental Materials A documents the list of all variables used in the upcoming analyses.
 
@@
 $$
-u_{0,i} \sim \mathcal{N}(0,\sigma_{\mu_0}^2)
+u_{0,i} \sim \mathcal{N}(0,\sigma_{\mu_0}^2) \\
 $$
-
 $$
 u_{1,i} \sim \mathcal{N}(0,\sigma_{u_1}^2)
-
 $$
-[KKNOTE: Is that right? I forget how to do that exactly.]
 
@@
 $$
-\tau = \frac{V(C)}{V(C) + V(D)}
+\omega = \frac{V(C)}{V(C) + V(D)}
 $$
 
-[TK Note: \\phi taken in Kiley & Vaisey, \\delta taken in Vaisey & Lizardo.
-I think using a Greek letter without strong associations (something like alpha, beta, or mu or whatever) is not good.
-So I settled on \\tau.
-Feel free to change it.]
-
-We can think of this number, $\tau$ as akin to an Intraclass Correlation Coefficient focused only on the systematic components of the model while ignoring residual variance that might be attributable to measurement error or non-persisting change [see also the variance decomposition of period and cohort by @vaisey2016].
+We can think of this number, $\omega$ as akin to an Intraclass Correlation Coefficient focused only on the systematic components of the model while ignoring residual variance that might be attributable to measurement error or non-persisting change [see also the variance decomposition of period and cohort by @vaisey2016].
 The higher this number is, the more systematic variance in responses is attributable to intrapersonal change during the panel.
 The lower the number, the less intrapersonal change seems to play a large role in explaining systematic cultural differences during the panel.
-Moving forward, we use $\tau$ as a placeholder for *systematic variance attributable to intrapersonal change*.
+Moving forward, we use $\omega$ as a placeholder for *systematic variance attributable to intrapersonal change*.
 
@@
 
-As a second step in our analysis, we use ordinary least squares regression to model $\tau$ as a function of features of the question and panel.
+As a second step in our analysis, we use ordinary least squares regression to model $\omega$ as a function of features of the question and panel.
 This allows us to explore how these features affect the prevalence of intrapersonal change.
 These covariates include the number of response options (measured as 2, 3-5, and more than 5), the panel, the log of the number of participants, the number of waves observed, the time since the question was first asked, and the total span of time the question was asked.
-
-[TK Note: I guess we need to say something here about the college stuff?
-KK Note: I didn't feel like that was necessary.
-It feels so distinct later on that I think it's OK to kind of "start over" there.
-I'm worried that people will forget the plan by the time they get there, and we'll have to reiterate everything anyway.
-But others should weigh in.].
 
@@
 To broadly summarize the results, all panels show a similar range of the proportion of systematic variance attributable to intrapersonal change.
-Across panels, the median proportion of $\tau$ is 0.094 (mean of 0.098), with an interquartile range of 0.053 to 0.139.
-All panels include questions where $\tau$ is essentially 0, and all panels include questions with $\tau$ values greater than 0.20.
-The maximum $\tau$ value is 0.26.
+Across panels, the median proportion of $\omega$ is 0.094 (mean of 0.098), with an interquartile range of 0.053 to 0.139.
+All panels include questions where $\omega$ is essentially 0, and all panels include questions with $\omega$ values greater than 0.20.
+The maximum $\omega$ value is 0.26.
 
@@
 <about here>
+Figure 1: Proportions of Systematic Variance in Personal Culture
 \end{center}
-
-Figure 1: Proportions of explained variance in items of personal culture; based on predictions at participants' wave mid-point.
 
@@
 
-To the extent that there are differences across the panels, the PSID has the highest $\tau$ values with mean .120 and median .135.
+To the extent that there are differences across the panels, the PSID has the highest $\omega$ values with mean .120 and median .135.
 While we cannot disentangle features of the sample from features of the questions asked to each sample, the specific samples for many PSID questions have lower average ages than those from other panels.
 To the extent that younger respondents might be more likely to make durable changes of opinion, this higher explanatory power of intrapersonal change might reflect the distinct age profile of respondents in this sample.
-At the other end, the GSS has the lowest range of $\tau$ with mean .073 and median .069.
+At the other end, the GSS has the lowest range of $\omega$ with mean .073 and median .069.
 This potentially reflects the fact that the GSS observes people for a shorter duration, on average, than the other panels.
-If, consistent with life course adaption theories, people are more likely to make significant cultural changes the longer we observe them, then duration likely affects the range of $\tau$ (a point we explore more below).
+If, consistent with life course adaption theories, people are more likely to make significant cultural changes the longer we observe them, then duration likely affects the range of $\omega$ (a point we explore more below).
 However, the GSS results are still consistent with results from the other panels.
@@
 While there are some differences between panels, these differences are small compared to the differences within panels.
-For about 6 percent of items explored here,  $\tau$ is greater than 0.20.
+For about 6 percent of items explored here,  $\omega$ is greater than 0.20.
 These questions tend to ask about external referents (e.g., confidence in specific government leaders or political parties), life satisfaction, or current financial position.
@@
 
-Appendix A presents the estimated proportion of variance attributable to interpersonal differences and intrapersonal change, the estimated values of $\tau$, and the proportion of residual variance for each question analyzed.
+Supplemental Materials A presents the estimated proportion of variance attributable to interpersonal differences and intrapersonal change, the estimated values of $\omega$, and the proportion of residual variance for each question analyzed.
 In general, interpersonal differences are almost always the largest component of the total variance and tend to explain between 55 and 70 percent of total variance, while intrapersonal change is always the smallest, typically explaining between 3 and 8 percent of total variance.
@@
 
-Figure 2 plots the results from an ordinary least squares regression of $\tau$ as a function of question, panel, and sample features.
-These models also include fixed effects for panels, so coefficients reflect the association between a particular sample or question characteristic within a panel.[KK NOTE: CAN WE SUMMARIZE THESE PANEL FIXED EFFECTS AT ALL?
-AE: WE COULD SHOW THAT THEY DON'T CAPTURE TOPIC EFFECTS BEYOND OUR CLASSIFICATION IF WE INCLUDE THOSE / THAT THEY POSSIBLY CAPTURE UNMEASURED COUNTRY/CONTEXT DIFFERENCES OR DIFFERENCES IN SURVEY METHODOLOGY?]
+Figure 2 plots the results from an ordinary least squares regression of $\omega$ as a function of question, panel, and sample features.
+These models also include fixed effects for panels and topics, so coefficients reflect the association within a panel and topical domain.
 
@@
 <about here>
+Figure 2: Regression Model Estimating $\omega$
 \end{center}
-Figure 2: Proportion of explained variance attributable to intrapersonal change; based on predictions at participants' wave mid-point; supressed intercept model; survey indicators not shown.
 
-Figure 2 shows that the more response options respondents are given and the larger the sample, the larger the values of $\tau$.
+Figure 2 shows that the more response options respondents are given and the larger the sample, the larger the values of $\omega$.
 We interpret these coefficients as suggesting greater resolution on a question makes it easier to detect and model change.
-The earlier a question was asked, even net of how long it was been asked, the lower the $\tau$.
+The earlier a question was asked, even net of how long it was been asked, the lower the $\omega$.
+
+While we did not generate expectations for how question content relates to $\omega$, some of the associations between question structure, panel duration, and panels themselves might be driven by differences in the topics addressed by each panel.
+To address this, we coded each question as falling into one of nine different content categories and included these as indicator variables in the meta analysis.
+Supplemental Materials B presents the estimates for these content indicators.
+They show some associations with $\omega$, with questions about subjective SES; social life, social cohesion, and trust; environment and climate; and health and morale showing larger $\omega$ values than questions about religion and spirituality; politics, government and the economy; and gender and family.
 
 The most notable result in the meta analysis is the negative coefficient attached to the duration of years covered by the question.
-The longer a question is observed, the less is the value of $\tau$.
+The longer a question is observed, the less is the value of $\omega$.
 Theories of personal cultural change that link changes in personal culture to social experiences, including the LCAM, suggest that the longer we observe respondents, the more likely people are to undergo potentially transformative experiences and therefore the more variance would be explained by intrapersonal change.
-The fact that we find a negative coefficient attached to this variable can be a challenge to that assumption [TK Note: since there is a complication of this, I changed is to can be, but feel free to ignore it.]
+The fact that we find a negative coefficient attached to this variable can be a challenge to that assumption.
 
-As a further test of this coefficient, we compared the $\tau$ values when using the full duration of a panel compared to when we dropped the final wave and therefore reduced the total duration of observation for the question.
-If the coefficient reflects a true negative effect of duration on $\tau$, we should see that same effect within questions.
-Results from this analysis are presented in Appendix B and contradict the coefficient from the regression model: we found that the longer we observe the same question, the higher the $\tau$ is.
+As a further test of this coefficient, we compared the $\omega$ values when using the full duration of a panel compared to when we dropped the final wave and therefore reduced the total duration of observation for the question.
+If the coefficient reflects a true negative effect of duration on $\omega$, we should see that same effect within questions.
+Results from this analysis are presented in Supplemental Materials C and contradict the coefficient from the regression model: we found that the longer we observe the same question, the higher the $\omega$ is.
 We interpret this combination of findings as suggesting that the kinds of questions asked for a longer time period tend to demonstrate less intrapersonal change than questions asked over shorter periods, rather than a true function of time.
-
 This reinforces the explanation for why the GSS displays the lowest average proportion of systematic variance attributable to intrapersonal change: its shorter duration.
-
-While we did not generate expectations for how question content relates to $\tau$, some of the associations between question structure, panel duration, and panels themselves might be driven by differences in the topics addressed by each panel.
-To address this, we coded each question as falling into one of nine different content categories and included these as indicator variables in the meta analysis.
-Appendix C presents the estimates for these content indicators.
-They show some associations with $\tau$, with questions about subjective SES; social life, social cohesion, and trust; environment and climate; and health and morale showing larger $\tau$ values than questions about religion and spirituality; politics, government and the economy; and gender and family.
-[KKNOTE: I FEEL LIKE WE WANTED TO SAY SOMETHING ELSE HERE BUT I CANT REMEMBER WHAT]
 
@@
 
-To test these competing propositions, we use the models presented above to separately calculate $\tau$ values for people with at least a bachelor's degree and people with less than a bachelor's degree at wave 1[^wave] of the General Social Survey's panels.[^missing] We focus on the GSS because it contains the largest number of questions tapping general political dispositions, which is the domain where education has proven particularly relevant for understanding attitude stability.
+To test these competing propositions, we use the models presented above to separately calculate $\omega$ values for people with at least a bachelor's degree and people with less than a bachelor's degree at wave 1[^wave] of the General Social Survey's panels.
+We focus on the GSS because it contains the largest number of questions tapping general political dispositions, which is the domain where education has proven particularly relevant for understanding attitude stability.
 The GSS also covers a turbulent window of American politics from 2006 to 2014.
@@
 
-Figure 3 plots the distribution of differences in $\tau$ values between people with at least a bachelor's degree and people with less than a bachelor's degree at wave 1 of the panel for all 183 GSS items we explored.
+Figure 3 plots the distribution of differences in $\omega$ values between people with at least a bachelor's degree and people with less than a bachelor's degree at wave 1 of the panel for all 183 GSS items we explored.
 Values greater than 0 indicate that intrapersonal change is more important in explaining the responses of college graduates than those of people without a college degree, while values less than 0 indicate that intrapersonal change is less important for explaining variation among people with a college degree than those without a college degree.
@@
 <about here>
+Figure 3: Difference in $\omega$ Across College Graduates and Non-Graduates
 \end{center}
-Figure 3: Difference between share of intrapersonal change and share of inter-personal differences in explained variance; based on predictions at participants' wave mid-point
 
@@
 conservative) on seven-point scales; four questions about the government's role in improving the condition of the poor, paying people's medical bills, giving special treatment to black people, and doing things that private businesses could do, measured on five-point scales; a question about whether the government should do more to reduce income differences, measured on a seven-point scale; and one question about whether black people should be given preferences in hiring, measured on a five-point scale.
-Estimates of $\tau$ for these eight questions, for both education groups, are presented in Figure 4.
+Estimates of $\omega$ for these eight questions, for both education groups, are presented in Figure 4.
 
@@
 <about here>
+Figure 4: Difference in $\omega$ Across College Graduates and Non-Graduates on Political Culture
 \end{center}
-Figure 4: Proportions of intrapersonal change of explained variance by college degree; based on predictions at participants' wave mid-point.
 
-On all eight questions presented in Figure 4, $\tau$ values are greater for people without a college degree than for people with a college degree, meaning intrapersonal change explains less systematic variance for people with a college degree than for people without one.
-There are also large differences in $\tau$ values across questions for both groups.
+On all eight questions presented in Figure 4, $\omega$ values are greater for people without a college degree than for people with a college degree, meaning intrapersonal change explains less systematic variance for people with a college degree than for people without one.
+There are also large differences in $\omega$ values across questions for both groups.
 For example, on the question of whether the government should try to solve more problems or leave those problems to be solved by private businesses ("government do more or less"), less than one percent of the systematic variance is attributable to intrapersonal change for both groups.
@@
 
-In contrast, partisan identification and political ideology both show larger values of $\tau$ than most other questions, as well as a larger absolute difference between people with a college degree and those without.
+In contrast, partisan identification and political ideology both show larger values of $\omega$ than most other questions, as well as a larger absolute difference between people with a college degree and those without.
 Compared to the other questions explored here, change as adults plays a much larger role in explaining why people report the partisan identification and political ideology they do.
 And this is particularly true for respondents who do not have a college degree.
-The $\tau$ value for ideological identification among non-college educated respondents is almost four times that of college-educated respondents.
+The $\omega$ value for ideological identification among non-college educated respondents is almost four times that of college-educated respondents.
 
-We note here that this meaningful difference in $\tau$ values across education groups and across questions would not be detectable using previous methods.
+We note here that this meaningful difference in $\omega$ values across education groups and across questions would not be detectable using previous methods.
 For partisan identification and political ideology, both college-educated respondents and people with less than college degree would likely favor the AUM or LCAM over the SDM because they these questions both show evidence of some members of the population making intrapersonal change.
@@
 
-Applying our proposed measure, $\tau$, to all questions about personal culture from the panel datasets previously discussed in this debate revealed a consistent pattern.
+Applying our proposed measure, $\omega$, to all questions about personal culture from the panel datasets previously discussed in this debate revealed a consistent pattern.
 Nearly all questions show evidence of people making durable, intrapersonal change over time, with some showing notably high intrapersonal change compared to interpersonal differences.
@@
 While some people clearly change as they transition into new roles or environments, this change often seems to be insufficient in magnitude and duration to explain what are often pronounced differences among people in diverse roles.
-[I'VE TRIED TO FIND A WAY TO GET BOTH IMPACT AND DURATION IN THERE; COULDN'T FIND A BETTER SOLUTION THOUGH--IDEAS?
-ALSO WE NEED TO BE CAREFUL WITH THIS SENTENCE, IT SEEMS WE ARE OVERSTATING HERE; LET'S FIND A WAY TO RELATIVIZE THE CLAIM HERE?].
 These and our results suggest that when observing differences in personal culture across social roles, such as parenthood, education, or professional authority [@longest2013], or across occupations [@weeden2005], selection likely plays a large role in explaining these difference, though exceptions always exist.
@@
 Differentiating between the two is therefore crucial for understanding cultural differences and cannot be done with cross-sectional data.
-[MAYBE THE COMPARISON BETWEEN NON-SYSTEMATIC VARIATION AND SYSTEMATIC (LINEAR+BASELINE DIFFERENCES) WOULD EVEN BE BETTER TO MAKE THAT POINT?].
 Instead we should extend panels beyond two waves to gain more leverage to understand when and in what form personal culture changes.
 
-Further, we align with others [CITES?] advocating for more studies -- including panel studies -- focusing on the socialization of youth.
+Further, we align with others [**WE NEED SOME CITES HERE**] advocating for more studies --including panel studies-- focusing on the socialization of youth.
 Our analysis points to the majority of variance among adults rooting in differences encountered before turning 18.
@@
 
-Finally, a key insight from our meta-analysis is that an increase in response resolution -- the range of response options survey respondents are given -- correlates with an increased share of systematic variance explained by intrapersonal change.
+Finally, a key insight from our meta-analysis is that an increase in response resolution --the range of response options survey respondents are given-- correlates with an increased share of systematic variance explained by intrapersonal change.
 This may reflect specific issues being assessed by survey items using different scales.
@@
 
-[^missing]: We estimate the models presented in Figures 1 and 2 using the full sample, but only use respondents who have information on their highest degree to generate the results presented in Figures 3 and 4.
-In total, ### respondents (%%% percent) reported no degree information.
-
 # References
@@
 
-# Appendix
+# Tables and Figures
 
@@
 ```{r fig1, fig.height=2}
-knitr::include_graphics("fig/variation_share_lm_mdpnt.png")
+knitr::include_graphics("../figures/figure_1_bw.png")
 ```
 ```{=latex}
-\textit{Notes:} The figure shows $\tau$ and 1 - $\tau$ as the proportion of systematic variance attributable to intrapersonal change and interpersonal differences, respectively.
-See Appendix A for the full set of item values.
+\textit{Notes:} The figure shows $\omega$ and 1 - $\omega$ as the proportion of systematic variance attributable to intrapersonal change and interpersonal differences, respectively.
+See Supplemental Materials A for the full set of item values.
 \end{figure}
@@
 ```{=latex}
-\begin{figure}[h]
-\caption{Regression Model Estimating $\tau$}
+\begin{figure}[ht]
+\caption{Regression Model Estimating $\omega$}
 
@@
 ```{r fig2, out.width = "300px"}
-knitr::include_graphics("fig/meta_no_int.png")
+knitr::include_graphics("../figures/figure_2.png")
 ```
@@
 Survey indicators and item topics not shown.
-See Appendix B for the full set of coefficients.
+See Supplemental Materials B for the full set of coefficients.
 \end{figure}
@@
 ```{=latex}
-\begin{figure}[h]
-\caption{Difference in $\tau$ Across College Graduates and Non-Graduates}
+\begin{figure}[ht]
+\caption{Difference in $\omega$ Across College Graduates and Non-Graduates}
 ```
 ```{r fig3, out.width = "450px"}
-knitr::include_graphics("fig/FIG_college1.png")
+knitr::include_graphics("../figures/figure_3_bw.png")
 ```
 ```{=latex}
-\textit{Notes:} The figure shows the difference in $\tau$ values across college graduates and college non-graduates.
+\textit{Notes:} The figure shows the difference in $\omega$ values across college graduates and college non-graduates.
 Values above (below) 0 means that those with college degree have higher (lower) variance of intrapersonal change.
@@
 ```{=latex}
-\begin{figure}[h]
-\caption{Difference in $\tau$ Across College Graduates and Non-Graduates on Political Culture}
+\begin{figure}[ht]
+\caption{Difference in $\omega$ Across College Graduates and Non-Graduates on Political Culture}
 ```
 ```{r fig4, out.width = "450px"}
-knitr::include_graphics("fig/FIG_college2.png")
+knitr::include_graphics("../figures/figure_4_bw.png")
 ```
 ```{=latex}
-\textit{Notes:} $\tau$ values across college graduates and college non-graduates on 8 political culture items from the General Social Survey (2006-2014).
+\textit{Notes:} $\omega$ values across college graduates and college non-graduates on 8 political culture items from the General Social Survey (2006-2014).
 \end{figure}

### 2023-09-27 manuscript/manuscript.tex
cleaning up and updates

--- 2023-09-27 manuscript.tex
@@
 
-A recurring focus in the sociology of culture is the question of whether
-people, as they move through their lives and transition between social
-roles, change their personal culture --- the declarative and
-non-declarative attitudes, beliefs, values, practices, and dispositions
-that exist at the individual level (Kiley and Vaisey 2020; Lersch 2023;
-Lizardo 2017).
-
-This seemingly simple question underlies a number of debates across
-sociological subfields. For example, pragmatist theories of action claim
-that changes in social environments cause people to adapt their views
-and make new cultural meanings (Gross 2009; Swidler 2001), while
-Bourdieusian practice theories argue that the ``past conditions of
-production'' leave a mark on people's personal culture that lasts
-throughout their lives (Bourdieu 1990). Similarly, models of social
-influence assume that people adapt their culture in the face of new
-information (Vaisey and Lizardo 2010 KK Note: I don't feel like this is
-a good cite for this.), while the emphasis on cohort effects in models
+Does personal culture --an individual's attitudes, beliefs, values, and
+practices--- change over the life course, or is it largely fixed once a
+person reaches adulthood?
+
+This question underlies several contemporary debates in sociology (Kiley
+and Vaisey 2020; Lersch 2023; Lizardo 2017). For example, pragmatist
+theories of action claim that changes in social environments cause
+people to adapt their views and make new cultural meanings (Gross 2009;
+Swidler 2001), while Bourdieusian practice theories argue that the
+``past conditions of production'' leave a mark on people's personal
+culture that lasts throughout their lives (Bourdieu 1990). Similarly,
+models of social influence assume that people adapt their culture in the
+face of new information, while the emphasis on cohort effects in models
 of aggregate social and cultural change requires them to be open to
@@
 Because all of the above theoretical perspectives have some empirical
-support, the current debate revolves around not whether they exist, but
-when and how they each apply and the contribution of each to explaining
-cultural differences. That being said, scholars have struggled to reach
-a consensus on the relative importance of change during adulthood. Over
-the span of a few years, most apparent changes in people's responses to
-personal culture items appear to be transitory, with little evidence of
-persistent change among adults (Kiley and Vaisey 2020; Vaisey and Kiley
-2021). This suggests that change during adulthood is not a strong force
-in explaining personal cultural differences. When looked at over longer
-time horizons, however, there is at least some evidence that adults make
-persistent cultural changes, especially as they move through distinct
-social roles (Lersch 2023).
+support, the current debate is not about whether they exist, but about
+their contribution to explaining cultural differences. That said,
+scholars have struggled to reach a consensus on the relative importance
+of change during adulthood. Over the span of a few years, most apparent
+changes in people's responses to personal survey items appear to be
+transitory, with little evidence of persistent change among adults
+(Kiley and Vaisey 2020; Vaisey and Kiley 2021). This suggests that
+change during adulthood is not a major factor in explaining personal
+cultural differences. When looked at over longer time horizons, however,
+there is at least some evidence that adults make persistent cultural
+changes (Lersch 2023).
 
@@
 adjudicating different theoretical perspectives (Lersch 2023: p.~228).
-Accordingly, researchers have formalized models of responses to cultural
-items over time as demonstrating either ``change'' or ``no change,''
-which essentially asks whether we \emph{ever} observe intrapersonal
-change in personal culture.
+That is, researchers have compared models of responses to cultural items
+over time as consistent with either ``change'' or ``no change,'' which
+essentially asks whether we observe \emph{any} intrapersonal change in
+personal culture.
 
@@
 change in a given dataset depends not only on sample populations, item
-measurement, statistical methods, and power, but crucially also on
-definitions of what counts as ``change.'' These methodological decisions
-potentially limit the possibility of convergence on the theoretical
-debates. Moreover, asking whether there is evidence of change limits the
-kind of questions that can be asked to, ``Do we see change in this
-measure?'' But such inquiries are ill-suited to address what's at the
-core of the theoretical debate. There is enough evidence to assume that
-neither socialization at an early age nor the influence of the
-contemporary environment alone can explain cultural differences. The
-question that remains is under what conditions, across which groups, and
-in what domains do we observe differences in the relative importance of
-intrapersonal change and interpersonal differences. This focus is
-necessarily comparative and requires an approach that quantifies forms
-of cultural change rather than calling victory for the one or the other
+measurement, statistical methods, and power, but also on definitions of
+what counts as ``change.'' These methodological decisions potentially
+limit the possibility of convergence on the theoretical debates.
+Moreover, asking whether there is evidence of change limits the kind of
+questions that can be asked to, ``Do we see change in this measure?''
+But such inquiries are ill-suited to address the core of the theoretical
+debate: determining \emph{how much} the process of intrapersonal change
+contributes to cultural differences.
+
+There is enough evidence to assert that neither pre-adult socialization
+nor the contemporary environment alone can fully explain cultural
+differences. The question that remains is about the relative
+contributions of these processes. In the current debate, for a single
+construct (e.g., support for gay rights), existing approaches can only
+make claims about \emph{whether} intrapersonal change is happening, not
+about how much. This limitation also prevents investigating under what
+conditions, across which groups, and in what domains do we see
+differences in the relative importance of intrapersonal change. These
+questions are about \emph{degree}, not existence, and thus require
+precise measurement rather than declarations of victory for a particular
 perspective.
 
-In this paper, we try to intervene in this debate by moving beyond the
+In this paper, we intervene in this debate by moving beyond the
 ``tournament of models'' approach toward a quantification of the
-processes that lead to differences in personal culture. Drawing on the
-seven panel datasets analyzed in the debate, we use established methods
-to quantify the amount of variance in responses that is due to
-between-people differences at a single point during the survey versus
-the amount of variance in responses that is due to linear intrapersonal
-change over time. This measure gives us a sense of the importance of
-change over the duration of the panel relative to the differences that
-exist between people assuming they never changed.
+processes that lead to differences in personal culture, and introducing
+a method for quantifying the relative contributions of interpersonal
+differences and intrapersonal change to the patterns we see in empirical
+data. Drawing on the seven panel datasets analyzed in the debate, we use
+established methods to quantify the amount of variance in responses that
+is due to between-people differences at a single point during the survey
+versus the amount of variance in responses that is due to linear
+intrapersonal change over time. This measure gives us a sense of the
+importance of change over the duration of the panel relative to the
+differences that exist between people assuming they never changed.
 
@@
 different interpretations, the results of previous work are in many ways
-highly consonant. {[}I think the revisions that were done to this
-paragraph were confusing and mostly reverted it.{]}
+highly consonant.
 
@@
 relative contributions of change and initial differences toward
-explaining differences in personal culture?'' {[}KK Note: I didn't like
-this change. It's not about collectively explaining, its about them
-relative to each other.{]}
+explaining differences in personal culture?''
 
@@
 start of the panel that were important in shaping their dispositions.
-{[}TK Note: I feel that this paragraph puts too much emphasis on
-pre-panel processes. It might be early imprinting or broad life-course,
-but why emphasize it this much? KK NOTE: Because it could be either, I
-don't want to treat it like it's just ``imprinting.''{]} {[}KK NOTE: You
-reversed what was being said here. You say ``early imprinting'' denotes
-intrapersonal change (process denotes number). We want to say the
-opposite -- what I had before -- number denotes process: ``interpersonal
-differences'' denotes ``early imprinting.'' I changed it back. I also
-took out reference to ``transformative'' experiences because I don't
-think we're necessarily talking about those specifically, just any
-experience that changes you. Because we don't say a specific theoretical
-process (``socialization'', etc.) I want to keep the language of each
-denoting combinations of processes. ``Persistent change'' isn't a
-theoretical process. It's a bunch of them.{]}
 
@@
 touch on the ongoing debate outlined here. Because of this we quantify
-this component but focus principally on the other two. {[}KK NOTE: I
-agreed with TK's note and mostly reverted this paragraph.{]}
+this component but focus principally on the other two.
 
@@
 
-Third, we expect the variance explained by each process --interpersonal
-differences and intrapersonal change-- to depend on specific attributes
-of the question and panel. First, the higher the response resolution for
-a given question and the larger the sample, the greater the variance
-explained by intrapersonal change should be, as smaller changes in
-intrapersonal change will be easier to detect. Second, the LCAM and
-various theoretical perspectives that underlie it posit that social
-transitions drive changes in personal culture. Since the probability
-that any person experiences such a transition goes up the longer we
-observe them, we expect panel duration to be positively related to the
-proportion of variation explained by intrapersonal change.
+Third, we expect the variance explained by each process
+-\/--interpersonal differences and intrapersonal change-\/-- to depend
+on specific attributes of the question and panel. First, the higher the
+response resolution for a given question and the larger the sample, the
+greater the variance explained by intrapersonal change should be, as
+smaller changes in intrapersonal change will be easier to detect.
+Second, the LCAM and various theoretical perspectives that underlie it
+posit that social transitions drive changes in personal culture. Since
+the probability that any person experiences such a transition goes up
+the longer we observe them, we expect panel duration to be positively
+related to the proportion of variation explained by intrapersonal
+change.
 
@@
 \] \[
-u_{0,i} \sim \mathcal{N}(0,\sigma_{\mu_0}^2)
-\]
-
-\[
+u_{0,i} \sim \mathcal{N}(0,\sigma_{\mu_0}^2) \\
+\] \[
 u_{1,i} \sim \mathcal{N}(0,\sigma_{u_1}^2)
 \]
-
-{[}KKNOTE: Is that right? I forget how to do that exactly.{]}
 
@@
 \[
-\tau = \frac{V(C)}{V(C) + V(D)}
+\omega = \frac{V(C)}{V(C) + V(D)}
 \]
 
-{[}TK Note: \textbackslash phi taken in Kiley \& Vaisey,
-\textbackslash delta taken in Vaisey \& Lizardo. I think using a Greek
-letter without strong associations (something like alpha, beta, or mu or
-whatever) is not good. So I settled on \textbackslash tau. Feel free to
-change it.{]}
-
-We can think of this number, \(\tau\) as akin to an Intraclass
+We can think of this number, \(\omega\) as akin to an Intraclass
 Correlation Coefficient focused only on the systematic components of the
@@
 explaining systematic cultural differences during the panel. Moving
-forward, we use \(\tau\) as a placeholder for \emph{systematic variance
-attributable to intrapersonal change}.
+forward, we use \(\omega\) as a placeholder for \emph{systematic
+variance attributable to intrapersonal change}.
 
@@
 As a second step in our analysis, we use ordinary least squares
-regression to model \(\tau\) as a function of features of the question
+regression to model \(\omega\) as a function of features of the question
 and panel. This allows us to explore how these features affect the
@@
 
-{[}TK Note: I guess we need to say something here about the college
-stuff? KK Note: I didn't feel like that was necessary. It feels so
-distinct later on that I think it's OK to kind of ``start over'' there.
-I'm worried that people will forget the plan by the time they get there,
-and we'll have to reiterate everything anyway. But others should weigh
-in.{]}.
-
 \hypertarget{results}{%
@@
 systematic variance attributable to intrapersonal change. Across panels,
-the median proportion of \(\tau\) is 0.094 (mean of 0.098), with an
+the median proportion of \(\omega\) is 0.094 (mean of 0.098), with an
 interquartile range of 0.053 to 0.139. All panels include questions
-where \(\tau\) is essentially 0, and all panels include questions with
-\(\tau\) values greater than 0.20. The maximum \(\tau\) value is 0.26.
+where \(\omega\) is essentially 0, and all panels include questions with
+\(\omega\) values greater than 0.20. The maximum \(\omega\) value is
+0.26.
 
@@
 <about here>
+Figure 1: Proportions of Systematic Variance in Personal Culture
 \end{center}
-
-Figure 1: Proportions of explained variance in items of personal
-culture; based on predictions at participants' wave mid-point.
 
@@
 To the extent that there are differences across the panels, the PSID has
-the highest \(\tau\) values with mean .120 and median .135. While we
+the highest \(\omega\) values with mean .120 and median .135. While we
 cannot disentangle features of the sample from features of the questions
@@
 reflect the distinct age profile of respondents in this sample. At the
-other end, the GSS has the lowest range of \(\tau\) with mean .073 and
+other end, the GSS has the lowest range of \(\omega\) with mean .073 and
 median .069. This potentially reflects the fact that the GSS observes
@@
 make significant cultural changes the longer we observe them, then
-duration likely affects the range of \(\tau\) (a point we explore more
+duration likely affects the range of \(\omega\) (a point we explore more
 below). However, the GSS results are still consistent with results from
@@
 small compared to the differences within panels. For about 6 percent of
-items explored here, \(\tau\) is greater than 0.20. These questions tend
-to ask about external referents (e.g., confidence in specific government
-leaders or political parties), life satisfaction, or current financial
-position. At the other end, questions about religious identification,
-views on gender roles, and support for civil liberties tend to have very
-low proportion of variance attributable to intrapersonal change.
+items explored here, \(\omega\) is greater than 0.20. These questions
+tend to ask about external referents (e.g., confidence in specific
+government leaders or political parties), life satisfaction, or current
+financial position. At the other end, questions about religious
+identification, views on gender roles, and support for civil liberties
+tend to have very low proportion of variance attributable to
+intrapersonal change.
 
@@
 
-Appendix A presents the estimated proportion of variance attributable to
-interpersonal differences and intrapersonal change, the estimated values
-of \(\tau\), and the proportion of residual variance for each question
-analyzed. In general, interpersonal differences are almost always the
-largest component of the total variance and tend to explain between 55
-and 70 percent of total variance, while intrapersonal change is always
-the smallest, typically explaining between 3 and 8 percent of total
-variance. Residual variance tends to account for between 22 and 37
-percent of variance, though on several questions residual variance is
-greater than 50 percent.
+Supplemental Materials A presents the estimated proportion of variance
+attributable to interpersonal differences and intrapersonal change, the
+estimated values of \(\omega\), and the proportion of residual variance
+for each question analyzed. In general, interpersonal differences are
+almost always the largest component of the total variance and tend to
+explain between 55 and 70 percent of total variance, while intrapersonal
+change is always the smallest, typically explaining between 3 and 8
+percent of total variance. Residual variance tends to account for
+between 22 and 37 percent of variance, though on several questions
+residual variance is greater than 50 percent.
 
@@
 Figure 2 plots the results from an ordinary least squares regression of
-\(\tau\) as a function of question, panel, and sample features. These
-models also include fixed effects for panels, so coefficients reflect
-the association between a particular sample or question characteristic
-within a panel.{[}KK NOTE: CAN WE SUMMARIZE THESE PANEL FIXED EFFECTS AT
-ALL? AE: WE COULD SHOW THAT THEY DON'T CAPTURE TOPIC EFFECTS BEYOND OUR
-CLASSIFICATION IF WE INCLUDE THOSE / THAT THEY POSSIBLY CAPTURE
-UNMEASURED COUNTRY/CONTEXT DIFFERENCES OR DIFFERENCES IN SURVEY
-METHODOLOGY?{]}
+\(\omega\) as a function of question, panel, and sample features. These
+models also include fixed effects for panels and topics, so coefficients
+reflect the association within a panel and topical domain.
 
@@
 <about here>
+Figure 2: Regression Model Estimating $\omega$
 \end{center}
 
-Figure 2: Proportion of explained variance attributable to intrapersonal
-change; based on predictions at participants' wave mid-point; supressed
-intercept model; survey indicators not shown.
-
 Figure 2 shows that the more response options respondents are given and
-the larger the sample, the larger the values of \(\tau\). We interpret
+the larger the sample, the larger the values of \(\omega\). We interpret
 these coefficients as suggesting greater resolution on a question makes
 it easier to detect and model change. The earlier a question was asked,
-even net of how long it was been asked, the lower the \(\tau\).
+even net of how long it was been asked, the lower the \(\omega\).
+
+While we did not generate expectations for how question content relates
+to \(\omega\), some of the associations between question structure,
+panel duration, and panels themselves might be driven by differences in
+the topics addressed by each panel. To address this, we coded each
+question as falling into one of nine different content categories and
+included these as indicator variables in the meta analysis. Supplemental
+Materials B presents the estimates for these content indicators. They
+show some associations with \(\omega\), with questions about subjective
+SES; social life, social cohesion, and trust; environment and climate;
+and health and morale showing larger \(\omega\) values than questions
+about religion and spirituality; politics, government and the economy;
+and gender and family.
 
@@
 attached to the duration of years covered by the question. The longer a
-question is observed, the less is the value of \(\tau\). Theories of
+question is observed, the less is the value of \(\omega\). Theories of
 personal cultural change that link changes in personal culture to social
@@
 coefficient attached to this variable can be a challenge to that
-assumption {[}TK Note: since there is a complication of this, I changed
-is to can be, but feel free to ignore it.{]}
-
-As a further test of this coefficient, we compared the \(\tau\) values
+assumption.
+
+As a further test of this coefficient, we compared the \(\omega\) values
 when using the full duration of a panel compared to when we dropped the
@@
 the question. If the coefficient reflects a true negative effect of
-duration on \(\tau\), we should see that same effect within questions.
-Results from this analysis are presented in Appendix B and contradict
-the coefficient from the regression model: we found that the longer we
-observe the same question, the higher the \(\tau\) is. We interpret this
-combination of findings as suggesting that the kinds of questions asked
-for a longer time period tend to demonstrate less intrapersonal change
-than questions asked over shorter periods, rather than a true function
-of time.
-
-This reinforces the explanation for why the GSS displays the lowest
-average proportion of systematic variance attributable to intrapersonal
-change: its shorter duration.
-
-While we did not generate expectations for how question content relates
-to \(\tau\), some of the associations between question structure, panel
-duration, and panels themselves might be driven by differences in the
-topics addressed by each panel. To address this, we coded each question
-as falling into one of nine different content categories and included
-these as indicator variables in the meta analysis. Appendix C presents
-the estimates for these content indicators. They show some associations
-with \(\tau\), with questions about subjective SES; social life, social
-cohesion, and trust; environment and climate; and health and morale
-showing larger \(\tau\) values than questions about religion and
-spirituality; politics, government and the economy; and gender and
-family. {[}KKNOTE: I FEEL LIKE WE WANTED TO SAY SOMETHING ELSE HERE BUT
-I CANT REMEMBER WHAT{]}
+duration on \(\omega\), we should see that same effect within questions.
+Results from this analysis are presented in Supplemental Materials C and
+contradict the coefficient from the regression model: we found that the
+longer we observe the same question, the higher the \(\omega\) is. We
+interpret this combination of findings as suggesting that the kinds of
+questions asked for a longer time period tend to demonstrate less
+intrapersonal change than questions asked over shorter periods, rather
+than a true function of time. This reinforces the explanation for why
+the GSS displays the lowest average proportion of systematic variance
+attributable to intrapersonal change: its shorter duration.
 
@@
 To test these competing propositions, we use the models presented above
-to separately calculate \(\tau\) values for people with at least a
+to separately calculate \(\omega\) values for people with at least a
 bachelor's degree and people with less than a bachelor's degree at wave
@@
   across the panel produces functionally identical results.} of the
-General Social Survey's panels.\footnote{We estimate the models
-  presented in Figures 1 and 2 using the full sample, but only use
-  respondents who have information on their highest degree to generate
-  the results presented in Figures 3 and 4. In total, \#\#\# respondents
-  (\%\%\% percent) reported no degree information.} We focus on the GSS
-because it contains the largest number of questions tapping general
-political dispositions, which is the domain where education has proven
-particularly relevant for understanding attitude stability. The GSS also
-covers a turbulent window of American politics from 2006 to 2014. This
-window covers the start of the Great Recession, debates about federal
-intervention in and regulation of Wall Street, the election of Barack
-Obama as the first black president in U.S. history, debates about the
-role of the federal government in the health care sector, the emergence
-of the Tea Party, and political realignment and clarification on the
-issue of gay marriage, among other topics.
-
-Figure 3 plots the distribution of differences in \(\tau\) values
+General Social Survey's panels. We focus on the GSS because it contains
+the largest number of questions tapping general political dispositions,
+which is the domain where education has proven particularly relevant for
+understanding attitude stability. The GSS also covers a turbulent window
+of American politics from 2006 to 2014. This window covers the start of
+the Great Recession, debates about federal intervention in and
+regulation of Wall Street, the election of Barack Obama as the first
+black president in U.S. history, debates about the role of the federal
+government in the health care sector, the emergence of the Tea Party,
+and political realignment and clarification on the issue of gay
+marriage, among other topics.
+
+Figure 3 plots the distribution of differences in \(\omega\) values
 between people with at least a bachelor's degree and people with less
@@
 <about here>
+Figure 3: Difference in $\omega$ Across College Graduates and Non-Graduates
 \end{center}
-
-Figure 3: Difference between share of intrapersonal change and share of
-inter-personal differences in explained variance; based on predictions
-at participants' wave mid-point
 
@@
 about whether black people should be given preferences in hiring,
-measured on a five-point scale. Estimates of \(\tau\) for these eight
+measured on a five-point scale. Estimates of \(\omega\) for these eight
 questions, for both education groups, are presented in Figure 4.
@@
 <about here>
+Figure 4: Difference in $\omega$ Across College Graduates and Non-Graduates on Political Culture
 \end{center}
 
-Figure 4: Proportions of intrapersonal change of explained variance by
-college degree; based on predictions at participants' wave mid-point.
-
-On all eight questions presented in Figure 4, \(\tau\) values are
+On all eight questions presented in Figure 4, \(\omega\) values are
 greater for people without a college degree than for people with a
@@
 variance for people with a college degree than for people without one.
-There are also large differences in \(\tau\) values across questions for
-both groups. For example, on the question of whether the government
+There are also large differences in \(\omega\) values across questions
+for both groups. For example, on the question of whether the government
 should try to solve more problems or leave those problems to be solved
@@
 In contrast, partisan identification and political ideology both show
-larger values of \(\tau\) than most other questions, as well as a larger
-absolute difference between people with a college degree and those
-without. Compared to the other questions explored here, change as adults
-plays a much larger role in explaining why people report the partisan
-identification and political ideology they do. And this is particularly
-true for respondents who do not have a college degree. The \(\tau\)
-value for ideological identification among non-college educated
-respondents is almost four times that of college-educated respondents.
-
-We note here that this meaningful difference in \(\tau\) values across
+larger values of \(\omega\) than most other questions, as well as a
+larger absolute difference between people with a college degree and
+those without. Compared to the other questions explored here, change as
+adults plays a much larger role in explaining why people report the
+partisan identification and political ideology they do. And this is
+particularly true for respondents who do not have a college degree. The
+\(\omega\) value for ideological identification among non-college
+educated respondents is almost four times that of college-educated
+respondents.
+
+We note here that this meaningful difference in \(\omega\) values across
 education groups and across questions would not be detectable using
@@
 
-Applying our proposed measure, \(\tau\), to all questions about personal
-culture from the panel datasets previously discussed in this debate
-revealed a consistent pattern. Nearly all questions show evidence of
-people making durable, intrapersonal change over time, with some showing
-notably high intrapersonal change compared to interpersonal differences.
-For some questions about factors that affect one's life satisfaction or
-views on government benefits, it seems plausible that differences in
-adult experiences predominantly account for the observed differences
-between people.
+Applying our proposed measure, \(\omega\), to all questions about
+personal culture from the panel datasets previously discussed in this
+debate revealed a consistent pattern. Nearly all questions show evidence
+of people making durable, intrapersonal change over time, with some
+showing notably high intrapersonal change compared to interpersonal
+differences. For some questions about factors that affect one's life
+satisfaction or views on government benefits, it seems plausible that
+differences in adult experiences predominantly account for the observed
+differences between people.
 
@@
 insufficient in magnitude and duration to explain what are often
-pronounced differences among people in diverse roles. {[}I'VE TRIED TO
-FIND A WAY TO GET BOTH IMPACT AND DURATION IN THERE; COULDN'T FIND A
-BETTER SOLUTION THOUGH--IDEAS? ALSO WE NEED TO BE CAREFUL WITH THIS
-SENTENCE, IT SEEMS WE ARE OVERSTATING HERE; LET'S FIND A WAY TO
-RELATIVIZE THE CLAIM HERE?{]}. These and our results suggest that when
-observing differences in personal culture across social roles, such as
-parenthood, education, or professional authority (Longest, Hitlin, and
-Vaisey 2013), or across occupations (Weeden and Grusky 2005), selection
-likely plays a large role in explaining these difference, though
-exceptions always exist.
+pronounced differences among people in diverse roles. These and our
+results suggest that when observing differences in personal culture
+across social roles, such as parenthood, education, or professional
+authority (Longest, Hitlin, and Vaisey 2013), or across occupations
+(Weeden and Grusky 2005), selection likely plays a large role in
+explaining these difference, though exceptions always exist.
 
@@
 the two is therefore crucial for understanding cultural differences and
-cannot be done with cross-sectional data. {[}MAYBE THE COMPARISON
-BETWEEN NON-SYSTEMATIC VARIATION AND SYSTEMATIC (LINEAR+BASELINE
-DIFFERENCES) WOULD EVEN BE BETTER TO MAKE THAT POINT?{]}. Instead we
-should extend panels beyond two waves to gain more leverage to
-understand when and in what form personal culture changes.
-
-Further, we align with others {[}CITES?{]} advocating for more studies
--- including panel studies -- focusing on the socialization of youth.
-Our analysis points to the majority of variance among adults rooting in
-differences encountered before turning 18. By empirically constraining
-their analysis to experiences after that point, social scientists may
-inadvertently concentrate on topics and questions that, while important,
-might not be able to help explain major cultural differences in adult
-populations.
+cannot be done with cross-sectional data. Instead we should extend
+panels beyond two waves to gain more leverage to understand when and in
+what form personal culture changes.
+
+Further, we align with others {[}\textbf{WE NEED SOME CITES HERE}{]}
+advocating for more studies --including panel studies-- focusing on the
+socialization of youth. Our analysis points to the majority of variance
+among adults rooting in differences encountered before turning 18. By
+empirically constraining their analysis to experiences after that point,
+social scientists may inadvertently concentrate on topics and questions
+that, while important, might not be able to help explain major cultural
+differences in adult populations.
 
 Finally, a key insight from our meta-analysis is that an increase in
-response resolution -- the range of response options survey respondents
-are given -- correlates with an increased share of systematic variance
+response resolution --the range of response options survey respondents
+are given-- correlates with an increased share of systematic variance
 explained by intrapersonal change. This may reflect specific issues
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-vaisey2010a}{}}%
-Vaisey, Stephen, and Omar Lizardo. 2010. {``Can {Cultural} {Worldviews}
-{Influence} {Network} {Composition}?''} \emph{Social Forces}
-88(4):1595--1618. doi:
-\href{https://doi.org/10.1353/sof.2010.0009}{10.1353/sof.2010.0009}.
-
 \leavevmode\vadjust pre{\hypertarget{ref-vaisey2016}{}}%
@@
 
-\hypertarget{appendix}{%
-\section{Appendix}\label{appendix}}
+\hypertarget{tables-and-figures}{%
+\section{Tables and Figures}\label{tables-and-figures}}
 
@@
 
-\begin{center}\includegraphics[width=20.83in]{fig/variation_share_lm_mdpnt} \end{center}
-
-\textit{Notes:} The figure shows $\tau$ and 1 - $\tau$ as the proportion of systematic variance attributable to intrapersonal change and interpersonal differences, respectively.
+\begin{center}\includegraphics[width=20.83in]{../figures/figure_1_bw} \end{center}
+
+\textit{Notes:} The figure shows $\omega$ and 1 - $\omega$ as the proportion of systematic variance attributable to intrapersonal change and interpersonal differences, respectively.
 See Appendix A for the full set of item values.
@@
 \begin{figure}[ht]
-\caption{Regression Model Estimating $\tau$}
-
-\begin{center}\includegraphics[width=300px]{fig/meta_no_int} \end{center}
+\caption{Regression Model Estimating $\omega$}
+
+\begin{center}\includegraphics[width=300px]{../figures/figure_2} \end{center}
 
@@
 \begin{figure}[ht]
-\caption{Difference in $\tau$ Across College Graduates and Non-Graduates}
-
-\begin{center}\includegraphics[width=450px]{fig/FIG_college1} \end{center}
-
-\textit{Notes:} The figure shows the difference in $\tau$ values across college graduates and college non-graduates.
+\caption{Difference in $\omega$ Across College Graduates and Non-Graduates}
+
+\begin{center}\includegraphics[width=450px]{../figures/figure_3_bw} \end{center}
+
+\textit{Notes:} The figure shows the difference in $\omega$ values across college graduates and college non-graduates.
 Values above (below) 0 means that those with college degree have higher (lower) variance of intrapersonal change.
@@
 \begin{figure}[ht]
-\caption{Difference in $\tau$ Across College Graduates and Non-Graduates on Political Culture}
-
-\begin{center}\includegraphics[width=450px]{fig/FIG_college2} \end{center}
-
-\textit{Notes:} $\tau$ values across college graduates and college non-graduates on 8 political culture items from the General Social Survey (2006-2014).
+\caption{Difference in $\omega$ Across College Graduates and Non-Graduates on Political Culture}
+
+\begin{center}\includegraphics[width=450px]{../figures/figure_4_bw} \end{center}
+
+\textit{Notes:} $\omega$ values across college graduates and college non-graduates on 8 political culture items from the General Social Survey (2006-2014).
 \end{figure}

### 2023-09-27 manuscript/manuscript.Rmd
first SV edits

@@
 
-Does personal culture --an individual's attitudes, beliefs, values, and practices--- change over the life course, or is it largely fixed once a person reaches adulthood?
+Does personal culture---an individual's attitudes, beliefs, values, and practices---change over the life course, or is it largely fixed by adulthood?
+This question underlies an important contemporary debate in sociology [@lizardo2017; @kiley2020; @lersch2023] and has deep roots in seemingly contradictory theoretical perspectives.
+For example, pragmatist theories of action claim that changes in social environments cause people to adapt their views and make new cultural meanings [@swidler2001; @gross2009], while Bourdieusian practice theories argue that the "past conditions of production" leave a mark on people's personal culture that lasts throughout their lives [@bourdieu1990].
+Models of social influence assume that people adapt their culture in the face of new information [@goldberg2018], while the emphasis on cohort effects in models of aggregate social and cultural change requires them to be open to change while young but become fairly resistant to it as they age [@ryder1965].
+Finally, life course theories posit important changes over time as people advance through important transitions in their lives [@elder2003; @bardi2009].
 
-This question underlies several contemporary debates in sociology [@lizardo2017; @kiley2020; @lersch2023].
-For example, pragmatist theories of action claim that changes in social environments cause people to adapt their views and make new cultural meanings [@swidler2001; @gross2009], while Bourdieusian practice theories argue that the “past conditions of production” leave a mark on people's personal culture that lasts throughout their lives [@bourdieu1990].
-Similarly, models of social influence assume that people adapt their culture in the face of new information, while the emphasis on cohort effects in models of aggregate social and cultural change requires them to be open to change while young but become fairly resistant to it as they age [@ryder1965].
-Finally, life course theories posit slower but still important changes over time as people advance through important transitions in their lives [@elder2003; @bardi2009].
+Because all of these theoretical perspectives have some empirical support, the current debate is not about whether their posited processes exist at all, but about their relative contribution to explaining the cultural differences we see in the world.
+That said, scholars have struggled to reach a consensus on the importance of change during adulthood.
+Over the span of a few years, most apparent changes in people's responses to personal survey items appear to be transitory, with little evidence of persistent change among adults [@kiley2020; @vaisey2021].
+This suggests that change during adulthood is not a major factor in explaining personal cultural differences.
+When we consider a longer time horizon, however, there is evidence that adults make at least some persistent changes [@lersch2023].
 
-Because all of the above theoretical perspectives have some empirical support, the current debate is not about whether they exist, but about their contribution to explaining cultural differences.
-That said, scholars have struggled to reach a consensus on the relative importance of change during adulthood.
-Over the span of a few years, most apparent changes in people’s responses to personal survey items appear to be transitory, with little evidence of persistent change among adults [@kiley2020; @vaisey2021].
-This suggests that change during adulthood is not a major factor in explaining personal cultural differences.
-When looked at over longer time horizons, however, there is at least some evidence that adults make persistent cultural changes [@lersch2023].
+These seemingly inconsistent findings are partly due to the fact that researchers have taken a "tournament of models" approach to adjudicating different theoretical perspectives [@lersch2023: p.
+228].
+That is, researchers have compared the fit of different statistical models to each item to determine whether it is more consistent with "no change" or "any change." The size of these piles then serves as the primary evidence for the truth of the theory [@kiley2020; @vaisey2021; @lersch2023].
 
-These seemingly inconsistent findings are partly due to the fact that researchers have taken a “tournament of models” approach in adjudicating different theoretical perspectives [@lersch2023: p.
-228].
-That is, researchers have compared models of responses to cultural items over time as consistent with either “change” or “no change,” which essentially asks whether we observe *any* intrapersonal change in personal culture.
+We argue, however, that asking *whether* people change on an item may cause more problems than it solves.
+Being able to detect *any* change on an item depends not only on sample populations, measurement, statistical methods, and power, but also on definitions of what counts as "change." These differences can limit the possibility of agreement on which "pile" a variable belongs in.
+Moreover, asking whether there is evidence of change reduces to, "Is there *any* evidence of change in this item?" But such binary inquiries are ill-suited to address the core of the theoretical debate: determining *how much* the process of intrapersonal change contributes to cultural differences.
 
-We argue that asking only *whether* people change obscures what are actually highly compatible empirical patterns.
-Being able to detect change in a given dataset depends not only on sample populations, item measurement, statistical methods, and power, but also on definitions of what counts as “change.” These methodological decisions potentially limit the possibility of convergence on the theoretical debates.
-Moreover, asking whether there is evidence of change limits the kind of questions that can be asked to, “Do we see change in this measure?” But such inquiries are ill-suited to address the core of the theoretical debate: determining *how much* the process of intrapersonal change contributes to cultural differences.
+There is enough evidence to assert that neither pre-adult socialization nor the contemporary social context alone can explain the full range of cultural differences.
+The question that remains is about the *relative* contributions of these processes.
+In the current debate, for a single construct (e.g., support for gay rights), existing approaches can only make claims about *whether* intrapersonal change is happening, not about how much.
+This also prevents investigating under what conditions, across which groups, and in what domains do we see differences in the relative importance of intrapersonal change and pre-existing interpersonal differences.
+These questions are about *degree*, not existence; taking this debate forward will require precise measurement rather than declarations of victory for a particular perspective.
 
-There is enough evidence to assert that neither pre-adult socialization nor the contemporary environment alone can fully explain cultural differences.
-The question that remains is about the relative contributions of these processes.
-In the current debate, for a single construct (e.g., support for gay rights), existing approaches can only make claims about *whether* intrapersonal change is happening, not about how much.
-This limitation also prevents investigating under what conditions, across which groups, and in what domains do we see differences in the relative importance of intrapersonal change.
-These questions are about *degree*, not existence, and thus require precise measurement rather than declarations of victory for a particular perspective.
+In this paper, we move beyond the "tournament of models" approach and introduce a method for quantifying the relative contributions of interpersonal differences and intrapersonal change for a single item.
+Drawing on seven panel data sets from five countries, we quantify the amount of variance in survey responses that is due to between-person differences at a single point during the survey versus the amount of variance in responses that is due to systematic intrapersonal change over the study period.
+This measure quantifies the amount of change over the duration of the panel relative to the stable differences that exist between people.
 
-In this paper, we intervene in this debate by moving beyond the “tournament of models” approach toward a quantification of the processes that lead to differences in personal culture, and introducing a method for quantifying the relative contributions of interpersonal differences and intrapersonal change to the patterns we see in empirical data.
-Drawing on the seven panel datasets analyzed in the debate, we use established methods to quantify the amount of variance in responses that is due to between-people differences at a single point during the survey versus the amount of variance in responses that is due to linear intrapersonal change over time.
-This measure gives us a sense of the importance of change over the duration of the panel relative to the differences that exist between people assuming they never changed.
+Using this measure, we observe similar patterns across the datasets that have previously yielded conflicting interpretations.
+In all datasets---despite their varied social contexts and duration---we find a similar distribution of intrapersonal change.
+In general, we find that intrapersonal change accounts for a substantially smaller amount of variance in personal culture survey item than stable interpersonal differences.
+Of course, whether the amount of intrapersonal change in an item is large enough to be "meaningful" is a substantive and theoretical question.
+Nevertheless, on many measures of personal culture (some observed for over a decade), we do not observe enough intrapersonal change for it to play a substantial role in explaining the differences we observe between adults.
+Notable, interesting, and important exceptions exist, of course, but they are departures from the overall pattern.
 
-We observe similar patterns across the datasets that yielded previously conflicting interpretations.
-In all datasets we find a similar distribution of how much explanatory power can be attributed to intrapersonal change.
-But quantifying these processes also reveals that intrapersonal change plays a substantially smaller role in explaining the variance in peoples’ personal culture than interpersonal differences at their baselines.
-Whether the amount of change is large enough to be “meaningful” ultimately remains a theoretical question reflected in prior expectations about particular cultural item and population.
-It also crucially depends on how long people are observed in a panel.
-Nevertheless, on many measures of personal culture we do not observe enough intrapersonal change, in some cases over more than 10 years, for it to play a substantial role in explaining the differences we observe between adults, though notable exceptions exist.
+In addition to advancing a theoretical debate, our method can also help address substantive questions.
+To illustrate how it might be used by a researcher with a specific topical interest, we investigate differences in the relative importance of intrapersonal change for people with and without a college degree.
+We find a smaller amount of systematic change among college-educated respondents, suggesting that college solidifies one's personal culture rather than fostering an openness to new information.
+This offers a novel empirical basis to theorize about the role of university education in shaping personal culture.
 
-These results help clarify and, hopefully, transcend previous debates about whether adults change.
-What is more, the approach we propose can be applied to a broader range of theoretical questions than approaches discerning whether durable change exists in a panel.
-To exemplify this, we explore differences in the explanatory power of intrapersonal change for people with college degrees and those without.
-We find that change has less explanatory power for people with college degrees than those without.
-This may imply that the college experience solidifies political dispositions rather than fosters an openness and adaptability to new information and influences later in life.
-This offers a novel, empirical basis to theorize about the role of college in shaping personal culture.
-We believe that this approach has the potential to advance the theoretical debate on the role of environments, social identities, and roles in facilitating personal cultural change and stability.
+Taken together, to goal of this paper is to advance---and, hopefully, transcend---debates about *whether* adults change.
+People do change, at least a little, on most things.
+Our method can help quantify exactly how much.
 
@@
 Recent debates about whether adults undergo intrapersonal cultural change emerged in part because theories of cultural change at the aggregate level tend to implicitly invoke one of two models of individual behavior.
-The first, what Kiley and Vaisey [-@kiley2020] call a “Settled Dispositions Model” (SDM), assumes that peoples’ personal culture is relatively fixed by the time they are adults.
+The first, what Kiley and Vaisey [-@kiley2020] call a "Settled Dispositions Model" (SDM), assumes that peoples' personal culture is relatively fixed by the time they are adults.
 While they might make temporary changes in their declarative culture in reaction to their environments, this model assumes that people return to a settled baseline over a short period of time.
-This model underlies theories of cultural change that suggest people are imprinted by early socialization experiences such as the “past conditions of production” a Bourdieusian practice theory [@bourdieu1990], cohort replacement theories of aggregate change [@mannheim1952; @ryder1965], and control theories in social psychology [@smith-lovin1988; @robinson2007].
+This model underlies theories of cultural change that suggest people are imprinted by early socialization experiences such as the "past conditions of production" a Bourdieusian practice theory [@bourdieu1990], cohort replacement theories of aggregate change [@mannheim1952; @ryder1965], and control theories in social psychology [@smith-lovin1988; @robinson2007].
 
-The second model summarized by Kiley and Vaisey [-@kiley2020], an “Active Updating Model” (AUM), posits that people continually update their personal culture as they move through life.
+The second model summarized by Kiley and Vaisey [-@kiley2020], an "Active Updating Model" (AUM), posits that people continually update their personal culture as they move through life.
 This model suggests people change their personal culture as they adapt and make new meanings when encountering new social environments, discourses, and information [@gross2009; @swidler2001].
@@
 
-There is no reason to believe that only one of these two models is “correct” at all times and in all places.
+There is no reason to believe that only one of these two models is "correct" at all times and in all places.
 A population observed over some period of time contains a mix of people who are noticeably changing and people who are not.
 Instead, different perspectives argue that each of these ideal-typical models is more operative at different times, for different people, and for different elements of personal culture.
-For example, adolescence and early adulthood is typically viewed as a “formative period” for personal cultural development and thus characterized by higher rates of active updating, while middle age and later life are potentially characterized more by settled dispositions [@alwin1991; @krosnick1989; @eaton2009].
+For example, adolescence and early adulthood is typically viewed as a "formative period" for personal cultural development and thus characterized by higher rates of active updating, while middle age and later life are potentially characterized more by settled dispositions [@alwin1991; @krosnick1989; @eaton2009].
 Similarly, salient issues, such as views around gay rights in the 2010s; issues that see substantial elite realignments, such as views around the Vietnam conflict in the 1970s [@zaller1992]; and novel issues of public opinion, such as views around vaccines during the Covid-19 pandemic [@scoville2022], might be characterized by active updating, while established issues of low salience are characterized by stability.
 
-Empirically comparing which of these models better fit a broad range of questions from the General Social Survey’s rotating panels, Kiley and Vaisey [-@kiley2020] found limited evidence of durable change.
+Empirically comparing which of these models better fit a broad range of questions from the General Social Survey's rotating panels, Kiley and Vaisey [-@kiley2020] found limited evidence of durable change.
 While the majority of items preferred the AUM, the amount of durable change detected on these items was small.
 A substantial minority of questions (39 percent) favored the SDM, meaning they showed no evidence of durable change.
-Questions with more evidence of durable change included salient issues like gay marriage and questions tapping “public” statements or behaviors such as partisan identification and religious service attendance.
+Questions with more evidence of durable change included salient issues like gay marriage and questions tapping "public" statements or behaviors such as partisan identification and religious service attendance.
 There was also more evidence of durable change among early adults (people ages 18-30) than the rest of the population.
-Overall, the researchers concluded that “results ultimately suggest that real, persistent attitude change is an uncommon phenomenon among adults'' [@kiley2020: p.
+Overall, the researchers concluded that "results ultimately suggest that real, persistent attitude change is an uncommon phenomenon among adults'' [@kiley2020: p.
 500; see also @vaisey2021].
@@
 On the other hand, the claim that change is a relatively infrequent phenomenon has been difficult to square with research identifying durable change as a result of social experiences across a number of cultural dimensions, such as morality [@brocic2021], trust [@mewes2021], and concerns about immigration [@kratz2021].
-Similarly, there are longitudinal studies showing that cues from political elites can change an individual’s position on specific policy issues [@slothuus2021; @zaller1992] and that changes in their close contacts and acquaintances can change individuals' attitudes on group-related politics [@gelman2021; @dellaposta2018].
+Similarly, there are longitudinal studies showing that cues from political elites can change an individual's position on specific policy issues [@slothuus2021; @zaller1992] and that changes in their close contacts and acquaintances can change individuals' attitudes on group-related politics [@gelman2021; @dellaposta2018].
 Given how often we observe people change their personal culture, it is hard to accept that adults do not change.
 
-Drawing on these findings, Lersch [-@lersch2023] challenged the SDM and AUM as a “needless dichotomy,” proposing the “Life Course Adaption Model” (LCAM) as an alternative.
+Drawing on these findings, Lersch [-@lersch2023] challenged the SDM and AUM as a "needless dichotomy," proposing the "Life Course Adaption Model" (LCAM) as an alternative.
 This model draws on the life course perspective to model personal culture as a different linear trajectory over the duration of a panel for each respondent.
@@
 No questions favored the AUM.
-He concludes that “new experiences over the life course [...] can persistently move individuals’ personal culture in novel directions” [@lersch2023: p.
+He concludes that "new experiences over the life course [...] can persistently move individuals' personal culture in novel directions" [@lersch2023: p.
 243-244].
@@
 Despite differences in how these researchers interpret their findings, the empirical results are not far apart.
-Kiley and Vaisey [-@kiley2020] found that the majority of items they tested favored the AUM, meaning there was evidence of durable change among adults, a pattern consistent with Lersch’s results.
+Kiley and Vaisey [-@kiley2020] found that the majority of items they tested favored the AUM, meaning there was evidence of durable change among adults, a pattern consistent with Lersch's results.
 While Lersch [-@lersch2023] found evidence of durable change on most items, this change only accounted for a small proportion of variance explained.
 On average, people changed only about .07 standard deviations over 10 years.
-This is consistent with Kiley and Vaisey’s finding that durable change is often small.
+This is consistent with Kiley and Vaisey's finding that durable change is often small.
 And a quarter of the items Lersch studied still preferred the SDM, suggesting that even with different assumptions assumptions about change and more extensive data, many questions show no evidence of durable change.
 Likewise, in studies showing evidence of durable change in specific measures of personal culture, the change is often small.
-For example, Brocic and Miles [-@brocic2021] estimate that completing graduate degrees in humanities, arts, and social sciences shift peoples’ moral relativism only about .2 standard deviations on average compared to people with no college degree, and this is the largest effect they identify.
+For example, Brocic and Miles [-@brocic2021] estimate that completing graduate degrees in humanities, arts, and social sciences shift peoples' moral relativism only about .2 standard deviations on average compared to people with no college degree, and this is the largest effect they identify.
 And studies of aggregate change show that, even on items where cohort effects explain more variance than period effects, there is always evidence that some people change over time [@vaisey2016].
@@
 
-Current debates often narrow down to ask whether people *ever* change, which is rarely theoretically interesting since the answer is almost always “yes.” Over sufficient time and in a large enough sample, researchers will likely observe some evidence for durable changes in individuals.
-A lack of evidence for change may simply due to a lack of clarity in the question asked or a low level of resolution in the possible answers, or that the question simply was not asked for long enough or to enough people.
-Conversely, finding evidence of change tells us little about how much change is present in a population and how meaningful that change is for social differences.
+The current debate can be simplified to asking whether people *ever* change.
+This is rarely theoretically interesting since the answer is almost always "yes." Over sufficient time and in a large enough sample, researchers will likely observe some evidence for durable individual change.
+A lack of evidence for change may be due to a poor survey question, a low resolution in response options, or that the question simply was not asked for long enough or to enough people.
+Conversely, finding evidence for some change tells us only that it was "not zero." From that alone, we learn little about how much change has happened in a population.
 
 A more theoretically productive approach begins with a model assuming that during an observed time period people might remain stable, might change a little, or might undergo significant shifts in their personal culture.
-Lersch’s LCAM formally does this by modeling each individual as following their own trajectory over time.
-But rather than debating whether this model fits a particular data set better than a model that assumes that people never change, researchers should quantify individual changes over time and compare those to the differences that exist between people at the panel's start.
-In other words, the focus should shift from asking, “do people change?” to asking, “what are the relative contributions of change and initial differences toward explaining differences in personal culture?”
+Lersch's LCAM does this by modeling each individual as following their own linear trajectory.
+But rather than debating whether this model fits a particular data set better than a model that assumes that people never change, researchers should quantify these individual changes and compare those to the differences that exist between people at the panel's start.
+In other words, the focus should shift from asking, "do people change?" to asking, "what are the relative contributions of change and initial differences for explaining differences in personal culture?"
 
 In this approach, there are two metrics that each reflect a combination of relevant theoretical processes.
-The first metric, interpersonal differences at the start of the survey, reflects the accumulated experiences of people prior to entering the panel survey that shape how they answer questions, what Lersch calls "early imprinting." While commonly associated with experiences during an early formative period that result in settled dispositions, this variation could also reflect experiences that happen at any stage in life as long as they predate the panel and consistently impact subsequent responses.
-For instance, for those people entering the panel post-retirement, this “imprinting” might reflect this pivotal life transition.
+The first metric, interpersonal differences at the start of the survey, reflects the accumulated experiences of people prior to entering the panel survey that shape how they answer questions.
+Lersch calls this "early imprinting." While commonly associated with experiences during a formative period that result in settled dispositions, this variation could also reflect experiences that happen at any stage in life as long as they predate the panel and consistently affect subsequent responses.
+For instance, for those people entering the panel post-retirement, this "imprinting" might reflect this pivotal life transition.
 Consequently, this metric also reflects variation in individuals' social roles or statuses at the start of the panel that were important in shaping their dispositions.
 
-The second metric, the amount of intrapersonal linear change people make during the panel, captures durable changes in personal culture over time, a set of processes collectively called “persistent change” or “adaption” by Lersch and "active updating" by Kiley and Vaisey.
-Lersch attributes these changes to social triggers such as moving into a new environment or adopting new social roles.
+The second metric, the amount of intrapersonal linear change people make during the panel, captures durable changes in personal culture over time, a set of processes collectively called "persistent change" or "adaption" by Lersch and "active updating" by Kiley and Vaisey.
+Lersch attributes these changes to social triggers such as moving into a new environment or adopting new social roles (although he never measures them directly).
 Additionally, they might reflect the diffusion of new cultural forms across social networks, cues from political elites or otherwise culturally influential leaders, the emergence of issues in politics or culture, or large-scale social shifts.
 
-A third metric, often called “residual variance” or “fluctuation,” accounts for the remainder of variance in peoples’ responses over time.
+A third metric, often called "residual variance" or "fluctuation," accounts for the remainder of variance in peoples' responses.
 This non-durable change emerges for a variety of reasons.
-For example, people might not have clear personal cultural dispositions as measured by surveys.
+For example, people might not have clear a clear disposition on a particular item as it is asked.
 Instead, they might internalize a broad set of considerations and construct an opinion in the context of the survey interview, with different considerations coming to the forefront of their cognition each interview [@tourangeau2000; @zaller1992; @feldman1992].
-This variance in considerations can produce variance in responses even if peoples’ “dispositions” --the full set of considerations they could draw on-- remains stable over time.
-Additionally, this variance can also include measurement errors, such as misinterpreted questions, erroneous response selections, or responses getting coded incorrectly.
-While this third process reflects important aspects of personal culture, it does not directly touch on the ongoing debate outlined here.
+Additionally, this variance can also include measurement error, such as misinterpreted questions, erroneous response selections, or responses getting coded incorrectly.
+While this third metric at least partially reflects important processes of personal culture, it does not directly touch on the ongoing debate outlined here.
 Because of this we quantify this component but focus principally on the other two.
@@
 Furthermore, these metrics are not necessarily useful in isolation or without context.
-What is important for the theoretical debate is the relative contribution of these two components in explaining cultural differences in a population over a specific duration of time.
-In other words, when we look at a population of adults over some time frame, is there enough intrapersonal change relative to interpersonal difference to view it as an important explanation for why people demonstrate cultural differences?
+What is important for the theoretical debate is the relative contribution of these two components in explaining cultural differences in a population.
+In other words, when we look at a population of adults over some time frame, is there enough intrapersonal change relative to interpersonal difference to view it as an important explanation for why people differ from each other?
 Only through quantifying the two processes we can get closer to a true answer to this question.
@@
 Yet the true utility of the proposed quantification goes beyond settling past debates; it allows researchers to ask questions that were difficult to explore with previous approaches.
-For example, classifying questions based on whether they show durable change overlooks possibly important differences among those that do.
+For example, classifying survey items based on whether they show durable change overlooks possibly important differences among those that do.
 By quantifying the relative contributions of interpersonal differences and intrapersonal change, researchers can gauge their relative importance in explaining differences among survey items, between groups within a given population, across time, and across societies.
@@
 This expectation should be uncontroversial, regardless of the theoretical process assumed to be most relevant for a particular question.
-Panel surveys do not encompass individuals’ entire lifetimes.
+Panel surveys do not encompass individuals' entire lifetimes.
 Thus, even if intrapersonal change during adulthood is the major driver of cultural differences, the change detected in a panel covering 20 years probably will not fully account for the entire variance of interpersonal differences.
@@
 However, theoretical perspectives differ in how much variance they would expect to be driven by intrapersonal change, enabling quantification to provide a new perspective on existing questions.
-“Formative period” and cohort theories, which emphasize early-life socialization as a strong force in shaping adult personal culture, suggest that the variance from intrapersonal change would still fall short of explaining interpersonal differences, even when we account for the duration of observation.
+"Formative period" and cohort theories, which emphasize early-life socialization as a strong force in shaping adult personal culture, suggest that the variance from intrapersonal change would still fall short of explaining interpersonal differences, even when we account for the duration of observation.
 Conversely, other theoretical perspectives emphasizing the importance of people's contemporaneous social environment for personal culture expect intrapersonal change to account for a large portion of the variance, regardless of the duration.
 
-Third, we expect the variance explained by each process -–interpersonal differences and intrapersonal change-– to depend on specific attributes of the question and panel.
+Third, we expect the variance explained by each process ---interpersonal differences and intrapersonal change--- to depend on specific attributes of the question and panel.
 First, the higher the response resolution for a given question and the larger the sample, the greater the variance explained by intrapersonal change should be, as smaller changes in intrapersonal change will be easier to detect.
@@
 
-# Analytical Strategy
+# Analytic Strategy
 
@@
 
-We use data from 7 nationally representative panel surveys from Australia, Germany, Great Britain, Switzerland, and the United States (summarized in Table 1), combining all the data files used in previous work [@kiley2020; @lersch2023].[^data] These studies cover a long period of time (the range of surveys spans from 1968 to 2021), with 609 personal culture items that capture attitudes, beliefs, values, self-assessments, self-descriptions, and behaviors [@alwin2007].
+We use data from 7 nationally representative panel surveys from Australia, Germany, Great Britain, Switzerland, and the United States (summarized in Table 1), combining all the data files used in previous work [@kiley2020; @lersch2023].[^1] These studies cover a long period of time (the range of surveys spans from 1968 to 2021), with 609 personal culture items that capture attitudes, beliefs, values, self-assessments, self-descriptions, and behaviors [@alwin2007].
 We restricted the sample such that individuals between the ages of 18 and 79 are included without further elimination, and in all surveys, we used all possible cases for which respondents provided responses.
@@
 
+[^1]: For more information on these data sources, see [@data_gss; @data_hilda; @data_bhps; @data_ukhls; @data_soep; @data_shp; @data_psid].
+
+```{=tex}
 \begin{center}
@@
 \end{center}
-
+```
 ## Life Course Adaption Model
 
-We start with Lersch’s [-@lersch2023] Life Course Adaption Model, which formalizes survey responses at time $t$ as a function of individual-level random intercepts and slopes for survey age, what is often called a mixed-effects growth curve model.
+We start with Lersch's [-@lersch2023] Life Course Adaption Model, which formalizes survey responses at time $t$ as a function of individual-level random intercepts and slopes for survey age, what is often called a mixed-effects growth curve model.
 This model formally assumes a set of propositions about change that reflect the theoretical debate to this point.
@@
 Second, it assumes that people change over time, taking the form of random slopes for each respondent as a linear function of time.
-Third, it assumes that people deviate around this baseline randomly over time, reflecting “fluctuation” or short-term non-persistent change.
+Third, it assumes that people deviate around this baseline randomly over time, reflecting "fluctuation" or short-term non-persistent change.
 Finally, it assumes that people draw on earlier time points in reacting to new events.
@@
 y_{i, t} = \beta_0 + u_{0, i} + \beta_{1} x_t + u_{1, i} x_{i, t} + \epsilon_{i, t}
-$$
-$$
+$$ $$
 u_{0,i} \sim \mathcal{N}(0,\sigma_{\mu_0}^2) \\
-$$
-$$
+$$ $$
 u_{1,i} \sim \mathcal{N}(0,\sigma_{u_1}^2)
@@
 We then use these models to derive two measures of systematic variance that correspond to the two collections of theoretical processes outlined above: interpersonal differences at baseline and intrapersonal change over time.
-First, to measure the importance of interpersonal differences, we calculate the sum of the squared residuals derived from measuring each respondent at the midpoint of their age trajectory.[^midpoint] We then divide this value by the total sum of squares and subtract it from from 1, producing a value we denote as V(D), or the proportion of variance attributable to interpersonal differences.
+First, to measure the importance of interpersonal differences, we calculate the sum of the squared residuals derived from measuring each respondent at the midpoint of their age trajectory.[^2] We then divide this value by the total sum of squares and subtract it from from 1, producing a value we denote as V(D), or the proportion of variance attributable to interpersonal differences.
 This number is equivalent to an $R^2$ value generated by using random intercepts for all respondents.
@@
 
-These two numbers reflect the total systematic variance of the model that can be attributed to either differences across people or linear change over the course of the panel.[^explained] Our principal outcome of interest is the proportion of systematic variance accounted for by intrapersonal change, which we calculate as
+[^2]: We use the scale midpoint, rather than the actual difference at baseline, to better capture the differences between people independent of their age trajectories.
+
+These two numbers reflect the total systematic variance of the model that can be attributed to either differences across people or linear change over the course of the panel.[^3] Our principal outcome of interest is the proportion of systematic variance accounted for by intrapersonal change, which we calculate as
+
+[^3]: This does not mean that other factors might not explain additional variance if they were included in the model.
+For example, peoples' responses about religious beliefs and participation might systematically vary by the day of the week or month in which they are surveyed.
+This variation is moved into the "random" or non-systematic component of the model for us.
 
@@
 
+```{=tex}
 \begin{center}
@@
 \end{center}
-
+```
 Across all questions, interpersonal differences explain a much larger share of the systematic variance in responses than intrapersonal change.
@@
 As noted above, the substantive importance of intrapersonal change for understanding cultural differences depends on a range of factors, including how long the panel runs, whether assumptions about linear change hold, and whether the panel is capturing a distinctly turbulent period or a distinctly stable period for the item under question.
-However, if we assume that the period under observation is “typical” for a question –-not a time of extremely heightened (or lowered) sensitivity or change-– then it does not seem realistic for this process of intrapersonal change to explain a large share of interpersonal differences for many of the questions we observe.
+However, if we assume that the period under observation is "typical" for a question ---not a time of extremely heightened (or lowered) sensitivity or change--- then it does not seem realistic for this process of intrapersonal change to explain a large share of interpersonal differences for many of the questions we observe.
 
@@
 \end{center}
-
+```
 Figure 2 shows that the more response options respondents are given and the larger the sample, the larger the values of $\omega$.
@@
 
-To test these competing propositions, we use the models presented above to separately calculate $\omega$ values for people with at least a bachelor's degree and people with less than a bachelor's degree at wave 1[^wave] of the General Social Survey's panels.
+To test these competing propositions, we use the models presented above to separately calculate $\omega$ values for people with at least a bachelor's degree and people with less than a bachelor's degree at wave 1[^4] of the General Social Survey's panels.
 We focus on the GSS because it contains the largest number of questions tapping general political dispositions, which is the domain where education has proven particularly relevant for understanding attitude stability.
@@
 
+[^4]: A small number of respondents report different highest degrees at wave 1, wave 2, and wave 3.
+Some of this is due to measurement error, and some of it is due to a small number of people obtaining a higher degree during the four years of the panel.
+Estimating the panel with highest at wave 1 or highest degree reported across the panel produces functionally identical results.
+
 Figure 3 plots the distribution of differences in $\omega$ values between people with at least a bachelor's degree and people with less than a bachelor's degree at wave 1 of the panel for all 183 GSS items we explored.
@@
 \end{center}
-
+```
 There is a clear pattern in Figure 3: for more than 80 percent of GSS items explored here, intrapersonal change is a larger component of systematic variance for people without a college degree than for people with a college degree (i.e., the difference is negative).
@@
 \end{center}
-
+```
 On all eight questions presented in Figure 4, $\omega$ values are greater for people without a college degree than for people with a college degree, meaning intrapersonal change explains less systematic variance for people with a college degree than for people without one.
@@
 Nearly all questions show evidence of people making durable, intrapersonal change over time, with some showing notably high intrapersonal change compared to interpersonal differences.
-For some questions about factors that affect one’s life satisfaction or views on government benefits, it seems plausible that differences in adult experiences predominantly account for the observed differences between people.
+For some questions about factors that affect one's life satisfaction or views on government benefits, it seems plausible that differences in adult experiences predominantly account for the observed differences between people.
 
@@
 This may reflect specific issues being assessed by survey items using different scales.
-Alternatively, it might imply that when intrapersonal changes occurs, they are often subtle with answers moving from “agree” to “strongly agree,” rather than from “strongly agree” to “strongly disagree.” Such subtle changes require finer response options to be detected.
+Alternatively, it might imply that when intrapersonal changes occurs, they are often subtle with answers moving from "agree" to "strongly agree," rather than from "strongly agree" to "strongly disagree." Such subtle changes require finer response options to be detected.
 
@@
 The approach outlined here generates estimates of how much of the observed variance in a population can be attributed to one of two sets of theoretical processes: intrapersonal change and interpersonal differences at baseline.
-It does not quantify the proportion of people who “change,” nor can we be sure that the amount of interpersonal change we detect is driven by many people making small changes or a few people making large changes.
+It does not quantify the proportion of people who "change," nor can we be sure that the amount of interpersonal change we detect is driven by many people making small changes or a few people making large changes.
 
@@
 This assumption simplifies reality in which change likely also takes non-linear and discontinuous forms.
-People might jump from one “stable” disposition to another or experience a "turning point" in their lives that upsets an otherwise stable trajectory.
+People might jump from one "stable" disposition to another or experience a "turning point" in their lives that upsets an otherwise stable trajectory.
 In parts, our assumption of linearity is a limitation of the data, as most questions are only observed for three to six waves.
@@
 
-We believe the approach outlined here can push past the “needless dichotomy” implicit in the question of whether people change or not and address more theoretically informed questions about the conditions under which the relative importance of intrapersonal change and interpersonal differences in explaining cultural differences shifts.
+We believe the approach outlined here can push past the "needless dichotomy" implicit in the question of whether people change or not and address more theoretically informed questions about the conditions under which the relative importance of intrapersonal change and interpersonal differences in explaining cultural differences shifts.
 Characterizing questions as displaying change or not can only take researchers so far, but the question of whether some questions demonstrate more change than others, or whether some groups are characterized by more stability than others, has the potential to weigh in on a broader range of theoretical debates.
@@
 
-[^data]: For more information on these data sources, see [@data_gss; @data_hilda; @data_bhps; @data_ukhls; @data_soep; @data_shp; @data_psid].
-
-[^midpoint]: We use the scale midpoint, rather than the actual difference at baseline, to better capture the differences between people independent of their age trajectories.
-
-[^explained]: This does not mean that other factors might not explain additional variance if they were included in the model.
-For example, peoples' responses about religious beliefs and participation might systematically vary by the day of the week or month in which they are surveyed.
-This variation is moved into the "random" or non-systematic component of the model for us.
-
-[^wave]: A small number of respondents report different highest degrees at wave 1, wave 2, and wave 3.
-Some of this is due to measurement error, and some of it is due to a small number of people obtaining a higher degree during the four years of the panel.
-Estimating the panel with highest at wave 1 or highest degree reported across the panel produces functionally identical results.
-
 # References
 
-<div id="refs"></div>
+::: {#refs}
+:::
 
@@
 ```
+
 ```{=latex}
@@
 ```
+\newpage
 
-\newpage
 ```{=latex}
@@
 \caption{Regression Model Estimating $\omega$}
-
 ```

### 2023-09-27 manuscript/manuscript.tex
first SV edits

@@
 
-Does personal culture --an individual's attitudes, beliefs, values, and
-practices--- change over the life course, or is it largely fixed once a
-person reaches adulthood?
-
-This question underlies several contemporary debates in sociology (Kiley
-and Vaisey 2020; Lersch 2023; Lizardo 2017). For example, pragmatist
-theories of action claim that changes in social environments cause
-people to adapt their views and make new cultural meanings (Gross 2009;
-Swidler 2001), while Bourdieusian practice theories argue that the
-``past conditions of production'' leave a mark on people's personal
-culture that lasts throughout their lives (Bourdieu 1990). Similarly,
-models of social influence assume that people adapt their culture in the
-face of new information, while the emphasis on cohort effects in models
-of aggregate social and cultural change requires them to be open to
-change while young but become fairly resistant to it as they age (Ryder
-1965). Finally, life course theories posit slower but still important
-changes over time as people advance through important transitions in
-their lives (Bardi et al. 2009; Elder, Johnson, and Crosnoe 2003).
-
-Because all of the above theoretical perspectives have some empirical
-support, the current debate is not about whether they exist, but about
-their contribution to explaining cultural differences. That said,
-scholars have struggled to reach a consensus on the relative importance
-of change during adulthood. Over the span of a few years, most apparent
-changes in people's responses to personal survey items appear to be
-transitory, with little evidence of persistent change among adults
-(Kiley and Vaisey 2020; Vaisey and Kiley 2021). This suggests that
-change during adulthood is not a major factor in explaining personal
-cultural differences. When looked at over longer time horizons, however,
-there is at least some evidence that adults make persistent cultural
-changes (Lersch 2023).
+Does personal culture---an individual's attitudes, beliefs, values, and
+practices---change over the life course, or is it largely fixed by
+adulthood? This question underlies an important contemporary debate in
+sociology (Kiley and Vaisey 2020; Lersch 2023; Lizardo 2017) and has
+deep roots in seemingly contradictory theoretical perspectives. For
+example, pragmatist theories of action claim that changes in social
+environments cause people to adapt their views and make new cultural
+meanings (Gross 2009; Swidler 2001), while Bourdieusian practice
+theories argue that the ``past conditions of production'' leave a mark
+on people's personal culture that lasts throughout their lives (Bourdieu
+1990). Models of social influence assume that people adapt their culture
+in the face of new information (Goldberg and Stein 2018), while the
+emphasis on cohort effects in models of aggregate social and cultural
+change requires them to be open to change while young but become fairly
+resistant to it as they age (Ryder 1965). Finally, life course theories
+posit important changes over time as people advance through important
+transitions in their lives (Bardi et al. 2009; Elder, Johnson, and
+Crosnoe 2003).
+
+Because all of these theoretical perspectives have some empirical
+support, the current debate is not about whether their posited processes
+exist at all, but about their relative contribution to explaining the
+cultural differences we see in the world. That said, scholars have
+struggled to reach a consensus on the importance of change during
+adulthood. Over the span of a few years, most apparent changes in
+people's responses to personal survey items appear to be transitory,
+with little evidence of persistent change among adults (Kiley and Vaisey
+2020; Vaisey and Kiley 2021). This suggests that change during adulthood
+is not a major factor in explaining personal cultural differences. When
+we consider a longer time horizon, however, there is evidence that
+adults make at least some persistent changes (Lersch 2023).
 
 These seemingly inconsistent findings are partly due to the fact that
-researchers have taken a ``tournament of models'' approach in
+researchers have taken a ``tournament of models'' approach to
 adjudicating different theoretical perspectives (Lersch 2023: p.~228).
-That is, researchers have compared models of responses to cultural items
-over time as consistent with either ``change'' or ``no change,'' which
-essentially asks whether we observe \emph{any} intrapersonal change in
-personal culture.
-
-We argue that asking only \emph{whether} people change obscures what are
-actually highly compatible empirical patterns. Being able to detect
-change in a given dataset depends not only on sample populations, item
-measurement, statistical methods, and power, but also on definitions of
-what counts as ``change.'' These methodological decisions potentially
-limit the possibility of convergence on the theoretical debates.
-Moreover, asking whether there is evidence of change limits the kind of
-questions that can be asked to, ``Do we see change in this measure?''
-But such inquiries are ill-suited to address the core of the theoretical
-debate: determining \emph{how much} the process of intrapersonal change
-contributes to cultural differences.
+That is, researchers have compared the fit of different statistical
+models to each item to determine whether it is more consistent with ``no
+change'' or ``any change.'' The size of these piles then serves as the
+primary evidence for the truth of the theory (Kiley and Vaisey 2020;
+Lersch 2023; Vaisey and Kiley 2021).
+
+We argue, however, that asking \emph{whether} people change on an item
+may cause more problems than it solves. Being able to detect \emph{any}
+change on an item depends not only on sample populations, measurement,
+statistical methods, and power, but also on definitions of what counts
+as ``change.'' These differences can limit the possibility of agreement
+on which ``pile'' a variable belongs in. Moreover, asking whether there
+is evidence of change reduces to, ``Is there \emph{any} evidence of
+change in this item?'' But such binary inquiries are ill-suited to
+address the core of the theoretical debate: determining \emph{how much}
+the process of intrapersonal change contributes to cultural differences.
 
 There is enough evidence to assert that neither pre-adult socialization
-nor the contemporary environment alone can fully explain cultural
-differences. The question that remains is about the relative
-contributions of these processes. In the current debate, for a single
-construct (e.g., support for gay rights), existing approaches can only
-make claims about \emph{whether} intrapersonal change is happening, not
-about how much. This limitation also prevents investigating under what
-conditions, across which groups, and in what domains do we see
-differences in the relative importance of intrapersonal change. These
-questions are about \emph{degree}, not existence, and thus require
-precise measurement rather than declarations of victory for a particular
-perspective.
-
-In this paper, we intervene in this debate by moving beyond the
-``tournament of models'' approach toward a quantification of the
-processes that lead to differences in personal culture, and introducing
-a method for quantifying the relative contributions of interpersonal
-differences and intrapersonal change to the patterns we see in empirical
-data. Drawing on the seven panel datasets analyzed in the debate, we use
-established methods to quantify the amount of variance in responses that
-is due to between-people differences at a single point during the survey
-versus the amount of variance in responses that is due to linear
-intrapersonal change over time. This measure gives us a sense of the
-importance of change over the duration of the panel relative to the
-differences that exist between people assuming they never changed.
-
-We observe similar patterns across the datasets that yielded previously
-conflicting interpretations. In all datasets we find a similar
-distribution of how much explanatory power can be attributed to
-intrapersonal change. But quantifying these processes also reveals that
-intrapersonal change plays a substantially smaller role in explaining
-the variance in peoples' personal culture than interpersonal differences
-at their baselines. Whether the amount of change is large enough to be
-``meaningful'' ultimately remains a theoretical question reflected in
-prior expectations about particular cultural item and population. It
-also crucially depends on how long people are observed in a panel.
-Nevertheless, on many measures of personal culture we do not observe
-enough intrapersonal change, in some cases over more than 10 years, for
-it to play a substantial role in explaining the differences we observe
-between adults, though notable exceptions exist.
-
-These results help clarify and, hopefully, transcend previous debates
-about whether adults change. What is more, the approach we propose can
-be applied to a broader range of theoretical questions than approaches
-discerning whether durable change exists in a panel. To exemplify this,
-we explore differences in the explanatory power of intrapersonal change
-for people with college degrees and those without. We find that change
-has less explanatory power for people with college degrees than those
-without. This may imply that the college experience solidifies political
-dispositions rather than fosters an openness and adaptability to new
-information and influences later in life. This offers a novel, empirical
-basis to theorize about the role of college in shaping personal culture.
-We believe that this approach has the potential to advance the
-theoretical debate on the role of environments, social identities, and
-roles in facilitating personal cultural change and stability.
+nor the contemporary social context alone can explain the full range of
+cultural differences. The question that remains is about the
+\emph{relative} contributions of these processes. In the current debate,
+for a single construct (e.g., support for gay rights), existing
+approaches can only make claims about \emph{whether} intrapersonal
+change is happening, not about how much. This also prevents
+investigating under what conditions, across which groups, and in what
+domains do we see differences in the relative importance of
+intrapersonal change and pre-existing interpersonal differences. These
+questions are about \emph{degree}, not existence; taking this debate
+forward will require precise measurement rather than declarations of
+victory for a particular perspective.
+
+In this paper, we move beyond the ``tournament of models'' approach and
+introduce a method for quantifying the relative contributions of
+interpersonal differences and intrapersonal change for a single item.
+Drawing on seven panel data sets from five countries, we quantify the
+amount of variance in survey responses that is due to between-person
+differences at a single point during the survey versus the amount of
+variance in responses that is due to systematic intrapersonal change
+over the study period. This measure quantifies the amount of change over
+the duration of the panel relative to the stable differences that exist
+between people.
+
+Using this measure, we observe similar patterns across the datasets that
+have previously yielded conflicting interpretations. In all
+datasets---despite their varied social contexts and duration---we find a
+similar distribution of intrapersonal change. In general, we find that
+intrapersonal change accounts for a substantially smaller amount of
+variance in personal culture survey item than stable interpersonal
+differences. Of course, whether the amount of intrapersonal change in an
+item is large enough to be ``meaningful'' is a substantive and
+theoretical question. Nevertheless, on many measures of personal culture
+(some observed for over a decade), we do not observe enough
+intrapersonal change for it to play a substantial role in explaining the
+differences we observe between adults. Notable, interesting, and
+important exceptions exist, of course, but they are departures from the
+overall pattern.
+
+Our method can also help address substantive questions about personal
+change. To illustrate this, we investigate differences in the relative
+importance of intrapersonal change on political items for people with
+and without a college degree. We find that change is less pronounced for
+college-educated respondents, suggesting that college solidifies
+political dispositions rather than fostering an openness to new
+information. This offers a novel empirical basis to theorize about the
+role of college in shaping personal culture . In summary, we believe
+that this approach has the potential to advance the theoretical debate
+on the role of environments, social identities, and roles in
+facilitating personal cultural change and stability.
+
+Taken together, to goal of this paper is to advance---and, hopefully,
+transcend---debates about \emph{whether} adults change. They do change,
+at least a little, on most things. Our method can help quantify exactly
+how much.
 
@@
 
-Third, we expect the variance explained by each process
--\/--interpersonal differences and intrapersonal change-\/-- to depend
-on specific attributes of the question and panel. First, the higher the
-response resolution for a given question and the larger the sample, the
-greater the variance explained by intrapersonal change should be, as
-smaller changes in intrapersonal change will be easier to detect.
-Second, the LCAM and various theoretical perspectives that underlie it
-posit that social transitions drive changes in personal culture. Since
-the probability that any person experiences such a transition goes up
-the longer we observe them, we expect panel duration to be positively
-related to the proportion of variation explained by intrapersonal
-change.
+Third, we expect the variance explained by each process ---interpersonal
+differences and intrapersonal change--- to depend on specific attributes
+of the question and panel. First, the higher the response resolution for
+a given question and the larger the sample, the greater the variance
+explained by intrapersonal change should be, as smaller changes in
+intrapersonal change will be easier to detect. Second, the LCAM and
+various theoretical perspectives that underlie it posit that social
+transitions drive changes in personal culture. Since the probability
+that any person experiences such a transition goes up the longer we
+observe them, we expect panel duration to be positively related to the
+proportion of variation explained by intrapersonal change.
 
@@
 cross-domain set of items to capture a broad range of individual
-personal culture. Appendix A documents the list of all variables used in
-the upcoming analyses.
+personal culture. Supplemental Materials A documents the list of all
+variables used in the upcoming analyses.
 
@@
 for a question ---not a time of extremely heightened (or lowered)
-sensitivity or change-\/-- then it does not seem realistic for this
+sensitivity or change--- then it does not seem realistic for this
 process of intrapersonal change to explain a large share of
@@
 National{ö}konomie Und Statistik} 239(2):345--60.
+
+\leavevmode\vadjust pre{\hypertarget{ref-goldberg2018}{}}%
+Goldberg, Amir, and Sarah K. Stein. 2018. {``Beyond {Social}
+{Contagion}: {Associative} {Diffusion} and the {Emergence} of {Cultural}
+{Variation}.''} \emph{American Sociological Review} 83(5):897--932. doi:
+\href{https://doi.org/10.1177/0003122418797576}{10.1177/0003122418797576}.
 
@@
 \textit{Notes:} The figure shows $\omega$ and 1 - $\omega$ as the proportion of systematic variance attributable to intrapersonal change and interpersonal differences, respectively.
-See Appendix A for the full set of item values.
+See Supplemental Materials A for the full set of item values.
 \end{figure}

### 2023-09-27 supplementals/manuscript_supplementals.Rmd
supplemental materials

---
title: |
  | Supplemental Materials for
  | Quantifying Stability and Change in Personal Culture Using Panel Data
author: 
- 
output:
  pdf_document:
    keep_tex: true
    latex_engine: xelatex
    toc: true
geometry: margin=1in 
mainfont: Times New Roman
fontsize: 12pt
header-includes:
  \usepackage{fontspec}
  \usepackage{longtable}
  \usepackage{multirow}
  \usepackage{array}
  \usepackage{booktabs}
  \usepackage{dcolumn}
  \usepackage{lscape}
  \usepackage{caption}
  \newcommand{\blandscape}{\begin{landscape}}
  \newcommand{\elandscape}{\end{landscape}}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F, message = F, warning = F, fig.align = "center")
```

\newpage
\blandscape

# Supplemental Materials A

In Table A1, we present the survey items, corresponding labels and topic indicators, as well as the variances calculated from our estimations: V(C) refers to intraindividual change, V(D) refers to interindividual differences, and V(R) refers to residual variance. We also present the $\omega$ values.

```{=latex}
\begin{scriptsize}
\begin{longtable}{|p{1.75in}|p{0.3in}|p{0.3in}|p{0.3in}|p{0.3in}|p{2.5in}|p{2.5in}}
\caption*{Table A1: The Descriptives of Variables Used}
\\
\toprule
\multicolumn{1}{l}{} & \emph{\textbf{ω}} & \textbf{V(C)} & \textbf{V(D)} & \textbf{V(R)} & \textbf{Label} & \textbf{Topic} \\ 
\midrule
\multicolumn{7}{l}{\textbf{BHPS}} \\ 
\midrule
cohadv & 0.09 & 0.05 & 0.53 & 0.42 & Advantages In Living As Couple & Gender \& Family \\ 
cohdis & 0.00 & 0.00 & 0.57 & 0.43 & Disadvantages In Living As Couple & Gender \& Family \\ 
crwora & 0.06 & 0.04 & 0.60 & 0.37 & Worry About Being Affected By Crime & Social Life, Social Cohesion \& Trust \\ 
crworb & 0.10 & 0.07 & 0.61 & 0.32 & Extent Of Concern About Crime & Social Life, Social Cohesion \& Trust \\ 
dfwld & 0.03 & 0.02 & 0.54 & 0.44 & Different World In Future For Young & Gender \& Family \\ 
lfimpa & 0.13 & 0.08 & 0.59 & 0.32 & Importance Of: Health & Health \& Morale \\ 
lfimpb & 0.19 & 0.16 & 0.67 & 0.16 & Importance Of: Money & Health \& Morale \\ 
lfimpc & 0.10 & 0.09 & 0.79 & 0.13 & Importance Of: Having Children & Health \& Morale \\ 
lfimpd & 0.20 & 0.15 & 0.60 & 0.25 & Importance Of: Having A Good Job & Health \& Morale \\ 
lfimpe & 0.16 & 0.12 & 0.63 & 0.25 & Importance Of: Being Independent & Health \& Morale \\ 
lfimpf & 0.19 & 0.17 & 0.70 & 0.13 & Importance Of: Owning Own Home & Health \& Morale \\ 
lfimpg & 0.14 & 0.11 & 0.69 & 0.19 & Importance Of: Good Partnership & Health \& Morale \\ 
lfimph & 0.20 & 0.16 & 0.66 & 0.17 & Importance Of: Good Friends & Health \& Morale \\ 
opchd1\_1 & 0.02 & 0.01 & 0.48 & 0.51 & Quality Best Prep. Child: Well Liked & Gender \& Family \\ 
opchd1\_2 & 0.02 & 0.01 & 0.61 & 0.38 & Quality Best Prep. Child: Think For Self & Gender \& Family \\ 
opchd1\_3 & 0.14 & 0.09 & 0.53 & 0.38 & Quality Best Prep. Child: Work Hard & Gender \& Family \\ 
opchd1\_4 & 0.00 & 0.00 & 0.54 & 0.46 & Quality Best Prep. Child: Help Others & Gender \& Family \\ 
opchd1\_5 & 0.01 & 0.00 & 0.58 & 0.41 & Quality Best Prep. Child: Obey Parents & Gender \& Family \\ 
opcls1 & 0.16 & 0.11 & 0.55 & 0.34 & Sees Self As Belonging To Social Class & Subjective Ses \\ 
opcls2 & 0.10 & 0.08 & 0.73 & 0.19 & Subjective Social Class Membership & Subjective Ses \\ 
opcls3 & 0.16 & 0.10 & 0.52 & 0.38 & Social Class Affects Opportunities & Subjective Ses \\ 
opeur1 & 0.13 & 0.11 & 0.70 & 0.19 & Opinion: Uk Membership Of Eu & Politics, Government \& Economy \\ 
opeur2 & 0.13 & 0.10 & 0.71 & 0.19 & Uk Benefited From Being In Eu & Politics, Government \& Economy \\ 
opeur3 & 0.05 & 0.03 & 0.65 & 0.32 & Uk Long Term Policy Re. Eu Should Be & Politics, Government \& Economy \\ 
opfama & 0.10 & 0.07 & 0.63 & 0.30 & Pre-School Child Suffers If Mother Works & Gender \& Family \\ 
opfamb & 0.09 & 0.07 & 0.64 & 0.29 & Family Suffers If Mother Works Full-Time & Gender \& Family \\ 
opfamc & 0.14 & 0.07 & 0.46 & 0.47 & Woman And Family Happier If She Works & Gender \& Family \\ 
opfamd & 0.12 & 0.07 & 0.54 & 0.39 & Husband And Wife Should Both Contribute & Gender \& Family \\ 
opfame & 0.12 & 0.07 & 0.53 & 0.40 & Full Time Job Makes Woman Independent & Gender \& Family \\ 
opfamf & 0.07 & 0.05 & 0.68 & 0.27 & Husband Should Earn, Wife Stay At Home & Gender \& Family \\ 
opfamg & 0.12 & 0.07 & 0.52 & 0.42 & Children Need Father As Much As Mother & Gender \& Family \\ 
opfamh & 0.11 & 0.07 & 0.59 & 0.34 & Employers Should Help With Childcare & Gender \& Family \\ 
opfami & 0.07 & 0.05 & 0.66 & 0.29 & Single Parents Are As Good As Couples & Gender \& Family \\ 
opfamj & 0.15 & 0.13 & 0.75 & 0.12 & Cohabiting Is Always Wrong & Gender \& Family \\ 
opfamk & 0.11 & 0.06 & 0.52 & 0.42 & Adult Children Should Care For Parents & Gender \& Family \\ 
opfaml & 0.11 & 0.07 & 0.56 & 0.37 & Divorce Better Than Unhappy Marriage & Gender \& Family \\ 
opfamm & 0.12 & 0.11 & 0.77 & 0.13 & The Man Should Be Head Of Household & Gender \& Family \\ 
opfamn & 0.08 & 0.07 & 0.82 & 0.11 & Bible Gods Word And True & Religion \& Spirituality \\ 
opfamo & 0.07 & 0.06 & 0.76 & 0.19 & Co-Habiting Is Alright & Gender \& Family \\ 
opfamp & 0.11 & 0.08 & 0.64 & 0.28 & Parents Ought Stay Together For Children & Gender \& Family \\ 
opfamq & 0.11 & 0.07 & 0.61 & 0.32 & Marital Status Is Irrelevant To Children & Gender \& Family \\ 
opfamr & 0.05 & 0.04 & 0.76 & 0.19 & Homosexual Relationships Are Wrong & Gender \& Family \\ 
ophla & 0.15 & 0.09 & 0.54 & 0.37 & All Health Care Should Be Free & Politics, Government \& Economy \\ 
ophlb & 0.12 & 0.08 & 0.59 & 0.34 & Compulsory Private Insurance If Can Pay & Politics, Government \& Economy \\ 
ophlc & 0.17 & 0.11 & 0.52 & 0.38 & Unfair That Wealth Buys Medical Priority & Politics, Government \& Economy \\ 
opiss1 & 0.07 & 0.05 & 0.68 & 0.28 & Rising Prices Are A Cause For Concern & Politics, Government \& Economy \\ 
opiss2 & 0.05 & 0.03 & 0.71 & 0.26 & Ozone Layer Is A Cause For Concern & Environment \& Climate \\ 
opiss3 & 0.10 & 0.07 & 0.62 & 0.31 & Unemployment Is A Cause For Concern & Politics, Government \& Economy \\ 
opiss4 & 0.01 & 0.00 & 0.71 & 0.29 & Threat To Nature Is A Cause For Concern & Environment \& Climate \\ 
opiss5 & 0.08 & 0.06 & 0.70 & 0.24 & Moral Standards Are A Cause For Concern & Gender \& Family \\ 
opnata & 0.18 & 0.13 & 0.56 & 0.31 & Britain Has Much To Learn & Politics, Government \& Economy \\ 
opnatb & 0.14 & 0.11 & 0.63 & 0.26 & British Citizenship Best & Politics, Government \& Economy \\ 
opnatc & 0.15 & 0.10 & 0.55 & 0.36 & Some Shameful Things About Britain & Politics, Government \& Economy \\ 
opnatd & 0.24 & 0.17 & 0.52 & 0.32 & People Too Ready To Criticise Britain & Politics, Government \& Economy \\ 
opnate & 0.17 & 0.12 & 0.60 & 0.28 & Britain Should Stay Single State & Politics, Government \& Economy \\ 
opnatf & 0.12 & 0.08 & 0.62 & 0.30 & Co-Operation Better Than Independence & Politics, Government \& Economy \\ 
opngbha & 0.17 & 0.13 & 0.65 & 0.22 & Belong To Neighbourhood & Social Life, Social Cohesion \& Trust \\ 
opngbhb & 0.20 & 0.16 & 0.66 & 0.18 & Local Friends Mean A Lot & Social Life, Social Cohesion \& Trust \\ 
opngbhc & 0.15 & 0.11 & 0.63 & 0.26 & Advice Obtainable Locally & Social Life, Social Cohesion \& Trust \\ 
opngbhd & 0.21 & 0.16 & 0.63 & 0.21 & Can Borrow Things From Neighbours & Social Life, Social Cohesion \& Trust \\ 
opngbhe & 0.20 & 0.15 & 0.59 & 0.26 & Willing To Improve Neighbourhood & Social Life, Social Cohesion \& Trust \\ 
opngbhf & 0.10 & 0.07 & 0.59 & 0.34 & Plan To Stay In Neighbourhood & Social Life, Social Cohesion \& Trust \\ 
opngbhg & 0.18 & 0.14 & 0.66 & 0.20 & Am Similar To Others In Neighbourhood & Social Life, Social Cohesion \& Trust \\ 
opngbhh & 0.19 & 0.15 & 0.65 & 0.20 & Talk Regularly To Neighbours & Social Life, Social Cohesion \& Trust \\ 
oppola & 0.19 & 0.11 & 0.48 & 0.41 & Government Reflects People's Wishes & Politics, Government \& Economy \\ 
oppolb & 0.17 & 0.09 & 0.47 & 0.44 & People Can't Influence Govt. Policy & Politics, Government \& Economy \\ 
oppolc & 0.06 & 0.04 & 0.60 & 0.36 & Should Be Max. Limit On Income & Politics, Government \& Economy \\ 
oppold & 0.18 & 0.10 & 0.47 & 0.43 & Govt. Puts Nation's Interests First & Politics, Government \& Economy \\ 
oprlg3 & 0.08 & 0.06 & 0.76 & 0.17 & Religion Makes Difference To Life & Religion \& Spirituality \\ 
opsoca & 0.13 & 0.08 & 0.54 & 0.38 & Ordinary People Share Nations Wealth & Politics, Government \& Economy \\ 
opsocb & 0.11 & 0.07 & 0.58 & 0.34 & One Law For Rich And One For Poor & Politics, Government \& Economy \\ 
opsocc & 0.09 & 0.06 & 0.58 & 0.36 & Private Enterprise Solves Economic Probs & Politics, Government \& Economy \\ 
opsocd & 0.12 & 0.07 & 0.53 & 0.40 & Public Services Ought To Be State Owned & Politics, Government \& Economy \\ 
opsoce & 0.10 & 0.07 & 0.64 & 0.29 & Govt. Has Obligation To Provide Jobs & Politics, Government \& Economy \\ 
opsocf & 0.09 & 0.06 & 0.65 & 0.28 & Strong Trade Unions Protect Employees & Politics, Government \& Economy \\ 
trust & 0.12 & 0.08 & 0.61 & 0.31 & Trustworthiness Of Others & Social Life, Social Cohesion \& Trust \\ 
vote1 & 0.10 & 0.06 & 0.54 & 0.40 & Supports A Particular Political Party & Politics, Government \& Economy \\ 
vote2 & 0.14 & 0.07 & 0.44 & 0.50 & Closer To One Political Party Than Other & Politics, Government \& Economy \\ 
vote4\_1 & 0.05 & 0.04 & 0.88 & 0.08 & Party Affinity: Conservative & Politics, Government \& Economy \\ 
vote4\_2 & 0.05 & 0.04 & 0.85 & 0.11 & Party Affinity: Labour & Politics, Government \& Economy \\ 
vote4\_3 & 0.08 & 0.07 & 0.72 & 0.21 & Party Affinity: Liberal Democrat/Lib/Sdp & Politics, Government \& Economy \\ 
vote5 & 0.10 & 0.06 & 0.54 & 0.41 & Strength Of Support For Stated Party & Politics, Government \& Economy \\ 
vote6 & 0.06 & 0.04 & 0.70 & 0.25 & Level Of Interest In Politics & Politics, Government \& Economy \\ 
\midrule
\multicolumn{7}{l}{\textbf{GSS}} \\ 
\midrule
abany & 0.03 & 0.02 & 0.76 & 0.22 & Abortion: Any Reason & Gender \& Family \\ 
abdefect & 0.10 & 0.08 & 0.75 & 0.17 & Abortion: Birth Defect & Gender \& Family \\ 
abhlth & 0.14 & 0.12 & 0.72 & 0.16 & Abortion: Mother's Health & Gender \& Family \\ 
abnomore & 0.10 & 0.08 & 0.76 & 0.16 & Abortion: Birth Control & Gender \& Family \\ 
abpoor & 0.09 & 0.08 & 0.77 & 0.15 & Abortion: Too Poor & Gender \& Family \\ 
abrape & 0.12 & 0.11 & 0.77 & 0.13 & Abortion: Rape & Gender \& Family \\ 
abscale6 & 0.08 & 0.07 & 0.84 & 0.09 & Abortion Scale (6 Items) & Gender \& Family \\ 
abscale7 & 0.07 & 0.06 & 0.84 & 0.10 & Abortion Scale (7 Items) & Gender \& Family \\ 
absingle & 0.05 & 0.04 & 0.77 & 0.19 & Abortion: Unmarried & Gender \& Family \\ 
affrmact & 0.12 & 0.09 & 0.68 & 0.22 & Hiring Preferences & National Identity, Ethnicity \& Immigration \\ 
aged & 0.01 & 0.00 & 0.60 & 0.40 & Seniors Should Live With Family & Gender \& Family \\ 
athscale & 0.00 & 0.00 & 0.74 & 0.26 & Athiest: Scale & Politics, Government \& Economy \\ 
attend & 0.08 & 0.08 & 0.85 & 0.07 & Attend Religious Services & Religion \& Spirituality \\ 
bible & 0.10 & 0.09 & 0.80 & 0.12 & View Of Bible & Religion \& Spirituality \\ 
cappun & 0.12 & 0.11 & 0.76 & 0.13 & Death Penalty & Politics, Government \& Economy \\ 
chldidel & 0.15 & 0.12 & 0.71 & 0.17 & Ideal \# Of Children & Gender \& Family \\ 
class & 0.09 & 0.07 & 0.73 & 0.20 & Social Class & Subjective Ses \\ 
closeblk & 0.11 & 0.09 & 0.71 & 0.20 & Feel Close To Blacks & National Identity, Ethnicity \& Immigration \\ 
closewht & 0.04 & 0.03 & 0.61 & 0.36 & Feel Close To Whites & National Identity, Ethnicity \& Immigration \\ 
colath & 0.00 & 0.00 & 0.64 & 0.36 & Athiest: Teach At College & Politics, Government \& Economy \\ 
colcom & 0.09 & 0.06 & 0.63 & 0.31 & Red: Teach At College & Politics, Government \& Economy \\ 
colhomo & 0.00 & 0.00 & 0.67 & 0.32 & Gay: Teach At College & Politics, Government \& Economy \\ 
colmil & 0.03 & 0.02 & 0.61 & 0.37 & Militarist: Teach At College & Politics, Government \& Economy \\ 
colmslm & 0.00 & 0.00 & 0.70 & 0.30 & Islamist: Teach At College & Politics, Government \& Economy \\ 
colrac & 0.00 & 0.00 & 0.60 & 0.40 & Racist: Teach At College & Politics, Government \& Economy \\ 
comscale & 0.03 & 0.02 & 0.76 & 0.22 & Red: Scale & Politics, Government \& Economy \\ 
conarmy & 0.12 & 0.08 & 0.63 & 0.29 & Military & Social Life, Social Cohesion \& Trust \\ 
conbus & 0.11 & 0.07 & 0.60 & 0.33 & Major Companies & Social Life, Social Cohesion \& Trust \\ 
conclerg & 0.05 & 0.03 & 0.66 & 0.31 & Organized Religion & Social Life, Social Cohesion \& Trust \\ 
coneduc & 0.02 & 0.01 & 0.60 & 0.38 & Education & Social Life, Social Cohesion \& Trust \\ 
confed & 0.24 & 0.17 & 0.56 & 0.27 & Executive Branch & Social Life, Social Cohesion \& Trust \\ 
confinan & 0.10 & 0.06 & 0.56 & 0.38 & Banks \& Finance & Social Life, Social Cohesion \& Trust \\ 
conjudge & 0.09 & 0.06 & 0.63 & 0.31 & Supreme Court & Social Life, Social Cohesion \& Trust \\ 
conlabor & 0.14 & 0.10 & 0.62 & 0.28 & Organized Labor & Social Life, Social Cohesion \& Trust \\ 
conlegis & 0.02 & 0.01 & 0.61 & 0.37 & Congress & Social Life, Social Cohesion \& Trust \\ 
conmedia & 0.17 & 0.13 & 0.66 & 0.21 & Tv \& Press & Social Life, Social Cohesion \& Trust \\ 
conmedic & 0.13 & 0.09 & 0.59 & 0.32 & Medicine & Social Life, Social Cohesion \& Trust \\ 
conpress & 0.16 & 0.12 & 0.62 & 0.26 & Press & Social Life, Social Cohesion \& Trust \\ 
consci & 0.04 & 0.02 & 0.61 & 0.36 & Scientific Community & Social Life, Social Cohesion \& Trust \\ 
contv & 0.18 & 0.13 & 0.59 & 0.27 & Television & Social Life, Social Cohesion \& Trust \\ 
courts & 0.06 & 0.04 & 0.65 & 0.32 & Courts Deal With Criminals & Politics, Government \& Economy \\ 
discaff & 0.00 & 0.00 & 0.56 & 0.44 & Affirm. Action Hurt Whites & National Identity, Ethnicity \& Immigration \\ 
discaffm & 0.09 & 0.06 & 0.55 & 0.40 & Affirm. Action Hurt Men & Gender \& Family \\ 
discaffw & 0.03 & 0.02 & 0.56 & 0.42 & Affirm. Action Hurt Women & Gender \& Family \\ 
divlaw & 0.00 & 0.00 & 0.71 & 0.29 & Divorce Laws Too Lenient & Politics, Government \& Economy \\ 
eqwlth & 0.05 & 0.04 & 0.68 & 0.28 & Government Reduce Inequality & Politics, Government \& Economy \\ 
fair & 0.04 & 0.03 & 0.68 & 0.30 & People Try To Be Fair & Social Life, Social Cohesion \& Trust \\ 
fear & 0.06 & 0.04 & 0.66 & 0.30 & Afraid To Walk At Night & Politics, Government \& Economy \\ 
fechld & 0.13 & 0.10 & 0.65 & 0.25 & Working Moms Warm & Gender \& Family \\ 
fefam & 0.02 & 0.01 & 0.69 & 0.30 & Better For Man To Work & Gender \& Family \\ 
fehire & 0.00 & 0.00 & 0.62 & 0.38 & Hire \& Promote Women & Gender \& Family \\ 
fejobaff & 0.16 & 0.13 & 0.69 & 0.18 & Preferences For Hiring Women & Gender \& Family \\ 
fepol & 0.00 & 0.00 & 0.64 & 0.36 & Men Better For Politics & Gender \& Family \\ 
fepresch & 0.02 & 0.01 & 0.66 & 0.33 & Pre-Ks Suffer If Mom Works & Gender \& Family \\ 
finalter & 0.20 & 0.14 & 0.54 & 0.32 & Finances Better Or Worse & Subjective Ses \\ 
finrela & 0.10 & 0.08 & 0.70 & 0.22 & Relative Income & Subjective Ses \\ 
fund & 0.08 & 0.07 & 0.84 & 0.09 & Fundamentalist & Religion \& Spirituality \\ 
genderscale & 0.03 & 0.02 & 0.76 & 0.22 & Scale Of Gender Role Attitudes & Gender \& Family \\ 
getahead & 0.19 & 0.13 & 0.56 & 0.31 & Work To Get Ahead & Subjective Ses \\ 
god & 0.05 & 0.04 & 0.82 & 0.14 & Belief About God & Religion \& Spirituality \\ 
goodlife & 0.05 & 0.03 & 0.60 & 0.37 & Finances Will Improve & Subjective Ses \\ 
grass & 0.03 & 0.03 & 0.78 & 0.20 & Legalize Marijuana & Politics, Government \& Economy \\ 
gunlaw & 0.00 & 0.00 & 0.63 & 0.37 & Require Gun Permits & Politics, Government \& Economy \\ 
hapmar & 0.08 & 0.06 & 0.70 & 0.24 & Happiness: Marriage & Health \& Morale \\ 
happy & 0.19 & 0.15 & 0.62 & 0.23 & Happiness: General & Health \& Morale \\ 
health & 0.10 & 0.08 & 0.75 & 0.17 & Health: Self-Rated & Health \& Morale \\ 
helpblk & 0.04 & 0.03 & 0.70 & 0.27 & Government Help Blacks & Politics, Government \& Economy \\ 
helpful & 0.01 & 0.01 & 0.65 & 0.34 & People Try To Be Helpful & Social Life, Social Cohesion \& Trust \\ 
helpnot & 0.00 & 0.00 & 0.62 & 0.37 & Government Do More Or Less & Politics, Government \& Economy \\ 
helpoth & 0.10 & 0.07 & 0.59 & 0.34 & Child Should Help Others & Gender \& Family \\ 
helppoor & 0.01 & 0.01 & 0.65 & 0.35 & Government Help Poor & Politics, Government \& Economy \\ 
helpsick & 0.02 & 0.01 & 0.66 & 0.33 & Government Pay Medical & Politics, Government \& Economy \\ 
homoscale & 0.00 & 0.00 & 0.74 & 0.26 & Gay: Scale & Politics, Government \& Economy \\ 
homosex & 0.04 & 0.04 & 0.83 & 0.13 & Morality: Same-Sex Sex & Gender \& Family \\ 
intlblks & 0.23 & 0.14 & 0.48 & 0.37 & Blacks Intelligent Or Not & National Identity, Ethnicity \& Immigration \\ 
intlwhts & 0.01 & 0.01 & 0.49 & 0.50 & Whites Intelligent Or Not & National Identity, Ethnicity \& Immigration \\ 
jobfind & 0.16 & 0.12 & 0.65 & 0.23 & Find A Good Job & Subjective Ses \\ 
joblose & 0.01 & 0.00 & 0.56 & 0.44 & Likely To Lose Job & Subjective Ses \\ 
kidssol & 0.04 & 0.03 & 0.69 & 0.28 & Mobility Next Generation & Subjective Ses \\ 
letdie1 & 0.14 & 0.12 & 0.72 & 0.17 & Let Die: Incurable & Health \& Morale \\ 
letin1a & 0.02 & 0.01 & 0.70 & 0.29 & How Many Immigrants? & National Identity, Ethnicity \& Immigration \\ 
libath & 0.00 & 0.00 & 0.61 & 0.39 & Athiest: Library Book & Politics, Government \& Economy \\ 
libcom & 0.05 & 0.04 & 0.65 & 0.31 & Red: Library Book & Politics, Government \& Economy \\ 
libhomo & 0.00 & 0.00 & 0.65 & 0.35 & Gay: Library Book & Politics, Government \& Economy \\ 
libmil & 0.02 & 0.01 & 0.60 & 0.38 & Militarist: Library Book & Politics, Government \& Economy \\ 
libmslm & 0.00 & 0.00 & 0.68 & 0.32 & Islamist: Library Book & Politics, Government \& Economy \\ 
librac & 0.00 & 0.00 & 0.59 & 0.40 & Racist: Library Book & Politics, Government \& Economy \\ 
life & 0.04 & 0.03 & 0.67 & 0.30 & Life: Exciting Or Dull & Health \& Morale \\ 
liveblks & 0.20 & 0.14 & 0.58 & 0.27 & Neighborhood Half Black & National Identity, Ethnicity \& Immigration \\ 
livewhts & 0.13 & 0.07 & 0.49 & 0.44 & Neighborhood Half White & National Identity, Ethnicity \& Immigration \\ 
marblk & 0.06 & 0.04 & 0.69 & 0.27 & Relative Marry Black & National Identity, Ethnicity \& Immigration \\ 
marhomo & 0.05 & 0.04 & 0.82 & 0.14 & Gay Marriage & Gender \& Family \\ 
marwht & 0.00 & 0.00 & 0.55 & 0.45 & Relative Marry White & National Identity, Ethnicity \& Immigration \\ 
meovrwrk & 0.03 & 0.02 & 0.58 & 0.40 & Men Hurt Family When Overwork & Gender \& Family \\ 
milscale & 0.03 & 0.02 & 0.69 & 0.29 & Militarist: Scale & Politics, Government \& Economy \\ 
misanthropy & 0.07 & 0.06 & 0.77 & 0.17 & Misanthropy Scale & Social Life, Social Cohesion \& Trust \\ 
mslmscale & 0.00 & 0.00 & 0.77 & 0.23 & Islamist: Scale & Politics, Government \& Economy \\ 
nataid & 0.07 & 0.05 & 0.65 & 0.30 & Foreign Aid & Politics, Government \& Economy \\ 
natarms & 0.08 & 0.06 & 0.67 & 0.28 & Defense & Politics, Government \& Economy \\ 
natchld & 0.05 & 0.04 & 0.63 & 0.34 & Child Care & Politics, Government \& Economy \\ 
natcity & 0.05 & 0.03 & 0.63 & 0.34 & Problems Of Cities & Politics, Government \& Economy \\ 
natcrime & 0.12 & 0.08 & 0.62 & 0.29 & Fighting Crime & Politics, Government \& Economy \\ 
natdrug & 0.12 & 0.08 & 0.60 & 0.32 & Dealing With Drugs & Politics, Government \& Economy \\ 
nateduc & 0.16 & 0.13 & 0.66 & 0.21 & Education & Politics, Government \& Economy \\ 
natenvir & 0.01 & 0.01 & 0.67 & 0.32 & Protect Environment & Politics, Government \& Economy \\ 
natfare & 0.08 & 0.06 & 0.75 & 0.19 & Welfare & Politics, Government \& Economy \\ 
natheal & 0.02 & 0.01 & 0.57 & 0.41 & Improving Health & Politics, Government \& Economy \\ 
natmass & 0.05 & 0.03 & 0.62 & 0.35 & Mass Transit & Politics, Government \& Economy \\ 
natpark & 0.19 & 0.14 & 0.58 & 0.28 & Parks \& Recreation & Politics, Government \& Economy \\ 
natrace & 0.13 & 0.11 & 0.72 & 0.18 & Improve Life For Blacks & Politics, Government \& Economy \\ 
natroad & 0.01 & 0.01 & 0.61 & 0.39 & Highways \& Bridges & Politics, Government \& Economy \\ 
natsci & 0.12 & 0.08 & 0.60 & 0.32 & Scientific Research & Politics, Government \& Economy \\ 
natsoc & 0.01 & 0.00 & 0.63 & 0.36 & Social Security & Politics, Government \& Economy \\ 
natspac & 0.10 & 0.07 & 0.70 & 0.23 & Space Exploration & Politics, Government \& Economy \\ 
neverrelact & 0.12 & 0.09 & 0.65 & 0.27 & Never Active In Religion & Religion \& Spirituality \\ 
news & 0.08 & 0.06 & 0.77 & 0.16 & Read Newspaper & Social Life, Social Cohesion \& Trust \\ 
obey & 0.08 & 0.07 & 0.72 & 0.22 & Child Should Obey & Gender \& Family \\ 
owngun & 0.08 & 0.08 & 0.86 & 0.06 & Gun Owner & Politics, Government \& Economy \\ 
parsol & 0.01 & 0.01 & 0.71 & 0.28 & Mobility & Subjective Ses \\ 
partyid & 0.07 & 0.06 & 0.88 & 0.05 & Political Party & Politics, Government \& Economy \\ 
pillok & 0.10 & 0.07 & 0.70 & 0.23 & Birth Control 14-16 Year Olds & Gender \& Family \\ 
polabuse & 0.00 & 0.00 & 0.50 & 0.50 & Police Hit Person Cursing & Politics, Government \& Economy \\ 
polattak & 0.00 & 0.00 & 0.54 & 0.46 & Police Hit Person Punching & Politics, Government \& Economy \\ 
polescap & 0.16 & 0.11 & 0.59 & 0.30 & Police Hit Person Fleeing & Politics, Government \& Economy \\ 
polhitok & 0.00 & 0.00 & 0.71 & 0.29 & Police Can Hit Citizens & Politics, Government \& Economy \\ 
polmurdr & 0.00 & 0.00 & 0.53 & 0.47 & Police Hit Murder Suspect & Politics, Government \& Economy \\ 
polviews & 0.08 & 0.07 & 0.75 & 0.18 & Political Views & Politics, Government \& Economy \\ 
polviolent & 0.01 & 0.01 & 0.62 & 0.38 & Police Hit: Scale & Politics, Government \& Economy \\ 
popespks & 0.15 & 0.12 & 0.70 & 0.19 & Pope Infallible & Religion \& Spirituality \\ 
popular & 0.23 & 0.17 & 0.55 & 0.28 & Child Should Be Popular & Gender \& Family \\ 
pornlaw & 0.04 & 0.02 & 0.66 & 0.32 & Restrict Pornography & Gender \& Family \\ 
postlife & 0.09 & 0.07 & 0.78 & 0.15 & Life After Death & Religion \& Spirituality \\ 
pray & 0.08 & 0.07 & 0.86 & 0.07 & Pray & Religion \& Spirituality \\ 
prayer & 0.13 & 0.10 & 0.68 & 0.22 & Bible Prayer In Schools & Politics, Government \& Economy \\ 
premarsx & 0.07 & 0.06 & 0.79 & 0.15 & Morality: Premarital Sex & Gender \& Family \\ 
racdif1 & 0.09 & 0.07 & 0.65 & 0.28 & Racial Diff.: Discrimination & National Identity, Ethnicity \& Immigration \\ 
racdif2 & 0.00 & 0.00 & 0.57 & 0.43 & Racial Diff.: Inborn & National Identity, Ethnicity \& Immigration \\ 
racdif3 & 0.08 & 0.05 & 0.64 & 0.30 & Racial Diff.: Lack Education & National Identity, Ethnicity \& Immigration \\ 
racdif4 & 0.06 & 0.04 & 0.64 & 0.32 & Racial Diff.: Lack Will & National Identity, Ethnicity \& Immigration \\ 
racopen & 0.13 & 0.09 & 0.61 & 0.30 & Open Housing Law & National Identity, Ethnicity \& Immigration \\ 
racscale & 0.01 & 0.01 & 0.69 & 0.31 & Racist: Scale & Politics, Government \& Economy \\ 
reborn & 0.07 & 0.06 & 0.81 & 0.13 & Born Again & Religion \& Spirituality \\ 
relactiv & 0.13 & 0.11 & 0.76 & 0.13 & Do Religious Activities & Religion \& Spirituality \\ 
relexp & 0.14 & 0.12 & 0.76 & 0.12 & Had Religious Experience & Religion \& Spirituality \\ 
relexper & 0.03 & 0.02 & 0.74 & 0.24 & Religious Turning Point & Religion \& Spirituality \\ 
reliten & 0.10 & 0.09 & 0.80 & 0.11 & Strength Of Affiliation & Religion \& Spirituality \\ 
relpersn & 0.06 & 0.05 & 0.81 & 0.14 & Religious Person & Religion \& Spirituality \\ 
richwork & 0.11 & 0.08 & 0.66 & 0.26 & Would Work If Rich & Health \& Morale \\ 
satfin & 0.17 & 0.14 & 0.68 & 0.18 & Satisfied With Finances & Health \& Morale \\ 
satjob & 0.20 & 0.15 & 0.60 & 0.25 & Satisfied With Work & Health \& Morale \\ 
selfhunt & 0.02 & 0.02 & 0.80 & 0.18 & Hunter: Self & Politics, Government \& Economy \\ 
sexeduc & 0.00 & 0.00 & 0.69 & 0.31 & Sex-Ed. In Public Schools & Gender \& Family \\ 
socbar & 0.11 & 0.10 & 0.80 & 0.10 & Meet Friends At Bar & Social Life, Social Cohesion \& Trust \\ 
socfrend & 0.09 & 0.06 & 0.65 & 0.28 & Meet Friends & Social Life, Social Cohesion \& Trust \\ 
socialscale & 0.09 & 0.07 & 0.74 & 0.18 & Social Life Scale & Social Life, Social Cohesion \& Trust \\ 
socommun & 0.16 & 0.12 & 0.64 & 0.25 & Meet Neighbors & Social Life, Social Cohesion \& Trust \\ 
socrel & 0.13 & 0.10 & 0.67 & 0.23 & Meet Relatives & Social Life, Social Cohesion \& Trust \\ 
spanking & 0.07 & 0.06 & 0.73 & 0.21 & Spank Children & Gender \& Family \\ 
spkath & 0.03 & 0.02 & 0.66 & 0.32 & Athiest: Make Speech & Politics, Government \& Economy \\ 
spkcom & 0.01 & 0.01 & 0.71 & 0.28 & Red: Make Speech & Politics, Government \& Economy \\ 
spkhomo & 0.07 & 0.05 & 0.65 & 0.30 & Gay: Make Speech & Politics, Government \& Economy \\ 
spkmil & 0.00 & 0.00 & 0.63 & 0.36 & Militarist: Make Speech & Politics, Government \& Economy \\ 
spkmslm & 0.02 & 0.01 & 0.73 & 0.26 & Islamist: Make Speech & Politics, Government \& Economy \\ 
spkrac & 0.06 & 0.04 & 0.63 & 0.33 & Racist: Make Speech & Politics, Government \& Economy \\ 
sprtprsn & 0.10 & 0.09 & 0.78 & 0.13 & Spiritual Person & Religion \& Spirituality \\ 
suicide1 & 0.01 & 0.00 & 0.73 & 0.27 & Suicide: Incurable & Health \& Morale \\ 
suicide2 & 0.08 & 0.06 & 0.65 & 0.29 & Suicide: Bankruptcy & Health \& Morale \\ 
suicide3 & 0.03 & 0.02 & 0.66 & 0.32 & Suicide: Dishonor & Health \& Morale \\ 
suicide4 & 0.08 & 0.06 & 0.68 & 0.26 & Suicide: Tired Of Living & Health \& Morale \\ 
suicidescale & 0.03 & 0.03 & 0.75 & 0.23 & Suicide Scale & Health \& Morale \\ 
teensex & 0.11 & 0.08 & 0.66 & 0.26 & Morality: Sex 14-16 Year Olds & Gender \& Family \\ 
thnkself & 0.10 & 0.07 & 0.65 & 0.27 & Child Should Think For Self & Gender \& Family \\ 
trust & 0.00 & 0.00 & 0.72 & 0.28 & People Can Be Trusted & Social Life, Social Cohesion \& Trust \\ 
tvhours & 0.15 & 0.13 & 0.75 & 0.13 & Watch Television & Social Life, Social Cohesion \& Trust \\ 
uswary & 0.04 & 0.03 & 0.67 & 0.30 & Us At War Next 10 Years & Politics, Government \& Economy \\ 
wlthblks & 0.02 & 0.01 & 0.52 & 0.47 & Blacks Rich Or Poor & National Identity, Ethnicity \& Immigration \\ 
wlthwhts & 0.23 & 0.16 & 0.53 & 0.30 & Whites Rich Or Poor & National Identity, Ethnicity \& Immigration \\ 
workblks & 0.00 & 0.00 & 0.56 & 0.44 & Blacks Work Hard & National Identity, Ethnicity \& Immigration \\ 
workhard & 0.01 & 0.00 & 0.57 & 0.42 & Child Should Work Hard & Gender \& Family \\ 
workwhts & 0.23 & 0.16 & 0.55 & 0.29 & Whites Work Hard & National Identity, Ethnicity \& Immigration \\ 
wrkwayup & 0.12 & 0.09 & 0.69 & 0.21 & No Racial Favors & National Identity, Ethnicity \& Immigration \\ 
xmarsex & 0.18 & 0.14 & 0.64 & 0.22 & Morality: Extramarital Affair & Gender \& Family \\ 
xmovie & 0.07 & 0.06 & 0.71 & 0.23 & Watch X-Rated Movie & Social Life, Social Cohesion \& Trust \\ 
\midrule
\multicolumn{7}{l}{\textbf{HILDA}} \\ 
\midrule
atwkadc & 0.13 & 0.09 & 0.58 & 0.33 & Fine For Child Under 3 To Be In All-Day Care & Gender \& Family \\ 
atwkbmw & 0.07 & 0.05 & 0.69 & 0.25 & Better If Man Earns Money, Woman Cares For Children & Gender \& Family \\ 
atwkcdw & 0.13 & 0.09 & 0.58 & 0.33 & As Well For Child If Mother Earns Money, Father Cares For Children & Gender \& Family \\ 
atwkcs & 0.12 & 0.08 & 0.59 & 0.33 & Children Suffer Bc Father Works Too Much & Gender \& Family \\ 
atwkdcm & 0.14 & 0.10 & 0.64 & 0.25 & Better For Child To Stay With Mother In Case Of Divorce & Gender \& Family \\ 
atwkfhi & 0.17 & 0.11 & 0.55 & 0.33 & Father Should Be As Involved As Mother & Gender \& Family \\ 
atwkmmf & 0.08 & 0.05 & 0.61 & 0.34 & Bad For Relationship If Woman Earns More & Gender \& Family \\ 
atwkmpl & 0.02 & 0.01 & 0.66 & 0.33 & Men Make Better Political Leaders & Gender \& Family \\ 
atwkmrl & 0.11 & 0.08 & 0.59 & 0.34 & Most Important Role For Men To Be Father & Gender \& Family \\ 
atwkmsw & 0.08 & 0.05 & 0.62 & 0.33 & Mothers Who Don't Need Money Shouldn't Work & Gender \& Family \\ 
atwkpsc & 0.08 & 0.06 & 0.63 & 0.32 & Pre-School Child Suffers If Mother Works Ft & Gender \& Family \\ 
atwkseh & 0.16 & 0.10 & 0.53 & 0.37 & If Both Partners Work They Should Share Domestic Tasks & Gender \& Family \\ 
atwkwfr & 0.12 & 0.07 & 0.52 & 0.41 & Working Father Can Have As Good Relationship To Child & Gender \& Family \\ 
atwkwfs & 0.15 & 0.09 & 0.52 & 0.39 & Working Fathers Care More About Success Than Needs Of Children & Gender \& Family \\ 
atwkwmr & 0.11 & 0.07 & 0.58 & 0.36 & Working Mother Can Have As Good Relationship To Child & Gender \& Family \\ 
atwkwms & 0.15 & 0.10 & 0.56 & 0.34 & Working Mothers Care More About Success Than Needs Of Children & Gender \& Family \\ 
atwkwrl & 0.10 & 0.07 & 0.60 & 0.33 & Most Important Role For Women To Be Mother & Gender \& Family \\ 
lskeep & 0.07 & 0.04 & 0.55 & 0.41 & Most People You Meet Keep Their Word & Social Life, Social Cohesion \& Trust \\ 
lssop & 0.13 & 0.08 & 0.58 & 0.34 & Most People You Meet Succeed By Stepping On Others & Social Life, Social Cohesion \& Trust \\ 
mcchmf & 0.12 & 0.09 & 0.66 & 0.25 & Children Happier With Both Father And Mother & Gender \& Family \\ 
mcdef & 0.06 & 0.05 & 0.80 & 0.15 & Cohabitation Alright Even Without Marriage Intention & Gender \& Family \\ 
mcdiv & 0.08 & 0.05 & 0.62 & 0.32 & Divorce Alright Even With Children & Gender \& Family \\ 
mchscr & 0.09 & 0.07 & 0.76 & 0.16 & Same Rights For Homosexual Couples & Gender \& Family \\ 
mcindpc & 0.15 & 0.11 & 0.63 & 0.26 & Child Should Start Living Independently By 18-20 & Gender \& Family \\ 
mcmhc & 0.18 & 0.14 & 0.66 & 0.20 & A Man Has To Have Children In Order To Be Fulfilled & Gender \& Family \\ 
mcmoi & 0.13 & 0.09 & 0.64 & 0.27 & Marriage Is Outdated & Gender \& Family \\ 
mcnodiv & 0.08 & 0.07 & 0.70 & 0.23 & Marriage Should Never Be Ended & Gender \& Family \\ 
mcsingp & 0.07 & 0.05 & 0.69 & 0.26 & Alright For Woman To Have Children As Single Parent & Gender \& Family \\ 
mcwhc & 0.16 & 0.12 & 0.67 & 0.21 & A Woman Has To Have Children In Order To Be Fulfilled & Gender \& Family \\ 
relimp & 0.05 & 0.05 & 0.87 & 0.08 & Importance Of Religion & Religion \& Spirituality \\ 
\midrule
\multicolumn{7}{l}{\textbf{PSID}} \\ 
\midrule
cds\_bothparentscare & 0.06 & 0.04 & 0.55 & 0.42 & Dad Should Be As Involved As Mom & Gender \& Family \\ 
cds\_daddevelopment & 0.08 & 0.05 & 0.53 & 0.42 & Dad Large Role In Child Development & Gender \& Family \\ 
cds\_dadoldkidbetter & 0.15 & 0.11 & 0.59 & 0.30 & Dad Able To Enjoy Older Children More & Gender \& Family \\ 
cds\_dadtooinvolved & 0.19 & 0.13 & 0.55 & 0.31 & Dad Too Involved If Keeps From Better Job & Gender \& Family \\ 
cds\_daycareok & 0.16 & 0.12 & 0.63 & 0.25 & Ok To Have Child Under 3 In All-Day Care & Gender \& Family \\ 
cds\_fatherhoodfulfill & 0.00 & 0.00 & 0.57 & 0.43 & Dadhood Most Fulfilling Experience & Gender \& Family \\ 
cds\_fatherinteractchild & 0.01 & 0.00 & 0.56 & 0.44 & Dads Should Interact With Child & Gender \& Family \\ 
cds\_firstyears & 0.01 & 0.00 & 0.59 & 0.40 & First 4 Years Most Import. In Child's Develop. & Gender \& Family \\ 
cds\_girlboyequal & 0.05 & 0.03 & 0.59 & 0.38 & Encourage As Much Independence In Daughters And Sons & Gender \& Family \\ 
cds\_husbandcareerimport & 0.13 & 0.09 & 0.61 & 0.30 & Better Wife Helps Husband Career & Gender \& Family \\ 
cds\_mostimportantchild\_1 & 0.10 & 0.06 & 0.55 & 0.39 & Importance For Child: Obey & Gender \& Family \\ 
cds\_mostimportantchild\_3 & 0.00 & 0.00 & 0.57 & 0.43 & Importance For Child: Think For Herself & Gender \& Family \\ 
cds\_motherfatherequal & 0.19 & 0.13 & 0.56 & 0.31 & Dad And Mom Equally Good Meeting Kid's Needs & Gender \& Family \\ 
cds\_motherhoodfulfill & 0.00 & 0.00 & 0.59 & 0.41 & Momhood Most Fulfilling Experience & Gender \& Family \\ 
cds\_mothernotft & 0.12 & 0.10 & 0.67 & 0.24 & Mom With Child Under 5 Shouldn't Work Full-Time & Gender \& Family \\ 
cds\_preschoolsuffer & 0.13 & 0.09 & 0.65 & 0.26 & Preschool Child Suffers If Mom Employed & Gender \& Family \\ 
cds\_sharehousehold & 0.00 & 0.00 & 0.48 & 0.52 & Husband And Wife In Both Ft Should Share Household & Gender \& Family \\ 
cds\_traditionalbest & 0.10 & 0.07 & 0.66 & 0.27 & Better If Man Earns Living & Gender \& Family \\ 
cds\_workingmotherssame & 0.05 & 0.03 & 0.55 & 0.42 & Working Mom As Warm As Nonworking Mom & Gender \& Family \\ 
gettingbetter & 0.18 & 0.11 & 0.50 & 0.38 & Life Getting Better & Health \& Morale \\ 
lifesatis & 0.17 & 0.11 & 0.53 & 0.35 & Life Satisfaction & Health \& Morale \\ 
likejobormoney & 0.17 & 0.09 & 0.46 & 0.44 & Satisfaction About Income Important In Job & Occupation \& Education \\ 
opinionimport & 0.17 & 0.12 & 0.56 & 0.32 & How Much Does It Matter What Other People Think About You? & Social Life, Social Cohesion \& Trust \\ 
satisfiedself & 0.22 & 0.14 & 0.50 & 0.36 & Are You More Often Satisfied Or Dissatisfied With Yourself? & Health \& Morale \\ 
savingspending & 0.20 & 0.12 & 0.49 & 0.38 & Spend Or Save Money & Health \& Morale \\ 
tas\_belongcommunity & 0.14 & 0.09 & 0.56 & 0.36 & Freq. Of Feeling Belonging To Communty & Social Life, Social Cohesion \& Trust \\ 
tas\_contributesociety & 0.14 & 0.09 & 0.53 & 0.38 & Freq. Of Feeling Contrib To Society & Social Life, Social Cohesion \& Trust \\ 
tas\_divorcehurts & 0.10 & 0.05 & 0.49 & 0.45 & Divorce Hurts Kids & Gender \& Family \\ 
tas\_divorcesuccess & 0.15 & 0.08 & 0.47 & 0.45 & Child From Divorced Family Can Be Successful & Gender \& Family \\ 
tas\_fathertimeoff & 0.20 & 0.13 & 0.52 & 0.36 & Dads Should Take Time With Young Child & Gender \& Family \\ 
tas\_importrel & 0.10 & 0.08 & 0.73 & 0.19 & Importance Of Religion & Religion \& Spirituality \\ 
tas\_importspirit & 0.13 & 0.10 & 0.67 & 0.24 & Importance Of Spirituality & Religion \& Spirituality \\ 
tas\_lifesatisfact & 0.19 & 0.14 & 0.61 & 0.25 & How Satisfied With Life As A Whole & Health \& Morale \\ 
tas\_mothersnodemanding & 0.17 & 0.11 & 0.55 & 0.34 & Moms Should Have Demandng Job & Gender \& Family \\ 
tas\_peoplearegood & 0.15 & 0.10 & 0.54 & 0.36 & Freq. Of Feeling People Basically Good & Social Life, Social Cohesion \& Trust \\ 
tas\_religpref0509\_2 & 0.04 & 0.03 & 0.76 & 0.21 & Religion: Protestant [2005-09] & Religion \& Spirituality \\ 
tas\_religpref0509\_3 & 0.07 & 0.06 & 0.89 & 0.05 & Religion: Catholic [2005-09] & Religion \& Spirituality \\ 
tas\_religpref1115\_2 & 0.11 & 0.10 & 0.83 & 0.07 & Religion: Protestant [2011-15] & Religion \& Spirituality \\ 
tas\_religpref1115\_3 & 0.03 & 0.02 & 0.87 & 0.11 & Religion: Catholic [2011-15] & Religion \& Spirituality \\ 
tas\_singleparentsok & 0.15 & 0.09 & 0.50 & 0.41 & Child From Single-Parent Family Can Do As Well & Gender \& Family \\ 
tas\_singleparentsuffer & 0.14 & 0.10 & 0.60 & 0.30 & Single Parent Hurts Kids & Gender \& Family \\ 
tas\_singlewomenkids & 0.12 & 0.07 & 0.51 & 0.42 & Single Women Should Have Kids & Gender \& Family \\ 
tas\_societybetter & 0.17 & 0.11 & 0.52 & 0.37 & Freq. Of Feeling Society Gettng Better & Social Life, Social Cohesion \& Trust \\ 
tas\_societymakessense & 0.14 & 0.09 & 0.53 & 0.38 & Feeling Way Society Works Makes Sense & Social Life, Social Cohesion \& Trust \\ 
tas\_spritual & 0.14 & 0.10 & 0.63 & 0.27 & Spiritual Person & Religion \& Spirituality \\ 
thinkfuture & 0.21 & 0.13 & 0.50 & 0.37 & Think About Future & Health \& Morale \\ 
trustothers & 0.13 & 0.10 & 0.65 & 0.26 & Trust Most Others & Social Life, Social Cohesion \& Trust \\ 
undeserved & 0.19 & 0.13 & 0.55 & 0.32 & Are There A Lot Of People Who Have Good Things They Don't Deserve? & Social Life, Social Cohesion \& Trust \\ 
\midrule
\multicolumn{7}{l}{\textbf{SHP}} \\ 
\midrule
pc70 & 0.05 & 0.03 & 0.56 & 0.40 & Self-Perception : Incapacity To Make Plans Because Of Unpredictability & Health \& Morale \\ 
pc71 & 0.16 & 0.10 & 0.55 & 0.35 & Self-Perception : Little Influence On Life Events & Health \& Morale \\ 
pc72 & 0.00 & 0.00 & 0.54 & 0.46 & Self-Perception : Capacity To Overcome Unexpected Problems & Health \& Morale \\ 
pc73 & 0.03 & 0.02 & 0.59 & 0.39 & Self-Perception : Capacity To Choose Between Two Possibilities & Health \& Morale \\ 
pc74 & 0.13 & 0.09 & 0.63 & 0.27 & Self-Perception : Feeling Of Uselessness & Health \& Morale \\ 
pc75 & 0.00 & 0.00 & 0.64 & 0.36 & Self-Perception : Feeling Of Selfsatisfaction & Health \& Morale \\ 
pl101 & 0.14 & 0.10 & 0.63 & 0.27 & Feeling Of Security (Victimization) & Health \& Morale \\ 
pn56 & 0.21 & 0.17 & 0.63 & 0.20 & Values: Lying In Own Interest & Social Life, Social Cohesion \& Trust \\ 
pn57 & 0.20 & 0.15 & 0.63 & 0.22 & Values: Cheating On Tax Declaration & Social Life, Social Cohesion \& Trust \\ 
pn58 & 0.20 & 0.16 & 0.64 & 0.20 & Values: Avoiding A Fare On A Public Transport & Social Life, Social Cohesion \& Trust \\ 
pn59 & 0.05 & 0.02 & 0.50 & 0.47 & Values: Claiming State Benefits Not Entitled To & Social Life, Social Cohesion \& Trust \\ 
pn60 & 0.11 & 0.08 & 0.67 & 0.25 & Values: Keeping Found Money & Social Life, Social Cohesion \& Trust \\ 
pn61 & 0.00 & 0.00 & 0.50 & 0.50 & Values: Failing To Report Damage To Parked Vehicle & Social Life, Social Cohesion \& Trust \\ 
pp01 & 0.04 & 0.03 & 0.79 & 0.17 & Interest In Politics & Politics, Government \& Economy \\ 
pp02 & 0.08 & 0.05 & 0.59 & 0.35 & Overall Satisfaction With Democracy & Politics, Government \& Economy \\ 
pp03 & 0.08 & 0.05 & 0.56 & 0.40 & Feeling About Political Influence & Politics, Government \& Economy \\ 
pp04 & 0.08 & 0.06 & 0.63 & 0.31 & Trust In Federal Government & Politics, Government \& Economy \\ 
pp07 & 0.06 & 0.04 & 0.67 & 0.28 & Future: Active In Boycott & Politics, Government \& Economy \\ 
pp08 & 0.06 & 0.05 & 0.70 & 0.25 & Future: Active In Strike & Politics, Government \& Economy \\ 
pp09 & 0.06 & 0.04 & 0.72 & 0.23 & Future: Active In Demonstration & Politics, Government \& Economy \\ 
pp10 & 0.05 & 0.04 & 0.74 & 0.21 & Political Position: Left, Right & Politics, Government \& Economy \\ 
pp12 & 0.06 & 0.05 & 0.71 & 0.24 & Opinion On Swiss Army: Direction & Politics, Government \& Economy \\ 
pp13 & 0.09 & 0.06 & 0.55 & 0.40 & Opinion On Social Expenses: Direction & Politics, Government \& Economy \\ 
pp14 & 0.15 & 0.12 & 0.65 & 0.24 & Opinion On Joining The Eu: Direction & Politics, Government \& Economy \\ 
pp15 & 0.08 & 0.05 & 0.59 & 0.35 & Opinion On Chances For Foreigners: Direction & National Identity, Ethnicity \& Immigration \\ 
pp16 & 0.09 & 0.05 & 0.52 & 0.43 & Opinion On Environmental Protection: Direction & Environment \& Climate \\ 
pp17 & 0.10 & 0.05 & 0.51 & 0.43 & Opinion On Taxes On High Income: Direction & Politics, Government \& Economy \\ 
pp18 & 0.09 & 0.06 & 0.62 & 0.32 & Opinion On Nuclear Energy: Direction & Environment \& Climate \\ 
pp19\_1 & 0.08 & 0.05 & 0.55 & 0.40 & Party Support: Swiss Radical-Democratic Party & Politics, Government \& Economy \\ 
pp19\_3 & 0.08 & 0.05 & 0.66 & 0.28 & Party Support: Pss Swiss Socialist Party & Politics, Government \& Economy \\ 
pp19\_4 & 0.08 & 0.06 & 0.67 & 0.28 & Party Support: Udc Democratic Union Of The Centre & Politics, Government \& Economy \\ 
pp20 & 0.09 & 0.05 & 0.53 & 0.42 & Gender: Women In General Penalized & Gender \& Family \\ 
pp21 & 0.11 & 0.06 & 0.51 & 0.42 & Gender: Personnally Penalized & Gender \& Family \\ 
pp22 & 0.07 & 0.04 & 0.62 & 0.34 & Gender: In Favour Of Measures & Gender \& Family \\ 
pp23 & 0.15 & 0.07 & 0.38 & 0.55 & Gender: Personal Action & Gender \& Family \\ 
pp45 & 0.06 & 0.04 & 0.65 & 0.31 & General Trust In People & Social Life, Social Cohesion \& Trust \\ 
pp50 & 0.14 & 0.10 & 0.61 & 0.30 & Class Identification: Yes-No & Subjective Ses \\ 
pp51 & 0.11 & 0.10 & 0.79 & 0.11 & Class Identification (Spontanously): Which & Subjective Ses \\ 
pp52 & 0.17 & 0.15 & 0.70 & 0.15 & Class Identification (Non Spontaneously): Which & Subjective Ses \\ 
pp53 & 0.00 & 0.00 & 0.68 & 0.32 & Public Expenses: Protection Of The Environment & Politics, Government \& Economy \\ 
pp54 & 0.15 & 0.11 & 0.65 & 0.24 & Public Expenses: Health & Politics, Government \& Economy \\ 
pp55 & 0.09 & 0.07 & 0.72 & 0.21 & Public Expenses: Day Nursery & Politics, Government \& Economy \\ 
pp56 & 0.01 & 0.01 & 0.64 & 0.36 & Public Expenses: Police And Public Order & Politics, Government \& Economy \\ 
pp57 & 0.13 & 0.08 & 0.59 & 0.33 & Public Expenses: Research And Education & Politics, Government \& Economy \\ 
pp58 & 0.06 & 0.05 & 0.74 & 0.22 & Public Expenses: Military And Defense & Politics, Government \& Economy \\ 
pp59 & 0.09 & 0.06 & 0.63 & 0.31 & Public Expenses: Retirement & Politics, Government \& Economy \\ 
pp60 & 0.08 & 0.07 & 0.72 & 0.21 & Public Expenses: Culture And Art & Politics, Government \& Economy \\ 
pp61 & 0.03 & 0.02 & 0.62 & 0.35 & Public Expenses: Unemployment Benefit & Politics, Government \& Economy \\ 
pp62 & 0.11 & 0.09 & 0.69 & 0.22 & Public Expenses: Aid To Developing Countries & Politics, Government \& Economy \\ 
pp63 & 0.00 & 0.00 & 0.67 & 0.33 & Public Expenses: Social Aid & Politics, Government \& Economy \\ 
pp64 & 0.09 & 0.06 & 0.62 & 0.31 & Opinion On Joining The Eu: Joining Or Bilateral Agreements & Politics, Government \& Economy \\ 
pp65 & 0.18 & 0.15 & 0.70 & 0.14 & Opinion On Joining The Eu: Continuation Or Dissolution Of Bilateral Agreements & Politics, Government \& Economy \\ 
pp66 & 0.17 & 0.15 & 0.72 & 0.13 & Party Identification: Yes-No & Politics, Government \& Economy \\ 
pp67\_2 & 0.04 & 0.04 & 0.93 & 0.03 & Party Identification: Swiss Christian-Democrat Party & Politics, Government \& Economy \\ 
pp67\_3 & 0.06 & 0.06 & 0.90 & 0.04 & Party Identification: Pss Swiss Socialist Party & Politics, Government \& Economy \\ 
pp67\_4 & 0.04 & 0.04 & 0.93 & 0.03 & Party Identification: Udc Democratic Union Of The Centre & Politics, Government \& Economy \\ 
pp68 & 0.00 & 0.00 & 0.65 & 0.35 & Party Identification: Proximity & Politics, Government \& Economy \\ 
pp69 & 0.11 & 0.09 & 0.76 & 0.15 & Sympathy Political Party: Liberal Party & Politics, Government \& Economy \\ 
pp70 & 0.11 & 0.09 & 0.76 & 0.15 & Sympathy Political Party: Christian Democrats & Politics, Government \& Economy \\ 
pp71 & 0.08 & 0.08 & 0.83 & 0.09 & Sympathy Political Party: Social Democrats & Politics, Government \& Economy \\ 
pp72 & 0.09 & 0.08 & 0.85 & 0.07 & Sympathy Political Party: Swiss People's Party & Politics, Government \& Economy \\ 
pp73 & 0.09 & 0.08 & 0.80 & 0.12 & Sympathy Political Party: Green Party & Politics, Government \& Economy \\ 
pp74 & 0.07 & 0.05 & 0.64 & 0.31 & Opinion On Federalism: Direction & Politics, Government \& Economy \\ 
pp76 & 0.03 & 0.02 & 0.75 & 0.23 & Protection Of Environment: Importance & Environment \& Climate \\ 
pp77 & 0.13 & 0.10 & 0.65 & 0.26 & Protection Of Environment: Recycling & Environment \& Climate \\ 
pp78 & 0.21 & 0.19 & 0.69 & 0.13 & Protection Of Environment: Bin Bag Fees & Environment \& Climate \\ 
pp79 & 0.08 & 0.07 & 0.76 & 0.17 & Protection Of Environment: Organic And Ecological Products & Environment \& Climate \\ 
pp80 & 0.10 & 0.08 & 0.72 & 0.20 & Protection Of Environment: Purchase Of Fruits And Vegetables & Environment \& Climate \\ 
pr16 & 0.03 & 0.03 & 0.89 & 0.08 & Belief In God Or Something Divine & Religion \& Spirituality \\ 
pr20 & 0.11 & 0.10 & 0.81 & 0.09 & Feeling Of Spirituality & Religion \& Spirituality \\ 
pr21 & 0.05 & 0.05 & 0.84 & 0.11 & Feeling Of Religiosity & Religion \& Spirituality \\ 
pr23 & 0.11 & 0.10 & 0.79 & 0.12 & Attitude Towards Christianism & Religion \& Spirituality \\ 
pr24 & 0.13 & 0.12 & 0.76 & 0.12 & Attitude Towards Islam & Religion \& Spirituality \\ 
pr25 & 0.11 & 0.09 & 0.73 & 0.18 & Attitude Towards Hinduism & Religion \& Spirituality \\ 
pr26 & 0.12 & 0.10 & 0.77 & 0.13 & Attitude Towards Buddhism & Religion \& Spirituality \\ 
pr27 & 0.10 & 0.08 & 0.75 & 0.17 & Attitude Towards Judaism & Religion \& Spirituality \\ 
pr28 & 0.10 & 0.08 & 0.72 & 0.20 & Attitude Towards Atheism & Religion \& Spirituality \\ 
\midrule
\multicolumn{7}{l}{\textbf{SOEP}} \\ 
\midrule
plb0104\_h & 0.23 & 0.16 & 0.55 & 0.29 & Importance Of Employment & Occupation \& Education \\ 
plb0528\_h & 0.16 & 0.12 & 0.64 & 0.24 & Priority In Career Choice: Income & Occupation \& Education \\ 
plb0530\_h & 0.06 & 0.04 & 0.62 & 0.34 & Priority In Career Choice: Status & Occupation \& Education \\ 
plb0531\_h & 0.06 & 0.04 & 0.67 & 0.28 & Priority In Career Choice: Social Contact & Occupation \& Education \\ 
plb0532\_h & 0.12 & 0.08 & 0.60 & 0.31 & Priority In Career Choice: Compatibility Of Family And Work & Occupation \& Education \\ 
plc0115 & 0.10 & 0.06 & 0.58 & 0.35 & Social Security Contributions & Politics, Government \& Economy \\ 
pld0043 & 0.13 & 0.09 & 0.64 & 0.27 & Freq. Lend Friends Pers. Belongings & Social Life, Social Cohesion \& Trust \\ 
pld0044 & 0.16 & 0.12 & 0.64 & 0.24 & Freq. Lend Friends Money & Social Life, Social Cohesion \& Trust \\ 
pld0045 & 0.09 & 0.06 & 0.62 & 0.32 & Freq. Leave Door Unlocked & Social Life, Social Cohesion \& Trust \\ 
plh0001 & 0.17 & 0.13 & 0.63 & 0.24 & Democratic Ideal & Politics, Government \& Economy \\ 
plh0004 & 0.10 & 0.08 & 0.69 & 0.23 & Political Attitude & Politics, Government \& Economy \\ 
plh0007 & 0.05 & 0.03 & 0.69 & 0.28 & Interest In Politics & Politics, Government \& Economy \\ 
plh0011\_h & 0.08 & 0.05 & 0.53 & 0.42 & Party Preference: Yes & Politics, Government \& Economy \\ 
plh0012\_h1 & 0.05 & 0.04 & 0.81 & 0.15 & Party Preference: Spd & Politics, Government \& Economy \\ 
plh0012\_h2 & 0.04 & 0.04 & 0.80 & 0.16 & Party Preference: Cdu & Politics, Government \& Economy \\ 
plh0012\_h3 & 0.04 & 0.03 & 0.76 & 0.21 & Party Preference: Csu & Politics, Government \& Economy \\ 
plh0012\_h4 & 0.09 & 0.06 & 0.66 & 0.28 & Party Preference: Fdp & Politics, Government \& Economy \\ 
plh0012\_h5 & 0.06 & 0.05 & 0.76 & 0.19 & Party Preference: The Greens & Politics, Government \& Economy \\ 
plh0012\_h6 & 0.06 & 0.05 & 0.78 & 0.17 & Party Preference: The Left & Politics, Government \& Economy \\ 
plh0013\_h & 0.09 & 0.05 & 0.49 & 0.46 & Party Preference: Intensity & Politics, Government \& Economy \\ 
plh0014 & 0.24 & 0.17 & 0.54 & 0.29 & Government Interference & Politics, Government \& Economy \\ 
plh0015 & 0.00 & 0.00 & 0.46 & 0.54 & Extent And Type Of State Powers & Politics, Government \& Economy \\ 
plh0016 & 0.03 & 0.02 & 0.56 & 0.42 & Financial Security For Family & Politics, Government \& Economy \\ 
plh0017 & 0.03 & 0.02 & 0.53 & 0.45 & Care For Pre-School Children & Politics, Government \& Economy \\ 
plh0018 & 0.01 & 0.00 & 0.49 & 0.50 & Care For School Children & Politics, Government \& Economy \\ 
plh0019 & 0.14 & 0.09 & 0.55 & 0.36 & Financial Security When Unemployed & Politics, Government \& Economy \\ 
plh0020 & 0.00 & 0.00 & 0.59 & 0.41 & Job Creation Schemes & Politics, Government \& Economy \\ 
plh0021 & 0.15 & 0.10 & 0.57 & 0.34 & Financial Security When Sick & Politics, Government \& Economy \\ 
plh0022 & 0.08 & 0.05 & 0.51 & 0.44 & Help For The Sick & Politics, Government \& Economy \\ 
plh0023 & 0.19 & 0.13 & 0.57 & 0.30 & Financial Security In Old-Age & Politics, Government \& Economy \\ 
plh0024 & 0.05 & 0.03 & 0.52 & 0.45 & Help For The Elderly & Politics, Government \& Economy \\ 
plh0025 & 0.09 & 0.05 & 0.55 & 0.40 & Financial Security When Requiring Care & Politics, Government \& Economy \\ 
plh0026 & 0.07 & 0.04 & 0.53 & 0.43 & Help For Those Requiring Care & Politics, Government \& Economy \\ 
plh0030 & 0.16 & 0.10 & 0.51 & 0.39 & Worried About Childcare & Gender \& Family \\ 
plh0031 & 0.10 & 0.07 & 0.58 & 0.35 & Worried About Consequences Of Eu- Enlargement & Politics, Government \& Economy \\ 
plh0032 & 0.11 & 0.05 & 0.38 & 0.57 & Worried About Economic Development & Politics, Government \& Economy \\ 
plh0033 & 0.09 & 0.05 & 0.50 & 0.46 & Worried About Finances & Health \& Morale \\ 
plh0034 & 0.07 & 0.04 & 0.53 & 0.43 & Worried About Stability Of Financial Markets & Politics, Government \& Economy \\ 
plh0035 & 0.09 & 0.05 & 0.55 & 0.40 & Worried About Own Health & Health \& Morale \\ 
plh0036 & 0.11 & 0.05 & 0.45 & 0.49 & Worried About Environment & Environment \& Climate \\ 
plh0037 & 0.09 & 0.06 & 0.56 & 0.38 & Worried About Consequences From Climate Change & Environment \& Climate \\ 
plh0038 & 0.09 & 0.04 & 0.42 & 0.54 & Worried About Peace & Politics, Government \& Economy \\ 
plh0039 & 0.08 & 0.05 & 0.58 & 0.37 & Worried About Global Terrorism & Politics, Government \& Economy \\ 
plh0040 & 0.08 & 0.04 & 0.52 & 0.44 & Worried About Crime In Germany & Politics, Government \& Economy \\ 
plh0041 & 0.10 & 0.06 & 0.54 & 0.40 & Worried About Introduction Of Euro & Politics, Government \& Economy \\ 
plh0042 & 0.10 & 0.05 & 0.48 & 0.46 & Worried About Job Security & Politics, Government \& Economy \\ 
plh0054 & 0.07 & 0.04 & 0.52 & 0.44 & Peace And Order & Politics, Government \& Economy \\ 
plh0054\_1 & 0.12 & 0.06 & 0.48 & 0.46 & Political Goals: Prority Peace And Order & Politics, Government \& Economy \\ 
plh0056 & 0.07 & 0.04 & 0.48 & 0.48 & Citizen Influence & Politics, Government \& Economy \\ 
plh0056\_1 & 0.09 & 0.05 & 0.46 & 0.49 & Political Goals: Prority Citizen Influence & Politics, Government \& Economy \\ 
plh0058 & 0.05 & 0.02 & 0.47 & 0.50 & Fight Against Inflation & Politics, Government \& Economy \\ 
plh0058\_1 & 0.01 & 0.00 & 0.38 & 0.62 & Political Goals: Prority Rising Prices & Politics, Government \& Economy \\ 
plh0061 & 0.08 & 0.04 & 0.51 & 0.44 & Freedom Of Speech & Politics, Government \& Economy \\ 
plh0061\_1 & 0.15 & 0.08 & 0.48 & 0.44 & Political Goals: Prority Freedom Of Speech & Politics, Government \& Economy \\ 
plh0090 & 0.19 & 0.15 & 0.66 & 0.19 & Importance Of Circle Of Friends & Social Life, Social Cohesion \& Trust \\ 
plh0091\_v2 & 0.16 & 0.14 & 0.69 & 0.18 & Importance For Satisfaction: Work & Health \& Morale \\ 
plh0092\_v2 & 0.21 & 0.17 & 0.64 & 0.20 & Importance For Satisfaction: Family & Health \& Morale \\ 
plh0093\_v2 & 0.20 & 0.16 & 0.65 & 0.19 & Importance For Satisfaction: Friends & Health \& Morale \\ 
plh0094\_v2 & 0.21 & 0.17 & 0.64 & 0.19 & Importance For Satisfaction: Income & Health \& Morale \\ 
plh0095\_v2 & 0.24 & 0.19 & 0.59 & 0.23 & Importance For Satisfaction: Dwelling & Health \& Morale \\ 
plh0096\_v2 & 0.20 & 0.15 & 0.62 & 0.22 & Importance For Satisfaction: Political Influence & Health \& Morale \\ 
plh0097\_v2 & 0.14 & 0.12 & 0.74 & 0.14 & Importance For Satisfaction: Professional Success & Health \& Morale \\ 
plh0098\_v2 & 0.25 & 0.19 & 0.60 & 0.21 & Importance For Satisfaction: Leisure & Health \& Morale \\ 
plh0099\_v2 & 0.26 & 0.20 & 0.57 & 0.23 & Importance For Satisfaction: Health & Health \& Morale \\ 
plh0100\_v2 & 0.11 & 0.08 & 0.62 & 0.30 & Importance For Satisfaction: Protection Of Environment & Health \& Morale \\ 
plh0101 & 0.09 & 0.08 & 0.82 & 0.09 & Import.-Rlgion, Belief For Satisfaction & Health \& Morale \\ 
plh0104 & 0.08 & 0.05 & 0.57 & 0.38 & Importance: To Be Able To Afford Sth. & Subjective Ses \\ 
plh0105 & 0.07 & 0.04 & 0.54 & 0.42 & Importance: To Help Others & Social Life, Social Cohesion \& Trust \\ 
plh0106 & 0.09 & 0.06 & 0.59 & 0.35 & Importance: To Fulfil Once Potential & Health \& Morale \\ 
plh0107 & 0.13 & 0.09 & 0.62 & 0.29 & Importance: To Have Success In The Job & Occupation \& Education \\ 
plh0108 & 0.08 & 0.06 & 0.71 & 0.23 & Importance: To Have An Own House & Subjective Ses \\ 
plh0109 & 0.12 & 0.08 & 0.61 & 0.31 & Importance: To Have A Happy Marriage & Gender \& Family \\ 
plh0110 & 0.08 & 0.06 & 0.70 & 0.23 & Importance: To Have Children & Gender \& Family \\ 
plh0111 & 0.05 & 0.03 & 0.62 & 0.34 & Importance: To Be Socially And Politically Active & Social Life, Social Cohesion \& Trust \\ 
plh0112 & 0.09 & 0.07 & 0.64 & 0.29 & Importance: To Be Able To Travel And See The World & Health \& Morale \\ 
plh0113 & 0.24 & 0.20 & 0.62 & 0.19 & Entitlement Received Are Deserved & Politics, Government \& Economy \\ 
plh0114 & 0.27 & 0.19 & 0.53 & 0.28 & German Citizen Benefits Are Equitable & Politics, Government \& Economy \\ 
plh0152\_v1 & 0.09 & 0.06 & 0.60 & 0.33 & Satisfaction With Democracy [1987-1989] & Politics, Government \& Economy \\ 
plh0152\_v2 & 0.09 & 0.07 & 0.70 & 0.23 & Satisfaction With Democracy [2005,2010,2016] & Politics, Government \& Economy \\ 
plh0188 & 0.07 & 0.04 & 0.46 & 0.51 & Confident About Future & Health \& Morale \\ 
plh0192 & 0.02 & 0.02 & 0.61 & 0.37 & On The Whole Trust People & Social Life, Social Cohesion \& Trust \\ 
plh0193 & 0.05 & 0.03 & 0.60 & 0.37 & Nowadays Can't Trust Anyone & Social Life, Social Cohesion \& Trust \\ 
plh0194 & 0.11 & 0.07 & 0.57 & 0.35 & Caution Towards Foreigners & Social Life, Social Cohesion \& Trust \\ 
plh0195 & 0.08 & 0.05 & 0.61 & 0.34 & Most People Are Exploitive, Fair & Social Life, Social Cohesion \& Trust \\ 
plh0196 & 0.08 & 0.05 & 0.57 & 0.39 & Most People Are Helpful, Act In Own Interest & Social Life, Social Cohesion \& Trust \\ 
plh0197 & 0.06 & 0.04 & 0.71 & 0.24 & Willingness To Take Risks While Driving & Health \& Morale \\ 
plh0198 & 0.11 & 0.08 & 0.66 & 0.26 & Willingness To Take Risks In Financial Matters & Health \& Morale \\ 
plh0199 & 0.02 & 0.01 & 0.70 & 0.28 & Willingness To Take Risks In Leisure And Sports & Health \& Morale \\ 
plh0200 & 0.06 & 0.04 & 0.67 & 0.29 & Willingness To Take Risks In Occupation & Occupation \& Education \\ 
plh0201 & 0.18 & 0.14 & 0.64 & 0.22 & Willingness To Take Health Risks & Health \& Morale \\ 
plh0202 & 0.21 & 0.17 & 0.64 & 0.19 & Willingness To Take Risks In Trusting Other People & Social Life, Social Cohesion \& Trust \\ 
plh0204\_h & 0.06 & 0.04 & 0.58 & 0.38 & Self-Evaluation: Risk & Health \& Morale \\ 
plh0212 & 0.11 & 0.08 & 0.62 & 0.31 & Thorough Worker & Health \& Morale \\ 
plh0213 & 0.10 & 0.08 & 0.70 & 0.22 & Am Communicative & Health \& Morale \\ 
plh0214 & 0.12 & 0.09 & 0.63 & 0.28 & Am Sometimes Too Coarse With Others & Health \& Morale \\ 
plh0215 & 0.12 & 0.09 & 0.64 & 0.27 & Am Original & Health \& Morale \\ 
plh0216 & 0.07 & 0.05 & 0.64 & 0.31 & Worry A Lot & Health \& Morale \\ 
plh0217 & 0.11 & 0.08 & 0.58 & 0.35 & Able To Forgive & Health \& Morale \\ 
plh0218 & 0.08 & 0.06 & 0.65 & 0.29 & Tend To Be Lazy & Health \& Morale \\ 
plh0219 & 0.11 & 0.08 & 0.67 & 0.24 & Am Sociable & Health \& Morale \\ 
plh0220 & 0.09 & 0.07 & 0.70 & 0.23 & Value Artistic Experiences & Health \& Morale \\ 
plh0221 & 0.11 & 0.08 & 0.64 & 0.28 & Somewhat Nervous & Health \& Morale \\ 
plh0222 & 0.14 & 0.09 & 0.58 & 0.33 & Carry Out Tasks Efficiently & Health \& Morale \\ 
plh0223 & 0.12 & 0.08 & 0.63 & 0.29 & Reserved & Health \& Morale \\ 
plh0224 & 0.14 & 0.09 & 0.58 & 0.32 & Friendly With Others & Health \& Morale \\ 
plh0225 & 0.11 & 0.08 & 0.65 & 0.27 & Have Lively Imagination & Health \& Morale \\ 
plh0226 & 0.12 & 0.09 & 0.63 & 0.28 & Deal Well With Stress & Health \& Morale \\ 
plh0244 & 0.00 & 0.00 & 0.58 & 0.42 & Attitude Towards Future & Health \& Morale \\ 
plh0253 & 0.11 & 0.08 & 0.70 & 0.22 & Personal Patience & Health \& Morale \\ 
plh0254 & 0.10 & 0.07 & 0.65 & 0.28 & Personal Impulsivness & Health \& Morale \\ 
plh0255 & 0.14 & 0.11 & 0.67 & 0.22 & Inquisitive & Health \& Morale \\ 
plh0332 & 0.06 & 0.04 & 0.60 & 0.35 & Concerns Child Health & Health \& Morale \\ 
plh0335 & 0.10 & 0.07 & 0.65 & 0.28 & Worried About Own Retirement Pension & Politics, Government \& Economy \\ 
plh0336 & 0.17 & 0.10 & 0.51 & 0.39 & Worried About Cohesion In Society & Social Life, Social Cohesion \& Trust \\ 
plj0040 & 0.12 & 0.08 & 0.60 & 0.31 & Ties To Local Area & National Identity, Ethnicity \& Immigration \\ 
plj0043 & 0.09 & 0.07 & 0.70 & 0.23 & Ties To Local Area & National Identity, Ethnicity \& Immigration \\ 
plj0046 & 0.09 & 0.05 & 0.54 & 0.41 & Worried About Immigration To Germany & National Identity, Ethnicity \& Immigration \\ 
plj0047 & 0.10 & 0.05 & 0.44 & 0.51 & Worried About Hostility To Foreigners & National Identity, Ethnicity \& Immigration \\ 
plj0078 & 0.03 & 0.02 & 0.74 & 0.24 & Feel German & National Identity, Ethnicity \& Immigration \\ 
plj0081 & 0.14 & 0.09 & 0.59 & 0.32 & Connected With Country Of Origin & National Identity, Ethnicity \& Immigration \\ 
plj0082 & 0.08 & 0.03 & 0.41 & 0.55 & Feeling Of Not Belonging & National Identity, Ethnicity \& Immigration \\ 
plj0083 & 0.11 & 0.06 & 0.53 & 0.40 & Feel At Home In Country Of Origin & National Identity, Ethnicity \& Immigration \\ 
\midrule
\multicolumn{7}{l}{\textbf{UKHLS}} \\ 
\midrule
ahvwell & 0.07 & 0.05 & 0.62 & 0.33 & Importance Of Doing Well In Quals & Occupation \& Education \\ 
apvocimp & 0.18 & 0.14 & 0.61 & 0.25 & Importance Of Apprenticeship/Vocational & Occupation \& Education \\ 
carbuy1 & 0.20 & 0.12 & 0.51 & 0.37 & Comfort & Environment \& Climate \\ 
carbuy10 & 0.23 & 0.15 & 0.50 & 0.34 & Safety & Environment \& Climate \\ 
carbuy11 & 0.20 & 0.14 & 0.59 & 0.26 & Speed/Performance & Environment \& Climate \\ 
carbuy12 & 0.12 & 0.08 & 0.56 & 0.36 & Features - Sat-Nav, Cd Player, Music System, Power Steering Etc... & Environment \& Climate \\ 
carbuy2 & 0.18 & 0.10 & 0.47 & 0.43 & Costs - Purchase/Running/Resale Value/Tax/Insurance & Environment \& Climate \\ 
carbuy3 & 0.19 & 0.13 & 0.57 & 0.30 & Small Engine & Environment \& Climate \\ 
carbuy4 & 0.22 & 0.15 & 0.55 & 0.30 & Large Engine & Environment \& Climate \\ 
carbuy5 & 0.22 & 0.16 & 0.56 & 0.29 & Environmentally Friendly/Low Co2 Emissions & Environment \& Climate \\ 
carbuy6 & 0.25 & 0.11 & 0.34 & 0.55 & Electric Car/Van - One That's Plugged Directly Into An Electricity Supply & Environment \& Climate \\ 
carbuy7 & 0.17 & 0.12 & 0.59 & 0.29 & Style/Design/Image Of Brand/Model & Environment \& Climate \\ 
carbuy8 & 0.23 & 0.15 & 0.52 & 0.32 & Interior Space/Functionality/Boot Size & Environment \& Climate \\ 
carbuy9 & 0.17 & 0.09 & 0.44 & 0.47 & Reliability & Environment \& Climate \\ 
civicduty & 0.03 & 0.02 & 0.73 & 0.24 & Sense Of Civic Duty & Politics, Government \& Economy \\ 
colbens1 & 0.15 & 0.13 & 0.75 & 0.12 & Party Sympathy: Conservative & Politics, Government \& Economy \\ 
colbens2 & 0.15 & 0.13 & 0.74 & 0.13 & Party Sympathy: Labour & Politics, Government \& Economy \\ 
colbens3 & 0.17 & 0.14 & 0.68 & 0.18 & Party Sympathy: Liberal Democrats & Politics, Government \& Economy \\ 
demorient & 0.22 & 0.17 & 0.62 & 0.20 & Democratic Orientation & Politics, Government \& Economy \\ 
eumem & 0.12 & 0.12 & 0.85 & 0.03 & Should Uk Remain A Member Of The Eu & Politics, Government \& Economy \\ 
ocimpa & 0.05 & 0.04 & 0.62 & 0.34 & Occupation: Important: Job Security & Occupation \& Education \\ 
ocimpb & 0.08 & 0.06 & 0.69 & 0.26 & Occupation: Important: High Income & Occupation \& Education \\ 
ocimpe & 0.01 & 0.00 & 0.61 & 0.38 & Occupation: Important: Lot Leisure Time & Occupation \& Education \\ 
ocimpf & 0.00 & 0.00 & 0.63 & 0.37 & Occupation: Important: Interesting Job & Occupation \& Education \\ 
ocimpi & 0.06 & 0.04 & 0.69 & 0.27 & Occupation: Important: Contribute To Soc. & Occupation \& Education \\ 
ocimpk & 0.05 & 0.03 & 0.62 & 0.34 & Occupation: Important: Time For Family & Occupation \& Education \\ 
ocimpl & 0.08 & 0.06 & 0.71 & 0.23 & Occupation: Important: Job Help Others & Occupation \& Education \\ 
oprlg & 0.11 & 0.10 & 0.81 & 0.09 & Whether Belong To A Religion & Religion \& Spirituality \\ 
oprlg1\_1 & 0.01 & 0.01 & 0.89 & 0.10 & Religion: Church Of England/Anglican & Religion \& Spirituality \\ 
oprlg1\_11 & 0.01 & 0.01 & 0.98 & 0.01 & Religion: Muslim/Islam & Religion \& Spirituality \\ 
oprlg1\_2 & 0.00 & 0.00 & 0.95 & 0.05 & Religion: Roman Catholic & Religion \& Spirituality \\ 
oprlg3 & 0.07 & 0.06 & 0.84 & 0.10 & Religion Makes A Difference To Life & Religion \& Spirituality \\ 
scenv\_ftst & 0.21 & 0.17 & 0.61 & 0.22 & How Feel About Current Lifestyle And The Environment & Environment \& Climate \\ 
scenv\_grn & 0.00 & 0.00 & 0.65 & 0.35 & Being Green Is An Alternative Lifestyle & Environment \& Climate \\ 
scfannoy & 0.14 & 0.10 & 0.60 & 0.31 & Friends Get On My Nerves & Social Life, Social Cohesion \& Trust \\ 
scfcritic & 0.01 & 0.01 & 0.61 & 0.38 & Friends Criticise Me & Social Life, Social Cohesion \& Trust \\ 
scopecl200 & 0.06 & 0.03 & 0.52 & 0.45 & Affected By Climate Change Next 200 Years & Environment \& Climate \\ 
scopecl30 & 0.04 & 0.02 & 0.58 & 0.40 & Affected By Climate Change Next 30 Years & Environment \& Climate \\ 
scopfamb & 0.07 & 0.05 & 0.72 & 0.23 & Family Suffers If Mother Works Full-Time & Gender \& Family \\ 
scopfamd & 0.03 & 0.02 & 0.61 & 0.37 & Husband And Wife Should Contribute To Hh Income & Gender \& Family \\ 
scopfamf & 0.03 & 0.02 & 0.71 & 0.27 & Husband Should Earn, Wife Should Stay At Home & Gender \& Family \\ 
scopngbha & 0.19 & 0.15 & 0.64 & 0.22 & Belong To Neighbourhood & Social Life, Social Cohesion \& Trust \\ 
scopngbhb & 0.16 & 0.13 & 0.65 & 0.23 & Local Friends Mean A Lot & Social Life, Social Cohesion \& Trust \\ 
scopngbhc & 0.18 & 0.14 & 0.63 & 0.23 & Advice Obtainable Locally & Social Life, Social Cohesion \& Trust \\ 
scopngbhd & 0.17 & 0.13 & 0.63 & 0.24 & Can Borrow Things From Neighbours & Social Life, Social Cohesion \& Trust \\ 
scopngbhe & 0.11 & 0.07 & 0.57 & 0.36 & Willing To Improve Neighbourhood & Social Life, Social Cohesion \& Trust \\ 
scopngbhf & 0.20 & 0.15 & 0.60 & 0.24 & Plan To Stay In Neighbourhood & Social Life, Social Cohesion \& Trust \\ 
scopngbhg & 0.19 & 0.15 & 0.63 & 0.23 & Am Similar To Others In Neighbourhood & Social Life, Social Cohesion \& Trust \\ 
scopngbhh & 0.18 & 0.14 & 0.65 & 0.21 & Talk Regularly To Neighbours & Social Life, Social Cohesion \& Trust \\ 
scpundstnd & 0.09 & 0.06 & 0.65 & 0.28 & Spouse/Partner Understands Way I Feel & Gender \& Family \\ 
scrannoy & 0.20 & 0.16 & 0.63 & 0.21 & Family Gets On My Nerves & Gender \& Family \\ 
scrcritic & 0.05 & 0.04 & 0.63 & 0.34 & Family Criticises Me & Gender \& Family \\ 
scrletdwn & 0.21 & 0.15 & 0.59 & 0.26 & Family Lets Me Down & Gender \& Family \\ 
scrrely & 0.18 & 0.14 & 0.63 & 0.22 & Can Rely On Family & Gender \& Family \\ 
vote1 & 0.13 & 0.08 & 0.57 & 0.34 & Supports A Particular Political Party & Politics, Government \& Economy \\ 
vote2 & 0.15 & 0.09 & 0.51 & 0.40 & Closer To One Political Party Than Others & Politics, Government \& Economy \\ 
vote3\_1 & 0.20 & 0.13 & 0.53 & 0.34 & Party Support: Conservatives & Politics, Government \& Economy \\ 
vote3\_2 & 0.14 & 0.09 & 0.55 & 0.36 & Party Support: Labour & Politics, Government \& Economy \\ 
vote3\_3 & 0.22 & 0.13 & 0.48 & 0.39 & Party Support: Liberal Democrat & Politics, Government \& Economy \\ 
vote4\_1 & 0.05 & 0.05 & 0.88 & 0.07 & Party Affinity: Conservatives & Politics, Government \& Economy \\ 
vote4\_2 & 0.06 & 0.05 & 0.86 & 0.09 & Party Affinity: Labour & Politics, Government \& Economy \\ 
vote4\_3 & 0.12 & 0.10 & 0.73 & 0.18 & Party Affinity: Liberal Democrat & Politics, Government \& Economy \\ 
vote5 & 0.12 & 0.08 & 0.57 & 0.35 & Strength Of Support For Stated Party & Politics, Government \& Economy \\ 
vote6 & 0.06 & 0.04 & 0.73 & 0.23 & Level Of Interest In Politics & Politics, Government \& Economy \\ 
voteintent & 0.06 & 0.04 & 0.76 & 0.20 & Voting Intention & Politics, Government \& Economy \\ 
votenorm & 0.03 & 0.02 & 0.56 & 0.42 & Voting As A Social Norm & Politics, Government \& Economy \\ 
\bottomrule
\end{longtable}
\end{scriptsize}
```

\elandscape
\newpage

# Supplemental Materials B

Table B1 presents the full set of coefficients from an OLS model regressing $\omega$ on multiple covariates, described in Figure 2 in the main manuscript.

```{=latex}
\begin{table}[ht]
\begin{center}
\caption*{Table B1: Regression Model} 
\label{tab:sometab}
\begin{tabular}{l c c}
\hline
 & Estimates & SE \\
\hline
N Part. (log)                                       & $0.01$        
                                                    & $(0.01)$      \\
Duration                                            & $-0.00$ 
                                                    & $(0.00)^{***}$ \\
N Waves (Part.)                                     & $0.00$        
                                                    & $(0.00)$      \\
N Waves (Survey)                                    & $-0.00$       
                                                    & $(0.00)$      \\
Wave 1 (Decade)                                     & $-0.01$  
                                                    & $(0.00)^{**}$  \\
Panels: BHPS                                        & $0.08$        
                                                    & $(0.05)$      \\
Panels: GSS                                         & $0.04$        
                                                    & $(0.05)$      \\
Panels: HILDA                                       & $0.08$        
                                                    & $(0.05)$      \\
Panels: PSID                                        & $0.09$    
                                                    & $(0.04)^{*}$  \\
Panels: SHP                                         & $0.06$        
                                                    & $(0.05)$      \\
Panels: SOEP                                        & $0.06$        
                                                    & $(0.05)$      \\
Panels: UKHLS                                       & $0.09$        
                                                    & $(0.06)$      \\
Topics: Environment and Climate                     & $0.04$   
                                                    & $(0.01)^{**}$      \\
Topics: Gender and Family                           & $0.00$        
                                                    & $(0.01)$      \\
Topics: Health and Morale                           & $0.03$   
                                                    & $(0.01)^{**}$      \\
Topics: National Identity, Ethnicity, Immigration   & $0.02$        
                                                    & $(0.01)$      \\
Topics: Occupation and Education                    & $-0.02$       
                                                    & $(0.02)$      \\
Topics: Religion and Spirituality                   & $-0.00$       
                                                    & $(0.01)$      \\
Topics: Social Life, Social Cohesion and Trust      & $0.03$  
                                                    & $(0.01)^{***}$      \\
Topics: Subjective SES                              & $0.03$    
                                                    & $(0.01)^{*}$      \\
\hline
Adjusted R$^2$                                      & $0.77$     &   \\
N                                                   & $609$      &  \\
\hline
\multicolumn{2}{l}{\scriptsize{$^{***}p<0.001$; $^{**}p<0.01$; $^{*}p<0.05$}}
\end{tabular}
\end{center}
\end{table}
```

\newpage

# Supplemental Materials C

Figure C1 presents the alternative strategy we used to understand the effects of duration on intraindividual change. First, we restricted the sample such that we have at least 4 waves of observations for all participants. This reduced the number of items to 249. We then fitted the Life Course Adaption models to this sample. Second, we removed the last wave from all the observations, and refit the same models once more. In the final step, we calculated the $\omega$ values for each item. The Figure C1 shows a scatterplot for these two sets of observations.

```{=latex}
\begin{figure}[ht]
\begin{center}
\caption*{Figure C1: Difference in $\omega$ Across Two Sample Specifications} 
```
```{r fig3, out.width = "5in"}
knitr::include_graphics("../figures/figure_S3.png")
```
```{=latex}
\end{center}
\end{figure}
```

### 2023-09-27 supplementals/manuscript_supplementals.tex
supplemental materials

--- 2023-09-27 supplementals/manuscript_supplementals.Rmd
@@
----
-title: |
-  | Supplemental Materials for
-  | Quantifying Stability and Change in Personal Culture Using Panel Data
-author: 
-- 
-output:
-  pdf_document:
-    keep_tex: true
-    latex_engine: xelatex
-    toc: true
-geometry: margin=1in 
-mainfont: Times New Roman
-fontsize: 12pt
-header-includes:
-  \usepackage{fontspec}
-  \usepackage{longtable}
-  \usepackage{multirow}
-  \usepackage{array}
-  \usepackage{booktabs}
-  \usepackage{dcolumn}
-  \usepackage{lscape}
-  \usepackage{caption}
-  \newcommand{\blandscape}{\begin{landscape}}
-  \newcommand{\elandscape}{\end{landscape}}
----
+% Options for packages loaded elsewhere
+\PassOptionsToPackage{unicode}{hyperref}
+\PassOptionsToPackage{hyphens}{url}
+%
+\documentclass[
+  12pt,
+]{article}
+\usepackage{amsmath,amssymb}
+\usepackage{iftex}
+\ifPDFTeX
+  \usepackage[T1]{fontenc}
+  \usepackage[utf8]{inputenc}
+  \usepackage{textcomp} % provide euro and other symbols
+\else % if luatex or xetex
+  \usepackage{unicode-math} % this also loads fontspec
+  \defaultfontfeatures{Scale=MatchLowercase}
+  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
+\fi
+\usepackage{lmodern}
+\ifPDFTeX\else
+  % xetex/luatex font selection
+  \setmainfont[]{Times New Roman}
+\fi
+% Use upquote if available, for straight quotes in verbatim environments
+\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
+\IfFileExists{microtype.sty}{% use microtype if available
+  \usepackage[]{microtype}
+  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
+}{}
+\makeatletter
+\@ifundefined{KOMAClassName}{% if non-KOMA class
+  \IfFileExists{parskip.sty}{%
+    \usepackage{parskip}
+  }{% else
+    \setlength{\parindent}{0pt}
+    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
+}{% if KOMA class
+  \KOMAoptions{parskip=half}}
+\makeatother
+\usepackage{xcolor}
+\usepackage[margin=1in]{geometry}
+\usepackage{graphicx}
+\makeatletter
+\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
+\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
+\makeatother
+% Scale images if necessary, so that they will not overflow the page
+% margins by default, and it is still possible to overwrite the defaults
+% using explicit options in \includegraphics[width, height, ...]{}
+\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
+% Set default figure placement to htbp
+\makeatletter
+\def\fps@figure{htbp}
+\makeatother
+\setlength{\emergencystretch}{3em} % prevent overfull lines
+\providecommand{\tightlist}{%
+  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
+\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
+\usepackage{fontspec} \usepackage{longtable} \usepackage{multirow} \usepackage{array} \usepackage{booktabs} \usepackage{dcolumn} \usepackage{lscape} \usepackage{caption} \newcommand{\blandscape}{\begin{landscape}} \newcommand{\elandscape}{\end{landscape}}
+\ifLuaTeX
+  \usepackage{selnolig}  % disable illegal ligatures
+\fi
+\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
+\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
+\urlstyle{same}
+\hypersetup{
+  hidelinks,
+  pdfcreator={LaTeX via pandoc}}
 
-```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = F, message = F, warning = F, fig.align = "center")
-```
+\title{Supplemental Materials for\\
+Quantifying Stability and Change in Personal Culture Using Panel Data}
+\author{}
+\date{\vspace{-2.5em}}
 
+\begin{document}
+\maketitle
+
+{
+\setcounter{tocdepth}{2}
+\tableofcontents
+}
 \newpage
-\blandscape
+\begin{landscape}
 
-# Supplemental Materials A
+\hypertarget{supplemental-materials-a}{%
+\section{Supplemental Materials A}\label{supplemental-materials-a}}
 
-In Table A1, we present the survey items, corresponding labels and topic indicators, as well as the variances calculated from our estimations: V(C) refers to intraindividual change, V(D) refers to interindividual differences, and V(R) refers to residual variance.
-We also present the $\omega$ values.
+In Table A1, we present the survey items, corresponding labels and topic
+indicators, as well as the variances calculated from our estimations:
+V(C) refers to intraindividual change, V(D) refers to interindividual
+differences, and V(R) refers to residual variance. We also present the
+\(\omega\) values.
 
-```{=latex}
 \begin{scriptsize}
@@
 \end{scriptsize}
-```
 
-\elandscape
+\end{landscape}
 \newpage
 
-# Supplemental Materials B
+\hypertarget{supplemental-materials-b}{%
+\section{Supplemental Materials B}\label{supplemental-materials-b}}
 
-Table B1 presents the full set of coefficients from an OLS model regressing $\omega$ on multiple covariates, described in Figure 2 in the main manuscript.
+Table B1 presents the full set of coefficients from an OLS model
+regressing \(\omega\) on multiple covariates, described in Figure 2 in
+the main manuscript.
 
-```{=latex}
 \begin{table}[ht]
@@
 \end{table}
-```
 
@@
 
-# Supplemental Materials C
+\hypertarget{supplemental-materials-c}{%
+\section{Supplemental Materials C}\label{supplemental-materials-c}}
 
-Figure C1 presents the alternative strategy we used to understand the effects of duration on intraindividual change.
-First, we restricted the sample such that we have at least 4 waves of observations for all participants.
-This reduced the number of items to 249.
-We then fitted the Life Course Adaption models to this sample.
-Second, we removed the last wave from all the observations, and refit the same models once more.
-In the final step, we calculated the $\omega$ values for each item.
-The Figure C1 shows a scatterplot for these two sets of observations.
+Figure C1 presents the alternative strategy we used to understand the
+effects of duration on intraindividual change. First, we restricted the
+sample such that we have at least 4 waves of observations for all
+participants. This reduced the number of items to 249. We then fitted
+the Life Course Adaption models to this sample. Second, we removed the
+last wave from all the observations, and refit the same models once
+more. In the final step, we calculated the \(\omega\) values for each
+item. The Figure C1 shows a scatterplot for these two sets of
+observations.
 
-```{=latex}
 \begin{figure}[ht]
@@
 \caption*{Figure C1: Difference in $\omega$ Across Two Sample Specifications}
-```
-```{r fig3, out.width = "5in"}
-knitr::include_graphics("../figures/figure_S3.png")
-```
-```{=latex}
+
+\begin{center}\includegraphics[width=5in]{../figures/figure_S3} \end{center}
+
 \end{center}
 \end{figure}
-```
+
+\end{document}

### 2023-09-27 tables/table_A1.tex
supplemental materials

--- 2023-09-27 supplementals/manuscript_supplementals.Rmd
@@
----
-title: |
-  | Supplemental Materials for
-  | Quantifying Stability and Change in Personal Culture Using Panel Data
-author: 
-- 
-output:
-  pdf_document:
-    keep_tex: true
-    latex_engine: xelatex
-    toc: true
-geometry: margin=1in 
-mainfont: Times New Roman
-fontsize: 12pt
-header-includes:
-  \usepackage{fontspec}
-  \usepackage{longtable}
-  \usepackage{multirow}
-  \usepackage{array}
-  \usepackage{booktabs}
-  \usepackage{dcolumn}
-  \usepackage{lscape}
-  \usepackage{caption}
-  \newcommand{\blandscape}{\begin{landscape}}
-  \newcommand{\elandscape}{\end{landscape}}
----
-
-```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = F, message = F, warning = F, fig.align = "center")
-```
-
-\newpage
-\blandscape
-
-# Supplemental Materials A
-
-In Table A1, we present the survey items, corresponding labels and topic indicators, as well as the variances calculated from our estimations: V(C) refers to intraindividual change, V(D) refers to interindividual differences, and V(R) refers to residual variance.
-We also present the $\omega$ values.
-
-```{=latex}
-\begin{scriptsize}
-\begin{longtable}{|p{1.75in}|p{0.3in}|p{0.3in}|p{0.3in}|p{0.3in}|p{2.5in}|p{2.5in}}
-\caption*{Table A1: The Descriptives of Variables Used}
-\\
+\begin{longtable}{l|rrrrll}
+\caption*{
+{\large \textbf{Descriptives of variables}}
+} \\ 
 \toprule
-\multicolumn{1}{l}{} & \emph{\textbf{ω}} & \textbf{V(C)} & \textbf{V(D)} & \textbf{V(R)} & \textbf{Label} & \textbf{Topic} \\ 
+\multicolumn{1}{l}{} & \emph{\textbf{ω}} & \textbf{Within variance} & \textbf{Between variance} & \textbf{Residual variance} & \textbf{Label} & \textbf{Topic} \\ 
 \midrule
@@
 \end{longtable}
-\end{scriptsize}
-```
-
-\elandscape
-\newpage
-
-# Supplemental Materials B
-
-Table B1 presents the full set of coefficients from an OLS model regressing $\omega$ on multiple covariates, described in Figure 2 in the main manuscript.
-
-```{=latex}
-\begin{table}[ht]
-\begin{center}
-\caption*{Table B1: Regression Model} 
-\label{tab:sometab}
-\begin{tabular}{l c c}
-\hline
- & Estimates & SE \\
-\hline
-N Part. (log)                                       & $0.01$        
-                                                    & $(0.01)$      \\
-Duration                                            & $-0.00$ 
-                                                    & $(0.00)^{***}$ \\
-N Waves (Part.)                                     & $0.00$        
-                                                    & $(0.00)$      \\
-N Waves (Survey)                                    & $-0.00$       
-                                                    & $(0.00)$      \\
-Wave 1 (Decade)                                     & $-0.01$  
-                                                    & $(0.00)^{**}$  \\
-Panels: BHPS                                        & $0.08$        
-                                                    & $(0.05)$      \\
-Panels: GSS                                         & $0.04$        
-                                                    & $(0.05)$      \\
-Panels: HILDA                                       & $0.08$        
-                                                    & $(0.05)$      \\
-Panels: PSID                                        & $0.09$    
-                                                    & $(0.04)^{*}$  \\
-Panels: SHP                                         & $0.06$        
-                                                    & $(0.05)$      \\
-Panels: SOEP                                        & $0.06$        
-                                                    & $(0.05)$      \\
-Panels: UKHLS                                       & $0.09$        
-                                                    & $(0.06)$      \\
-Topics: Environment and Climate                     & $0.04$   
-                                                    & $(0.01)^{**}$      \\
-Topics: Gender and Family                           & $0.00$        
-                                                    & $(0.01)$      \\
-Topics: Health and Morale                           & $0.03$   
-                                                    & $(0.01)^{**}$      \\
-Topics: National Identity, Ethnicity, Immigration   & $0.02$        
-                                                    & $(0.01)$      \\
-Topics: Occupation and Education                    & $-0.02$       
-                                                    & $(0.02)$      \\
-Topics: Religion and Spirituality                   & $-0.00$       
-                                                    & $(0.01)$      \\
-Topics: Social Life, Social Cohesion and Trust      & $0.03$  
-                                                    & $(0.01)^{***}$      \\
-Topics: Subjective SES                              & $0.03$    
-                                                    & $(0.01)^{*}$      \\
-\hline
-Adjusted R$^2$                                      & $0.77$     &   \\
-N                                                   & $609$      &  \\
-\hline
-\multicolumn{2}{l}{\scriptsize{$^{***}p<0.001$; $^{**}p<0.01$; $^{*}p<0.05$}}
-\end{tabular}
-\end{center}
-\end{table}
-```
-
-\newpage
-
-# Supplemental Materials C
-
-Figure C1 presents the alternative strategy we used to understand the effects of duration on intraindividual change.
-First, we restricted the sample such that we have at least 4 waves of observations for all participants.
-This reduced the number of items to 249.
-We then fitted the Life Course Adaption models to this sample.
-Second, we removed the last wave from all the observations, and refit the same models once more.
-In the final step, we calculated the $\omega$ values for each item.
-The Figure C1 shows a scatterplot for these two sets of observations.
-
-```{=latex}
-\begin{figure}[ht]
-\begin{center}
-\caption*{Figure C1: Difference in $\omega$ Across Two Sample Specifications} 
-```
-```{r fig3, out.width = "5in"}
-knitr::include_graphics("../figures/figure_S3.png")
-```
-```{=latex}
-\end{center}
-\end{figure}
-```

### 2023-09-27 manuscript/manuscript.Rmd
ceiled to capped

@@
 \textit{Notes:} The model estimates the proportion of systematic variance attributable to intrapersonal change.
-It includes the log of the number of participants, the number of waves the variable was asked (ceiled at t = 10 for each item), the date of the first wave (measured as the year minus 1968 divided by 10), number of waves the question was asked averaged across participants, duration in years per item averaged across participants, and panel and topic fixed effects.
+It includes the log of the number of participants, the number of waves the variable was asked (capped at t = 10 for each item), the date of the first wave (measured as the year minus 1968 divided by 10), number of waves the question was asked averaged across participants, duration in years per item averaged across participants, and panel and topic fixed effects.
 Coefficients are estimated from suppressed intercept model based on predictions at participants' wave mid-point.

### 2023-09-27 manuscript/manuscript.Rmd
color figs, inline, and endnotes.

@@
 fontsize: 12pt
-header-includes: \usepackage{setspace}\doublespacing
+header-includes:
+  \usepackage{setspace}\doublespacing
+  \usepackage{endnotes}
+  \renewcommand{\enotesize}{\normalsize}
+  \let\footnote=\endnote
 bibliography: kileybib.bib
@@
 ```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = F, message = F, warning = F, fig.align = "center")
+knitr::opts_chunk$set(echo = F, message = F, warning = F)
 ```
@@
 
-```{=tex}
-\begin{center}
-<about here> 
-Table 1: The Description of the Data Sources
-\end{center}
+```{r, message = FALSE}
+
+library(tidyverse)
+library(knitr)
+library(kableExtra)
+
+table <- tibble(
+  Country = c("Australia", "Germany", "Great Britian", "Great Britian", "Switzerland", "United States", "United States"),
+  Survey = c("Household, Income and Labor Dynamics in Australia (HILDA)",
+             "Socio-Economic Panel Study (SOEP)",
+             "British Household Panel Survey (BHPS)",
+             "Understanding Society/UK Household Longitudinal Study (UKHLS)",
+             "Swiss Household Panel (SHP)",
+             "General Social Survey (GSS)",
+             "Panel Study of Income Dynamics (PSID)"),
+  Period = c("2001-2021", "1984-2020", "1991-2008", "2009-2020", "1999-2019", "2006-2012", "1968-2019"),
+  Outcomes = c(30, 122, 83, 66, 77, 183, 48)
+)
+
+kableExtra::kable(table,
+                  booktabs = TRUE,
+                  format = "latex",
+                  align = "llcc",
+                  caption = "The Description of the Data Sources") |> 
+  kable_styling(latex_options = c("scale_down", "HOLD_position"))
+
 ```
+
 ## Life Course Adaption Model
@@
 
-```{=tex}
+```{=latex}
+\begin{figure}[htp]
 \begin{center}
-<about here>
-Figure 1: Proportions of Systematic Variance in Personal Culture
+\caption{Proportions of Systematic Variance in Personal Culture}
+```
+```{r fig1, out.width = "100%", fig.pos = "H"}
+knitr::include_graphics("../figures/figure_1.png")
+```
+```{=latex}
 \end{center}
+\footnotesize{\textit{Notes:} The figure shows $\omega$ and 1 - $\omega$ as the proportion of systematic variance attributable to intrapersonal change and interpersonal differences.
+See Supplemental Materials A for the full set of item values.}
+\end{figure}
 ```
+
 Across all questions, interpersonal differences explain a much larger share of the systematic variance in responses than intrapersonal change.
@@
 
-```{=tex}
+```{=latex}
+\begin{figure}[htp]
 \begin{center}
-<about here>
-Figure 2: Regression Model Estimating $\omega$
+\caption{Regression Model Estimating $\omega$}
+```
+```{r fig2, out.width = "450px"}
+knitr::include_graphics("../figures/figure_2.png")
+```
+```{=latex}
 \end{center}
+\footnotesize{\textit{Notes:} The model estimates the proportion of systematic variance attributable to intrapersonal change.
+It includes the log of the number of participants, the number of waves the variable was asked (capped at t = 10 for each item), the date of the first wave (measured as the year minus 1968 divided by 10), number of waves the question was asked averaged across participants, duration in years per item averaged across participants, and panel and topic fixed effects.
+Coefficients are estimated from suppressed intercept model based on predictions at participants' wave mid-point.
+Survey indicators and item topics not shown.
+See Supplemental Materials B for the full set of coefficients.}
+\end{figure}
 ```
+
 Figure 2 shows that the more response options respondents are given and the larger the sample, the larger the values of $\omega$.
@@
 
-```{=tex}
+```{=latex}
+\begin{figure}[hbt]
 \begin{center}
-<about here>
-Figure 3: Difference in $\omega$ Across College Graduates and Non-Graduates
+\caption{Difference in $\omega$ Across College Graduates and Non-Graduates}
+```
+```{r fig3, out.width = "450px"}
+knitr::include_graphics("../figures/figure_3.png")
+```
+
+```{=latex}
 \end{center}
+\footnotesize{\textit{Notes:} The figure shows the difference in $\omega$ values across college graduates and college non-graduates.
+Values above (below) 0 means that those with college degree have higher (lower) variance of intrapersonal change.
+The dashed red line marks 0 difference.}
+\end{figure}
 ```
+
 There is a clear pattern in Figure 3: for more than 80 percent of GSS items explored here, intrapersonal change is a larger component of systematic variance for people without a college degree than for people with a college degree (i.e., the difference is negative).
@@
 
-```{=tex}
+```{=latex}
+\begin{figure}[hbt]
 \begin{center}
-<about here>
-Figure 4: Difference in $\omega$ Across College Graduates and Non-Graduates on Political Culture
+\caption{Difference in $\omega$ Across College Graduates and Non-Graduates on Political Culture}
+```
+```{r fig4, out.width = "450px"}
+knitr::include_graphics("../figures/figure_4.png")
+```
+
+```{=latex}
 \end{center}
+\footnotesize{\textit{Notes:} $\omega$ values across college graduates and college non-graduates on 8 political culture items from the General Social Survey (2006-2014).}
+\end{figure}
 ```
+
 On all eight questions presented in Figure 4, $\omega$ values are greater for people without a college degree than for people with a college degree, meaning intrapersonal change explains less systematic variance for people with a college degree than for people without one.
@@
 
-# Endnotes
+\theendnotes
 
@@
 :::
-
-\newpage
-
-# Tables and Figures
-
-## Tables
-
-```{r, message = FALSE}
-
-library(tidyverse)
-library(knitr)
-library(kableExtra)
-
-table <- tibble(
-  Country = c("Australia", "Germany", "Great Britian", "Great Britian", "Switzerland", "United States", "United States"),
-  Survey = c("Household, Income and Labor Dynamics in Australia (HILDA)",
-             "Socio-Economic Panel Study (SOEP)",
-             "British Household Panel Survey (BHPS)",
-             "Understanding Society/UK Household Longitudinal Study (UKHLS)",
-             "Swiss Household Panel (SHP)",
-             "General Social Survey (GSS)",
-             "Panel Study of Income Dynamics (PSID)"),
-  Period = c("2001-2021", "1984-2020", "1991-2008", "2009-2020", "1999-2019", "2006-2012", "1968-2019"),
-  Outcomes = c(30, 122, 83, 66, 77, 183, 48)
-)
-
-kableExtra::kable(table, 
-                  format = "latex", 
-                  caption = "The Description of the Data Sources") |> 
-  kable_styling(latex_options = c("scale_down", "HOLD_position"))
-
-```
-
-\newpage
-
-## Figures
-
-```{=latex}
-\begin{figure}[ht]
-\caption{Proportions of Systematic Variance in Personal Culture}
-```
-```{r fig1, fig.height=2}
-knitr::include_graphics("../figures/figure_1_bw.png")
-```
-
-```{=latex}
-\textit{Notes:} The figure shows $\omega$ and 1 - $\omega$ as the proportion of systematic variance attributable to intrapersonal change and interpersonal differences, respectively.
-See Supplemental Materials A for the full set of item values.
-\end{figure}
-```
-\newpage
-
-```{=latex}
-\begin{figure}[ht]
-\caption{Regression Model Estimating $\omega$}
-```
-```{r fig2, out.width = "300px"}
-knitr::include_graphics("../figures/figure_2.png")
-```
-
-```{=latex}
-\textit{Notes:} The model estimates the proportion of systematic variance attributable to intrapersonal change.
-It includes the log of the number of participants, the number of waves the variable was asked (capped at t = 10 for each item), the date of the first wave (measured as the year minus 1968 divided by 10), number of waves the question was asked averaged across participants, duration in years per item averaged across participants, and panel and topic fixed effects.
-Coefficients are estimated from suppressed intercept model based on predictions at participants' wave mid-point.
-Survey indicators and item topics not shown.
-See Supplemental Materials B for the full set of coefficients.
-\end{figure}
-```
-\newpage
-
-```{=latex}
-\begin{figure}[ht]
-\caption{Difference in $\omega$ Across College Graduates and Non-Graduates}
-```
-```{r fig3, out.width = "450px"}
-knitr::include_graphics("../figures/figure_3_bw.png")
-```
-
-```{=latex}
-\textit{Notes:} The figure shows the difference in $\omega$ values across college graduates and college non-graduates.
-Values above (below) 0 means that those with college degree have higher (lower) variance of intrapersonal change.
-The dashed red line marks 0 difference.
-\end{figure}
-```
-\newpage
-
-```{=latex}
-\begin{figure}[ht]
-\caption{Difference in $\omega$ Across College Graduates and Non-Graduates on Political Culture}
-```
-```{r fig4, out.width = "450px"}
-knitr::include_graphics("../figures/figure_4_bw.png")
-```
-
-```{=latex}
-\textit{Notes:} $\omega$ values across college graduates and college non-graduates on 8 political culture items from the General Social Survey (2006-2014).
-\end{figure}
-```

### 2023-09-27 manuscript/manuscript.tex
color figs, inline, and endnotes.

@@
 \newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}
-\usepackage{setspace}\doublespacing
+\usepackage{setspace}\doublespacing \usepackage{endnotes} \renewcommand{\enotesize}{\normalsize} \let\footnote=\endnote
 \usepackage{booktabs}
@@
 
-Our method can also help address substantive questions about personal
-change. To illustrate this, we investigate differences in the relative
-importance of intrapersonal change on political items for people with
-and without a college degree. We find that change is less pronounced for
-college-educated respondents, suggesting that college solidifies
-political dispositions rather than fostering an openness to new
+In addition to advancing a theoretical debate, our method can also help
+address substantive questions. To illustrate how it might be used by a
+researcher with a specific topical interest, we investigate differences
+in the relative importance of intrapersonal change for people with and
+without a college degree. We find a smaller amount of systematic change
+among college-educated respondents, suggesting that college solidifies
+one's personal culture rather than fostering an openness to new
 information. This offers a novel empirical basis to theorize about the
-role of college in shaping personal culture . In summary, we believe
-that this approach has the potential to advance the theoretical debate
-on the role of environments, social identities, and roles in
-facilitating personal cultural change and stability.
+role of university education in shaping personal culture.
 
 Taken together, to goal of this paper is to advance---and, hopefully,
-transcend---debates about \emph{whether} adults change. They do change,
-at least a little, on most things. Our method can help quantify exactly
-how much.
+transcend---debates about \emph{whether} adults change. People do
+change, at least a little, on most things. Our method can help quantify
+exactly how much.
 
@@
 
-Current debates often narrow down to ask whether people \emph{ever}
-change, which is rarely theoretically interesting since the answer is
-almost always ``yes.'' Over sufficient time and in a large enough
-sample, researchers will likely observe some evidence for durable
-changes in individuals. A lack of evidence for change may simply due to
-a lack of clarity in the question asked or a low level of resolution in
-the possible answers, or that the question simply was not asked for long
-enough or to enough people. Conversely, finding evidence of change tells
-us little about how much change is present in a population and how
-meaningful that change is for social differences.
+The current debate can be simplified to asking whether people
+\emph{ever} change. This is rarely theoretically interesting since the
+answer is almost always ``yes.'' Over sufficient time and in a large
+enough sample, researchers will likely observe some evidence for durable
+individual change. A lack of evidence for change may be due to a poor
+survey question, a low resolution in response options, or that the
+question simply was not asked for long enough or to enough people.
+Conversely, finding evidence for some change tells us only that it was
+``not zero.'' From that alone, we learn little about how much change has
+happened in a population.
 
@@
 change a little, or might undergo significant shifts in their personal
-culture. Lersch's LCAM formally does this by modeling each individual as
-following their own trajectory over time. But rather than debating
-whether this model fits a particular data set better than a model that
-assumes that people never change, researchers should quantify individual
-changes over time and compare those to the differences that exist
-between people at the panel's start. In other words, the focus should
-shift from asking, ``do people change?'' to asking, ``what are the
-relative contributions of change and initial differences toward
-explaining differences in personal culture?''
+culture. Lersch's LCAM does this by modeling each individual as
+following their own linear trajectory. But rather than debating whether
+this model fits a particular data set better than a model that assumes
+that people never change, researchers should quantify these individual
+changes and compare those to the differences that exist between people
+at the panel's start. In other words, the focus should shift from
+asking, ``do people change?'' to asking, ``what are the relative
+contributions of change and initial differences for explaining
+differences in personal culture?''
 
@@
 experiences of people prior to entering the panel survey that shape how
-they answer questions, what Lersch calls ``early imprinting.'' While
-commonly associated with experiences during an early formative period
-that result in settled dispositions, this variation could also reflect
+they answer questions. Lersch calls this ``early imprinting.'' While
+commonly associated with experiences during a formative period that
+result in settled dispositions, this variation could also reflect
 experiences that happen at any stage in life as long as they predate the
-panel and consistently impact subsequent responses. For instance, for
+panel and consistently affect subsequent responses. For instance, for
 those people entering the panel post-retirement, this ``imprinting''
@@
 Lersch attributes these changes to social triggers such as moving into a
-new environment or adopting new social roles. Additionally, they might
-reflect the diffusion of new cultural forms across social networks, cues
-from political elites or otherwise culturally influential leaders, the
-emergence of issues in politics or culture, or large-scale social
-shifts.
+new environment or adopting new social roles (although he never measures
+them directly). Additionally, they might reflect the diffusion of new
+cultural forms across social networks, cues from political elites or
+otherwise culturally influential leaders, the emergence of issues in
+politics or culture, or large-scale social shifts.
 
 A third metric, often called ``residual variance'' or ``fluctuation,''
-accounts for the remainder of variance in peoples' responses over time.
-This non-durable change emerges for a variety of reasons. For example,
-people might not have clear personal cultural dispositions as measured
-by surveys. Instead, they might internalize a broad set of
-considerations and construct an opinion in the context of the survey
-interview, with different considerations coming to the forefront of
-their cognition each interview (Feldman and Zaller 1992; Tourangeau,
-Rips, and Rasinski 2000; Zaller 1992). This variance in considerations
-can produce variance in responses even if peoples' ``dispositions''
---the full set of considerations they could draw on-- remains stable
-over time. Additionally, this variance can also include measurement
-errors, such as misinterpreted questions, erroneous response selections,
-or responses getting coded incorrectly. While this third process
-reflects important aspects of personal culture, it does not directly
-touch on the ongoing debate outlined here. Because of this we quantify
-this component but focus principally on the other two.
+accounts for the remainder of variance in peoples' responses. This
+non-durable change emerges for a variety of reasons. For example, people
+might not have clear a clear disposition on a particular item as it is
+asked. Instead, they might internalize a broad set of considerations and
+construct an opinion in the context of the survey interview, with
+different considerations coming to the forefront of their cognition each
+interview (Feldman and Zaller 1992; Tourangeau, Rips, and Rasinski 2000;
+Zaller 1992). Additionally, this variance can also include measurement
+error, such as misinterpreted questions, erroneous response selections,
+or responses getting coded incorrectly. While this third metric at least
+partially reflects important processes of personal culture, it does not
+directly touch on the ongoing debate outlined here. Because of this we
+quantify this component but focus principally on the other two.
 
@@
 relative contribution of these two components in explaining cultural
-differences in a population over a specific duration of time. In other
-words, when we look at a population of adults over some time frame, is
-there enough intrapersonal change relative to interpersonal difference
-to view it as an important explanation for why people demonstrate
-cultural differences? Only through quantifying the two processes we can
-get closer to a true answer to this question.
+differences in a population. In other words, when we look at a
+population of adults over some time frame, is there enough intrapersonal
+change relative to interpersonal difference to view it as an important
+explanation for why people differ from each other? Only through
+quantifying the two processes we can get closer to a true answer to this
+question.
 
@@
 past debates; it allows researchers to ask questions that were difficult
-to explore with previous approaches. For example, classifying questions
-based on whether they show durable change overlooks possibly important
-differences among those that do. By quantifying the relative
+to explore with previous approaches. For example, classifying survey
+items based on whether they show durable change overlooks possibly
+important differences among those that do. By quantifying the relative
 contributions of interpersonal differences and intrapersonal change,
@@
 
-\hypertarget{analytical-strategy}{%
-\section{Analytical Strategy}\label{analytical-strategy}}
+\hypertarget{analytic-strategy}{%
+\section{Analytic Strategy}\label{analytic-strategy}}
 
@@
 
-\begin{center}
-<about here> 
-Table 1: The Description of the Data Sources
-\end{center}
+\begin{table}[H]
+
+\caption{\label{tab:unnamed-chunk-1}The Description of the Data Sources}
+\centering
+\resizebox{\linewidth}{!}{
+\begin{tabular}[t]{llcc}
+\toprule
+Country & Survey & Period & Outcomes\\
+\midrule
+Australia & Household, Income and Labor Dynamics in Australia (HILDA) & 2001-2021 & 30\\
+Germany & Socio-Economic Panel Study (SOEP) & 1984-2020 & 122\\
+Great Britian & British Household Panel Survey (BHPS) & 1991-2008 & 83\\
+Great Britian & Understanding Society/UK Household Longitudinal Study (UKHLS) & 2009-2020 & 66\\
+Switzerland & Swiss Household Panel (SHP) & 1999-2019 & 77\\
+\addlinespace
+United States & General Social Survey (GSS) & 2006-2012 & 183\\
+United States & Panel Study of Income Dynamics (PSID) & 1968-2019 & 48\\
+\bottomrule
+\end{tabular}}
+\end{table}
 
@@
 
+\begin{figure}[htp]
 \begin{center}
-<about here>
-Figure 1: Proportions of Systematic Variance in Personal Culture
+\caption{Proportions of Systematic Variance in Personal Culture}
+
+\includegraphics[width=1\linewidth]{../figures/figure_1}
+
 \end{center}
+\footnotesize{\textit{Notes:} The figure shows $\omega$ and 1 - $\omega$ as the proportion of systematic variance attributable to intrapersonal change and interpersonal differences.
+See Supplemental Materials A for the full set of item values.}
+\end{figure}
 
@@
 
+\begin{figure}[htp]
 \begin{center}
-<about here>
-Figure 2: Regression Model Estimating $\omega$
+\caption{Regression Model Estimating $\omega$}
+
+\includegraphics[width=450px]{../figures/figure_2}
+
 \end{center}
+\footnotesize{\textit{Notes:} The model estimates the proportion of systematic variance attributable to intrapersonal change.
+It includes the log of the number of participants, the number of waves the variable was asked (capped at t = 10 for each item), the date of the first wave (measured as the year minus 1968 divided by 10), number of waves the question was asked averaged across participants, duration in years per item averaged across participants, and panel and topic fixed effects.
+Coefficients are estimated from suppressed intercept model based on predictions at participants' wave mid-point.
+Survey indicators and item topics not shown.
+See Supplemental Materials B for the full set of coefficients.}
+\end{figure}
 
@@
 
+\begin{figure}[hbt]
 \begin{center}
-<about here>
-Figure 3: Difference in $\omega$ Across College Graduates and Non-Graduates
+\caption{Difference in $\omega$ Across College Graduates and Non-Graduates}
+
+\includegraphics[width=450px]{../figures/figure_3}
+
 \end{center}
+\footnotesize{\textit{Notes:} The figure shows the difference in $\omega$ values across college graduates and college non-graduates.
+Values above (below) 0 means that those with college degree have higher (lower) variance of intrapersonal change.
+The dashed red line marks 0 difference.}
+\end{figure}
 
@@
 
+\begin{figure}[hbt]
 \begin{center}
-<about here>
-Figure 4: Difference in $\omega$ Across College Graduates and Non-Graduates on Political Culture
+\caption{Difference in $\omega$ Across College Graduates and Non-Graduates on Political Culture}
+
+\includegraphics[width=450px]{../figures/figure_4}
+
 \end{center}
+\footnotesize{\textit{Notes:} $\omega$ values across college graduates and college non-graduates on 8 political culture items from the General Social Survey (2006-2014).}
+\end{figure}
 
@@
 
-\hypertarget{endnotes}{%
-\section{Endnotes}\label{endnotes}}
+\theendnotes
 
@@
 
-\newpage
-
-\hypertarget{tables-and-figures}{%
-\section{Tables and Figures}\label{tables-and-figures}}
-
-\hypertarget{tables}{%
-\subsection{Tables}\label{tables}}
-
-\begin{table}[H]
-
-\caption{\label{tab:unnamed-chunk-1}The Description of the Data Sources}
-\centering
-\resizebox{\linewidth}{!}{
-\begin{tabular}[t]{l|l|l|r}
-\hline
-Country & Survey & Period & Outcomes\\
-\hline
-Australia & Household, Income and Labor Dynamics in Australia (HILDA) & 2001-2021 & 30\\
-\hline
-Germany & Socio-Economic Panel Study (SOEP) & 1984-2020 & 122\\
-\hline
-Great Britian & British Household Panel Survey (BHPS) & 1991-2008 & 83\\
-\hline
-Great Britian & Understanding Society/UK Household Longitudinal Study (UKHLS) & 2009-2020 & 66\\
-\hline
-Switzerland & Swiss Household Panel (SHP) & 1999-2019 & 77\\
-\hline
-United States & General Social Survey (GSS) & 2006-2012 & 183\\
-\hline
-United States & Panel Study of Income Dynamics (PSID) & 1968-2019 & 48\\
-\hline
-\end{tabular}}
-\end{table}
-
-\newpage
-
-\hypertarget{figures}{%
-\subsection{Figures}\label{figures}}
-
-\begin{figure}[ht]
-\caption{Proportions of Systematic Variance in Personal Culture}
-
-\begin{center}\includegraphics[width=20.83in]{../figures/figure_1_bw} \end{center}
-
-\textit{Notes:} The figure shows $\omega$ and 1 - $\omega$ as the proportion of systematic variance attributable to intrapersonal change and interpersonal differences, respectively.
-See Supplemental Materials A for the full set of item values.
-\end{figure}
-
-\newpage
-
-\begin{figure}[ht]
-\caption{Regression Model Estimating $\omega$}
-
-\begin{center}\includegraphics[width=300px]{../figures/figure_2} \end{center}
-
-\textit{Notes:} The model estimates the proportion of systematic variance attributable to intrapersonal change.
-It includes the log of the number of participants, the number of waves the variable was asked (ceiled at t = 10 for each item), the date of the first wave (measured as the year minus 1968 divided by 10), number of waves the question was asked averaged across participants, duration in years per item averaged across participants, and panel and topic fixed effects.
-Coefficients are estimated from suppressed intercept model based on predictions at participants' wave mid-point.
-Survey indicators and item topics not shown.
-See Supplemental Materials B for the full set of coefficients.
-\end{figure}
-
-\newpage
-
-\begin{figure}[ht]
-\caption{Difference in $\omega$ Across College Graduates and Non-Graduates}
-
-\begin{center}\includegraphics[width=450px]{../figures/figure_3_bw} \end{center}
-
-\textit{Notes:} The figure shows the difference in $\omega$ values across college graduates and college non-graduates.
-Values above (below) 0 means that those with college degree have higher (lower) variance of intrapersonal change.
-The dashed red line marks 0 difference.
-\end{figure}
-
-\newpage
-
-\begin{figure}[ht]
-\caption{Difference in $\omega$ Across College Graduates and Non-Graduates on Political Culture}
-
-\begin{center}\includegraphics[width=450px]{../figures/figure_4_bw} \end{center}
-
-\textit{Notes:} $\omega$ values across college graduates and college non-graduates on 8 political culture items from the General Social Survey (2006-2014).
-\end{figure}
-
 \end{document}

### 2023-09-28 manuscript/manuscript.Rmd
Minor tweaks and inclusion of refernces re topics

@@
 In doing so, we compare how much cultural difference is explained by intrapersonal change and interpersonal differences.
-Our results harmonize recent findings by showing that, while most measures of personal culture show change over time, the variance explained by intrapersonal change is typically small compared to interpersonal differences at baseline.
+Our results harmonize recent findings by showing that, while most measures of personal culture show durable change over time, the variance explained by intrapersonal change is typically small compared to that of interpersonal differences at baseline.
 Quantifying these processes offers new ways to address theoretical questions of when and how their relative importance shifts.
-We examplify this approach by exploring how the importance of these processes differs between college graduates and the rest of the population."
+We exemplify this approach by exploring how the importance of these processes differs between college graduates and the rest of the population."
 ---
@@
 
-These seemingly inconsistent findings are partly due to the fact that researchers have taken a "tournament of models" approach to adjudicating different theoretical perspectives [@lersch2023: p.
+These seemingly inconsistent findings are partly due to the fact that researchers have taken a "tournament of models" approach to adjudicate different theoretical perspectives [@lersch2023: p.
 228].
@@
 We argue, however, that asking *whether* people change on an item may cause more problems than it solves.
-Being able to detect *any* change on an item depends not only on sample populations, measurement, statistical methods, and power, but also on definitions of what counts as "change." These differences can limit the possibility of agreement on which "pile" a variable belongs in.
+Being able to detect *any* change on an item depends not only on sample populations, measurement, statistical methods, and power, but also on definitions of what counts as "change." These differences can limit the possibility of agreement on which "pile" a variable belongs.
 Moreover, asking whether there is evidence of change reduces to, "Is there *any* evidence of change in this item?" But such binary inquiries are ill-suited to address the core of the theoretical debate: determining *how much* the process of intrapersonal change contributes to cultural differences.
@@
 This also prevents investigating under what conditions, across which groups, and in what domains do we see differences in the relative importance of intrapersonal change and pre-existing interpersonal differences.
-These questions are about *degree*, not existence; taking this debate forward will require precise measurement rather than declarations of victory for a particular perspective.
+These questions are about *degree*, not existence; to take this debate forward therefore requires precise measurement rather than declarations of victory for a particular perspective.
 
@@
 
-Taken together, to goal of this paper is to advance---and, hopefully, transcend---debates about *whether* adults change.
+Taken together, the goal of this paper is to advance---and, hopefully, transcend---debates about *whether* adults change.
 People do change, at least a little, on most things.
-Our method can help quantify exactly how much.
+Our method can help quantify exactly *how much*.
 
@@
 Kiley and Vaisey [-@kiley2020] found that the majority of items they tested favored the AUM, meaning there was evidence of durable change among adults, a pattern consistent with Lersch's results.
-While Lersch [-@lersch2023] found evidence of durable change on most items, this change only accounted for a small proportion of variance explained.
+Also Lersch [-@lersch2023] found evidence of durable change on most items, but this change only accounted for a small proportion of variance explained.
 On average, people changed only about .07 standard deviations over 10 years.
 This is consistent with Kiley and Vaisey's finding that durable change is often small.
-And a quarter of the items Lersch studied still preferred the SDM, suggesting that even with different assumptions assumptions about change and more extensive data, many questions show no evidence of durable change.
-Likewise, in studies showing evidence of durable change in specific measures of personal culture, the change is often small.
+Moreover, a quarter of the items Lersch studied still preferred the SDM, suggesting that even with different assumptions assumptions about change and more extensive data, many questions show no evidence of durable change.
+Consistent with this, in studies showing evidence of durable change in specific measures of personal culture, the change is often small.
 For example, Brocic and Miles [-@brocic2021] estimate that completing graduate degrees in humanities, arts, and social sciences shift peoples' moral relativism only about .2 standard deviations on average compared to people with no college degree, and this is the largest effect they identify.
@@
 
+We propose an approach that offers such a quantification.
 In this approach, there are two metrics that each reflect a combination of relevant theoretical processes.
 The first metric, interpersonal differences at the start of the survey, reflects the accumulated experiences of people prior to entering the panel survey that shape how they answer questions.
-Lersch calls this "early imprinting." While commonly associated with experiences during a formative period that result in settled dispositions, this variation could also reflect experiences that happen at any stage in life as long as they predate the panel and consistently affect subsequent responses.
+Lersch calls this "early imprinting." While commonly associated with experiences during a formative period that result in settled dispositions, this variation could also reflect experiences that happened at any stage in life as long as they predate the panel and consistently affect subsequent responses.
 For instance, for those people entering the panel post-retirement, this "imprinting" might reflect this pivotal life transition.
@@
 This non-durable change emerges for a variety of reasons.
-For example, people might not have clear a clear disposition on a particular item as it is asked.
-Instead, they might internalize a broad set of considerations and construct an opinion in the context of the survey interview, with different considerations coming to the forefront of their cognition each interview [@tourangeau2000; @zaller1992; @feldman1992].
+For example, people might not have a clear disposition on a particular item as it is asked.
+Instead, they might internalize a broad set of considerations and construct an opinion in the context of the survey interview, with different considerations coming to the forefront of their cognition during each interview [@tourangeau2000; @zaller1992; @feldman1992].
 Additionally, this variance can also include measurement error, such as misinterpreted questions, erroneous response selections, or responses getting coded incorrectly.
@@
 What is important for the theoretical debate is the relative contribution of these two components in explaining cultural differences in a population.
-In other words, when we look at a population of adults over some time frame, is there enough intrapersonal change relative to interpersonal difference to view it as an important explanation for why people differ from each other?
+In other words, when we look at a population of adults over some time frame, is there enough intrapersonal change relative to interpersonal difference to view it as an important explanation for why people differ from each other as observed?
 Only through quantifying the two processes we can get closer to a true answer to this question.
@@
 
-We use data from 7 nationally representative panel surveys from Australia, Germany, Great Britain, Switzerland, and the United States (summarized in Table 1), combining all the data files used in previous work [@kiley2020; @lersch2023].[^1] These studies cover a long period of time (the range of surveys spans from 1968 to 2021), with 609 personal culture items that capture attitudes, beliefs, values, self-assessments, self-descriptions, and behaviors [@alwin2007].
+We use data from seven nationally representative panel surveys from Australia, Germany, Great Britain, Switzerland, and the United States (summarized in Table 1), combining all the data files used in previous work [@kiley2020; @lersch2023].[^1] These studies cover a long period of time (the range of surveys spans from 1968 to 2021), with 609 personal culture items that capture attitudes, beliefs, values, self-assessments, self-descriptions, and behaviors [@alwin2007].
 We restricted the sample such that individuals between the ages of 18 and 79 are included without further elimination, and in all surveys, we used all possible cases for which respondents provided responses.
@@
 
-We can think of this number, $\omega$ as akin to an Intraclass Correlation Coefficient focused only on the systematic components of the model while ignoring residual variance that might be attributable to measurement error or non-persisting change [see also the variance decomposition of period and cohort by @vaisey2016].
-The higher this number is, the more systematic variance in responses is attributable to intrapersonal change during the panel.
-The lower the number, the less intrapersonal change seems to play a large role in explaining systematic cultural differences during the panel.
+We can think of this proportion $\omega$ as akin to an Intraclass Correlation Coefficient focused only on the systematic components of the model while ignoring residual variance that might be attributable to measurement error or non-persisting change [see also the variance decomposition of period and cohort by @vaisey2016].
+The higher this proportion is, the more systematic variance in responses is attributable to intrapersonal change during the panel.
+The lower this proportion, the less intrapersonal change seems to play a large role in explaining systematic cultural differences during the panel.
 Moving forward, we use $\omega$ as a placeholder for *systematic variance attributable to intrapersonal change*.
@@
 This allows us to explore how these features affect the prevalence of intrapersonal change.
-These covariates include the number of response options (measured as 2, 3-5, and more than 5), the panel, the log of the number of participants, the number of waves observed, the time since the question was first asked, and the total span of time the question was asked.
+These covariates include the number of response options (measured as 2, 3-5, and more than 5), the panel, the log of the number of participants, the number of waves observed, the time since the question was first asked, and the total duration of time the question was asked.
 
@@
 To broadly summarize the results, all panels show a similar range of the proportion of systematic variance attributable to intrapersonal change.
-Across panels, the median proportion of $\omega$ is 0.094 (mean of 0.098), with an interquartile range of 0.053 to 0.139.
+Across panels, the median value of $\omega$ is 0.094 (mean of 0.098), with an interquartile range of 0.053 to 0.139.
 All panels include questions where $\omega$ is essentially 0, and all panels include questions with $\omega$ values greater than 0.20.
@@
 
-In contrast to the tournament of models approach, quantifying change this way allows us to explore variation in the relative importance of intrapersonal change across questions that all both show evidence of change.
+In contrast to the tournament of models approach, quantifying change this way allows us to explore variation in the relative importance of intrapersonal change across questions that all show evidence of change.
 For example, Kiley and Vaisey [-@kiley2020] found that confidence in the press and confidence in religious leaders were both characterized by active updating.
@@
 We interpret these coefficients as suggesting greater resolution on a question makes it easier to detect and model change.
-The earlier a question was asked, even net of how long it was been asked, the lower the $\omega$.
+The earlier a question was asked, even net of how long it was been asked, the lower $\omega$.
 
 While we did not generate expectations for how question content relates to $\omega$, some of the associations between question structure, panel duration, and panels themselves might be driven by differences in the topics addressed by each panel.
-To address this, we coded each question as falling into one of nine different content categories and included these as indicator variables in the meta analysis.
+To address this, we followed Hout et al.
+[-@hout2016] and Lersch [-@lersch2023] in coding each question as falling into one of nine different content categories and included these as indicator variables in the meta analysis.
 Supplemental Materials B presents the estimates for these content indicators.
@@
 
-On all eight questions presented in Figure 4, $\omega$ values are greater for people without a college degree than for people with a college degree, meaning intrapersonal change explains less systematic variance for people with a college degree than for people without one.
+On all eight questions presented in Figure 4, $\omega$ values are smaller for people with a college degree than for people without a college degree, meaning intrapersonal change explains less systematic variance for people with a college degree than for people without one.
 There are also large differences in $\omega$ values across questions for both groups.
@@
 We believe these patterns shed new light on the mechanisms underlying differences in attitude behavior across groups.
-Something about college attendance limits the ability of experiences later in life to drive changes in personal culture.
+Something about college attendance seems to limit the ability of experiences later in life to drive changes in personal culture.
 While these results should not be interpreted as causal effects of attending college -- they are potentially confounded by age, social class, race, gender, and other factors that explain selection into higher education -- they open up a set of new questions and dynamics to explore.
@@
 More than adjudicating past positions in this debate, the approach presented in this paper enables scholars to more clearly explore the relative contributions of change and stability in explaining variation in personal culture, as well as how these contributions might differ across populations.
-To exemplify this, we detailed how amount of cultural difference explained by intrapersonal change varies substantially by survey item but also by individuals' characteristics such as their education.
+To exemplify this, we detailed how the amount of cultural difference explained by intrapersonal change varies substantially by survey item but also by individuals' characteristics such as their education.
 Intrapersonal change plays almost no part in explaining sentiments regarding the government's economic role, but it is more defining for ideological and partisan identities.
@@
 This conclusion aligns with a range of recent causal inference work suggesting that selection effects, rather than treatment effects, predominantly account for personal cultural differences among individuals in varied social roles and positions [@campbell2016; @wodtke2018].
-While some people clearly change as they transition into new roles or environments, this change often seems to be insufficient in magnitude and duration to explain what are often pronounced differences among people in diverse roles.
+While some people clearly change as they transition into new roles or environments, this change seems to be insufficient in magnitude and duration to explain what are often pronounced differences among people in diverse roles.
 These and our results suggest that when observing differences in personal culture across social roles, such as parenthood, education, or professional authority [@longest2013], or across occupations [@weeden2005], selection likely plays a large role in explaining these difference, though exceptions always exist.
@@
 
-Further, we align with others [**WE NEED SOME CITES HERE**] advocating for more studies --including panel studies-- focusing on the socialization of youth.
+Further, we align with others [**WE NEED SOME CITES HERE**] advocating for more studies, including panel studies, focusing on the socialization of youth.
 Our analysis points to the majority of variance among adults rooting in differences encountered before turning 18.
@@
 
-Finally, a key insight from our meta-analysis is that an increase in response resolution --the range of response options survey respondents are given-- correlates with an increased share of systematic variance explained by intrapersonal change.
+Finally, a key insight from our meta-analysis is that an increase in response resolution (the range of response options survey respondents are given) correlates with an increased share of systematic variance explained by intrapersonal change.
 This may reflect specific issues being assessed by survey items using different scales.
@@
 This is a sensible assumption on short panels where classifying change that lasts less than two years as durable seems unreasonable.
-Practically, it means that the variance produced by people making durable changes in their cultural dispositions to then returning to a previous state later in life is classified as residual variance, rather than intrapersonal change.
+Practically, it means that the variance produced by people making durable changes in their cultural dispositions to then return to a previous state later in life is classified as residual variance, rather than intrapersonal change.
 Longer panels and more flexible definitions of "change" would allow us to account for such trajectories and thus potentially also identify more durable change even in the same population.
@@
 
-While we focused on education as a determinant of the explanatory power of intrapersonal change, our approach can readily be extended to other socio-demographic categories that might affect intrapersonal change --race/ethnicity, religiosity, social class, gender, and political affiliation, for example-- or their intersections.
+While we focused on education as a determinant of the explanatory power of intrapersonal change, our approach can readily be extended to other socio-demographic categories that might affect intrapersonal change -- race/ethnicity, religiosity, social class, gender, and political affiliation, for example -- or their intersections.
 Researchers with different data might be able to evaluate the relative importance of intrapersonal change and interpersonal differences across contexts as well, either national contexts if the same questions are administered across countries, or organizational contexts if data traces students within schools, workers within workplaces, or congregation members within congregations.

### 2023-09-28 supplementals/manuscript_supplementals.Rmd
changes in supplements.

@@
   \usepackage{lscape}
+  \usepackage{siunitx}
   \usepackage{caption}
@@
 
-Table B1 presents the full set of coefficients from an OLS model regressing $\omega$ on multiple covariates, described in Figure 2 in the main manuscript.
+Table B1 presents the full set of coefficients from an OLS model regressing $\omega$ on multiple covariates, described in Figure 2. We multiplied the covariates by 100 to increase legibility.
 
 ```{=latex}
+\newcolumntype{d}[1]{D{.}{.}{#1}}
+\newcommand\mc[1]{\multicolumn{1}{c}{#1}} % handy shortcut macro
 \begin{table}[ht]
 \begin{center}
-\caption*{Table B1: Regression Model} 
+\caption*{Table B1: Regression Model Estimating $\omega$}
 \label{tab:sometab}
-\begin{tabular}{l c c}
+\begin{tabular}{l *{4}{d{1.2}} }
 \hline
- & Estimates & SE \\
+ & \mc{Estimates} & \mc{SE} \\
 \hline
-N Part. (log)                                       & $0.01$        
-                                                    & $(0.01)$      \\
-Duration                                            & $-0.00$ 
-                                                    & $(0.00)^{***}$ \\
-N Waves (Part.)                                     & $0.00$        
-                                                    & $(0.00)$      \\
-N Waves (Survey)                                    & $-0.00$       
-                                                    & $(0.00)$      \\
-Wave 1 (Decade)                                     & $-0.01$  
-                                                    & $(0.00)^{**}$  \\
-Panels: BHPS                                        & $0.08$        
-                                                    & $(0.05)$      \\
-Panels: GSS                                         & $0.04$        
-                                                    & $(0.05)$      \\
-Panels: HILDA                                       & $0.08$        
-                                                    & $(0.05)$      \\
-Panels: PSID                                        & $0.09$    
-                                                    & $(0.04)^{*}$  \\
-Panels: SHP                                         & $0.06$        
-                                                    & $(0.05)$      \\
-Panels: SOEP                                        & $0.06$        
-                                                    & $(0.05)$      \\
-Panels: UKHLS                                       & $0.09$        
-                                                    & $(0.06)$      \\
-Topics: Environment and Climate                     & $0.04$   
-                                                    & $(0.01)^{**}$      \\
-Topics: Gender and Family                           & $0.00$        
-                                                    & $(0.01)$      \\
-Topics: Health and Morale                           & $0.03$   
-                                                    & $(0.01)^{**}$      \\
-Topics: National Identity, Ethnicity, Immigration   & $0.02$        
-                                                    & $(0.01)$      \\
-Topics: Occupation and Education                    & $-0.02$       
-                                                    & $(0.02)$      \\
-Topics: Religion and Spirituality                   & $-0.00$       
-                                                    & $(0.01)$      \\
-Topics: Social Life, Social Cohesion and Trust      & $0.03$  
-                                                    & $(0.01)^{***}$      \\
-Topics: Subjective SES                              & $0.03$    
-                                                    & $(0.01)^{*}$      \\
+Number of Participants (Logged)                       & 0.99        
+                                                      & 0.53      \\
+Duration                                              & -0.49^{***} 
+                                                      & 0.13      \\
+Average Number of Waves Across Participants           & 0.09        
+                                                      & 0.20      \\
+Average Number of Waves per Item                      & -0.20       
+                                                      & 0.23      \\
+Wave 1 (Decade)                                       & -1.17^{**}  
+                                                      & 0.44      \\
+Panels: BHPS                                          & 0.08        
+                                                      & 0.05      \\
+Panels: GSS                                           & 0.04        
+                                                      & 0.05      \\
+Panels: HILDA                                         & 0.08        
+                                                      & 0.05      \\
+Panels: PSID                                          & 0.09^{*}    
+                                                      & 0.04      \\
+Panels: SHP                                           & 0.06        
+                                                      & 0.05      \\
+Panels: SOEP                                          & 0.06        
+                                                      & 0.05      \\
+Panels: UKHLS                                         & 0.09        
+                                                      & 0.06      \\
+Topics: Environment and Climate                       & 0.04^{**}   
+                                                      & 0.01      \\
+Topics: Gender and Family                             & 0.00        
+                                                      & 0.01      \\
+Topics: Health and Morale                             & 0.03^{**}   
+                                                      & 0.01      \\
+Topics: National Identity, Ethnicity and Immigration  & 0.02        
+                                                      & 0.01      \\
+Topics: Occupation and Education                      & -0.02       
+                                                      & 0.02      \\
+Topics: Religion and Spirituality                     & -0.00       
+                                                      & 0.01      \\
+Topics: Social Life, Social Cohesion and Trust        & 0.03^{***}  
+                                                      & 0.01      \\
+Topics: Subjective SES                                & 0.03^{*}    
+                                                      & 0.01      \\
 \hline
-Adjusted R$^2$                                      & $0.77$     &   \\
-N                                                   & $609$      &  \\
+Adjusted R-Squared                                    & 0.77 &    \\
 \hline
-\multicolumn{2}{l}{\scriptsize{$^{***}p<0.001$; $^{**}p<0.01$; $^{*}p<0.05$}}
+\multicolumn{3}{l}{\scriptsize{$^{***}p<0.001$; $^{**}p<0.01$; $^{*}p<0.05$}. N = 609.}
 \end{tabular}
@@
 \end{table}
+
 ```

### 2023-09-28 supplementals/manuscript_supplementals.tex
changes in supplements.

@@
 \setcounter{secnumdepth}{-\maxdimen} % remove section numbering
-\usepackage{fontspec} \usepackage{longtable} \usepackage{multirow} \usepackage{array} \usepackage{booktabs} \usepackage{dcolumn} \usepackage{lscape} \usepackage{caption} \newcommand{\blandscape}{\begin{landscape}} \newcommand{\elandscape}{\end{landscape}}
+\usepackage{fontspec} \usepackage{longtable} \usepackage{multirow} \usepackage{array} \usepackage{booktabs} \usepackage{dcolumn} \usepackage{lscape} \usepackage{siunitx} \usepackage{caption} \newcommand{\blandscape}{\begin{landscape}} \newcommand{\elandscape}{\end{landscape}}
 \ifLuaTeX
@@
 Table B1 presents the full set of coefficients from an OLS model
-regressing \(\omega\) on multiple covariates, described in Figure 2 in
-the main manuscript.
+regressing \(\omega\) on multiple covariates, described in Figure 2. We
+multiplied the covariates by 100 to increase legibility.
 
+\newcolumntype{d}[1]{D{.}{.}{#1}}
+\newcommand\mc[1]{\multicolumn{1}{c}{#1}} % handy shortcut macro
 \begin{table}[ht]
 \begin{center}
-\caption*{Table B1: Regression Model} 
+\caption*{Table B1: Regression Model Estimating $\omega$}
 \label{tab:sometab}
-\begin{tabular}{l c c}
+\begin{tabular}{l *{4}{d{1.2}} }
 \hline
- & Estimates & SE \\
+ & \mc{Estimates} & \mc{SE} \\
 \hline
-N Part. (log)                                       & $0.01$        
-                                                    & $(0.01)$      \\
-Duration                                            & $-0.00$ 
-                                                    & $(0.00)^{***}$ \\
-N Waves (Part.)                                     & $0.00$        
-                                                    & $(0.00)$      \\
-N Waves (Survey)                                    & $-0.00$       
-                                                    & $(0.00)$      \\
-Wave 1 (Decade)                                     & $-0.01$  
-                                                    & $(0.00)^{**}$  \\
-Panels: BHPS                                        & $0.08$        
-                                                    & $(0.05)$      \\
-Panels: GSS                                         & $0.04$        
-                                                    & $(0.05)$      \\
-Panels: HILDA                                       & $0.08$        
-                                                    & $(0.05)$      \\
-Panels: PSID                                        & $0.09$    
-                                                    & $(0.04)^{*}$  \\
-Panels: SHP                                         & $0.06$        
-                                                    & $(0.05)$      \\
-Panels: SOEP                                        & $0.06$        
-                                                    & $(0.05)$      \\
-Panels: UKHLS                                       & $0.09$        
-                                                    & $(0.06)$      \\
-Topics: Environment and Climate                     & $0.04$   
-                                                    & $(0.01)^{**}$      \\
-Topics: Gender and Family                           & $0.00$        
-                                                    & $(0.01)$      \\
-Topics: Health and Morale                           & $0.03$   
-                                                    & $(0.01)^{**}$      \\
-Topics: National Identity, Ethnicity, Immigration   & $0.02$        
-                                                    & $(0.01)$      \\
-Topics: Occupation and Education                    & $-0.02$       
-                                                    & $(0.02)$      \\
-Topics: Religion and Spirituality                   & $-0.00$       
-                                                    & $(0.01)$      \\
-Topics: Social Life, Social Cohesion and Trust      & $0.03$  
-                                                    & $(0.01)^{***}$      \\
-Topics: Subjective SES                              & $0.03$    
-                                                    & $(0.01)^{*}$      \\
+Number of Participants (Logged)                       & 0.99        
+                                                      & 0.53      \\
+Duration                                              & -0.49^{***} 
+                                                      & 0.13      \\
+Average Number of Waves Across Participants           & 0.09        
+                                                      & 0.20      \\
+Average Number of Waves per Item                      & -0.20       
+                                                      & 0.23      \\
+Wave 1 (Decade)                                       & -1.17^{**}  
+                                                      & 0.44      \\
+Panels: BHPS                                          & 0.08        
+                                                      & 0.05      \\
+Panels: GSS                                           & 0.04        
+                                                      & 0.05      \\
+Panels: HILDA                                         & 0.08        
+                                                      & 0.05      \\
+Panels: PSID                                          & 0.09^{*}    
+                                                      & 0.04      \\
+Panels: SHP                                           & 0.06        
+                                                      & 0.05      \\
+Panels: SOEP                                          & 0.06        
+                                                      & 0.05      \\
+Panels: UKHLS                                         & 0.09        
+                                                      & 0.06      \\
+Topics: Environment and Climate                       & 0.04^{**}   
+                                                      & 0.01      \\
+Topics: Gender and Family                             & 0.00        
+                                                      & 0.01      \\
+Topics: Health and Morale                             & 0.03^{**}   
+                                                      & 0.01      \\
+Topics: National Identity, Ethnicity and Immigration  & 0.02        
+                                                      & 0.01      \\
+Topics: Occupation and Education                      & -0.02       
+                                                      & 0.02      \\
+Topics: Religion and Spirituality                     & -0.00       
+                                                      & 0.01      \\
+Topics: Social Life, Social Cohesion and Trust        & 0.03^{***}  
+                                                      & 0.01      \\
+Topics: Subjective SES                                & 0.03^{*}    
+                                                      & 0.01      \\
 \hline
-Adjusted R$^2$                                      & $0.77$     &   \\
-N                                                   & $609$      &  \\
+Adjusted R-Squared                                    & 0.77 &    \\
 \hline
-\multicolumn{2}{l}{\scriptsize{$^{***}p<0.001$; $^{**}p<0.01$; $^{*}p<0.05$}}
+\multicolumn{3}{l}{\scriptsize{$^{***}p<0.001$; $^{**}p<0.01$; $^{*}p<0.05$}. N = 609.}
 \end{tabular}

### 2023-09-28 manuscript/manuscript.Rmd
reknitted the manuscript.

@@
 ```
-```{r fig2, out.width = "450px"}
+```{r fig2, out.width = "350px"}
 knitr::include_graphics("../figures/figure_2.png")

### 2023-09-28 manuscript/manuscript.tex
reknitted the manuscript.

@@
 interpersonal differences. Our results harmonize recent findings by
-showing that, while most measures of personal culture show change over
-time, the variance explained by intrapersonal change is typically small
-compared to interpersonal differences at baseline. Quantifying these
-processes offers new ways to address theoretical questions of when and
-how their relative importance shifts. We examplify this approach by
-exploring how the importance of these processes differs between college
-graduates and the rest of the population.
+showing that, while most measures of personal culture show durable
+change over time, the variance explained by intrapersonal change is
+typically small compared to that of interpersonal differences at
+baseline. Quantifying these processes offers new ways to address
+theoretical questions of when and how their relative importance shifts.
+We exemplify this approach by exploring how the importance of these
+processes differs between college graduates and the rest of the
+population.
 \end{abstract}
@@
 These seemingly inconsistent findings are partly due to the fact that
-researchers have taken a ``tournament of models'' approach to
-adjudicating different theoretical perspectives (Lersch 2023: p.~228).
-That is, researchers have compared the fit of different statistical
-models to each item to determine whether it is more consistent with ``no
-change'' or ``any change.'' The size of these piles then serves as the
-primary evidence for the truth of the theory (Kiley and Vaisey 2020;
-Lersch 2023; Vaisey and Kiley 2021).
+researchers have taken a ``tournament of models'' approach to adjudicate
+different theoretical perspectives (Lersch 2023: p.~228). That is,
+researchers have compared the fit of different statistical models to
+each item to determine whether it is more consistent with ``no change''
+or ``any change.'' The size of these piles then serves as the primary
+evidence for the truth of the theory (Kiley and Vaisey 2020; Lersch
+2023; Vaisey and Kiley 2021).
 
@@
 as ``change.'' These differences can limit the possibility of agreement
-on which ``pile'' a variable belongs in. Moreover, asking whether there
-is evidence of change reduces to, ``Is there \emph{any} evidence of
-change in this item?'' But such binary inquiries are ill-suited to
-address the core of the theoretical debate: determining \emph{how much}
-the process of intrapersonal change contributes to cultural differences.
+on which ``pile'' a variable belongs. Moreover, asking whether there is
+evidence of change reduces to, ``Is there \emph{any} evidence of change
+in this item?'' But such binary inquiries are ill-suited to address the
+core of the theoretical debate: determining \emph{how much} the process
+of intrapersonal change contributes to cultural differences.
 
@@
 intrapersonal change and pre-existing interpersonal differences. These
-questions are about \emph{degree}, not existence; taking this debate
-forward will require precise measurement rather than declarations of
-victory for a particular perspective.
+questions are about \emph{degree}, not existence; to take this debate
+forward therefore requires precise measurement rather than declarations
+of victory for a particular perspective.
 
@@
 
-Taken together, to goal of this paper is to advance---and, hopefully,
+Taken together, the goal of this paper is to advance---and, hopefully,
 transcend---debates about \emph{whether} adults change. People do
 change, at least a little, on most things. Our method can help quantify
-exactly how much.
+exactly \emph{how much}.
 
@@
 was evidence of durable change among adults, a pattern consistent with
-Lersch's results. While Lersch (2023) found evidence of durable change
-on most items, this change only accounted for a small proportion of
+Lersch's results. Also Lersch (2023) found evidence of durable change on
+most items, but this change only accounted for a small proportion of
 variance explained. On average, people changed only about .07 standard
 deviations over 10 years. This is consistent with Kiley and Vaisey's
-finding that durable change is often small. And a quarter of the items
-Lersch studied still preferred the SDM, suggesting that even with
+finding that durable change is often small. Moreover, a quarter of the
+items Lersch studied still preferred the SDM, suggesting that even with
 different assumptions assumptions about change and more extensive data,
-many questions show no evidence of durable change. Likewise, in studies
-showing evidence of durable change in specific measures of personal
-culture, the change is often small. For example, Brocic and Miles (2021)
-estimate that completing graduate degrees in humanities, arts, and
-social sciences shift peoples' moral relativism only about .2 standard
-deviations on average compared to people with no college degree, and
-this is the largest effect they identify. And studies of aggregate
-change show that, even on items where cohort effects explain more
-variance than period effects, there is always evidence that some people
-change over time (Vaisey and Lizardo 2016). In other words, despite
-different interpretations, the results of previous work are in many ways
-highly consonant.
+many questions show no evidence of durable change. Consistent with this,
+in studies showing evidence of durable change in specific measures of
+personal culture, the change is often small. For example, Brocic and
+Miles (2021) estimate that completing graduate degrees in humanities,
+arts, and social sciences shift peoples' moral relativism only about .2
+standard deviations on average compared to people with no college
+degree, and this is the largest effect they identify. And studies of
+aggregate change show that, even on items where cohort effects explain
+more variance than period effects, there is always evidence that some
+people change over time (Vaisey and Lizardo 2016). In other words,
+despite different interpretations, the results of previous work are in
+many ways highly consonant.
 
@@
 
-In this approach, there are two metrics that each reflect a combination
-of relevant theoretical processes. The first metric, interpersonal
+We propose an approach that offers such a quantification. In this
+approach, there are two metrics that each reflect a combination of
+relevant theoretical processes. The first metric, interpersonal
 differences at the start of the survey, reflects the accumulated
@@
 result in settled dispositions, this variation could also reflect
-experiences that happen at any stage in life as long as they predate the
-panel and consistently affect subsequent responses. For instance, for
-those people entering the panel post-retirement, this ``imprinting''
+experiences that happened at any stage in life as long as they predate
+the panel and consistently affect subsequent responses. For instance,
+for those people entering the panel post-retirement, this ``imprinting''
 might reflect this pivotal life transition. Consequently, this metric
@@
 non-durable change emerges for a variety of reasons. For example, people
-might not have clear a clear disposition on a particular item as it is
-asked. Instead, they might internalize a broad set of considerations and
+might not have a clear disposition on a particular item as it is asked.
+Instead, they might internalize a broad set of considerations and
 construct an opinion in the context of the survey interview, with
-different considerations coming to the forefront of their cognition each
-interview (Feldman and Zaller 1992; Tourangeau, Rips, and Rasinski 2000;
-Zaller 1992). Additionally, this variance can also include measurement
-error, such as misinterpreted questions, erroneous response selections,
-or responses getting coded incorrectly. While this third metric at least
-partially reflects important processes of personal culture, it does not
-directly touch on the ongoing debate outlined here. Because of this we
-quantify this component but focus principally on the other two.
+different considerations coming to the forefront of their cognition
+during each interview (Feldman and Zaller 1992; Tourangeau, Rips, and
+Rasinski 2000; Zaller 1992). Additionally, this variance can also
+include measurement error, such as misinterpreted questions, erroneous
+response selections, or responses getting coded incorrectly. While this
+third metric at least partially reflects important processes of personal
+culture, it does not directly touch on the ongoing debate outlined here.
+Because of this we quantify this component but focus principally on the
+other two.
 
@@
 change relative to interpersonal difference to view it as an important
-explanation for why people differ from each other? Only through
-quantifying the two processes we can get closer to a true answer to this
-question.
+explanation for why people differ from each other as observed? Only
+through quantifying the two processes we can get closer to a true answer
+to this question.
 
@@
 
-We use data from 7 nationally representative panel surveys from
+We use data from seven nationally representative panel surveys from
 Australia, Germany, Great Britain, Switzerland, and the United States
@@
 
-We can think of this number, \(\omega\) as akin to an Intraclass
+We can think of this proportion \(\omega\) as akin to an Intraclass
 Correlation Coefficient focused only on the systematic components of the
@@
 decomposition of period and cohort by Vaisey and Lizardo 2016). The
-higher this number is, the more systematic variance in responses is
-attributable to intrapersonal change during the panel. The lower the
-number, the less intrapersonal change seems to play a large role in
+higher this proportion is, the more systematic variance in responses is
+attributable to intrapersonal change during the panel. The lower this
+proportion, the less intrapersonal change seems to play a large role in
 explaining systematic cultural differences during the panel. Moving
@@
 the log of the number of participants, the number of waves observed, the
-time since the question was first asked, and the total span of time the
-question was asked.
+time since the question was first asked, and the total duration of time
+the question was asked.
 
@@
 systematic variance attributable to intrapersonal change. Across panels,
-the median proportion of \(\omega\) is 0.094 (mean of 0.098), with an
+the median value of \(\omega\) is 0.094 (mean of 0.098), with an
 interquartile range of 0.053 to 0.139. All panels include questions
@@
 this way allows us to explore variation in the relative importance of
-intrapersonal change across questions that all both show evidence of
-change. For example, Kiley and Vaisey (2020) found that confidence in
-the press and confidence in religious leaders were both characterized by
-active updating. Our results show that intrapersonal change is much more
+intrapersonal change across questions that all show evidence of change.
+For example, Kiley and Vaisey (2020) found that confidence in the press
+and confidence in religious leaders were both characterized by active
+updating. Our results show that intrapersonal change is much more
 important for explaining variance in confidence in the press (0.164)
@@
 
-\includegraphics[width=450px]{../figures/figure_2}
+\includegraphics[width=350px]{../figures/figure_2}
 
@@
 it easier to detect and model change. The earlier a question was asked,
-even net of how long it was been asked, the lower the \(\omega\).
+even net of how long it was been asked, the lower \(\omega\).
 
@@
 panel duration, and panels themselves might be driven by differences in
-the topics addressed by each panel. To address this, we coded each
-question as falling into one of nine different content categories and
-included these as indicator variables in the meta analysis. Supplemental
-Materials B presents the estimates for these content indicators. They
-show some associations with \(\omega\), with questions about subjective
-SES; social life, social cohesion, and trust; environment and climate;
-and health and morale showing larger \(\omega\) values than questions
-about religion and spirituality; politics, government and the economy;
-and gender and family.
+the topics addressed by each panel. To address this, we followed Hout et
+al. (2016) and Lersch (2023) in coding each question as falling into one
+of nine different content categories and included these as indicator
+variables in the meta analysis. Supplemental Materials B presents the
+estimates for these content indicators. They show some associations with
+\(\omega\), with questions about subjective SES; social life, social
+cohesion, and trust; environment and climate; and health and morale
+showing larger \(\omega\) values than questions about religion and
+spirituality; politics, government and the economy; and gender and
+family.
 
@@
 On all eight questions presented in Figure 4, \(\omega\) values are
-greater for people without a college degree than for people with a
+smaller for people with a college degree than for people without a
 college degree, meaning intrapersonal change explains less systematic
@@
 differences in attitude behavior across groups. Something about college
-attendance limits the ability of experiences later in life to drive
-changes in personal culture. While these results should not be
+attendance seems to limit the ability of experiences later in life to
+drive changes in personal culture. While these results should not be
 interpreted as causal effects of attending college -- they are
@@
 in personal culture, as well as how these contributions might differ
-across populations. To exemplify this, we detailed how amount of
+across populations. To exemplify this, we detailed how the amount of
 cultural difference explained by intrapersonal change varies
@@
 2016; Wodtke 2018). While some people clearly change as they transition
-into new roles or environments, this change often seems to be
-insufficient in magnitude and duration to explain what are often
-pronounced differences among people in diverse roles. These and our
-results suggest that when observing differences in personal culture
-across social roles, such as parenthood, education, or professional
-authority (Longest, Hitlin, and Vaisey 2013), or across occupations
-(Weeden and Grusky 2005), selection likely plays a large role in
-explaining these difference, though exceptions always exist.
+into new roles or environments, this change seems to be insufficient in
+magnitude and duration to explain what are often pronounced differences
+among people in diverse roles. These and our results suggest that when
+observing differences in personal culture across social roles, such as
+parenthood, education, or professional authority (Longest, Hitlin, and
+Vaisey 2013), or across occupations (Weeden and Grusky 2005), selection
+likely plays a large role in explaining these difference, though
+exceptions always exist.
 
@@
 Further, we align with others {[}\textbf{WE NEED SOME CITES HERE}{]}
-advocating for more studies --including panel studies-- focusing on the
+advocating for more studies, including panel studies, focusing on the
 socialization of youth. Our analysis points to the majority of variance
@@
 Finally, a key insight from our meta-analysis is that an increase in
-response resolution --the range of response options survey respondents
-are given-- correlates with an increased share of systematic variance
+response resolution (the range of response options survey respondents
+are given) correlates with an increased share of systematic variance
 explained by intrapersonal change. This may reflect specific issues
@@
 Practically, it means that the variance produced by people making
-durable changes in their cultural dispositions to then returning to a
+durable changes in their cultural dispositions to then return to a
 previous state later in life is classified as residual variance, rather
@@
 of intrapersonal change, our approach can readily be extended to other
-socio-demographic categories that might affect intrapersonal change
---race/ethnicity, religiosity, social class, gender, and political
-affiliation, for example-- or their intersections. Researchers with
+socio-demographic categories that might affect intrapersonal change --
+race/ethnicity, religiosity, social class, gender, and political
+affiliation, for example -- or their intersections. Researchers with
 different data might be able to evaluate the relative importance of
@@
 
+\leavevmode\vadjust pre{\hypertarget{ref-hout2016}{}}%
+Hout, Michael, and Orestes P. Hastings. 2016. {``Reliability of the
+{Core} {Items} in the {General} {Social} {Survey}: {Estimates} from the
+{Three}-{Wave} {Panels}, 2006--2014.''} \emph{Sociological Science}
+3:971--1002. doi:
+\href{https://doi.org/10.15195/v3.a43}{10.15195/v3.a43}.
+
 \leavevmode\vadjust pre{\hypertarget{ref-data_psid}{}}%

### 2023-09-29 manuscript/manuscript.Rmd
small update in the manuscript.
SV edits

@@
 500; see also @vaisey2021].
-This lack of durable change is consistent with other recent findings that cohort replacement plays a larger role than period effects in explaining differences in personal culture [@vaisey2016].
+This lack of durable change is consistent with other recent findings that cohort replacement plays a somewhat larger role than period effects in explaining differences in personal culture [@vaisey2016].
 
@@
 A more theoretically productive approach begins with a model assuming that during an observed time period people might remain stable, might change a little, or might undergo significant shifts in their personal culture.
-Lersch's LCAM does this by modeling each individual as following their own linear trajectory.
+Lersch's LCAM does this by modeling each individual as following a personal linear trajectory.
 But rather than debating whether this model fits a particular data set better than a model that assumes that people never change, researchers should quantify these individual changes and compare those to the differences that exist between people at the panel's start.
-In other words, the focus should shift from asking, "do people change?" to asking, "what are the relative contributions of change and initial differences for explaining differences in personal culture?"
+In other words, the focus should shift from asking, "do people change?" to asking, "what are the relative contributions of change and initial differences for explaining variance in personal culture?"
 
 We propose an approach that offers such a quantification.
-In this approach, there are two metrics that each reflect a combination of relevant theoretical processes.
+Our measure relates two metrics that each reflect a combination of relevant theoretical processes.
 The first metric, interpersonal differences at the start of the survey, reflects the accumulated experiences of people prior to entering the panel survey that shape how they answer questions.
@@
 
-The second metric, the amount of intrapersonal linear change people make during the panel, captures durable changes in personal culture over time, a set of processes collectively called "persistent change" or "adaption" by Lersch and "active updating" by Kiley and Vaisey.
-Lersch attributes these changes to social triggers such as moving into a new environment or adopting new social roles (although he never measures them directly).
+The second metric, the amount of intrapersonal linear change people make during the panel, captures durable changes in personal culture over time.
+This captures the set of processes collectively called "persistent change" or "adaption" by Lersch and "active updating" by Kiley and Vaisey.
+Lersch attributes these changes to social triggers such as moving into a new environment or adopting new social roles (although he never measures these directly).
 Additionally, they might reflect the diffusion of new cultural forms across social networks, cues from political elites or otherwise culturally influential leaders, the emergence of issues in politics or culture, or large-scale social shifts.
 
-A third metric, often called "residual variance" or "fluctuation," accounts for the remainder of variance in peoples' responses.
-This non-durable change emerges for a variety of reasons.
+A third metric, which we might call "residual variance" or "fluctuation," accounts for the remainder of variance in peoples' responses.
+These non-durable changes emerge for a variety of reasons.
 For example, people might not have a clear disposition on a particular item as it is asked.
 Instead, they might internalize a broad set of considerations and construct an opinion in the context of the survey interview, with different considerations coming to the forefront of their cognition during each interview [@tourangeau2000; @zaller1992; @feldman1992].
-Additionally, this variance can also include measurement error, such as misinterpreted questions, erroneous response selections, or responses getting coded incorrectly.
-While this third metric at least partially reflects important processes of personal culture, it does not directly touch on the ongoing debate outlined here.
-Because of this we quantify this component but focus principally on the other two.
+This variance can also include measurement error, such as misinterpreted questions, erroneous response selections, or responses getting coded incorrectly.
+Although this third metric at least partially reflects important processes of personal culture, it does not directly touch on the ongoing debate outlined here.
+Because of this, we quantify this component but focus principally on the other two.
 
@@
 By shifting the question and adjusting the approach, we expect consistent patterns across datasets previously analysed in the debate.
-First, because previous work found evidence of durable change on most measures of personal culture, we expect that intrapersonal change will have some explanatory power for most items.
-But because these earlier studies also found many questions (between 25 and 35 percent) lacking evidence of durable change, we expect considerable variation in the variance explained by intrapersonal change, with some questions indicating essentially no variance attributable to such change.
+First, because previous work found evidence of durable change on most measures of personal culture, we expect that we will find some intrapersonal change in most items.
+But because these earlier studies also found many questions (between 25 and 35 percent) lacking evidence of *any* durable change, we expect considerable variation in the variance accounted for by intrapersonal change, with some questions indicating essentially no variance attributable to such change.
 
-Second, we expect the variance explained by intrapersonal change to be modest compared to interpersonal differences at the start of the survey.
+Second, we expect the variance attributable to intrapersonal change to be modest compared to interpersonal differences at the start of the survey.
 This expectation should be uncontroversial, regardless of the theoretical process assumed to be most relevant for a particular question.
@@
 However, theoretical perspectives differ in how much variance they would expect to be driven by intrapersonal change, enabling quantification to provide a new perspective on existing questions.
-"Formative period" and cohort theories, which emphasize early-life socialization as a strong force in shaping adult personal culture, suggest that the variance from intrapersonal change would still fall short of explaining interpersonal differences, even when we account for the duration of observation.
-Conversely, other theoretical perspectives emphasizing the importance of people's contemporaneous social environment for personal culture expect intrapersonal change to account for a large portion of the variance, regardless of the duration.
+"Formative period" and cohort theories, which emphasize early-life socialization as a strong influence on adult personal culture, suggest that the variance from intrapersonal change would still fall short of explaining interpersonal differences, even when we account for the duration of observation.
+Conversely, other theoretical perspectives emphasizing the importance of people's contemporaneous social environment for personal culture might expect intrapersonal change to account for a large portion of the variance, regardless of the duration.
 
@@
 We restricted the sample such that individuals between the ages of 18 and 79 are included without further elimination, and in all surveys, we used all possible cases for which respondents provided responses.
-In the end, the analyses that follow rely on a cross-cultural sample with a cross-domain set of items to capture a broad range of individual personal culture.
+In the end, the analyses that follow rely on a cross-national sample with a cross-domain set of items to capture a broad range of individual personal culture.
 Supplemental Materials A documents the list of all variables used in the upcoming analyses.
@@
 Third, it assumes that people deviate around this baseline randomly over time, reflecting "fluctuation" or short-term non-persistent change.
-Finally, it assumes that people draw on earlier time points in reacting to new events.
 Formally, this can be written as
@@
 $$
-y_{i, t} = \beta_0 + u_{0, i} + \beta_{1} x_t + u_{1, i} x_{i, t} + \epsilon_{i, t}
-$$ $$
-u_{0,i} \sim \mathcal{N}(0,\sigma_{\mu_0}^2) \\
-$$ $$
-u_{1,i} \sim \mathcal{N}(0,\sigma_{u_1}^2)
+y_{it} = \beta_0 + \alpha_{0i} + (\beta_1 + \alpha_{1i}) \text{year}_{it} + \epsilon_{it}
 $$
 
-where $\beta_0$ is the average intercept, $u_{0, i}$ is the random intercept for individuals, $\beta_1 x_t$ is the average slope, $u_{1, i} x_{i, t}$ is the random slope for individuals, and $\epsilon_{i, t}$ is the random error term that captures transitory disturbances.
+$$
+\begin{aligned}
+\alpha_{0i} &\sim \mathcal{N}(0,\tau^2_{0}) \\
+\alpha_{1i} &\sim \mathcal{N}(0,\tau^2_{1}) \\
+\epsilon_{it} &\sim \mathcal{N}(0,\sigma^2)
+\end{aligned}
+$$
+
+where $\beta_0$ is the average intercept, $\alpha_{0i}$ is the random intercept component for individuals, $\beta_1$ is the average yearly change in the outcome, $\alpha_{1i}$ is the random slope component for individuals, and $\epsilon_{it}$ is the random error term that captures transitory disturbances.
+The $\alpha_{0i}$ and $\alpha_{1i}$ terms are also allowed to covary.
 
@@
 
-As a first step in our analysis, we fit a LCAM to each of the 609 measures of personal culture outlined above.
+As a first step in our analysis, we fit this LCAM to each of the 609 measures of personal culture outlined above.
 We then use these models to derive two measures of systematic variance that correspond to the two collections of theoretical processes outlined above: interpersonal differences at baseline and intrapersonal change over time.
-First, to measure the importance of interpersonal differences, we calculate the sum of the squared residuals derived from measuring each respondent at the midpoint of their age trajectory.[^2] We then divide this value by the total sum of squares and subtract it from from 1, producing a value we denote as V(D), or the proportion of variance attributable to interpersonal differences.
-This number is equivalent to an $R^2$ value generated by using random intercepts for all respondents.
-Second, we calculate the sum of squared residuals generated by additionally accounting for the age trajectory.
+
+We estimate these using a form of variance decomposition.
+First, to measure stable interpersonal differences, we calculate the sum of the squared residuals derived from assuming that each person spends the entire survey at the estimated midpoint of their LCAM trajectory.[^2] We then divide this value by the total sum of squares and subtract this number from from 1, producing a value we denote as $V(D)$.
+This is the proportion of total variance attributable to interpersonal differences.
+(This number is equivalent to the conditional $R^2$ value generated by a model with only random intercepts for all respondents.) Second, we calculate another sum of squared residuals generated by allowing the predictions to vary by estimated LCAM trajectory.
 We divide this number by the total sum of squares and subtract it from 1.
 We then subtract $V(D)$ from this value.
-This gives us a value we denote as $V(C)$, the additional proportion of variance explained by intrapersonal change.
+This gives us a value we denote as $V(C)$, the additional proportion of variance accounted for by intrapersonal change.
 
-[^2]: We use the scale midpoint, rather than the actual difference at baseline, to better capture the differences between people independent of their age trajectories.
+[^2]: We use the midpoint estimate, rather than the value at the first wave, because this is the best measure of "baseline" available under the assumption that the person does not change.
+The first wave measure alone contains an unknown amount of measurement error and transient fluctuations.
 
-These two numbers reflect the total systematic variance of the model that can be attributed to either differences across people or linear change over the course of the panel.[^3] Our principal outcome of interest is the proportion of systematic variance accounted for by intrapersonal change, which we calculate as
+These two numbers together reflect the total systematic variance of the LCAM that can be attributed to either differences between people or linear intrapersonal change over the course of the panel.[^3] Our principal outcome of interest is the proportion of systematic variance accounted for by intrapersonal change, which we calculate as
 
@@
 
-We can think of this proportion $\omega$ as akin to an Intraclass Correlation Coefficient focused only on the systematic components of the model while ignoring residual variance that might be attributable to measurement error or non-persisting change [see also the variance decomposition of period and cohort by @vaisey2016].
+We can think of $\omega$ as a type of Intraclass Correlation Coefficient focused only on the systematic components of the model while ignoring residual variance that might be attributable to measurement error or non-persisting change [see also the variance decomposition of period and cohort by @vaisey2016].
 The higher this proportion is, the more systematic variance in responses is attributable to intrapersonal change during the panel.
-The lower this proportion, the less intrapersonal change seems to play a large role in explaining systematic cultural differences during the panel.
-Moving forward, we use $\omega$ as a placeholder for *systematic variance attributable to intrapersonal change*.
+From now on, we use $\omega$ as a placeholder for *systematic variance attributable to intrapersonal change*.
 
 Because this quotient summarizes variance and therefore has no natural referent, it is hard to say at which values we observe "a lot" of intrapersonal change.
-But summarizing questions in this way allows us to compare the relative importance of our two broad processes across a range of questions that might have different levels of non-systematic variance, giving us a sense of the relative prevalence of these two processes for each question (something not achievable under the tournament of models approach), and allowing us to compare which kinds of questions show more or less intrapersonal change relative to interpersonal differences.
-While the tournament of models approach allowed researchers to say whether questions in general showed evidence of updating, this approach allows us to make more specific claims about the prevalence of these theoretical processes for particular questions relative to each other.
+But summarizing questions in this way allows us to compare the relative importance of our two broad processes across a range of questions that might have different levels of non-systematic variance.
+This gives us a sense of the relative prevalence of these two processes for each question, which is not achievable under the tournament of models approach.
+This also allows us to compare which kinds of questions show more or less intrapersonal change relative to interpersonal differences.
+While the tournament of models approach allowed researchers to say whether questions in general showed *any* evidence of updating, this approach allows us to make more specific claims about the prevalence of these theoretical processes for particular questions relative to each other.
 
-As a second step in our analysis, we use ordinary least squares regression to model $\omega$ as a function of features of the question and panel.
-This allows us to explore how these features affect the prevalence of intrapersonal change.
-These covariates include the number of response options (measured as 2, 3-5, and more than 5), the panel, the log of the number of participants, the number of waves observed, the time since the question was first asked, and the total duration of time the question was asked.
+As a second step in our analysis, we use linear regression to conduct a meta-analysis, modeling $\omega$ as a function of features of the question and panel.
+This allows us to explore how these features are associated with the degree of intrapersonal change.
+These covariates include the number of response options (measured as 2, 3-5, and more than 5), the specific survey, the log of the number of participants, the number of waves observed, the time period (i.e., decade) the question was first asked, and the total duration of time the question was asked.
 
@@
 
-Figure 1 plots the proportion of systematic variance attributable to interpersonal differences and intrapersonal change for each of the 609 questions analyzed, plotted separately by panel.
+Figure 1 plots the proportion of systematic variance attributable to interpersonal differences and intrapersonal change for each of the 609 questions, plotted separately by panel.
 To broadly summarize the results, all panels show a similar range of the proportion of systematic variance attributable to intrapersonal change.
@@
 ```
-
-Across all questions, interpersonal differences explain a much larger share of the systematic variance in responses than intrapersonal change.
+Across all questions, interpersonal differences account for a much larger share of the systematic variance in responses than intrapersonal change.
 Again, this is to be expected.
-Interpersonal differences capture not just pre-adult socialization, but all accumulated experiences up to the start of the panel that might have a bearing on personal culture.
+Interpersonal differences capture not just pre-adult socialization, but all accumulated experiences up to the start of the panel that might influence personal culture.
 
@@
 While we cannot disentangle features of the sample from features of the questions asked to each sample, the specific samples for many PSID questions have lower average ages than those from other panels.
-To the extent that younger respondents might be more likely to make durable changes of opinion, this higher explanatory power of intrapersonal change might reflect the distinct age profile of respondents in this sample.
+To the extent that younger respondents might be more likely to make durable changes of opinion, these higher estimates of intrapersonal change might reflect the distinct age profile of respondents in this sample.
 At the other end, the GSS has the lowest range of $\omega$ with mean .073 and median .069.
@@
 While there are some differences between panels, these differences are small compared to the differences within panels.
-For about 6 percent of items explored here, $\omega$ is greater than 0.20.
-These questions tend to ask about external referents (e.g., confidence in specific government leaders or political parties), life satisfaction, or current financial position.
-At the other end, questions about religious identification, views on gender roles, and support for civil liberties tend to have very low proportion of variance attributable to intrapersonal change.
+For about 6 percent of items, $\omega$ is greater than 0.20.
+These questions tend to ask about objectively changing external referents (e.g., confidence in specific government leaders or political parties), life satisfaction, or current financial position.
+At the other end, questions about religious identification, views on gender roles, and support for civil liberties tend to have very low estimates of intrapersonal change.
 
@@
 For example, Kiley and Vaisey [-@kiley2020] found that confidence in the press and confidence in religious leaders were both characterized by active updating.
-Our results show that intrapersonal change is much more important for explaining variance in confidence in the press (0.164) than confidence in religion (0.049).
+Our results show that intrapersonal change is much more important for explaining variance in confidence in the press (0.164) than confidence in religion (0.049), even though both are updating.
 
-Supplemental Materials A presents the estimated proportion of variance attributable to interpersonal differences and intrapersonal change, the estimated values of $\omega$, and the proportion of residual variance for each question analyzed.
-In general, interpersonal differences are almost always the largest component of the total variance and tend to explain between 55 and 70 percent of total variance, while intrapersonal change is always the smallest, typically explaining between 3 and 8 percent of total variance.
+Supplemental Materials A presents the estimated proportion of variance attributable to interpersonal differences and intrapersonal change, the estimated values of $\omega$, and the proportion of residual variance for each question.
+Interpersonal differences are almost always the largest component of the total variance and tend to account for between 55 and 70 percent of total variance, while intrapersonal change is always the smallest, typically accounting for between 3 and 8 percent of total variance.
 Residual variance tends to account for between 22 and 37 percent of variance, though on several questions residual variance is greater than 50 percent.
+This might indicate survey items with low reliability or ones that capture genuinely rapid fluctuations.
 
-As noted above, the substantive importance of intrapersonal change for understanding cultural differences depends on a range of factors, including how long the panel runs, whether assumptions about linear change hold, and whether the panel is capturing a distinctly turbulent period or a distinctly stable period for the item under question.
-However, if we assume that the period under observation is "typical" for a question ---not a time of extremely heightened (or lowered) sensitivity or change--- then it does not seem realistic for this process of intrapersonal change to explain a large share of interpersonal differences for many of the questions we observe.
+As we noted above, the substantive importance of a given amount of intrapersonal change depends on a range of factors, including how long the panel runs, whether assumptions about linear change hold, and whether the panel is capturing a distinctly turbulent period or a distinctly stable period for the relevant item.
+However, if we assume that the period under observation is "typical" for a question---not a time of extremely heightened (or lowered) sensitivity or change---then it does not seem realistic for that intrapersonal change accounts for a large share of the cultural differences we see in the world.
 
@@
 
-Figure 2 plots the results from an ordinary least squares regression of $\omega$ as a function of question, panel, and sample features.
+Figure 2 plots the results from a linear regression of $\omega$ as a function of question, panel, and sample features.
 These models also include fixed effects for panels and topics, so coefficients reflect the association within a panel and topical domain.
@@
 ```
-
 Figure 2 shows that the more response options respondents are given and the larger the sample, the larger the values of $\omega$.
@@
 
-While we did not generate expectations for how question content relates to $\omega$, some of the associations between question structure, panel duration, and panels themselves might be driven by differences in the topics addressed by each panel.
+Although we did not state expectations for how question content would relate to $\omega$, some of the associations between question structure, panel duration, and panels themselves might be driven by differences in the topics addressed by each panel.
 To address this, we followed Hout et al.
@@
 The longer a question is observed, the less is the value of $\omega$.
-Theories of personal cultural change that link changes in personal culture to social experiences, including the LCAM, suggest that the longer we observe respondents, the more likely people are to undergo potentially transformative experiences and therefore the more variance would be explained by intrapersonal change.
-The fact that we find a negative coefficient attached to this variable can be a challenge to that assumption.
+Theories of personal cultural change that link changes in personal culture to social experiences, including the LCAM, suggest that the longer we observe respondents, the more likely people are to undergo potentially transformative experiences and therefore the more variance would be attributable to  intrapersonal change.
+Finding a negative coefficient here seems to challenge to that assumption.
 
-As a further test of this coefficient, we compared the $\omega$ values when using the full duration of a panel compared to when we dropped the final wave and therefore reduced the total duration of observation for the question.
+As a further test of this finding, we compared the $\omega$ values when using the full duration of a panel compared to when we dropped the final wave and therefore reduced the total duration of observation for the question.
 If the coefficient reflects a true negative effect of duration on $\omega$, we should see that same effect within questions.
-Results from this analysis are presented in Supplemental Materials C and contradict the coefficient from the regression model: we found that the longer we observe the same question, the higher the $\omega$ is.
+Results from this analysis are presented in Supplemental Materials C and contradict the coefficient from the regression model; with a few exceptions, we found that the longer we observe the same question, the higher $\omega$ is.
 We interpret this combination of findings as suggesting that the kinds of questions asked for a longer time period tend to demonstrate less intrapersonal change than questions asked over shorter periods, rather than a true function of time.
-This reinforces the explanation for why the GSS displays the lowest average proportion of systematic variance attributable to intrapersonal change: its shorter duration.
+This supports the conclusion that the GSS shows less intrapersonal change because of its shorter duration.
 
@@
 These results mostly re-frame and align previous findings, but the value of our approach lies in its ability to extend the debate to a broader set of theoretical questions.
-To show this potential, we turn to a specific empirical example: the relative importance of intrapersonal change and interpersonal difference for explaining cultural variation by education levels.
+To show this potential, we turn to a specific empirical example: the relative importance of intrapersonal change and interpersonal difference for explaining cultural variation by level of education.
 
@@
 This work has tended to focus on the fact that college-educated Americans give responses that are less likely to be affected by measurement error or short-term influences than the rest of the population [@zaller1992; @alwin2007].
-But focus on the difference in the non-systematic or residual component of variance across groups obscures the fact that the relative importance of different components of systematic variance, intrapersonal change and interpersonal differences, might also differ across these groups.
+But this focus on the non-systematic or residual component of variance across groups lumps the two systematic forms of difference together.
+That is, it obscures the fact that the amount of interpersonal differences and intrapersonal change might also differ across these groups.
 
-There are theoretical reasons to believe that education might be associated with either more or less systematic variance attributable to intrapersonal change.
-On one hand, because college graduates are more connected to mainstream discourse and elite signals, they might be more likely to make durable changes in response to the emergence of new information, new issues, or political realignments, while those without chronic information display more variance around their baselines without their baselines changing much [@zaller1992].
-In contrast, it could be the case that the stability demonstrated by educated respondents reflects the fact that they have already formed durable opinions and are relatively closed off to new information.
-In this perspective, college could serve as a formative experience that solidifies some dimensions of personal culture for people who attend.
-But among people who do not attend college, later life experiences might prove more important in forming or changing personal culture, as these experiences potentially provide information that college-educated peers received earlier.
+There are theoretical reasons to believe that education might be associated with either more or less intrapersonal change.
+On one hand, because college graduates are more connected to mainstream discourse and elite signals, they might be more likely to make durable changes in response to the emergence of new information, new issues, or political realignments, while those without chronic information might display more variance around an unchanging baseline [@zaller1992].
+Conversely, it could be that the observed stability of the college educated reflects the fact that they have already formed durable opinions and are relatively closed off to new information.
+If this is true, college could be understood as a formative experience that solidifies some dimensions of personal culture.
+Perhaps for those who do not attend college, later life experiences might prove more important in forming or changing personal culture, as these experiences potentially provide information that college-educated peers have already received.
 
-To test these competing propositions, we use the models presented above to separately calculate $\omega$ values for people with at least a bachelor's degree and people with less than a bachelor's degree at wave 1[^4] of the General Social Survey's panels.
+To compare these competing propositions, we calculate $\omega$ values separately for people with at least a bachelor's degree and people with less than a bachelor's degree at wave 1[^4] of the three General Social Survey's panels.
 We focus on the GSS because it contains the largest number of questions tapping general political dispositions, which is the domain where education has proven particularly relevant for understanding attitude stability.
 The GSS also covers a turbulent window of American politics from 2006 to 2014.
-This window covers the start of the Great Recession, debates about federal intervention in and regulation of Wall Street, the election of Barack Obama as the first black president in U.S.
-history, debates about the role of the federal government in the health care sector, the emergence of the Tea Party, and political realignment and clarification on the issue of gay marriage, among other topics.
+This window covers the start of the Great Recession, debates about federal intervention in and regulation of Wall Street, the election of Barack Obama as the first black U.S.
+president, debates about the role of the federal government in the health care sector, the emergence of the Tea Party, and political realignment and clarification on the issue of gay marriage, among other topics.
 
@@
 
-Figure 3 plots the distribution of differences in $\omega$ values between people with at least a bachelor's degree and people with less than a bachelor's degree at wave 1 of the panel for all 183 GSS items we explored.
-Values greater than 0 indicate that intrapersonal change is more important in explaining the responses of college graduates than those of people without a college degree, while values less than 0 indicate that intrapersonal change is less important for explaining variation among people with a college degree than those without a college degree.
+Figure 3 plots the distribution of differences in $\omega$ values between people with at least a bachelor's degree and people with less than a bachelor's degree at wave 1 of the panel for 183 GSS items.
+Values greater than 0 indicate that intrapersonal change accounts for more systematic variance among college graduates than among those without a college degree, while values less than 0 indicate the opposite.
 
@@
 ```
-
-There is a clear pattern in Figure 3: for more than 80 percent of GSS items explored here, intrapersonal change is a larger component of systematic variance for people without a college degree than for people with a college degree (i.e., the difference is negative).
-While most of these differences are small in absolute terms, less than 2 percentage points, several are greater than 5 percentage points.
+There is a clear pattern in Figure 3: for more than 80 percent of these GSS items, intrapersonal change is a larger component of systematic variance for people without a college degree.
+While most of these differences are small in absolute terms (less than 2 percentage points), several are greater than 5 percentage points.
 Given the distribution observed in Figure 1 showing that the systematic variance attributable to intrapersonal change averages around 0.09, a 5 percentage point difference between groups is quite substantial.
@@
 conservative) on seven-point scales; four questions about the government's role in improving the condition of the poor, paying people's medical bills, giving special treatment to black people, and doing things that private businesses could do, measured on five-point scales; a question about whether the government should do more to reduce income differences, measured on a seven-point scale; and one question about whether black people should be given preferences in hiring, measured on a five-point scale.
-Estimates of $\omega$ for these eight questions, for both education groups, are presented in Figure 4.
+We present estimates of $\omega$ for these eight questions, for both education groups, in Figure 4.
 
@@
 ```
-
-On all eight questions presented in Figure 4, $\omega$ values are smaller for people with a college degree than for people without a college degree, meaning intrapersonal change explains less systematic variance for people with a college degree than for people without one.
+On all eight questions presented in Figure 4, $\omega$ values are smaller for people with a college degree, meaning intrapersonal change is less common among them.
 There are also large differences in $\omega$ values across questions for both groups.
@@
 
-In contrast, partisan identification and political ideology both show larger values of $\omega$ than most other questions, as well as a larger absolute difference between people with a college degree and those without.
-Compared to the other questions explored here, change as adults plays a much larger role in explaining why people report the partisan identification and political ideology they do.
-And this is particularly true for respondents who do not have a college degree.
-The $\omega$ value for ideological identification among non-college educated respondents is almost four times that of college-educated respondents.
+In contrast, partisan identification and political ideology both show larger values of $\omega$ than most other questions, as well as a larger absolute difference by degree status.
+Compared to the other questions, intrapersonal change plays a much larger role in accounting for partisan identification and political ideology.
+And this is particularly true for respondents who do not have a college degree; the $\omega$ value for ideological identification among non-college educated respondents is almost four times that of college-educated respondents.
 
-We note here that this meaningful difference in $\omega$ values across education groups and across questions would not be detectable using previous methods.
-For partisan identification and political ideology, both college-educated respondents and people with less than college degree would likely favor the AUM or LCAM over the SDM because they these questions both show evidence of some members of the population making intrapersonal change.
-In other words, using the tournament of models approach obscures the fact that the relative importance of intrapersonal change in explaining differences on these two items is different across these two groups, that intrapersonal change explains more systematic variance for ideological identification than partisan identification for people without a college degree, and that there appears to be more durable change on questions of affirmative action than on questions of government aid to black Americans.
+It is worth pointing out that this meaningful difference in $\omega$ values across education groups and across questions would not have been detectable using previous methods.
+For partisan identification and political ideology, both college-educated respondents and people with less than college degree would likely favor the AUM or LCAM over the SDM because they these questions both show evidence of some members of the population making *some* intrapersonal change.
+In other words, the tournament of models obscures the fact that the relative importance of intrapersonal change differs across these two groups, that intrapersonal change explains more systematic variance for ideological identification and partisan identification for people without a college degree, and that there appears to be more durable change on questions of affirmative action than on questions of government aid to black Americans.
 
-We believe these patterns shed new light on the mechanisms underlying differences in attitude behavior across groups.
-Something about college attendance seems to limit the ability of experiences later in life to drive changes in personal culture.
+We believe these patterns shed new light on the mechanisms underlying differences in attitudes and behavior across groups.
+Something about college attendance seems to crystallize personal culture.
 While these results should not be interpreted as causal effects of attending college -- they are potentially confounded by age, social class, race, gender, and other factors that explain selection into higher education -- they open up a set of new questions and dynamics to explore.
 
-# Discussion
+# Discussion and Conclusion
 
-This paper set out intervene in the debate on whether people change their personal culture -- their attitudes, beliefs, values, practices, and dispositions -- as they move through their adult life.
+The objective of this paper is to intervene in the recent debate on whether people change their personal culture -- their attitudes, beliefs, values, practices, and dispositions -- as they move through their adult life.
 Instead of falling back on a "tournament of models" approach [@lersch2023: p.
-228] to proclaim victory for only one answer to this question, it seeks to refocus the debate to the theoretically more productive question of the relative importance of intrapersonal change for explaining differences in the personal culture.
-It proposes a new approach that quantifies the amount of observed systematic variance that is attributable to either interpersonal differences at baseline or intrapersonal change over the duration of a panel.
+228] to proclaim victory for a particular answer to this question, our goal is to refocus the debate to the theoretically more productive question of the relative importance of intrapersonal change for explaining differences in the personal culture.
+We propose a new approach that quantifies the amount of observed systematic variance that is attributable to either interpersonal differences at baseline or intrapersonal change over the duration of a panel.
 
-Applying our proposed measure, $\omega$, to all questions about personal culture from the panel datasets previously discussed in this debate revealed a consistent pattern.
-Nearly all questions show evidence of people making durable, intrapersonal change over time, with some showing notably high intrapersonal change compared to interpersonal differences.
-For some questions about factors that affect one's life satisfaction or views on government benefits, it seems plausible that differences in adult experiences predominantly account for the observed differences between people.
+Applying our proposed measure, $\omega$, to 609 survey items from all of the panel datasets previously discussed in this debate revealed a consistent pattern.
+Nearly all questions show evidence of people making some durable, intrapersonal change over time.
+Some show notably high amounts of intrapersonal change.
+For some questions about life satisfaction or views on government benefits, it seems plausible that differences in adult experiences predominantly account for observed differences between people.
 
@@
 On numerous questions, like those on civil liberties, abortion, generalized trust, and civic duty, the systematic variance attributable to intrapersonal change is essentially zero.
-For these questions, it seems there is not enough cultural change during adulthood to warrant attributing differences in personal culture mainly to experiences and social transitions; instead, the primary source of the observed differences appears to stem from experiences in childhood, adolescence, and early adulthood.
+For these questions, it seems there is not enough cultural change during adulthood to warrant attributing the differences we observe to experiences and social transitions; instead, the primary source of the observed differences appears to stem from experiences in childhood, adolescence, and early adulthood.
 
-More than adjudicating past positions in this debate, the approach presented in this paper enables scholars to more clearly explore the relative contributions of change and stability in explaining variation in personal culture, as well as how these contributions might differ across populations.
-To exemplify this, we detailed how the amount of cultural difference explained by intrapersonal change varies substantially by survey item but also by individuals' characteristics such as their education.
-Intrapersonal change plays almost no part in explaining sentiments regarding the government's economic role, but it is more defining for ideological and partisan identities.
-Intriguingly, it is more important in explaining cultural variation among non-college graduates compared to college graduates in the United States.
-While age and other factors related to college completion might confound this pattern, it suggests that college completion may solidify personal culture to an extent that renders the effect of other adult experiences on political dispositions less influential for college graduates relative to non-college graduates.
+More than adjudicating past positions in this debate, the approach we presented in this paper will enable scholars to explore more precisely the relative contributions of change and stability in explaining variation in personal culture, as well as how these contributions might differ across populations.
+To illustrate this, we showed how the amount of cultural difference explained by intrapersonal change varies substantially by survey item but also by individuals' characteristics such as their education.
+While age and other factors related to college completion might confound this pattern, it suggests that college completion may crystallize personal culture to an extent that renders later adult experiences on political dispositions less influential.
+
+## Limitations
+
+Although we believe it is a major advance on previous work, the approach we outline here has important limitations.
+We allocate all systematic variance to one of two sets of theoretical processes: intrapersonal change and interpersonal differences at baseline.
+Our approach does not quantify the proportion of people who "change," nor can we be sure that the amount of interpersonal change we detect is driven by many people making small changes or a few people making large changes.
+
+Our results and interpretations also hinge on how we have defined change.
+As does the LCAM, our approach treats cultural trajectories as varying linear slopes for each respondent, thus assuming that change is a linear function of time.
+This assumption simplifies reality in which change likely also takes non-linear and discontinuous forms.
+People might jump from one "stable" disposition to another or experience a "turning point" in their lives that upsets an otherwise stable trajectory.
+In parts, our assumption of linearity is a limitation of the data, as most questions are only observed for three to six waves.
+Panels with more waves might allow researchers to loosen this assumption to test alternative, more flexible models of change.
+
+Similarly, our approach assumes that durable change is unidirectional.
+This is a sensible assumption on short panels where classifying change that lasts less than two years as durable seems unreasonable.
+Practically, it means that the variance produced by people making durable changes in their cultural dispositions to then return to a previous state later in life is classified as residual variance, rather than intrapersonal change.
+Longer panels and more flexible definitions of "change" might allow us to account for such trajectories.
+
+Finally, we had chosen to examine the broadest array of measures of personal culture available to us, ranging from religious beliefs and core values, to policy preferences, and even to the importance of different features when buying a new car across five countries (Australia, Germany, Switzerland, the United Kingdom and the United States).
+Nevertheless, our findings remain limited to the kinds of questions that are asked in panel surveys and in the contexts they were administered, reflecting issues of general (national) politics, gender roles, immigration and race relations, and general well being.
+Although we have no reason to believe results to be different, our findings do not directly speak to other dimensions of culture such as artistic tastes, leisure activities, and time use.
 
@@
 
-Sociologists interested in understanding cultural differences have largely asked about the presence and significance of cultural change in adults [e.g., @kiley2020; @vaisey2021; @lersch2023].
-Yet doing so has unnecessarily limited and might even set the wrong focus.
-In any population, some degree of cultural change in adults is inevitable.
-Likewise, no viable theoretical perspective would expect just one process -- stable differences or internal change -- to shape their personal culture.
-Our results reinforce this point, showing the relevance of both interpersonal differences and intrapersonal change in understanding the differences in personal culture among adults across all items analyzed.
+Despite these limitations, we believe our method and findings have important implications for social science research.
+Sociologists interested in understanding cultural differences have largely asked about the *existence* of cultural change in adults [e.g., @kiley2020; @vaisey2021; @lersch2023].
+Yet this approach has inadvertently limited the debate.
+In any population, *some* degree of adult cultural change is inevitable.
+Although ideal types like the settled dispositions and active updating models are useful, no theoretical perspective would expect either early life socialization or adult intrapersonal change to be the sole source of one's personal culture.
+Our results reinforce this point, showing the relevance of both factors and allowing their precise quantification.
 
-This and other recent findings [@stewart2023; @quinn2023] suggest that it is theoretically more productive to ask about the relative importance of these two processes in understanding cultural variation.
+This and other recent findings [@stewart2023; @quinn2023] suggest that it is theoretically more productive to measure the relative importance of these two components in concrete, substantive cases.
 Drawing a unified conclusion from survey items measuring various cultural forms on different scales and across different time frames is challenging.
-However, the general pattern suggests that, for most items, intrapersonal change in adulthood is not the primary reason for cultural differences among people.
+Nevertheless, the general pattern suggests that, for most items, intrapersonal change in adulthood is not the primary reason we see the differences we see between people in the world.
 
-To clarify, our findings do not imply that adults remain static or that their changes are inconsequential.
-Even minor shifts in personal culture, such as a 2 percentage-point change in support for gay marriage, can have significant ramifications.
-While the majority may remain consistent in their views, understanding the underlying mechanisms of even such minor shifts remains a crucial task for the sociology of culture.
+We are not claiming that adults remain static or that their changes are inconsequential.
+Even quantitatively minor shifts, such as a 2 percentage-point change in support for gay marriage, can have massive ramifications.
+While the majority may remain consistent in their views, understanding the underlying mechanisms of even such numerically minor shifts remains a crucial task for the sociology of culture.
 
-Our findings primarily suggest that understanding variations in personal culture requires scholars to examine the conditions and experiences of early life.
-While transitions between social roles, changes in social networks, and the experience of organizational environments are often emphasized in sociological research, these factors seem to account for a smaller proportion of differences among adults than the formative experiences early in life.
+Our findings suggest that understanding variation in personal culture requires examining the conditions and experiences of early life.
+While sociological research often focuses on transitions between social roles, changes in social networks, and the experience of organizational environments, these factors seem to account for a smaller proportion of adult differences than early-life experience.
+Simply put, we need more work on early-life socialization [@guhin2020].
 
@@
 Our findings also have implications for understanding cultural change at the aggregate level.
-Given that intrapersonal change does not appear as the primary driver for many cultural differences among adults, our results suggest that many forms of cultural change at the aggregate level are more likely driven by processes of cohort replacement than the by contemporaneous social conditions changing people's cultural dispositions [@vaisey2016; @underwood2022].
+Given that intrapersonal change accounts for a relatively small amount of variance among adults, many forms of cultural change at the aggregate level are necessarily more likely driven cohort replacement than by contemporaneous social conditions [@vaisey2016; @underwood2022].
 This likely holds true for cultural change at the macro societal level as well as the micro level such as within organizations, political parties, and professions.
@@
 The weight of these two factors might thus change as the prominence of education shifts in the life course.
-If scholars find similar differences across other social categorizations like class or race, it would advocate for a more thorough integration of demographic processes into the analysis of cultural change at the aggregate level.
+If scholars find similar differences across other social categorizations like class or race, it might necessitate a more thorough integration of demographic processes into the analysis of cultural change at the population level.
 
@@
 
-Further, we align with others [**WE NEED SOME CITES HERE**] advocating for more studies, including panel studies, focusing on the socialization of youth.
-Our analysis points to the majority of variance among adults rooting in differences encountered before turning 18.
-By empirically constraining their analysis to experiences after that point, social scientists may inadvertently concentrate on topics and questions that, while important, might not be able to help explain major cultural differences in adult populations.
-
-Finally, a key insight from our meta-analysis is that an increase in response resolution (the range of response options survey respondents are given) correlates with an increased share of systematic variance explained by intrapersonal change.
+A key insight from our meta-analysis is that an increase in response resolution (the range of response options survey respondents are given) correlates with an increased share of systematic variance explained by intrapersonal change.
 This may reflect specific issues being assessed by survey items using different scales.
@@
 
-## Limitations
+Perhaps most important, we need more panel studies on youth.
+It is very likely that the majority of adult differences in personal culture are rooted in different experiences before age 18.
+By empirically constraining our analyses to adult experiences, social scientists may inadvertently concentrate on topics and questions that, while important, might not be able to help explain major cultural differences in adult populations.
 
-The approach outlined here generates estimates of how much of the observed variance in a population can be attributed to one of two sets of theoretical processes: intrapersonal change and interpersonal differences at baseline.
-It does not quantify the proportion of people who "change," nor can we be sure that the amount of interpersonal change we detect is driven by many people making small changes or a few people making large changes.
+## Conclusion
 
-Our results and interpretations also hinge on how we have defined change.
-Similarly to the LCAM, our approach treats cultural trajectories as varying linear slopes for each respondent, thus assuming that change is a linear function of time.
-This assumption simplifies reality in which change likely also takes non-linear and discontinuous forms.
-People might jump from one "stable" disposition to another or experience a "turning point" in their lives that upsets an otherwise stable trajectory.
-In parts, our assumption of linearity is a limitation of the data, as most questions are only observed for three to six waves.
-Panels with more waves might allow researchers to loosen this assumption to test alternative, more flexible models of change.
-
-Similarly, our approach assumes that durable change is unidirectional.
-This is a sensible assumption on short panels where classifying change that lasts less than two years as durable seems unreasonable.
-Practically, it means that the variance produced by people making durable changes in their cultural dispositions to then return to a previous state later in life is classified as residual variance, rather than intrapersonal change.
-Longer panels and more flexible definitions of "change" would allow us to account for such trajectories and thus potentially also identify more durable change even in the same population.
-
-Finally, we had chosen to examine the broadest array of measures of personal culture available to us, ranging from religious beliefs and core values, to policy preferences, and even to the importance of different features when buying a new car across five countries (Australia, Germany, Switzerland, the United Kingdom and the United States).
-Nevertheless, our findings remain limited to the kinds of questions that are asked in panel surveys and in the contexts they were administered, reflecting issues of general (national) politics, gender roles, immigration and race relations, and general well being.
-Although we have no reason to believe results to be different, our findings do not directly speak to other dimensions of culture such as artistic tastes, leisure activities, and time use.
-
-## Implications for Future Research
-
-We believe the approach outlined here can push past the "needless dichotomy" implicit in the question of whether people change or not and address more theoretically informed questions about the conditions under which the relative importance of intrapersonal change and interpersonal differences in explaining cultural differences shifts.
+We believe the approach we outlined here can push past the "needless dichotomy" implicit in the question of whether people change or not [@lersch2023].
 Characterizing questions as displaying change or not can only take researchers so far, but the question of whether some questions demonstrate more change than others, or whether some groups are characterized by more stability than others, has the potential to weigh in on a broader range of theoretical debates.
-We hope researchers move toward those questions.
-
-While we focused on education as a determinant of the explanatory power of intrapersonal change, our approach can readily be extended to other socio-demographic categories that might affect intrapersonal change -- race/ethnicity, religiosity, social class, gender, and political affiliation, for example -- or their intersections.
-Researchers with different data might be able to evaluate the relative importance of intrapersonal change and interpersonal differences across contexts as well, either national contexts if the same questions are administered across countries, or organizational contexts if data traces students within schools, workers within workplaces, or congregation members within congregations.
-
-The wide variation across questions in the proportion of systematic variation attributable to intrapersonal change also calls for a more systematic exploration for why some forms of culture are characterized by stability while others are characterized by change, and what the consequences of these differences for the organization of beliefs in the public might be.
+We hope researchers find our approach useful as they investigate these questions.
 

### 2023-09-29 manuscript/manuscript.tex
small update in the manuscript.
SV edits

@@
 lack of durable change is consistent with other recent findings that
-cohort replacement plays a larger role than period effects in explaining
-differences in personal culture (Vaisey and Lizardo 2016).
+cohort replacement plays a somewhat larger role than period effects in
+explaining differences in personal culture (Vaisey and Lizardo 2016).
 
@@
 culture. Lersch's LCAM does this by modeling each individual as
-following their own linear trajectory. But rather than debating whether
+following a personal linear trajectory. But rather than debating whether
 this model fits a particular data set better than a model that assumes
@@
 asking, ``do people change?'' to asking, ``what are the relative
-contributions of change and initial differences for explaining
-differences in personal culture?''
-
-We propose an approach that offers such a quantification. In this
-approach, there are two metrics that each reflect a combination of
-relevant theoretical processes. The first metric, interpersonal
-differences at the start of the survey, reflects the accumulated
-experiences of people prior to entering the panel survey that shape how
-they answer questions. Lersch calls this ``early imprinting.'' While
-commonly associated with experiences during a formative period that
-result in settled dispositions, this variation could also reflect
-experiences that happened at any stage in life as long as they predate
-the panel and consistently affect subsequent responses. For instance,
-for those people entering the panel post-retirement, this ``imprinting''
-might reflect this pivotal life transition. Consequently, this metric
-also reflects variation in individuals' social roles or statuses at the
-start of the panel that were important in shaping their dispositions.
+contributions of change and initial differences for explaining variance
+in personal culture?''
+
+We propose an approach that offers such a quantification. Our measure
+relates two metrics that each reflect a combination of relevant
+theoretical processes. The first metric, interpersonal differences at
+the start of the survey, reflects the accumulated experiences of people
+prior to entering the panel survey that shape how they answer questions.
+Lersch calls this ``early imprinting.'' While commonly associated with
+experiences during a formative period that result in settled
+dispositions, this variation could also reflect experiences that
+happened at any stage in life as long as they predate the panel and
+consistently affect subsequent responses. For instance, for those people
+entering the panel post-retirement, this ``imprinting'' might reflect
+this pivotal life transition. Consequently, this metric also reflects
+variation in individuals' social roles or statuses at the start of the
+panel that were important in shaping their dispositions.
 
@@
 during the panel, captures durable changes in personal culture over
-time, a set of processes collectively called ``persistent change'' or
-``adaption'' by Lersch and ``active updating'' by Kiley and Vaisey.
-Lersch attributes these changes to social triggers such as moving into a
-new environment or adopting new social roles (although he never measures
-them directly). Additionally, they might reflect the diffusion of new
-cultural forms across social networks, cues from political elites or
-otherwise culturally influential leaders, the emergence of issues in
-politics or culture, or large-scale social shifts.
-
-A third metric, often called ``residual variance'' or ``fluctuation,''
-accounts for the remainder of variance in peoples' responses. This
-non-durable change emerges for a variety of reasons. For example, people
-might not have a clear disposition on a particular item as it is asked.
-Instead, they might internalize a broad set of considerations and
-construct an opinion in the context of the survey interview, with
-different considerations coming to the forefront of their cognition
-during each interview (Feldman and Zaller 1992; Tourangeau, Rips, and
-Rasinski 2000; Zaller 1992). Additionally, this variance can also
-include measurement error, such as misinterpreted questions, erroneous
-response selections, or responses getting coded incorrectly. While this
-third metric at least partially reflects important processes of personal
-culture, it does not directly touch on the ongoing debate outlined here.
-Because of this we quantify this component but focus principally on the
-other two.
+time. This captures the set of processes collectively called
+``persistent change'' or ``adaption'' by Lersch and ``active updating''
+by Kiley and Vaisey. Lersch attributes these changes to social triggers
+such as moving into a new environment or adopting new social roles
+(although he never measures these directly). Additionally, they might
+reflect the diffusion of new cultural forms across social networks, cues
+from political elites or otherwise culturally influential leaders, the
+emergence of issues in politics or culture, or large-scale social
+shifts.
+
+A third metric, which we might call ``residual variance'' or
+``fluctuation,'' accounts for the remainder of variance in peoples'
+responses. These non-durable changes emerge for a variety of reasons.
+For example, people might not have a clear disposition on a particular
+item as it is asked. Instead, they might internalize a broad set of
+considerations and construct an opinion in the context of the survey
+interview, with different considerations coming to the forefront of
+their cognition during each interview (Feldman and Zaller 1992;
+Tourangeau, Rips, and Rasinski 2000; Zaller 1992). This variance can
+also include measurement error, such as misinterpreted questions,
+erroneous response selections, or responses getting coded incorrectly.
+Although this third metric at least partially reflects important
+processes of personal culture, it does not directly touch on the ongoing
+debate outlined here. Because of this, we quantify this component but
+focus principally on the other two.
 
@@
 First, because previous work found evidence of durable change on most
-measures of personal culture, we expect that intrapersonal change will
-have some explanatory power for most items. But because these earlier
-studies also found many questions (between 25 and 35 percent) lacking
-evidence of durable change, we expect considerable variation in the
-variance explained by intrapersonal change, with some questions
+measures of personal culture, we expect that we will find some
+intrapersonal change in most items. But because these earlier studies
+also found many questions (between 25 and 35 percent) lacking evidence
+of \emph{any} durable change, we expect considerable variation in the
+variance accounted for by intrapersonal change, with some questions
 indicating essentially no variance attributable to such change.
 
-Second, we expect the variance explained by intrapersonal change to be
-modest compared to interpersonal differences at the start of the survey.
-This expectation should be uncontroversial, regardless of the
+Second, we expect the variance attributable to intrapersonal change to
+be modest compared to interpersonal differences at the start of the
+survey. This expectation should be uncontroversial, regardless of the
 theoretical process assumed to be most relevant for a particular
@@
 ``Formative period'' and cohort theories, which emphasize early-life
-socialization as a strong force in shaping adult personal culture,
-suggest that the variance from intrapersonal change would still fall
-short of explaining interpersonal differences, even when we account for
-the duration of observation. Conversely, other theoretical perspectives
+socialization as a strong influence on adult personal culture, suggest
+that the variance from intrapersonal change would still fall short of
+explaining interpersonal differences, even when we account for the
+duration of observation. Conversely, other theoretical perspectives
 emphasizing the importance of people's contemporaneous social
-environment for personal culture expect intrapersonal change to account
-for a large portion of the variance, regardless of the duration.
+environment for personal culture might expect intrapersonal change to
+account for a large portion of the variance, regardless of the duration.
 
@@
 all possible cases for which respondents provided responses. In the end,
-the analyses that follow rely on a cross-cultural sample with a
+the analyses that follow rely on a cross-national sample with a
 cross-domain set of items to capture a broad range of individual
@@
 assumes that people deviate around this baseline randomly over time,
-reflecting ``fluctuation'' or short-term non-persistent change. Finally,
-it assumes that people draw on earlier time points in reacting to new
-events. Formally, this can be written as
+reflecting ``fluctuation'' or short-term non-persistent change.
+Formally, this can be written as
 
 \[
-y_{i, t} = \beta_0 + u_{0, i} + \beta_{1} x_t + u_{1, i} x_{i, t} + \epsilon_{i, t}
-\] \[
-u_{0,i} \sim \mathcal{N}(0,\sigma_{\mu_0}^2) \\
-\] \[
-u_{1,i} \sim \mathcal{N}(0,\sigma_{u_1}^2)
+y_{it} = \beta_0 + \alpha_{0i} + (\beta_1 + \alpha_{1i}) \text{year}_{it} + \epsilon_{it}
 \]
 
-where \(\beta_0\) is the average intercept, \(u_{0, i}\) is the random
-intercept for individuals, \(\beta_1 x_t\) is the average slope,
-\(u_{1, i} x_{i, t}\) is the random slope for individuals, and
-\(\epsilon_{i, t}\) is the random error term that captures transitory
-disturbances.
+\[
+\begin{aligned}
+\alpha_{0i} &\sim \mathcal{N}(0,\tau^2_{0}) \\
+\alpha_{1i} &\sim \mathcal{N}(0,\tau^2_{1}) \\
+\epsilon_{it} &\sim \mathcal{N}(0,\sigma^2)
+\end{aligned}
+\]
+
+where \(\beta_0\) is the average intercept, \(\alpha_{0i}\) is the
+random intercept component for individuals, \(\beta_1\) is the average
+yearly change in the outcome, \(\alpha_{1i}\) is the random slope
+component for individuals, and \(\epsilon_{it}\) is the random error
+term that captures transitory disturbances. The \(\alpha_{0i}\) and
+\(\alpha_{1i}\) terms are also allowed to covary.
 
@@
 
-As a first step in our analysis, we fit a LCAM to each of the 609
+As a first step in our analysis, we fit this LCAM to each of the 609
 measures of personal culture outlined above. We then use these models to
@@
 collections of theoretical processes outlined above: interpersonal
-differences at baseline and intrapersonal change over time. First, to
-measure the importance of interpersonal differences, we calculate the
-sum of the squared residuals derived from measuring each respondent at
-the midpoint of their age trajectory.\footnote{We use the scale
-  midpoint, rather than the actual difference at baseline, to better
-  capture the differences between people independent of their age
-  trajectories.} We then divide this value by the total sum of squares
-and subtract it from from 1, producing a value we denote as V(D), or the
-proportion of variance attributable to interpersonal differences. This
-number is equivalent to an \(R^2\) value generated by using random
-intercepts for all respondents. Second, we calculate the sum of squared
-residuals generated by additionally accounting for the age trajectory.
-We divide this number by the total sum of squares and subtract it from
-1. We then subtract \(V(D)\) from this value. This gives us a value we
-denote as \(V(C)\), the additional proportion of variance explained by
-intrapersonal change.
-
-These two numbers reflect the total systematic variance of the model
-that can be attributed to either differences across people or linear
-change over the course of the panel.\footnote{This does not mean that
-  other factors might not explain additional variance if they were
-  included in the model. For example, peoples' responses about religious
-  beliefs and participation might systematically vary by the day of the
-  week or month in which they are surveyed. This variation is moved into
-  the ``random'' or non-systematic component of the model for us.} Our
-principal outcome of interest is the proportion of systematic variance
-accounted for by intrapersonal change, which we calculate as
+differences at baseline and intrapersonal change over time.
+
+We estimate these using a form of variance decomposition. First, to
+measure stable interpersonal differences, we calculate the sum of the
+squared residuals derived from assuming that each person spends the
+entire survey at the estimated midpoint of their LCAM
+trajectory.\footnote{We use the midpoint estimate, rather than the value
+  at the first wave, because this is the best measure of ``baseline''
+  available under the assumption that the person does not change. The
+  first wave measure alone contains an unknown amount of measurement
+  error and transient fluctuations.} We then divide this value by the
+total sum of squares and subtract this number from from 1, producing a
+value we denote as \(V(D)\). This is the proportion of total variance
+attributable to interpersonal differences. (This number is equivalent to
+the conditional \(R^2\) value generated by a model with only random
+intercepts for all respondents.) Second, we calculate another sum of
+squared residuals generated by allowing the predictions to vary by
+estimated LCAM trajectory. We divide this number by the total sum of
+squares and subtract it from 1. We then subtract \(V(D)\) from this
+value. This gives us a value we denote as \(V(C)\), the additional
+proportion of variance accounted for by intrapersonal change.
+
+These two numbers together reflect the total systematic variance of the
+LCAM that can be attributed to either differences between people or
+linear intrapersonal change over the course of the panel.\footnote{This
+  does not mean that other factors might not explain additional variance
+  if they were included in the model. For example, peoples' responses
+  about religious beliefs and participation might systematically vary by
+  the day of the week or month in which they are surveyed. This
+  variation is moved into the ``random'' or non-systematic component of
+  the model for us.} Our principal outcome of interest is the proportion
+of systematic variance accounted for by intrapersonal change, which we
+calculate as
 
@@
 
-We can think of this proportion \(\omega\) as akin to an Intraclass
-Correlation Coefficient focused only on the systematic components of the
-model while ignoring residual variance that might be attributable to
-measurement error or non-persisting change (see also the variance
-decomposition of period and cohort by Vaisey and Lizardo 2016). The
-higher this proportion is, the more systematic variance in responses is
-attributable to intrapersonal change during the panel. The lower this
-proportion, the less intrapersonal change seems to play a large role in
-explaining systematic cultural differences during the panel. Moving
-forward, we use \(\omega\) as a placeholder for \emph{systematic
-variance attributable to intrapersonal change}.
+We can think of \(\omega\) as a type of Intraclass Correlation
+Coefficient focused only on the systematic components of the model while
+ignoring residual variance that might be attributable to measurement
+error or non-persisting change (see also the variance decomposition of
+period and cohort by Vaisey and Lizardo 2016). The higher this
+proportion is, the more systematic variance in responses is attributable
+to intrapersonal change during the panel. From now on, we use \(\omega\)
+as a placeholder for \emph{systematic variance attributable to
+intrapersonal change}.
 
@@
 range of questions that might have different levels of non-systematic
-variance, giving us a sense of the relative prevalence of these two
-processes for each question (something not achievable under the
-tournament of models approach), and allowing us to compare which kinds
-of questions show more or less intrapersonal change relative to
+variance. This gives us a sense of the relative prevalence of these two
+processes for each question, which is not achievable under the
+tournament of models approach. This also allows us to compare which
+kinds of questions show more or less intrapersonal change relative to
 interpersonal differences. While the tournament of models approach
-allowed researchers to say whether questions in general showed evidence
-of updating, this approach allows us to make more specific claims about
-the prevalence of these theoretical processes for particular questions
-relative to each other.
-
-As a second step in our analysis, we use ordinary least squares
-regression to model \(\omega\) as a function of features of the question
-and panel. This allows us to explore how these features affect the
-prevalence of intrapersonal change. These covariates include the number
-of response options (measured as 2, 3-5, and more than 5), the panel,
-the log of the number of participants, the number of waves observed, the
-time since the question was first asked, and the total duration of time
-the question was asked.
+allowed researchers to say whether questions in general showed
+\emph{any} evidence of updating, this approach allows us to make more
+specific claims about the prevalence of these theoretical processes for
+particular questions relative to each other.
+
+As a second step in our analysis, we use linear regression to conduct a
+meta-analysis, modeling \(\omega\) as a function of features of the
+question and panel. This allows us to explore how these features are
+associated with the degree of intrapersonal change. These covariates
+include the number of response options (measured as 2, 3-5, and more
+than 5), the specific survey, the log of the number of participants, the
+number of waves observed, the time period (i.e., decade) the question
+was first asked, and the total duration of time the question was asked.
 
@@
 interpersonal differences and intrapersonal change for each of the 609
-questions analyzed, plotted separately by panel. To broadly summarize
-the results, all panels show a similar range of the proportion of
-systematic variance attributable to intrapersonal change. Across panels,
-the median value of \(\omega\) is 0.094 (mean of 0.098), with an
-interquartile range of 0.053 to 0.139. All panels include questions
-where \(\omega\) is essentially 0, and all panels include questions with
-\(\omega\) values greater than 0.20. The maximum \(\omega\) value is
-0.26.
+questions, plotted separately by panel. To broadly summarize the
+results, all panels show a similar range of the proportion of systematic
+variance attributable to intrapersonal change. Across panels, the median
+value of \(\omega\) is 0.094 (mean of 0.098), with an interquartile
+range of 0.053 to 0.139. All panels include questions where \(\omega\)
+is essentially 0, and all panels include questions with \(\omega\)
+values greater than 0.20. The maximum \(\omega\) value is 0.26.
 
@@
 
-Across all questions, interpersonal differences explain a much larger
-share of the systematic variance in responses than intrapersonal change.
-Again, this is to be expected. Interpersonal differences capture not
-just pre-adult socialization, but all accumulated experiences up to the
-start of the panel that might have a bearing on personal culture.
+Across all questions, interpersonal differences account for a much
+larger share of the systematic variance in responses than intrapersonal
+change. Again, this is to be expected. Interpersonal differences capture
+not just pre-adult socialization, but all accumulated experiences up to
+the start of the panel that might influence personal culture.
 
@@
 younger respondents might be more likely to make durable changes of
-opinion, this higher explanatory power of intrapersonal change might
-reflect the distinct age profile of respondents in this sample. At the
-other end, the GSS has the lowest range of \(\omega\) with mean .073 and
+opinion, these higher estimates of intrapersonal change might reflect
+the distinct age profile of respondents in this sample. At the other
+end, the GSS has the lowest range of \(\omega\) with mean .073 and
 median .069. This potentially reflects the fact that the GSS observes
@@
 small compared to the differences within panels. For about 6 percent of
-items explored here, \(\omega\) is greater than 0.20. These questions
-tend to ask about external referents (e.g., confidence in specific
-government leaders or political parties), life satisfaction, or current
-financial position. At the other end, questions about religious
+items, \(\omega\) is greater than 0.20. These questions tend to ask
+about objectively changing external referents (e.g., confidence in
+specific government leaders or political parties), life satisfaction, or
+current financial position. At the other end, questions about religious
 identification, views on gender roles, and support for civil liberties
-tend to have very low proportion of variance attributable to
-intrapersonal change.
+tend to have very low estimates of intrapersonal change.
 
@@
 important for explaining variance in confidence in the press (0.164)
-than confidence in religion (0.049).
+than confidence in religion (0.049), even though both are updating.
 
@@
 estimated values of \(\omega\), and the proportion of residual variance
-for each question analyzed. In general, interpersonal differences are
-almost always the largest component of the total variance and tend to
-explain between 55 and 70 percent of total variance, while intrapersonal
-change is always the smallest, typically explaining between 3 and 8
-percent of total variance. Residual variance tends to account for
-between 22 and 37 percent of variance, though on several questions
-residual variance is greater than 50 percent.
-
-As noted above, the substantive importance of intrapersonal change for
-understanding cultural differences depends on a range of factors,
-including how long the panel runs, whether assumptions about linear
-change hold, and whether the panel is capturing a distinctly turbulent
-period or a distinctly stable period for the item under question.
-However, if we assume that the period under observation is ``typical''
-for a question ---not a time of extremely heightened (or lowered)
-sensitivity or change--- then it does not seem realistic for this
-process of intrapersonal change to explain a large share of
-interpersonal differences for many of the questions we observe.
+for each question. Interpersonal differences are almost always the
+largest component of the total variance and tend to account for between
+55 and 70 percent of total variance, while intrapersonal change is
+always the smallest, typically accounting for between 3 and 8 percent of
+total variance. Residual variance tends to account for between 22 and 37
+percent of variance, though on several questions residual variance is
+greater than 50 percent. This might indicate survey items with low
+reliability or ones that capture genuinely rapid fluctuations.
+
+As we noted above, the substantive importance of a given amount of
+intrapersonal change depends on a range of factors, including how long
+the panel runs, whether assumptions about linear change hold, and
+whether the panel is capturing a distinctly turbulent period or a
+distinctly stable period for the relevant item. However, if we assume
+that the period under observation is ``typical'' for a question---not a
+time of extremely heightened (or lowered) sensitivity or change---then
+it does not seem realistic for that intrapersonal change accounts for a
+large share of the cultural differences we see in the world.
 
@@
 
-Figure 2 plots the results from an ordinary least squares regression of
-\(\omega\) as a function of question, panel, and sample features. These
-models also include fixed effects for panels and topics, so coefficients
-reflect the association within a panel and topical domain.
+Figure 2 plots the results from a linear regression of \(\omega\) as a
+function of question, panel, and sample features. These models also
+include fixed effects for panels and topics, so coefficients reflect the
+association within a panel and topical domain.
 
@@
 
-While we did not generate expectations for how question content relates
-to \(\omega\), some of the associations between question structure,
-panel duration, and panels themselves might be driven by differences in
-the topics addressed by each panel. To address this, we followed Hout et
-al. (2016) and Lersch (2023) in coding each question as falling into one
-of nine different content categories and included these as indicator
-variables in the meta analysis. Supplemental Materials B presents the
-estimates for these content indicators. They show some associations with
-\(\omega\), with questions about subjective SES; social life, social
-cohesion, and trust; environment and climate; and health and morale
-showing larger \(\omega\) values than questions about religion and
-spirituality; politics, government and the economy; and gender and
-family.
+Although we did not state expectations for how question content would
+relate to \(\omega\), some of the associations between question
+structure, panel duration, and panels themselves might be driven by
+differences in the topics addressed by each panel. To address this, we
+followed Hout et al. (2016) and Lersch (2023) in coding each question as
+falling into one of nine different content categories and included these
+as indicator variables in the meta analysis. Supplemental Materials B
+presents the estimates for these content indicators. They show some
+associations with \(\omega\), with questions about subjective SES;
+social life, social cohesion, and trust; environment and climate; and
+health and morale showing larger \(\omega\) values than questions about
+religion and spirituality; politics, government and the economy; and
+gender and family.
 
@@
 transformative experiences and therefore the more variance would be
-explained by intrapersonal change. The fact that we find a negative
-coefficient attached to this variable can be a challenge to that
-assumption.
-
-As a further test of this coefficient, we compared the \(\omega\) values
+attributable to intrapersonal change. Finding a negative coefficient
+here seems to challenge to that assumption.
+
+As a further test of this finding, we compared the \(\omega\) values
 when using the full duration of a panel compared to when we dropped the
@@
 Results from this analysis are presented in Supplemental Materials C and
-contradict the coefficient from the regression model: we found that the
-longer we observe the same question, the higher the \(\omega\) is. We
-interpret this combination of findings as suggesting that the kinds of
-questions asked for a longer time period tend to demonstrate less
-intrapersonal change than questions asked over shorter periods, rather
-than a true function of time. This reinforces the explanation for why
-the GSS displays the lowest average proportion of systematic variance
-attributable to intrapersonal change: its shorter duration.
+contradict the coefficient from the regression model; with a few
+exceptions, we found that the longer we observe the same question, the
+higher \(\omega\) is. We interpret this combination of findings as
+suggesting that the kinds of questions asked for a longer time period
+tend to demonstrate less intrapersonal change than questions asked over
+shorter periods, rather than a true function of time. This supports the
+conclusion that the GSS shows less intrapersonal change because of its
+shorter duration.
 
@@
 change and interpersonal difference for explaining cultural variation by
-education levels.
+level of education.
 
@@
 measurement error or short-term influences than the rest of the
-population (Alwin 2007; Zaller 1992). But focus on the difference in the
-non-systematic or residual component of variance across groups obscures
-the fact that the relative importance of different components of
-systematic variance, intrapersonal change and interpersonal differences,
-might also differ across these groups.
+population (Alwin 2007; Zaller 1992). But this focus on the
+non-systematic or residual component of variance across groups lumps the
+two systematic forms of difference together. That is, it obscures the
+fact that the amount of interpersonal differences and intrapersonal
+change might also differ across these groups.
 
 There are theoretical reasons to believe that education might be
-associated with either more or less systematic variance attributable to
-intrapersonal change. On one hand, because college graduates are more
-connected to mainstream discourse and elite signals, they might be more
-likely to make durable changes in response to the emergence of new
-information, new issues, or political realignments, while those without
-chronic information display more variance around their baselines without
-their baselines changing much (Zaller 1992). In contrast, it could be
-the case that the stability demonstrated by educated respondents
-reflects the fact that they have already formed durable opinions and are
-relatively closed off to new information. In this perspective, college
-could serve as a formative experience that solidifies some dimensions of
-personal culture for people who attend. But among people who do not
-attend college, later life experiences might prove more important in
-forming or changing personal culture, as these experiences potentially
-provide information that college-educated peers received earlier.
-
-To test these competing propositions, we use the models presented above
-to separately calculate \(\omega\) values for people with at least a
-bachelor's degree and people with less than a bachelor's degree at wave
-1\footnote{A small number of respondents report different highest
-  degrees at wave 1, wave 2, and wave 3. Some of this is due to
-  measurement error, and some of it is due to a small number of people
-  obtaining a higher degree during the four years of the panel.
-  Estimating the panel with highest at wave 1 or highest degree reported
-  across the panel produces functionally identical results.} of the
-General Social Survey's panels. We focus on the GSS because it contains
-the largest number of questions tapping general political dispositions,
-which is the domain where education has proven particularly relevant for
-understanding attitude stability. The GSS also covers a turbulent window
-of American politics from 2006 to 2014. This window covers the start of
-the Great Recession, debates about federal intervention in and
-regulation of Wall Street, the election of Barack Obama as the first
-black president in U.S. history, debates about the role of the federal
-government in the health care sector, the emergence of the Tea Party,
-and political realignment and clarification on the issue of gay
-marriage, among other topics.
+associated with either more or less intrapersonal change. On one hand,
+because college graduates are more connected to mainstream discourse and
+elite signals, they might be more likely to make durable changes in
+response to the emergence of new information, new issues, or political
+realignments, while those without chronic information might display more
+variance around an unchanging baseline (Zaller 1992). Conversely, it
+could be that the observed stability of the college educated reflects
+the fact that they have already formed durable opinions and are
+relatively closed off to new information. If this is true, college could
+be understood as a formative experience that solidifies some dimensions
+of personal culture. Perhaps for those who do not attend college, later
+life experiences might prove more important in forming or changing
+personal culture, as these experiences potentially provide information
+that college-educated peers have already received.
+
+To compare these competing propositions, we calculate \(\omega\) values
+separately for people with at least a bachelor's degree and people with
+less than a bachelor's degree at wave 1\footnote{A small number of
+  respondents report different highest degrees at wave 1, wave 2, and
+  wave 3. Some of this is due to measurement error, and some of it is
+  due to a small number of people obtaining a higher degree during the
+  four years of the panel. Estimating the panel with highest at wave 1
+  or highest degree reported across the panel produces functionally
+  identical results.} of the three General Social Survey's panels. We
+focus on the GSS because it contains the largest number of questions
+tapping general political dispositions, which is the domain where
+education has proven particularly relevant for understanding attitude
+stability. The GSS also covers a turbulent window of American politics
+from 2006 to 2014. This window covers the start of the Great Recession,
+debates about federal intervention in and regulation of Wall Street, the
+election of Barack Obama as the first black U.S. president, debates
+about the role of the federal government in the health care sector, the
+emergence of the Tea Party, and political realignment and clarification
+on the issue of gay marriage, among other topics.
 
@@
 between people with at least a bachelor's degree and people with less
-than a bachelor's degree at wave 1 of the panel for all 183 GSS items we
-explored. Values greater than 0 indicate that intrapersonal change is
-more important in explaining the responses of college graduates than
-those of people without a college degree, while values less than 0
-indicate that intrapersonal change is less important for explaining
-variation among people with a college degree than those without a
-college degree.
+than a bachelor's degree at wave 1 of the panel for 183 GSS items.
+Values greater than 0 indicate that intrapersonal change accounts for
+more systematic variance among college graduates than among those
+without a college degree, while values less than 0 indicate the
+opposite.
 
@@
 
-There is a clear pattern in Figure 3: for more than 80 percent of GSS
-items explored here, intrapersonal change is a larger component of
-systematic variance for people without a college degree than for people
-with a college degree (i.e., the difference is negative). While most of
-these differences are small in absolute terms, less than 2 percentage
-points, several are greater than 5 percentage points. Given the
-distribution observed in Figure 1 showing that the systematic variance
-attributable to intrapersonal change averages around 0.09, a 5
-percentage point difference between groups is quite substantial.
+There is a clear pattern in Figure 3: for more than 80 percent of these
+GSS items, intrapersonal change is a larger component of systematic
+variance for people without a college degree. While most of these
+differences are small in absolute terms (less than 2 percentage points),
+several are greater than 5 percentage points. Given the distribution
+observed in Figure 1 showing that the systematic variance attributable
+to intrapersonal change averages around 0.09, a 5 percentage point
+difference between groups is quite substantial.
 
@@
 about whether black people should be given preferences in hiring,
-measured on a five-point scale. Estimates of \(\omega\) for these eight
-questions, for both education groups, are presented in Figure 4.
+measured on a five-point scale. We present estimates of \(\omega\) for
+these eight questions, for both education groups, in Figure 4.
 
@@
 On all eight questions presented in Figure 4, \(\omega\) values are
-smaller for people with a college degree than for people without a
-college degree, meaning intrapersonal change explains less systematic
-variance for people with a college degree than for people without one.
-There are also large differences in \(\omega\) values across questions
-for both groups. For example, on the question of whether the government
-should try to solve more problems or leave those problems to be solved
-by private businesses (``government do more or less''), less than one
-percent of the systematic variance is attributable to intrapersonal
-change for both groups. In other words, while people might vacillate on
-this question at random (37 percent of variance is residual for this
-question), there is functionally no evidence that people make systematic
-changes of opinion on this issue during the GSS panel.
+smaller for people with a college degree, meaning intrapersonal change
+is less common among them. There are also large differences in
+\(\omega\) values across questions for both groups. For example, on the
+question of whether the government should try to solve more problems or
+leave those problems to be solved by private businesses (``government do
+more or less''), less than one percent of the systematic variance is
+attributable to intrapersonal change for both groups. In other words,
+while people might vacillate on this question at random (37 percent of
+variance is residual for this question), there is functionally no
+evidence that people make systematic changes of opinion on this issue
+during the GSS panel.
 
@@
 larger values of \(\omega\) than most other questions, as well as a
-larger absolute difference between people with a college degree and
-those without. Compared to the other questions explored here, change as
-adults plays a much larger role in explaining why people report the
-partisan identification and political ideology they do. And this is
-particularly true for respondents who do not have a college degree. The
+larger absolute difference by degree status. Compared to the other
+questions, intrapersonal change plays a much larger role in accounting
+for partisan identification and political ideology. And this is
+particularly true for respondents who do not have a college degree; the
 \(\omega\) value for ideological identification among non-college
@@
 
-We note here that this meaningful difference in \(\omega\) values across
-education groups and across questions would not be detectable using
-previous methods. For partisan identification and political ideology,
-both college-educated respondents and people with less than college
-degree would likely favor the AUM or LCAM over the SDM because they
-these questions both show evidence of some members of the population
-making intrapersonal change. In other words, using the tournament of
-models approach obscures the fact that the relative importance of
-intrapersonal change in explaining differences on these two items is
-different across these two groups, that intrapersonal change explains
-more systematic variance for ideological identification than partisan
-identification for people without a college degree, and that there
-appears to be more durable change on questions of affirmative action
-than on questions of government aid to black Americans.
+It is worth pointing out that this meaningful difference in \(\omega\)
+values across education groups and across questions would not have been
+detectable using previous methods. For partisan identification and
+political ideology, both college-educated respondents and people with
+less than college degree would likely favor the AUM or LCAM over the SDM
+because they these questions both show evidence of some members of the
+population making \emph{some} intrapersonal change. In other words, the
+tournament of models obscures the fact that the relative importance of
+intrapersonal change differs across these two groups, that intrapersonal
+change explains more systematic variance for ideological identification
+and partisan identification for people without a college degree, and
+that there appears to be more durable change on questions of affirmative
+action than on questions of government aid to black Americans.
 
 We believe these patterns shed new light on the mechanisms underlying
-differences in attitude behavior across groups. Something about college
-attendance seems to limit the ability of experiences later in life to
-drive changes in personal culture. While these results should not be
-interpreted as causal effects of attending college -- they are
-potentially confounded by age, social class, race, gender, and other
-factors that explain selection into higher education -- they open up a
-set of new questions and dynamics to explore.
-
-\hypertarget{discussion}{%
-\section{Discussion}\label{discussion}}
-
-This paper set out intervene in the debate on whether people change
-their personal culture -- their attitudes, beliefs, values, practices,
-and dispositions -- as they move through their adult life. Instead of
-falling back on a ``tournament of models'' approach (Lersch 2023:
-p.~228) to proclaim victory for only one answer to this question, it
-seeks to refocus the debate to the theoretically more productive
-question of the relative importance of intrapersonal change for
-explaining differences in the personal culture. It proposes a new
-approach that quantifies the amount of observed systematic variance that
-is attributable to either interpersonal differences at baseline or
-intrapersonal change over the duration of a panel.
-
-Applying our proposed measure, \(\omega\), to all questions about
-personal culture from the panel datasets previously discussed in this
-debate revealed a consistent pattern. Nearly all questions show evidence
-of people making durable, intrapersonal change over time, with some
-showing notably high intrapersonal change compared to interpersonal
-differences. For some questions about factors that affect one's life
+differences in attitudes and behavior across groups. Something about
+college attendance seems to crystallize personal culture. While these
+results should not be interpreted as causal effects of attending college
+-- they are potentially confounded by age, social class, race, gender,
+and other factors that explain selection into higher education -- they
+open up a set of new questions and dynamics to explore.
+
+\hypertarget{discussion-and-conclusion}{%
+\section{Discussion and Conclusion}\label{discussion-and-conclusion}}
+
+The objective of this paper is to intervene in the recent debate on
+whether people change their personal culture -- their attitudes,
+beliefs, values, practices, and dispositions -- as they move through
+their adult life. Instead of falling back on a ``tournament of models''
+approach (Lersch 2023: p.~228) to proclaim victory for a particular
+answer to this question, our goal is to refocus the debate to the
+theoretically more productive question of the relative importance of
+intrapersonal change for explaining differences in the personal culture.
+We propose a new approach that quantifies the amount of observed
+systematic variance that is attributable to either interpersonal
+differences at baseline or intrapersonal change over the duration of a
+panel.
+
+Applying our proposed measure, \(\omega\), to 609 survey items from all
+of the panel datasets previously discussed in this debate revealed a
+consistent pattern. Nearly all questions show evidence of people making
+some durable, intrapersonal change over time. Some show notably high
+amounts of intrapersonal change. For some questions about life
 satisfaction or views on government benefits, it seems plausible that
-differences in adult experiences predominantly account for the observed
+differences in adult experiences predominantly account for observed
 differences between people.
@@
 change is essentially zero. For these questions, it seems there is not
-enough cultural change during adulthood to warrant attributing
-differences in personal culture mainly to experiences and social
-transitions; instead, the primary source of the observed differences
-appears to stem from experiences in childhood, adolescence, and early
-adulthood.
-
-More than adjudicating past positions in this debate, the approach
-presented in this paper enables scholars to more clearly explore the
-relative contributions of change and stability in explaining variation
-in personal culture, as well as how these contributions might differ
-across populations. To exemplify this, we detailed how the amount of
-cultural difference explained by intrapersonal change varies
+enough cultural change during adulthood to warrant attributing the
+differences we observe to experiences and social transitions; instead,
+the primary source of the observed differences appears to stem from
+experiences in childhood, adolescence, and early adulthood.
+
+More than adjudicating past positions in this debate, the approach we
+presented in this paper will enable scholars to explore more precisely
+the relative contributions of change and stability in explaining
+variation in personal culture, as well as how these contributions might
+differ across populations. To illustrate this, we showed how the amount
+of cultural difference explained by intrapersonal change varies
 substantially by survey item but also by individuals' characteristics
-such as their education. Intrapersonal change plays almost no part in
-explaining sentiments regarding the government's economic role, but it
-is more defining for ideological and partisan identities. Intriguingly,
-it is more important in explaining cultural variation among non-college
-graduates compared to college graduates in the United States. While age
-and other factors related to college completion might confound this
-pattern, it suggests that college completion may solidify personal
-culture to an extent that renders the effect of other adult experiences
-on political dispositions less influential for college graduates
-relative to non-college graduates.
+such as their education. While age and other factors related to college
+completion might confound this pattern, it suggests that college
+completion may crystallize personal culture to an extent that renders
+later adult experiences on political dispositions less influential.
+
+\hypertarget{limitations}{%
+\subsection{Limitations}\label{limitations}}
+
+Although we believe it is a major advance on previous work, the approach
+we outline here has important limitations. We allocate all systematic
+variance to one of two sets of theoretical processes: intrapersonal
+change and interpersonal differences at baseline. Our approach does not
+quantify the proportion of people who ``change,'' nor can we be sure
+that the amount of interpersonal change we detect is driven by many
+people making small changes or a few people making large changes.
+
+Our results and interpretations also hinge on how we have defined
+change. As does the LCAM, our approach treats cultural trajectories as
+varying linear slopes for each respondent, thus assuming that change is
+a linear function of time. This assumption simplifies reality in which
+change likely also takes non-linear and discontinuous forms. People
+might jump from one ``stable'' disposition to another or experience a
+``turning point'' in their lives that upsets an otherwise stable
+trajectory. In parts, our assumption of linearity is a limitation of the
+data, as most questions are only observed for three to six waves. Panels
+with more waves might allow researchers to loosen this assumption to
+test alternative, more flexible models of change.
+
+Similarly, our approach assumes that durable change is unidirectional.
+This is a sensible assumption on short panels where classifying change
+that lasts less than two years as durable seems unreasonable.
+Practically, it means that the variance produced by people making
+durable changes in their cultural dispositions to then return to a
+previous state later in life is classified as residual variance, rather
+than intrapersonal change. Longer panels and more flexible definitions
+of ``change'' might allow us to account for such trajectories.
+
+Finally, we had chosen to examine the broadest array of measures of
+personal culture available to us, ranging from religious beliefs and
+core values, to policy preferences, and even to the importance of
+different features when buying a new car across five countries
+(Australia, Germany, Switzerland, the United Kingdom and the United
+States). Nevertheless, our findings remain limited to the kinds of
+questions that are asked in panel surveys and in the contexts they were
+administered, reflecting issues of general (national) politics, gender
+roles, immigration and race relations, and general well being. Although
+we have no reason to believe results to be different, our findings do
+not directly speak to other dimensions of culture such as artistic
+tastes, leisure activities, and time use.
 
@@
 
-Sociologists interested in understanding cultural differences have
-largely asked about the presence and significance of cultural change in
-adults (e.g., Kiley and Vaisey 2020; Lersch 2023; Vaisey and Kiley
-2021). Yet doing so has unnecessarily limited and might even set the
-wrong focus. In any population, some degree of cultural change in adults
-is inevitable. Likewise, no viable theoretical perspective would expect
-just one process -- stable differences or internal change -- to shape
-their personal culture. Our results reinforce this point, showing the
-relevance of both interpersonal differences and intrapersonal change in
-understanding the differences in personal culture among adults across
-all items analyzed.
+Despite these limitations, we believe our method and findings have
+important implications for social science research. Sociologists
+interested in understanding cultural differences have largely asked
+about the \emph{existence} of cultural change in adults (e.g., Kiley and
+Vaisey 2020; Lersch 2023; Vaisey and Kiley 2021). Yet this approach has
+inadvertently limited the debate. In any population, \emph{some} degree
+of adult cultural change is inevitable. Although ideal types like the
+settled dispositions and active updating models are useful, no
+theoretical perspective would expect either early life socialization or
+adult intrapersonal change to be the sole source of one's personal
+culture. Our results reinforce this point, showing the relevance of both
+factors and allowing their precise quantification.
 
 This and other recent findings (Quinn et al. 2023; Stewart and Berkman
-2023) suggest that it is theoretically more productive to ask about the
-relative importance of these two processes in understanding cultural
-variation. Drawing a unified conclusion from survey items measuring
-various cultural forms on different scales and across different time
-frames is challenging. However, the general pattern suggests that, for
-most items, intrapersonal change in adulthood is not the primary reason
-for cultural differences among people.
-
-To clarify, our findings do not imply that adults remain static or that
-their changes are inconsequential. Even minor shifts in personal
-culture, such as a 2 percentage-point change in support for gay
-marriage, can have significant ramifications. While the majority may
-remain consistent in their views, understanding the underlying
-mechanisms of even such minor shifts remains a crucial task for the
-sociology of culture.
-
-Our findings primarily suggest that understanding variations in personal
-culture requires scholars to examine the conditions and experiences of
-early life. While transitions between social roles, changes in social
-networks, and the experience of organizational environments are often
-emphasized in sociological research, these factors seem to account for a
-smaller proportion of differences among adults than the formative
-experiences early in life.
+2023) suggest that it is theoretically more productive to measure the
+relative importance of these two components in concrete, substantive
+cases. Drawing a unified conclusion from survey items measuring various
+cultural forms on different scales and across different time frames is
+challenging. Nevertheless, the general pattern suggests that, for most
+items, intrapersonal change in adulthood is not the primary reason we
+see the differences we see between people in the world.
+
+We are not claiming that adults remain static or that their changes are
+inconsequential. Even quantitatively minor shifts, such as a 2
+percentage-point change in support for gay marriage, can have massive
+ramifications. While the majority may remain consistent in their views,
+understanding the underlying mechanisms of even such numerically minor
+shifts remains a crucial task for the sociology of culture.
+
+Our findings suggest that understanding variation in personal culture
+requires examining the conditions and experiences of early life. While
+sociological research often focuses on transitions between social roles,
+changes in social networks, and the experience of organizational
+environments, these factors seem to account for a smaller proportion of
+adult differences than early-life experience. Simply put, we need more
+work on early-life socialization (Guhin, Calarco, and Miller-Idriss
+2020).
 
@@
 Our findings also have implications for understanding cultural change at
-the aggregate level. Given that intrapersonal change does not appear as
-the primary driver for many cultural differences among adults, our
-results suggest that many forms of cultural change at the aggregate
-level are more likely driven by processes of cohort replacement than the
-by contemporaneous social conditions changing people's cultural
-dispositions (Underwoood et al. 2022; Vaisey and Lizardo 2016). This
-likely holds true for cultural change at the macro societal level as
-well as the micro level such as within organizations, political parties,
-and professions. Our findings also indicate that the impact of formative
-experiences, as opposed to contemporaneous social factors, on cultural
-change is contingent upon an individuals' level of education. The weight
-of these two factors might thus change as the prominence of education
-shifts in the life course. If scholars find similar differences across
-other social categorizations like class or race, it would advocate for a
-more thorough integration of demographic processes into the analysis of
-cultural change at the aggregate level.
+the aggregate level. Given that intrapersonal change accounts for a
+relatively small amount of variance among adults, many forms of cultural
+change at the aggregate level are necessarily more likely driven cohort
+replacement than by contemporaneous social conditions (Underwoood et al.
+2022; Vaisey and Lizardo 2016). This likely holds true for cultural
+change at the macro societal level as well as the micro level such as
+within organizations, political parties, and professions. Our findings
+also indicate that the impact of formative experiences, as opposed to
+contemporaneous social factors, on cultural change is contingent upon an
+individuals' level of education. The weight of these two factors might
+thus change as the prominence of education shifts in the life course. If
+scholars find similar differences across other social categorizations
+like class or race, it might necessitate a more thorough integration of
+demographic processes into the analysis of cultural change at the
+population level.
 
@@
 
-Further, we align with others {[}\textbf{WE NEED SOME CITES HERE}{]}
-advocating for more studies, including panel studies, focusing on the
-socialization of youth. Our analysis points to the majority of variance
-among adults rooting in differences encountered before turning 18. By
-empirically constraining their analysis to experiences after that point,
-social scientists may inadvertently concentrate on topics and questions
-that, while important, might not be able to help explain major cultural
-differences in adult populations.
-
-Finally, a key insight from our meta-analysis is that an increase in
-response resolution (the range of response options survey respondents
-are given) correlates with an increased share of systematic variance
-explained by intrapersonal change. This may reflect specific issues
-being assessed by survey items using different scales. Alternatively, it
-might imply that when intrapersonal changes occurs, they are often
-subtle with answers moving from ``agree'' to ``strongly agree,'' rather
-than from ``strongly agree'' to ``strongly disagree.'' Such subtle
-changes require finer response options to be detected.
-
-\hypertarget{limitations}{%
-\subsection{Limitations}\label{limitations}}
-
-The approach outlined here generates estimates of how much of the
-observed variance in a population can be attributed to one of two sets
-of theoretical processes: intrapersonal change and interpersonal
-differences at baseline. It does not quantify the proportion of people
-who ``change,'' nor can we be sure that the amount of interpersonal
-change we detect is driven by many people making small changes or a few
-people making large changes.
-
-Our results and interpretations also hinge on how we have defined
-change. Similarly to the LCAM, our approach treats cultural trajectories
-as varying linear slopes for each respondent, thus assuming that change
-is a linear function of time. This assumption simplifies reality in
-which change likely also takes non-linear and discontinuous forms.
-People might jump from one ``stable'' disposition to another or
-experience a ``turning point'' in their lives that upsets an otherwise
-stable trajectory. In parts, our assumption of linearity is a limitation
-of the data, as most questions are only observed for three to six waves.
-Panels with more waves might allow researchers to loosen this assumption
-to test alternative, more flexible models of change.
-
-Similarly, our approach assumes that durable change is unidirectional.
-This is a sensible assumption on short panels where classifying change
-that lasts less than two years as durable seems unreasonable.
-Practically, it means that the variance produced by people making
-durable changes in their cultural dispositions to then return to a
-previous state later in life is classified as residual variance, rather
-than intrapersonal change. Longer panels and more flexible definitions
-of ``change'' would allow us to account for such trajectories and thus
-potentially also identify more durable change even in the same
-population.
-
-Finally, we had chosen to examine the broadest array of measures of
-personal culture available to us, ranging from religious beliefs and
-core values, to policy preferences, and even to the importance of
-different features when buying a new car across five countries
-(Australia, Germany, Switzerland, the United Kingdom and the United
-States). Nevertheless, our findings remain limited to the kinds of
-questions that are asked in panel surveys and in the contexts they were
-administered, reflecting issues of general (national) politics, gender
-roles, immigration and race relations, and general well being. Although
-we have no reason to believe results to be different, our findings do
-not directly speak to other dimensions of culture such as artistic
-tastes, leisure activities, and time use.
-
-\hypertarget{implications-for-future-research}{%
-\subsection{Implications for Future
-Research}\label{implications-for-future-research}}
-
-We believe the approach outlined here can push past the ``needless
-dichotomy'' implicit in the question of whether people change or not and
-address more theoretically informed questions about the conditions under
-which the relative importance of intrapersonal change and interpersonal
-differences in explaining cultural differences shifts. Characterizing
-questions as displaying change or not can only take researchers so far,
-but the question of whether some questions demonstrate more change than
-others, or whether some groups are characterized by more stability than
-others, has the potential to weigh in on a broader range of theoretical
-debates. We hope researchers move toward those questions.
-
-While we focused on education as a determinant of the explanatory power
-of intrapersonal change, our approach can readily be extended to other
-socio-demographic categories that might affect intrapersonal change --
-race/ethnicity, religiosity, social class, gender, and political
-affiliation, for example -- or their intersections. Researchers with
-different data might be able to evaluate the relative importance of
-intrapersonal change and interpersonal differences across contexts as
-well, either national contexts if the same questions are administered
-across countries, or organizational contexts if data traces students
-within schools, workers within workplaces, or congregation members
-within congregations.
-
-The wide variation across questions in the proportion of systematic
-variation attributable to intrapersonal change also calls for a more
-systematic exploration for why some forms of culture are characterized
-by stability while others are characterized by change, and what the
-consequences of these differences for the organization of beliefs in the
-public might be.
+A key insight from our meta-analysis is that an increase in response
+resolution (the range of response options survey respondents are given)
+correlates with an increased share of systematic variance explained by
+intrapersonal change. This may reflect specific issues being assessed by
+survey items using different scales. Alternatively, it might imply that
+when intrapersonal changes occurs, they are often subtle with answers
+moving from ``agree'' to ``strongly agree,'' rather than from ``strongly
+agree'' to ``strongly disagree.'' Such subtle changes require finer
+response options to be detected.
+
+Perhaps most important, we need more panel studies on youth. It is very
+likely that the majority of adult differences in personal culture are
+rooted in different experiences before age 18. By empirically
+constraining our analyses to adult experiences, social scientists may
+inadvertently concentrate on topics and questions that, while important,
+might not be able to help explain major cultural differences in adult
+populations.
+
+\hypertarget{conclusion}{%
+\subsection{Conclusion}\label{conclusion}}
+
+We believe the approach we outlined here can push past the ``needless
+dichotomy'' implicit in the question of whether people change or not
+(Lersch 2023). Characterizing questions as displaying change or not can
+only take researchers so far, but the question of whether some questions
+demonstrate more change than others, or whether some groups are
+characterized by more stability than others, has the potential to weigh
+in on a broader range of theoretical debates. We hope researchers find
+our approach useful as they investigate these questions.
 
@@
 
+\leavevmode\vadjust pre{\hypertarget{ref-guhin2020}{}}%
+Guhin, Jeffrey, Jessica McCrory Calarco, and Cynthia Miller-Idriss.
+2020. \emph{Whatever {Happened} to {Socialization}?} \emph{preprint}.
+SocArXiv. doi:
+\href{https://doi.org/10.31235/osf.io/zp2wy}{10.31235/osf.io/zp2wy}.
+
 \leavevmode\vadjust pre{\hypertarget{ref-hout2016}{}}%

### 2023-09-29 manuscript/manuscript.Rmd
subsection title (change/stability) was inconsistent with paper title
more SV edits

@@
 csl: asa.csl
-abstract: "Recent work has produced conflicting interpretations on the question of whether adults undergo durable changes in their personal culture over time.
-This paper asserts that this divergence is less due to researchers' findings than the approach taken -- a tournament of models focusing on the presence or absence of change.
-To move beyond this, we reanalyze the data used previously in this debate with an approach that quantifies the processes of cultural difference.
-In doing so, we compare how much cultural difference is explained by intrapersonal change and interpersonal differences.
-Our results harmonize recent findings by showing that, while most measures of personal culture show durable change over time, the variance explained by intrapersonal change is typically small compared to that of interpersonal differences at baseline.
-Quantifying these processes offers new ways to address theoretical questions of when and how their relative importance shifts.
-We exemplify this approach by exploring how the importance of these processes differs between college graduates and the rest of the population."
+abstract: "Does personal culture---a person's attitudes, beliefs, values, and practices---change over the life course, or is it largely fixed by adulthood?
+Recent work has reached substantially different conclusions on this question.
+We argue that this disagreement is an unintended consequence of the \"tournament of models\" approach researchers have used.
+We introduce a new measure for quantifying the proportion of systematic variance in panel data attributable to intrapersonal change versus stable interpersonal differences.
+Applying this measure to 609 items from seven surveys in five countries, we find that although intrapersonal change is common, it is generally small in magnitude.
+We also demonstrate that this measure is useful for comparing social groups by showing that intrapersonal change is less common among U.S.
+college graduates than among those without a college degree.
+Overall, we believe this new measure will be a useful methodological tool and help reconcile an important theoretical debate."
 ---
@@
 
-Does personal culture---an individual's attitudes, beliefs, values, and practices---change over the life course, or is it largely fixed by adulthood?
+Does personal culture---a person's attitudes, beliefs, values, and practices---change over the life course, or is it largely fixed by adulthood?
 This question underlies an important contemporary debate in sociology [@lizardo2017; @kiley2020; @lersch2023] and has deep roots in seemingly contradictory theoretical perspectives.
@@
 
-Because all of these theoretical perspectives have some empirical support, the current debate is not about whether their posited processes exist at all, but about their relative contribution to explaining the cultural differences we see in the world.
-That said, scholars have struggled to reach a consensus on the importance of change during adulthood.
-Over the span of a few years, most apparent changes in people's responses to personal survey items appear to be transitory, with little evidence of persistent change among adults [@kiley2020; @vaisey2021].
-This suggests that change during adulthood is not a major factor in explaining personal cultural differences.
-When we consider a longer time horizon, however, there is evidence that adults make at least some persistent changes [@lersch2023].
+Because all of these perspectives have some empirical support, the current debate is not about whether the processes they posit exist at all, but about their relative contribution to explaining the cultural differences we see in the world.
+Nevertheless, researchers have struggled to reach a consensus on the importance of cultural change during adulthood.
+Over the span of a few years, most apparent changes in adults' survey responses appear to be transitory, with little evidence of persistent change [@kiley2020; @vaisey2021].
+This suggests that change during adulthood may not be a major factor in explaining contemporary cultural divides.
+By contrast, when we consider a longer time horizon, there is evidence that adults make at least some persistent changes [@lersch2023].
 
@@
 228].
-That is, researchers have compared the fit of different statistical models to each item to determine whether it is more consistent with "no change" or "any change." The size of these piles then serves as the primary evidence for the truth of the theory [@kiley2020; @vaisey2021; @lersch2023].
+That is, researchers have relied on model selection criteria---primarily the Bayesian Information Criterion [@raftery1995]---to sort variables into the "no change" or "any change" piles.
+The size of these piles then serves as the primary evidence for the truth of the theory [@kiley2020; @vaisey2021; @lersch2023].
 
-We argue, however, that asking *whether* people change on an item may cause more problems than it solves.
-Being able to detect *any* change on an item depends not only on sample populations, measurement, statistical methods, and power, but also on definitions of what counts as "change." These differences can limit the possibility of agreement on which "pile" a variable belongs.
-Moreover, asking whether there is evidence of change reduces to, "Is there *any* evidence of change in this item?" But such binary inquiries are ill-suited to address the core of the theoretical debate: determining *how much* the process of intrapersonal change contributes to cultural differences.
+We argue, however, that asking *whether* people change on an item is an unintended source of confusion.
+Being able to detect *any* change on an item depends not only on sample populations, measurement, statistical methods, and power, but also on definitions of what counts as "change." These differences can limit the possibility of agreement on the assignment of a variable to the right pile.
+Moreover, asking whether there is evidence of change reduces to, "Is there *any* evidence of change in this item?" But such binary questions are ill-suited to address the core of the theoretical debate: determining *how much* the process of intrapersonal change contributes to cultural differences.
 
-There is enough evidence to assert that neither pre-adult socialization nor the contemporary social context alone can explain the full range of cultural differences.
-The question that remains is about the *relative* contributions of these processes.
-In the current debate, for a single construct (e.g., support for gay rights), existing approaches can only make claims about *whether* intrapersonal change is happening, not about how much.
-This also prevents investigating under what conditions, across which groups, and in what domains do we see differences in the relative importance of intrapersonal change and pre-existing interpersonal differences.
-These questions are about *degree*, not existence; to take this debate forward therefore requires precise measurement rather than declarations of victory for a particular perspective.
+Of course neither pre-adult socialization nor the contemporary social context alone can explain the full range of cultural differences.
+The relevant question is about the *relative* contributions of these processes.
+In the current debate, for a single construct (e.g., support for gay rights), model comparison approaches can only help decide *whether* intrapersonal change is happening, not about how much.
+This also prevents investigating under what conditions, across which groups, and in what domains we might see differences in the the amount of intrapersonal change.
+These questions are about *degree*, not existence.
+Taking this debate forward therefore requires precise measurement rather than declarations of victory for a particular perspective.
 
-In this paper, we move beyond the "tournament of models" approach and introduce a method for quantifying the relative contributions of interpersonal differences and intrapersonal change for a single item.
-Drawing on seven panel data sets from five countries, we quantify the amount of variance in survey responses that is due to between-person differences at a single point during the survey versus the amount of variance in responses that is due to systematic intrapersonal change over the study period.
-This measure quantifies the amount of change over the duration of the panel relative to the stable differences that exist between people.
+In this paper, we move beyond the "tournament of models" approach and introduce a new method for quantifying the relative contributions of interpersonal differences and intrapersonal change for a single item.
+We use 609 items from seven panel surveys from five countries to quantify the proportion of variance in that is attributable to systematic intrapersonal change over the study period rather than to stable interpersonal differences.
 
-Using this measure, we observe similar patterns across the datasets that have previously yielded conflicting interpretations.
-In all datasets---despite their varied social contexts and duration---we find a similar distribution of intrapersonal change.
-In general, we find that intrapersonal change accounts for a substantially smaller amount of variance in personal culture survey item than stable interpersonal differences.
-Of course, whether the amount of intrapersonal change in an item is large enough to be "meaningful" is a substantive and theoretical question.
-Nevertheless, on many measures of personal culture (some observed for over a decade), we do not observe enough intrapersonal change for it to play a substantial role in explaining the differences we observe between adults.
-Notable, interesting, and important exceptions exist, of course, but they are departures from the overall pattern.
+We observe similar results across all datasets despite their varied social contexts and duration.
+In general, we find that intrapersonal change accounts for a small amount of variance in personal culture survey items.
+Of course, any question about whether a given amount of intrapersonal change is "meaningful" or "important" is a substantive and theoretical.
+Nevertheless, on many measures (some observed for over a decade), we do not see enough change to believe that intrapersonal change processes play a substantial role in explaining the differences we observe between adults.
+Notable, interesting, and important exceptions exist, but they are departures from the overall pattern.
 
-In addition to advancing a theoretical debate, our method can also help address substantive questions.
+In addition to advancing a theoretical debate, our method can also help address specific substantive questions.
 To illustrate how it might be used by a researcher with a specific topical interest, we investigate differences in the relative importance of intrapersonal change for people with and without a college degree.
-We find a smaller amount of systematic change among college-educated respondents, suggesting that college solidifies one's personal culture rather than fostering an openness to new information.
-This offers a novel empirical basis to theorize about the role of university education in shaping personal culture.
+We find a smaller amount of systematic change among college-educated respondents, suggesting that college crystallizes one's personal culture rather than fostering an openness to new information.
+This offers a new empirical basis to theorize about the role of education in influencing personal culture.
 
-Taken together, the goal of this paper is to advance---and, hopefully, transcend---debates about *whether* adults change.
+Taken together, our goal in this paper is to advance---and, hopefully, transcend---debates about *whether* adults change.
 People do change, at least a little, on most things.
@@
 
-## Change and Stability in Personal Culture
+## Stability and Change in Personal Culture
 
@@
 There is no reason to believe that only one of these two models is "correct" at all times and in all places.
-A population observed over some period of time contains a mix of people who are noticeably changing and people who are not.
+A population observed over some period of time contains a mix of people who are changing and people who are not.
 Instead, different perspectives argue that each of these ideal-typical models is more operative at different times, for different people, and for different elements of personal culture.
@@
 Empirically comparing which of these models better fit a broad range of questions from the General Social Survey's rotating panels, Kiley and Vaisey [-@kiley2020] found limited evidence of durable change.
-While the majority of items preferred the AUM, the amount of durable change detected on these items was small.
-A substantial minority of questions (39 percent) favored the SDM, meaning they showed no evidence of durable change.
+While the majority of items did prefer the AUM, the amount of durable change detected on these items was small.
+A substantial minority of questions (39 percent) favored the SDM, meaning they were more consistent with zero durable change.
 Questions with more evidence of durable change included salient issues like gay marriage and questions tapping "public" statements or behaviors such as partisan identification and religious service attendance.
-There was also more evidence of durable change among early adults (people ages 18-30) than the rest of the population.
+There was also more evidence of durable change among early adults (people ages 18-30) than among the rest of the population.
 Overall, the researchers concluded that "results ultimately suggest that real, persistent attitude change is an uncommon phenomenon among adults'' [@kiley2020: p.
@@
 
-Second, Kiley and Vaisey's analysis of the AUM and SDM is based on three-wave panel data over four years, which might not be extensive enough to adjudicate the two models.
+Second, Kiley and Vaisey's analysis of the AUM and SDM is based on three-wave panel data over four years, which might not be intensive or extensive enough to adjudicate the two models.
 Lersch evaluates the LCAM against the AUM and SDM using panel data spanning a wider duration (from 3 to 36 years) and more waves (3 to 18), offering a better chance to observe durable change.
@@
 Kiley and Vaisey [-@kiley2020] found that the majority of items they tested favored the AUM, meaning there was evidence of durable change among adults, a pattern consistent with Lersch's results.
-Also Lersch [-@lersch2023] found evidence of durable change on most items, but this change only accounted for a small proportion of variance explained.
+Also Lersch [-@lersch2023] found evidence of durable change on most items, but these changes were small.
 On average, people changed only about .07 standard deviations over 10 years.
 This is consistent with Kiley and Vaisey's finding that durable change is often small.
-Moreover, a quarter of the items Lersch studied still preferred the SDM, suggesting that even with different assumptions assumptions about change and more extensive data, many questions show no evidence of durable change.
-Consistent with this, in studies showing evidence of durable change in specific measures of personal culture, the change is often small.
+Moreover, a quarter of the items Lersch studied still preferred the SDM, suggesting that even with different assumptions about change and more extensive data, many questions still are consistent with a model that allows zero intrapersonal change.
+
+Other studies are compatible with these findings as well.
 For example, Brocic and Miles [-@brocic2021] estimate that completing graduate degrees in humanities, arts, and social sciences shift peoples' moral relativism only about .2 standard deviations on average compared to people with no college degree, and this is the largest effect they identify.
@@
 
-The current debate can be simplified to asking whether people *ever* change.
-This is rarely theoretically interesting since the answer is almost always "yes." Over sufficient time and in a large enough sample, researchers will likely observe some evidence for durable individual change.
+To this point, the theoretical debate has been about whether people *ever* change.
+As we argued above, this is rarely the right question since the answer is almost always "yes." Over sufficient time and in a large enough sample, researchers will likely observe some evidence for durable individual change.
 A lack of evidence for change may be due to a poor survey question, a low resolution in response options, or that the question simply was not asked for long enough or to enough people.
-Conversely, finding evidence for some change tells us only that it was "not zero." From that alone, we learn little about how much change has happened in a population.
+Conversely, finding evidence for some change tells us only that it was "not zero." From that alone, we learn little about how much intrapersonal change has happened in a population.
 
@@
 Lersch's LCAM does this by modeling each individual as following a personal linear trajectory.
-But rather than debating whether this model fits a particular data set better than a model that assumes that people never change, researchers should quantify these individual changes and compare those to the differences that exist between people at the panel's start.
-In other words, the focus should shift from asking, "do people change?" to asking, "what are the relative contributions of change and initial differences for explaining variance in personal culture?"
+But rather than debating whether this model fits a particular data set better than a model that assumes that people never change, it would be better to quantify these individual changes and compare them to the stable differences that exist between people.
+In other words, the focus should shift from asking, "do people change?" to asking, "what are the relative contributions of change and pre-existing differences for explaining variance in personal culture?"
 
-We propose an approach that offers such a quantification.
-Our measure relates two metrics that each reflect a combination of relevant theoretical processes.
-The first metric, interpersonal differences at the start of the survey, reflects the accumulated experiences of people prior to entering the panel survey that shape how they answer questions.
-Lersch calls this "early imprinting." While commonly associated with experiences during a formative period that result in settled dispositions, this variation could also reflect experiences that happened at any stage in life as long as they predate the panel and consistently affect subsequent responses.
+We propose an approach---which we formalize below---that offers such a quantification.
+Our measure relates two variance components that each reflect a combination of relevant theoretical processes.
+The first component, stable interpersonal differences, reflects the accumulated experiences of people prior to entering the panel survey.
+Lersch [-@lersch2023: 24] calls this "early imprinting." While commonly associated with experiences during a formative period that result in settled dispositions, this variation could also reflect experiences that happened at any time as long as they predate the panel and consistently affect subsequent responses.
 For instance, for those people entering the panel post-retirement, this "imprinting" might reflect this pivotal life transition.
-Consequently, this metric also reflects variation in individuals' social roles or statuses at the start of the panel that were important in shaping their dispositions.
+Consequently, this component also reflects variation in individuals' social roles or statuses at the start of the panel that were important in shaping their dispositions.
 
-The second metric, the amount of intrapersonal linear change people make during the panel, captures durable changes in personal culture over time.
+The second component, the amount of intrapersonal linear change people make during the panel, captures durable changes in personal culture over time.
 This captures the set of processes collectively called "persistent change" or "adaption" by Lersch and "active updating" by Kiley and Vaisey.
-Lersch attributes these changes to social triggers such as moving into a new environment or adopting new social roles (although he never measures these directly).
-Additionally, they might reflect the diffusion of new cultural forms across social networks, cues from political elites or otherwise culturally influential leaders, the emergence of issues in politics or culture, or large-scale social shifts.
+Lersch attributes these changes to social triggers such as moving into a new environment or adopting new social roles (although he does not measure these directly).
+They might also reflect the diffusion of new cultural forms across social networks, cues from political elites or otherwise culturally influential leaders, the emergence of issues in politics or culture, or large-scale social shifts.
 
-A third metric, which we might call "residual variance" or "fluctuation," accounts for the remainder of variance in peoples' responses.
+A third component, which we might call "residual variance" or "fluctuation," accounts for the remainder of variance in peoples' responses.
 These non-durable changes emerge for a variety of reasons.
@@
 This variance can also include measurement error, such as misinterpreted questions, erroneous response selections, or responses getting coded incorrectly.
-Although this third metric at least partially reflects important processes of personal culture, it does not directly touch on the ongoing debate outlined here.
+Although this third component at least partially reflects important processes of personal culture, it does not directly touch on the ongoing debate outlined here.
 Because of this, we quantify this component but focus principally on the other two.
 
-At a theoretical level, the coexistence of the sets of processes captured in the two main metrics, interpersonal differences and intrapersonal change, is undeniable, and they are linked in many ways.
+At a theoretical level, the coexistence of the sets of processes captured in the two main components, interpersonal differences and intrapersonal change, is undeniable, and they are linked in many ways.
 Any intrapersonal change during one's life will likely manifest as interpersonal differences by the time people enter a panel survey.
@@
 
-Furthermore, these metrics are not necessarily useful in isolation or without context.
+Furthermore, these components are not necessarily useful in isolation or without context.
 What is important for the theoretical debate is the relative contribution of these two components in explaining cultural differences in a population.
-In other words, when we look at a population of adults over some time frame, is there enough intrapersonal change relative to interpersonal difference to view it as an important explanation for why people differ from each other as observed?
-Only through quantifying the two processes we can get closer to a true answer to this question.
+In other words, when we look at a population of adults over some time frame, is there enough intrapersonal change relative to interpersonal difference to view it as an important process for why people differ from each other?
+Only through quantifying the two components we can get closer to a true answer to this question.
 
-Yet the true utility of the proposed quantification goes beyond settling past debates; it allows researchers to ask questions that were difficult to explore with previous approaches.
-For example, classifying survey items based on whether they show durable change overlooks possibly important differences among those that do.
+Yet the true utility of the proposed quantification goes beyond settling past debates; it allows researchers to ask new questions.
+For example, classifying survey items based on whether they show durable change overlooks possibly important differences among those that do show such change.
 By quantifying the relative contributions of interpersonal differences and intrapersonal change, researchers can gauge their relative importance in explaining differences among survey items, between groups within a given population, across time, and across societies.
@@
 
-By shifting the question and adjusting the approach, we expect consistent patterns across datasets previously analysed in the debate.
+By shifting the question and adjusting the approach, we expect consistent patterns across datasets.
 First, because previous work found evidence of durable change on most measures of personal culture, we expect that we will find some intrapersonal change in most items.
-But because these earlier studies also found many questions (between 25 and 35 percent) lacking evidence of *any* durable change, we expect considerable variation in the variance accounted for by intrapersonal change, with some questions indicating essentially no variance attributable to such change.
+But because these earlier studies also found many questions (between 25 and 35 percent) lacking evidence of *any* durable change, we expect considerable variation in the proportion of variance accounted for by intrapersonal change, with some questions indicating essentially none at all.
 
-Second, we expect the variance attributable to intrapersonal change to be modest compared to interpersonal differences at the start of the survey.
+Second, we expect the variance attributable to intrapersonal change to be modest compared to pre-existing interpersonal differences.
 This expectation should be uncontroversial, regardless of the theoretical process assumed to be most relevant for a particular question.
@@
 Our expectations so far largely echo what was observed in previous studies.
-However, theoretical perspectives differ in how much variance they would expect to be driven by intrapersonal change, enabling quantification to provide a new perspective on existing questions.
-"Formative period" and cohort theories, which emphasize early-life socialization as a strong influence on adult personal culture, suggest that the variance from intrapersonal change would still fall short of explaining interpersonal differences, even when we account for the duration of observation.
-Conversely, other theoretical perspectives emphasizing the importance of people's contemporaneous social environment for personal culture might expect intrapersonal change to account for a large portion of the variance, regardless of the duration.
+However, theoretical perspectives differ in how much change they expect, enabling quantification to provide a new perspective on existing questions.
+"Formative period" and cohort theories, which emphasize early-life socialization as a strong influence on adult personal culture, imply low amounts of intrapersonal change, even when we account for the duration of observation.
+Conversely, other theoretical perspectives emphasizing the importance of people's contemporaneous social environment might expect a larger degree of intrapersonal change, even in a relatively short panel.
 
-Third, we expect the variance explained by each process ---interpersonal differences and intrapersonal change--- to depend on specific attributes of the question and panel.
-First, the higher the response resolution for a given question and the larger the sample, the greater the variance explained by intrapersonal change should be, as smaller changes in intrapersonal change will be easier to detect.
-Second, the LCAM and various theoretical perspectives that underlie it posit that social transitions drive changes in personal culture.
+Third, we expect the variance explained by each process---interpersonal differences and intrapersonal change---to depend on specific attributes of the question and panel.
+First, the higher the response resolution for a given question and the larger the sample, the greater the variance attributed to intrapersonal change should be, as smaller changes in intrapersonal change will be easier to detect.
+Second, life course theories posit that social transitions drive changes in personal culture.
 Since the probability that any person experiences such a transition goes up the longer we observe them, we expect panel duration to be positively related to the proportion of variation explained by intrapersonal change.
@@
 
-We use data from seven nationally representative panel surveys from Australia, Germany, Great Britain, Switzerland, and the United States (summarized in Table 1), combining all the data files used in previous work [@kiley2020; @lersch2023].[^1] These studies cover a long period of time (the range of surveys spans from 1968 to 2021), with 609 personal culture items that capture attitudes, beliefs, values, self-assessments, self-descriptions, and behaviors [@alwin2007].
+We use data from seven nationally representative panel surveys from Australia, Germany, Great Britain, Switzerland, and the United States (summarized in Table 1), combining all the data files used in previous work [@kiley2020; @lersch2023].[^1] These studies cover a long period of time (the range of surveys spans from 1968 to 2021), with 609 personal culture items capturing attitudes, beliefs, values, self-assessments, self-descriptions, and behaviors [@alwin2007].
 We restricted the sample such that individuals between the ages of 18 and 79 are included without further elimination, and in all surveys, we used all possible cases for which respondents provided responses.
@@
 
-We start with Lersch's [-@lersch2023] Life Course Adaption Model, which formalizes survey responses at time $t$ as a function of individual-level random intercepts and slopes for survey age, what is often called a mixed-effects growth curve model.
-This model formally assumes a set of propositions about change that reflect the theoretical debate to this point.
+The estimation procedure begins with with Lersch's [-@lersch2023] Life Course Adaption Model, which formalizes survey responses at time $t$ as a function of individual-level random intercepts and slopes for survey age.
+In some contexts, this model is called a mixed-effects growth curve model.
+This model assumes a set of propositions about change that reflect the theoretical debate to this point.
 First, consistent with the settled disposition model, it assumes that people start the survey with cultural differences, modeled as random intercepts for each respondent.
@@
 As a first step in our analysis, we fit this LCAM to each of the 609 measures of personal culture outlined above.
-We then use these models to derive two measures of systematic variance that correspond to the two collections of theoretical processes outlined above: interpersonal differences at baseline and intrapersonal change over time.
+We then use these models to derive two measures of the components we outlined above: pre-existing interpersonal differences and intrapersonal change over time.
 
@@
 This is the proportion of total variance attributable to interpersonal differences.
-(This number is equivalent to the conditional $R^2$ value generated by a model with only random intercepts for all respondents.) Second, we calculate another sum of squared residuals generated by allowing the predictions to vary by estimated LCAM trajectory.
+(This number is equivalent to the conditional $R^2$ value generated by a model with only random intercepts for all respondents.)
+
+Second, we calculate another sum of squared residuals generated by allowing the predictions to vary by estimated LCAM trajectory.
 We divide this number by the total sum of squares and subtract it from 1.
@@
 Theories of personal cultural change that link changes in personal culture to social experiences, including the LCAM, suggest that the longer we observe respondents, the more likely people are to undergo potentially transformative experiences and therefore the more variance would be attributable to  intrapersonal change.
-Finding a negative coefficient here seems to challenge to that assumption.
+Finding a negative coefficient here seems to challenge that assumption.
 
@@
 It is worth pointing out that this meaningful difference in $\omega$ values across education groups and across questions would not have been detectable using previous methods.
-For partisan identification and political ideology, both college-educated respondents and people with less than college degree would likely favor the AUM or LCAM over the SDM because they these questions both show evidence of some members of the population making *some* intrapersonal change.
+For partisan identification and political ideology, both college-educated respondents and people with less than college degree would likely favor the AUM or LCAM over the SDM because these questions both show evidence of some members of the population making *some* intrapersonal change.
 In other words, the tournament of models obscures the fact that the relative importance of intrapersonal change differs across these two groups, that intrapersonal change explains more systematic variance for ideological identification and partisan identification for people without a college degree, and that there appears to be more durable change on questions of affirmative action than on questions of government aid to black Americans.

### 2023-09-29 manuscript/manuscript.tex
more SV edits

@@
 \begin{abstract}
-Recent work has produced conflicting interpretations on the question of
-whether adults undergo durable changes in their personal culture over
-time. This paper asserts that this divergence is less due to
-researchers' findings than the approach taken -- a tournament of models
-focusing on the presence or absence of change. To move beyond this, we
-reanalyze the data used previously in this debate with an approach that
-quantifies the processes of cultural difference. In doing so, we compare
-how much cultural difference is explained by intrapersonal change and
-interpersonal differences. Our results harmonize recent findings by
-showing that, while most measures of personal culture show durable
-change over time, the variance explained by intrapersonal change is
-typically small compared to that of interpersonal differences at
-baseline. Quantifying these processes offers new ways to address
-theoretical questions of when and how their relative importance shifts.
-We exemplify this approach by exploring how the importance of these
-processes differs between college graduates and the rest of the
-population.
+Does personal culture---a person's attitudes, beliefs, values, and
+practices---change over the life course, or is it largely fixed by
+adulthood? Recent work has reached substantially different conclusions
+on this question. We argue that this disagreement is an unintended
+consequence of the ``tournament of models'' approach researchers have
+used. We introduce a new measure for quantifying the proportion of
+systematic variance in panel data attributable to intrapersonal change
+versus stable interpersonal differences. Applying this measure to 609
+items from seven surveys in five countries, we find that although
+intrapersonal change is common, it is generally small in magnitude. We
+also demonstrate that this measure is useful for comparing social groups
+by showing that intrapersonal change is less common among U.S. college
+graduates than among those without a college degree. Overall, we believe
+this new measure will be a useful methodological tool and help reconcile
+an important theoretical debate.
 \end{abstract}
@@
 
-Does personal culture---an individual's attitudes, beliefs, values, and
+Does personal culture---a person's attitudes, beliefs, values, and
 practices---change over the life course, or is it largely fixed by
@@
 
-Because all of these theoretical perspectives have some empirical
-support, the current debate is not about whether their posited processes
-exist at all, but about their relative contribution to explaining the
-cultural differences we see in the world. That said, scholars have
-struggled to reach a consensus on the importance of change during
-adulthood. Over the span of a few years, most apparent changes in
-people's responses to personal survey items appear to be transitory,
-with little evidence of persistent change among adults (Kiley and Vaisey
-2020; Vaisey and Kiley 2021). This suggests that change during adulthood
-is not a major factor in explaining personal cultural differences. When
-we consider a longer time horizon, however, there is evidence that
-adults make at least some persistent changes (Lersch 2023).
+Because all of these perspectives have some empirical support, the
+current debate is not about whether the processes they posit exist at
+all, but about their relative contribution to explaining the cultural
+differences we see in the world. Nevertheless, researchers have
+struggled to reach a consensus on the importance of cultural change
+during adulthood. Over the span of a few years, most apparent changes in
+adults' survey responses appear to be transitory, with little evidence
+of persistent change (Kiley and Vaisey 2020; Vaisey and Kiley 2021).
+This suggests that change during adulthood may not be a major factor in
+explaining contemporary cultural divides. By contrast, when we consider
+a longer time horizon, there is evidence that adults make at least some
+persistent changes (Lersch 2023).
 
@@
 different theoretical perspectives (Lersch 2023: p.~228). That is,
-researchers have compared the fit of different statistical models to
-each item to determine whether it is more consistent with ``no change''
-or ``any change.'' The size of these piles then serves as the primary
-evidence for the truth of the theory (Kiley and Vaisey 2020; Lersch
-2023; Vaisey and Kiley 2021).
+researchers have relied on model selection criteria---primarily the
+Bayesian Information Criterion (Raftery 1995)---to sort variables into
+the ``no change'' or ``any change'' piles. The size of these piles then
+serves as the primary evidence for the truth of the theory (Kiley and
+Vaisey 2020; Lersch 2023; Vaisey and Kiley 2021).
 
 We argue, however, that asking \emph{whether} people change on an item
-may cause more problems than it solves. Being able to detect \emph{any}
+is an unintended source of confusion. Being able to detect \emph{any}
 change on an item depends not only on sample populations, measurement,
@@
 as ``change.'' These differences can limit the possibility of agreement
-on which ``pile'' a variable belongs. Moreover, asking whether there is
-evidence of change reduces to, ``Is there \emph{any} evidence of change
-in this item?'' But such binary inquiries are ill-suited to address the
-core of the theoretical debate: determining \emph{how much} the process
-of intrapersonal change contributes to cultural differences.
-
-There is enough evidence to assert that neither pre-adult socialization
-nor the contemporary social context alone can explain the full range of
-cultural differences. The question that remains is about the
-\emph{relative} contributions of these processes. In the current debate,
-for a single construct (e.g., support for gay rights), existing
-approaches can only make claims about \emph{whether} intrapersonal
-change is happening, not about how much. This also prevents
-investigating under what conditions, across which groups, and in what
-domains do we see differences in the relative importance of
-intrapersonal change and pre-existing interpersonal differences. These
-questions are about \emph{degree}, not existence; to take this debate
-forward therefore requires precise measurement rather than declarations
-of victory for a particular perspective.
+on the assignment of a variable to the right pile. Moreover, asking
+whether there is evidence of change reduces to, ``Is there \emph{any}
+evidence of change in this item?'' But such binary questions are
+ill-suited to address the core of the theoretical debate: determining
+\emph{how much} the process of intrapersonal change contributes to
+cultural differences.
+
+Of course neither pre-adult socialization nor the contemporary social
+context alone can explain the full range of cultural differences. The
+relevant question is about the \emph{relative} contributions of these
+processes. In the current debate, for a single construct (e.g., support
+for gay rights), model comparison approaches can only help decide
+\emph{whether} intrapersonal change is happening, not about how much.
+This also prevents investigating under what conditions, across which
+groups, and in what domains we might see differences in the the amount
+of intrapersonal change. These questions are about \emph{degree}, not
+existence. Taking this debate forward therefore requires precise
+measurement rather than declarations of victory for a particular
+perspective.
 
 In this paper, we move beyond the ``tournament of models'' approach and
-introduce a method for quantifying the relative contributions of
-interpersonal differences and intrapersonal change for a single item.
-Drawing on seven panel data sets from five countries, we quantify the
-amount of variance in survey responses that is due to between-person
-differences at a single point during the survey versus the amount of
-variance in responses that is due to systematic intrapersonal change
-over the study period. This measure quantifies the amount of change over
-the duration of the panel relative to the stable differences that exist
-between people.
-
-Using this measure, we observe similar patterns across the datasets that
-have previously yielded conflicting interpretations. In all
-datasets---despite their varied social contexts and duration---we find a
-similar distribution of intrapersonal change. In general, we find that
-intrapersonal change accounts for a substantially smaller amount of
-variance in personal culture survey item than stable interpersonal
-differences. Of course, whether the amount of intrapersonal change in an
-item is large enough to be ``meaningful'' is a substantive and
-theoretical question. Nevertheless, on many measures of personal culture
-(some observed for over a decade), we do not observe enough
-intrapersonal change for it to play a substantial role in explaining the
-differences we observe between adults. Notable, interesting, and
-important exceptions exist, of course, but they are departures from the
-overall pattern.
+introduce a new method for quantifying the relative contributions of
+interpersonal differences and intrapersonal change for a single item. We
+use 609 items from seven panel surveys from five countries to quantify
+the proportion of variance in that is attributable to systematic
+intrapersonal change over the study period rather than to stable
+interpersonal differences.
+
+We observe similar results across all datasets despite their varied
+social contexts and duration. In general, we find that intrapersonal
+change accounts for a small amount of variance in personal culture
+survey items. Of course, any question about whether a given amount of
+intrapersonal change is ``meaningful'' or ``important'' is a substantive
+and theoretical. Nevertheless, on many measures (some observed for over
+a decade), we do not see enough change to believe that intrapersonal
+change processes play a substantial role in explaining the differences
+we observe between adults. Notable, interesting, and important
+exceptions exist, but they are departures from the overall pattern.
 
 In addition to advancing a theoretical debate, our method can also help
-address substantive questions. To illustrate how it might be used by a
-researcher with a specific topical interest, we investigate differences
-in the relative importance of intrapersonal change for people with and
-without a college degree. We find a smaller amount of systematic change
-among college-educated respondents, suggesting that college solidifies
-one's personal culture rather than fostering an openness to new
-information. This offers a novel empirical basis to theorize about the
-role of university education in shaping personal culture.
-
-Taken together, the goal of this paper is to advance---and, hopefully,
+address specific substantive questions. To illustrate how it might be
+used by a researcher with a specific topical interest, we investigate
+differences in the relative importance of intrapersonal change for
+people with and without a college degree. We find a smaller amount of
+systematic change among college-educated respondents, suggesting that
+college crystallizes one's personal culture rather than fostering an
+openness to new information. This offers a new empirical basis to
+theorize about the role of education in influencing personal culture.
+
+Taken together, our goal in this paper is to advance---and, hopefully,
 transcend---debates about \emph{whether} adults change. People do
@@
 
-\hypertarget{change-and-stability-in-personal-culture}{%
-\subsection{Change and Stability in Personal
-Culture}\label{change-and-stability-in-personal-culture}}
+\hypertarget{stability-and-change-in-personal-culture}{%
+\subsection{Stability and Change in Personal
+Culture}\label{stability-and-change-in-personal-culture}}
 
@@
 ``correct'' at all times and in all places. A population observed over
-some period of time contains a mix of people who are noticeably changing
-and people who are not. Instead, different perspectives argue that each
-of these ideal-typical models is more operative at different times, for
-different people, and for different elements of personal culture. For
-example, adolescence and early adulthood is typically viewed as a
-``formative period'' for personal cultural development and thus
-characterized by higher rates of active updating, while middle age and
-later life are potentially characterized more by settled dispositions
-(Alwin and Krosnick 1991; Eaton et al. 2009; Krosnick and Alwin 1989).
-Similarly, salient issues, such as views around gay rights in the 2010s;
-issues that see substantial elite realignments, such as views around the
+some period of time contains a mix of people who are changing and people
+who are not. Instead, different perspectives argue that each of these
+ideal-typical models is more operative at different times, for different
+people, and for different elements of personal culture. For example,
+adolescence and early adulthood is typically viewed as a ``formative
+period'' for personal cultural development and thus characterized by
+higher rates of active updating, while middle age and later life are
+potentially characterized more by settled dispositions (Alwin and
+Krosnick 1991; Eaton et al. 2009; Krosnick and Alwin 1989). Similarly,
+salient issues, such as views around gay rights in the 2010s; issues
+that see substantial elite realignments, such as views around the
 Vietnam conflict in the 1970s (Zaller 1992); and novel issues of public
@@
 Vaisey (2020) found limited evidence of durable change. While the
-majority of items preferred the AUM, the amount of durable change
+majority of items did prefer the AUM, the amount of durable change
 detected on these items was small. A substantial minority of questions
-(39 percent) favored the SDM, meaning they showed no evidence of durable
-change. Questions with more evidence of durable change included salient
-issues like gay marriage and questions tapping ``public'' statements or
-behaviors such as partisan identification and religious service
-attendance. There was also more evidence of durable change among early
-adults (people ages 18-30) than the rest of the population. Overall, the
-researchers concluded that ``results ultimately suggest that real,
-persistent attitude change is an uncommon phenomenon among adults'\,'
-(Kiley and Vaisey 2020: p.~500; see also Vaisey and Kiley 2021). This
-lack of durable change is consistent with other recent findings that
-cohort replacement plays a somewhat larger role than period effects in
-explaining differences in personal culture (Vaisey and Lizardo 2016).
+(39 percent) favored the SDM, meaning they were more consistent with
+zero durable change. Questions with more evidence of durable change
+included salient issues like gay marriage and questions tapping
+``public'' statements or behaviors such as partisan identification and
+religious service attendance. There was also more evidence of durable
+change among early adults (people ages 18-30) than among the rest of the
+population. Overall, the researchers concluded that ``results ultimately
+suggest that real, persistent attitude change is an uncommon phenomenon
+among adults'\,' (Kiley and Vaisey 2020: p.~500; see also Vaisey and
+Kiley 2021). This lack of durable change is consistent with other recent
+findings that cohort replacement plays a somewhat larger role than
+period effects in explaining differences in personal culture (Vaisey and
+Lizardo 2016).
 
@@
 Second, Kiley and Vaisey's analysis of the AUM and SDM is based on
-three-wave panel data over four years, which might not be extensive
-enough to adjudicate the two models. Lersch evaluates the LCAM against
-the AUM and SDM using panel data spanning a wider duration (from 3 to 36
-years) and more waves (3 to 18), offering a better chance to observe
-durable change.
+three-wave panel data over four years, which might not be intensive or
+extensive enough to adjudicate the two models. Lersch evaluates the LCAM
+against the AUM and SDM using panel data spanning a wider duration (from
+3 to 36 years) and more waves (3 to 18), offering a better chance to
+observe durable change.
 
@@
 Lersch's results. Also Lersch (2023) found evidence of durable change on
-most items, but this change only accounted for a small proportion of
-variance explained. On average, people changed only about .07 standard
-deviations over 10 years. This is consistent with Kiley and Vaisey's
-finding that durable change is often small. Moreover, a quarter of the
-items Lersch studied still preferred the SDM, suggesting that even with
-different assumptions assumptions about change and more extensive data,
-many questions show no evidence of durable change. Consistent with this,
-in studies showing evidence of durable change in specific measures of
-personal culture, the change is often small. For example, Brocic and
-Miles (2021) estimate that completing graduate degrees in humanities,
-arts, and social sciences shift peoples' moral relativism only about .2
-standard deviations on average compared to people with no college
-degree, and this is the largest effect they identify. And studies of
-aggregate change show that, even on items where cohort effects explain
-more variance than period effects, there is always evidence that some
-people change over time (Vaisey and Lizardo 2016). In other words,
-despite different interpretations, the results of previous work are in
-many ways highly consonant.
+most items, but these changes were small. On average, people changed
+only about .07 standard deviations over 10 years. This is consistent
+with Kiley and Vaisey's finding that durable change is often small.
+Moreover, a quarter of the items Lersch studied still preferred the SDM,
+suggesting that even with different assumptions about change and more
+extensive data, many questions still are consistent with a model that
+allows zero intrapersonal change.
+
+Other studies are compatible with these findings as well. For example,
+Brocic and Miles (2021) estimate that completing graduate degrees in
+humanities, arts, and social sciences shift peoples' moral relativism
+only about .2 standard deviations on average compared to people with no
+college degree, and this is the largest effect they identify. And
+studies of aggregate change show that, even on items where cohort
+effects explain more variance than period effects, there is always
+evidence that some people change over time (Vaisey and Lizardo 2016). In
+other words, despite different interpretations, the results of previous
+work are in many ways highly consonant.
 
@@
 
-The current debate can be simplified to asking whether people
-\emph{ever} change. This is rarely theoretically interesting since the
-answer is almost always ``yes.'' Over sufficient time and in a large
-enough sample, researchers will likely observe some evidence for durable
-individual change. A lack of evidence for change may be due to a poor
-survey question, a low resolution in response options, or that the
-question simply was not asked for long enough or to enough people.
-Conversely, finding evidence for some change tells us only that it was
-``not zero.'' From that alone, we learn little about how much change has
-happened in a population.
+To this point, the theoretical debate has been about whether people
+\emph{ever} change. As we argued above, this is rarely the right
+question since the answer is almost always ``yes.'' Over sufficient time
+and in a large enough sample, researchers will likely observe some
+evidence for durable individual change. A lack of evidence for change
+may be due to a poor survey question, a low resolution in response
+options, or that the question simply was not asked for long enough or to
+enough people. Conversely, finding evidence for some change tells us
+only that it was ``not zero.'' From that alone, we learn little about
+how much intrapersonal change has happened in a population.
 
@@
 this model fits a particular data set better than a model that assumes
-that people never change, researchers should quantify these individual
-changes and compare those to the differences that exist between people
-at the panel's start. In other words, the focus should shift from
-asking, ``do people change?'' to asking, ``what are the relative
-contributions of change and initial differences for explaining variance
-in personal culture?''
-
-We propose an approach that offers such a quantification. Our measure
-relates two metrics that each reflect a combination of relevant
-theoretical processes. The first metric, interpersonal differences at
-the start of the survey, reflects the accumulated experiences of people
-prior to entering the panel survey that shape how they answer questions.
-Lersch calls this ``early imprinting.'' While commonly associated with
+that people never change, it would be better to quantify these
+individual changes and compare them to the stable differences that exist
+between people. In other words, the focus should shift from asking, ``do
+people change?'' to asking, ``what are the relative contributions of
+change and pre-existing differences for explaining variance in personal
+culture?''
+
+We propose an approach---which we formalize below---that offers such a
+quantification. Our measure relates two variance components that each
+reflect a combination of relevant theoretical processes. The first
+component, stable interpersonal differences, reflects the accumulated
+experiences of people prior to entering the panel survey. Lersch (2023:
+24) calls this ``early imprinting.'' While commonly associated with
 experiences during a formative period that result in settled
 dispositions, this variation could also reflect experiences that
-happened at any stage in life as long as they predate the panel and
-consistently affect subsequent responses. For instance, for those people
-entering the panel post-retirement, this ``imprinting'' might reflect
-this pivotal life transition. Consequently, this metric also reflects
-variation in individuals' social roles or statuses at the start of the
-panel that were important in shaping their dispositions.
-
-The second metric, the amount of intrapersonal linear change people make
-during the panel, captures durable changes in personal culture over
+happened at any time as long as they predate the panel and consistently
+affect subsequent responses. For instance, for those people entering the
+panel post-retirement, this ``imprinting'' might reflect this pivotal
+life transition. Consequently, this component also reflects variation in
+individuals' social roles or statuses at the start of the panel that
+were important in shaping their dispositions.
+
+The second component, the amount of intrapersonal linear change people
+make during the panel, captures durable changes in personal culture over
 time. This captures the set of processes collectively called
@@
 such as moving into a new environment or adopting new social roles
-(although he never measures these directly). Additionally, they might
-reflect the diffusion of new cultural forms across social networks, cues
-from political elites or otherwise culturally influential leaders, the
+(although he does not measure these directly). They might also reflect
+the diffusion of new cultural forms across social networks, cues from
+political elites or otherwise culturally influential leaders, the
 emergence of issues in politics or culture, or large-scale social
@@
 
-A third metric, which we might call ``residual variance'' or
+A third component, which we might call ``residual variance'' or
 ``fluctuation,'' accounts for the remainder of variance in peoples'
@@
 erroneous response selections, or responses getting coded incorrectly.
-Although this third metric at least partially reflects important
+Although this third component at least partially reflects important
 processes of personal culture, it does not directly touch on the ongoing
@@
 At a theoretical level, the coexistence of the sets of processes
-captured in the two main metrics, interpersonal differences and
+captured in the two main components, interpersonal differences and
 intrapersonal change, is undeniable, and they are linked in many ways.
@@
 
-Furthermore, these metrics are not necessarily useful in isolation or
+Furthermore, these components are not necessarily useful in isolation or
 without context. What is important for the theoretical debate is the
@@
 change relative to interpersonal difference to view it as an important
-explanation for why people differ from each other as observed? Only
-through quantifying the two processes we can get closer to a true answer
-to this question.
+process for why people differ from each other? Only through quantifying
+the two components we can get closer to a true answer to this question.
 
 Yet the true utility of the proposed quantification goes beyond settling
-past debates; it allows researchers to ask questions that were difficult
-to explore with previous approaches. For example, classifying survey
-items based on whether they show durable change overlooks possibly
-important differences among those that do. By quantifying the relative
-contributions of interpersonal differences and intrapersonal change,
-researchers can gauge their relative importance in explaining
-differences among survey items, between groups within a given
-population, across time, and across societies.
+past debates; it allows researchers to ask new questions. For example,
+classifying survey items based on whether they show durable change
+overlooks possibly important differences among those that do show such
+change. By quantifying the relative contributions of interpersonal
+differences and intrapersonal change, researchers can gauge their
+relative importance in explaining differences among survey items,
+between groups within a given population, across time, and across
+societies.
 
@@
 By shifting the question and adjusting the approach, we expect
-consistent patterns across datasets previously analysed in the debate.
-First, because previous work found evidence of durable change on most
-measures of personal culture, we expect that we will find some
-intrapersonal change in most items. But because these earlier studies
-also found many questions (between 25 and 35 percent) lacking evidence
-of \emph{any} durable change, we expect considerable variation in the
-variance accounted for by intrapersonal change, with some questions
-indicating essentially no variance attributable to such change.
+consistent patterns across datasets. First, because previous work found
+evidence of durable change on most measures of personal culture, we
+expect that we will find some intrapersonal change in most items. But
+because these earlier studies also found many questions (between 25 and
+35 percent) lacking evidence of \emph{any} durable change, we expect
+considerable variation in the proportion of variance accounted for by
+intrapersonal change, with some questions indicating essentially none at
+all.
 
 Second, we expect the variance attributable to intrapersonal change to
-be modest compared to interpersonal differences at the start of the
-survey. This expectation should be uncontroversial, regardless of the
-theoretical process assumed to be most relevant for a particular
-question. Panel surveys do not encompass individuals' entire lifetimes.
-Thus, even if intrapersonal change during adulthood is the major driver
-of cultural differences, the change detected in a panel covering 20
-years probably will not fully account for the entire variance of
-interpersonal differences.
+be modest compared to pre-existing interpersonal differences. This
+expectation should be uncontroversial, regardless of the theoretical
+process assumed to be most relevant for a particular question. Panel
+surveys do not encompass individuals' entire lifetimes. Thus, even if
+intrapersonal change during adulthood is the major driver of cultural
+differences, the change detected in a panel covering 20 years probably
+will not fully account for the entire variance of interpersonal
+differences.
 
 Our expectations so far largely echo what was observed in previous
-studies. However, theoretical perspectives differ in how much variance
-they would expect to be driven by intrapersonal change, enabling
-quantification to provide a new perspective on existing questions.
-``Formative period'' and cohort theories, which emphasize early-life
-socialization as a strong influence on adult personal culture, suggest
-that the variance from intrapersonal change would still fall short of
-explaining interpersonal differences, even when we account for the
-duration of observation. Conversely, other theoretical perspectives
-emphasizing the importance of people's contemporaneous social
-environment for personal culture might expect intrapersonal change to
-account for a large portion of the variance, regardless of the duration.
-
-Third, we expect the variance explained by each process ---interpersonal
-differences and intrapersonal change--- to depend on specific attributes
+studies. However, theoretical perspectives differ in how much change
+they expect, enabling quantification to provide a new perspective on
+existing questions. ``Formative period'' and cohort theories, which
+emphasize early-life socialization as a strong influence on adult
+personal culture, imply low amounts of intrapersonal change, even when
+we account for the duration of observation. Conversely, other
+theoretical perspectives emphasizing the importance of people's
+contemporaneous social environment might expect a larger degree of
+intrapersonal change, even in a relatively short panel.
+
+Third, we expect the variance explained by each process---interpersonal
+differences and intrapersonal change---to depend on specific attributes
 of the question and panel. First, the higher the response resolution for
 a given question and the larger the sample, the greater the variance
-explained by intrapersonal change should be, as smaller changes in
-intrapersonal change will be easier to detect. Second, the LCAM and
-various theoretical perspectives that underlie it posit that social
-transitions drive changes in personal culture. Since the probability
-that any person experiences such a transition goes up the longer we
-observe them, we expect panel duration to be positively related to the
-proportion of variation explained by intrapersonal change.
+attributed to intrapersonal change should be, as smaller changes in
+intrapersonal change will be easier to detect. Second, life course
+theories posit that social transitions drive changes in personal
+culture. Since the probability that any person experiences such a
+transition goes up the longer we observe them, we expect panel duration
+to be positively related to the proportion of variation explained by
+intrapersonal change.
 
@@
 a long period of time (the range of surveys spans from 1968 to 2021),
-with 609 personal culture items that capture attitudes, beliefs, values,
+with 609 personal culture items capturing attitudes, beliefs, values,
 self-assessments, self-descriptions, and behaviors (Alwin 2007). We
@@
 
-We start with Lersch's (2023) Life Course Adaption Model, which
-formalizes survey responses at time \(t\) as a function of
-individual-level random intercepts and slopes for survey age, what is
-often called a mixed-effects growth curve model. This model formally
-assumes a set of propositions about change that reflect the theoretical
-debate to this point. First, consistent with the settled disposition
-model, it assumes that people start the survey with cultural
-differences, modeled as random intercepts for each respondent. Second,
-it assumes that people change over time, taking the form of random
-slopes for each respondent as a linear function of time. Third, it
-assumes that people deviate around this baseline randomly over time,
+The estimation procedure begins with with Lersch's (2023) Life Course
+Adaption Model, which formalizes survey responses at time \(t\) as a
+function of individual-level random intercepts and slopes for survey
+age. In some contexts, this model is called a mixed-effects growth curve
+model. This model assumes a set of propositions about change that
+reflect the theoretical debate to this point. First, consistent with the
+settled disposition model, it assumes that people start the survey with
+cultural differences, modeled as random intercepts for each respondent.
+Second, it assumes that people change over time, taking the form of
+random slopes for each respondent as a linear function of time. Third,
+it assumes that people deviate around this baseline randomly over time,
 reflecting ``fluctuation'' or short-term non-persistent change.
@@
 measures of personal culture outlined above. We then use these models to
-derive two measures of systematic variance that correspond to the two
-collections of theoretical processes outlined above: interpersonal
-differences at baseline and intrapersonal change over time.
+derive two measures of the components we outlined above: pre-existing
+interpersonal differences and intrapersonal change over time.
 
@@
 the conditional \(R^2\) value generated by a model with only random
-intercepts for all respondents.) Second, we calculate another sum of
-squared residuals generated by allowing the predictions to vary by
-estimated LCAM trajectory. We divide this number by the total sum of
-squares and subtract it from 1. We then subtract \(V(D)\) from this
-value. This gives us a value we denote as \(V(C)\), the additional
-proportion of variance accounted for by intrapersonal change.
+intercepts for all respondents.)
+
+Second, we calculate another sum of squared residuals generated by
+allowing the predictions to vary by estimated LCAM trajectory. We divide
+this number by the total sum of squares and subtract it from 1. We then
+subtract \(V(D)\) from this value. This gives us a value we denote as
+\(V(C)\), the additional proportion of variance accounted for by
+intrapersonal change.
 
@@
 attributable to intrapersonal change. Finding a negative coefficient
-here seems to challenge to that assumption.
+here seems to challenge that assumption.
 
@@
 less than college degree would likely favor the AUM or LCAM over the SDM
-because they these questions both show evidence of some members of the
+because these questions both show evidence of some members of the
 population making \emph{some} intrapersonal change. In other words, the
@@
 
+\leavevmode\vadjust pre{\hypertarget{ref-raftery1995}{}}%
+Raftery, Adrian E. 1995. {``Bayesian {Model} {Selection} in {Social}
+{Research}.''} \emph{Sociological Methodology} 25:111--63. doi:
+\href{https://doi.org/10.2307/271063}{10.2307/271063}.
+
 \leavevmode\vadjust pre{\hypertarget{ref-robinson2007}{}}%

### 2023-09-29 manuscript/manuscript.Rmd
added equations for V(D) and V(C)

@@
 
-We estimate these using a form of variance decomposition.
-First, to measure stable interpersonal differences, we calculate the sum of the squared residuals derived from assuming that each person spends the entire survey at the estimated midpoint of their LCAM trajectory.[^2] We then divide this value by the total sum of squares and subtract this number from from 1, producing a value we denote as $V(D)$.
-This is the proportion of total variance attributable to interpersonal differences.
-(This number is equivalent to the conditional $R^2$ value generated by a model with only random intercepts for all respondents.)
+We estimate these using a form of variance decomposition. First, to measure stable interpersonal differences, we calculate $V(D)$ according to the equation below.[^2] 
+$$
+V(D) = 1 - \frac{\sum_{i=1}^N \sum_{t=1}^T [y_{it} - (\hat{\beta}_0 + \hat{\alpha}_{0i} + (\hat{\beta}_1 + \hat{\alpha}_{1i}) \overline{\text{year}}_{it})]^2}{\sum_{i=1}^N \sum_{t=1}^T [y_{it} - \bar{y}_{it}]^2}
+$$
 
-Second, we calculate another sum of squared residuals generated by allowing the predictions to vary by estimated LCAM trajectory.
-We divide this number by the total sum of squares and subtract it from 1.
-We then subtract $V(D)$ from this value.
-This gives us a value we denote as $V(C)$, the additional proportion of variance accounted for by intrapersonal change.
-
-[^2]: We use the midpoint estimate, rather than the value at the first wave, because this is the best measure of "baseline" available under the assumption that the person does not change.
+[^2]: We use the model estimate from the midpoint year, rather than the value at the first wave, because this is the best measure of "baseline" available under the assumption that the person does not change.
 The first wave measure alone contains an unknown amount of measurement error and transient fluctuations.
 
-These two numbers together reflect the total systematic variance of the LCAM that can be attributed to either differences between people or linear intrapersonal change over the course of the panel.[^3] Our principal outcome of interest is the proportion of systematic variance accounted for by intrapersonal change, which we calculate as
+This gives us the proportion of total variance in an item attributable to stable interpersonal differences.
 
-[^3]: This does not mean that other factors might not explain additional variance if they were included in the model.
-For example, peoples' responses about religious beliefs and participation might systematically vary by the day of the week or month in which they are surveyed.
-This variation is moved into the "random" or non-systematic component of the model for us.
+To calculate the second component, which is the variance attributable to systematic intrapersonal change, we calculate $V(C)$ as follows:
+
+$$
+V(C) = 1 - \frac{\sum_{i=1}^N \sum_{t=1}^T [y_{it} - (\hat{\beta}_0 + \hat{\alpha}_{0i} + (\hat{\beta}_1 + \hat{\alpha}_{1i}) \text{year}_{it})]^2}{\sum_{i=1}^N \sum_{t=1}^T [y_{it} - \bar{y}_{it}]^2} - V(D)
+$$
+
+This is the incremental proportion of variance accounted for when we allow the model predictions to change over time for each person.
+
+The third component, which is the proportion of variance attributable to measurement error or transient fluctuations, is what's left, such that these three numbers sum to 1.
+As we discussed above, this component is not of core interest here.
+
+Together, $V(D)$ and $V(C)$ reflect the total systematic variance of the outcome that the LCAM can attribute to either differences between people or linear intrapersonal change over the course of the panel.
+Our principal measure of interest is the proportion of systematic variance accounted for by intrapersonal change, which we calculate as
 
@@
 The higher this proportion is, the more systematic variance in responses is attributable to intrapersonal change during the panel.
-From now on, we use $\omega$ as a placeholder for *systematic variance attributable to intrapersonal change*.
+From now on, we use $\omega$ to stand for *the proportion of systematic variance attributable to intrapersonal change*.
 
-Because this quotient summarizes variance and therefore has no natural referent, it is hard to say at which values we observe "a lot" of intrapersonal change.
+Because this quotient summarizes variance proportions and therefore has no natural referent, it is hard to say at which values we observe "a lot" of intrapersonal change.
 But summarizing questions in this way allows us to compare the relative importance of our two broad processes across a range of questions that might have different levels of non-systematic variance.

### 2023-09-29 manuscript/manuscript.tex
added equations for V(D) and V(C)

@@
 We estimate these using a form of variance decomposition. First, to
-measure stable interpersonal differences, we calculate the sum of the
-squared residuals derived from assuming that each person spends the
-entire survey at the estimated midpoint of their LCAM
-trajectory.\footnote{We use the midpoint estimate, rather than the value
-  at the first wave, because this is the best measure of ``baseline''
-  available under the assumption that the person does not change. The
-  first wave measure alone contains an unknown amount of measurement
-  error and transient fluctuations.} We then divide this value by the
-total sum of squares and subtract this number from from 1, producing a
-value we denote as \(V(D)\). This is the proportion of total variance
-attributable to interpersonal differences. (This number is equivalent to
-the conditional \(R^2\) value generated by a model with only random
-intercepts for all respondents.)
-
-Second, we calculate another sum of squared residuals generated by
-allowing the predictions to vary by estimated LCAM trajectory. We divide
-this number by the total sum of squares and subtract it from 1. We then
-subtract \(V(D)\) from this value. This gives us a value we denote as
-\(V(C)\), the additional proportion of variance accounted for by
-intrapersonal change.
-
-These two numbers together reflect the total systematic variance of the
-LCAM that can be attributed to either differences between people or
-linear intrapersonal change over the course of the panel.\footnote{This
-  does not mean that other factors might not explain additional variance
-  if they were included in the model. For example, peoples' responses
-  about religious beliefs and participation might systematically vary by
-  the day of the week or month in which they are surveyed. This
-  variation is moved into the ``random'' or non-systematic component of
-  the model for us.} Our principal outcome of interest is the proportion
-of systematic variance accounted for by intrapersonal change, which we
-calculate as
+measure stable interpersonal differences, we calculate \(V(D)\)
+according to the equation below.\footnote{We use the model estimate from
+  the midpoint year, rather than the value at the first wave, because
+  this is the best measure of ``baseline'' available under the
+  assumption that the person does not change. The first wave measure
+  alone contains an unknown amount of measurement error and transient
+  fluctuations.} \[
+V(D) = 1 - \frac{\sum_{i=1}^N \sum_{t=1}^T [y_{it} - (\hat{\beta}_0 + \hat{\alpha}_{0i} + (\hat{\beta}_1 + \hat{\alpha}_{1i}) \overline{\text{year}}_{it})]^2}{\sum_{i=1}^N \sum_{t=1}^T [y_{it} - \bar{y}_{it}]^2}
+\]
+
+This gives us the proportion of total variance in an item attributable
+to stable interpersonal differences.
+
+To calculate the second component, which is the variance attributable to
+systematic intrapersonal change, we calculate \(V(C)\) as follows:
+
+\[
+V(C) = 1 - \frac{\sum_{i=1}^N \sum_{t=1}^T [y_{it} - (\hat{\beta}_0 + \hat{\alpha}_{0i} + (\hat{\beta}_1 + \hat{\alpha}_{1i}) \text{year}_{it})]^2}{\sum_{i=1}^N \sum_{t=1}^T [y_{it} - \bar{y}_{it}]^2} - V(D)
+\]
+
+This is the incremental proportion of variance accounted for when we
+allow the model predictions to change over time for each person.
+
+The third component, which is the proportion of variance attributable to
+measurement error or transient fluctuations, is what's left, such that
+these three numbers sum to 1. As we discussed above, this component is
+not of core interest here.
+
+Together, \(V(D)\) and \(V(C)\) reflect the total systematic variance of
+the outcome that the LCAM can attribute to either differences between
+people or linear intrapersonal change over the course of the panel. Our
+principal measure of interest is the proportion of systematic variance
+accounted for by intrapersonal change, which we calculate as
 
@@
 to intrapersonal change during the panel. From now on, we use \(\omega\)
-as a placeholder for \emph{systematic variance attributable to
+to stand for \emph{the proportion of systematic variance attributable to
 intrapersonal change}.
 
-Because this quotient summarizes variance and therefore has no natural
-referent, it is hard to say at which values we observe ``a lot'' of
-intrapersonal change. But summarizing questions in this way allows us to
-compare the relative importance of our two broad processes across a
-range of questions that might have different levels of non-systematic
-variance. This gives us a sense of the relative prevalence of these two
-processes for each question, which is not achievable under the
-tournament of models approach. This also allows us to compare which
-kinds of questions show more or less intrapersonal change relative to
-interpersonal differences. While the tournament of models approach
-allowed researchers to say whether questions in general showed
-\emph{any} evidence of updating, this approach allows us to make more
-specific claims about the prevalence of these theoretical processes for
-particular questions relative to each other.
+Because this quotient summarizes variance proportions and therefore has
+no natural referent, it is hard to say at which values we observe ``a
+lot'' of intrapersonal change. But summarizing questions in this way
+allows us to compare the relative importance of our two broad processes
+across a range of questions that might have different levels of
+non-systematic variance. This gives us a sense of the relative
+prevalence of these two processes for each question, which is not
+achievable under the tournament of models approach. This also allows us
+to compare which kinds of questions show more or less intrapersonal
+change relative to interpersonal differences. While the tournament of
+models approach allowed researchers to say whether questions in general
+showed \emph{any} evidence of updating, this approach allows us to make
+more specific claims about the prevalence of these theoretical processes
+for particular questions relative to each other.
 

### 2023-09-30 manuscript/manuscript.Rmd
fixed equation

@@
 
-We estimate these using a form of variance decomposition. First, to measure stable interpersonal differences, we calculate $V(D)$ according to the equation below.[^2] 
+We estimate these using a form of variance decomposition.
+First, to measure stable interpersonal differences, we calculate $V(D)$ according to the equation below, where $\widetilde{year_i}$ is the midpoint of the observed years for each respondent for that particular item.[^2]
 $$
-V(D) = 1 - \frac{\sum_{i=1}^N \sum_{t=1}^T [y_{it} - (\hat{\beta}_0 + \hat{\alpha}_{0i} + (\hat{\beta}_1 + \hat{\alpha}_{1i}) \overline{\text{year}}_{it})]^2}{\sum_{i=1}^N \sum_{t=1}^T [y_{it} - \bar{y}_{it}]^2}
+V(D) = 1 - \frac{\sum_{i=1}^N \sum_{t=1}^T [y_{it} - (\hat{\beta}_0 + \hat{\alpha}_{0i} + (\hat{\beta}_1 + \hat{\alpha}_{1i}) \widetilde{\text{year}}_{i})]^2}{\sum_{i=1}^N \sum_{t=1}^T [y_{it} - \bar{y}_{it}]^2}
 $$

### 2023-09-30 manuscript/manuscript.tex
fixed equation

@@
 measure stable interpersonal differences, we calculate \(V(D)\)
-according to the equation below.\footnote{We use the model estimate from
-  the midpoint year, rather than the value at the first wave, because
-  this is the best measure of ``baseline'' available under the
-  assumption that the person does not change. The first wave measure
-  alone contains an unknown amount of measurement error and transient
-  fluctuations.} \[
-V(D) = 1 - \frac{\sum_{i=1}^N \sum_{t=1}^T [y_{it} - (\hat{\beta}_0 + \hat{\alpha}_{0i} + (\hat{\beta}_1 + \hat{\alpha}_{1i}) \overline{\text{year}}_{it})]^2}{\sum_{i=1}^N \sum_{t=1}^T [y_{it} - \bar{y}_{it}]^2}
+according to the equation below, where \(\widetilde{year_i}\) is the
+midpoint of the observed years for each respondent for that particular
+item.\footnote{We use the model estimate from the midpoint year, rather
+  than the value at the first wave, because this is the best measure of
+  ``baseline'' available under the assumption that the person does not
+  change. The first wave measure alone contains an unknown amount of
+  measurement error and transient fluctuations.} \[
+V(D) = 1 - \frac{\sum_{i=1}^N \sum_{t=1}^T [y_{it} - (\hat{\beta}_0 + \hat{\alpha}_{0i} + (\hat{\beta}_1 + \hat{\alpha}_{1i}) \widetilde{\text{year}}_{i})]^2}{\sum_{i=1}^N \sum_{t=1}^T [y_{it} - \bar{y}_{it}]^2}
 \]

### 2023-09-30 manuscript/manuscript.Rmd
changes in appendices and supplements

@@
 header-includes:
+  \usepackage{caption}
   \usepackage{setspace}\doublespacing
@@
 
-Supplemental Materials A presents the estimated proportion of variance attributable to interpersonal differences and intrapersonal change, the estimated values of $\omega$, and the proportion of residual variance for each question.
+Appendix A shows the distribution of $V(D)$ and $V(C)$ across panels, and Supplemental Materials A presents the estimated proportion of variance attributable to interpersonal differences and intrapersonal change, the estimated values of $\omega$, and the proportion of residual variance for each question.
 Interpersonal differences are almost always the largest component of the total variance and tend to account for between 55 and 70 percent of total variance, while intrapersonal change is always the smallest, typically accounting for between 3 and 8 percent of total variance.
@@
 If the coefficient reflects a true negative effect of duration on $\omega$, we should see that same effect within questions.
-Results from this analysis are presented in Supplemental Materials C and contradict the coefficient from the regression model; with a few exceptions, we found that the longer we observe the same question, the higher $\omega$ is.
+Results from this analysis are presented in Appendix B and contradict the coefficient from the regression model; with a few exceptions, we found that the longer we observe the same question, the higher $\omega$ is.
 We interpret this combination of findings as suggesting that the kinds of questions asked for a longer time period tend to demonstrate less intrapersonal change than questions asked over shorter periods, rather than a true function of time.
@@
 
+# Appendix
+
+## Appendix A: The Distribution of Variation Components
+
+Figure A1 shows the distribution of $V(D)$ and $V(C)$ across all panels.
+
+```{=latex}
+\begin{figure}[htp]
+\begin{center}
+\caption*{Figure A1: The Distribution of V(D) and V(C)}
+```
+```{r figA1, out.width = "350px"}
+knitr::include_graphics("../figures/figure_A1.png")
+```
+```{=latex}
+\end{center}
+\end{figure}
+```
+
+## Appendix B: $\omega$ Values When the Last Wave Removed
+
+Figure A2 presents the alternative strategy we used to understand the effects of duration on intraindividual change.
+First, we restricted the sample such that we have at least 4 waves of observations for all participants.
+This reduced the number of items to 249.
+We then fitted the Life Course Adaption models to this sample.
+Second, we removed the last wave from all the observations, and refit the same models once more.
+In the final step, we calculated the $\omega$ values for each item.
+The Figure C1 shows a scatterplot for these two sets of observations.
+
+```{=latex}
+\begin{figure}[tb]
+\begin{center}
+\caption*{Figure A2: Difference in $\omega$ Across Two Sample Specifications} 
+```
+```{r figA2, out.width = "400px"}
+knitr::include_graphics("../figures/figure_A2.png")
+```
+```{=latex}
+\end{center}
+\end{figure}
+```
+
 # References

### 2023-09-30 manuscript/manuscript.tex
changes in appendices and supplements

@@
 \newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}
-\usepackage{setspace}\doublespacing \usepackage{endnotes} \renewcommand{\enotesize}{\normalsize} \let\footnote=\endnote
+\usepackage{caption} \usepackage{setspace}\doublespacing \usepackage{endnotes} \renewcommand{\enotesize}{\normalsize} \let\footnote=\endnote
 \usepackage{booktabs}
@@
 
-Supplemental Materials A presents the estimated proportion of variance
-attributable to interpersonal differences and intrapersonal change, the
-estimated values of \(\omega\), and the proportion of residual variance
-for each question. Interpersonal differences are almost always the
-largest component of the total variance and tend to account for between
-55 and 70 percent of total variance, while intrapersonal change is
-always the smallest, typically accounting for between 3 and 8 percent of
-total variance. Residual variance tends to account for between 22 and 37
-percent of variance, though on several questions residual variance is
-greater than 50 percent. This might indicate survey items with low
-reliability or ones that capture genuinely rapid fluctuations.
+Appendix A shows the distribution of \(V(D)\) and \(V(C)\) across
+panels, and Supplemental Materials A presents the estimated proportion
+of variance attributable to interpersonal differences and intrapersonal
+change, the estimated values of \(\omega\), and the proportion of
+residual variance for each question. Interpersonal differences are
+almost always the largest component of the total variance and tend to
+account for between 55 and 70 percent of total variance, while
+intrapersonal change is always the smallest, typically accounting for
+between 3 and 8 percent of total variance. Residual variance tends to
+account for between 22 and 37 percent of variance, though on several
+questions residual variance is greater than 50 percent. This might
+indicate survey items with low reliability or ones that capture
+genuinely rapid fluctuations.
 
@@
 duration on \(\omega\), we should see that same effect within questions.
-Results from this analysis are presented in Supplemental Materials C and
-contradict the coefficient from the regression model; with a few
-exceptions, we found that the longer we observe the same question, the
-higher \(\omega\) is. We interpret this combination of findings as
-suggesting that the kinds of questions asked for a longer time period
-tend to demonstrate less intrapersonal change than questions asked over
-shorter periods, rather than a true function of time. This supports the
+Results from this analysis are presented in Appendix B and contradict
+the coefficient from the regression model; with a few exceptions, we
+found that the longer we observe the same question, the higher
+\(\omega\) is. We interpret this combination of findings as suggesting
+that the kinds of questions asked for a longer time period tend to
+demonstrate less intrapersonal change than questions asked over shorter
+periods, rather than a true function of time. This supports the
 conclusion that the GSS shows less intrapersonal change because of its
@@
 
+\hypertarget{appendix}{%
+\section{Appendix}\label{appendix}}
+
+\hypertarget{appendix-a-the-distribution-of-variation-components}{%
+\subsection{Appendix A: The Distribution of Variation
+Components}\label{appendix-a-the-distribution-of-variation-components}}
+
+Figure A1 shows the distribution of \(V(D)\) and \(V(C)\) across all
+panels.
+
+\begin{figure}[htp]
+\begin{center}
+\caption*{Figure A1: The Distribution of V(D) and V(C)}
+
+\includegraphics[width=350px]{../figures/figure_A1}
+
+\end{center}
+\end{figure}
+
+\hypertarget{appendix-b-omega-values-when-the-last-wave-removed}{%
+\subsection{\texorpdfstring{Appendix B: \(\omega\) Values When the Last
+Wave
+Removed}{Appendix B: \textbackslash omega Values When the Last Wave Removed}}\label{appendix-b-omega-values-when-the-last-wave-removed}}
+
+Figure A2 presents the alternative strategy we used to understand the
+effects of duration on intraindividual change. First, we restricted the
+sample such that we have at least 4 waves of observations for all
+participants. This reduced the number of items to 249. We then fitted
+the Life Course Adaption models to this sample. Second, we removed the
+last wave from all the observations, and refit the same models once
+more. In the final step, we calculated the \(\omega\) values for each
+item. The Figure C1 shows a scatterplot for these two sets of
+observations.
+
+\begin{figure}[tb]
+\begin{center}
+\caption*{Figure A2: Difference in $\omega$ Across Two Sample Specifications}
+
+\includegraphics[width=400px]{../figures/figure_A2}
+
+\end{center}
+\end{figure}
+
 \hypertarget{references}{%

### 2023-09-30 supplementals/manuscript_supplementals.Rmd
changes in appendices and supplements

@@
 ```
-
-\newpage
-
-# Supplemental Materials C
-
-Figure C1 presents the alternative strategy we used to understand the effects of duration on intraindividual change.
-First, we restricted the sample such that we have at least 4 waves of observations for all participants.
-This reduced the number of items to 249.
-We then fitted the Life Course Adaption models to this sample.
-Second, we removed the last wave from all the observations, and refit the same models once more.
-In the final step, we calculated the $\omega$ values for each item.
-The Figure C1 shows a scatterplot for these two sets of observations.
-
-```{=latex}
-\begin{figure}[ht]
-\begin{center}
-\caption*{Figure C1: Difference in $\omega$ Across Two Sample Specifications} 
-```
-```{r fig3, out.width = "5in"}
-knitr::include_graphics("../figures/figure_S3.png")
-```
-```{=latex}
-\end{center}
-\end{figure}
-```

### 2023-09-30 supplementals/manuscript_supplementals.tex
changes in appendices and supplements

@@
 
-\newpage
-
-\hypertarget{supplemental-materials-c}{%
-\section{Supplemental Materials C}\label{supplemental-materials-c}}
-
-Figure C1 presents the alternative strategy we used to understand the
-effects of duration on intraindividual change. First, we restricted the
-sample such that we have at least 4 waves of observations for all
-participants. This reduced the number of items to 249. We then fitted
-the Life Course Adaption models to this sample. Second, we removed the
-last wave from all the observations, and refit the same models once
-more. In the final step, we calculated the \(\omega\) values for each
-item. The Figure C1 shows a scatterplot for these two sets of
-observations.
-
-\begin{figure}[ht]
-\begin{center}
-\caption*{Figure C1: Difference in $\omega$ Across Two Sample Specifications}
-
-\begin{center}\includegraphics[width=5in]{../figures/figure_S3} \end{center}
-
-\end{center}
-\end{figure}
-
 \end{document}

### 2023-09-30 manuscript/manuscript.Rmd
small changes and typos.

@@
 In the current debate, for a single construct (e.g., support for gay rights), model comparison approaches can only help decide *whether* intrapersonal change is happening, not about how much.
-This also prevents investigating under what conditions, across which groups, and in what domains we might see differences in the the amount of intrapersonal change.
+This also prevents investigating under what conditions, across which groups, and in what domains we might see differences in the amount of intrapersonal change.
 These questions are about *degree*, not existence.
@@
 In this paper, we move beyond the "tournament of models" approach and introduce a new method for quantifying the relative contributions of interpersonal differences and intrapersonal change for a single item.
-We use 609 items from seven panel surveys from five countries to quantify the proportion of variance in that is attributable to systematic intrapersonal change over the study period rather than to stable interpersonal differences.
+We use 609 items from seven panel surveys from five countries to quantify the proportion of variance that is attributable to systematic intrapersonal change over the study period rather than to stable interpersonal differences.
 
@@
 In general, we find that intrapersonal change accounts for a small amount of variance in personal culture survey items.
-Of course, any question about whether a given amount of intrapersonal change is "meaningful" or "important" is a substantive and theoretical.
+Of course, any question about whether a given amount of intrapersonal change is "meaningful" or "important" is substantive and theoretical.
 Nevertheless, on many measures (some observed for over a decade), we do not see enough change to believe that intrapersonal change processes play a substantial role in explaining the differences we observe between adults.
@@
 While they might make temporary changes in their declarative culture in reaction to their environments, this model assumes that people return to a settled baseline over a short period of time.
-This model underlies theories of cultural change that suggest people are imprinted by early socialization experiences such as the "past conditions of production" a Bourdieusian practice theory [@bourdieu1990], cohort replacement theories of aggregate change [@mannheim1952; @ryder1965], and control theories in social psychology [@smith-lovin1988; @robinson2007].
+This model underlies theories of cultural change that suggest people are imprinted by early socialization experiences such as the "past conditions of production" in Bourdieusian practice theory [@bourdieu1990], cohort replacement theories of aggregate change [@mannheim1952; @ryder1965], and control theories in social psychology [@smith-lovin1988; @robinson2007].
 
@@
 
-The estimation procedure begins with with Lersch's [-@lersch2023] Life Course Adaption Model, which formalizes survey responses at time $t$ as a function of individual-level random intercepts and slopes for survey age.
+The estimation procedure begins with Lersch's [-@lersch2023] Life Course Adaption Model, which formalizes survey responses at time $t$ as a function of individual-level random intercepts and slopes for survey age.
 In some contexts, this model is called a mixed-effects growth curve model.
@@
 First, to measure stable interpersonal differences, we calculate $V(D)$ according to the equation below, where $\widetilde{year_i}$ is the midpoint of the observed years for each respondent for that particular item.[^2]
+
 $$
@@
 As we noted above, the substantive importance of a given amount of intrapersonal change depends on a range of factors, including how long the panel runs, whether assumptions about linear change hold, and whether the panel is capturing a distinctly turbulent period or a distinctly stable period for the relevant item.
-However, if we assume that the period under observation is "typical" for a question---not a time of extremely heightened (or lowered) sensitivity or change---then it does not seem realistic for that intrapersonal change accounts for a large share of the cultural differences we see in the world.
+However, if we assume that the period under observation is "typical" for a question---not a time of extremely heightened (or lowered) sensitivity or change---then it does not seem realistic that intrapersonal change accounts for a large share of the cultural differences we see in the world.
 
@@
 
-The objective of this paper is to intervene in the recent debate on whether people change their personal culture -- their attitudes, beliefs, values, practices, and dispositions -- as they move through their adult life.
+The objective of this paper is to intervene in the recent debate on whether people change their personal culture -- their attitudes, beliefs, values, and practices -- as they move through their adult life.
 Instead of falling back on a "tournament of models" approach [@lersch2023: p.
@@
 Drawing a unified conclusion from survey items measuring various cultural forms on different scales and across different time frames is challenging.
-Nevertheless, the general pattern suggests that, for most items, intrapersonal change in adulthood is not the primary reason we see the differences we see between people in the world.
+Nevertheless, the general pattern suggests that, for most items, intrapersonal change in adulthood is not the primary reason for the differences we see between people in the world.
 
@@
 Our findings suggest that understanding variation in personal culture requires examining the conditions and experiences of early life.
-While sociological research often focuses on transitions between social roles, changes in social networks, and the experience of organizational environments, these factors seem to account for a smaller proportion of adult differences than early-life experience.
+While sociological research often focuses on transitions between social roles, changes in social networks, or the experience of organizational environments, these factors seem to account for a smaller proportion of adult differences than early-life experience.
 Simply put, we need more work on early-life socialization [@guhin2020].
@@
 
+# References
+
+::: {#refs}
+:::
+
 # Appendix
@@
 ```
-
-# References
-
-::: {#refs}
-:::

### 2023-09-30 manuscript/manuscript.tex
small changes and typos.

@@
 This also prevents investigating under what conditions, across which
-groups, and in what domains we might see differences in the the amount
-of intrapersonal change. These questions are about \emph{degree}, not
+groups, and in what domains we might see differences in the amount of
+intrapersonal change. These questions are about \emph{degree}, not
 existence. Taking this debate forward therefore requires precise
@@
 use 609 items from seven panel surveys from five countries to quantify
-the proportion of variance in that is attributable to systematic
+the proportion of variance that is attributable to systematic
 intrapersonal change over the study period rather than to stable
@@
 survey items. Of course, any question about whether a given amount of
-intrapersonal change is ``meaningful'' or ``important'' is a substantive
+intrapersonal change is ``meaningful'' or ``important'' is substantive
 and theoretical. Nevertheless, on many measures (some observed for over
@@
 cultural change that suggest people are imprinted by early socialization
-experiences such as the ``past conditions of production'' a Bourdieusian
-practice theory (Bourdieu 1990), cohort replacement theories of
-aggregate change (Mannheim 1952; Ryder 1965), and control theories in
-social psychology (Robinson 2007; Smith-Lovin and Heise 1988).
+experiences such as the ``past conditions of production'' in
+Bourdieusian practice theory (Bourdieu 1990), cohort replacement
+theories of aggregate change (Mannheim 1952; Ryder 1965), and control
+theories in social psychology (Robinson 2007; Smith-Lovin and Heise
+1988).
 
@@
 
-The estimation procedure begins with with Lersch's (2023) Life Course
+The estimation procedure begins with Lersch's (2023) Life Course
 Adaption Model, which formalizes survey responses at time \(t\) as a
@@
   change. The first wave measure alone contains an unknown amount of
-  measurement error and transient fluctuations.} \[
+  measurement error and transient fluctuations.}
+
+\[
 V(D) = 1 - \frac{\sum_{i=1}^N \sum_{t=1}^T [y_{it} - (\hat{\beta}_0 + \hat{\alpha}_{0i} + (\hat{\beta}_1 + \hat{\alpha}_{1i}) \widetilde{\text{year}}_{i})]^2}{\sum_{i=1}^N \sum_{t=1}^T [y_{it} - \bar{y}_{it}]^2}
@@
 time of extremely heightened (or lowered) sensitivity or change---then
-it does not seem realistic for that intrapersonal change accounts for a
+it does not seem realistic that intrapersonal change accounts for a
 large share of the cultural differences we see in the world.
@@
 whether people change their personal culture -- their attitudes,
-beliefs, values, practices, and dispositions -- as they move through
-their adult life. Instead of falling back on a ``tournament of models''
-approach (Lersch 2023: p.~228) to proclaim victory for a particular
-answer to this question, our goal is to refocus the debate to the
-theoretically more productive question of the relative importance of
-intrapersonal change for explaining differences in the personal culture.
-We propose a new approach that quantifies the amount of observed
-systematic variance that is attributable to either interpersonal
-differences at baseline or intrapersonal change over the duration of a
-panel.
+beliefs, values, and practices -- as they move through their adult life.
+Instead of falling back on a ``tournament of models'' approach (Lersch
+2023: p.~228) to proclaim victory for a particular answer to this
+question, our goal is to refocus the debate to the theoretically more
+productive question of the relative importance of intrapersonal change
+for explaining differences in the personal culture. We propose a new
+approach that quantifies the amount of observed systematic variance that
+is attributable to either interpersonal differences at baseline or
+intrapersonal change over the duration of a panel.
 
@@
 challenging. Nevertheless, the general pattern suggests that, for most
-items, intrapersonal change in adulthood is not the primary reason we
-see the differences we see between people in the world.
+items, intrapersonal change in adulthood is not the primary reason for
+the differences we see between people in the world.
 
@@
 sociological research often focuses on transitions between social roles,
-changes in social networks, and the experience of organizational
+changes in social networks, or the experience of organizational
 environments, these factors seem to account for a smaller proportion of
@@
 
+\hypertarget{references}{%
+\section{References}\label{references}}
+
+\hypertarget{refs}{}
+\begin{CSLReferences}{1}{0}
+\leavevmode\vadjust pre{\hypertarget{ref-alvarez2002}{}}%
+Alvarez, R. Michael, and John Brehm. 2002. \emph{Hard Choices, Easy
+Answers: Values, Information, and {American} Public Opinion}. Princeton,
+N.J: Princeton University Press.
+
+\leavevmode\vadjust pre{\hypertarget{ref-alwin2007}{}}%
+Alwin, Duane F. 2007. \emph{Margins of {Error}: {A} {Study} of
+{Reliability} in {Survey} {Measurement}}. Hoboken, N.J.: John Wiley \&
+Sons.
+
+\leavevmode\vadjust pre{\hypertarget{ref-alwin1991}{}}%
+Alwin, Duane F., and Jon A. Krosnick. 1991.
+{``\href{https://www.jstor.org/stable/2781642}{Aging, {Cohorts}, and the
+{Stability} of {Sociopolitical} {Orientations} {Over} the {Life}
+{Span}}.''} \emph{American Journal of Sociology} 97(1):169--95.
+
+\leavevmode\vadjust pre{\hypertarget{ref-bail2018}{}}%
+Bail, Christopher A., Lisa P. Argyle, Taylor W. Brown, John P. Bumpus,
+Haohan Chen, M. B. Fallin Hunzaker, Jaemin Lee, Marcus Mann, Friedolin
+Merhout, and Alexander Volfovsky. 2018. {``Exposure to Opposing Views on
+Social Media Can Increase Political Polarization.''} \emph{Proceedings
+of the National Academy of Sciences} 115(37):9216--21. doi:
+\href{https://doi.org/10.1073/pnas.1804840115}{10.1073/pnas.1804840115}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-bardi2009}{}}%
+Bardi, Anat, Julie Anne Lee, Nadi Hofmann-Towfigh, and Geoffrey Soutar.
+2009. {``The Structure of Intraindividual Value Change.''} \emph{Journal
+of Personality and Social Psychology} 97(5):913--29.
+
+\leavevmode\vadjust pre{\hypertarget{ref-bourdieu1990}{}}%
+Bourdieu, Pierre. 1990. \emph{The Logic of Practice}. Stanford, Calif. :
+Stanford University Press, 1990.
+
+\leavevmode\vadjust pre{\hypertarget{ref-boutyline2017}{}}%
+Boutyline, Andrei, and Stephen Vaisey. 2017. {``Belief {Network}
+{Analysis}: {A} {Relational} {Approach} to {Understanding} the
+{Structure} of {Attitudes}.''} \emph{American Journal of Sociology}
+122(5):1371--1447. doi:
+\href{https://doi.org/10.1086/691274}{10.1086/691274}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-brocic2021}{}}%
+Broćić, Miloš, and Andrew Miles. 2021. {``College and the {`{Culture}
+{War}'}: {Assessing} {Higher} {Education}'s {Influence} on {Moral}
+{Attitudes}.''} \emph{American Sociological Review} 00031224211041094.
+doi:
+\href{https://doi.org/10.1177/00031224211041094}{10.1177/00031224211041094}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-campbell2016}{}}%
+Campbell, Colin, and Jonathan Horowitz. 2016. {``Does {College}
+{Influence} {Sociopolitical} {Attitudes}?''} \emph{Sociology of
+Education} 89(1):40--58. doi:
+\href{https://doi.org/10.1177/0038040715617224}{10.1177/0038040715617224}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-christakis2010}{}}%
+Christakis, Nicholas, and James Fowler. 2010. \emph{Connected: {The}
+{Amazing} {Power} of {Social} {Networks} and {How} {They} {Shape} {Our}
+{Lives}}. HarperCollins Publishers.
+
+\leavevmode\vadjust pre{\hypertarget{ref-dellaposta2018}{}}%
+DellaPosta, Daniel. 2018. {``Gay {Acquaintanceship} and {Attitudes}
+Toward {Homosexuality}: {A} {Conservative} {Test}.''} \emph{Socius}
+4:2378023118798959. doi:
+\href{https://doi.org/10.1177/2378023118798959}{10.1177/2378023118798959}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-dellaposta2020}{}}%
+DellaPosta, Daniel. 2020. {``Pluralistic {Collapse}: {The} {`{Oil}
+{Spill}'} {Model} of {Mass} {Opinion} {Polarization}.''} \emph{American
+Sociological Review} 85(3):507--36. doi:
+\href{https://doi.org/10.1177/0003122420922989}{10.1177/0003122420922989}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-eaton2009}{}}%
+Eaton, Asia A., Penny S. Visser, Jon A. Krosnick, and Sowmya Anand.
+2009. {``Social {Power} and {Attitude} {Strength} {Over} the {Life}
+{Course}.''} \emph{Personality and Social Psychology Bulletin}
+35(12):1646--60. doi:
+\href{https://doi.org/10.1177/0146167209349114}{10.1177/0146167209349114}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-elder2003}{}}%
+Elder, Glen H., Monica Kirkpatrick Johnson, and Robert Crosnoe. 2003.
+{``\href{https://doi.org/10.1007/978-0-306-48247-2_1}{The {Emergence}
+and {Development} of {Life} {Course} {Theory}}.''} Pp. 3--19 in
+\emph{Handbook of the {Life} {Course}}, \emph{Handbooks of {Sociology}
+and {Social} {Research}}, edited by J. T. Mortimer and M. J. Shanahan.
+Boston, MA: Springer US.
+
+\leavevmode\vadjust pre{\hypertarget{ref-feldman1992}{}}%
+Feldman, Stanley, and John Zaller. 1992. {``The {Political} {Culture} of
+{Ambivalence}: {Ideological} {Responses} to the {Welfare} {State}.''}
+\emph{American Journal of Political Science} 36(1):268--307. doi:
+\href{https://doi.org/10.2307/2111433}{10.2307/2111433}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-gelman2021}{}}%
+Gelman, Andrew, and Yotam Margalit. 2021. {``Social Penumbras Predict
+Political Attitudes.''} \emph{Proceedings of the National Academy of
+Sciences} 118(6):e2019375118. doi:
+\href{https://doi.org/10.1073/pnas.2019375118}{10.1073/pnas.2019375118}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-data_soep}{}}%
+Goebel, Jan, Markus M. Grabka, Stefan Liebig, Martin Kroh, David
+Richter, Carsten Schröder, and Jürgen Schupp. 2019. {``The German
+Socio-Economic Panel (SOEP).''} \emph{Jahrb{ü}cher f{ü}r
+National{ö}konomie Und Statistik} 239(2):345--60.
+
+\leavevmode\vadjust pre{\hypertarget{ref-goldberg2018}{}}%
+Goldberg, Amir, and Sarah K. Stein. 2018. {``Beyond {Social}
+{Contagion}: {Associative} {Diffusion} and the {Emergence} of {Cultural}
+{Variation}.''} \emph{American Sociological Review} 83(5):897--932. doi:
+\href{https://doi.org/10.1177/0003122418797576}{10.1177/0003122418797576}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-gross2009}{}}%
+Gross, Neil. 2009. {``A {Pragmatist} {Theory} of {Social}
+{Mechanisms}.''} \emph{American Sociological Review} 74(3):358--79. doi:
+\href{https://doi.org/10.1177/000312240907400302}{10.1177/000312240907400302}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-guhin2020}{}}%
+Guhin, Jeffrey, Jessica McCrory Calarco, and Cynthia Miller-Idriss.
+2020. \emph{Whatever {Happened} to {Socialization}?} \emph{preprint}.
+SocArXiv. doi:
+\href{https://doi.org/10.31235/osf.io/zp2wy}{10.31235/osf.io/zp2wy}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-hout2016}{}}%
+Hout, Michael, and Orestes P. Hastings. 2016. {``Reliability of the
+{Core} {Items} in the {General} {Social} {Survey}: {Estimates} from the
+{Three}-{Wave} {Panels}, 2006--2014.''} \emph{Sociological Science}
+3:971--1002. doi:
+\href{https://doi.org/10.15195/v3.a43}{10.15195/v3.a43}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-data_psid}{}}%
+Income Dynamics, Panel Study of. 2013. {``Public Use Dataset.''}
+\emph{The Survey Research Center Institute for Social Research,
+University of Michgan}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-kiley2020}{}}%
+Kiley, Kevin, and Stephen Vaisey. 2020. {``Measuring {Stability} and
+{Change} in {Personal} {Culture} {Using} {Panel} {Data}.''}
+\emph{American Sociological Review} 85(3):477--506. doi:
+\href{https://doi.org/10.1177/0003122420921538}{10.1177/0003122420921538}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-kratz2021}{}}%
+Kratz, Fabian. 2021. {``Do Concerns about Immigration Change After
+Adolescence? How Education and Critical Life Events Affect Concerns
+about Immigration.''} \emph{European Sociological Review}
+37(6):987--1003.
+
+\leavevmode\vadjust pre{\hypertarget{ref-krosnick1989}{}}%
+Krosnick, Jon, and Duane F. Alwin. 1989. {``Aging and {Susceptibility}
+to {Attitude} {Change}.''} \emph{Journal of Personality and Social
+Psychology} 57:416--25. doi:
+\href{https://doi.org/10.1037//0022-3514.57.3.416}{10.1037//0022-3514.57.3.416}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-lersch2023}{}}%
+Lersch, Philipp M. 2023. {``Change in {Personal} {Culture} over the
+{Life} {Course}.''} \emph{American Sociological Review} 88(2):220--51.
+doi:
+\href{https://doi.org/10.1177/00031224231156456}{10.1177/00031224231156456}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-lizardo2017}{}}%
+Lizardo, Omar. 2017. {``Improving {Cultural} {Analysis}: {Considering}
+{Personal} {Culture} in Its {Declarative} and {Nondeclarative}
+{Modes}.''} \emph{American Sociological Review} 82(1):88--115. doi:
+\href{https://doi.org/10.1177/0003122416675175}{10.1177/0003122416675175}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-longest2013}{}}%
+Longest, K. C., S. Hitlin, and S. Vaisey. 2013. {``Position and
+{Disposition}: {The} {Contextual} {Development} of {Human} {Values}.''}
+\emph{Social Forces} 91(4):1499--1528. doi:
+\href{https://doi.org/10.1093/sf/sot045}{10.1093/sf/sot045}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-mannheim1952}{}}%
+Mannheim, Karl. 1952. {``The {Problem} of {Generations}.''} Pp. 276--322
+in \emph{Essays on the {Sociology} of {Knowledge}}. New York: Oxford
+University Press.
+
+\leavevmode\vadjust pre{\hypertarget{ref-mewes2021}{}}%
+Mewes, Jan, Malcolm Fairbrother, Giuseppe Nicola Giordano, Cary Wu, and
+Rima Wilkes. 2021. {``Experiences Matter: A Longitudinal Study of
+Individual-Level Sources of Declining Social Trust in the United
+States.''} \emph{Social Science Research} 95:102537.
+
+\leavevmode\vadjust pre{\hypertarget{ref-quinn2023}{}}%
+Quinn, Joseph M., Robert E. Freeland, Kimberly B. Rogers, Jesse Hoey,
+and Lynn Smith-Lovin. 2023. {``How Cultural Meanings of Occupations in
+the u.s. Changed During the Covid-19 Pandemic.''} \emph{American
+Behavioral Scientist} 67(1):125--47.
+
+\leavevmode\vadjust pre{\hypertarget{ref-raftery1995}{}}%
+Raftery, Adrian E. 1995. {``Bayesian {Model} {Selection} in {Social}
+{Research}.''} \emph{Sociological Methodology} 25:111--63. doi:
+\href{https://doi.org/10.2307/271063}{10.2307/271063}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-robinson2007}{}}%
+Robinson, Dawn T. 2007. {``Control {Theories} in {Sociology}.''}
+\emph{Annual Review of Sociology} 33(1):157--74. doi:
+\href{https://doi.org/10.1146/annurev.soc.32.061604.123110}{10.1146/annurev.soc.32.061604.123110}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-ryder1965}{}}%
+Ryder, Norman B. 1965. {``The {Cohort} as a {Concept} in the {Study} of
+{Social} {Change}.''} \emph{American Sociological Review} 30(6):843--61.
+doi: \href{https://doi.org/10.2307/2090964}{10.2307/2090964}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-scoville2022}{}}%
+Scoville, Caleb, Andrew McCumber, Razvan Amironesei, and June Jeon.
+2022. {``Mask {Refusal} {Backlash}: {The} {Politicization} of {Face}
+{Masks} in the {American} {Public} {Sphere} During the {Early} {Stages}
+of the {COVID}-19 {Pandemic}.''} \emph{Socius: Sociological Research for
+a Dynamic World} 8:22.
+
+\leavevmode\vadjust pre{\hypertarget{ref-slothuus2021}{}}%
+Slothuus, Rune, and Martin Bisgaard. 2021. {``How {Political} {Parties}
+{Shape} {Public} {Opinion} in the {Real} {World}.''} \emph{American
+Journal of Political Science} 65(4):896--911. doi:
+\href{https://doi.org/10.1111/AJPS.12550}{10.1111/AJPS.12550}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-data_gss}{}}%
+Smith, Tom W., Michael Davern, Jeremy Freese, and Stephen L. Morgan.
+2022. {``General Social Surveys, 1972-2022: Cumulative Codebook /
+Principal Investigator, Tom w. Smith; Co-Principal Investigators,
+Michael Davern, Jeremy Freese and Stephen l. Morgan.''} \emph{Chicago:
+NORC}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-smith-lovin1988}{}}%
+Smith-Lovin, Lynn, and David R. Heise. 1988. \emph{Analyzing {Social}
+{Interaction}: {Advances} in {Affect} {Control} {Theory}}. New York:
+Gordon; Breach Sciences Publishers.
+
+\leavevmode\vadjust pre{\hypertarget{ref-stewart2023}{}}%
+Stewart, Evan, and Diane Berkman. 2023. {``Shifting or Settled? Tracking
+Racial Animus During COVID-19.''} \emph{Social Psychology Quarterly}
+86(3):312--33.
+
+\leavevmode\vadjust pre{\hypertarget{ref-data_hilda}{}}%
+Summerfield, Michelle, Ross Dunn, Simon Freidin, Markus Hahn, Peter
+Ittak, Milica Kecmanovic, Ning Li, Ninette Macalalad, Nicole Watson,
+Roger Wilkins, and others. 2011. {``HILDA User Manual -- Release 10.''}
+\emph{Melbourne Institute of Applied Economic and Social Research,
+University of Melbourne}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-swidler2001}{}}%
+Swidler, Ann. 2001. \emph{Talk of {Love}: {How} {Culture} {Matters}}.
+Chicago: University of Chicago Press.
+
+\leavevmode\vadjust pre{\hypertarget{ref-data_bhps}{}}%
+Taylor, Marcia Freed. 1996. {``British Household Panel Survey User
+Manual: Volume a-Introduction, Technical Report and Appendices.''}
+
+\leavevmode\vadjust pre{\hypertarget{ref-tourangeau2000}{}}%
+Tourangeau, Roger, Lance J. Rips, and Kenneth A. Rasinski. 2000.
+\emph{The Psychology of Survey Response}. Cambridge, U.K. ; New York:
+Cambridge University Press.
+
+\leavevmode\vadjust pre{\hypertarget{ref-underwood2022}{}}%
+Underwoood, Ted, Kevin Kiley, Wenyi Shang, and Stephen Vaisey. 2022.
+{``Cohort Succession Explains Most Change in Literary Culture.''}
+\emph{Sociological Science} 9:184--205.
+
+\leavevmode\vadjust pre{\hypertarget{ref-data_ukhls}{}}%
+University of Essex, Institute for Social, and Economic Research. 2019.
+{``Understanding Society: Waves 1--9, 2009--2018 and Harmonised BHPS:
+Waves 1--18, 1991--2009.''} \emph{UK Data Service}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-vaisey2021}{}}%
+Vaisey, Stephen, and Kevin Kiley. 2021. {``A {Model}-{Based} {Method}
+for {Detecting} {Persistent} {Cultural} {Change} {Using} {Panel}
+{Data}.''} \emph{Sociological Science} 8:83--95. doi:
+\href{https://doi.org/10.15195/v8.a5}{10.15195/v8.a5}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-vaisey2016}{}}%
+Vaisey, Stephen, and Omar Lizardo. 2016. {``Cultural {Fragmentation} or
+{Acquired} {Dispositions}? {A} {New} {Approach} to {Accounting} for
+{Patterns} of {Cultural} {Change}.''} \emph{Socius} 2:2378023116669726.
+doi:
+\href{https://doi.org/10.1177/2378023116669726}{10.1177/2378023116669726}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-visser2004}{}}%
+Visser, Penny S., and Robert R. Mirabile. 2004. {``Attitudes in the
+Social Context: {The} Impact of Social Network Composition on
+Individual-Level Attitude Strength.''} \emph{Journal of Personality and
+Social Psychology} 779--95.
+
+\leavevmode\vadjust pre{\hypertarget{ref-data_shp}{}}%
+Voorpostel, Marieke, Robin Tillmann, Florence Lebert, Ursina Kuhn,
+Oliver Lipps, Valérie-Anne Ryser, and B. Wernli. 2016. {``Swiss
+Household Panel User Guide (1999--2015).''} \emph{Lausanne: FORS}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-weeden2005}{}}%
+Weeden, Kim A., and David B. Grusky. 2005. {``The {Case} for a {New}
+{Class} {Map}.''} \emph{American Journal of Sociology} 111(1):141--212.
+doi: \href{https://doi.org/10.1086/428815}{10.1086/428815}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-wodtke2018}{}}%
+Wodtke, Geoffrey T. 2018. {``The {Effects} of {Education} on {Beliefs}
+about {Racial} {Inequality}.''} \emph{Social Psychology Quarterly}
+81(4):273--94. doi:
+\href{https://doi.org/10.1177/0190272518804145}{10.1177/0190272518804145}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-zaller1992}{}}%
+Zaller, John. 1992.
+\emph{\href{https://doi.org/10.1017/CBO9780511818691}{The {Nature} and
+{Origins} of {Mass} {Opinion}}}. Cambridge: Cambridge University Press.
+
+\end{CSLReferences}
+
 \hypertarget{appendix}{%
@@
 
-\hypertarget{references}{%
-\section{References}\label{references}}
-
-\hypertarget{refs}{}
-\begin{CSLReferences}{1}{0}
-\leavevmode\vadjust pre{\hypertarget{ref-alvarez2002}{}}%
-Alvarez, R. Michael, and John Brehm. 2002. \emph{Hard Choices, Easy
-Answers: Values, Information, and {American} Public Opinion}. Princeton,
-N.J: Princeton University Press.
-
-\leavevmode\vadjust pre{\hypertarget{ref-alwin2007}{}}%
-Alwin, Duane F. 2007. \emph{Margins of {Error}: {A} {Study} of
-{Reliability} in {Survey} {Measurement}}. Hoboken, N.J.: John Wiley \&
-Sons.
-
-\leavevmode\vadjust pre{\hypertarget{ref-alwin1991}{}}%
-Alwin, Duane F., and Jon A. Krosnick. 1991.
-{``\href{https://www.jstor.org/stable/2781642}{Aging, {Cohorts}, and the
-{Stability} of {Sociopolitical} {Orientations} {Over} the {Life}
-{Span}}.''} \emph{American Journal of Sociology} 97(1):169--95.
-
-\leavevmode\vadjust pre{\hypertarget{ref-bail2018}{}}%
-Bail, Christopher A., Lisa P. Argyle, Taylor W. Brown, John P. Bumpus,
-Haohan Chen, M. B. Fallin Hunzaker, Jaemin Lee, Marcus Mann, Friedolin
-Merhout, and Alexander Volfovsky. 2018. {``Exposure to Opposing Views on
-Social Media Can Increase Political Polarization.''} \emph{Proceedings
-of the National Academy of Sciences} 115(37):9216--21. doi:
-\href{https://doi.org/10.1073/pnas.1804840115}{10.1073/pnas.1804840115}.
-
-\leavevmode\vadjust pre{\hypertarget{ref-bardi2009}{}}%
-Bardi, Anat, Julie Anne Lee, Nadi Hofmann-Towfigh, and Geoffrey Soutar.
-2009. {``The Structure of Intraindividual Value Change.''} \emph{Journal
-of Personality and Social Psychology} 97(5):913--29.
-
-\leavevmode\vadjust pre{\hypertarget{ref-bourdieu1990}{}}%
-Bourdieu, Pierre. 1990. \emph{The Logic of Practice}. Stanford, Calif. :
-Stanford University Press, 1990.
-
-\leavevmode\vadjust pre{\hypertarget{ref-boutyline2017}{}}%
-Boutyline, Andrei, and Stephen Vaisey. 2017. {``Belief {Network}
-{Analysis}: {A} {Relational} {Approach} to {Understanding} the
-{Structure} of {Attitudes}.''} \emph{American Journal of Sociology}
-122(5):1371--1447. doi:
-\href{https://doi.org/10.1086/691274}{10.1086/691274}.
-
-\leavevmode\vadjust pre{\hypertarget{ref-brocic2021}{}}%
-Broćić, Miloš, and Andrew Miles. 2021. {``College and the {`{Culture}
-{War}'}: {Assessing} {Higher} {Education}'s {Influence} on {Moral}
-{Attitudes}.''} \emph{American Sociological Review} 00031224211041094.
-doi:
-\href{https://doi.org/10.1177/00031224211041094}{10.1177/00031224211041094}.
-
-\leavevmode\vadjust pre{\hypertarget{ref-campbell2016}{}}%
-Campbell, Colin, and Jonathan Horowitz. 2016. {``Does {College}
-{Influence} {Sociopolitical} {Attitudes}?''} \emph{Sociology of
-Education} 89(1):40--58. doi:
-\href{https://doi.org/10.1177/0038040715617224}{10.1177/0038040715617224}.
-
-\leavevmode\vadjust pre{\hypertarget{ref-christakis2010}{}}%
-Christakis, Nicholas, and James Fowler. 2010. \emph{Connected: {The}
-{Amazing} {Power} of {Social} {Networks} and {How} {They} {Shape} {Our}
-{Lives}}. HarperCollins Publishers.
-
-\leavevmode\vadjust pre{\hypertarget{ref-dellaposta2018}{}}%
-DellaPosta, Daniel. 2018. {``Gay {Acquaintanceship} and {Attitudes}
-Toward {Homosexuality}: {A} {Conservative} {Test}.''} \emph{Socius}
-4:2378023118798959. doi:
-\href{https://doi.org/10.1177/2378023118798959}{10.1177/2378023118798959}.
-
-\leavevmode\vadjust pre{\hypertarget{ref-dellaposta2020}{}}%
-DellaPosta, Daniel. 2020. {``Pluralistic {Collapse}: {The} {`{Oil}
-{Spill}'} {Model} of {Mass} {Opinion} {Polarization}.''} \emph{American
-Sociological Review} 85(3):507--36. doi:
-\href{https://doi.org/10.1177/0003122420922989}{10.1177/0003122420922989}.
-
-\leavevmode\vadjust pre{\hypertarget{ref-eaton2009}{}}%
-Eaton, Asia A., Penny S. Visser, Jon A. Krosnick, and Sowmya Anand.
-2009. {``Social {Power} and {Attitude} {Strength} {Over} the {Life}
-{Course}.''} \emph{Personality and Social Psychology Bulletin}
-35(12):1646--60. doi:
-\href{https://doi.org/10.1177/0146167209349114}{10.1177/0146167209349114}.
-
-\leavevmode\vadjust pre{\hypertarget{ref-elder2003}{}}%
-Elder, Glen H., Monica Kirkpatrick Johnson, and Robert Crosnoe. 2003.
-{``\href{https://doi.org/10.1007/978-0-306-48247-2_1}{The {Emergence}
-and {Development} of {Life} {Course} {Theory}}.''} Pp. 3--19 in
-\emph{Handbook of the {Life} {Course}}, \emph{Handbooks of {Sociology}
-and {Social} {Research}}, edited by J. T. Mortimer and M. J. Shanahan.
-Boston, MA: Springer US.
-
-\leavevmode\vadjust pre{\hypertarget{ref-feldman1992}{}}%
-Feldman, Stanley, and John Zaller. 1992. {``The {Political} {Culture} of
-{Ambivalence}: {Ideological} {Responses} to the {Welfare} {State}.''}
-\emph{American Journal of Political Science} 36(1):268--307. doi:
-\href{https://doi.org/10.2307/2111433}{10.2307/2111433}.
-
-\leavevmode\vadjust pre{\hypertarget{ref-gelman2021}{}}%
-Gelman, Andrew, and Yotam Margalit. 2021. {``Social Penumbras Predict
-Political Attitudes.''} \emph{Proceedings of the National Academy of
-Sciences} 118(6):e2019375118. doi:
-\href{https://doi.org/10.1073/pnas.2019375118}{10.1073/pnas.2019375118}.
-
-\leavevmode\vadjust pre{\hypertarget{ref-data_soep}{}}%
-Goebel, Jan, Markus M. Grabka, Stefan Liebig, Martin Kroh, David
-Richter, Carsten Schröder, and Jürgen Schupp. 2019. {``The German
-Socio-Economic Panel (SOEP).''} \emph{Jahrb{ü}cher f{ü}r
-National{ö}konomie Und Statistik} 239(2):345--60.
-
-\leavevmode\vadjust pre{\hypertarget{ref-goldberg2018}{}}%
-Goldberg, Amir, and Sarah K. Stein. 2018. {``Beyond {Social}
-{Contagion}: {Associative} {Diffusion} and the {Emergence} of {Cultural}
-{Variation}.''} \emph{American Sociological Review} 83(5):897--932. doi:
-\href{https://doi.org/10.1177/0003122418797576}{10.1177/0003122418797576}.
-
-\leavevmode\vadjust pre{\hypertarget{ref-gross2009}{}}%
-Gross, Neil. 2009. {``A {Pragmatist} {Theory} of {Social}
-{Mechanisms}.''} \emph{American Sociological Review} 74(3):358--79. doi:
-\href{https://doi.org/10.1177/000312240907400302}{10.1177/000312240907400302}.
-
-\leavevmode\vadjust pre{\hypertarget{ref-guhin2020}{}}%
-Guhin, Jeffrey, Jessica McCrory Calarco, and Cynthia Miller-Idriss.
-2020. \emph{Whatever {Happened} to {Socialization}?} \emph{preprint}.
-SocArXiv. doi:
-\href{https://doi.org/10.31235/osf.io/zp2wy}{10.31235/osf.io/zp2wy}.
-
-\leavevmode\vadjust pre{\hypertarget{ref-hout2016}{}}%
-Hout, Michael, and Orestes P. Hastings. 2016. {``Reliability of the
-{Core} {Items} in the {General} {Social} {Survey}: {Estimates} from the
-{Three}-{Wave} {Panels}, 2006--2014.''} \emph{Sociological Science}
-3:971--1002. doi:
-\href{https://doi.org/10.15195/v3.a43}{10.15195/v3.a43}.
-
-\leavevmode\vadjust pre{\hypertarget{ref-data_psid}{}}%
-Income Dynamics, Panel Study of. 2013. {``Public Use Dataset.''}
-\emph{The Survey Research Center Institute for Social Research,
-University of Michgan}.
-
-\leavevmode\vadjust pre{\hypertarget{ref-kiley2020}{}}%
-Kiley, Kevin, and Stephen Vaisey. 2020. {``Measuring {Stability} and
-{Change} in {Personal} {Culture} {Using} {Panel} {Data}.''}
-\emph{American Sociological Review} 85(3):477--506. doi:
-\href{https://doi.org/10.1177/0003122420921538}{10.1177/0003122420921538}.
-
-\leavevmode\vadjust pre{\hypertarget{ref-kratz2021}{}}%
-Kratz, Fabian. 2021. {``Do Concerns about Immigration Change After
-Adolescence? How Education and Critical Life Events Affect Concerns
-about Immigration.''} \emph{European Sociological Review}
-37(6):987--1003.
-
-\leavevmode\vadjust pre{\hypertarget{ref-krosnick1989}{}}%
-Krosnick, Jon, and Duane F. Alwin. 1989. {``Aging and {Susceptibility}
-to {Attitude} {Change}.''} \emph{Journal of Personality and Social
-Psychology} 57:416--25. doi:
-\href{https://doi.org/10.1037//0022-3514.57.3.416}{10.1037//0022-3514.57.3.416}.
-
-\leavevmode\vadjust pre{\hypertarget{ref-lersch2023}{}}%
-Lersch, Philipp M. 2023. {``Change in {Personal} {Culture} over the
-{Life} {Course}.''} \emph{American Sociological Review} 88(2):220--51.
-doi:
-\href{https://doi.org/10.1177/00031224231156456}{10.1177/00031224231156456}.
-
-\leavevmode\vadjust pre{\hypertarget{ref-lizardo2017}{}}%
-Lizardo, Omar. 2017. {``Improving {Cultural} {Analysis}: {Considering}
-{Personal} {Culture} in Its {Declarative} and {Nondeclarative}
-{Modes}.''} \emph{American Sociological Review} 82(1):88--115. doi:
-\href{https://doi.org/10.1177/0003122416675175}{10.1177/0003122416675175}.
-
-\leavevmode\vadjust pre{\hypertarget{ref-longest2013}{}}%
-Longest, K. C., S. Hitlin, and S. Vaisey. 2013. {``Position and
-{Disposition}: {The} {Contextual} {Development} of {Human} {Values}.''}
-\emph{Social Forces} 91(4):1499--1528. doi:
-\href{https://doi.org/10.1093/sf/sot045}{10.1093/sf/sot045}.
-
-\leavevmode\vadjust pre{\hypertarget{ref-mannheim1952}{}}%
-Mannheim, Karl. 1952. {``The {Problem} of {Generations}.''} Pp. 276--322
-in \emph{Essays on the {Sociology} of {Knowledge}}. New York: Oxford
-University Press.
-
-\leavevmode\vadjust pre{\hypertarget{ref-mewes2021}{}}%
-Mewes, Jan, Malcolm Fairbrother, Giuseppe Nicola Giordano, Cary Wu, and
-Rima Wilkes. 2021. {``Experiences Matter: A Longitudinal Study of
-Individual-Level Sources of Declining Social Trust in the United
-States.''} \emph{Social Science Research} 95:102537.
-
-\leavevmode\vadjust pre{\hypertarget{ref-quinn2023}{}}%
-Quinn, Joseph M., Robert E. Freeland, Kimberly B. Rogers, Jesse Hoey,
-and Lynn Smith-Lovin. 2023. {``How Cultural Meanings of Occupations in
-the u.s. Changed During the Covid-19 Pandemic.''} \emph{American
-Behavioral Scientist} 67(1):125--47.
-
-\leavevmode\vadjust pre{\hypertarget{ref-raftery1995}{}}%
-Raftery, Adrian E. 1995. {``Bayesian {Model} {Selection} in {Social}
-{Research}.''} \emph{Sociological Methodology} 25:111--63. doi:
-\href{https://doi.org/10.2307/271063}{10.2307/271063}.
-
-\leavevmode\vadjust pre{\hypertarget{ref-robinson2007}{}}%
-Robinson, Dawn T. 2007. {``Control {Theories} in {Sociology}.''}
-\emph{Annual Review of Sociology} 33(1):157--74. doi:
-\href{https://doi.org/10.1146/annurev.soc.32.061604.123110}{10.1146/annurev.soc.32.061604.123110}.
-
-\leavevmode\vadjust pre{\hypertarget{ref-ryder1965}{}}%
-Ryder, Norman B. 1965. {``The {Cohort} as a {Concept} in the {Study} of
-{Social} {Change}.''} \emph{American Sociological Review} 30(6):843--61.
-doi: \href{https://doi.org/10.2307/2090964}{10.2307/2090964}.
-
-\leavevmode\vadjust pre{\hypertarget{ref-scoville2022}{}}%
-Scoville, Caleb, Andrew McCumber, Razvan Amironesei, and June Jeon.
-2022. {``Mask {Refusal} {Backlash}: {The} {Politicization} of {Face}
-{Masks} in the {American} {Public} {Sphere} During the {Early} {Stages}
-of the {COVID}-19 {Pandemic}.''} \emph{Socius: Sociological Research for
-a Dynamic World} 8:22.
-
-\leavevmode\vadjust pre{\hypertarget{ref-slothuus2021}{}}%
-Slothuus, Rune, and Martin Bisgaard. 2021. {``How {Political} {Parties}
-{Shape} {Public} {Opinion} in the {Real} {World}.''} \emph{American
-Journal of Political Science} 65(4):896--911. doi:
-\href{https://doi.org/10.1111/AJPS.12550}{10.1111/AJPS.12550}.
-
-\leavevmode\vadjust pre{\hypertarget{ref-data_gss}{}}%
-Smith, Tom W., Michael Davern, Jeremy Freese, and Stephen L. Morgan.
-2022. {``General Social Surveys, 1972-2022: Cumulative Codebook /
-Principal Investigator, Tom w. Smith; Co-Principal Investigators,
-Michael Davern, Jeremy Freese and Stephen l. Morgan.''} \emph{Chicago:
-NORC}.
-
-\leavevmode\vadjust pre{\hypertarget{ref-smith-lovin1988}{}}%
-Smith-Lovin, Lynn, and David R. Heise. 1988. \emph{Analyzing {Social}
-{Interaction}: {Advances} in {Affect} {Control} {Theory}}. New York:
-Gordon; Breach Sciences Publishers.
-
-\leavevmode\vadjust pre{\hypertarget{ref-stewart2023}{}}%
-Stewart, Evan, and Diane Berkman. 2023. {``Shifting or Settled? Tracking
-Racial Animus During COVID-19.''} \emph{Social Psychology Quarterly}
-86(3):312--33.
-
-\leavevmode\vadjust pre{\hypertarget{ref-data_hilda}{}}%
-Summerfield, Michelle, Ross Dunn, Simon Freidin, Markus Hahn, Peter
-Ittak, Milica Kecmanovic, Ning Li, Ninette Macalalad, Nicole Watson,
-Roger Wilkins, and others. 2011. {``HILDA User Manual -- Release 10.''}
-\emph{Melbourne Institute of Applied Economic and Social Research,
-University of Melbourne}.
-
-\leavevmode\vadjust pre{\hypertarget{ref-swidler2001}{}}%
-Swidler, Ann. 2001. \emph{Talk of {Love}: {How} {Culture} {Matters}}.
-Chicago: University of Chicago Press.
-
-\leavevmode\vadjust pre{\hypertarget{ref-data_bhps}{}}%
-Taylor, Marcia Freed. 1996. {``British Household Panel Survey User
-Manual: Volume a-Introduction, Technical Report and Appendices.''}
-
-\leavevmode\vadjust pre{\hypertarget{ref-tourangeau2000}{}}%
-Tourangeau, Roger, Lance J. Rips, and Kenneth A. Rasinski. 2000.
-\emph{The Psychology of Survey Response}. Cambridge, U.K. ; New York:
-Cambridge University Press.
-
-\leavevmode\vadjust pre{\hypertarget{ref-underwood2022}{}}%
-Underwoood, Ted, Kevin Kiley, Wenyi Shang, and Stephen Vaisey. 2022.
-{``Cohort Succession Explains Most Change in Literary Culture.''}
-\emph{Sociological Science} 9:184--205.
-
-\leavevmode\vadjust pre{\hypertarget{ref-data_ukhls}{}}%
-University of Essex, Institute for Social, and Economic Research. 2019.
-{``Understanding Society: Waves 1--9, 2009--2018 and Harmonised BHPS:
-Waves 1--18, 1991--2009.''} \emph{UK Data Service}.
-
-\leavevmode\vadjust pre{\hypertarget{ref-vaisey2021}{}}%
-Vaisey, Stephen, and Kevin Kiley. 2021. {``A {Model}-{Based} {Method}
-for {Detecting} {Persistent} {Cultural} {Change} {Using} {Panel}
-{Data}.''} \emph{Sociological Science} 8:83--95. doi:
-\href{https://doi.org/10.15195/v8.a5}{10.15195/v8.a5}.
-
-\leavevmode\vadjust pre{\hypertarget{ref-vaisey2016}{}}%
-Vaisey, Stephen, and Omar Lizardo. 2016. {``Cultural {Fragmentation} or
-{Acquired} {Dispositions}? {A} {New} {Approach} to {Accounting} for
-{Patterns} of {Cultural} {Change}.''} \emph{Socius} 2:2378023116669726.
-doi:
-\href{https://doi.org/10.1177/2378023116669726}{10.1177/2378023116669726}.
-
-\leavevmode\vadjust pre{\hypertarget{ref-visser2004}{}}%
-Visser, Penny S., and Robert R. Mirabile. 2004. {``Attitudes in the
-Social Context: {The} Impact of Social Network Composition on
-Individual-Level Attitude Strength.''} \emph{Journal of Personality and
-Social Psychology} 779--95.
-
-\leavevmode\vadjust pre{\hypertarget{ref-data_shp}{}}%
-Voorpostel, Marieke, Robin Tillmann, Florence Lebert, Ursina Kuhn,
-Oliver Lipps, Valérie-Anne Ryser, and B. Wernli. 2016. {``Swiss
-Household Panel User Guide (1999--2015).''} \emph{Lausanne: FORS}.
-
-\leavevmode\vadjust pre{\hypertarget{ref-weeden2005}{}}%
-Weeden, Kim A., and David B. Grusky. 2005. {``The {Case} for a {New}
-{Class} {Map}.''} \emph{American Journal of Sociology} 111(1):141--212.
-doi: \href{https://doi.org/10.1086/428815}{10.1086/428815}.
-
-\leavevmode\vadjust pre{\hypertarget{ref-wodtke2018}{}}%
-Wodtke, Geoffrey T. 2018. {``The {Effects} of {Education} on {Beliefs}
-about {Racial} {Inequality}.''} \emph{Social Psychology Quarterly}
-81(4):273--94. doi:
-\href{https://doi.org/10.1177/0190272518804145}{10.1177/0190272518804145}.
-
-\leavevmode\vadjust pre{\hypertarget{ref-zaller1992}{}}%
-Zaller, John. 1992.
-\emph{\href{https://doi.org/10.1017/CBO9780511818691}{The {Nature} and
-{Origins} of {Mass} {Opinion}}}. Cambridge: Cambridge University Press.
-
-\end{CSLReferences}
-
 \end{document}

### 2023-09-30 manuscript/manuscript.Rmd
minor revisions in the write-up.

@@
 
-Because all of these perspectives have some empirical support, the current debate is not about whether the processes they posit exist at all, but about their relative contribution to explaining the cultural differences we see in the world.
+Because all of these perspectives have some empirical support, the theoretical debate is not about whether the processes they posit exist at all, but about their relative contribution to explaining the cultural differences we see in the world.
 Nevertheless, researchers have struggled to reach a consensus on the importance of cultural change during adulthood.
@@
 Other studies are compatible with these findings as well.
-For example, Brocic and Miles [-@brocic2021] estimate that completing graduate degrees in humanities, arts, and social sciences shift peoples' moral relativism only about .2 standard deviations on average compared to people with no college degree, and this is the largest effect they identify.
+For example, Brocic and Miles [-@brocic2021] estimate that completing graduate degrees in humanities, arts, and social sciences shift peoples' moral relativism only about 0.2 standard deviations on average compared to people with no college degree, and this is the largest effect they identify.
 And studies of aggregate change show that, even on items where cohort effects explain more variance than period effects, there is always evidence that some people change over time [@vaisey2016].
@@
 
-To this point, the theoretical debate has been about whether people *ever* change.
+To this point, the debate has been about whether people *ever* change.
 As we argued above, this is rarely the right question since the answer is almost always "yes." Over sufficient time and in a large enough sample, researchers will likely observe some evidence for durable individual change.
@@
 To address this, we followed Hout et al.
-[-@hout2016] and Lersch [-@lersch2023] in coding each question as falling into one of nine different content categories and included these as indicator variables in the meta analysis.
-Supplemental Materials B presents the estimates for these content indicators.
+[-@hout2016] and Lersch [-@lersch2023] in coding each question as falling into one of nine different topical domains and included these as indicator variables in the meta analysis.
+Supplemental Materials B presents the estimates for these topic indicators.
 They show some associations with $\omega$, with questions about subjective SES; social life, social cohesion, and trust; environment and climate; and health and morale showing larger $\omega$ values than questions about religion and spirituality; politics, government and the economy; and gender and family.
 
-The most notable result in the meta analysis is the negative coefficient attached to the duration of years covered by the question.
+The most notable result in the meta analysis is the negative coefficient attached to the duration of years covered by the question on average across participants.
 The longer a question is observed, the less is the value of $\omega$.
@@
 
-As a further test of this finding, we compared the $\omega$ values when using the full duration of a panel compared to when we dropped the final wave and therefore reduced the total duration of observation for the question.
+As a further test of this finding, we compared the $\omega$ values when using the full duration of a panel compared to when we dropped the final wave for each participant and therefore reduced the total duration of observation for the question.
 If the coefficient reflects a true negative effect of duration on $\omega$, we should see that same effect within questions.
@@
 There are also large differences in $\omega$ values across questions for both groups.
-For example, on the question of whether the government should try to solve more problems or leave those problems to be solved by private businesses ("government do more or less"), less than one percent of the systematic variance is attributable to intrapersonal change for both groups.
+For example, on the question of whether the government should try to solve more problems or leave those problems to be solved by private businesses ("government do more or less"), less than 1 percent of the systematic variance is attributable to intrapersonal change for both groups.
 In other words, while people might vacillate on this question at random (37 percent of variance is residual for this question), there is functionally no evidence that people make systematic changes of opinion on this issue during the GSS panel.
@@
 Figure A2 presents the alternative strategy we used to understand the effects of duration on intraindividual change.
-First, we restricted the sample such that we have at least 4 waves of observations for all participants.
-This reduced the number of items to 249.
-We then fitted the Life Course Adaption models to this sample.
-Second, we removed the last wave from all the observations, and refit the same models once more.
+First, we removed the last wave from all the observations and fitted the Life Course Adaption models.
+This effectively reduced the number of items to 249.
+We then refitted the model for all participants in this sample using the unrestricted data.
 In the final step, we calculated the $\omega$ values for each item.
-The Figure C1 shows a scatterplot for these two sets of observations.
+The Figure A2 shows a scatterplot for these two sets of observations.
 

### 2023-09-30 manuscript/manuscript.tex
minor revisions in the write-up.

@@
 Because all of these perspectives have some empirical support, the
-current debate is not about whether the processes they posit exist at
-all, but about their relative contribution to explaining the cultural
+theoretical debate is not about whether the processes they posit exist
+at all, but about their relative contribution to explaining the cultural
 differences we see in the world. Nevertheless, researchers have
@@
 humanities, arts, and social sciences shift peoples' moral relativism
-only about .2 standard deviations on average compared to people with no
+only about 0.2 standard deviations on average compared to people with no
 college degree, and this is the largest effect they identify. And
@@
 
-To this point, the theoretical debate has been about whether people
-\emph{ever} change. As we argued above, this is rarely the right
-question since the answer is almost always ``yes.'' Over sufficient time
-and in a large enough sample, researchers will likely observe some
-evidence for durable individual change. A lack of evidence for change
-may be due to a poor survey question, a low resolution in response
-options, or that the question simply was not asked for long enough or to
-enough people. Conversely, finding evidence for some change tells us
-only that it was ``not zero.'' From that alone, we learn little about
-how much intrapersonal change has happened in a population.
+To this point, the debate has been about whether people \emph{ever}
+change. As we argued above, this is rarely the right question since the
+answer is almost always ``yes.'' Over sufficient time and in a large
+enough sample, researchers will likely observe some evidence for durable
+individual change. A lack of evidence for change may be due to a poor
+survey question, a low resolution in response options, or that the
+question simply was not asked for long enough or to enough people.
+Conversely, finding evidence for some change tells us only that it was
+``not zero.'' From that alone, we learn little about how much
+intrapersonal change has happened in a population.
 
@@
 followed Hout et al. (2016) and Lersch (2023) in coding each question as
-falling into one of nine different content categories and included these
-as indicator variables in the meta analysis. Supplemental Materials B
-presents the estimates for these content indicators. They show some
+falling into one of nine different topical domains and included these as
+indicator variables in the meta analysis. Supplemental Materials B
+presents the estimates for these topic indicators. They show some
 associations with \(\omega\), with questions about subjective SES;
@@
 The most notable result in the meta analysis is the negative coefficient
-attached to the duration of years covered by the question. The longer a
-question is observed, the less is the value of \(\omega\). Theories of
-personal cultural change that link changes in personal culture to social
-experiences, including the LCAM, suggest that the longer we observe
-respondents, the more likely people are to undergo potentially
-transformative experiences and therefore the more variance would be
-attributable to intrapersonal change. Finding a negative coefficient
-here seems to challenge that assumption.
+attached to the duration of years covered by the question on average
+across participants. The longer a question is observed, the less is the
+value of \(\omega\). Theories of personal cultural change that link
+changes in personal culture to social experiences, including the LCAM,
+suggest that the longer we observe respondents, the more likely people
+are to undergo potentially transformative experiences and therefore the
+more variance would be attributable to intrapersonal change. Finding a
+negative coefficient here seems to challenge that assumption.
 
@@
 when using the full duration of a panel compared to when we dropped the
-final wave and therefore reduced the total duration of observation for
-the question. If the coefficient reflects a true negative effect of
-duration on \(\omega\), we should see that same effect within questions.
-Results from this analysis are presented in Appendix B and contradict
-the coefficient from the regression model; with a few exceptions, we
-found that the longer we observe the same question, the higher
-\(\omega\) is. We interpret this combination of findings as suggesting
-that the kinds of questions asked for a longer time period tend to
-demonstrate less intrapersonal change than questions asked over shorter
-periods, rather than a true function of time. This supports the
-conclusion that the GSS shows less intrapersonal change because of its
-shorter duration.
+final wave for each participant and therefore reduced the total duration
+of observation for the question. If the coefficient reflects a true
+negative effect of duration on \(\omega\), we should see that same
+effect within questions. Results from this analysis are presented in
+Appendix B and contradict the coefficient from the regression model;
+with a few exceptions, we found that the longer we observe the same
+question, the higher \(\omega\) is. We interpret this combination of
+findings as suggesting that the kinds of questions asked for a longer
+time period tend to demonstrate less intrapersonal change than questions
+asked over shorter periods, rather than a true function of time. This
+supports the conclusion that the GSS shows less intrapersonal change
+because of its shorter duration.
 
@@
 leave those problems to be solved by private businesses (``government do
-more or less''), less than one percent of the systematic variance is
+more or less''), less than 1 percent of the systematic variance is
 attributable to intrapersonal change for both groups. In other words,
@@
 Figure A2 presents the alternative strategy we used to understand the
-effects of duration on intraindividual change. First, we restricted the
-sample such that we have at least 4 waves of observations for all
-participants. This reduced the number of items to 249. We then fitted
-the Life Course Adaption models to this sample. Second, we removed the
-last wave from all the observations, and refit the same models once
-more. In the final step, we calculated the \(\omega\) values for each
-item. The Figure C1 shows a scatterplot for these two sets of
-observations.
+effects of duration on intraindividual change. First, we removed the
+last wave from all the observations and fitted the Life Course Adaption
+models. This effectively reduced the number of items to 249. We then
+refitted the model for all participants in this sample using the
+unrestricted data. In the final step, we calculated the \(\omega\)
+values for each item. The Figure A2 shows a scatterplot for these two
+sets of observations.
 

### 2023-10-02 manuscript/manuscript.Rmd
moved note

@@
 
-To compare these competing propositions, we calculate $\omega$ values separately for people with at least a bachelor's degree and people with less than a bachelor's degree at wave 1[^4] of the three General Social Survey's panels.
-We focus on the GSS because it contains the largest number of questions tapping general political dispositions, which is the domain where education has proven particularly relevant for understanding attitude stability.
+To compare these competing propositions, we calculate $\omega$ values separately for people with at least a bachelor's degree and people with less than a bachelor's degree at wave 1 of the three General Social Survey's panels.[^4] We focus on the GSS because it contains the largest number of questions tapping general political dispositions, which is the domain where education has proven particularly relevant for understanding attitude stability.
 The GSS also covers a turbulent window of American politics from 2006 to 2014.

### 2023-10-02 manuscript/manuscript.tex
moved note

@@
 separately for people with at least a bachelor's degree and people with
-less than a bachelor's degree at wave 1\footnote{A small number of
-  respondents report different highest degrees at wave 1, wave 2, and
-  wave 3. Some of this is due to measurement error, and some of it is
-  due to a small number of people obtaining a higher degree during the
-  four years of the panel. Estimating the panel with highest at wave 1
-  or highest degree reported across the panel produces functionally
-  identical results.} of the three General Social Survey's panels. We
-focus on the GSS because it contains the largest number of questions
-tapping general political dispositions, which is the domain where
-education has proven particularly relevant for understanding attitude
-stability. The GSS also covers a turbulent window of American politics
-from 2006 to 2014. This window covers the start of the Great Recession,
-debates about federal intervention in and regulation of Wall Street, the
-election of Barack Obama as the first black U.S. president, debates
-about the role of the federal government in the health care sector, the
-emergence of the Tea Party, and political realignment and clarification
-on the issue of gay marriage, among other topics.
+less than a bachelor's degree at wave 1 of the three General Social
+Survey's panels.\footnote{A small number of respondents report different
+  highest degrees at wave 1, wave 2, and wave 3. Some of this is due to
+  measurement error, and some of it is due to a small number of people
+  obtaining a higher degree during the four years of the panel.
+  Estimating the panel with highest at wave 1 or highest degree reported
+  across the panel produces functionally identical results.} We focus on
+the GSS because it contains the largest number of questions tapping
+general political dispositions, which is the domain where education has
+proven particularly relevant for understanding attitude stability. The
+GSS also covers a turbulent window of American politics from 2006 to
+2014. This window covers the start of the Great Recession, debates about
+federal intervention in and regulation of Wall Street, the election of
+Barack Obama as the first black U.S. president, debates about the role
+of the federal government in the health care sector, the emergence of
+the Tea Party, and political realignment and clarification on the issue
+of gay marriage, among other topics.
 

### 2023-10-02 manuscript/manuscript.Rmd
format changes.

@@
   pdf_document:
+    latex_engine: xelatex
     keep_tex: true
 geometry: margin=1in 
-mainfont: Times New Roman
 fontsize: 12pt
 header-includes:
+  \usepackage{fontspec}
+  \setmainfont{Times New Roman}
+  \setlength{\parindent}{3em}
   \usepackage{caption}
@@
   \let\footnote=\endnote
+  \pagenumbering{gobble}
 bibliography: kileybib.bib
@@
 
-```{r, message = FALSE}
-
-library(tidyverse)
-library(knitr)
-library(kableExtra)
-
-table <- tibble(
-  Country = c("Australia", "Germany", "Great Britian", "Great Britian", "Switzerland", "United States", "United States"),
-  Survey = c("Household, Income and Labor Dynamics in Australia (HILDA)",
-             "Socio-Economic Panel Study (SOEP)",
-             "British Household Panel Survey (BHPS)",
-             "Understanding Society/UK Household Longitudinal Study (UKHLS)",
-             "Swiss Household Panel (SHP)",
-             "General Social Survey (GSS)",
-             "Panel Study of Income Dynamics (PSID)"),
-  Period = c("2001-2021", "1984-2020", "1991-2008", "2009-2020", "1999-2019", "2006-2012", "1968-2019"),
-  Outcomes = c(30, 122, 83, 66, 77, 183, 48)
-)
-
-kableExtra::kable(table,
-                  booktabs = TRUE,
-                  format = "latex",
-                  align = "llcc",
-                  caption = "The Description of the Data Sources") |> 
-  kable_styling(latex_options = c("scale_down", "HOLD_position"))
-
+```{=tex}
+\begin{center}
+<Table 1 about here>
+\end{center}
 ```
@@
 
-```{=latex}
-\begin{figure}[htp]
+```{=tex}
 \begin{center}
-\caption{Proportions of Systematic Variance in Personal Culture}
-```
-```{r fig1, out.width = "100%", fig.pos = "H"}
-knitr::include_graphics("../figures/figure_1.png")
+<Figure 1 about here>
+\end{center}
 ```
 
-```{=latex}
-\end{center}
-\footnotesize{\textit{Notes:} The figure shows $\omega$ and 1 - $\omega$ as the proportion of systematic variance attributable to intrapersonal change and interpersonal differences.
-See Supplemental Materials A for the full set of item values.}
-\end{figure}
-```
 Across all questions, interpersonal differences account for a much larger share of the systematic variance in responses than intrapersonal change.
@@
 
-```{=latex}
-\begin{figure}[htp]
+```{=tex}
 \begin{center}
-\caption{Regression Model Estimating $\omega$}
-```
-```{r fig2, out.width = "350px"}
-knitr::include_graphics("../figures/figure_2.png")
+<Figure 2 about here>
+\end{center}
 ```
 
-```{=latex}
-\end{center}
-\footnotesize{\textit{Notes:} The model estimates the proportion of systematic variance attributable to intrapersonal change.
-It includes the log of the number of participants, the number of waves the variable was asked (capped at t = 10 for each item), the date of the first wave (measured as the year minus 1968 divided by 10), number of waves the question was asked averaged across participants, duration in years per item averaged across participants, and panel and topic fixed effects.
-Coefficients are estimated from suppressed intercept model based on predictions at participants' wave mid-point.
-Survey indicators and item topics not shown.
-See Supplemental Materials B for the full set of coefficients.}
-\end{figure}
-```
 Figure 2 shows that the more response options respondents are given and the larger the sample, the larger the values of $\omega$.
@@
 
-```{=latex}
-\begin{figure}[hbt]
+```{=tex}
 \begin{center}
-\caption{Difference in $\omega$ Across College Graduates and Non-Graduates}
-```
-```{r fig3, out.width = "450px"}
-knitr::include_graphics("../figures/figure_3.png")
+<Figure 3 about here>
+\end{center}
 ```
 
-```{=latex}
-\end{center}
-\footnotesize{\textit{Notes:} The figure shows the difference in $\omega$ values across college graduates and college non-graduates.
-Values above (below) 0 means that those with college degree have higher (lower) variance of intrapersonal change.
-The dashed red line marks 0 difference.}
-\end{figure}
-```
 There is a clear pattern in Figure 3: for more than 80 percent of these GSS items, intrapersonal change is a larger component of systematic variance for people without a college degree.
@@
 
-```{=latex}
-\begin{figure}[hbt]
+```{=tex}
 \begin{center}
-\caption{Difference in $\omega$ Across College Graduates and Non-Graduates on Political Culture}
-```
-```{r fig4, out.width = "450px"}
-knitr::include_graphics("../figures/figure_4.png")
+<Figure 4 about here>
+\end{center}
 ```
 
-```{=latex}
-\end{center}
-\footnotesize{\textit{Notes:} $\omega$ values across college graduates and college non-graduates on 8 political culture items from the General Social Survey (2006-2014).}
-\end{figure}
-```
 On all eight questions presented in Figure 4, $\omega$ values are smaller for people with a college degree, meaning intrapersonal change is less common among them.
@@
 
+\newpage
+
 # Appendix
@@
 
-```{=latex}
-\begin{figure}[htp]
+```{=tex}
 \begin{center}
-\caption*{Figure A1: The Distribution of V(D) and V(C)}
-```
-```{r figA1, out.width = "350px"}
-knitr::include_graphics("../figures/figure_A1.png")
-```
-```{=latex}
+<Figure A1 about here>
 \end{center}
-\end{figure}
 ```
@@
 
+```{=tex}
+\begin{center}
+<Figure B1 about here>
+\end{center}
+```
+
+\newpage
+
+# Tables and Figures
+
+```{r, message = FALSE}
+
+library(tidyverse)
+library(knitr)
+library(kableExtra)
+
+table <- tibble(
+  Country = c("Australia", "Germany", "Great Britian", "Great Britian", "Switzerland", "United States", "United States"),
+  Survey = c("Household, Income and Labor Dynamics in Australia (HILDA)",
+             "Socio-Economic Panel Study (SOEP)",
+             "British Household Panel Survey (BHPS)",
+             "Understanding Society/UK Household Longitudinal Study (UKHLS)",
+             "Swiss Household Panel (SHP)",
+             "General Social Survey (GSS)",
+             "Panel Study of Income Dynamics (PSID)"),
+  Period = c("2001-2021", "1984-2020", "1991-2008", "2009-2020", "1999-2019", "2006-2012", "1968-2019"),
+  Outcomes = c(30, 122, 83, 66, 77, 183, 48)
+)
+
+kableExtra::kable(table,
+                  booktabs = TRUE,
+                  format = "latex",
+                  align = "llcc",
+                  caption = "The Description of the Data Sources") |> 
+  kable_styling(latex_options = c("scale_down", "HOLD_position"))
+
+```
+
+\newpage
+
 ```{=latex}
-\begin{figure}[tb]
+\begin{figure}[ht]
+\begin{center}
+\caption{Proportions of Systematic Variance in Personal Culture}
+```
+```{r fig1, out.width = "100%", fig.pos = "H"}
+knitr::include_graphics("../figures/figure_1.png")
+```
+
+```{=latex}
+\end{center}
+\footnotesize{\textit{Notes:} The figure shows $\omega$ and 1 - $\omega$ as the proportion of systematic variance attributable to intrapersonal change and interpersonal differences.
+See Supplemental Materials A for the full set of item values.}
+\end{figure}
+```
+
+\newpage
+
+```{=latex}
+\begin{figure}[ht]
+\begin{center}
+\caption{Regression Model Estimating $\omega$}
+```
+```{r fig2}
+knitr::include_graphics("../figures/figure_2.png")
+```
+
+```{=latex}
+\end{center}
+\footnotesize{\textit{Notes:} The model estimates the proportion of systematic variance attributable to intrapersonal change.
+It includes the log of the number of participants, the number of waves the variable was asked (capped at t = 10 for each item), the date of the first wave (measured as the year minus 1968 divided by 10), number of waves the question was asked averaged across participants, duration in years per item averaged across participants, and panel and topic fixed effects.
+Coefficients are estimated from suppressed intercept model based on predictions at participants' wave mid-point.
+Survey indicators and item topics not shown.
+See Supplemental Materials B for the full set of coefficients.}
+\end{figure}
+```
+
+\newpage
+
+```{=latex}
+\begin{figure}[ht]
+\begin{center}
+\caption{Difference in $\omega$ Across College Graduates and Non-Graduates}
+```
+```{r fig3}
+knitr::include_graphics("../figures/figure_3.png")
+```
+
+```{=latex}
+\end{center}
+\footnotesize{\textit{Notes:} The figure shows the difference in $\omega$ values across college graduates and college non-graduates.
+Values above (below) 0 means that those with college degree have higher (lower) variance of intrapersonal change.
+The dashed red line marks 0 difference.}
+\end{figure}
+```
+
+\newpage
+
+```{=latex}
+\begin{figure}[ht]
+\begin{center}
+\caption{Difference in $\omega$ Across College Graduates and Non-Graduates on Political Culture}
+```
+```{r fig4}
+knitr::include_graphics("../figures/figure_4.png")
+```
+
+```{=latex}
+\end{center}
+\footnotesize{\textit{Notes:} $\omega$ values across college graduates and college non-graduates on 8 political culture items from the General Social Survey (2006-2014).}
+\end{figure}
+```
+
+\newpage
+
+```{=latex}
+\begin{figure}[htp]
+\begin{center}
+\caption*{Figure A1: The Distribution of V(D) and V(C)}
+```
+```{r figA1}
+knitr::include_graphics("../figures/figure_A1.png")
+```
+```{=latex}
+\end{center}
+\end{figure}
+```
+
+\newpage
+
+```{=latex}
+\begin{figure}[htp]
 \begin{center}
@@
 ```
-```{r figA2, out.width = "400px"}
+```{r figA2}
 knitr::include_graphics("../figures/figure_A2.png")

### 2023-10-02 supplementals/manuscript_supplementals.tex
page numbers

@@
 \setcounter{secnumdepth}{-\maxdimen} % remove section numbering
-\usepackage{fontspec} \usepackage{longtable} \usepackage{multirow} \usepackage{array} \usepackage{booktabs} \usepackage{dcolumn} \usepackage{lscape} \usepackage{siunitx} \usepackage{caption} \newcommand{\blandscape}{\begin{landscape}} \newcommand{\elandscape}{\end{landscape}}
+\usepackage{fontspec} \setmainfont{Times New Roman} \usepackage{longtable} \usepackage{multirow} \usepackage{array} \usepackage{booktabs} \usepackage{dcolumn} \usepackage{lscape} \usepackage{siunitx} \usepackage{caption} \newcommand{\blandscape}{\begin{landscape}} \newcommand{\elandscape}{\end{landscape}} \pagenumbering{gobble}
 \ifLuaTeX

### 2023-10-11 manuscript/manuscript.Rmd
abstract tweaks

@@
 ---
-title: "Quantifying Stability and Change in Personal Culture Using Panel Data"
+title: "Quantifying The Importance of Change for Understanding Differences in Personal Culture"
 author: 
@@
 csl: asa.csl
-abstract: "Does personal culture---a person's attitudes, beliefs, values, and practices---change over the life course, or is it largely fixed by adulthood?
+abstract: "When researchers want to understand differences in personal culture---a person's attitudes, beliefs, values, and practices--how much attention should they pay to adult experience?
 Recent work has reached substantially different conclusions on this question.
-We argue that this disagreement is an unintended consequence of the \"tournament of models\" approach researchers have used.
-We introduce a new measure for quantifying the proportion of systematic variance in panel data attributable to intrapersonal change versus stable interpersonal differences.
+We argue that this disagreement is an unintended consequence of the \"tournament of models\" approach researchers have used, which focuses on whether people change and not how much they change.
+To advance the theoretical debate, we refocus attention on the relative importance of personal change over time for explaining differences between people.
+We introduce a new measure for quantifying the proportion of systematic variance in panel data attributable to intrapersonal change.
 Applying this measure to 609 items from seven surveys in five countries, we find that although intrapersonal change is common, it is generally small in magnitude.
-We also demonstrate that this measure is useful for comparing social groups by showing that intrapersonal change is less common among U.S.
+As an extension of the theoretical debate, we demonstrate that this measure provides new insights when comparing social groups, showing that intrapersonal change is less common among U.S.
 college graduates than among those without a college degree.
-Overall, we believe this new measure will be a useful methodological tool and help reconcile an important theoretical debate."
+Our findings provide a new perspective on several important theoretical debates, as well as a tool to address new questions."
 ---

### 2023-10-11 manuscript/manuscript.tex
format changes.

@@
   % xetex/luatex font selection
-  \setmainfont[]{Times New Roman}
 \fi
@@
 \newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}
-\usepackage{caption} \usepackage{setspace}\doublespacing \usepackage{endnotes} \renewcommand{\enotesize}{\normalsize} \let\footnote=\endnote
+\usepackage{fontspec} \setmainfont{Times New Roman} \setlength{\parindent}{3em} \usepackage{caption} \usepackage{setspace}\doublespacing \usepackage{endnotes} \renewcommand{\enotesize}{\normalsize} \let\footnote=\endnote \pagenumbering{gobble}
 \usepackage{booktabs}
@@
 \hypersetup{
-  pdftitle={Quantifying Stability and Change in Personal Culture Using Panel Data},
+  pdftitle={Quantifying The Importance of Change for Understanding Differences in Personal Culture},
   hidelinks,
@@
 
-\title{Quantifying Stability and Change in Personal Culture Using Panel
-Data}
+\title{Quantifying The Importance of Change for Understanding
+Differences in Personal Culture}
 \author{}
@@
 \begin{abstract}
-Does personal culture---a person's attitudes, beliefs, values, and
-practices---change over the life course, or is it largely fixed by
-adulthood? Recent work has reached substantially different conclusions
-on this question. We argue that this disagreement is an unintended
-consequence of the ``tournament of models'' approach researchers have
-used. We introduce a new measure for quantifying the proportion of
-systematic variance in panel data attributable to intrapersonal change
-versus stable interpersonal differences. Applying this measure to 609
+When researchers want to understand differences in personal culture---a
+person's attitudes, beliefs, values, and practices--how much attention
+should they pay to adult experience? Recent work has reached
+substantially different conclusions on this question. We argue that this
+disagreement is an unintended consequence of the ``tournament of
+models'' approach researchers have used, which focuses on whether people
+change and not how much they change. To advance the theoretical debate,
+we refocus attention on the relative importance of personal change over
+time for explaining differences between people. We introduce a new
+measure for quantifying the proportion of systematic variance in panel
+data attributable to intrapersonal change. Applying this measure to 609
 items from seven surveys in five countries, we find that although
-intrapersonal change is common, it is generally small in magnitude. We
-also demonstrate that this measure is useful for comparing social groups
-by showing that intrapersonal change is less common among U.S. college
-graduates than among those without a college degree. Overall, we believe
-this new measure will be a useful methodological tool and help reconcile
-an important theoretical debate.
+intrapersonal change is common, it is generally small in magnitude. As
+an extension of the theoretical debate, we demonstrate that this measure
+provides new insights when comparing social groups, showing that
+intrapersonal change is less common among U.S. college graduates than
+among those without a college degree. Our findings provide a new
+perspective on several important theoretical debates, as well as a tool
+to address new questions.
 \end{abstract}
@@
 
-\begin{table}[H]
-
-\caption{\label{tab:unnamed-chunk-1}The Description of the Data Sources}
-\centering
-\resizebox{\linewidth}{!}{
-\begin{tabular}[t]{llcc}
-\toprule
-Country & Survey & Period & Outcomes\\
-\midrule
-Australia & Household, Income and Labor Dynamics in Australia (HILDA) & 2001-2021 & 30\\
-Germany & Socio-Economic Panel Study (SOEP) & 1984-2020 & 122\\
-Great Britian & British Household Panel Survey (BHPS) & 1991-2008 & 83\\
-Great Britian & Understanding Society/UK Household Longitudinal Study (UKHLS) & 2009-2020 & 66\\
-Switzerland & Swiss Household Panel (SHP) & 1999-2019 & 77\\
-\addlinespace
-United States & General Social Survey (GSS) & 2006-2012 & 183\\
-United States & Panel Study of Income Dynamics (PSID) & 1968-2019 & 48\\
-\bottomrule
-\end{tabular}}
-\end{table}
+\begin{center}
+<Table 1 about here>
+\end{center}
 
@@
 
-\begin{figure}[htp]
 \begin{center}
-\caption{Proportions of Systematic Variance in Personal Culture}
-
-\includegraphics[width=1\linewidth]{../figures/figure_1}
-
+<Figure 1 about here>
 \end{center}
-\footnotesize{\textit{Notes:} The figure shows $\omega$ and 1 - $\omega$ as the proportion of systematic variance attributable to intrapersonal change and interpersonal differences.
-See Supplemental Materials A for the full set of item values.}
-\end{figure}
 
@@
 
-\begin{figure}[htp]
 \begin{center}
-\caption{Regression Model Estimating $\omega$}
-
-\includegraphics[width=350px]{../figures/figure_2}
-
+<Figure 2 about here>
 \end{center}
-\footnotesize{\textit{Notes:} The model estimates the proportion of systematic variance attributable to intrapersonal change.
-It includes the log of the number of participants, the number of waves the variable was asked (capped at t = 10 for each item), the date of the first wave (measured as the year minus 1968 divided by 10), number of waves the question was asked averaged across participants, duration in years per item averaged across participants, and panel and topic fixed effects.
-Coefficients are estimated from suppressed intercept model based on predictions at participants' wave mid-point.
-Survey indicators and item topics not shown.
-See Supplemental Materials B for the full set of coefficients.}
-\end{figure}
 
@@
 
-\begin{figure}[hbt]
 \begin{center}
-\caption{Difference in $\omega$ Across College Graduates and Non-Graduates}
-
-\includegraphics[width=450px]{../figures/figure_3}
-
+<Figure 3 about here>
 \end{center}
-\footnotesize{\textit{Notes:} The figure shows the difference in $\omega$ values across college graduates and college non-graduates.
-Values above (below) 0 means that those with college degree have higher (lower) variance of intrapersonal change.
-The dashed red line marks 0 difference.}
-\end{figure}
 
@@
 
-\begin{figure}[hbt]
 \begin{center}
-\caption{Difference in $\omega$ Across College Graduates and Non-Graduates on Political Culture}
-
-\includegraphics[width=450px]{../figures/figure_4}
-
+<Figure 4 about here>
 \end{center}
-\footnotesize{\textit{Notes:} $\omega$ values across college graduates and college non-graduates on 8 political culture items from the General Social Survey (2006-2014).}
-\end{figure}
 
@@
 
+\newpage
+
 \hypertarget{appendix}{%
@@
 
-\begin{figure}[htp]
 \begin{center}
-\caption*{Figure A1: The Distribution of V(D) and V(C)}
-
-\includegraphics[width=350px]{../figures/figure_A1}
-
+<Figure A1 about here>
 \end{center}
-\end{figure}
 
@@
 
-\begin{figure}[tb]
+\begin{center}
+<Figure B1 about here>
+\end{center}
+
+\newpage
+
+\hypertarget{tables-and-figures}{%
+\section{Tables and Figures}\label{tables-and-figures}}
+
+\begin{table}[H]
+
+\caption{\label{tab:unnamed-chunk-1}The Description of the Data Sources}
+\centering
+\resizebox{\linewidth}{!}{
+\begin{tabular}[t]{llcc}
+\toprule
+Country & Survey & Period & Outcomes\\
+\midrule
+Australia & Household, Income and Labor Dynamics in Australia (HILDA) & 2001-2021 & 30\\
+Germany & Socio-Economic Panel Study (SOEP) & 1984-2020 & 122\\
+Great Britian & British Household Panel Survey (BHPS) & 1991-2008 & 83\\
+Great Britian & Understanding Society/UK Household Longitudinal Study (UKHLS) & 2009-2020 & 66\\
+Switzerland & Swiss Household Panel (SHP) & 1999-2019 & 77\\
+\addlinespace
+United States & General Social Survey (GSS) & 2006-2012 & 183\\
+United States & Panel Study of Income Dynamics (PSID) & 1968-2019 & 48\\
+\bottomrule
+\end{tabular}}
+\end{table}
+
+\newpage
+
+\begin{figure}[ht]
+\begin{center}
+\caption{Proportions of Systematic Variance in Personal Culture}
+
+\includegraphics[width=1\linewidth]{../figures/figure_1}
+
+\end{center}
+\footnotesize{\textit{Notes:} The figure shows $\omega$ and 1 - $\omega$ as the proportion of systematic variance attributable to intrapersonal change and interpersonal differences.
+See Supplemental Materials A for the full set of item values.}
+\end{figure}
+
+\newpage
+
+\begin{figure}[ht]
+\begin{center}
+\caption{Regression Model Estimating $\omega$}
+
+\includegraphics{../figures/figure_2.png}
+
+\end{center}
+\footnotesize{\textit{Notes:} The model estimates the proportion of systematic variance attributable to intrapersonal change.
+It includes the log of the number of participants, the number of waves the variable was asked (capped at t = 10 for each item), the date of the first wave (measured as the year minus 1968 divided by 10), number of waves the question was asked averaged across participants, duration in years per item averaged across participants, and panel and topic fixed effects.
+Coefficients are estimated from suppressed intercept model based on predictions at participants' wave mid-point.
+Survey indicators and item topics not shown.
+See Supplemental Materials B for the full set of coefficients.}
+\end{figure}
+
+\newpage
+
+\begin{figure}[ht]
+\begin{center}
+\caption{Difference in $\omega$ Across College Graduates and Non-Graduates}
+
+\includegraphics{../figures/figure_3.png}
+
+\end{center}
+\footnotesize{\textit{Notes:} The figure shows the difference in $\omega$ values across college graduates and college non-graduates.
+Values above (below) 0 means that those with college degree have higher (lower) variance of intrapersonal change.
+The dashed red line marks 0 difference.}
+\end{figure}
+
+\newpage
+
+\begin{figure}[ht]
+\begin{center}
+\caption{Difference in $\omega$ Across College Graduates and Non-Graduates on Political Culture}
+
+\includegraphics{../figures/figure_4.png}
+
+\end{center}
+\footnotesize{\textit{Notes:} $\omega$ values across college graduates and college non-graduates on 8 political culture items from the General Social Survey (2006-2014).}
+\end{figure}
+
+\newpage
+
+\begin{figure}[htp]
+\begin{center}
+\caption*{Figure A1: The Distribution of V(D) and V(C)}
+
+\includegraphics{../figures/figure_A1.png}
+
+\end{center}
+\end{figure}
+
+\newpage
+
+\begin{figure}[htp]
 \begin{center}
@@
 
-\includegraphics[width=400px]{../figures/figure_A2}
+\includegraphics{../figures/figure_A2.png}
 

### 2023-10-14 manuscript/manuscript.Rmd
added names

@@
 author: 
-- 
+  - Achim Edelmann^[médialab, Science Po]
+  - Kevin Kiley^[Department of Sociology and Anthropology, North Carolina State University]
+  - Turgut Keskintürk^[Department of Sociology, Duke University]
+  - Isabella Bouklas^[Department of Sociology, Duke University]
+  - Stephen Vaisey^[Department of Sociology, Duke University]
 output:
@@
   \let\footnote=\endnote
-  \pagenumbering{gobble}
 bibliography: kileybib.bib

### 2023-10-14 manuscript/manuscript.tex

@@
 \newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}
-\usepackage{fontspec} \setmainfont{Times New Roman} \setlength{\parindent}{3em} \usepackage{caption} \usepackage{setspace}\doublespacing \usepackage{endnotes} \renewcommand{\enotesize}{\normalsize} \let\footnote=\endnote \pagenumbering{gobble}
+\usepackage{fontspec} \setmainfont{Times New Roman} \setlength{\parindent}{3em} \usepackage{caption} \usepackage{setspace}\doublespacing \usepackage{endnotes} \renewcommand{\enotesize}{\normalsize} \let\footnote=\endnote
 \usepackage{booktabs}
@@
   pdftitle={Quantifying The Importance of Change for Understanding Differences in Personal Culture},
+  pdfauthor={Achim Edelmann; Kevin Kiley; Turgut Keskintürk; Isabella Bouklas; Stephen Vaisey},
   hidelinks,
@@
 Differences in Personal Culture}
-\author{}
-\date{\vspace{-2.5em}}
+\author{Achim Edelmann\footnote{médialab, Science Po} \and Kevin
+Kiley\footnote{Department of Sociology and Anthropology, North Carolina
+  State University} \and Turgut Keskintürk\footnote{Department of
+  Sociology, Duke University} \and Isabella
+Bouklas\footnote{Department of Sociology, Duke University} \and Stephen
+Vaisey\footnote{Department of Sociology, Duke University}}
+\date{}
 

### 2024-03-14 manuscript/manuscript.Rmd
kkedits314

--- 2023-10-11 manuscript/manuscript.Rmd
@@
 ---
-title: "Quantifying The Importance of Change for Understanding Differences in Personal Culture"
+title: "Quantifying the Importance of Change for Understanding Differences in Personal Culture"
 author: 
-- 
+  - Achim Edelmann^[médialab, Science Po]
+  - Kevin Kiley^[Department of Sociology and Anthropology, North Carolina State University]
+  - Turgut Keskintürk^[Department of Sociology, Duke University]
+  - Isabella Bouklas^[Department of Sociology, Duke University]
+  - Stephen Vaisey^[Department of Sociology, Duke University]
 output:
@@
   \usepackage{caption}
-  \usepackage{setspace}\doublespacing
+  \usepackage{setspace}
   \usepackage{endnotes}
@@
   \let\footnote=\endnote
-  \pagenumbering{gobble}
 bibliography: kileybib.bib
 csl: asa.csl
-abstract: "When researchers want to understand differences in personal culture---a person's attitudes, beliefs, values, and practices--how much attention should they pay to adult experience?
-Recent work has reached substantially different conclusions on this question.
-We argue that this disagreement is an unintended consequence of the \"tournament of models\" approach researchers have used, which focuses on whether people change and not how much they change.
-To advance the theoretical debate, we refocus attention on the relative importance of personal change over time for explaining differences between people.
-We introduce a new measure for quantifying the proportion of systematic variance in panel data attributable to intrapersonal change.
-Applying this measure to 609 items from seven surveys in five countries, we find that although intrapersonal change is common, it is generally small in magnitude.
+abstract: "Work in the sociology of culture has converged on a general conclusion that while people sometimes make substantial shifts in their personal culture over time, most elements of personal culture are characterized by low rates of persistent change during adulthood.
+Recent developments have begun to quantify change over time and initial differences, but it is unclear how to use these components to understand the relative importance of change, especially when change is modeled non-linearly.
+To advance this debate, we introduce a measure for quantifying the proportion of systematic variance in panel data attributable to intrapersonal change.
+Applying this measure to 609 items from seven surveys in five countries, we find that although intrapersonal change is common, it does not explain a large share of systematic variance.
 As an extension of the theoretical debate, we demonstrate that this measure provides new insights when comparing social groups, showing that intrapersonal change is less common among U.S.
@@
 
+\doublespacing
+
 # Introduction
 
-Does personal culture---a person's attitudes, beliefs, values, and practices---change over the life course, or is it largely fixed by adulthood?
-This question underlies an important contemporary debate in sociology [@lizardo2017; @kiley2020; @lersch2023] and has deep roots in seemingly contradictory theoretical perspectives.
+An important contemporary debate in sociology centers on the importance of intrapersonal change for understanding differences in "personal culture" -- attitudes, beliefs, values, and practices [@lizardo2017; @kiley2020; @lersch2023; @underwood2021; @vaisey2016].
+This question has deep roots in seemingly contradictory theoretical perspectives.
 For example, pragmatist theories of action claim that changes in social environments cause people to adapt their views and make new cultural meanings [@swidler2001; @gross2009], while Bourdieusian practice theories argue that the "past conditions of production" leave a mark on people's personal culture that lasts throughout their lives [@bourdieu1990].
 Models of social influence assume that people adapt their culture in the face of new information [@goldberg2018], while the emphasis on cohort effects in models of aggregate social and cultural change requires them to be open to change while young but become fairly resistant to it as they age [@ryder1965].
-Finally, life course theories posit important changes over time as people advance through important transitions in their lives [@elder2003; @bardi2009].
+Finally, life course theories posit important changes over time as people advance through important transitions in their lives, but that certain life stages have more lasting influence than others [@elder2003; @bardi2009].
 
-Because all of these perspectives have some empirical support, the theoretical debate is not about whether the processes they posit exist at all, but about their relative contribution to explaining the cultural differences we see in the world.
-Nevertheless, researchers have struggled to reach a consensus on the importance of cultural change during adulthood.
-Over the span of a few years, most apparent changes in adults' survey responses appear to be transitory, with little evidence of persistent change [@kiley2020; @vaisey2021].
-This suggests that change during adulthood may not be a major factor in explaining contemporary cultural divides.
-By contrast, when we consider a longer time horizon, there is evidence that adults make at least some persistent changes [@lersch2023].
+Recent work on this topic has converged on a general conclusion that people do sometimes make substantial shifts in their personal culture over time but that most elements of personal culture are characterized by low rates of persistent change during adulthood [@kiley2020; @lersch2023].
+Until recently, this work has adopted a "tournament of models" approach to adjudicating different theoretical perspectives [@lersch2023: p.
+228].
+In this framework, researchers formalized different models of change over time and relied on model selection criteria---primarily the Bayesian Information Criterion [@raftery1995]---to classify different questions as demonstrating either "no change" or "any change." The size of these piles then serves as the primary evidence for the truth of different arguments [@kiley2020; @vaisey2021].
 
-These seemingly inconsistent findings are partly due to the fact that researchers have taken a "tournament of models" approach to adjudicate different theoretical perspectives [@lersch2023: p.
-228].
-That is, researchers have relied on model selection criteria---primarily the Bayesian Information Criterion [@raftery1995]---to sort variables into the "no change" or "any change" piles.
-The size of these piles then serves as the primary evidence for the truth of the theory [@kiley2020; @vaisey2021; @lersch2023].
+However, as consensus in the overall picture emerges, researchers are now turning toward understanding the conditions that produce persistent change in personal culture, including age, education, and life-course transitions; whether certain kinds of people are more likely to make persistent changes in personal culture; and whether certain kinds of personal culture are more likely to be characterized by persistent change or stability.
+The tournament of models approach, useful for the broad question of whether change happens, is ill-suited to addressing these more theoretically grounded questions.
 
-We argue, however, that asking *whether* people change on an item is an unintended source of confusion.
-Being able to detect *any* change on an item depends not only on sample populations, measurement, statistical methods, and power, but also on definitions of what counts as "change." These differences can limit the possibility of agreement on the assignment of a variable to the right pile.
-Moreover, asking whether there is evidence of change reduces to, "Is there *any* evidence of change in this item?" But such binary questions are ill-suited to address the core of the theoretical debate: determining *how much* the process of intrapersonal change contributes to cultural differences.
+In a recent paper, Lersch [-@lersch2023] takes an important step forward by devising a Life Course Adaption Model (LCAM), a mixed-effects growth curve model that quantifies various components of cultural difference: early imprinting, persistent change, biographical experience, and fluctuation.
+In quantifying these aspects of cultural difference, the LCAM lets researchers address a broader range of questions than the single question of whether a measure is characterized by stability or change.
+However, because the LCAM produces so many different measures, it is not obvious how to use these components to understand the relative importance of change, espeically when change is measured in non-linear ways.
 
-Of course neither pre-adult socialization nor the contemporary social context alone can explain the full range of cultural differences.
-The relevant question is about the *relative* contributions of these processes.
-In the current debate, for a single construct (e.g., support for gay rights), model comparison approaches can only help decide *whether* intrapersonal change is happening, not about how much.
-This also prevents investigating under what conditions, across which groups, and in what domains we might see differences in the amount of intrapersonal change.
-These questions are about *degree*, not existence.
-Taking this debate forward therefore requires precise measurement rather than declarations of victory for a particular perspective.
+In this paper, we build on the LCAM by outlining a single metric that combines diverse components of the LCAM to quantify the relative importance of intrapersonal change and interpersonal difference for explaining variation in a population for a variable over time.
+Our approach isolates the "systematic" components of variance (i.e., it removes random change and measurement error), to provide a single succinct summary of how much systematic variation can reasonably be attributed to people making persistent changes over time.
+We demonstrate the utility of this quantity in two ways.
 
-In this paper, we move beyond the "tournament of models" approach and introduce a new method for quantifying the relative contributions of interpersonal differences and intrapersonal change for a single item.
-We use 609 items from seven panel surveys from five countries to quantify the proportion of variance that is attributable to systematic intrapersonal change over the study period rather than to stable interpersonal differences.
+First, we use 609 items from seven panel surveys from five countries to quantify the proportion of variance that is attributable to systematic intrapersonal change over the study period rather than to interpersonal differences at baseline.
+This allows us to generate an overall picture of how much of the total systematic variation in personal culture change tends to explain.
+These findings largely echo previous findings but enable clearer statements about the relative importance of change across questions, rather than binary claims about whether change "happens" or specific quantification of how much change occurs.
 
-We observe similar results across all datasets despite their varied social contexts and duration.
-In general, we find that intrapersonal change accounts for a small amount of variance in personal culture survey items.
-Of course, any question about whether a given amount of intrapersonal change is "meaningful" or "important" is substantive and theoretical.
-Nevertheless, on many measures (some observed for over a decade), we do not see enough change to believe that intrapersonal change processes play a substantial role in explaining the differences we observe between adults.
-Notable, interesting, and important exceptions exist, but they are departures from the overall pattern.
-
-In addition to advancing a theoretical debate, our method can also help address specific substantive questions.
-To illustrate how it might be used by a researcher with a specific topical interest, we investigate differences in the relative importance of intrapersonal change for people with and without a college degree.
-We find a smaller amount of systematic change among college-educated respondents, suggesting that college crystallizes one's personal culture rather than fostering an openness to new information.
+Second, to illustrate how this quantity might be used by a researcher with a specific topical interest, we investigate differences in the relative importance of intrapersonal change for people with and without a college degree.
+We find that change explains a smaller proportion of systematic variance among college-educated respondents, suggesting that college crystallizes one's personal culture rather than fostering an openness to new information.
 This offers a new empirical basis to theorize about the role of education in influencing personal culture.
-
-Taken together, our goal in this paper is to advance---and, hopefully, transcend---debates about *whether* adults change.
-People do change, at least a little, on most things.
-Our method can help quantify exactly *how much*.
 
@@
 500; see also @vaisey2021].
-This lack of durable change is consistent with other recent findings that cohort replacement plays a somewhat larger role than period effects in explaining differences in personal culture [@vaisey2016].
+This lack of durable change is consistent with other recent findings that cohort replacement plays a larger role than period effects in explaining differences in personal culture [@vaisey2016].
 
@@
 Drawing on these findings, Lersch [-@lersch2023] challenged the SDM and AUM as a "needless dichotomy," proposing the "Life Course Adaption Model" (LCAM) as an alternative.
-This model draws on the life course perspective to model personal culture as a different linear trajectory over the duration of a panel for each respondent.
-In doing this, Lersch rectified two shortcomings of the AUM and SDM.
+This model draws on the life course perspective to model personal culture as a different linear trajectory[^linear] over the duration of a panel for each respondent.
+In doing this, Lersch rectified three shortcomings of the AUM and SDM.
 First, the AUM, as described by Kiley and Vaisey, posits that changes follow a Markov process where responses at time $t$ are a function of responses at time $t-1$ but not earlier time points.
@@
 
-Second, Kiley and Vaisey's analysis of the AUM and SDM is based on three-wave panel data over four years, which might not be intensive or extensive enough to adjudicate the two models.
+[^linear]: In supplemental materials, Lersch models age using splines as well, with results generally showing the same conclusions.
+
+Second, Kiley and Vaisey's conclusions about the AUM and SDM were based on examining three-wave panel data over four years, which might not be intensive or extensive enough to adjudicate the two models.
 Lersch evaluates the LCAM against the AUM and SDM using panel data spanning a wider duration (from 3 to 36 years) and more waves (3 to 18), offering a better chance to observe durable change.
 
-When Lersch compared the LCAM to the AUM and SDM on data from five countries, 297 of the 428 questions he analyzed preferred the LCAM, suggesting that we observe some linear change for adults on most items.
-The SDM was preferred on 112 items, and the rest did not yield conclusive results.
-No questions favored the AUM.
+Third, and most importantly, the LCAM quantifies various components of cultural difference, allowing researchers to ask a more varied set of questions than simply whether people change.
+For example, researchers can use the LCAM framework to quantify the magnitude of change in a measure of culture over time (i.e., how much do people change on average), the variance in age-based slopes over time (i.e., do people tend to change in similar ways as they age), or how much personal culture tends to fluctuate around these age-based trends (how consistent are responses over time).
+
+When Lersch compared the LCAM to the AUM and SDM on data from five countries, 198 of the 428 questions he analyzed preferred the LCAM and 113 preferred the AUM, suggesting that we observe some change for adults on most items.
+The SDM was preferred on 93 items, and the rest did not yield conclusive results.
 He concludes that "new experiences over the life course [...] can persistently move individuals' personal culture in novel directions" [@lersch2023: p.
@@
 Despite differences in how these researchers interpret their findings, the empirical results are not far apart.
-Kiley and Vaisey [-@kiley2020] found that the majority of items they tested favored the AUM, meaning there was evidence of durable change among adults, a pattern consistent with Lersch's results.
-Also Lersch [-@lersch2023] found evidence of durable change on most items, but these changes were small.
+Kiley and Vaisey [-@kiley2020] found that the majority of items they tested favored the AUM, meaning there was evidence of persisting change among adults, a pattern consistent with Lersch's results.
+Also Lersch [-@lersch2023] found evidence of persisting change on most items, but these changes were small.
 On average, people changed only about .07 standard deviations over 10 years.
-This is consistent with Kiley and Vaisey's finding that durable change is often small.
+This is consistent with Kiley and Vaisey's finding that persisting change is often small.
 Moreover, a quarter of the items Lersch studied still preferred the SDM, suggesting that even with different assumptions about change and more extensive data, many questions still are consistent with a model that allows zero intrapersonal change.
@@
 
-To this point, the debate has been about whether people *ever* change.
-As we argued above, this is rarely the right question since the answer is almost always "yes." Over sufficient time and in a large enough sample, researchers will likely observe some evidence for durable individual change.
+To this point, the debate has broadly centered on drawing conclusions about whether people *ever* change.
+But this is rarely the right question since the answer is almost always "yes." Over sufficient time and in a large enough sample, researchers will likely observe some evidence for durable change in the population.
 A lack of evidence for change may be due to a poor survey question, a low resolution in response options, or that the question simply was not asked for long enough or to enough people.
@@
 Lersch's LCAM does this by modeling each individual as following a personal linear trajectory.
-But rather than debating whether this model fits a particular data set better than a model that assumes that people never change, it would be better to quantify these individual changes and compare them to the stable differences that exist between people.
-In other words, the focus should shift from asking, "do people change?" to asking, "what are the relative contributions of change and pre-existing differences for explaining variance in personal culture?"
+In doing so, the LCAM opens the door to a variety of questions.
+We assert that among the many things that can be done with the LCAM, it provides the opportunity to shift from asking, "do people change?" to asking, "what are the relative contributions of change and pre-existing differences for explaining variance in personal culture?"
 
-We propose an approach---which we formalize below---that offers such a quantification.
+Our approach---which we formalize below---offers such a quantification.
 Our measure relates two variance components that each reflect a combination of relevant theoretical processes.
@@
 
-The second component, the amount of intrapersonal linear change people make during the panel, captures durable changes in personal culture over time.
-This captures the set of processes collectively called "persistent change" or "adaption" by Lersch and "active updating" by Kiley and Vaisey.
+The second component, the amount of intrapersonal change people make during the panel, captures persisting changes in personal culture over time.
+This component captures the set of processes collectively called "persistent change" or "adaption" by Lersch and "active updating" by Kiley and Vaisey.
 Lersch attributes these changes to social triggers such as moving into a new environment or adopting new social roles (although he does not measure these directly).
@@
 Furthermore, these components are not necessarily useful in isolation or without context.
+If absolute change is about .07 standard deviations every 10 years, but the variance of starting points at age 18 is very high, then change over time will not produce significant differences among people, even over a long time frame.
+If, on the other hand, people tend to start at very similar positions and change at a rate of .07 standard deviations, then that change is very useful for understanding difference.
 What is important for the theoretical debate is the relative contribution of these two components in explaining cultural differences in a population.
@@
 By quantifying the relative contributions of interpersonal differences and intrapersonal change, researchers can gauge their relative importance in explaining differences among survey items, between groups within a given population, across time, and across societies.
+
+## Formalization
+
+The starting point for our quantity is Lersch's [-@lersch2023] Life Course Adaption Model, which formalizes survey responses at time $t$ as a function of individual-level random intercepts and slopes for survey age.
+In some contexts, this model is called a mixed-effects growth curve model.
+This model assumes a set of propositions about change that reflect the theoretical debate to this point.
+First, consistent with the settled disposition model, it assumes that people start the survey with cultural differences, modeled as random intercepts for each respondent.
+Second, it assumes that people change over time, taking the form of random slopes for each respondent as a linear function of time.
+Third, it assumes that people deviate around this baseline randomly over time, reflecting "fluctuation" or short-term non-persistent change.
+Formally, this can be written as
+
+$$
+y_{it} = \beta_0 + \alpha_{0i} + (\beta_1 + \alpha_{1i}) \text{year}_{it} + \epsilon_{it}
+$$
+
+$$
+\begin{aligned}
+\alpha_{0i} &\sim \mathcal{N}(0,\tau^2_{0}) \\
+\alpha_{1i} &\sim \mathcal{N}(0,\tau^2_{1}) \\
+\epsilon_{it} &\sim \mathcal{N}(0,\sigma^2)
+\end{aligned}
+$$
+
+where $\beta_0$ is the average intercept, $\alpha_{0i}$ is the random intercept component for individuals, $\beta_1$ is the average yearly change in the outcome, $\alpha_{1i}$ is the random slope component for individuals, and $\epsilon_{it}$ is the random error term that captures transitory disturbances.
+The $\alpha_{0i}$ and $\alpha_{1i}$ terms are also allowed to covary.
+
+We then use these models to derive two measures of the components we outlined above: pre-existing interpersonal differences and intrapersonal change over time.
+
+We estimate these using a form of variance decomposition.
+First, to measure stable interpersonal differences, we calculate $V(D)$ according to the equation below, where $\widetilde{year_i}$ is the midpoint of the observed years for each respondent for that particular item.[^2]
+
+$$
+V(D) = 1 - \frac{\sum_{i=1}^N \sum_{t=1}^T [y_{it} - (\hat{\beta}_0 + \hat{\alpha}_{0i} + (\hat{\beta}_1 + \hat{\alpha}_{1i}) \widetilde{\text{year}}_{i})]^2}{\sum_{i=1}^N \sum_{t=1}^T [y_{it} - \bar{y}_{it}]^2}
+$$
+
+[^2]: We use the model estimate from the midpoint year, rather than the value at the first wave, because this is the best measure of "baseline" available under the assumption that the person does not change.
+The first wave measure alone contains an unknown amount of measurement error and transient fluctuations.
+
+This gives us the proportion of total variance in an item attributable to stable interpersonal differences.
+
+To calculate the second component, which is the variance attributable to systematic intrapersonal change, we calculate $V(C)$ as follows:
+
+$$
+V(C) = 1 - \frac{\sum_{i=1}^N \sum_{t=1}^T [y_{it} - (\hat{\beta}_0 + \hat{\alpha}_{0i} + (\hat{\beta}_1 + \hat{\alpha}_{1i}) \text{year}_{it})]^2}{\sum_{i=1}^N \sum_{t=1}^T [y_{it} - \bar{y}_{it}]^2} - V(D)
+$$
+
+This is the incremental proportion of variance accounted for when we allow the model predictions to change over time for each person.
+
+The third component, which is the proportion of variance attributable to measurement error or transient fluctuations, is what's left, such that these three numbers sum to 1.
+As we discussed above, this component is not of core interest here, so we set it aside in our analysis.
+
+Together, $V(D)$ and $V(C)$ reflect the total systematic variance of the outcome that the LCAM can attribute to either differences between people or linear intrapersonal change over the course of the panel.
+Our principal measure of interest is the proportion of systematic variance accounted for by intrapersonal change, which we calculate as
+
+$$
+\omega = \frac{V(C)}{V(C) + V(D)}
+$$
+
+We can think of $\omega$ as a type of Intraclass Correlation Coefficient focused only on the systematic components of the model while ignoring residual variance that might be attributable to measurement error or non-persisting change [see also the variance decomposition of period and cohort by @vaisey2016].
+The higher this proportion is, the more systematic variance in responses is attributable to intrapersonal change during the panel.
+From now on, we use $\omega$ to stand for *the proportion of systematic variance attributable to intrapersonal change*.
+
+Because this quotient summarizes variance proportions and therefore has no natural referent, it is hard to say at which values we observe "a lot" of intrapersonal change.
+But summarizing questions in this way allows us to compare the relative importance of our two broad processes across a range of questions that might have different levels of non-systematic variance.
+This gives us a sense of the relative prevalence of these two processes for each question, which is not achievable under the tournament of models approach.
+This also allows us to compare which kinds of questions show more or less intrapersonal change relative to interpersonal differences.
+While the tournament of models approach allowed researchers to say whether questions in general showed *any* evidence of updating, this approach allows us to make more specific claims about the prevalence of these theoretical processes for particular questions relative to each other.
+
+## Bootstrapping for Confidence Intervals
+
+[Explain how we do that here]
 
@@
 
-## Life Course Adaption Model
-
-The estimation procedure begins with Lersch's [-@lersch2023] Life Course Adaption Model, which formalizes survey responses at time $t$ as a function of individual-level random intercepts and slopes for survey age.
-In some contexts, this model is called a mixed-effects growth curve model.
-This model assumes a set of propositions about change that reflect the theoretical debate to this point.
-First, consistent with the settled disposition model, it assumes that people start the survey with cultural differences, modeled as random intercepts for each respondent.
-Second, it assumes that people change over time, taking the form of random slopes for each respondent as a linear function of time.
-Third, it assumes that people deviate around this baseline randomly over time, reflecting "fluctuation" or short-term non-persistent change.
-Formally, this can be written as
-
-$$
-y_{it} = \beta_0 + \alpha_{0i} + (\beta_1 + \alpha_{1i}) \text{year}_{it} + \epsilon_{it}
-$$
-
-$$
-\begin{aligned}
-\alpha_{0i} &\sim \mathcal{N}(0,\tau^2_{0}) \\
-\alpha_{1i} &\sim \mathcal{N}(0,\tau^2_{1}) \\
-\epsilon_{it} &\sim \mathcal{N}(0,\sigma^2)
-\end{aligned}
-$$
-
-where $\beta_0$ is the average intercept, $\alpha_{0i}$ is the random intercept component for individuals, $\beta_1$ is the average yearly change in the outcome, $\alpha_{1i}$ is the random slope component for individuals, and $\epsilon_{it}$ is the random error term that captures transitory disturbances.
-The $\alpha_{0i}$ and $\alpha_{1i}$ terms are also allowed to covary.
-
 ## Analysis Steps
 
-As a first step in our analysis, we fit this LCAM to each of the 609 measures of personal culture outlined above.
-We then use these models to derive two measures of the components we outlined above: pre-existing interpersonal differences and intrapersonal change over time.
+As a first step in our analysis, we fit this LCAM to each of the 609 measures of personal culture
+outlined above. We then derive $\omega$ for each of these items.
 
-We estimate these using a form of variance decomposition.
-First, to measure stable interpersonal differences, we calculate $V(D)$ according to the equation below, where $\widetilde{year_i}$ is the midpoint of the observed years for each respondent for that particular item.[^2]
-
-$$
-V(D) = 1 - \frac{\sum_{i=1}^N \sum_{t=1}^T [y_{it} - (\hat{\beta}_0 + \hat{\alpha}_{0i} + (\hat{\beta}_1 + \hat{\alpha}_{1i}) \widetilde{\text{year}}_{i})]^2}{\sum_{i=1}^N \sum_{t=1}^T [y_{it} - \bar{y}_{it}]^2}
-$$
-
-[^2]: We use the model estimate from the midpoint year, rather than the value at the first wave, because this is the best measure of "baseline" available under the assumption that the person does not change.
-The first wave measure alone contains an unknown amount of measurement error and transient fluctuations.
-
-This gives us the proportion of total variance in an item attributable to stable interpersonal differences.
-
-To calculate the second component, which is the variance attributable to systematic intrapersonal change, we calculate $V(C)$ as follows:
-
-$$
-V(C) = 1 - \frac{\sum_{i=1}^N \sum_{t=1}^T [y_{it} - (\hat{\beta}_0 + \hat{\alpha}_{0i} + (\hat{\beta}_1 + \hat{\alpha}_{1i}) \text{year}_{it})]^2}{\sum_{i=1}^N \sum_{t=1}^T [y_{it} - \bar{y}_{it}]^2} - V(D)
-$$
-
-This is the incremental proportion of variance accounted for when we allow the model predictions to change over time for each person.
-
-The third component, which is the proportion of variance attributable to measurement error or transient fluctuations, is what's left, such that these three numbers sum to 1.
-As we discussed above, this component is not of core interest here.
-
-Together, $V(D)$ and $V(C)$ reflect the total systematic variance of the outcome that the LCAM can attribute to either differences between people or linear intrapersonal change over the course of the panel.
-Our principal measure of interest is the proportion of systematic variance accounted for by intrapersonal change, which we calculate as
-
-$$
-\omega = \frac{V(C)}{V(C) + V(D)}
-$$
-
-We can think of $\omega$ as a type of Intraclass Correlation Coefficient focused only on the systematic components of the model while ignoring residual variance that might be attributable to measurement error or non-persisting change [see also the variance decomposition of period and cohort by @vaisey2016].
-The higher this proportion is, the more systematic variance in responses is attributable to intrapersonal change during the panel.
-From now on, we use $\omega$ to stand for *the proportion of systematic variance attributable to intrapersonal change*.
-
-Because this quotient summarizes variance proportions and therefore has no natural referent, it is hard to say at which values we observe "a lot" of intrapersonal change.
-But summarizing questions in this way allows us to compare the relative importance of our two broad processes across a range of questions that might have different levels of non-systematic variance.
-This gives us a sense of the relative prevalence of these two processes for each question, which is not achievable under the tournament of models approach.
-This also allows us to compare which kinds of questions show more or less intrapersonal change relative to interpersonal differences.
-While the tournament of models approach allowed researchers to say whether questions in general showed *any* evidence of updating, this approach allows us to make more specific claims about the prevalence of these theoretical processes for particular questions relative to each other.
-
-As a second step in our analysis, we use linear regression to conduct a meta-analysis, modeling $\omega$ as a function of features of the question and panel.
+As a second step in our analysis, we use linear regression to model $\omega$ as a function of features of the question and panel.
 This allows us to explore how these features are associated with the degree of intrapersonal change.
@@
 
-The objective of this paper is to intervene in the recent debate on whether people change their personal culture -- their attitudes, beliefs, values, and practices -- as they move through their adult life.
+The primary objective of this paper is to introduce a new quantity that can advance debate on whether people change their personal culture -- their attitudes, beliefs, values, and practices -- as they move through their adult life.
 Instead of falling back on a "tournament of models" approach [@lersch2023: p.
-228] to proclaim victory for a particular answer to this question, our goal is to refocus the debate to the theoretically more productive question of the relative importance of intrapersonal change for explaining differences in the personal culture.
+228] to proclaim victory for a particular answer to this question, our goal is to refocus the debate to the theoretically more productive question of the relative importance of intrapersonal change for explaining differences in personal culture.
 We propose a new approach that quantifies the amount of observed systematic variance that is attributable to either interpersonal differences at baseline or intrapersonal change over the duration of a panel.
@@
 
-More than adjudicating past positions in this debate, the approach we presented in this paper will enable scholars to explore more precisely the relative contributions of change and stability in explaining variation in personal culture, as well as how these contributions might differ across populations.
+More than adjudicating past positions in this debate, the approach we presented in this paper will enable researchers to explore more precisely the relative contributions of change and stability in explaining variation in personal culture, as well as how these contributions might differ across populations.
 To illustrate this, we showed how the amount of cultural difference explained by intrapersonal change varies substantially by survey item but also by individuals' characteristics such as their education.
@@
 
-Although we believe it is a major advance on previous work, the approach we outline here has important limitations.
+Although we believe the approach we outline here advances previous work, it has important limitations.
 We allocate all systematic variance to one of two sets of theoretical processes: intrapersonal change and interpersonal differences at baseline.
@@
 Our results and interpretations also hinge on how we have defined change.
-As does the LCAM, our approach treats cultural trajectories as varying linear slopes for each respondent, thus assuming that change is a linear function of time.
+As does the LCAM, our analysis here treats cultural trajectories as varying linear slopes for each respondent, thus assuming that change is a linear function of time.
 This assumption simplifies reality in which change likely also takes non-linear and discontinuous forms.
@@
 In parts, our assumption of linearity is a limitation of the data, as most questions are only observed for three to six waves.
-Panels with more waves might allow researchers to loosen this assumption to test alternative, more flexible models of change.
+However, our overall approach does not require modeling change as a linear function of age.
+Change can be modeled in non-linear forms, such as the splines used by Lersch [-@lersch2023], and accordingly incorporated into the calculation of variance explained.
 
@@
 Longer panels and more flexible definitions of "change" might allow us to account for such trajectories.
+Again, while we model change in a single way, the general variance decomposition approach we outline accommodates different definitions and functional forms of change.
 
@@
 Sociologists interested in understanding cultural differences have largely asked about the *existence* of cultural change in adults [e.g., @kiley2020; @vaisey2021; @lersch2023].
-Yet this approach has inadvertently limited the debate.
 In any population, *some* degree of adult cultural change is inevitable.
@@
 
-This and other recent findings [@stewart2023; @quinn2023] suggest that it is theoretically more productive to measure the relative importance of these two components in concrete, substantive cases.
+This and other recent findings [@stewart2023; @quinn2023; @quinn2024] suggest that it is theoretically more productive to measure the relative importance of these two components in concrete, substantive cases.
 Drawing a unified conclusion from survey items measuring various cultural forms on different scales and across different time frames is challenging.
@@
 
-We believe the approach we outlined here can push past the "needless dichotomy" implicit in the question of whether people change or not [@lersch2023].
+We believe the approach we outlined here can push past the "needless dichotomy" implicit in the question of whether people change or not.
 Characterizing questions as displaying change or not can only take researchers so far, but the question of whether some questions demonstrate more change than others, or whether some groups are characterized by more stability than others, has the potential to weigh in on a broader range of theoretical debates.
@@
 \theendnotes
+
+\singlespace
 

### 2024-03-14 manuscript/manuscript.tex
kkedits314

@@
 \newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}
-\usepackage{fontspec} \setmainfont{Times New Roman} \setlength{\parindent}{3em} \usepackage{caption} \usepackage{setspace}\doublespacing \usepackage{endnotes} \renewcommand{\enotesize}{\normalsize} \let\footnote=\endnote
+\usepackage{fontspec} \setmainfont{Times New Roman} \setlength{\parindent}{3em} \usepackage{caption} \usepackage{setspace} \usepackage{endnotes} \renewcommand{\enotesize}{\normalsize} \let\footnote=\endnote
 \usepackage{booktabs}
@@
 \hypersetup{
-  pdftitle={Quantifying The Importance of Change for Understanding Differences in Personal Culture},
+  pdftitle={Quantifying the Importance of Change for Understanding Differences in Personal Culture},
   pdfauthor={Achim Edelmann; Kevin Kiley; Turgut Keskintürk; Isabella Bouklas; Stephen Vaisey},
@@
 
-\title{Quantifying The Importance of Change for Understanding
+\title{Quantifying the Importance of Change for Understanding
 Differences in Personal Culture}
@@
 \begin{abstract}
-When researchers want to understand differences in personal culture---a
-person's attitudes, beliefs, values, and practices--how much attention
-should they pay to adult experience? Recent work has reached
-substantially different conclusions on this question. We argue that this
-disagreement is an unintended consequence of the ``tournament of
-models'' approach researchers have used, which focuses on whether people
-change and not how much they change. To advance the theoretical debate,
-we refocus attention on the relative importance of personal change over
-time for explaining differences between people. We introduce a new
-measure for quantifying the proportion of systematic variance in panel
-data attributable to intrapersonal change. Applying this measure to 609
-items from seven surveys in five countries, we find that although
-intrapersonal change is common, it is generally small in magnitude. As
-an extension of the theoretical debate, we demonstrate that this measure
+Work in the sociology of culture has converged on a general conclusion
+that while people sometimes make substantial shifts in their personal
+culture over time, most elements of personal culture are characterized
+by low rates of persistent change during adulthood. Recent developments
+have begun to quantify change over time and initial differences, but it
+is unclear how to use these components to understand the relative
+importance of change, especially when change is modeled non-linearly. To
+advance this debate, we introduce a measure for quantifying the
+proportion of systematic variance in panel data attributable to
+intrapersonal change. Applying this measure to 609 items from seven
+surveys in five countries, we find that although intrapersonal change is
+common, it does not explain a large share of systematic variance. As an
+extension of the theoretical debate, we demonstrate that this measure
 provides new insights when comparing social groups, showing that
@@
 
+\doublespacing
+
 \hypertarget{introduction}{%
@@
 
-Does personal culture---a person's attitudes, beliefs, values, and
-practices---change over the life course, or is it largely fixed by
-adulthood? This question underlies an important contemporary debate in
-sociology (Kiley and Vaisey 2020; Lersch 2023; Lizardo 2017) and has
-deep roots in seemingly contradictory theoretical perspectives. For
-example, pragmatist theories of action claim that changes in social
-environments cause people to adapt their views and make new cultural
-meanings (Gross 2009; Swidler 2001), while Bourdieusian practice
-theories argue that the ``past conditions of production'' leave a mark
-on people's personal culture that lasts throughout their lives (Bourdieu
-1990). Models of social influence assume that people adapt their culture
-in the face of new information (Goldberg and Stein 2018), while the
-emphasis on cohort effects in models of aggregate social and cultural
-change requires them to be open to change while young but become fairly
-resistant to it as they age (Ryder 1965). Finally, life course theories
-posit important changes over time as people advance through important
-transitions in their lives (Bardi et al. 2009; Elder, Johnson, and
-Crosnoe 2003).
-
-Because all of these perspectives have some empirical support, the
-theoretical debate is not about whether the processes they posit exist
-at all, but about their relative contribution to explaining the cultural
-differences we see in the world. Nevertheless, researchers have
-struggled to reach a consensus on the importance of cultural change
-during adulthood. Over the span of a few years, most apparent changes in
-adults' survey responses appear to be transitory, with little evidence
-of persistent change (Kiley and Vaisey 2020; Vaisey and Kiley 2021).
-This suggests that change during adulthood may not be a major factor in
-explaining contemporary cultural divides. By contrast, when we consider
-a longer time horizon, there is evidence that adults make at least some
-persistent changes (Lersch 2023).
-
-These seemingly inconsistent findings are partly due to the fact that
-researchers have taken a ``tournament of models'' approach to adjudicate
-different theoretical perspectives (Lersch 2023: p.~228). That is,
-researchers have relied on model selection criteria---primarily the
-Bayesian Information Criterion (Raftery 1995)---to sort variables into
-the ``no change'' or ``any change'' piles. The size of these piles then
-serves as the primary evidence for the truth of the theory (Kiley and
-Vaisey 2020; Lersch 2023; Vaisey and Kiley 2021).
-
-We argue, however, that asking \emph{whether} people change on an item
-is an unintended source of confusion. Being able to detect \emph{any}
-change on an item depends not only on sample populations, measurement,
-statistical methods, and power, but also on definitions of what counts
-as ``change.'' These differences can limit the possibility of agreement
-on the assignment of a variable to the right pile. Moreover, asking
-whether there is evidence of change reduces to, ``Is there \emph{any}
-evidence of change in this item?'' But such binary questions are
-ill-suited to address the core of the theoretical debate: determining
-\emph{how much} the process of intrapersonal change contributes to
-cultural differences.
-
-Of course neither pre-adult socialization nor the contemporary social
-context alone can explain the full range of cultural differences. The
-relevant question is about the \emph{relative} contributions of these
-processes. In the current debate, for a single construct (e.g., support
-for gay rights), model comparison approaches can only help decide
-\emph{whether} intrapersonal change is happening, not about how much.
-This also prevents investigating under what conditions, across which
-groups, and in what domains we might see differences in the amount of
-intrapersonal change. These questions are about \emph{degree}, not
-existence. Taking this debate forward therefore requires precise
-measurement rather than declarations of victory for a particular
-perspective.
-
-In this paper, we move beyond the ``tournament of models'' approach and
-introduce a new method for quantifying the relative contributions of
-interpersonal differences and intrapersonal change for a single item. We
-use 609 items from seven panel surveys from five countries to quantify
-the proportion of variance that is attributable to systematic
-intrapersonal change over the study period rather than to stable
-interpersonal differences.
-
-We observe similar results across all datasets despite their varied
-social contexts and duration. In general, we find that intrapersonal
-change accounts for a small amount of variance in personal culture
-survey items. Of course, any question about whether a given amount of
-intrapersonal change is ``meaningful'' or ``important'' is substantive
-and theoretical. Nevertheless, on many measures (some observed for over
-a decade), we do not see enough change to believe that intrapersonal
-change processes play a substantial role in explaining the differences
-we observe between adults. Notable, interesting, and important
-exceptions exist, but they are departures from the overall pattern.
-
-In addition to advancing a theoretical debate, our method can also help
-address specific substantive questions. To illustrate how it might be
-used by a researcher with a specific topical interest, we investigate
-differences in the relative importance of intrapersonal change for
-people with and without a college degree. We find a smaller amount of
-systematic change among college-educated respondents, suggesting that
+An important contemporary debate in sociology centers on the importance
+of intrapersonal change for understanding differences in ``personal
+culture'' -- attitudes, beliefs, values, and practices (Kiley and Vaisey
+2020; Lersch 2023; Lizardo 2017; Vaisey and Lizardo 2016;
+\textbf{underwood2021?}). This question has deep roots in seemingly
+contradictory theoretical perspectives. For example, pragmatist theories
+of action claim that changes in social environments cause people to
+adapt their views and make new cultural meanings (Gross 2009; Swidler
+2001), while Bourdieusian practice theories argue that the ``past
+conditions of production'' leave a mark on people's personal culture
+that lasts throughout their lives (Bourdieu 1990). Models of social
+influence assume that people adapt their culture in the face of new
+information (Goldberg and Stein 2018), while the emphasis on cohort
+effects in models of aggregate social and cultural change requires them
+to be open to change while young but become fairly resistant to it as
+they age (Ryder 1965). Finally, life course theories posit important
+changes over time as people advance through important transitions in
+their lives, but that certain life stages have more lasting influence
+than others (Bardi et al. 2009; Elder, Johnson, and Crosnoe 2003).
+
+Recent work on this topic has converged on a general conclusion that
+people do sometimes make substantial shifts in their personal culture
+over time but that most elements of personal culture are characterized
+by low rates of persistent change during adulthood (Kiley and Vaisey
+2020; Lersch 2023). Until recently, this work has adopted a ``tournament
+of models'' approach to adjudicating different theoretical perspectives
+(Lersch 2023: p.~228). In this framework, researchers formalized
+different models of change over time and relied on model selection
+criteria---primarily the Bayesian Information Criterion (Raftery
+1995)---to classify different questions as demonstrating either ``no
+change'' or ``any change.'' The size of these piles then serves as the
+primary evidence for the truth of different arguments (Kiley and Vaisey
+2020; Vaisey and Kiley 2021).
+
+However, as consensus in the overall picture emerges, researchers are
+now turning toward understanding the conditions that produce persistent
+change in personal culture, including age, education, and life-course
+transitions; whether certain kinds of people are more likely to make
+persistent changes in personal culture; and whether certain kinds of
+personal culture are more likely to be characterized by persistent
+change or stability. The tournament of models approach, useful for the
+broad question of whether change happens, is ill-suited to addressing
+these more theoretically grounded questions.
+
+In a recent paper, Lersch (2023) takes an important step forward by
+devising a Life Course Adaption Model (LCAM), a mixed-effects growth
+curve model that quantifies various components of cultural difference:
+early imprinting, persistent change, biographical experience, and
+fluctuation. In quantifying these aspects of cultural difference, the
+LCAM lets researchers address a broader range of questions than the
+single question of whether a measure is characterized by stability or
+change. However, because the LCAM produces so many different measures,
+it is not obvious how to use these components to understand the relative
+importance of change, espeically when change is measured in non-linear
+ways.
+
+In this paper, we build on the LCAM by outlining a single metric that
+combines diverse components of the LCAM to quantify the relative
+importance of intrapersonal change and interpersonal difference for
+explaining variation in a population for a variable over time. Our
+approach isolates the ``systematic'' components of variance (i.e., it
+removes random change and measurement error), to provide a single
+succinct summary of how much systematic variation can reasonably be
+attributed to people making persistent changes over time. We demonstrate
+the utility of this quantity in two ways.
+
+First, we use 609 items from seven panel surveys from five countries to
+quantify the proportion of variance that is attributable to systematic
+intrapersonal change over the study period rather than to interpersonal
+differences at baseline. This allows us to generate an overall picture
+of how much of the total systematic variation in personal culture change
+tends to explain. These findings largely echo previous findings but
+enable clearer statements about the relative importance of change across
+questions, rather than binary claims about whether change ``happens'' or
+specific quantification of how much change occurs.
+
+Second, to illustrate how this quantity might be used by a researcher
+with a specific topical interest, we investigate differences in the
+relative importance of intrapersonal change for people with and without
+a college degree. We find that change explains a smaller proportion of
+systematic variance among college-educated respondents, suggesting that
 college crystallizes one's personal culture rather than fostering an
@@
 theorize about the role of education in influencing personal culture.
-
-Taken together, our goal in this paper is to advance---and, hopefully,
-transcend---debates about \emph{whether} adults change. People do
-change, at least a little, on most things. Our method can help quantify
-exactly \emph{how much}.
 
@@
 Kiley 2021). This lack of durable change is consistent with other recent
-findings that cohort replacement plays a somewhat larger role than
-period effects in explaining differences in personal culture (Vaisey and
-Lizardo 2016).
+findings that cohort replacement plays a larger role than period effects
+in explaining differences in personal culture (Vaisey and Lizardo 2016).
 
@@
 (LCAM) as an alternative. This model draws on the life course
-perspective to model personal culture as a different linear trajectory
+perspective to model personal culture as a different linear
+trajectory\footnote{In supplemental materials, Lersch models age using
+  splines as well, with results generally showing the same conclusions.}
 over the duration of a panel for each respondent. In doing this, Lersch
-rectified two shortcomings of the AUM and SDM. First, the AUM, as
+rectified three shortcomings of the AUM and SDM. First, the AUM, as
 described by Kiley and Vaisey, posits that changes follow a Markov
@@
 
-Second, Kiley and Vaisey's analysis of the AUM and SDM is based on
-three-wave panel data over four years, which might not be intensive or
-extensive enough to adjudicate the two models. Lersch evaluates the LCAM
-against the AUM and SDM using panel data spanning a wider duration (from
-3 to 36 years) and more waves (3 to 18), offering a better chance to
-observe durable change.
+Second, Kiley and Vaisey's conclusions about the AUM and SDM were based
+on examining three-wave panel data over four years, which might not be
+intensive or extensive enough to adjudicate the two models. Lersch
+evaluates the LCAM against the AUM and SDM using panel data spanning a
+wider duration (from 3 to 36 years) and more waves (3 to 18), offering a
+better chance to observe durable change.
+
+Third, and most importantly, the LCAM quantifies various components of
+cultural difference, allowing researchers to ask a more varied set of
+questions than simply whether people change. For example, researchers
+can use the LCAM framework to quantify the magnitude of change in a
+measure of culture over time (i.e., how much do people change on
+average), the variance in age-based slopes over time (i.e., do people
+tend to change in similar ways as they age), or how much personal
+culture tends to fluctuate around these age-based trends (how consistent
+are responses over time).
 
 When Lersch compared the LCAM to the AUM and SDM on data from five
-countries, 297 of the 428 questions he analyzed preferred the LCAM,
-suggesting that we observe some linear change for adults on most items.
-The SDM was preferred on 112 items, and the rest did not yield
-conclusive results. No questions favored the AUM. He concludes that
-``new experiences over the life course {[}\ldots{]} can persistently
-move individuals' personal culture in novel directions'' (Lersch 2023:
-p.~243-244).
+countries, 198 of the 428 questions he analyzed preferred the LCAM and
+113 preferred the AUM, suggesting that we observe some change for adults
+on most items. The SDM was preferred on 93 items, and the rest did not
+yield conclusive results. He concludes that ``new experiences over the
+life course {[}\ldots{]} can persistently move individuals' personal
+culture in novel directions'' (Lersch 2023: p.~243-244).
 
@@
 that the majority of items they tested favored the AUM, meaning there
-was evidence of durable change among adults, a pattern consistent with
-Lersch's results. Also Lersch (2023) found evidence of durable change on
-most items, but these changes were small. On average, people changed
-only about .07 standard deviations over 10 years. This is consistent
-with Kiley and Vaisey's finding that durable change is often small.
-Moreover, a quarter of the items Lersch studied still preferred the SDM,
-suggesting that even with different assumptions about change and more
-extensive data, many questions still are consistent with a model that
-allows zero intrapersonal change.
+was evidence of persisting change among adults, a pattern consistent
+with Lersch's results. Also Lersch (2023) found evidence of persisting
+change on most items, but these changes were small. On average, people
+changed only about .07 standard deviations over 10 years. This is
+consistent with Kiley and Vaisey's finding that persisting change is
+often small. Moreover, a quarter of the items Lersch studied still
+preferred the SDM, suggesting that even with different assumptions about
+change and more extensive data, many questions still are consistent with
+a model that allows zero intrapersonal change.
 
@@
 
-To this point, the debate has been about whether people \emph{ever}
-change. As we argued above, this is rarely the right question since the
-answer is almost always ``yes.'' Over sufficient time and in a large
-enough sample, researchers will likely observe some evidence for durable
-individual change. A lack of evidence for change may be due to a poor
-survey question, a low resolution in response options, or that the
-question simply was not asked for long enough or to enough people.
-Conversely, finding evidence for some change tells us only that it was
-``not zero.'' From that alone, we learn little about how much
-intrapersonal change has happened in a population.
+To this point, the debate has broadly centered on drawing conclusions
+about whether people \emph{ever} change. But this is rarely the right
+question since the answer is almost always ``yes.'' Over sufficient time
+and in a large enough sample, researchers will likely observe some
+evidence for durable change in the population. A lack of evidence for
+change may be due to a poor survey question, a low resolution in
+response options, or that the question simply was not asked for long
+enough or to enough people. Conversely, finding evidence for some change
+tells us only that it was ``not zero.'' From that alone, we learn little
+about how much intrapersonal change has happened in a population.
 
@@
 culture. Lersch's LCAM does this by modeling each individual as
-following a personal linear trajectory. But rather than debating whether
-this model fits a particular data set better than a model that assumes
-that people never change, it would be better to quantify these
-individual changes and compare them to the stable differences that exist
-between people. In other words, the focus should shift from asking, ``do
-people change?'' to asking, ``what are the relative contributions of
-change and pre-existing differences for explaining variance in personal
-culture?''
-
-We propose an approach---which we formalize below---that offers such a
-quantification. Our measure relates two variance components that each
-reflect a combination of relevant theoretical processes. The first
-component, stable interpersonal differences, reflects the accumulated
-experiences of people prior to entering the panel survey. Lersch (2023:
-24) calls this ``early imprinting.'' While commonly associated with
-experiences during a formative period that result in settled
-dispositions, this variation could also reflect experiences that
-happened at any time as long as they predate the panel and consistently
-affect subsequent responses. For instance, for those people entering the
-panel post-retirement, this ``imprinting'' might reflect this pivotal
-life transition. Consequently, this component also reflects variation in
+following a personal linear trajectory. In doing so, the LCAM opens the
+door to a variety of questions. We assert that among the many things
+that can be done with the LCAM, it provides the opportunity to shift
+from asking, ``do people change?'' to asking, ``what are the relative
+contributions of change and pre-existing differences for explaining
+variance in personal culture?''
+
+Our approach---which we formalize below---offers such a quantification.
+Our measure relates two variance components that each reflect a
+combination of relevant theoretical processes. The first component,
+stable interpersonal differences, reflects the accumulated experiences
+of people prior to entering the panel survey. Lersch (2023: 24) calls
+this ``early imprinting.'' While commonly associated with experiences
+during a formative period that result in settled dispositions, this
+variation could also reflect experiences that happened at any time as
+long as they predate the panel and consistently affect subsequent
+responses. For instance, for those people entering the panel
+post-retirement, this ``imprinting'' might reflect this pivotal life
+transition. Consequently, this component also reflects variation in
 individuals' social roles or statuses at the start of the panel that
@@
 
-The second component, the amount of intrapersonal linear change people
-make during the panel, captures durable changes in personal culture over
-time. This captures the set of processes collectively called
+The second component, the amount of intrapersonal change people make
+during the panel, captures persisting changes in personal culture over
+time. This component captures the set of processes collectively called
 ``persistent change'' or ``adaption'' by Lersch and ``active updating''
@@
 Furthermore, these components are not necessarily useful in isolation or
-without context. What is important for the theoretical debate is the
-relative contribution of these two components in explaining cultural
-differences in a population. In other words, when we look at a
-population of adults over some time frame, is there enough intrapersonal
-change relative to interpersonal difference to view it as an important
-process for why people differ from each other? Only through quantifying
-the two components we can get closer to a true answer to this question.
+without context. If absolute change is about .07 standard deviations
+every 10 years, but the variance of starting points at age 18 is very
+high, then change over time will not produce significant differences
+among people, even over a long time frame. If, on the other hand, people
+tend to start at very similar positions and change at a rate of .07
+standard deviations, then that change is very useful for understanding
+difference. What is important for the theoretical debate is the relative
+contribution of these two components in explaining cultural differences
+in a population. In other words, when we look at a population of adults
+over some time frame, is there enough intrapersonal change relative to
+interpersonal difference to view it as an important process for why
+people differ from each other? Only through quantifying the two
+components we can get closer to a true answer to this question.
 
@@
 societies.
+
+\hypertarget{formalization}{%
+\subsection{Formalization}\label{formalization}}
+
+The starting point for our quantity is Lersch's (2023) Life Course
+Adaption Model, which formalizes survey responses at time \(t\) as a
+function of individual-level random intercepts and slopes for survey
+age. In some contexts, this model is called a mixed-effects growth curve
+model. This model assumes a set of propositions about change that
+reflect the theoretical debate to this point. First, consistent with the
+settled disposition model, it assumes that people start the survey with
+cultural differences, modeled as random intercepts for each respondent.
+Second, it assumes that people change over time, taking the form of
+random slopes for each respondent as a linear function of time. Third,
+it assumes that people deviate around this baseline randomly over time,
+reflecting ``fluctuation'' or short-term non-persistent change.
+Formally, this can be written as
+
+\[
+y_{it} = \beta_0 + \alpha_{0i} + (\beta_1 + \alpha_{1i}) \text{year}_{it} + \epsilon_{it}
+\]
+
+\[
+\begin{aligned}
+\alpha_{0i} &\sim \mathcal{N}(0,\tau^2_{0}) \\
+\alpha_{1i} &\sim \mathcal{N}(0,\tau^2_{1}) \\
+\epsilon_{it} &\sim \mathcal{N}(0,\sigma^2)
+\end{aligned}
+\]
+
+where \(\beta_0\) is the average intercept, \(\alpha_{0i}\) is the
+random intercept component for individuals, \(\beta_1\) is the average
+yearly change in the outcome, \(\alpha_{1i}\) is the random slope
+component for individuals, and \(\epsilon_{it}\) is the random error
+term that captures transitory disturbances. The \(\alpha_{0i}\) and
+\(\alpha_{1i}\) terms are also allowed to covary.
+
+We then use these models to derive two measures of the components we
+outlined above: pre-existing interpersonal differences and intrapersonal
+change over time.
+
+We estimate these using a form of variance decomposition. First, to
+measure stable interpersonal differences, we calculate \(V(D)\)
+according to the equation below, where \(\widetilde{year_i}\) is the
+midpoint of the observed years for each respondent for that particular
+item.\footnote{We use the model estimate from the midpoint year, rather
+  than the value at the first wave, because this is the best measure of
+  ``baseline'' available under the assumption that the person does not
+  change. The first wave measure alone contains an unknown amount of
+  measurement error and transient fluctuations.}
+
+\[
+V(D) = 1 - \frac{\sum_{i=1}^N \sum_{t=1}^T [y_{it} - (\hat{\beta}_0 + \hat{\alpha}_{0i} + (\hat{\beta}_1 + \hat{\alpha}_{1i}) \widetilde{\text{year}}_{i})]^2}{\sum_{i=1}^N \sum_{t=1}^T [y_{it} - \bar{y}_{it}]^2}
+\]
+
+This gives us the proportion of total variance in an item attributable
+to stable interpersonal differences.
+
+To calculate the second component, which is the variance attributable to
+systematic intrapersonal change, we calculate \(V(C)\) as follows:
+
+\[
+V(C) = 1 - \frac{\sum_{i=1}^N \sum_{t=1}^T [y_{it} - (\hat{\beta}_0 + \hat{\alpha}_{0i} + (\hat{\beta}_1 + \hat{\alpha}_{1i}) \text{year}_{it})]^2}{\sum_{i=1}^N \sum_{t=1}^T [y_{it} - \bar{y}_{it}]^2} - V(D)
+\]
+
+This is the incremental proportion of variance accounted for when we
+allow the model predictions to change over time for each person.
+
+The third component, which is the proportion of variance attributable to
+measurement error or transient fluctuations, is what's left, such that
+these three numbers sum to 1. As we discussed above, this component is
+not of core interest here, so we set it aside in our analysis.
+
+Together, \(V(D)\) and \(V(C)\) reflect the total systematic variance of
+the outcome that the LCAM can attribute to either differences between
+people or linear intrapersonal change over the course of the panel. Our
+principal measure of interest is the proportion of systematic variance
+accounted for by intrapersonal change, which we calculate as
+
+\[
+\omega = \frac{V(C)}{V(C) + V(D)}
+\]
+
+We can think of \(\omega\) as a type of Intraclass Correlation
+Coefficient focused only on the systematic components of the model while
+ignoring residual variance that might be attributable to measurement
+error or non-persisting change (see also the variance decomposition of
+period and cohort by Vaisey and Lizardo 2016). The higher this
+proportion is, the more systematic variance in responses is attributable
+to intrapersonal change during the panel. From now on, we use \(\omega\)
+to stand for \emph{the proportion of systematic variance attributable to
+intrapersonal change}.
+
+Because this quotient summarizes variance proportions and therefore has
+no natural referent, it is hard to say at which values we observe ``a
+lot'' of intrapersonal change. But summarizing questions in this way
+allows us to compare the relative importance of our two broad processes
+across a range of questions that might have different levels of
+non-systematic variance. This gives us a sense of the relative
+prevalence of these two processes for each question, which is not
+achievable under the tournament of models approach. This also allows us
+to compare which kinds of questions show more or less intrapersonal
+change relative to interpersonal differences. While the tournament of
+models approach allowed researchers to say whether questions in general
+showed \emph{any} evidence of updating, this approach allows us to make
+more specific claims about the prevalence of these theoretical processes
+for particular questions relative to each other.
+
+\hypertarget{bootstrapping-for-confidence-intervals}{%
+\subsection{Bootstrapping for Confidence
+Intervals}\label{bootstrapping-for-confidence-intervals}}
+
+{[}Explain how we do that here{]}
 
@@
 
-\hypertarget{life-course-adaption-model}{%
-\subsection{Life Course Adaption
-Model}\label{life-course-adaption-model}}
-
-The estimation procedure begins with Lersch's (2023) Life Course
-Adaption Model, which formalizes survey responses at time \(t\) as a
-function of individual-level random intercepts and slopes for survey
-age. In some contexts, this model is called a mixed-effects growth curve
-model. This model assumes a set of propositions about change that
-reflect the theoretical debate to this point. First, consistent with the
-settled disposition model, it assumes that people start the survey with
-cultural differences, modeled as random intercepts for each respondent.
-Second, it assumes that people change over time, taking the form of
-random slopes for each respondent as a linear function of time. Third,
-it assumes that people deviate around this baseline randomly over time,
-reflecting ``fluctuation'' or short-term non-persistent change.
-Formally, this can be written as
-
-\[
-y_{it} = \beta_0 + \alpha_{0i} + (\beta_1 + \alpha_{1i}) \text{year}_{it} + \epsilon_{it}
-\]
-
-\[
-\begin{aligned}
-\alpha_{0i} &\sim \mathcal{N}(0,\tau^2_{0}) \\
-\alpha_{1i} &\sim \mathcal{N}(0,\tau^2_{1}) \\
-\epsilon_{it} &\sim \mathcal{N}(0,\sigma^2)
-\end{aligned}
-\]
-
-where \(\beta_0\) is the average intercept, \(\alpha_{0i}\) is the
-random intercept component for individuals, \(\beta_1\) is the average
-yearly change in the outcome, \(\alpha_{1i}\) is the random slope
-component for individuals, and \(\epsilon_{it}\) is the random error
-term that captures transitory disturbances. The \(\alpha_{0i}\) and
-\(\alpha_{1i}\) terms are also allowed to covary.
-
 \hypertarget{analysis-steps}{%
@@
 As a first step in our analysis, we fit this LCAM to each of the 609
-measures of personal culture outlined above. We then use these models to
-derive two measures of the components we outlined above: pre-existing
-interpersonal differences and intrapersonal change over time.
-
-We estimate these using a form of variance decomposition. First, to
-measure stable interpersonal differences, we calculate \(V(D)\)
-according to the equation below, where \(\widetilde{year_i}\) is the
-midpoint of the observed years for each respondent for that particular
-item.\footnote{We use the model estimate from the midpoint year, rather
-  than the value at the first wave, because this is the best measure of
-  ``baseline'' available under the assumption that the person does not
-  change. The first wave measure alone contains an unknown amount of
-  measurement error and transient fluctuations.}
-
-\[
-V(D) = 1 - \frac{\sum_{i=1}^N \sum_{t=1}^T [y_{it} - (\hat{\beta}_0 + \hat{\alpha}_{0i} + (\hat{\beta}_1 + \hat{\alpha}_{1i}) \widetilde{\text{year}}_{i})]^2}{\sum_{i=1}^N \sum_{t=1}^T [y_{it} - \bar{y}_{it}]^2}
-\]
-
-This gives us the proportion of total variance in an item attributable
-to stable interpersonal differences.
-
-To calculate the second component, which is the variance attributable to
-systematic intrapersonal change, we calculate \(V(C)\) as follows:
-
-\[
-V(C) = 1 - \frac{\sum_{i=1}^N \sum_{t=1}^T [y_{it} - (\hat{\beta}_0 + \hat{\alpha}_{0i} + (\hat{\beta}_1 + \hat{\alpha}_{1i}) \text{year}_{it})]^2}{\sum_{i=1}^N \sum_{t=1}^T [y_{it} - \bar{y}_{it}]^2} - V(D)
-\]
-
-This is the incremental proportion of variance accounted for when we
-allow the model predictions to change over time for each person.
-
-The third component, which is the proportion of variance attributable to
-measurement error or transient fluctuations, is what's left, such that
-these three numbers sum to 1. As we discussed above, this component is
-not of core interest here.
-
-Together, \(V(D)\) and \(V(C)\) reflect the total systematic variance of
-the outcome that the LCAM can attribute to either differences between
-people or linear intrapersonal change over the course of the panel. Our
-principal measure of interest is the proportion of systematic variance
-accounted for by intrapersonal change, which we calculate as
-
-\[
-\omega = \frac{V(C)}{V(C) + V(D)}
-\]
-
-We can think of \(\omega\) as a type of Intraclass Correlation
-Coefficient focused only on the systematic components of the model while
-ignoring residual variance that might be attributable to measurement
-error or non-persisting change (see also the variance decomposition of
-period and cohort by Vaisey and Lizardo 2016). The higher this
-proportion is, the more systematic variance in responses is attributable
-to intrapersonal change during the panel. From now on, we use \(\omega\)
-to stand for \emph{the proportion of systematic variance attributable to
-intrapersonal change}.
-
-Because this quotient summarizes variance proportions and therefore has
-no natural referent, it is hard to say at which values we observe ``a
-lot'' of intrapersonal change. But summarizing questions in this way
-allows us to compare the relative importance of our two broad processes
-across a range of questions that might have different levels of
-non-systematic variance. This gives us a sense of the relative
-prevalence of these two processes for each question, which is not
-achievable under the tournament of models approach. This also allows us
-to compare which kinds of questions show more or less intrapersonal
-change relative to interpersonal differences. While the tournament of
-models approach allowed researchers to say whether questions in general
-showed \emph{any} evidence of updating, this approach allows us to make
-more specific claims about the prevalence of these theoretical processes
-for particular questions relative to each other.
-
-As a second step in our analysis, we use linear regression to conduct a
-meta-analysis, modeling \(\omega\) as a function of features of the
-question and panel. This allows us to explore how these features are
-associated with the degree of intrapersonal change. These covariates
-include the number of response options (measured as 2, 3-5, and more
-than 5), the specific survey, the log of the number of participants, the
-number of waves observed, the time period (i.e., decade) the question
-was first asked, and the total duration of time the question was asked.
+measures of personal culture outlined above. We then derive \(\omega\)
+for each of these items.
+
+As a second step in our analysis, we use linear regression to model
+\(\omega\) as a function of features of the question and panel. This
+allows us to explore how these features are associated with the degree
+of intrapersonal change. These covariates include the number of response
+options (measured as 2, 3-5, and more than 5), the specific survey, the
+log of the number of participants, the number of waves observed, the
+time period (i.e., decade) the question was first asked, and the total
+duration of time the question was asked.
 
@@
 
-The objective of this paper is to intervene in the recent debate on
-whether people change their personal culture -- their attitudes,
-beliefs, values, and practices -- as they move through their adult life.
-Instead of falling back on a ``tournament of models'' approach (Lersch
-2023: p.~228) to proclaim victory for a particular answer to this
-question, our goal is to refocus the debate to the theoretically more
-productive question of the relative importance of intrapersonal change
-for explaining differences in the personal culture. We propose a new
-approach that quantifies the amount of observed systematic variance that
-is attributable to either interpersonal differences at baseline or
-intrapersonal change over the duration of a panel.
+The primary objective of this paper is to introduce a new quantity that
+can advance debate on whether people change their personal culture --
+their attitudes, beliefs, values, and practices -- as they move through
+their adult life. Instead of falling back on a ``tournament of models''
+approach (Lersch 2023: p.~228) to proclaim victory for a particular
+answer to this question, our goal is to refocus the debate to the
+theoretically more productive question of the relative importance of
+intrapersonal change for explaining differences in personal culture. We
+propose a new approach that quantifies the amount of observed systematic
+variance that is attributable to either interpersonal differences at
+baseline or intrapersonal change over the duration of a panel.
 
@@
 More than adjudicating past positions in this debate, the approach we
-presented in this paper will enable scholars to explore more precisely
-the relative contributions of change and stability in explaining
-variation in personal culture, as well as how these contributions might
-differ across populations. To illustrate this, we showed how the amount
-of cultural difference explained by intrapersonal change varies
-substantially by survey item but also by individuals' characteristics
-such as their education. While age and other factors related to college
-completion might confound this pattern, it suggests that college
-completion may crystallize personal culture to an extent that renders
-later adult experiences on political dispositions less influential.
+presented in this paper will enable researchers to explore more
+precisely the relative contributions of change and stability in
+explaining variation in personal culture, as well as how these
+contributions might differ across populations. To illustrate this, we
+showed how the amount of cultural difference explained by intrapersonal
+change varies substantially by survey item but also by individuals'
+characteristics such as their education. While age and other factors
+related to college completion might confound this pattern, it suggests
+that college completion may crystallize personal culture to an extent
+that renders later adult experiences on political dispositions less
+influential.
 
@@
 
-Although we believe it is a major advance on previous work, the approach
-we outline here has important limitations. We allocate all systematic
-variance to one of two sets of theoretical processes: intrapersonal
-change and interpersonal differences at baseline. Our approach does not
-quantify the proportion of people who ``change,'' nor can we be sure
-that the amount of interpersonal change we detect is driven by many
-people making small changes or a few people making large changes.
+Although we believe the approach we outline here advances previous work,
+it has important limitations. We allocate all systematic variance to one
+of two sets of theoretical processes: intrapersonal change and
+interpersonal differences at baseline. Our approach does not quantify
+the proportion of people who ``change,'' nor can we be sure that the
+amount of interpersonal change we detect is driven by many people making
+small changes or a few people making large changes.
 
 Our results and interpretations also hinge on how we have defined
-change. As does the LCAM, our approach treats cultural trajectories as
-varying linear slopes for each respondent, thus assuming that change is
-a linear function of time. This assumption simplifies reality in which
-change likely also takes non-linear and discontinuous forms. People
-might jump from one ``stable'' disposition to another or experience a
-``turning point'' in their lives that upsets an otherwise stable
-trajectory. In parts, our assumption of linearity is a limitation of the
-data, as most questions are only observed for three to six waves. Panels
-with more waves might allow researchers to loosen this assumption to
-test alternative, more flexible models of change.
+change. As does the LCAM, our analysis here treats cultural trajectories
+as varying linear slopes for each respondent, thus assuming that change
+is a linear function of time. This assumption simplifies reality in
+which change likely also takes non-linear and discontinuous forms.
+People might jump from one ``stable'' disposition to another or
+experience a ``turning point'' in their lives that upsets an otherwise
+stable trajectory. In parts, our assumption of linearity is a limitation
+of the data, as most questions are only observed for three to six waves.
+However, our overall approach does not require modeling change as a
+linear function of age. Change can be modeled in non-linear forms, such
+as the splines used by Lersch (2023), and accordingly incorporated into
+the calculation of variance explained.
 
@@
 than intrapersonal change. Longer panels and more flexible definitions
-of ``change'' might allow us to account for such trajectories.
+of ``change'' might allow us to account for such trajectories. Again,
+while we model change in a single way, the general variance
+decomposition approach we outline accommodates different definitions and
+functional forms of change.
 
@@
 about the \emph{existence} of cultural change in adults (e.g., Kiley and
-Vaisey 2020; Lersch 2023; Vaisey and Kiley 2021). Yet this approach has
-inadvertently limited the debate. In any population, \emph{some} degree
-of adult cultural change is inevitable. Although ideal types like the
-settled dispositions and active updating models are useful, no
-theoretical perspective would expect either early life socialization or
-adult intrapersonal change to be the sole source of one's personal
-culture. Our results reinforce this point, showing the relevance of both
-factors and allowing their precise quantification.
+Vaisey 2020; Lersch 2023; Vaisey and Kiley 2021). In any population,
+\emph{some} degree of adult cultural change is inevitable. Although
+ideal types like the settled dispositions and active updating models are
+useful, no theoretical perspective would expect either early life
+socialization or adult intrapersonal change to be the sole source of
+one's personal culture. Our results reinforce this point, showing the
+relevance of both factors and allowing their precise quantification.
 
 This and other recent findings (Quinn et al. 2023; Stewart and Berkman
-2023) suggest that it is theoretically more productive to measure the
-relative importance of these two components in concrete, substantive
-cases. Drawing a unified conclusion from survey items measuring various
-cultural forms on different scales and across different time frames is
-challenging. Nevertheless, the general pattern suggests that, for most
-items, intrapersonal change in adulthood is not the primary reason for
-the differences we see between people in the world.
+2023; \textbf{quinn2024?}) suggest that it is theoretically more
+productive to measure the relative importance of these two components in
+concrete, substantive cases. Drawing a unified conclusion from survey
+items measuring various cultural forms on different scales and across
+different time frames is challenging. Nevertheless, the general pattern
+suggests that, for most items, intrapersonal change in adulthood is not
+the primary reason for the differences we see between people in the
+world.
 
@@
 We believe the approach we outlined here can push past the ``needless
-dichotomy'' implicit in the question of whether people change or not
-(Lersch 2023). Characterizing questions as displaying change or not can
-only take researchers so far, but the question of whether some questions
+dichotomy'' implicit in the question of whether people change or not.
+Characterizing questions as displaying change or not can only take
+researchers so far, but the question of whether some questions
 demonstrate more change than others, or whether some groups are
@@
 \begin{table}[H]
-
+\centering
 \caption{\label{tab:unnamed-chunk-1}The Description of the Data Sources}
 \centering
-\resizebox{\linewidth}{!}{
+\resizebox{\ifdim\width>\linewidth\linewidth\else\width\fi}{!}{
 \begin{tabular}[t]{llcc}

### 2024-03-15 manuscript/manuscript.Rmd
kkedits315

@@
 
-An important contemporary debate in sociology centers on the importance of intrapersonal change for understanding differences in "personal culture" -- attitudes, beliefs, values, and practices [@lizardo2017; @kiley2020; @lersch2023; @underwood2021; @vaisey2016].
+An important contemporary debate in sociology centers on the importance of individual change over time for understanding differences in "personal culture"---attitudes, beliefs, values, and practices [@lizardo2017; @kiley2020; @lersch2023; @underwood2022; @vaisey2016].
 This question has deep roots in seemingly contradictory theoretical perspectives.
@@
 Recent work on this topic has converged on a general conclusion that people do sometimes make substantial shifts in their personal culture over time but that most elements of personal culture are characterized by low rates of persistent change during adulthood [@kiley2020; @lersch2023].
-Until recently, this work has adopted a "tournament of models" approach to adjudicating different theoretical perspectives [@lersch2023: p.
+Work leading to this conclusion adopted a "tournament of models" approach to adjudicating different theoretical perspectives [@lersch2023: p.
 228].
@@
 In quantifying these aspects of cultural difference, the LCAM lets researchers address a broader range of questions than the single question of whether a measure is characterized by stability or change.
-However, because the LCAM produces so many different measures, it is not obvious how to use these components to understand the relative importance of change, espeically when change is measured in non-linear ways.
+However, because the LCAM produces so many different measures, it is not obvious how to use these components to understand the relative importance of within-person change, especially when change is measured in non-linear ways.
 
-In this paper, we build on the LCAM by outlining a single metric that combines diverse components of the LCAM to quantify the relative importance of intrapersonal change and interpersonal difference for explaining variation in a population for a variable over time.
+In this paper, we build on the LCAM by outlining a single metric that combines components of the LCAM to quantify the relative importance of intrapersonal change and interpersonal difference for explaining variation in a population for a variable over time.
 Our approach isolates the "systematic" components of variance (i.e., it removes random change and measurement error), to provide a single succinct summary of how much systematic variation can reasonably be attributed to people making persistent changes over time.
+This quantity can then be used as a quick diagnostic to compare across questions, groups, periods, or samples to isolate where within-person change is more or less useful for explaining variation.
 We demonstrate the utility of this quantity in two ways.
@@
 Third, and most importantly, the LCAM quantifies various components of cultural difference, allowing researchers to ask a more varied set of questions than simply whether people change.
-For example, researchers can use the LCAM framework to quantify the magnitude of change in a measure of culture over time (i.e., how much do people change on average), the variance in age-based slopes over time (i.e., do people tend to change in similar ways as they age), or how much personal culture tends to fluctuate around these age-based trends (how consistent are responses over time).
+For example, researchers can use the LCAM framework to quantify the magnitude of change in a measure of culture over time (i.e., how much do people change on average?), the variance in age-based slopes over time (i.e., do people tend to change in similar ways as they age?), or how much personal culture tends to fluctuate around these age-based trends (i.e., how consistent are responses over time?).
 
@@
 If absolute change is about .07 standard deviations every 10 years, but the variance of starting points at age 18 is very high, then change over time will not produce significant differences among people, even over a long time frame.
-If, on the other hand, people tend to start at very similar positions and change at a rate of .07 standard deviations, then that change is very useful for understanding difference.
+If, on the other hand, people tend to start at very similar positions and change at a rate of .07 standard deviations for every 10 years, then that change is very useful for understanding difference later in life.
 What is important for the theoretical debate is the relative contribution of these two components in explaining cultural differences in a population.
@@
 
-Yet the true utility of the proposed quantification goes beyond settling past debates; it allows researchers to ask new questions.
+Yet the true utility of the proposed quantification goes beyond settling past debates; it allows researchers to identify new questions.
 For example, classifying survey items based on whether they show durable change overlooks possibly important differences among those that do show such change.
@@
 
+This effectively models respondents' culture under the assumption that they make no systematic change over time, or that the expected value at any time point is the same as the expected value at any other.
 This gives us the proportion of total variance in an item attributable to stable interpersonal differences.
@@
 
-## Expectations
+# Example 1: Cross-Question Comparisons
 
-By shifting the question and adjusting the approach, we expect consistent patterns across datasets.
-First, because previous work found evidence of durable change on most measures of personal culture, we expect that we will find some intrapersonal change in most items.
-But because these earlier studies also found many questions (between 25 and 35 percent) lacking evidence of *any* durable change, we expect considerable variation in the proportion of variance accounted for by intrapersonal change, with some questions indicating essentially none at all.
-
-Second, we expect the variance attributable to intrapersonal change to be modest compared to pre-existing interpersonal differences.
-This expectation should be uncontroversial, regardless of the theoretical process assumed to be most relevant for a particular question.
-Panel surveys do not encompass individuals' entire lifetimes.
-Thus, even if intrapersonal change during adulthood is the major driver of cultural differences, the change detected in a panel covering 20 years probably will not fully account for the entire variance of interpersonal differences.
-
-Our expectations so far largely echo what was observed in previous studies.
-However, theoretical perspectives differ in how much change they expect, enabling quantification to provide a new perspective on existing questions.
-"Formative period" and cohort theories, which emphasize early-life socialization as a strong influence on adult personal culture, imply low amounts of intrapersonal change, even when we account for the duration of observation.
-Conversely, other theoretical perspectives emphasizing the importance of people's contemporaneous social environment might expect a larger degree of intrapersonal change, even in a relatively short panel.
-
-Third, we expect the variance explained by each process---interpersonal differences and intrapersonal change---to depend on specific attributes of the question and panel.
-First, the higher the response resolution for a given question and the larger the sample, the greater the variance attributed to intrapersonal change should be, as smaller changes in intrapersonal change will be easier to detect.
-Second, life course theories posit that social transitions drive changes in personal culture.
-Since the probability that any person experiences such a transition goes up the longer we observe them, we expect panel duration to be positively related to the proportion of variation explained by intrapersonal change.
-
-# Analytic Strategy
+To demonstrate the utility of this metric, we provide two examples. In the first, we compare the proportion of systematic variance explained by intrapersonal change across a broad array of questions.
 
@@
 
-## Analysis Steps
+## Analysis
 
-As a first step in our analysis, we fit this LCAM to each of the 609 measures of personal culture
-outlined above. We then derive $\omega$ for each of these items.
-
-As a second step in our analysis, we use linear regression to model $\omega$ as a function of features of the question and panel.
+As a first step in our analysis, we fit this LCAM to each of the 609 measures of personal culture outlined above.
+We then derive $\omega$ for each of these items.
+As a second step, we use linear regression to model $\omega$ as a function of features of the question and panel.
 This allows us to explore how these features are associated with the degree of intrapersonal change.
@@
 
-# Results
-
-## Variance Decomposition
+## Results
 
@@
 
-## Meta-Analysis
-
 Figure 2 plots the results from a linear regression of $\omega$ as a function of question, panel, and sample features.
@@
 
-## College and Change in Political Culture
+# Example 2: College and Change in Political Culture
 
 These results mostly re-frame and align previous findings, but the value of our approach lies in its ability to extend the debate to a broader set of theoretical questions.
-To show this potential, we turn to a specific empirical example: the relative importance of intrapersonal change and interpersonal difference for explaining cultural variation by level of education.
+As a second empirical application of $\omega$, we explore the relative importance of intrapersonal change and interpersonal difference for explaining cultural variation by level of education.
 
@@
 
-There are theoretical reasons to believe that education might be associated with either more or less intrapersonal change.
+There are theoretical reasons to believe that education might be associated with either more or less intrapersonal change over time.
 On one hand, because college graduates are more connected to mainstream discourse and elite signals, they might be more likely to make durable changes in response to the emergence of new information, new issues, or political realignments, while those without chronic information might display more variance around an unchanging baseline [@zaller1992].
-Conversely, it could be that the observed stability of the college educated reflects the fact that they have already formed durable opinions and are relatively closed off to new information.
+Conversely, because people with less education appear more likely to engage in automatic cognition in social situations (including survey interviews) than people with more education [@brett2021], they might be more influenced by changes in the social environment, producing more temporary change and more durable change.
+It could be that the observed stability of the college educated also reflects the fact that they have already formed durable opinions across various issues and are relatively closed off to new information.
 If this is true, college could be understood as a formative experience that solidifies some dimensions of personal culture.
 Perhaps for those who do not attend college, later life experiences might prove more important in forming or changing personal culture, as these experiences potentially provide information that college-educated peers have already received.
+
+## Data
 
@@
 Estimating the panel with highest at wave 1 or highest degree reported across the panel produces functionally identical results.
+
+## Results
 
@@
 
-On all eight questions presented in Figure 4, $\omega$ values are smaller for people with a college degree, meaning intrapersonal change is less common among them.
+On all eight questions presented in Figure 4, $\omega$ values are smaller for people with a college degree, meaning intrapersonal change is less important for explaining variation among them.
 There are also large differences in $\omega$ values across questions for both groups.
@@
 In contrast, partisan identification and political ideology both show larger values of $\omega$ than most other questions, as well as a larger absolute difference by degree status.
-Compared to the other questions, intrapersonal change plays a much larger role in accounting for partisan identification and political ideology.
+Compared to the other questions, intrapersonal change plays a much larger role in accounting for variation in partisan identification and political ideology.
 And this is particularly true for respondents who do not have a college degree; the $\omega$ value for ideological identification among non-college educated respondents is almost four times that of college-educated respondents.
@@
 We believe these patterns shed new light on the mechanisms underlying differences in attitudes and behavior across groups.
-Something about college attendance seems to crystallize personal culture.
+Something about college attendance seems to affect the role played by change in explaining variance.
+It could to this in either two ways: by creating greater variation in baseline views or by reducing change over time relative to people without college degrees.
 While these results should not be interpreted as causal effects of attending college -- they are potentially confounded by age, social class, race, gender, and other factors that explain selection into higher education -- they open up a set of new questions and dynamics to explore.
@@
 Instead of falling back on a "tournament of models" approach [@lersch2023: p.
-228] to proclaim victory for a particular answer to this question, our goal is to refocus the debate to the theoretically more productive question of the relative importance of intrapersonal change for explaining differences in personal culture.
-We propose a new approach that quantifies the amount of observed systematic variance that is attributable to either interpersonal differences at baseline or intrapersonal change over the duration of a panel.
+228] to proclaim a yes or no answer to this question, or only quantifying individual components of cultural difference to show how much people change, our metric contextualizes the amount of observed systematic variance that is attributable to intrapersonal change over the duration of a panel.
+In doing so, it provides a tool to identify questions, groups, or periods where change is a more meaningful explanation of cultural difference.
 
-Applying our proposed measure, $\omega$, to 609 survey items from all of the panel datasets previously discussed in this debate revealed a consistent pattern.
+We demonstrated the utility of this measure in two ways.
+First, applying our proposed measure, $\omega$, to 609 survey items from the panel datasets previously studied in this debate revealed a consistent pattern.
 Nearly all questions show evidence of people making some durable, intrapersonal change over time.
@@
 For some questions about life satisfaction or views on government benefits, it seems plausible that differences in adult experiences predominantly account for observed differences between people.
-
 However, intrapersonal change is often substantially less pronounced than interpersonal differences, accounting for less than 10 percent of systematic variance on average across all datasets.
-On numerous questions, like those on civil liberties, abortion, generalized trust, and civic duty, the systematic variance attributable to intrapersonal change is essentially zero.
+On questions related to civil liberties, abortion, generalized trust, and civic duty,  systematic variance attributable to intrapersonal change is essentially zero.
 For these questions, it seems there is not enough cultural change during adulthood to warrant attributing the differences we observe to experiences and social transitions; instead, the primary source of the observed differences appears to stem from experiences in childhood, adolescence, and early adulthood.
 
-More than adjudicating past positions in this debate, the approach we presented in this paper will enable researchers to explore more precisely the relative contributions of change and stability in explaining variation in personal culture, as well as how these contributions might differ across populations.
-To illustrate this, we showed how the amount of cultural difference explained by intrapersonal change varies substantially by survey item but also by individuals' characteristics such as their education.
-While age and other factors related to college completion might confound this pattern, it suggests that college completion may crystallize personal culture to an extent that renders later adult experiences on political dispositions less influential.
+While these findings generally echo the general consensus that has emerged from previous work, it adds two dimensions.
+First, it quantifies the relative importance of change across these questions, providing more resolution within what were mainly piles of questions either showing evidence of change or no evidence of change.
+This allows for greater comparison across questions and it provides a benchmark for researchers exploring new questions.
+Second, our metric can help contextualize the quantities identified in the LCAM.
+Is .07 standard deviations of change every 10 years a lot of change?
+For some questions, our metric will say this change explains a large component of variance, and for others it will not.
+
+In a second application of the quantity, we showed how it can open up new lines of inquiry.
+We showed how the amount of cultural difference explained by intrapersonal change varies substantially not just by survey item but also by individuals' characteristics such as their education.
+While age and other factors related to college completion might confound this pattern, it suggests that college completion may crystallize personal culture to an extent that renders later adult experiences on political dispositions less influential in explaining overall differences.
 
@@
 
+This is in fact one of the strengths of our approach.
+When change over time is modeled non-linearly, it can be difficult to summarize its relative importance.
+With our metric, researchers can quantify the additional explanatory power of modeling change using splies, polynomials, or other non-linear approaches relative to a linear trend.
+
 Similarly, our approach assumes that durable change is unidirectional.
@@
 
-Despite these limitations, we believe our method and findings have important implications for social science research.
+Despite these limitations, we believe our method and findings have implications for social science research.
 Sociologists interested in understanding cultural differences have largely asked about the *existence* of cultural change in adults [e.g., @kiley2020; @vaisey2021; @lersch2023].
@@
 Nevertheless, the general pattern suggests that, for most items, intrapersonal change in adulthood is not the primary reason for the differences we see between people in the world.
-
-We are not claiming that adults remain static or that their changes are inconsequential.
-Even quantitatively minor shifts, such as a 2 percentage-point change in support for gay marriage, can have massive ramifications.
-While the majority may remain consistent in their views, understanding the underlying mechanisms of even such numerically minor shifts remains a crucial task for the sociology of culture.
 
@@
 In trying to understand differences in personal culture, sociologists should therefore pay more attention to the heterogeneous effects that various factors including social events, encounters, and situations can have.
-
-Our findings also have implications for understanding cultural change at the aggregate level.
-Given that intrapersonal change accounts for a relatively small amount of variance among adults, many forms of cultural change at the aggregate level are necessarily more likely driven cohort replacement than by contemporaneous social conditions [@vaisey2016; @underwood2022].
-This likely holds true for cultural change at the macro societal level as well as the micro level such as within organizations, political parties, and professions.
-Our findings also indicate that the impact of formative experiences, as opposed to contemporaneous social factors, on cultural change is contingent upon an individuals' level of education.
-The weight of these two factors might thus change as the prominence of education shifts in the life course.
-If scholars find similar differences across other social categorizations like class or race, it might necessitate a more thorough integration of demographic processes into the analysis of cultural change at the population level.
 

### 2024-03-22 manuscript/manuscript.tex

@@
 An important contemporary debate in sociology centers on the importance
-of intrapersonal change for understanding differences in ``personal
-culture'' -- attitudes, beliefs, values, and practices (Kiley and Vaisey
-2020; Lersch 2023; Lizardo 2017; Vaisey and Lizardo 2016;
-\textbf{underwood2021?}). This question has deep roots in seemingly
+of individual change over time for understanding differences in
+``personal culture''---attitudes, beliefs, values, and practices (Kiley
+and Vaisey 2020; Lersch 2023; Lizardo 2017; Underwoood et al. 2022;
+Vaisey and Lizardo 2016). This question has deep roots in seemingly
 contradictory theoretical perspectives. For example, pragmatist theories
@@
 by low rates of persistent change during adulthood (Kiley and Vaisey
-2020; Lersch 2023). Until recently, this work has adopted a ``tournament
-of models'' approach to adjudicating different theoretical perspectives
-(Lersch 2023: p.~228). In this framework, researchers formalized
-different models of change over time and relied on model selection
-criteria---primarily the Bayesian Information Criterion (Raftery
-1995)---to classify different questions as demonstrating either ``no
-change'' or ``any change.'' The size of these piles then serves as the
-primary evidence for the truth of different arguments (Kiley and Vaisey
-2020; Vaisey and Kiley 2021).
+2020; Lersch 2023). Work leading to this conclusion adopted a
+``tournament of models'' approach to adjudicating different theoretical
+perspectives (Lersch 2023: p.~228). In this framework, researchers
+formalized different models of change over time and relied on model
+selection criteria---primarily the Bayesian Information Criterion
+(Raftery 1995)---to classify different questions as demonstrating either
+``no change'' or ``any change.'' The size of these piles then serves as
+the primary evidence for the truth of different arguments (Kiley and
+Vaisey 2020; Vaisey and Kiley 2021).
 
@@
 it is not obvious how to use these components to understand the relative
-importance of change, espeically when change is measured in non-linear
-ways.
+importance of within-person change, especially when change is measured
+in non-linear ways.
 
 In this paper, we build on the LCAM by outlining a single metric that
-combines diverse components of the LCAM to quantify the relative
-importance of intrapersonal change and interpersonal difference for
-explaining variation in a population for a variable over time. Our
-approach isolates the ``systematic'' components of variance (i.e., it
-removes random change and measurement error), to provide a single
-succinct summary of how much systematic variation can reasonably be
-attributed to people making persistent changes over time. We demonstrate
-the utility of this quantity in two ways.
+combines components of the LCAM to quantify the relative importance of
+intrapersonal change and interpersonal difference for explaining
+variation in a population for a variable over time. Our approach
+isolates the ``systematic'' components of variance (i.e., it removes
+random change and measurement error), to provide a single succinct
+summary of how much systematic variation can reasonably be attributed to
+people making persistent changes over time. This quantity can then be
+used as a quick diagnostic to compare across questions, groups, periods,
+or samples to isolate where within-person change is more or less useful
+for explaining variation. We demonstrate the utility of this quantity in
+two ways.
 
@@
 measure of culture over time (i.e., how much do people change on
-average), the variance in age-based slopes over time (i.e., do people
-tend to change in similar ways as they age), or how much personal
-culture tends to fluctuate around these age-based trends (how consistent
-are responses over time).
+average?), the variance in age-based slopes over time (i.e., do people
+tend to change in similar ways as they age?), or how much personal
+culture tends to fluctuate around these age-based trends (i.e., how
+consistent are responses over time?).
 
@@
 tend to start at very similar positions and change at a rate of .07
-standard deviations, then that change is very useful for understanding
-difference. What is important for the theoretical debate is the relative
-contribution of these two components in explaining cultural differences
-in a population. In other words, when we look at a population of adults
-over some time frame, is there enough intrapersonal change relative to
-interpersonal difference to view it as an important process for why
-people differ from each other? Only through quantifying the two
-components we can get closer to a true answer to this question.
+standard deviations for every 10 years, then that change is very useful
+for understanding difference later in life. What is important for the
+theoretical debate is the relative contribution of these two components
+in explaining cultural differences in a population. In other words, when
+we look at a population of adults over some time frame, is there enough
+intrapersonal change relative to interpersonal difference to view it as
+an important process for why people differ from each other? Only through
+quantifying the two components we can get closer to a true answer to
+this question.
 
 Yet the true utility of the proposed quantification goes beyond settling
-past debates; it allows researchers to ask new questions. For example,
-classifying survey items based on whether they show durable change
-overlooks possibly important differences among those that do show such
-change. By quantifying the relative contributions of interpersonal
+past debates; it allows researchers to identify new questions. For
+example, classifying survey items based on whether they show durable
+change overlooks possibly important differences among those that do show
+such change. By quantifying the relative contributions of interpersonal
 differences and intrapersonal change, researchers can gauge their
@@
 
-This gives us the proportion of total variance in an item attributable
-to stable interpersonal differences.
+This effectively models respondents' culture under the assumption that
+they make no systematic change over time, or that the expected value at
+any time point is the same as the expected value at any other. This
+gives us the proportion of total variance in an item attributable to
+stable interpersonal differences.
 
@@
 
-\hypertarget{expectations}{%
-\subsection{Expectations}\label{expectations}}
-
-By shifting the question and adjusting the approach, we expect
-consistent patterns across datasets. First, because previous work found
-evidence of durable change on most measures of personal culture, we
-expect that we will find some intrapersonal change in most items. But
-because these earlier studies also found many questions (between 25 and
-35 percent) lacking evidence of \emph{any} durable change, we expect
-considerable variation in the proportion of variance accounted for by
-intrapersonal change, with some questions indicating essentially none at
-all.
-
-Second, we expect the variance attributable to intrapersonal change to
-be modest compared to pre-existing interpersonal differences. This
-expectation should be uncontroversial, regardless of the theoretical
-process assumed to be most relevant for a particular question. Panel
-surveys do not encompass individuals' entire lifetimes. Thus, even if
-intrapersonal change during adulthood is the major driver of cultural
-differences, the change detected in a panel covering 20 years probably
-will not fully account for the entire variance of interpersonal
-differences.
-
-Our expectations so far largely echo what was observed in previous
-studies. However, theoretical perspectives differ in how much change
-they expect, enabling quantification to provide a new perspective on
-existing questions. ``Formative period'' and cohort theories, which
-emphasize early-life socialization as a strong influence on adult
-personal culture, imply low amounts of intrapersonal change, even when
-we account for the duration of observation. Conversely, other
-theoretical perspectives emphasizing the importance of people's
-contemporaneous social environment might expect a larger degree of
-intrapersonal change, even in a relatively short panel.
-
-Third, we expect the variance explained by each process---interpersonal
-differences and intrapersonal change---to depend on specific attributes
-of the question and panel. First, the higher the response resolution for
-a given question and the larger the sample, the greater the variance
-attributed to intrapersonal change should be, as smaller changes in
-intrapersonal change will be easier to detect. Second, life course
-theories posit that social transitions drive changes in personal
-culture. Since the probability that any person experiences such a
-transition goes up the longer we observe them, we expect panel duration
-to be positively related to the proportion of variation explained by
-intrapersonal change.
-
-\hypertarget{analytic-strategy}{%
-\section{Analytic Strategy}\label{analytic-strategy}}
+\hypertarget{example-1-cross-question-comparisons}{%
+\section{Example 1: Cross-Question
+Comparisons}\label{example-1-cross-question-comparisons}}
+
+To demonstrate the utility of this metric, we provide two examples. In
+the first, we compare the proportion of systematic variance explained by
+intrapersonal change across a broad array of questions.
 
@@
 
-\hypertarget{analysis-steps}{%
-\subsection{Analysis Steps}\label{analysis-steps}}
+\hypertarget{analysis}{%
+\subsection{Analysis}\label{analysis}}
 
@@
 measures of personal culture outlined above. We then derive \(\omega\)
-for each of these items.
-
-As a second step in our analysis, we use linear regression to model
-\(\omega\) as a function of features of the question and panel. This
-allows us to explore how these features are associated with the degree
-of intrapersonal change. These covariates include the number of response
-options (measured as 2, 3-5, and more than 5), the specific survey, the
-log of the number of participants, the number of waves observed, the
-time period (i.e., decade) the question was first asked, and the total
-duration of time the question was asked.
+for each of these items. As a second step, we use linear regression to
+model \(\omega\) as a function of features of the question and panel.
+This allows us to explore how these features are associated with the
+degree of intrapersonal change. These covariates include the number of
+response options (measured as 2, 3-5, and more than 5), the specific
+survey, the log of the number of participants, the number of waves
+observed, the time period (i.e., decade) the question was first asked,
+and the total duration of time the question was asked.
 
 \hypertarget{results}{%
-\section{Results}\label{results}}
-
-\hypertarget{variance-decomposition}{%
-\subsection{Variance Decomposition}\label{variance-decomposition}}
+\subsection{Results}\label{results}}
 
@@
 
-\hypertarget{meta-analysis}{%
-\subsection{Meta-Analysis}\label{meta-analysis}}
-
 Figure 2 plots the results from a linear regression of \(\omega\) as a
@@
 
-\hypertarget{college-and-change-in-political-culture}{%
-\subsection{College and Change in Political
-Culture}\label{college-and-change-in-political-culture}}
+\hypertarget{example-2-college-and-change-in-political-culture}{%
+\section{Example 2: College and Change in Political
+Culture}\label{example-2-college-and-change-in-political-culture}}
 
@@
 of our approach lies in its ability to extend the debate to a broader
-set of theoretical questions. To show this potential, we turn to a
-specific empirical example: the relative importance of intrapersonal
-change and interpersonal difference for explaining cultural variation by
-level of education.
+set of theoretical questions. As a second empirical application of
+\(\omega\), we explore the relative importance of intrapersonal change
+and interpersonal difference for explaining cultural variation by level
+of education.
 
@@
 There are theoretical reasons to believe that education might be
-associated with either more or less intrapersonal change. On one hand,
-because college graduates are more connected to mainstream discourse and
-elite signals, they might be more likely to make durable changes in
-response to the emergence of new information, new issues, or political
-realignments, while those without chronic information might display more
-variance around an unchanging baseline (Zaller 1992). Conversely, it
-could be that the observed stability of the college educated reflects
-the fact that they have already formed durable opinions and are
+associated with either more or less intrapersonal change over time. On
+one hand, because college graduates are more connected to mainstream
+discourse and elite signals, they might be more likely to make durable
+changes in response to the emergence of new information, new issues, or
+political realignments, while those without chronic information might
+display more variance around an unchanging baseline (Zaller 1992).
+Conversely, because people with less education appear more likely to
+engage in automatic cognition in social situations (including survey
+interviews) than people with more education (Brett and Miles 2021), they
+might be more influenced by changes in the social environment, producing
+more temporary change and more durable change. It could be that the
+observed stability of the college educated also reflects the fact that
+they have already formed durable opinions across various issues and are
 relatively closed off to new information. If this is true, college could
@@
 that college-educated peers have already received.
+
+\hypertarget{data-1}{%
+\subsection{Data}\label{data-1}}
 
@@
 
+\hypertarget{results-1}{%
+\subsection{Results}\label{results-1}}
+
 Figure 3 plots the distribution of differences in \(\omega\) values
@@
 smaller for people with a college degree, meaning intrapersonal change
-is less common among them. There are also large differences in
-\(\omega\) values across questions for both groups. For example, on the
-question of whether the government should try to solve more problems or
-leave those problems to be solved by private businesses (``government do
-more or less''), less than 1 percent of the systematic variance is
-attributable to intrapersonal change for both groups. In other words,
-while people might vacillate on this question at random (37 percent of
-variance is residual for this question), there is functionally no
-evidence that people make systematic changes of opinion on this issue
-during the GSS panel.
+is less important for explaining variation among them. There are also
+large differences in \(\omega\) values across questions for both groups.
+For example, on the question of whether the government should try to
+solve more problems or leave those problems to be solved by private
+businesses (``government do more or less''), less than 1 percent of the
+systematic variance is attributable to intrapersonal change for both
+groups. In other words, while people might vacillate on this question at
+random (37 percent of variance is residual for this question), there is
+functionally no evidence that people make systematic changes of opinion
+on this issue during the GSS panel.
 
@@
 questions, intrapersonal change plays a much larger role in accounting
-for partisan identification and political ideology. And this is
-particularly true for respondents who do not have a college degree; the
-\(\omega\) value for ideological identification among non-college
-educated respondents is almost four times that of college-educated
-respondents.
+for variation in partisan identification and political ideology. And
+this is particularly true for respondents who do not have a college
+degree; the \(\omega\) value for ideological identification among
+non-college educated respondents is almost four times that of
+college-educated respondents.
 
@@
 differences in attitudes and behavior across groups. Something about
-college attendance seems to crystallize personal culture. While these
-results should not be interpreted as causal effects of attending college
--- they are potentially confounded by age, social class, race, gender,
-and other factors that explain selection into higher education -- they
-open up a set of new questions and dynamics to explore.
+college attendance seems to affect the role played by change in
+explaining variance. It could to this in either two ways: by creating
+greater variation in baseline views or by reducing change over time
+relative to people without college degrees. While these results should
+not be interpreted as causal effects of attending college -- they are
+potentially confounded by age, social class, race, gender, and other
+factors that explain selection into higher education -- they open up a
+set of new questions and dynamics to explore.
 
@@
 their adult life. Instead of falling back on a ``tournament of models''
-approach (Lersch 2023: p.~228) to proclaim victory for a particular
-answer to this question, our goal is to refocus the debate to the
-theoretically more productive question of the relative importance of
-intrapersonal change for explaining differences in personal culture. We
-propose a new approach that quantifies the amount of observed systematic
-variance that is attributable to either interpersonal differences at
-baseline or intrapersonal change over the duration of a panel.
-
-Applying our proposed measure, \(\omega\), to 609 survey items from all
-of the panel datasets previously discussed in this debate revealed a
-consistent pattern. Nearly all questions show evidence of people making
-some durable, intrapersonal change over time. Some show notably high
-amounts of intrapersonal change. For some questions about life
-satisfaction or views on government benefits, it seems plausible that
-differences in adult experiences predominantly account for observed
-differences between people.
-
-However, intrapersonal change is often substantially less pronounced
-than interpersonal differences, accounting for less than 10 percent of
-systematic variance on average across all datasets. On numerous
-questions, like those on civil liberties, abortion, generalized trust,
-and civic duty, the systematic variance attributable to intrapersonal
-change is essentially zero. For these questions, it seems there is not
-enough cultural change during adulthood to warrant attributing the
-differences we observe to experiences and social transitions; instead,
-the primary source of the observed differences appears to stem from
-experiences in childhood, adolescence, and early adulthood.
-
-More than adjudicating past positions in this debate, the approach we
-presented in this paper will enable researchers to explore more
-precisely the relative contributions of change and stability in
-explaining variation in personal culture, as well as how these
-contributions might differ across populations. To illustrate this, we
-showed how the amount of cultural difference explained by intrapersonal
-change varies substantially by survey item but also by individuals'
-characteristics such as their education. While age and other factors
-related to college completion might confound this pattern, it suggests
-that college completion may crystallize personal culture to an extent
-that renders later adult experiences on political dispositions less
-influential.
+approach (Lersch 2023: p.~228) to proclaim a yes or no answer to this
+question, or only quantifying individual components of cultural
+difference to show how much people change, our metric contextualizes the
+amount of observed systematic variance that is attributable to
+intrapersonal change over the duration of a panel. In doing so, it
+provides a tool to identify questions, groups, or periods where change
+is a more meaningful explanation of cultural difference.
+
+We demonstrated the utility of this measure in two ways. First, applying
+our proposed measure, \(\omega\), to 609 survey items from the panel
+datasets previously studied in this debate revealed a consistent
+pattern. Nearly all questions show evidence of people making some
+durable, intrapersonal change over time. Some show notably high amounts
+of intrapersonal change. For some questions about life satisfaction or
+views on government benefits, it seems plausible that differences in
+adult experiences predominantly account for observed differences between
+people. However, intrapersonal change is often substantially less
+pronounced than interpersonal differences, accounting for less than 10
+percent of systematic variance on average across all datasets. On
+questions related to civil liberties, abortion, generalized trust, and
+civic duty, systematic variance attributable to intrapersonal change is
+essentially zero. For these questions, it seems there is not enough
+cultural change during adulthood to warrant attributing the differences
+we observe to experiences and social transitions; instead, the primary
+source of the observed differences appears to stem from experiences in
+childhood, adolescence, and early adulthood.
+
+While these findings generally echo the general consensus that has
+emerged from previous work, it adds two dimensions. First, it quantifies
+the relative importance of change across these questions, providing more
+resolution within what were mainly piles of questions either showing
+evidence of change or no evidence of change. This allows for greater
+comparison across questions and it provides a benchmark for researchers
+exploring new questions. Second, our metric can help contextualize the
+quantities identified in the LCAM. Is .07 standard deviations of change
+every 10 years a lot of change? For some questions, our metric will say
+this change explains a large component of variance, and for others it
+will not.
+
+In a second application of the quantity, we showed how it can open up
+new lines of inquiry. We showed how the amount of cultural difference
+explained by intrapersonal change varies substantially not just by
+survey item but also by individuals' characteristics such as their
+education. While age and other factors related to college completion
+might confound this pattern, it suggests that college completion may
+crystallize personal culture to an extent that renders later adult
+experiences on political dispositions less influential in explaining
+overall differences.
 
@@
 
+This is in fact one of the strengths of our approach. When change over
+time is modeled non-linearly, it can be difficult to summarize its
+relative importance. With our metric, researchers can quantify the
+additional explanatory power of modeling change using splies,
+polynomials, or other non-linear approaches relative to a linear trend.
+
 Similarly, our approach assumes that durable change is unidirectional.
@@
 Despite these limitations, we believe our method and findings have
-important implications for social science research. Sociologists
-interested in understanding cultural differences have largely asked
-about the \emph{existence} of cultural change in adults (e.g., Kiley and
-Vaisey 2020; Lersch 2023; Vaisey and Kiley 2021). In any population,
+implications for social science research. Sociologists interested in
+understanding cultural differences have largely asked about the
+\emph{existence} of cultural change in adults (e.g., Kiley and Vaisey
+2020; Lersch 2023; Vaisey and Kiley 2021). In any population,
 \emph{some} degree of adult cultural change is inevitable. Although
@@
 
-This and other recent findings (Quinn et al. 2023; Stewart and Berkman
-2023; \textbf{quinn2024?}) suggest that it is theoretically more
-productive to measure the relative importance of these two components in
-concrete, substantive cases. Drawing a unified conclusion from survey
-items measuring various cultural forms on different scales and across
+This and other recent findings (Quinn et al. 2023, 2024; Stewart and
+Berkman 2023) suggest that it is theoretically more productive to
+measure the relative importance of these two components in concrete,
+substantive cases. Drawing a unified conclusion from survey items
+measuring various cultural forms on different scales and across
 different time frames is challenging. Nevertheless, the general pattern
@@
 world.
-
-We are not claiming that adults remain static or that their changes are
-inconsequential. Even quantitatively minor shifts, such as a 2
-percentage-point change in support for gay marriage, can have massive
-ramifications. While the majority may remain consistent in their views,
-understanding the underlying mechanisms of even such numerically minor
-shifts remains a crucial task for the sociology of culture.
 
@@
 
-Our findings also have implications for understanding cultural change at
-the aggregate level. Given that intrapersonal change accounts for a
-relatively small amount of variance among adults, many forms of cultural
-change at the aggregate level are necessarily more likely driven cohort
-replacement than by contemporaneous social conditions (Underwoood et al.
-2022; Vaisey and Lizardo 2016). This likely holds true for cultural
-change at the macro societal level as well as the micro level such as
-within organizations, political parties, and professions. Our findings
-also indicate that the impact of formative experiences, as opposed to
-contemporaneous social factors, on cultural change is contingent upon an
-individuals' level of education. The weight of these two factors might
-thus change as the prominence of education shifts in the life course. If
-scholars find similar differences across other social categorizations
-like class or race, it might necessitate a more thorough integration of
-demographic processes into the analysis of cultural change at the
-population level.
-
 \hypertarget{implications-for-survey-research}{%
@@
 
+\leavevmode\vadjust pre{\hypertarget{ref-brett2021}{}}%
+Brett, Gordon, and Andrew Miles. 2021. {``Who {Thinks} {How}? {Social}
+{Patterns} in {Reliance} on {Automatic} and {Deliberate} {Cognition}.''}
+\emph{Sociological Science} 8.
+
 \leavevmode\vadjust pre{\hypertarget{ref-brocic2021}{}}%
@@
 States.''} \emph{Social Science Research} 95:102537.
+
+\leavevmode\vadjust pre{\hypertarget{ref-quinn2024}{}}%
+Quinn, Joseph M., Robert E. Freeland, E. K. Maloney, Kimberly B. Rogers,
+and Lynn Smith-Lovin. 2024. {``Meaning {Change} in {U}.{S}.
+{Occupational} {Identities} During the {COVID}-19 {Pandemic}: {Was} {It}
+{Temporary} or {Durable}?''} \emph{Social Psychology Quarterly}
+01902725241228529. doi:
+\href{https://doi.org/10.1177/01902725241228529}{10.1177/01902725241228529}.
 

### 2024-03-22 manuscript/manuscript.Rmd
data note

@@
 
+# Data Note
+
+This paper uses unit record data from Household, Income and Labour Dynamics in Australia Survey [HILDA] conducted by the Australian Government Department of Social Services (DSS).
+The findings and views reported in this paper, however, are those of the authors and should not be attributed to the Australian Government, DSS, or any of DSS’ contractors or partners.
+DOI: 10.26193/0LPD4U.
+The SOEP is administered at the German Institute for Economic Research (DIW Berlin) and funded by the Federal Ministry for Education and Research (BMBF) and the state governments under the umbrella of the Leibniz Association.
+The BHPS data were made available through the UK data archive.
+The data were originally collected by the ESRC research center on micro-social change at the University of Essex, now incorporated within the Institute for Social and Economic Research.
+The authors bear full responsibility for the analyses and interpretation of the data presented in this article.
+Understanding Society is an initiative funded by the Economic and Social Research Council and various Government Departments, with scientific leadership by the Institute for Social and Economic Research, University of Essex, and survey delivery by NatCen Social Research and Kantar Public.
+The research data are distributed by the UK Data Service.
+This study has been realized using data collected by the Swiss Household Panel (SHP), which is based at the Swiss Centre of Expertise in the Social Sciences FORS.
+The project is supported by the Swiss National Science Foundation.
+This study also uses data from The General Social Survey (GSS), which is a project of the independent research organization NORC at the University of Chicago, with principal funding from the National Science Foundation.
+The collection of data used in the PSID was partly supported by the National Institutes of Health under grant number R01 HD069609 and R01 AG040213, and the National Science Foundation under award numbers SES 1157698 and 1623684.
+
 \theendnotes

### 2024-03-22 manuscript/manuscript.tex
data note

@@
 
+\hypertarget{data-note}{%
+\section{Data Note}\label{data-note}}
+
+This paper uses unit record data from Household, Income and Labour
+Dynamics in Australia Survey {[}HILDA{]} conducted by the Australian
+Government Department of Social Services (DSS). The findings and views
+reported in this paper, however, are those of the authors and should not
+be attributed to the Australian Government, DSS, or any of DSS'
+contractors or partners. DOI: 10.26193/0LPD4U. The SOEP is administered
+at the German Institute for Economic Research (DIW Berlin) and funded by
+the Federal Ministry for Education and Research (BMBF) and the state
+governments under the umbrella of the Leibniz Association. The BHPS data
+were made available through the UK data archive. The data were
+originally collected by the ESRC research center on micro-social change
+at the University of Essex, now incorporated within the Institute for
+Social and Economic Research. The authors bear full responsibility for
+the analyses and interpretation of the data presented in this article.
+Understanding Society is an initiative funded by the Economic and Social
+Research Council and various Government Departments, with scientific
+leadership by the Institute for Social and Economic Research, University
+of Essex, and survey delivery by NatCen Social Research and Kantar
+Public. The research data are distributed by the UK Data Service. This
+study has been realized using data collected by the Swiss Household
+Panel (SHP), which is based at the Swiss Centre of Expertise in the
+Social Sciences FORS. The project is supported by the Swiss National
+Science Foundation. This study also uses data from The General Social
+Survey (GSS), which is a project of the independent research
+organization NORC at the University of Chicago, with principal funding
+from the National Science Foundation. The collection of data used in the
+PSID was partly supported by the National Institutes of Health under
+grant number R01 HD069609 and R01 AG040213, and the National Science
+Foundation under award numbers SES 1157698 and 1623684.
+
 \theendnotes

### 2024-08-22 write-ups/manuscript/manuscript.Rmd
adding code; changing repo structure.

--- 2024-03-22 manuscript/manuscript.Rmd
@@
 In this paper, we build on the LCAM by outlining a single metric that combines components of the LCAM to quantify the relative importance of intrapersonal change and interpersonal difference for explaining variation in a population for a variable over time.
-Our approach isolates the "systematic" components of variance (i.e., it removes random change and measurement error), to provide a single succinct summary of how much systematic variation can reasonably be attributed to people making persistent changes over time.
+Our approach isolates the "systematic" components of variance (i.e., it removes random change and measurement error), to provide a single summary of how much systematic variation can reasonably be attributed to people making persistent changes over time.
 This quantity can then be used as a quick diagnostic to compare across questions, groups, periods, or samples to isolate where within-person change is more or less useful for explaining variation.
@@
 
-Second, to illustrate how this quantity might be used by a researcher with a specific topical interest, we investigate differences in the relative importance of intrapersonal change for people with and without a college degree.
+Second, to illustrate how this quantity might be used by a researcher with a specific topical or theoretical interest, we investigate differences in the relative importance of intrapersonal change for people with and without a college degree.
 We find that change explains a smaller proportion of systematic variance among college-educated respondents, suggesting that college crystallizes one's personal culture rather than fostering an openness to new information.

### 2024-08-22 scripts/01_sim.R

@@ -1,0 +1,178 @@
+
+# - quantifying change ------------------------------------------------------- #
+# - simulations -------------------------------------------------------------- #
+
+### note: this script
+###       (a) generates dataframes with varying DGPs,
+###       (b) calculates omega values across different parameter combinations,
+###       (c) draws average omega values under different specifications.
+
+# ---------------------------------------------------------------------------- #
+
+rm(list = ls())
+pacman::p_load(
+  purrr,
+  furrr, 
+  hrbrthemes,
+  conflicted,
+  tidyverse)
+conflicts_prefer(dplyr::select())
+conflicts_prefer(dplyr::filter())
+
+# --- helpers
+my_oka <-
+  c("#eebd64",
+    "#8b3b36",
+    "#35455d")
+theme_set(theme_ipsum_rc())
+`%nin%` = Negate(`%in%`)
+
+# --- scripts
+suppressMessages(
+  source("./scripts/helper_scripts/build_DGP.R"))
+suppressMessages(
+  source("./scripts/helper_scripts/omega.R"))
+
+# --- parallel sessions
+plan(multisession); 
+set.seed(11235) # set seed for simulation protocol
+
+# ---------------------------------------------------------------------------- #
+# PART 1: BUILD DATAFRAMES --------------------------------------------------- #
+# ---------------------------------------------------------------------------- #
+
+# --- PARAMETER SPACE
+
+d <-
+  expand_grid(
+    # varying rates of change
+    p_rate = c(.10,
+               .25,
+               .50),
+    # varying strengths of change
+    p_strength = c(1, 
+                   2,
+                   4),
+    # varying balance of change
+    p_balance = c( 0,
+                  .5),
+    # 500 simulation runs per condition
+    sim = c(1:500)
+  )
+
+# --- DATA GENERATION
+
+d <- d |>
+  mutate(
+    data =
+      furrr::future_pmap(
+        ## mapping list
+        .l = list(p_rate, p_strength, p_balance),
+        ## refer to list based on index
+        .f = ~ generate_data(
+          n = 500,
+          reliable = 0.9,
+          t = 3,
+          rate = ..1,
+          strength = ..2,
+          balance = ..3
+        ),
+        ## for reproducibility
+        .options = furrr_options(seed = TRUE),
+        .progress = TRUE
+      )
+  )
+
[…]

### 2024-08-22 scripts/02_omega.R

@@ -1,0 +1,165 @@
+
+# - quantifying change ------------------------------------------------------- #
+# - omega calculations ------------------------------------------------------- #
+
+### note: this script
+###       (a) calculates omega values across survey items,
+###       (b) draws the relative shares of intrapersonal/interpersonal culture,
+###       (c) exports V(D), V(C) and Omega as a LaTeX table.
+
+# ---------------------------------------------------------------------------- #
+
+rm(list = ls())
+pacman::p_load(
+  purrr,
+  furrr, 
+  hrbrthemes,
+  conflicted,
+  tidyverse)
+conflicts_prefer(dplyr::select())
+conflicts_prefer(dplyr::filter())
+
+# --- helpers
+my_oka <-
+  c("#eebd64",
+    "#8b3b36",
+    "#35455d",
+    "#144525")
+theme_set(theme_ipsum_rc())
+`%nin%` = Negate(`%in%`)
+
+# --- scripts
+suppressMessages(
+  source("./scripts/helper_scripts/omega.R"))
+
+# --- data
+d <- readRDS("./data/data.rds") # nested survey x item panels
+
+# --- parallel sessions
+plan(multisession); 
+set.seed(11235) # set seed for simulation protocol
+options(future.globals.maxSize=5000*1024^2)
+
+# ---------------------------------------------------------------------------- #
+# PART 1: CALCULATE OMEGAS --------------------------------------------------- #
+# ---------------------------------------------------------------------------- #
+
+d <- d |>
+  mutate(omega = future_map(
+    .x = data,
+    .f = purrr::possibly(
+      ~ calculate_omega(
+        df = .,
+        yname = y,
+        tname = wave,
+        pname = pid,
+        boot = FALSE
+      )
+    ),
+    ## for reproducibility
+    .options = furrr_options(seed = TRUE),
+    .progress = TRUE
+  ))
+
+# ---------------------------------------------------------------------------- #
+# PART 2: PRETTIFY AND EXPORT ------------------------------------------------ #
+# ---------------------------------------------------------------------------- #
+
+d <- d |>
+  ## drop data
+  dplyr::select(-data) |>
+  ## unnest
+  unnest_wider(omega) |>
+  ## add variable labels
+  left_join(read_csv("./misc/var_labels.csv"),
+            by = c("surveys", "variable"))
+
+saveRDS(d, "./data/output/omegas.rds")
+
[…]

### 2024-08-22 scripts/03_educ.R

@@ -1,0 +1,190 @@
+
+# - quantifying change ------------------------------------------------------- #
+# - college example ---------------------------------------------------------- #
+
+### note: this script
+###       (a) prepares 8 political items from the General Social Survey,
+###       (b) calculates omega values across college and no-college subgroups,
+###       (c) compares the differences of omega values in booted samples.
+
+# ---------------------------------------------------------------------------- #
+
+rm(list = ls())
+pacman::p_load(
+  purrr,
+  furrr, 
+  hrbrthemes,
+  conflicted,
+  gssr,
+  tidyverse)
+conflicts_prefer(dplyr::select())
+conflicts_prefer(dplyr::filter())
+
+# --- helpers
+my_oka <-
+  c("#eebd64",
+    "#8b3b36",
+    "#35455d",
+    "#144525")
+theme_set(theme_ipsum_rc())
+`%nin%` = Negate(`%in%`)
+
+# --- scripts
+suppressMessages(
+  source("./scripts/helper_scripts/omega.R"))
+
+# --- data
+data("gss_panel06_long")
+data("gss_panel08_long")
+data("gss_panel10_long")
+
+# --- parallel sessions
+plan(multisession); 
+set.seed(11235) # set seed for simulation protocol
+options(future.globals.maxSize=5000*1024^2)
+
+# ---------------------------------------------------------------------------- #
+# PART 1: DATA PREPS --------------------------------------------------------- #
+# ---------------------------------------------------------------------------- #
+
+# --- get the data
+gss <- bind_rows(
+  gss_panel06_long |>
+    select(
+      firstid,
+      wave,
+      age,
+      degree,
+      partyid,
+      polviews,
+      eqwlth,
+      helpsick,
+      helpnot,
+      helpblk,
+      helppoor,
+      affrmact
+    ) |>
+    mutate(pid = paste("06", firstid, sep = "-")),
+  gss_panel08_long |>
+    select(
+      firstid,
+      wave,
+      age,
+      degree,
+      partyid,
+      polviews,
+      eqwlth,
+      helpsick,
+      helpnot,
+      helpblk,
+      helppoor,
+      affrmact
+    ) |>
+    mutate(pid = paste("08", firstid, sep = "-")),
+  gss_panel10_long |>
+    select(
+      firstid,
+      wave,
+      age,
+      degree,
+      partyid,
+      polviews,
+      eqwlth,
+      helpsick,
+      helpnot,
[…]

### 2024-08-22 scripts/helper_scripts/build_DGP.R

@@ -1,0 +1,117 @@
+
+# - quantifying change ------------------------------------------------------- #
+# - DGPs --------------------------------------------------------------------- #
+
+### note: this script
+###       (a) simulates panel data values using a set of parameters,
+###       (b) exports a dataframe with individual ID, time, and observed `y`.
+
+# ---------------------------------------------------------------------------- #
+
+# --- packages when the script is loaded
+pacman::p_load(tidyverse)
+
+# --- function call
+generate_data <-
+  
+  function(
+    
+    ## the number of N
+    n = 1000,
+    ## the number of T
+    t = 3,
+    ## the rate of change
+    rate = 0.25,
+    ## strength of change
+    strength = 1,
+    ## change balance
+    balance = 0.5,
+    ## reliability score
+    reliable = 0.8) {
+    
+    # ----------------------------------------------------- #
+    # BUILDING BLOCKS                                       #
+    # ----------------------------------------------------- #
+    
+    # --- part 1: generate actors
+    if (rate == 0) {
+      u <- 
+        tibble(
+          ## actors
+          pid = c(1:n),
+          ## central tendency
+          u = rnorm(n = n, mean = 0, sd = 1)
+        )
+    } else {
+      u <-
+        tibble(
+          ## actors
+          pid = c(1:n),
+          ## central tendency
+          u = rnorm(n = n, mean = 0, sd = 1)
+        ) |> 
+        mutate(
+          ## changer status
+          changer = sample(0:1, 
+                           n(), 
+                           replace = TRUE, 
+                           prob = c(1 - rate, rate))
+        )
+    }
+    
+    # --- part 2: generate the longitudinal data setup
+    
+    ## 2.1: build the window
+    span_window <- seq(
+      ## first t
+      from = 0,
+      ## final t                 
+      to =   1,
+      ## the number of ts
+      length.out = t)
+    
+    ## 2.2: generate pid x time grid
+    data <- expand_grid(u, t = span_window)
+    
+    # ----------------------------------------------------- #
+    # RESPONSE CONSTRUCTION                                 #
+    # ----------------------------------------------------- #
+    
+    # --- part 3: generate the observed scores
+    
+    ## 3.1: add change scores
+    data <- data |> 
+      left_join(u |> select(pid) |>
+                  mutate(upper = 
+                           rbinom(n = n, size = 1, prob = balance)) |>
+                  mutate(upper = ifelse(upper == 1, 1, -1)),
[…]

### 2024-08-22 scripts/helper_scripts/omega.R

@@ -1,0 +1,184 @@
+
+# - quantifying change ------------------------------------------------------- #
+# - omega -------------------------------------------------------------------- #
+
+### note: this script
+###       (a) processes the input data for omega calculation,
+###       (b) calculates within variance, between variance and omega from MLMs,
+###       (c) generates bootstrapped distributions for omega values.
+
+# ---------------------------------------------------------------------------- #
+
+# --- packages when the script is loaded
+pacman::p_load(glmmTMB,
+               lme4,
+               performance,
+               Matrix,
+               tidyverse)
+
+# --- function call
+calculate_omega <- 
+  
+  function(
+    
+    # dataframe
+    df,
+    # outcome variable
+    yname,
+    # time variable
+    tname,
+    # person identifier
+    pname,
+    # bootstrap
+    boot = TRUE,
+    # number of boots
+    nrep = 500) {
+    
+    # ----------------------------------------------------- #
+    # DATA PROCESSING                                       #
+    # ----------------------------------------------------- #
+    
+    df <- df |>
+      mutate(
+        # get person identifier
+        pid = {
+          {
+            pname
+          }
+        },
+        # get time identifier
+        time = {
+          {
+            tname
+          }
+        },
+        # get outcome identifier
+        y  = {
+          {
+            yname
+          }
+        }) |>
+      
+      # unique numeric identifier
+      mutate(id = cur_group_id(), .by = "pid") |>
+      
+      # drop missing values
+      select(id, time, y) |> drop_na() |>
+      
+      # drop respondents with only one observation
+      mutate(n = n(), .by = "id") |> dplyr::filter(n != 1) |> 
+      
+      # center time at respondent midpoint
+      mutate(
+        time = (time - (max(time) + min(time)) / 2),
+        .by = "id")
+    
+    # ----------------------------------------------------- #
+    # MODELING                                              #
+    # ----------------------------------------------------- #
+    
+    ## estimate LCAM
+    model <- glmmTMB(y ~ time + (1 + time | id), data = df)
+    
+    ## prior specification
+    prior_sd <- round(sd(df$y) * 10, 5) ## 10 times the observed slopes
+    prior_sd <- paste("normal(0, ", prior_sd, ")", sep = "")
+    
+    ## impose priors
[…]

### 2024-08-22 write-ups/manuscript/manuscript.Rmd
changes until examples.

@@
 ---
-title: "Quantifying the Importance of Change for Understanding Differences in Personal Culture"
+title: |
+  | Quantifying the Importance of Change for Understanding 
+  | Differences in Personal Culture
 author: 
-  - Achim Edelmann^[médialab, Science Po]
-  - Kevin Kiley^[Department of Sociology and Anthropology, North Carolina State University]
+  - Achim Edelmann^[médialab, Science Po; equal authorship.]
+  - Kevin Kiley^[Department of Sociology and Anthropology, North Carolina State University; equal authorship.]
   - Turgut Keskintürk^[Department of Sociology, Duke University]
@@
   pdf_document:
-    latex_engine: xelatex
     keep_tex: true
 geometry: margin=1in 
-fontsize: 12pt
+fontsize: 11pt
+fontfamily: mathpazo
 header-includes:
-  \usepackage{fontspec}
-  \setmainfont{Times New Roman}
-  \setlength{\parindent}{3em}
-  \usepackage{caption}
-  \usepackage{setspace}
-  \usepackage{endnotes}
-  \renewcommand{\enotesize}{\normalsize}
-  \let\footnote=\endnote
+- \usepackage{array}
+- \usepackage{longtable}
+- \usepackage{booktabs}
+- \usepackage{float}
+- \usepackage{caption}
+- \usepackage{amsmath}
+- \captionsetup{justification=raggedright,
+                singlelinecheck=false}
 bibliography: kileybib.bib
@@
 To advance this debate, we introduce a measure for quantifying the proportion of systematic variance in panel data attributable to intrapersonal change.
-Applying this measure to 609 items from seven surveys in five countries, we find that although intrapersonal change is common, it does not explain a large share of systematic variance.
+Applying this measure to 610 items from seven surveys in five countries, we find that although intrapersonal change is common, it does not explain a large share of systematic variance.
 As an extension of the theoretical debate, we demonstrate that this measure provides new insights when comparing social groups, showing that intrapersonal change is less common among U.S.
@@
 ```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = F, message = F, warning = F)
+knitr::opts_chunk$set(echo = F, 
+                      message = F, 
+                      warning = F)
+knitr::opts_chunk$set(fig.align = "center")
 ```
-
-\doublespacing
 
@@
 228].
-In this framework, researchers formalized different models of change over time and relied on model selection criteria---primarily the Bayesian Information Criterion [@raftery1995]---to classify different questions as demonstrating either "no change" or "any change." The size of these piles then serves as the primary evidence for the truth of different arguments [@kiley2020; @vaisey2021].
+In this framework, researchers formalized different models of change over time and relied on model selection criteria---primarily the Bayesian Information Criterion [@raftery1995]---to classify different questions as demonstrating either "no change" or "any change." The size of these piles then serves as the primary evidence for the truth of different arguments [@kiley2020; @vaisey2021; @lersch2023].
 
@@
 
-First, we use 609 items from seven panel surveys from five countries to quantify the proportion of variance that is attributable to systematic intrapersonal change over the study period rather than to interpersonal differences at baseline.
+First, we use 610 items from seven panel surveys from five countries to quantify the proportion of variance that is attributable to systematic intrapersonal change over the study period rather than to interpersonal differences at baseline.
 This allows us to generate an overall picture of how much of the total systematic variation in personal culture change tends to explain.
@@
 Drawing on these findings, Lersch [-@lersch2023] challenged the SDM and AUM as a "needless dichotomy," proposing the "Life Course Adaption Model" (LCAM) as an alternative.
-This model draws on the life course perspective to model personal culture as a different linear trajectory[^linear] over the duration of a panel for each respondent.
+This model draws on the life course perspective to model personal culture as a linear trajectory[^linear] over the duration of a panel for each respondent.
 In doing this, Lersch rectified three shortcomings of the AUM and SDM.
@@
 
-## Bootstrapping for Confidence Intervals
+## Bootstrapping for Quantifying Sampling Variability
 
-[Explain how we do that here]
+Of course, the $\omega$ measure is itself variable---most particularly due to sampling variability---such that the composition of durable changers and stable respondents in a particular sample might affect the systematic components $V(D)$ and $V(C)$.
+It is thus important to establish confidence intervals to reflect the uncertainty in the $\omega$.
+We achieve this by incorporating nonparametric bootstraps with case resampling.
+In doing this, we first resample the respondents with replacement and assign a new identifier to the set of new *respondents*, allowing us to change the composition of level-2 units in our data.
+Second, we estimate the LCAM in this slightly perturbed sample, and calculate the $\omega$.
+We then replicate this procedure $N$ times and establish confidence intervals using the distribution of bootstrapped omega values.
 
@@
 
-```{=tex}
-\begin{center}
-<Table 1 about here>
-\end{center}
+```{r, message = FALSE}
+library(tidyverse)
+library(knitr)
+library(kableExtra)
+table <- tibble(
+  Country = c("Australia", "Germany", "Great Britian", "Great Britian", "Switzerland", "United States", "United States"),
+  Survey = c("Household, Income and Labor Dynamics in Australia (HILDA)",
+             "Socio-Economic Panel Study (SOEP)",
+             "British Household Panel Survey (BHPS)",
+             "Understanding Society/UK Household Longitudinal Study (UKHLS)",
+             "Swiss Household Panel (SHP)",
+             "General Social Survey (GSS)",
+             "Panel Study of Income Dynamics (PSID)"),
+  Period = c("2001-2021", "1984-2020", "1991-2008", "2009-2020", "1999-2019", "2006-2012", "1968-2019"),
+  Outcomes = c(30, 122, 83, 66, 77, 183, 49)
+)
+kableExtra::kable(table,
+                  booktabs = TRUE,
+                  format = "latex",
+                  align = "llcc",
+                  caption = "The Description of the Data Sources") |> 
+  kable_styling(latex_options = c("scale_down", "HOLD_position"))
+
 ```
@@
 
-```{=tex}
+```{=latex}
+\begin{figure}[ht]
 \begin{center}
-<Figure 1 about here>
+\caption{Proportions of Systematic Variance in Personal Culture}
+```
+```{r fig1, out.width = "100%", fig.pos = "H"}
+knitr::include_graphics("../../figures/omegas.png")
+```
+```{=latex}
 \end{center}
+\footnotesize{\textit{Notes:} The figure shows $\omega$ and 1 - $\omega$ as the proportion of systematic variance attributable to intrapersonal change and interpersonal differences.
+See Supplemental Materials A for the full set of item values.}
+\end{figure}
 ```
@@
 
-```{=tex}
+```{=latex}
+\begin{figure}[ht]
 \begin{center}
-<Figure 3 about here>
+\caption{Difference in $\omega$ Across College Graduates and Non-Graduates}
+```
+```{r fig3}
+knitr::include_graphics("../../figures/colleges.png")
+```
+```{=latex}
 \end{center}
+\footnotesize{\textit{Notes:} The figure shows the difference in $\omega$ values across college graduates and college non-graduates.
+Values above (below) 0 means that those with college degree have higher (lower) variance of intrapersonal change.
+The dashed red line marks 0 difference.}
+\end{figure}
 ```
@@
 
-\theendnotes
-
-\singlespace
-
 # References
@@
 :::
-
-\newpage
-
-# Appendix
-
-## Appendix A: The Distribution of Variation Components
-
-Figure A1 shows the distribution of $V(D)$ and $V(C)$ across all panels.
-
-```{=tex}
-\begin{center}
-<Figure A1 about here>
-\end{center}
-```
-
-## Appendix B: $\omega$ Values When the Last Wave Removed
-
-Figure A2 presents the alternative strategy we used to understand the effects of duration on intraindividual change.
-First, we removed the last wave from all the observations and fitted the Life Course Adaption models.
-This effectively reduced the number of items to 249.
-We then refitted the model for all participants in this sample using the unrestricted data.
-In the final step, we calculated the $\omega$ values for each item.
-The Figure A2 shows a scatterplot for these two sets of observations.
-
-```{=tex}
-\begin{center}
-<Figure B1 about here>
-\end{center}
-```
-
-\newpage
-
-# Tables and Figures
-
-```{r, message = FALSE}
-
-library(tidyverse)
-library(knitr)
-library(kableExtra)
-
-table <- tibble(
-  Country = c("Australia", "Germany", "Great Britian", "Great Britian", "Switzerland", "United States", "United States"),
-  Survey = c("Household, Income and Labor Dynamics in Australia (HILDA)",
-             "Socio-Economic Panel Study (SOEP)",
-             "British Household Panel Survey (BHPS)",
-             "Understanding Society/UK Household Longitudinal Study (UKHLS)",
-             "Swiss Household Panel (SHP)",
-             "General Social Survey (GSS)",
-             "Panel Study of Income Dynamics (PSID)"),
-  Period = c("2001-2021", "1984-2020", "1991-2008", "2009-2020", "1999-2019", "2006-2012", "1968-2019"),
-  Outcomes = c(30, 122, 83, 66, 77, 183, 48)
-)
-
-kableExtra::kable(table,
-                  booktabs = TRUE,
-                  format = "latex",
-                  align = "llcc",
-                  caption = "The Description of the Data Sources") |> 
-  kable_styling(latex_options = c("scale_down", "HOLD_position"))
-
-```
-
-\newpage
-
-```{=latex}
-\begin{figure}[ht]
-\begin{center}
-\caption{Proportions of Systematic Variance in Personal Culture}
-```
-```{r fig1, out.width = "100%", fig.pos = "H"}
-knitr::include_graphics("../figures/figure_1.png")
-```
-
-```{=latex}
-\end{center}
-\footnotesize{\textit{Notes:} The figure shows $\omega$ and 1 - $\omega$ as the proportion of systematic variance attributable to intrapersonal change and interpersonal differences.
-See Supplemental Materials A for the full set of item values.}
-\end{figure}
-```
-
-\newpage
-
-```{=latex}
-\begin{figure}[ht]
-\begin{center}
-\caption{Regression Model Estimating $\omega$}
-```
-```{r fig2}
-knitr::include_graphics("../figures/figure_2.png")
-```
-
-```{=latex}
-\end{center}
-\footnotesize{\textit{Notes:} The model estimates the proportion of systematic variance attributable to intrapersonal change.
-It includes the log of the number of participants, the number of waves the variable was asked (capped at t = 10 for each item), the date of the first wave (measured as the year minus 1968 divided by 10), number of waves the question was asked averaged across participants, duration in years per item averaged across participants, and panel and topic fixed effects.
-Coefficients are estimated from suppressed intercept model based on predictions at participants' wave mid-point.
-Survey indicators and item topics not shown.
-See Supplemental Materials B for the full set of coefficients.}
-\end{figure}
-```
-
-\newpage
-
-```{=latex}
-\begin{figure}[ht]
-\begin{center}
-\caption{Difference in $\omega$ Across College Graduates and Non-Graduates}
-```
-```{r fig3}
-knitr::include_graphics("../figures/figure_3.png")
-```
-
-```{=latex}
-\end{center}
-\footnotesize{\textit{Notes:} The figure shows the difference in $\omega$ values across college graduates and college non-graduates.
-Values above (below) 0 means that those with college degree have higher (lower) variance of intrapersonal change.
-The dashed red line marks 0 difference.}
-\end{figure}
-```
-
-\newpage
-
-```{=latex}
-\begin{figure}[ht]
-\begin{center}
-\caption{Difference in $\omega$ Across College Graduates and Non-Graduates on Political Culture}
-```
-```{r fig4}
-knitr::include_graphics("../figures/figure_4.png")
-```
-
-```{=latex}
-\end{center}
-\footnotesize{\textit{Notes:} $\omega$ values across college graduates and college non-graduates on 8 political culture items from the General Social Survey (2006-2014).}
-\end{figure}
-```
-
-\newpage
-
-```{=latex}
-\begin{figure}[htp]
-\begin{center}
-\caption*{Figure A1: The Distribution of V(D) and V(C)}
-```
-```{r figA1}
-knitr::include_graphics("../figures/figure_A1.png")
-```
-```{=latex}
-\end{center}
-\end{figure}
-```
-
-\newpage
-
-```{=latex}
-\begin{figure}[htp]
-\begin{center}
-\caption*{Figure A2: Difference in $\omega$ Across Two Sample Specifications} 
-```
-```{r figA2}
-knitr::include_graphics("../figures/figure_A2.png")
-```
-```{=latex}
-\end{center}
-\end{figure}
-```

### 2024-08-22 write-ups/manuscript/manuscript.tex
changes until examples.

--- 2024-03-22 manuscript/manuscript.tex
@@
 \documentclass[
-  12pt,
+  11pt,
 ]{article}
@@
 \fi
-\usepackage{lmodern}
+\usepackage[]{mathpazo}
 \ifPDFTeX\else
@@
 \setcounter{secnumdepth}{-\maxdimen} % remove section numbering
+% definitions for citeproc citations
+\NewDocumentCommand\citeproctext{}{}
+\NewDocumentCommand\citeproc{mm}{%
+  \begingroup\def\citeproctext{#2}\cite{#1}\endgroup}
+\makeatletter
+ % allow citations to break across lines
+ \let\@cite@ofmt\@firstofone
+ % avoid brackets around text for \cite:
+ \def\@biblabel#1{}
+ \def\@cite#1#2{{#1\if@tempswa , #2\fi}}
+\makeatother
 \newlength{\cslhangindent}
@@
 \setlength{\csllabelwidth}{3em}
-\newlength{\cslentryspacingunit} % times entry-spacing
-\setlength{\cslentryspacingunit}{\parskip}
-\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
- {% don't indent paragraphs
-  \setlength{\parindent}{0pt}
+\newenvironment{CSLReferences}[2] % #1 hanging-indent, #2 entry-spacing
+ {\begin{list}{}{%
+  \setlength{\itemindent}{0pt}
+  \setlength{\leftmargin}{0pt}
+  \setlength{\parsep}{0pt}
   % turn on hanging indent if param 1 is 1
   \ifodd #1
-  \let\oldpar\par
-  \def\par{\hangindent=\cslhangindent\oldpar}
+   \setlength{\leftmargin}{\cslhangindent}
+   \setlength{\itemindent}{-1\cslhangindent}
   \fi
   % set entry spacing
-  \setlength{\parskip}{#2\cslentryspacingunit}
- }%
- {}
+  \setlength{\itemsep}{#2\baselineskip}}}
+ {\end{list}}
 \usepackage{calc}
-\newcommand{\CSLBlock}[1]{#1\hfill\break}
-\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
-\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}
+\newcommand{\CSLBlock}[1]{\hfill\break\parbox[t]{\linewidth}{\strut\ignorespaces#1\strut}}
+\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{\strut#1\strut}}
+\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{\strut#1\strut}}
 \newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}
-\usepackage{fontspec} \setmainfont{Times New Roman} \setlength{\parindent}{3em} \usepackage{caption} \usepackage{setspace} \usepackage{endnotes} \renewcommand{\enotesize}{\normalsize} \let\footnote=\endnote
+\usepackage{array}
+\usepackage{longtable}
+\usepackage{booktabs}
+\usepackage{float}
+\usepackage{caption}
+\usepackage{amsmath}
+\captionsetup{justification=raggedright, singlelinecheck=false}
 \usepackage{booktabs}
@@
 \fi
-\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
+\usepackage{bookmark}
 \IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
@@
 \hypersetup{
-  pdftitle={Quantifying the Importance of Change for Understanding Differences in Personal Culture},
   pdfauthor={Achim Edelmann; Kevin Kiley; Turgut Keskintürk; Isabella Bouklas; Stephen Vaisey},
@@
 
-\title{Quantifying the Importance of Change for Understanding
+\title{Quantifying the Importance of Change for Understanding\\
 Differences in Personal Culture}
-\author{Achim Edelmann\footnote{médialab, Science Po} \and Kevin
+\author{Achim Edelmann\footnote{médialab, Science Po; equal authorship.} \and Kevin
 Kiley\footnote{Department of Sociology and Anthropology, North Carolina
-  State University} \and Turgut Keskintürk\footnote{Department of
-  Sociology, Duke University} \and Isabella
+  State University; equal authorship.} \and Turgut
+Keskintürk\footnote{Department of Sociology, Duke University} \and Isabella
 Bouklas\footnote{Department of Sociology, Duke University} \and Stephen
@@
 proportion of systematic variance in panel data attributable to
-intrapersonal change. Applying this measure to 609 items from seven
+intrapersonal change. Applying this measure to 610 items from seven
 surveys in five countries, we find that although intrapersonal change is
@@
 
-\doublespacing
-
-\hypertarget{introduction}{%
-\section{Introduction}\label{introduction}}
+\section{Introduction}\label{introduction}
 
@@
 the primary evidence for the truth of different arguments (Kiley and
-Vaisey 2020; Vaisey and Kiley 2021).
+Vaisey 2020; Lersch 2023; Vaisey and Kiley 2021).
 
@@
 isolates the ``systematic'' components of variance (i.e., it removes
-random change and measurement error), to provide a single succinct
-summary of how much systematic variation can reasonably be attributed to
-people making persistent changes over time. This quantity can then be
-used as a quick diagnostic to compare across questions, groups, periods,
-or samples to isolate where within-person change is more or less useful
-for explaining variation. We demonstrate the utility of this quantity in
-two ways.
-
-First, we use 609 items from seven panel surveys from five countries to
+random change and measurement error), to provide a single summary of how
+much systematic variation can reasonably be attributed to people making
+persistent changes over time. This quantity can then be used as a quick
+diagnostic to compare across questions, groups, periods, or samples to
+isolate where within-person change is more or less useful for explaining
+variation. We demonstrate the utility of this quantity in two ways.
+
+First, we use 610 items from seven panel surveys from five countries to
 quantify the proportion of variance that is attributable to systematic
@@
 Second, to illustrate how this quantity might be used by a researcher
-with a specific topical interest, we investigate differences in the
-relative importance of intrapersonal change for people with and without
-a college degree. We find that change explains a smaller proportion of
-systematic variance among college-educated respondents, suggesting that
-college crystallizes one's personal culture rather than fostering an
-openness to new information. This offers a new empirical basis to
-theorize about the role of education in influencing personal culture.
-
-\hypertarget{background}{%
-\section{Background}\label{background}}
-
-\hypertarget{stability-and-change-in-personal-culture}{%
+with a specific topical or theoretical interest, we investigate
+differences in the relative importance of intrapersonal change for
+people with and without a college degree. We find that change explains a
+smaller proportion of systematic variance among college-educated
+respondents, suggesting that college crystallizes one's personal culture
+rather than fostering an openness to new information. This offers a new
+empirical basis to theorize about the role of education in influencing
+personal culture.
+
+\section{Background}\label{background}
+
 \subsection{Stability and Change in Personal
-Culture}\label{stability-and-change-in-personal-culture}}
+Culture}\label{stability-and-change-in-personal-culture}
 
@@
 (LCAM) as an alternative. This model draws on the life course
-perspective to model personal culture as a different linear
-trajectory\footnote{In supplemental materials, Lersch models age using
-  splines as well, with results generally showing the same conclusions.}
-over the duration of a panel for each respondent. In doing this, Lersch
-rectified three shortcomings of the AUM and SDM. First, the AUM, as
-described by Kiley and Vaisey, posits that changes follow a Markov
-process where responses at time \(t\) are a function of responses at
-time \(t-1\) but not earlier time points. However, earlier life
-experiences can (directly and indirectly) mold personal culture when
-transitioning to new social roles or into new environments, even if
-their initial impact is delayed. For example, childhood events might
-influence views on family structures later when individuals form their
-own families. Therefore, the LCAM considers influences from earlier than
-just \(t-1\) on responses at time \(t\).
+perspective to model personal culture as a linear trajectory\footnote{In
+  supplemental materials, Lersch models age using splines as well, with
+  results generally showing the same conclusions.} over the duration of
+a panel for each respondent. In doing this, Lersch rectified three
+shortcomings of the AUM and SDM. First, the AUM, as described by Kiley
+and Vaisey, posits that changes follow a Markov process where responses
+at time \(t\) are a function of responses at time \(t-1\) but not
+earlier time points. However, earlier life experiences can (directly and
+indirectly) mold personal culture when transitioning to new social roles
+or into new environments, even if their initial impact is delayed. For
+example, childhood events might influence views on family structures
+later when individuals form their own families. Therefore, the LCAM
+considers influences from earlier than just \(t-1\) on responses at time
+\(t\).
 
@@
 
-\hypertarget{quantification-not-adjudication}{%
 \subsection{Quantification, Not
-Adjudication}\label{quantification-not-adjudication}}
+Adjudication}\label{quantification-not-adjudication}
 
@@
 
-\hypertarget{formalization}{%
-\subsection{Formalization}\label{formalization}}
+\subsection{Formalization}\label{formalization}
 
@@
 
-\hypertarget{bootstrapping-for-confidence-intervals}{%
-\subsection{Bootstrapping for Confidence
-Intervals}\label{bootstrapping-for-confidence-intervals}}
-
-{[}Explain how we do that here{]}
-
-\hypertarget{example-1-cross-question-comparisons}{%
+\subsection{Bootstrapping for Quantifying Sampling
+Variability}\label{bootstrapping-for-quantifying-sampling-variability}
+
+Of course, the \(\omega\) measure is itself variable---most particularly
+due to sampling variability---such that the composition of durable
+changers and stable respondents in a particular sample might affect the
+systematic components \(V(D)\) and \(V(C)\). It is thus important to
+establish confidence intervals to reflect the uncertainty in the
+\(\omega\). We achieve this by incorporating nonparametric bootstraps
+with case resampling. In doing this, we first resample the respondents
+with replacement and assign a new identifier to the set of new
+\emph{respondents}, allowing us to change the composition of level-2
+units in our data. Second, we estimate the LCAM in this slightly
+perturbed sample, and calculate the \(\omega\). We then replicate this
+procedure \(N\) times and establish confidence intervals using the
+distribution of bootstrapped omega values.
+
 \section{Example 1: Cross-Question
-Comparisons}\label{example-1-cross-question-comparisons}}
+Comparisons}\label{example-1-cross-question-comparisons}
 
@@
 
-\hypertarget{data}{%
-\subsection{Data}\label{data}}
+\subsection{Data}\label{data}
 
@@
 
-\begin{center}
-<Table 1 about here>
-\end{center}
-
-\hypertarget{analysis}{%
-\subsection{Analysis}\label{analysis}}
+\begin{table}[H]
+\centering
+\caption{\label{tab:unnamed-chunk-1}The Description of the Data Sources}
+\centering
+\resizebox{\ifdim\width>\linewidth\linewidth\else\width\fi}{!}{
+\begin{tabular}[t]{llcc}
+\toprule
+Country & Survey & Period & Outcomes\\
+\midrule
+Australia & Household, Income and Labor Dynamics in Australia (HILDA) & 2001-2021 & 30\\
+Germany & Socio-Economic Panel Study (SOEP) & 1984-2020 & 122\\
+Great Britian & British Household Panel Survey (BHPS) & 1991-2008 & 83\\
+Great Britian & Understanding Society/UK Household Longitudinal Study (UKHLS) & 2009-2020 & 66\\
+Switzerland & Swiss Household Panel (SHP) & 1999-2019 & 77\\
+\addlinespace
+United States & General Social Survey (GSS) & 2006-2012 & 183\\
+United States & Panel Study of Income Dynamics (PSID) & 1968-2019 & 49\\
+\bottomrule
+\end{tabular}}
+\end{table}
+
+\subsection{Analysis}\label{analysis}
 
@@
 
-\hypertarget{results}{%
-\subsection{Results}\label{results}}
+\subsection{Results}\label{results}
 
@@
 
+\begin{figure}[ht]
 \begin{center}
-<Figure 1 about here>
+\caption{Proportions of Systematic Variance in Personal Culture}
+
+\begin{center}\includegraphics[width=1\linewidth]{../../figures/omegas} \end{center}
+
 \end{center}
+\footnotesize{\textit{Notes:} The figure shows $\omega$ and 1 - $\omega$ as the proportion of systematic variance attributable to intrapersonal change and interpersonal differences.
+See Supplemental Materials A for the full set of item values.}
+\end{figure}
 
@@
 
-\hypertarget{example-2-college-and-change-in-political-culture}{%
 \section{Example 2: College and Change in Political
-Culture}\label{example-2-college-and-change-in-political-culture}}
+Culture}\label{example-2-college-and-change-in-political-culture}
 
@@
 
-\hypertarget{data-1}{%
-\subsection{Data}\label{data-1}}
+\subsection{Data}\label{data-1}
 
@@
 
-\hypertarget{results-1}{%
-\subsection{Results}\label{results-1}}
+\subsection{Results}\label{results-1}
 
@@
 
+\begin{figure}[ht]
 \begin{center}
-<Figure 3 about here>
+\caption{Difference in $\omega$ Across College Graduates and Non-Graduates}
+
+\begin{center}\includegraphics[width=62.5in]{../../figures/colleges} \end{center}
+
 \end{center}
+\footnotesize{\textit{Notes:} The figure shows the difference in $\omega$ values across college graduates and college non-graduates.
+Values above (below) 0 means that those with college degree have higher (lower) variance of intrapersonal change.
+The dashed red line marks 0 difference.}
+\end{figure}
 
@@
 
-\hypertarget{discussion-and-conclusion}{%
-\section{Discussion and Conclusion}\label{discussion-and-conclusion}}
+\section{Discussion and Conclusion}\label{discussion-and-conclusion}
 
@@
 
-\hypertarget{limitations}{%
-\subsection{Limitations}\label{limitations}}
+\subsection{Limitations}\label{limitations}
 
@@
 
-\hypertarget{implications-for-cultural-sociology}{%
 \subsection{Implications for Cultural
-Sociology}\label{implications-for-cultural-sociology}}
+Sociology}\label{implications-for-cultural-sociology}
 
@@
 
-\hypertarget{implications-for-survey-research}{%
 \subsection{Implications for Survey
-Research}\label{implications-for-survey-research}}
+Research}\label{implications-for-survey-research}
 
@@
 
-\hypertarget{conclusion}{%
-\subsection{Conclusion}\label{conclusion}}
+\subsection{Conclusion}\label{conclusion}
 
@@
 
-\hypertarget{data-note}{%
-\section{Data Note}\label{data-note}}
+\section{Data Note}\label{data-note}
 
@@
 
-\theendnotes
-
-\singlespace
-
-\hypertarget{references}{%
-\section{References}\label{references}}
-
-\hypertarget{refs}{}
-\begin{CSLReferences}{1}{0}
-\leavevmode\vadjust pre{\hypertarget{ref-alvarez2002}{}}%
+\section{References}\label{references}
+
+\phantomsection\label{refs}
+\begin{CSLReferences}{1}{1}
+\bibitem[\citeproctext]{ref-alvarez2002}
 Alvarez, R. Michael, and John Brehm. 2002. \emph{Hard Choices, Easy
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-alwin2007}{}}%
+\bibitem[\citeproctext]{ref-alwin2007}
 Alwin, Duane F. 2007. \emph{Margins of {Error}: {A} {Study} of
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-alwin1991}{}}%
+\bibitem[\citeproctext]{ref-alwin1991}
 Alwin, Duane F., and Jon A. Krosnick. 1991.
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-bail2018}{}}%
+\bibitem[\citeproctext]{ref-bail2018}
 Bail, Christopher A., Lisa P. Argyle, Taylor W. Brown, John P. Bumpus,
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-bardi2009}{}}%
+\bibitem[\citeproctext]{ref-bardi2009}
 Bardi, Anat, Julie Anne Lee, Nadi Hofmann-Towfigh, and Geoffrey Soutar.
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-bourdieu1990}{}}%
+\bibitem[\citeproctext]{ref-bourdieu1990}
 Bourdieu, Pierre. 1990. \emph{The Logic of Practice}. Stanford, Calif. :
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-boutyline2017}{}}%
+\bibitem[\citeproctext]{ref-boutyline2017}
 Boutyline, Andrei, and Stephen Vaisey. 2017. {``Belief {Network}
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-brett2021}{}}%
+\bibitem[\citeproctext]{ref-brett2021}
 Brett, Gordon, and Andrew Miles. 2021. {``Who {Thinks} {How}? {Social}
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-brocic2021}{}}%
+\bibitem[\citeproctext]{ref-brocic2021}
 Broćić, Miloš, and Andrew Miles. 2021. {``College and the {`{Culture}
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-campbell2016}{}}%
+\bibitem[\citeproctext]{ref-campbell2016}
 Campbell, Colin, and Jonathan Horowitz. 2016. {``Does {College}
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-christakis2010}{}}%
+\bibitem[\citeproctext]{ref-christakis2010}
 Christakis, Nicholas, and James Fowler. 2010. \emph{Connected: {The}
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-dellaposta2018}{}}%
+\bibitem[\citeproctext]{ref-dellaposta2018}
 DellaPosta, Daniel. 2018. {``Gay {Acquaintanceship} and {Attitudes}
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-dellaposta2020}{}}%
+\bibitem[\citeproctext]{ref-dellaposta2020}
 DellaPosta, Daniel. 2020. {``Pluralistic {Collapse}: {The} {`{Oil}
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-eaton2009}{}}%
+\bibitem[\citeproctext]{ref-eaton2009}
 Eaton, Asia A., Penny S. Visser, Jon A. Krosnick, and Sowmya Anand.
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-elder2003}{}}%
+\bibitem[\citeproctext]{ref-elder2003}
 Elder, Glen H., Monica Kirkpatrick Johnson, and Robert Crosnoe. 2003.
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-feldman1992}{}}%
+\bibitem[\citeproctext]{ref-feldman1992}
 Feldman, Stanley, and John Zaller. 1992. {``The {Political} {Culture} of
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-gelman2021}{}}%
+\bibitem[\citeproctext]{ref-gelman2021}
 Gelman, Andrew, and Yotam Margalit. 2021. {``Social Penumbras Predict
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-data_soep}{}}%
+\bibitem[\citeproctext]{ref-data_soep}
 Goebel, Jan, Markus M. Grabka, Stefan Liebig, Martin Kroh, David
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-goldberg2018}{}}%
+\bibitem[\citeproctext]{ref-goldberg2018}
 Goldberg, Amir, and Sarah K. Stein. 2018. {``Beyond {Social}
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-gross2009}{}}%
+\bibitem[\citeproctext]{ref-gross2009}
 Gross, Neil. 2009. {``A {Pragmatist} {Theory} of {Social}
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-guhin2020}{}}%
+\bibitem[\citeproctext]{ref-guhin2020}
 Guhin, Jeffrey, Jessica McCrory Calarco, and Cynthia Miller-Idriss.
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-hout2016}{}}%
+\bibitem[\citeproctext]{ref-hout2016}
 Hout, Michael, and Orestes P. Hastings. 2016. {``Reliability of the
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-data_psid}{}}%
+\bibitem[\citeproctext]{ref-data_psid}
 Income Dynamics, Panel Study of. 2013. {``Public Use Dataset.''}
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-kiley2020}{}}%
+\bibitem[\citeproctext]{ref-kiley2020}
 Kiley, Kevin, and Stephen Vaisey. 2020. {``Measuring {Stability} and
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-kratz2021}{}}%
+\bibitem[\citeproctext]{ref-kratz2021}
 Kratz, Fabian. 2021. {``Do Concerns about Immigration Change After
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-krosnick1989}{}}%
+\bibitem[\citeproctext]{ref-krosnick1989}
 Krosnick, Jon, and Duane F. Alwin. 1989. {``Aging and {Susceptibility}
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-lersch2023}{}}%
+\bibitem[\citeproctext]{ref-lersch2023}
 Lersch, Philipp M. 2023. {``Change in {Personal} {Culture} over the
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-lizardo2017}{}}%
+\bibitem[\citeproctext]{ref-lizardo2017}
 Lizardo, Omar. 2017. {``Improving {Cultural} {Analysis}: {Considering}
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-longest2013}{}}%
+\bibitem[\citeproctext]{ref-longest2013}
 Longest, K. C., S. Hitlin, and S. Vaisey. 2013. {``Position and
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-mannheim1952}{}}%
+\bibitem[\citeproctext]{ref-mannheim1952}
 Mannheim, Karl. 1952. {``The {Problem} of {Generations}.''} Pp. 276--322
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-mewes2021}{}}%
+\bibitem[\citeproctext]{ref-mewes2021}
 Mewes, Jan, Malcolm Fairbrother, Giuseppe Nicola Giordano, Cary Wu, and
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-quinn2024}{}}%
+\bibitem[\citeproctext]{ref-quinn2024}
 Quinn, Joseph M., Robert E. Freeland, E. K. Maloney, Kimberly B. Rogers,
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-quinn2023}{}}%
+\bibitem[\citeproctext]{ref-quinn2023}
 Quinn, Joseph M., Robert E. Freeland, Kimberly B. Rogers, Jesse Hoey,
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-raftery1995}{}}%
+\bibitem[\citeproctext]{ref-raftery1995}
 Raftery, Adrian E. 1995. {``Bayesian {Model} {Selection} in {Social}
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-robinson2007}{}}%
+\bibitem[\citeproctext]{ref-robinson2007}
 Robinson, Dawn T. 2007. {``Control {Theories} in {Sociology}.''}
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-ryder1965}{}}%
+\bibitem[\citeproctext]{ref-ryder1965}
 Ryder, Norman B. 1965. {``The {Cohort} as a {Concept} in the {Study} of
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-scoville2022}{}}%
+\bibitem[\citeproctext]{ref-scoville2022}
 Scoville, Caleb, Andrew McCumber, Razvan Amironesei, and June Jeon.
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-slothuus2021}{}}%
+\bibitem[\citeproctext]{ref-slothuus2021}
 Slothuus, Rune, and Martin Bisgaard. 2021. {``How {Political} {Parties}
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-data_gss}{}}%
+\bibitem[\citeproctext]{ref-data_gss}
 Smith, Tom W., Michael Davern, Jeremy Freese, and Stephen L. Morgan.
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-smith-lovin1988}{}}%
+\bibitem[\citeproctext]{ref-smith-lovin1988}
 Smith-Lovin, Lynn, and David R. Heise. 1988. \emph{Analyzing {Social}
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-stewart2023}{}}%
+\bibitem[\citeproctext]{ref-stewart2023}
 Stewart, Evan, and Diane Berkman. 2023. {``Shifting or Settled? Tracking
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-data_hilda}{}}%
+\bibitem[\citeproctext]{ref-data_hilda}
 Summerfield, Michelle, Ross Dunn, Simon Freidin, Markus Hahn, Peter
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-swidler2001}{}}%
+\bibitem[\citeproctext]{ref-swidler2001}
 Swidler, Ann. 2001. \emph{Talk of {Love}: {How} {Culture} {Matters}}.
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-data_bhps}{}}%
+\bibitem[\citeproctext]{ref-data_bhps}
 Taylor, Marcia Freed. 1996. {``British Household Panel Survey User
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-tourangeau2000}{}}%
+\bibitem[\citeproctext]{ref-tourangeau2000}
 Tourangeau, Roger, Lance J. Rips, and Kenneth A. Rasinski. 2000.
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-underwood2022}{}}%
+\bibitem[\citeproctext]{ref-underwood2022}
 Underwoood, Ted, Kevin Kiley, Wenyi Shang, and Stephen Vaisey. 2022.
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-data_ukhls}{}}%
+\bibitem[\citeproctext]{ref-data_ukhls}
 University of Essex, Institute for Social, and Economic Research. 2019.
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-vaisey2021}{}}%
+\bibitem[\citeproctext]{ref-vaisey2021}
 Vaisey, Stephen, and Kevin Kiley. 2021. {``A {Model}-{Based} {Method}
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-vaisey2016}{}}%
+\bibitem[\citeproctext]{ref-vaisey2016}
 Vaisey, Stephen, and Omar Lizardo. 2016. {``Cultural {Fragmentation} or
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-visser2004}{}}%
+\bibitem[\citeproctext]{ref-visser2004}
 Visser, Penny S., and Robert R. Mirabile. 2004. {``Attitudes in the
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-data_shp}{}}%
+\bibitem[\citeproctext]{ref-data_shp}
 Voorpostel, Marieke, Robin Tillmann, Florence Lebert, Ursina Kuhn,
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-weeden2005}{}}%
+\bibitem[\citeproctext]{ref-weeden2005}
 Weeden, Kim A., and David B. Grusky. 2005. {``The {Case} for a {New}
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-wodtke2018}{}}%
+\bibitem[\citeproctext]{ref-wodtke2018}
 Wodtke, Geoffrey T. 2018. {``The {Effects} of {Education} on {Beliefs}
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-zaller1992}{}}%
+\bibitem[\citeproctext]{ref-zaller1992}
 Zaller, John. 1992.
@@
 
-\newpage
-
-\hypertarget{appendix}{%
-\section{Appendix}\label{appendix}}
-
-\hypertarget{appendix-a-the-distribution-of-variation-components}{%
-\subsection{Appendix A: The Distribution of Variation
-Components}\label{appendix-a-the-distribution-of-variation-components}}
-
-Figure A1 shows the distribution of \(V(D)\) and \(V(C)\) across all
-panels.
-
-\begin{center}
-<Figure A1 about here>
-\end{center}
-
-\hypertarget{appendix-b-omega-values-when-the-last-wave-removed}{%
-\subsection{\texorpdfstring{Appendix B: \(\omega\) Values When the Last
-Wave
-Removed}{Appendix B: \textbackslash omega Values When the Last Wave Removed}}\label{appendix-b-omega-values-when-the-last-wave-removed}}
-
-Figure A2 presents the alternative strategy we used to understand the
-effects of duration on intraindividual change. First, we removed the
-last wave from all the observations and fitted the Life Course Adaption
-models. This effectively reduced the number of items to 249. We then
-refitted the model for all participants in this sample using the
-unrestricted data. In the final step, we calculated the \(\omega\)
-values for each item. The Figure A2 shows a scatterplot for these two
-sets of observations.
-
-\begin{center}
-<Figure B1 about here>
-\end{center}
-
-\newpage
-
-\hypertarget{tables-and-figures}{%
-\section{Tables and Figures}\label{tables-and-figures}}
-
-\begin{table}[H]
-\centering
-\caption{\label{tab:unnamed-chunk-1}The Description of the Data Sources}
-\centering
-\resizebox{\ifdim\width>\linewidth\linewidth\else\width\fi}{!}{
-\begin{tabular}[t]{llcc}
-\toprule
-Country & Survey & Period & Outcomes\\
-\midrule
-Australia & Household, Income and Labor Dynamics in Australia (HILDA) & 2001-2021 & 30\\
-Germany & Socio-Economic Panel Study (SOEP) & 1984-2020 & 122\\
-Great Britian & British Household Panel Survey (BHPS) & 1991-2008 & 83\\
-Great Britian & Understanding Society/UK Household Longitudinal Study (UKHLS) & 2009-2020 & 66\\
-Switzerland & Swiss Household Panel (SHP) & 1999-2019 & 77\\
-\addlinespace
-United States & General Social Survey (GSS) & 2006-2012 & 183\\
-United States & Panel Study of Income Dynamics (PSID) & 1968-2019 & 48\\
-\bottomrule
-\end{tabular}}
-\end{table}
-
-\newpage
-
-\begin{figure}[ht]
-\begin{center}
-\caption{Proportions of Systematic Variance in Personal Culture}
-
-\includegraphics[width=1\linewidth]{../figures/figure_1}
-
-\end{center}
-\footnotesize{\textit{Notes:} The figure shows $\omega$ and 1 - $\omega$ as the proportion of systematic variance attributable to intrapersonal change and interpersonal differences.
-See Supplemental Materials A for the full set of item values.}
-\end{figure}
-
-\newpage
-
-\begin{figure}[ht]
-\begin{center}
-\caption{Regression Model Estimating $\omega$}
-
-\includegraphics{../figures/figure_2.png}
-
-\end{center}
-\footnotesize{\textit{Notes:} The model estimates the proportion of systematic variance attributable to intrapersonal change.
-It includes the log of the number of participants, the number of waves the variable was asked (capped at t = 10 for each item), the date of the first wave (measured as the year minus 1968 divided by 10), number of waves the question was asked averaged across participants, duration in years per item averaged across participants, and panel and topic fixed effects.
-Coefficients are estimated from suppressed intercept model based on predictions at participants' wave mid-point.
-Survey indicators and item topics not shown.
-See Supplemental Materials B for the full set of coefficients.}
-\end{figure}
-
-\newpage
-
-\begin{figure}[ht]
-\begin{center}
-\caption{Difference in $\omega$ Across College Graduates and Non-Graduates}
-
-\includegraphics{../figures/figure_3.png}
-
-\end{center}
-\footnotesize{\textit{Notes:} The figure shows the difference in $\omega$ values across college graduates and college non-graduates.
-Values above (below) 0 means that those with college degree have higher (lower) variance of intrapersonal change.
-The dashed red line marks 0 difference.}
-\end{figure}
-
-\newpage
-
-\begin{figure}[ht]
-\begin{center}
-\caption{Difference in $\omega$ Across College Graduates and Non-Graduates on Political Culture}
-
-\includegraphics{../figures/figure_4.png}
-
-\end{center}
-\footnotesize{\textit{Notes:} $\omega$ values across college graduates and college non-graduates on 8 political culture items from the General Social Survey (2006-2014).}
-\end{figure}
-
-\newpage
-
-\begin{figure}[htp]
-\begin{center}
-\caption*{Figure A1: The Distribution of V(D) and V(C)}
-
-\includegraphics{../figures/figure_A1.png}
-
-\end{center}
-\end{figure}
-
-\newpage
-
-\begin{figure}[htp]
-\begin{center}
-\caption*{Figure A2: Difference in $\omega$ Across Two Sample Specifications}
-
-\includegraphics{../figures/figure_A2.png}
-
-\end{center}
-\end{figure}
-
 \end{document}

### 2024-09-13 write-ups/manuscript.Rmd
updates (pending case study).

--- 2024-08-22 write-ups/manuscript/manuscript.Rmd
@@
     keep_tex: true
+bibliography: kileybib.bib
+csl: asa.csl
+date: "`r format(Sys.time(), '%B %Y')`"
 geometry: margin=1in 
@@
 fontfamily: mathpazo
+abstract: "Work in the sociology of culture has converged on a general conclusion that while people sometimes make substantial shifts in their personal culture over time, most elements of personal culture are characterized by low rates of persistent change during adulthood.
+Recent developments have begun to quantify change over time and initial differences, but it is unclear how to use these components to understand the relative importance of change, especially when change is modeled non-linearly.
+To advance this debate, we introduce a measure for quantifying the proportion of systematic variance in panel data attributable to intrapersonal change.
+Applying this measure to 610 items from seven surveys in five countries, we find that although intrapersonal change is common, it does not explain a large share of systematic variance.
+As an extension of the theoretical debate, we demonstrate that this measure provides new insights when comparing social groups, showing that intrapersonal change is less common among U.S.
+college graduates than among those without a college degree.
+Our findings provide a new perspective on several important theoretical debates, as well as a tool to address new questions."
 header-includes:
@@
                 singlelinecheck=false}
-bibliography: kileybib.bib
-csl: asa.csl
-abstract: "Work in the sociology of culture has converged on a general conclusion that while people sometimes make substantial shifts in their personal culture over time, most elements of personal culture are characterized by low rates of persistent change during adulthood.
-Recent developments have begun to quantify change over time and initial differences, but it is unclear how to use these components to understand the relative importance of change, especially when change is modeled non-linearly.
-To advance this debate, we introduce a measure for quantifying the proportion of systematic variance in panel data attributable to intrapersonal change.
-Applying this measure to 610 items from seven surveys in five countries, we find that although intrapersonal change is common, it does not explain a large share of systematic variance.
-As an extension of the theoretical debate, we demonstrate that this measure provides new insights when comparing social groups, showing that intrapersonal change is less common among U.S.
-college graduates than among those without a college degree.
-Our findings provide a new perspective on several important theoretical debates, as well as a tool to address new questions."
 ---
@@
 Recent work on this topic has converged on a general conclusion that people do sometimes make substantial shifts in their personal culture over time but that most elements of personal culture are characterized by low rates of persistent change during adulthood [@kiley2020; @lersch2023].
-Work leading to this conclusion adopted a "tournament of models" approach to adjudicating different theoretical perspectives [@lersch2023: p.
+Work leading to this conclusion adopted a "tournament of models" approach to adjudicating different theoretical perspectives [@lersch2023, p.
 228].
@@
 
-However, as consensus in the overall picture emerges, researchers are now turning toward understanding the conditions that produce persistent change in personal culture, including age, education, and life-course transitions; whether certain kinds of people are more likely to make persistent changes in personal culture; and whether certain kinds of personal culture are more likely to be characterized by persistent change or stability.
+However, as consensus emerges, researchers are now turning toward understanding the conditions that produce persistent change in personal culture, including age, education, and life-course transitions; whether certain kinds of people are more likely to make persistent changes in personal culture; and whether certain kinds of personal culture are more likely to be characterized by persistent change or stability.
 The tournament of models approach, useful for the broad question of whether change happens, is ill-suited to addressing these more theoretically grounded questions.
@@
 In quantifying these aspects of cultural difference, the LCAM lets researchers address a broader range of questions than the single question of whether a measure is characterized by stability or change.
-However, because the LCAM produces so many different measures, it is not obvious how to use these components to understand the relative importance of within-person change, especially when change is measured in non-linear ways.
+However, because the LCAM produces so many different measures, it is not obvious how to use these components to understand the relative importance of within-person change across items, groups, or time periods, especially when cultural change is measured in non-linear ways.
 
@@
 Our approach isolates the "systematic" components of variance (i.e., it removes random change and measurement error), to provide a single summary of how much systematic variation can reasonably be attributed to people making persistent changes over time.
-This quantity can then be used as a quick diagnostic to compare across questions, groups, periods, or samples to isolate where within-person change is more or less useful for explaining variation.
+This quantity---what we call $\omega$---can then be used as a diagnostic to compare across questions, groups, or periods to isolate where within-person change is more or less useful for explaining aggregate variation.
 We demonstrate the utility of this quantity in two ways.
@@
 This allows us to generate an overall picture of how much of the total systematic variation in personal culture change tends to explain.
-These findings largely echo previous findings but enable clearer statements about the relative importance of change across questions, rather than binary claims about whether change "happens" or specific quantification of how much change occurs.
+These findings largely echo previous findings, but enable clearer statements about the relative importance of change across questions, rather than binary claims about whether change "happens."
 
-Second, to illustrate how this quantity might be used by a researcher with a specific topical or theoretical interest, we investigate differences in the relative importance of intrapersonal change for people with and without a college degree.
+Second, to illustrate how this quantity might be used by a researcher with a specific theoretical interest, we investigate differences in the relative importance of intrapersonal change for people with and without a college degree, using several political preferences in the Swiss Household Panel as a case study.
 We find that change explains a smaller proportion of systematic variance among college-educated respondents, suggesting that college crystallizes one's personal culture rather than fostering an openness to new information.
-This offers a new empirical basis to theorize about the role of education in influencing personal culture.
+This offers a new empirical basis to theorize about the role of education in influencing personal culture, while at the same time showcasing a simple use of our proposed $\omega$ measure in a substantive theoretical context.
 
@@
 This model underlies, among others, theories of cultural diffusion [@christakis2010], attitude alignment [@dellaposta2020], and polarization [@bail2018].
-It is also implicit in most studies that ask whether specific experiences, changes in social roles, or political events, affect personal culture [@gelman2021; @visser2004; @slothuus2021].
+It is also implicit in most studies that ask whether specific experiences, changes in social roles, or political events, affect personal culture [@gelman2021; @slothuus2021].
 
@@
 There was also more evidence of durable change among early adults (people ages 18-30) than among the rest of the population.
-Overall, the researchers concluded that "results ultimately suggest that real, persistent attitude change is an uncommon phenomenon among adults'' [@kiley2020: p.
-500; see also @vaisey2021].
+Overall, the researchers concluded that "results ultimately suggest that real, persistent attitude change is an uncommon phenomenon among adults" [@kiley2020, p.
+500].
 This lack of durable change is consistent with other recent findings that cohort replacement plays a larger role than period effects in explaining differences in personal culture [@vaisey2016].
@@
 In doing this, Lersch rectified three shortcomings of the AUM and SDM.
-First, the AUM, as described by Kiley and Vaisey, posits that changes follow a Markov process where responses at time $t$ are a function of responses at time $t-1$ but not earlier time points.
+First, the AUM, as described by Kiley and Vaisey, posits that changes follow a Markov process where responses at time $t$ are a function of responses at time $t-1$, but not earlier time points.
 However, earlier life experiences can (directly and indirectly) mold personal culture when transitioning to new social roles or into new environments, even if their initial impact is delayed.
 For example, childhood events might influence views on family structures later when individuals form their own families.
-Therefore, the LCAM considers influences from earlier than just $t-1$ on responses at time $t$.
+Therefore, the LCAM considers earlier influences on responses at time $t$.
 
-[^linear]: In supplemental materials, Lersch models age using splines as well, with results generally showing the same conclusions.
+[^linear]: In supplemental materials, Lersch uses splines as well, with results generally showing the same conclusions.
 
 Second, Kiley and Vaisey's conclusions about the AUM and SDM were based on examining three-wave panel data over four years, which might not be intensive or extensive enough to adjudicate the two models.
-Lersch evaluates the LCAM against the AUM and SDM using panel data spanning a wider duration (from 3 to 36 years) and more waves (3 to 18), offering a better chance to observe durable change.
+Lersch evaluates the LCAM against the AUM and SDM using panel data spanning a wider duration (from 3 to 36 years) and more waves at the respondent level (3 to 18), offering a better chance to observe durable change in a longer time frame.
 
 Third, and most importantly, the LCAM quantifies various components of cultural difference, allowing researchers to ask a more varied set of questions than simply whether people change.
-For example, researchers can use the LCAM framework to quantify the magnitude of change in a measure of culture over time (i.e., how much do people change on average?), the variance in age-based slopes over time (i.e., do people tend to change in similar ways as they age?), or how much personal culture tends to fluctuate around these age-based trends (i.e., how consistent are responses over time?).
+For example, researchers can use the LCAM framework to quantify the magnitude of change in a measure of culture over time (how much do people change on average?), the variance in age-based slopes over time (do people tend to change in similar ways as they age?), or how much personal culture tends to fluctuate around these age-based trends (how consistent are responses over time?).
 
@@
 Kiley and Vaisey [-@kiley2020] found that the majority of items they tested favored the AUM, meaning there was evidence of persisting change among adults, a pattern consistent with Lersch's results.
-Also Lersch [-@lersch2023] found evidence of persisting change on most items, but these changes were small.
+Also, Lersch [-@lersch2023] found evidence of persisting change on most items, but these changes were relatively small.
 On average, people changed only about .07 standard deviations over 10 years.
-This is consistent with Kiley and Vaisey's finding that persisting change is often small.
+This is consistent with Kiley and Vaisey's finding that persisting change is often small in magnitude.
 Moreover, a quarter of the items Lersch studied still preferred the SDM, suggesting that even with different assumptions about change and more extensive data, many questions still are consistent with a model that allows zero intrapersonal change.
@@
 Other studies are compatible with these findings as well.
-For example, Brocic and Miles [-@brocic2021] estimate that completing graduate degrees in humanities, arts, and social sciences shift peoples' moral relativism only about 0.2 standard deviations on average compared to people with no college degree, and this is the largest effect they identify.
+For example, Brocic and Miles [-@brocic2021] estimate that completing graduate degrees in humanities, arts, and social sciences shift peoples' moral relativism only about .20 standard deviations on average compared to people with no college degree, and this is the largest effect they identify.
 And studies of aggregate change show that, even on items where cohort effects explain more variance than period effects, there is always evidence that some people change over time [@vaisey2016].
@@
 To this point, the debate has broadly centered on drawing conclusions about whether people *ever* change.
-But this is rarely the right question since the answer is almost always "yes." Over sufficient time and in a large enough sample, researchers will likely observe some evidence for durable change in the population.
+But this is rarely the right question since the answer is almost always "yes." Over sufficient time and in a large enough sample, researchers will likely observe some evidence for change in the population.
 A lack of evidence for change may be due to a poor survey question, a low resolution in response options, or that the question simply was not asked for long enough or to enough people.
@@
 The first component, stable interpersonal differences, reflects the accumulated experiences of people prior to entering the panel survey.
-Lersch [-@lersch2023: 24] calls this "early imprinting." While commonly associated with experiences during a formative period that result in settled dispositions, this variation could also reflect experiences that happened at any time as long as they predate the panel and consistently affect subsequent responses.
-For instance, for those people entering the panel post-retirement, this "imprinting" might reflect this pivotal life transition.
+Lersch [-@lersch2023, p.
+24] calls this "early imprinting." While commonly associated with experiences during a formative period that result in settled dispositions, this variation could also reflect experiences that happened at any time as long as they predate the panel and consistently affect subsequent responses.
+For instance, for those people entering the panel post-retirement, this "imprinting" might reflect this pivotal life transition and its subsequent effects observed in the panel window.
 Consequently, this component also reflects variation in individuals' social roles or statuses at the start of the panel that were important in shaping their dispositions.
@@
 The second component, the amount of intrapersonal change people make during the panel, captures persisting changes in personal culture over time.
-This component captures the set of processes collectively called "persistent change" or "adaption" by Lersch and "active updating" by Kiley and Vaisey.
-Lersch attributes these changes to social triggers such as moving into a new environment or adopting new social roles (although he does not measure these directly).
-They might also reflect the diffusion of new cultural forms across social networks, cues from political elites or otherwise culturally influential leaders, the emergence of issues in politics or culture, or large-scale social shifts.
+This component captures the set of processes collectively called "persistent change" or "adaption" by Lersch [-@lersch2023] and "active updating" by Kiley and Vaisey [-@kiley2020].
+Lersch [-@lersch2023] attributes these changes to social triggers such as moving into a new environment or adopting new social roles (although he does not measure these directly).
+They might also reflect the diffusion of new cultural forms or ideas across social networks, cues from political elites or otherwise culturally influential leaders, the emergence of issues in politics or culture, or large-scale social shifts.
 
@@
 Although this third component at least partially reflects important processes of personal culture, it does not directly touch on the ongoing debate outlined here.
-Because of this, we quantify this component but focus principally on the other two.
+Hence, we quantify this component but focus principally on the other two.
 
@@
 In other words, when we look at a population of adults over some time frame, is there enough intrapersonal change relative to interpersonal difference to view it as an important process for why people differ from each other?
-Only through quantifying the two components we can get closer to a true answer to this question.
+Only through quantifying the two components we can get closer to a true answer.
 
@@
 This model assumes a set of propositions about change that reflect the theoretical debate to this point.
-First, consistent with the settled disposition model, it assumes that people start the survey with cultural differences, modeled as random intercepts for each respondent.
+First, consistent with the SDM, it assumes that people start the survey with cultural differences, modeled as random intercepts for each respondent.
 Second, it assumes that people change over time, taking the form of random slopes for each respondent as a linear function of time.
@@
 
-where $\beta_0$ is the average intercept, $\alpha_{0i}$ is the random intercept component for individuals, $\beta_1$ is the average yearly change in the outcome, $\alpha_{1i}$ is the random slope component for individuals, and $\epsilon_{it}$ is the random error term that captures transitory disturbances.
+where $\beta_0$ is the average intercept, $\alpha_{0i}$ is the random intercept component for individuals, $\beta_1$ is the average yearly change in the outcome, $\alpha_{1i}$ is the random slope component for individuals, and $\epsilon_{it}$ is the random error term that captures fluctuations.
 The $\alpha_{0i}$ and $\alpha_{1i}$ terms are also allowed to covary.
@@
 
-This effectively models respondents' culture under the assumption that they make no systematic change over time, or that the expected value at any time point is the same as the expected value at any other.
-This gives us the proportion of total variance in an item attributable to stable interpersonal differences.
+This effectively models respondents' responses under the assumption that they make no systematic change over time, or that the expected value at any time point is the same as the expected value at any other.
+Therefore, this measure---V(D)---gives us the proportion of total variance in an item attributable to stable interpersonal differences.
 
@@
 
-This is the incremental proportion of variance accounted for when we allow the model predictions to change over time for each person.
+This is the incremental proportion of variance accounted for when we allow the model predictions to change over time for each person, allowing us to quantify the explanatory power of change.
 
@@
 
-Together, $V(D)$ and $V(C)$ reflect the total systematic variance of the outcome that the LCAM can attribute to either differences between people or linear intrapersonal change over the course of the panel.
+Together, $V(D)$ and $V(C)$ reflect the total systematic variance of the outcome that the LCAM can attribute to either differences between people or linear intrapersonal change[^alternatives] over the course of the panel.
 Our principal measure of interest is the proportion of systematic variance accounted for by intrapersonal change, which we calculate as
+
+[^alternatives]: The decomposition of variance is agnostic to this choice.
+One advantage is that the modeling assumptions can be relaxed to accommodate non-linear terms or splines, but the decomposition will follow the same procedure.
 
@@
 Second, we estimate the LCAM in this slightly perturbed sample, and calculate the $\omega$.
-We then replicate this procedure $N$ times and establish confidence intervals using the distribution of bootstrapped omega values.
+We then replicate this procedure $N$ times and establish confidence intervals around our measure.
 
@@
 
-To demonstrate the utility of this metric, we provide two examples. In the first, we compare the proportion of systematic variance explained by intrapersonal change across a broad array of questions.
+To demonstrate the utility of $\omega$, we provide two examples. In the first, we compare the proportion of systematic variance explained by intrapersonal change across a broad array of questions.
 
-## Data
+## Data and Analysis
 
-We use data from seven nationally representative panel surveys from Australia, Germany, Great Britain, Switzerland, and the United States (summarized in Table 1), combining all the data files used in previous work [@kiley2020; @lersch2023].[^1] These studies cover a long period of time (the range of surveys spans from 1968 to 2021), with 609 personal culture items capturing attitudes, beliefs, values, self-assessments, self-descriptions, and behaviors [@alwin2007].
+We use data from seven nationally representative panel surveys from Australia, Germany, Great Britain, Switzerland, and the United States (summarized in Table 1), combining all the data files used in previous work [@kiley2020; @lersch2023].[^1] These studies cover a long period of time (the range of surveys spans from 1968 to 2021), with 610 personal culture items capturing attitudes, beliefs, values, self-assessments, self-descriptions, and behaviors [@alwin2007].
 We restricted the sample such that individuals between the ages of 18 and 79 are included without further elimination, and in all surveys, we used all possible cases for which respondents provided responses.
 In the end, the analyses that follow rely on a cross-national sample with a cross-domain set of items to capture a broad range of individual personal culture.
-Supplemental Materials A documents the list of all variables used in the upcoming analyses.
+We document the list of all variables used in the upcoming analyses in the Appendix.
 
-[^1]: For more information on these data sources, see [@data_gss; @data_hilda; @data_bhps; @data_ukhls; @data_soep; @data_shp; @data_psid].
+[^1]: For more information on these data sources, see the relevant panel files [@data_gss; @data_hilda; @data_bhps; @data_ukhls; @data_soep; @data_shp; @data_psid].
 
@@
                   align = "llcc",
+                  linesep = "",
                   caption = "The Description of the Data Sources") |> 
@@
 
-## Analysis
-
-As a first step in our analysis, we fit this LCAM to each of the 609 measures of personal culture outlined above.
-We then derive $\omega$ for each of these items.
-As a second step, we use linear regression to model $\omega$ as a function of features of the question and panel.
-This allows us to explore how these features are associated with the degree of intrapersonal change.
-These covariates include the number of response options (measured as 2, 3-5, and more than 5), the specific survey, the log of the number of participants, the number of waves observed, the time period (i.e., decade) the question was first asked, and the total duration of time the question was asked.
+As a first step in our analysis, we fit this LCAM to each of the 610 measures of personal culture outlined above. We then derive $\omega$ for each of these items using the procedure above.
 
@@
 
-Figure 1 plots the proportion of systematic variance attributable to interpersonal differences and intrapersonal change for each of the 609 questions, plotted separately by panel.
+Figure 1 plots the proportion of systematic variance attributable to interpersonal differences and intrapersonal change for each of the 610 questions, plotted separately by panel survey.
 To broadly summarize the results, all panels show a similar range of the proportion of systematic variance attributable to intrapersonal change.
-Across panels, the median value of $\omega$ is 0.094 (mean of 0.098), with an interquartile range of 0.053 to 0.139.
-All panels include questions where $\omega$ is essentially 0, and all panels include questions with $\omega$ values greater than 0.20.
-The maximum $\omega$ value is 0.26.
+Across panels, the median value of $\omega$ is .053 (mean of .056), with an interquartile range of .032 to .077.
+All panels include questions where $\omega$ is essentially 0, and some panels include questions with $\omega$ values greater than .20.
+The maximum $\omega$ value is .23.
 
 ```{=latex}
-\begin{figure}[ht]
+\begin{figure}[!p]
 \begin{center}
-\caption{Proportions of Systematic Variance in Personal Culture}
 ```
@@
 \end{center}
+\caption*{Figure 1: Proportions of Systematic Variance in Personal Culture}
 \footnotesize{\textit{Notes:} The figure shows $\omega$ and 1 - $\omega$ as the proportion of systematic variance attributable to intrapersonal change and interpersonal differences.
-See Supplemental Materials A for the full set of item values.}
+See the Appendix for the full set of item values.}
 \end{figure}
@@
 
-Across all questions, interpersonal differences account for a much larger share of the systematic variance in responses than intrapersonal change.
+Across all questions and all surveys, interpersonal differences account for a much larger share of the systematic variance in responses than intrapersonal change.
 Again, this is to be expected.
@@
 
-To the extent that there are differences across the panels, the PSID has the highest $\omega$ values with mean .120 and median .135.
+To the extent that there are differences across the panels, the PSID has the highest $\omega$ values with mean .068 and median .076.
 While we cannot disentangle features of the sample from features of the questions asked to each sample, the specific samples for many PSID questions have lower average ages than those from other panels.
 To the extent that younger respondents might be more likely to make durable changes of opinion, these higher estimates of intrapersonal change might reflect the distinct age profile of respondents in this sample.
-At the other end, the GSS has the lowest range of $\omega$ with mean .073 and median .069.
-This potentially reflects the fact that the GSS observes people for a shorter duration, on average, than the other panels.
-If, consistent with life course adaption theories, people are more likely to make significant cultural changes the longer we observe them, then duration likely affects the range of $\omega$ (a point we explore more below).
-However, the GSS results are still consistent with results from the other panels.
+At the other end, the SHP has the lowest range with mean .042 and median .037, followed by GSS, with mean .044 and median .043.
 
 While there are some differences between panels, these differences are small compared to the differences within panels.
-For about 6 percent of items, $\omega$ is greater than 0.20.
+For about 9 percent of items, $\omega$ is greater than .10.
 These questions tend to ask about objectively changing external referents (e.g., confidence in specific government leaders or political parties), life satisfaction, or current financial position.
-At the other end, questions about religious identification, views on gender roles, and support for civil liberties tend to have very low estimates of intrapersonal change.
+At the other end, questions about religious identification, views on gender roles, and support for civil liberties tend to have very low estimates of intrapersonal change, suggesting that core preferences might be less variable.
 
@@
 For example, Kiley and Vaisey [-@kiley2020] found that confidence in the press and confidence in religious leaders were both characterized by active updating.
-Our results show that intrapersonal change is much more important for explaining variance in confidence in the press (0.164) than confidence in religion (0.049), even though both are updating.
+Our results show that intrapersonal change is much more important for explaining variance in confidence in the press (.091) than confidence in religion (.051), even though in both cases people are updating their beliefs.
 
-Appendix A shows the distribution of $V(D)$ and $V(C)$ across panels, and Supplemental Materials A presents the estimated proportion of variance attributable to interpersonal differences and intrapersonal change, the estimated values of $\omega$, and the proportion of residual variance for each question.
-Interpersonal differences are almost always the largest component of the total variance and tend to account for between 55 and 70 percent of total variance, while intrapersonal change is always the smallest, typically accounting for between 3 and 8 percent of total variance.
-Residual variance tends to account for between 22 and 37 percent of variance, though on several questions residual variance is greater than 50 percent.
+In the Appendix, we show the distribution of $V(D)$, $V(C)$ and the residual variance $V(E)$ for each question across panels.
+Interpersonal differences are almost always the largest component of the total variance and tend to account for between 50 and 66 percent of total variance, while intrapersonal change is always the smallest, typically accounting for between 2 and 5 percent of total variance.
+Residual variance tends to account for between 30 and 46 percent of variance, though on several questions residual variance is greater than 50 percent.
 This might indicate survey items with low reliability or ones that capture genuinely rapid fluctuations.
-
-As we noted above, the substantive importance of a given amount of intrapersonal change depends on a range of factors, including how long the panel runs, whether assumptions about linear change hold, and whether the panel is capturing a distinctly turbulent period or a distinctly stable period for the relevant item.
-However, if we assume that the period under observation is "typical" for a question---not a time of extremely heightened (or lowered) sensitivity or change---then it does not seem realistic that intrapersonal change accounts for a large share of the cultural differences we see in the world.
-
-Figure 2 plots the results from a linear regression of $\omega$ as a function of question, panel, and sample features.
-These models also include fixed effects for panels and topics, so coefficients reflect the association within a panel and topical domain.
-
-```{=tex}
-\begin{center}
-<Figure 2 about here>
-\end{center}
-```
-
-Figure 2 shows that the more response options respondents are given and the larger the sample, the larger the values of $\omega$.
-We interpret these coefficients as suggesting greater resolution on a question makes it easier to detect and model change.
-The earlier a question was asked, even net of how long it was been asked, the lower $\omega$.
-
-Although we did not state expectations for how question content would relate to $\omega$, some of the associations between question structure, panel duration, and panels themselves might be driven by differences in the topics addressed by each panel.
-To address this, we followed Hout et al.
-[-@hout2016] and Lersch [-@lersch2023] in coding each question as falling into one of nine different topical domains and included these as indicator variables in the meta analysis.
-Supplemental Materials B presents the estimates for these topic indicators.
-They show some associations with $\omega$, with questions about subjective SES; social life, social cohesion, and trust; environment and climate; and health and morale showing larger $\omega$ values than questions about religion and spirituality; politics, government and the economy; and gender and family.
-
-The most notable result in the meta analysis is the negative coefficient attached to the duration of years covered by the question on average across participants.
-The longer a question is observed, the less is the value of $\omega$.
-Theories of personal cultural change that link changes in personal culture to social experiences, including the LCAM, suggest that the longer we observe respondents, the more likely people are to undergo potentially transformative experiences and therefore the more variance would be attributable to intrapersonal change.
-Finding a negative coefficient here seems to challenge that assumption.
-
-As a further test of this finding, we compared the $\omega$ values when using the full duration of a panel compared to when we dropped the final wave for each participant and therefore reduced the total duration of observation for the question.
-If the coefficient reflects a true negative effect of duration on $\omega$, we should see that same effect within questions.
-Results from this analysis are presented in Appendix B and contradict the coefficient from the regression model; with a few exceptions, we found that the longer we observe the same question, the higher $\omega$ is.
-We interpret this combination of findings as suggesting that the kinds of questions asked for a longer time period tend to demonstrate less intrapersonal change than questions asked over shorter periods, rather than a true function of time.
-This supports the conclusion that the GSS shows less intrapersonal change because of its shorter duration.
 
@@
 
-```{=latex}
-\begin{figure}[ht]
-\begin{center}
-\caption{Difference in $\omega$ Across College Graduates and Non-Graduates}
-```
-```{r fig3}
-knitr::include_graphics("../../figures/colleges.png")
-```
-```{=latex}
-\end{center}
-\footnotesize{\textit{Notes:} The figure shows the difference in $\omega$ values across college graduates and college non-graduates.
+<!-- ```{=latex} -->
+<!-- \begin{figure}[ht] -->
+<!-- \begin{center} -->
+<!-- \caption{Difference in $\omega$ Across College Graduates and Non-Graduates} -->
+<!-- ``` -->
+<!-- ```{r fig3} -->
+<!-- knitr::include_graphics("../../figures/colleges.png") -->
+<!-- ``` -->
+<!-- ```{=latex} -->
+<!-- \end{center} -->
+<!-- \footnotesize{\textit{Notes:} The figure shows the difference in $\omega$ values across college graduates and college non-graduates.
 Values above (below) 0 means that those with college degree have higher (lower) variance of intrapersonal change.
-The dashed red line marks 0 difference.}
-\end{figure}
-```
+The dashed red line marks 0 difference.} -->
+<!-- \end{figure} -->
+<!-- ``` -->
 
@@
 We present estimates of $\omega$ for these eight questions, for both education groups, in Figure 4.
-
-```{=tex}
-\begin{center}
-<Figure 4 about here>
-\end{center}
-```
 
@@
 
-The primary objective of this paper is to introduce a new quantity that can advance debate on whether people change their personal culture -- their attitudes, beliefs, values, and practices -- as they move through their adult life.
-Instead of falling back on a "tournament of models" approach [@lersch2023: p.
+The primary objective of this paper is to introduce a new quantity that can advance debate on whether people change their personal culture---their attitudes, beliefs, values, and practices---as they move through their adult life.
+Instead of falling back on a "tournament of models" approach [@lersch2023, p.
 228] to proclaim a yes or no answer to this question, or only quantifying individual components of cultural difference to show how much people change, our metric contextualizes the amount of observed systematic variance that is attributable to intrapersonal change over the duration of a panel.
@@
 We demonstrated the utility of this measure in two ways.
-First, applying our proposed measure, $\omega$, to 609 survey items from the panel datasets previously studied in this debate revealed a consistent pattern.
+First, applying our proposed measure, $\omega$, to 610 survey items from the panel datasets previously studied in this debate revealed a consistent pattern.
 Nearly all questions show evidence of people making some durable, intrapersonal change over time.
 Some show notably high amounts of intrapersonal change.
-For some questions about life satisfaction or views on government benefits, it seems plausible that differences in adult experiences predominantly account for observed differences between people.
 However, intrapersonal change is often substantially less pronounced than interpersonal differences, accounting for less than 10 percent of systematic variance on average across all datasets.
-On questions related to civil liberties, abortion, generalized trust, and civic duty,  systematic variance attributable to intrapersonal change is essentially zero.
+On some questions, systematic variance attributable to intrapersonal change is essentially zero.
 For these questions, it seems there is not enough cultural change during adulthood to warrant attributing the differences we observe to experiences and social transitions; instead, the primary source of the observed differences appears to stem from experiences in childhood, adolescence, and early adulthood.
@@
 This allows for greater comparison across questions and it provides a benchmark for researchers exploring new questions.
-Second, our metric can help contextualize the quantities identified in the LCAM.
+Second, our metric might help contextualize the quantities identified in the LCAM.
 Is .07 standard deviations of change every 10 years a lot of change?
-For some questions, our metric will say this change explains a large component of variance, and for others it will not.
+For some questions or groups, our metric will say this change explains a large component of variance, and for others it will not.
 
-In a second application of the quantity, we showed how it can open up new lines of inquiry.
+In a second application of our metric, we showed how an analysis of $\omega$ can open up new lines of inquiry.
 We showed how the amount of cultural difference explained by intrapersonal change varies substantially not just by survey item but also by individuals' characteristics such as their education.
@@
 Longer panels and more flexible definitions of "change" might allow us to account for such trajectories.
-Again, while we model change in a single way, the general variance decomposition approach we outline accommodates different definitions and functional forms of change.
+Again, while we model change in a single way, the general variance decomposition approach we outline accommodates different functional forms.
 
@@
 Although ideal types like the settled dispositions and active updating models are useful, no theoretical perspective would expect either early life socialization or adult intrapersonal change to be the sole source of one's personal culture.
-Our results reinforce this point, showing the relevance of both factors and allowing their precise quantification.
+Our results reinforce this point, showing the relevance of both factors and allowing their precise quantification for the explanation of cultural differences.
 
@@
 The situations that promote durable change in personal culture might simply be rare during adulthood.
-Alternatively, it is possible that adults do encounter opportunities, necessities, and incentives for change, but their ability or willingness to change decreases.
+Alternatively, it is possible that adults do encounter opportunities, necessities, and incentives for change, but their ability or willingness to change decreases as a function of age.
 All the more it is important to research when and how social situations can provoke durable change in adults.
@@
 Related to this, our results regarding education and political views suggest that the importance of processes that lead to such change can vary by group.
-Aligning with life course theories, it seems the significance of experiences for cultural change may be contingent on other, prior experiences.
+Aligning with life course theories, the significance of experiences for cultural change may be contingent on other, prior experiences.
 For example, factors that shape cultural dispositions may likely differ for college and non-college graduates; the latter might be more profoundly influenced by mid-life experiences than the former in this regard.
-In trying to understand differences in personal culture, sociologists should therefore pay more attention to the heterogeneous effects that various factors including social events, encounters, and situations can have.
+In trying to understand cultural differences, sociologists should pay more attention to the heterogeneous effects that various factors including social events and encounters can have.
 
@@
 Differentiating between the two is therefore crucial for understanding cultural differences and cannot be done with cross-sectional data.
-Instead we should extend panels beyond two waves to gain more leverage to understand when and in what form personal culture changes.
-
-A key insight from our meta-analysis is that an increase in response resolution (the range of response options survey respondents are given) correlates with an increased share of systematic variance explained by intrapersonal change.
-This may reflect specific issues being assessed by survey items using different scales.
-Alternatively, it might imply that when intrapersonal changes occurs, they are often subtle with answers moving from "agree" to "strongly agree," rather than from "strongly agree" to "strongly disagree." Such subtle changes require finer response options to be detected.
+Instead, we should extend panels beyond two waves and longer time periods to gain more leverage to understand when and in what form personal culture changes.
 
@@
 It is very likely that the majority of adult differences in personal culture are rooted in different experiences before age 18.
-By empirically constraining our analyses to adult experiences, social scientists may inadvertently concentrate on topics and questions that, while important, might not be able to help explain major cultural differences in adult populations.
+By empirically constraining our analyses to adult experiences, social scientists may inadvertently concentrate on topics and questions that, while important, might not be able to help explain major cultural differences in adult populations, and the reasons why people believe in some things but not others.
 
@@
 Characterizing questions as displaying change or not can only take researchers so far, but the question of whether some questions demonstrate more change than others, or whether some groups are characterized by more stability than others, has the potential to weigh in on a broader range of theoretical debates.
-We hope researchers find our approach useful as they investigate these questions.
+We hope researchers find our approach useful as they investigate these questions across a variety of problems.
 

### 2024-09-13 write-ups/manuscript.tex
updates (pending case study).

--- 2024-08-22 write-ups/manuscript/manuscript.tex
@@
 Vaisey\footnote{Department of Sociology, Duke University}}
-\date{}
+\date{September 2024}
 
@@
 ``tournament of models'' approach to adjudicating different theoretical
-perspectives (Lersch 2023: p.~228). In this framework, researchers
+perspectives (Lersch 2023:228). In this framework, researchers
 formalized different models of change over time and relied on model
@@
 
-However, as consensus in the overall picture emerges, researchers are
-now turning toward understanding the conditions that produce persistent
-change in personal culture, including age, education, and life-course
-transitions; whether certain kinds of people are more likely to make
-persistent changes in personal culture; and whether certain kinds of
-personal culture are more likely to be characterized by persistent
-change or stability. The tournament of models approach, useful for the
-broad question of whether change happens, is ill-suited to addressing
-these more theoretically grounded questions.
+However, as consensus emerges, researchers are now turning toward
+understanding the conditions that produce persistent change in personal
+culture, including age, education, and life-course transitions; whether
+certain kinds of people are more likely to make persistent changes in
+personal culture; and whether certain kinds of personal culture are more
+likely to be characterized by persistent change or stability. The
+tournament of models approach, useful for the broad question of whether
+change happens, is ill-suited to addressing these more theoretically
+grounded questions.
 
@@
 it is not obvious how to use these components to understand the relative
-importance of within-person change, especially when change is measured
-in non-linear ways.
+importance of within-person change across items, groups, or time
+periods, especially when cultural change is measured in non-linear ways.
 
@@
 much systematic variation can reasonably be attributed to people making
-persistent changes over time. This quantity can then be used as a quick
-diagnostic to compare across questions, groups, periods, or samples to
-isolate where within-person change is more or less useful for explaining
-variation. We demonstrate the utility of this quantity in two ways.
+persistent changes over time. This quantity---what we call
+\(\omega\)---can then be used as a diagnostic to compare across
+questions, groups, or periods to isolate where within-person change is
+more or less useful for explaining aggregate variation. We demonstrate
+the utility of this quantity in two ways.
 
@@
 of how much of the total systematic variation in personal culture change
-tends to explain. These findings largely echo previous findings but
+tends to explain. These findings largely echo previous findings, but
 enable clearer statements about the relative importance of change across
-questions, rather than binary claims about whether change ``happens'' or
-specific quantification of how much change occurs.
+questions, rather than binary claims about whether change ``happens.''
 
 Second, to illustrate how this quantity might be used by a researcher
-with a specific topical or theoretical interest, we investigate
-differences in the relative importance of intrapersonal change for
-people with and without a college degree. We find that change explains a
-smaller proportion of systematic variance among college-educated
-respondents, suggesting that college crystallizes one's personal culture
-rather than fostering an openness to new information. This offers a new
-empirical basis to theorize about the role of education in influencing
-personal culture.
+with a specific theoretical interest, we investigate differences in the
+relative importance of intrapersonal change for people with and without
+a college degree, using several political preferences in the Swiss
+Household Panel as a case study. We find that change explains a smaller
+proportion of systematic variance among college-educated respondents,
+suggesting that college crystallizes one's personal culture rather than
+fostering an openness to new information. This offers a new empirical
+basis to theorize about the role of education in influencing personal
+culture, while at the same time showcasing a simple use of our proposed
+\(\omega\) measure in a substantive theoretical context.
 
@@
 in social roles, or political events, affect personal culture (Gelman
-and Margalit 2021; Slothuus and Bisgaard 2021; Visser and Mirabile
-2004).
+and Margalit 2021; Slothuus and Bisgaard 2021).
 
@@
 suggest that real, persistent attitude change is an uncommon phenomenon
-among adults'\,' (Kiley and Vaisey 2020: p.~500; see also Vaisey and
-Kiley 2021). This lack of durable change is consistent with other recent
-findings that cohort replacement plays a larger role than period effects
-in explaining differences in personal culture (Vaisey and Lizardo 2016).
+among adults'' (Kiley and Vaisey 2020:500). This lack of durable change
+is consistent with other recent findings that cohort replacement plays a
+larger role than period effects in explaining differences in personal
+culture (Vaisey and Lizardo 2016).
 
@@
 perspective to model personal culture as a linear trajectory\footnote{In
-  supplemental materials, Lersch models age using splines as well, with
-  results generally showing the same conclusions.} over the duration of
-a panel for each respondent. In doing this, Lersch rectified three
-shortcomings of the AUM and SDM. First, the AUM, as described by Kiley
-and Vaisey, posits that changes follow a Markov process where responses
-at time \(t\) are a function of responses at time \(t-1\) but not
-earlier time points. However, earlier life experiences can (directly and
-indirectly) mold personal culture when transitioning to new social roles
-or into new environments, even if their initial impact is delayed. For
-example, childhood events might influence views on family structures
-later when individuals form their own families. Therefore, the LCAM
-considers influences from earlier than just \(t-1\) on responses at time
-\(t\).
+  supplemental materials, Lersch uses splines as well, with results
+  generally showing the same conclusions.} over the duration of a panel
+for each respondent. In doing this, Lersch rectified three shortcomings
+of the AUM and SDM. First, the AUM, as described by Kiley and Vaisey,
+posits that changes follow a Markov process where responses at time
+\(t\) are a function of responses at time \(t-1\), but not earlier time
+points. However, earlier life experiences can (directly and indirectly)
+mold personal culture when transitioning to new social roles or into new
+environments, even if their initial impact is delayed. For example,
+childhood events might influence views on family structures later when
+individuals form their own families. Therefore, the LCAM considers
+earlier influences on responses at time \(t\).
 
@@
 evaluates the LCAM against the AUM and SDM using panel data spanning a
-wider duration (from 3 to 36 years) and more waves (3 to 18), offering a
-better chance to observe durable change.
+wider duration (from 3 to 36 years) and more waves at the respondent
+level (3 to 18), offering a better chance to observe durable change in a
+longer time frame.
 
@@
 can use the LCAM framework to quantify the magnitude of change in a
-measure of culture over time (i.e., how much do people change on
-average?), the variance in age-based slopes over time (i.e., do people
-tend to change in similar ways as they age?), or how much personal
-culture tends to fluctuate around these age-based trends (i.e., how
-consistent are responses over time?).
+measure of culture over time (how much do people change on average?),
+the variance in age-based slopes over time (do people tend to change in
+similar ways as they age?), or how much personal culture tends to
+fluctuate around these age-based trends (how consistent are responses
+over time?).
 
@@
 was evidence of persisting change among adults, a pattern consistent
-with Lersch's results. Also Lersch (2023) found evidence of persisting
-change on most items, but these changes were small. On average, people
-changed only about .07 standard deviations over 10 years. This is
-consistent with Kiley and Vaisey's finding that persisting change is
-often small. Moreover, a quarter of the items Lersch studied still
-preferred the SDM, suggesting that even with different assumptions about
-change and more extensive data, many questions still are consistent with
-a model that allows zero intrapersonal change.
+with Lersch's results. Also, Lersch (2023) found evidence of persisting
+change on most items, but these changes were relatively small. On
+average, people changed only about .07 standard deviations over 10
+years. This is consistent with Kiley and Vaisey's finding that
+persisting change is often small in magnitude. Moreover, a quarter of
+the items Lersch studied still preferred the SDM, suggesting that even
+with different assumptions about change and more extensive data, many
+questions still are consistent with a model that allows zero
+intrapersonal change.
 
@@
 humanities, arts, and social sciences shift peoples' moral relativism
-only about 0.2 standard deviations on average compared to people with no
+only about .20 standard deviations on average compared to people with no
 college degree, and this is the largest effect they identify. And
@@
 and in a large enough sample, researchers will likely observe some
-evidence for durable change in the population. A lack of evidence for
-change may be due to a poor survey question, a low resolution in
-response options, or that the question simply was not asked for long
-enough or to enough people. Conversely, finding evidence for some change
-tells us only that it was ``not zero.'' From that alone, we learn little
-about how much intrapersonal change has happened in a population.
+evidence for change in the population. A lack of evidence for change may
+be due to a poor survey question, a low resolution in response options,
+or that the question simply was not asked for long enough or to enough
+people. Conversely, finding evidence for some change tells us only that
+it was ``not zero.'' From that alone, we learn little about how much
+intrapersonal change has happened in a population.
 
@@
 stable interpersonal differences, reflects the accumulated experiences
-of people prior to entering the panel survey. Lersch (2023: 24) calls
+of people prior to entering the panel survey. Lersch (2023:24) calls
 this ``early imprinting.'' While commonly associated with experiences
@@
 post-retirement, this ``imprinting'' might reflect this pivotal life
-transition. Consequently, this component also reflects variation in
-individuals' social roles or statuses at the start of the panel that
-were important in shaping their dispositions.
+transition and its subsequent effects observed in the panel window.
+Consequently, this component also reflects variation in individuals'
+social roles or statuses at the start of the panel that were important
+in shaping their dispositions.
 
@@
 time. This component captures the set of processes collectively called
-``persistent change'' or ``adaption'' by Lersch and ``active updating''
-by Kiley and Vaisey. Lersch attributes these changes to social triggers
-such as moving into a new environment or adopting new social roles
-(although he does not measure these directly). They might also reflect
-the diffusion of new cultural forms across social networks, cues from
-political elites or otherwise culturally influential leaders, the
-emergence of issues in politics or culture, or large-scale social
-shifts.
+``persistent change'' or ``adaption'' by Lersch (2023) and ``active
+updating'' by Kiley and Vaisey (2020). Lersch (2023) attributes these
+changes to social triggers such as moving into a new environment or
+adopting new social roles (although he does not measure these directly).
+They might also reflect the diffusion of new cultural forms or ideas
+across social networks, cues from political elites or otherwise
+culturally influential leaders, the emergence of issues in politics or
+culture, or large-scale social shifts.
 
@@
 processes of personal culture, it does not directly touch on the ongoing
-debate outlined here. Because of this, we quantify this component but
-focus principally on the other two.
+debate outlined here. Hence, we quantify this component but focus
+principally on the other two.
 
@@
 an important process for why people differ from each other? Only through
-quantifying the two components we can get closer to a true answer to
-this question.
+quantifying the two components we can get closer to a true answer.
 
@@
 reflect the theoretical debate to this point. First, consistent with the
-settled disposition model, it assumes that people start the survey with
-cultural differences, modeled as random intercepts for each respondent.
-Second, it assumes that people change over time, taking the form of
-random slopes for each respondent as a linear function of time. Third,
-it assumes that people deviate around this baseline randomly over time,
-reflecting ``fluctuation'' or short-term non-persistent change.
-Formally, this can be written as
+SDM, it assumes that people start the survey with cultural differences,
+modeled as random intercepts for each respondent. Second, it assumes
+that people change over time, taking the form of random slopes for each
+respondent as a linear function of time. Third, it assumes that people
+deviate around this baseline randomly over time, reflecting
+``fluctuation'' or short-term non-persistent change. Formally, this can
+be written as
 
@@
 component for individuals, and \(\epsilon_{it}\) is the random error
-term that captures transitory disturbances. The \(\alpha_{0i}\) and
-\(\alpha_{1i}\) terms are also allowed to covary.
+term that captures fluctuations. The \(\alpha_{0i}\) and \(\alpha_{1i}\)
+terms are also allowed to covary.
 
@@
 
-This effectively models respondents' culture under the assumption that
+This effectively models respondents' responses under the assumption that
 they make no systematic change over time, or that the expected value at
-any time point is the same as the expected value at any other. This
-gives us the proportion of total variance in an item attributable to
-stable interpersonal differences.
+any time point is the same as the expected value at any other.
+Therefore, this measure---V(D)---gives us the proportion of total
+variance in an item attributable to stable interpersonal differences.
 
@@
 This is the incremental proportion of variance accounted for when we
-allow the model predictions to change over time for each person.
+allow the model predictions to change over time for each person,
+allowing us to quantify the explanatory power of change.
 
@@
 the outcome that the LCAM can attribute to either differences between
-people or linear intrapersonal change over the course of the panel. Our
-principal measure of interest is the proportion of systematic variance
-accounted for by intrapersonal change, which we calculate as
+people or linear intrapersonal change\footnote{The decomposition of
+  variance is agnostic to this choice. One advantage is that the
+  modeling assumptions can be relaxed to accommodate non-linear terms or
+  splines, but the decomposition will follow the same procedure.} over
+the course of the panel. Our principal measure of interest is the
+proportion of systematic variance accounted for by intrapersonal change,
+which we calculate as
 
@@
 perturbed sample, and calculate the \(\omega\). We then replicate this
-procedure \(N\) times and establish confidence intervals using the
-distribution of bootstrapped omega values.
+procedure \(N\) times and establish confidence intervals around our
+measure.
 
@@
 
-To demonstrate the utility of this metric, we provide two examples. In
+To demonstrate the utility of \(\omega\), we provide two examples. In
 the first, we compare the proportion of systematic variance explained by
@@
 
-\subsection{Data}\label{data}
+\subsection{Data and Analysis}\label{data-and-analysis}
 
@@
 work (Kiley and Vaisey 2020; Lersch 2023).\footnote{For more information
-  on these data sources, see (Goebel et al. 2019; Income Dynamics 2013;
-  Smith et al. 2022; Summerfield et al. 2011; Taylor 1996; University of
-  Essex and Research 2019; Voorpostel et al. 2016).} These studies cover
-a long period of time (the range of surveys spans from 1968 to 2021),
-with 609 personal culture items capturing attitudes, beliefs, values,
-self-assessments, self-descriptions, and behaviors (Alwin 2007). We
-restricted the sample such that individuals between the ages of 18 and
-79 are included without further elimination, and in all surveys, we used
-all possible cases for which respondents provided responses. In the end,
-the analyses that follow rely on a cross-national sample with a
-cross-domain set of items to capture a broad range of individual
-personal culture. Supplemental Materials A documents the list of all
-variables used in the upcoming analyses.
+  on these data sources, see the relevant panel files (Goebel et al.
+  2019; Income Dynamics 2013; Smith et al. 2022; Summerfield et al.
+  2011; Taylor 1996; University of Essex and Research 2019; Voorpostel
+  et al. 2016).} These studies cover a long period of time (the range of
+surveys spans from 1968 to 2021), with 610 personal culture items
+capturing attitudes, beliefs, values, self-assessments,
+self-descriptions, and behaviors (Alwin 2007). We restricted the sample
+such that individuals between the ages of 18 and 79 are included without
+further elimination, and in all surveys, we used all possible cases for
+which respondents provided responses. In the end, the analyses that
+follow rely on a cross-national sample with a cross-domain set of items
+to capture a broad range of individual personal culture. We document the
+list of all variables used in the upcoming analyses in the Appendix.
 
@@
 Switzerland & Swiss Household Panel (SHP) & 1999-2019 & 77\\
-\addlinespace
 United States & General Social Survey (GSS) & 2006-2012 & 183\\
@@
 
-\subsection{Analysis}\label{analysis}
-
-As a first step in our analysis, we fit this LCAM to each of the 609
+As a first step in our analysis, we fit this LCAM to each of the 610
 measures of personal culture outlined above. We then derive \(\omega\)
-for each of these items. As a second step, we use linear regression to
-model \(\omega\) as a function of features of the question and panel.
-This allows us to explore how these features are associated with the
-degree of intrapersonal change. These covariates include the number of
-response options (measured as 2, 3-5, and more than 5), the specific
-survey, the log of the number of participants, the number of waves
-observed, the time period (i.e., decade) the question was first asked,
-and the total duration of time the question was asked.
+for each of these items using the procedure above.
 
@@
 Figure 1 plots the proportion of systematic variance attributable to
-interpersonal differences and intrapersonal change for each of the 609
-questions, plotted separately by panel. To broadly summarize the
+interpersonal differences and intrapersonal change for each of the 610
+questions, plotted separately by panel survey. To broadly summarize the
 results, all panels show a similar range of the proportion of systematic
 variance attributable to intrapersonal change. Across panels, the median
-value of \(\omega\) is 0.094 (mean of 0.098), with an interquartile
-range of 0.053 to 0.139. All panels include questions where \(\omega\)
-is essentially 0, and all panels include questions with \(\omega\)
-values greater than 0.20. The maximum \(\omega\) value is 0.26.
-
-\begin{figure}[ht]
+value of \(\omega\) is .053 (mean of .056), with an interquartile range
+of .032 to .077. All panels include questions where \(\omega\) is
+essentially 0, and some panels include questions with \(\omega\) values
+greater than .20. The maximum \(\omega\) value is .23.
+
+\begin{figure}[!p]
 \begin{center}
-\caption{Proportions of Systematic Variance in Personal Culture}
 
@@
 \end{center}
+\caption*{Figure 1: Proportions of Systematic Variance in Personal Culture}
 \footnotesize{\textit{Notes:} The figure shows $\omega$ and 1 - $\omega$ as the proportion of systematic variance attributable to intrapersonal change and interpersonal differences.
-See Supplemental Materials A for the full set of item values.}
+See the Appendix for the full set of item values.}
 \end{figure}
 
-Across all questions, interpersonal differences account for a much
-larger share of the systematic variance in responses than intrapersonal
-change. Again, this is to be expected. Interpersonal differences capture
-not just pre-adult socialization, but all accumulated experiences up to
-the start of the panel that might influence personal culture.
+Across all questions and all surveys, interpersonal differences account
+for a much larger share of the systematic variance in responses than
+intrapersonal change. Again, this is to be expected. Interpersonal
+differences capture not just pre-adult socialization, but all
+accumulated experiences up to the start of the panel that might
+influence personal culture.
 
 To the extent that there are differences across the panels, the PSID has
-the highest \(\omega\) values with mean .120 and median .135. While we
+the highest \(\omega\) values with mean .068 and median .076. While we
 cannot disentangle features of the sample from features of the questions
@@
 the distinct age profile of respondents in this sample. At the other
-end, the GSS has the lowest range of \(\omega\) with mean .073 and
-median .069. This potentially reflects the fact that the GSS observes
-people for a shorter duration, on average, than the other panels. If,
-consistent with life course adaption theories, people are more likely to
-make significant cultural changes the longer we observe them, then
-duration likely affects the range of \(\omega\) (a point we explore more
-below). However, the GSS results are still consistent with results from
-the other panels.
+end, the SHP has the lowest range with mean .042 and median .037,
+followed by GSS, with mean .044 and median .043.
 
 While there are some differences between panels, these differences are
-small compared to the differences within panels. For about 6 percent of
-items, \(\omega\) is greater than 0.20. These questions tend to ask
-about objectively changing external referents (e.g., confidence in
-specific government leaders or political parties), life satisfaction, or
-current financial position. At the other end, questions about religious
+small compared to the differences within panels. For about 9 percent of
+items, \(\omega\) is greater than .10. These questions tend to ask about
+objectively changing external referents (e.g., confidence in specific
+government leaders or political parties), life satisfaction, or current
+financial position. At the other end, questions about religious
 identification, views on gender roles, and support for civil liberties
-tend to have very low estimates of intrapersonal change.
+tend to have very low estimates of intrapersonal change, suggesting that
+core preferences might be less variable.
 
@@
 updating. Our results show that intrapersonal change is much more
-important for explaining variance in confidence in the press (0.164)
-than confidence in religion (0.049), even though both are updating.
-
-Appendix A shows the distribution of \(V(D)\) and \(V(C)\) across
-panels, and Supplemental Materials A presents the estimated proportion
-of variance attributable to interpersonal differences and intrapersonal
-change, the estimated values of \(\omega\), and the proportion of
-residual variance for each question. Interpersonal differences are
-almost always the largest component of the total variance and tend to
-account for between 55 and 70 percent of total variance, while
-intrapersonal change is always the smallest, typically accounting for
-between 3 and 8 percent of total variance. Residual variance tends to
-account for between 22 and 37 percent of variance, though on several
-questions residual variance is greater than 50 percent. This might
-indicate survey items with low reliability or ones that capture
-genuinely rapid fluctuations.
-
-As we noted above, the substantive importance of a given amount of
-intrapersonal change depends on a range of factors, including how long
-the panel runs, whether assumptions about linear change hold, and
-whether the panel is capturing a distinctly turbulent period or a
-distinctly stable period for the relevant item. However, if we assume
-that the period under observation is ``typical'' for a question---not a
-time of extremely heightened (or lowered) sensitivity or change---then
-it does not seem realistic that intrapersonal change accounts for a
-large share of the cultural differences we see in the world.
-
-Figure 2 plots the results from a linear regression of \(\omega\) as a
-function of question, panel, and sample features. These models also
-include fixed effects for panels and topics, so coefficients reflect the
-association within a panel and topical domain.
-
-\begin{center}
-<Figure 2 about here>
-\end{center}
-
-Figure 2 shows that the more response options respondents are given and
-the larger the sample, the larger the values of \(\omega\). We interpret
-these coefficients as suggesting greater resolution on a question makes
-it easier to detect and model change. The earlier a question was asked,
-even net of how long it was been asked, the lower \(\omega\).
-
-Although we did not state expectations for how question content would
-relate to \(\omega\), some of the associations between question
-structure, panel duration, and panels themselves might be driven by
-differences in the topics addressed by each panel. To address this, we
-followed Hout et al. (2016) and Lersch (2023) in coding each question as
-falling into one of nine different topical domains and included these as
-indicator variables in the meta analysis. Supplemental Materials B
-presents the estimates for these topic indicators. They show some
-associations with \(\omega\), with questions about subjective SES;
-social life, social cohesion, and trust; environment and climate; and
-health and morale showing larger \(\omega\) values than questions about
-religion and spirituality; politics, government and the economy; and
-gender and family.
-
-The most notable result in the meta analysis is the negative coefficient
-attached to the duration of years covered by the question on average
-across participants. The longer a question is observed, the less is the
-value of \(\omega\). Theories of personal cultural change that link
-changes in personal culture to social experiences, including the LCAM,
-suggest that the longer we observe respondents, the more likely people
-are to undergo potentially transformative experiences and therefore the
-more variance would be attributable to intrapersonal change. Finding a
-negative coefficient here seems to challenge that assumption.
-
-As a further test of this finding, we compared the \(\omega\) values
-when using the full duration of a panel compared to when we dropped the
-final wave for each participant and therefore reduced the total duration
-of observation for the question. If the coefficient reflects a true
-negative effect of duration on \(\omega\), we should see that same
-effect within questions. Results from this analysis are presented in
-Appendix B and contradict the coefficient from the regression model;
-with a few exceptions, we found that the longer we observe the same
-question, the higher \(\omega\) is. We interpret this combination of
-findings as suggesting that the kinds of questions asked for a longer
-time period tend to demonstrate less intrapersonal change than questions
-asked over shorter periods, rather than a true function of time. This
-supports the conclusion that the GSS shows less intrapersonal change
-because of its shorter duration.
+important for explaining variance in confidence in the press (0.091)
+than confidence in religion (0.051), even though in both cases people
+are updating their beliefs.
+
+In the Appendix, we show the distribution of \(V(D)\), \(V(C)\) and the
+residual variance \(V(E)\) for each question across panels.
+Interpersonal differences are almost always the largest component of the
+total variance and tend to account for between 50 and 66 percent of
+total variance, while intrapersonal change is always the smallest,
+typically accounting for between 2 and 5 percent of total variance.
+Residual variance tends to account for between 30 and 46 percent of
+variance, though on several questions residual variance is greater than
+50 percent. This might indicate survey items with low reliability or
+ones that capture genuinely rapid fluctuations.
 
@@
 
-\subsection{Data}\label{data-1}
+\subsection{Data}\label{data}
 
@@
 
-\begin{figure}[ht]
-\begin{center}
-\caption{Difference in $\omega$ Across College Graduates and Non-Graduates}
-
-\begin{center}\includegraphics[width=62.5in]{../../figures/colleges} \end{center}
-
-\end{center}
-\footnotesize{\textit{Notes:} The figure shows the difference in $\omega$ values across college graduates and college non-graduates.
-Values above (below) 0 means that those with college degree have higher (lower) variance of intrapersonal change.
-The dashed red line marks 0 difference.}
-\end{figure}
-
 There is a clear pattern in Figure 3: for more than 80 percent of these
@@
 these eight questions, for both education groups, in Figure 4.
-
-\begin{center}
-<Figure 4 about here>
-\end{center}
 
@@
 The primary objective of this paper is to introduce a new quantity that
-can advance debate on whether people change their personal culture --
-their attitudes, beliefs, values, and practices -- as they move through
-their adult life. Instead of falling back on a ``tournament of models''
-approach (Lersch 2023: p.~228) to proclaim a yes or no answer to this
-question, or only quantifying individual components of cultural
+can advance debate on whether people change their personal
+culture---their attitudes, beliefs, values, and practices---as they move
+through their adult life. Instead of falling back on a ``tournament of
+models'' approach (Lersch 2023:228) to proclaim a yes or no answer to
+this question, or only quantifying individual components of cultural
 difference to show how much people change, our metric contextualizes the
@@
 We demonstrated the utility of this measure in two ways. First, applying
-our proposed measure, \(\omega\), to 609 survey items from the panel
+our proposed measure, \(\omega\), to 610 survey items from the panel
 datasets previously studied in this debate revealed a consistent
@@
 durable, intrapersonal change over time. Some show notably high amounts
-of intrapersonal change. For some questions about life satisfaction or
-views on government benefits, it seems plausible that differences in
-adult experiences predominantly account for observed differences between
-people. However, intrapersonal change is often substantially less
-pronounced than interpersonal differences, accounting for less than 10
-percent of systematic variance on average across all datasets. On
-questions related to civil liberties, abortion, generalized trust, and
-civic duty, systematic variance attributable to intrapersonal change is
-essentially zero. For these questions, it seems there is not enough
-cultural change during adulthood to warrant attributing the differences
-we observe to experiences and social transitions; instead, the primary
-source of the observed differences appears to stem from experiences in
-childhood, adolescence, and early adulthood.
+of intrapersonal change. However, intrapersonal change is often
+substantially less pronounced than interpersonal differences, accounting
+for less than 10 percent of systematic variance on average across all
+datasets. On some questions, systematic variance attributable to
+intrapersonal change is essentially zero. For these questions, it seems
+there is not enough cultural change during adulthood to warrant
+attributing the differences we observe to experiences and social
+transitions; instead, the primary source of the observed differences
+appears to stem from experiences in childhood, adolescence, and early
+adulthood.
 
@@
 comparison across questions and it provides a benchmark for researchers
-exploring new questions. Second, our metric can help contextualize the
+exploring new questions. Second, our metric might help contextualize the
 quantities identified in the LCAM. Is .07 standard deviations of change
-every 10 years a lot of change? For some questions, our metric will say
-this change explains a large component of variance, and for others it
-will not.
-
-In a second application of the quantity, we showed how it can open up
-new lines of inquiry. We showed how the amount of cultural difference
-explained by intrapersonal change varies substantially not just by
-survey item but also by individuals' characteristics such as their
-education. While age and other factors related to college completion
-might confound this pattern, it suggests that college completion may
-crystallize personal culture to an extent that renders later adult
-experiences on political dispositions less influential in explaining
-overall differences.
+every 10 years a lot of change? For some questions or groups, our metric
+will say this change explains a large component of variance, and for
+others it will not.
+
+In a second application of our metric, we showed how an analysis of
+\(\omega\) can open up new lines of inquiry. We showed how the amount of
+cultural difference explained by intrapersonal change varies
+substantially not just by survey item but also by individuals'
+characteristics such as their education. While age and other factors
+related to college completion might confound this pattern, it suggests
+that college completion may crystallize personal culture to an extent
+that renders later adult experiences on political dispositions less
+influential in explaining overall differences.
 
@@
 while we model change in a single way, the general variance
-decomposition approach we outline accommodates different definitions and
-functional forms of change.
+decomposition approach we outline accommodates different functional
+forms.
 
@@
 one's personal culture. Our results reinforce this point, showing the
-relevance of both factors and allowing their precise quantification.
+relevance of both factors and allowing their precise quantification for
+the explanation of cultural differences.
 
@@
 that adults do encounter opportunities, necessities, and incentives for
-change, but their ability or willingness to change decreases. All the
-more it is important to research when and how social situations can
-provoke durable change in adults.
+change, but their ability or willingness to change decreases as a
+function of age. All the more it is important to research when and how
+social situations can provoke durable change in adults.
 
@@
 suggest that the importance of processes that lead to such change can
-vary by group. Aligning with life course theories, it seems the
-significance of experiences for cultural change may be contingent on
-other, prior experiences. For example, factors that shape cultural
-dispositions may likely differ for college and non-college graduates;
-the latter might be more profoundly influenced by mid-life experiences
-than the former in this regard. In trying to understand differences in
-personal culture, sociologists should therefore pay more attention to
-the heterogeneous effects that various factors including social events,
-encounters, and situations can have.
+vary by group. Aligning with life course theories, the significance of
+experiences for cultural change may be contingent on other, prior
+experiences. For example, factors that shape cultural dispositions may
+likely differ for college and non-college graduates; the latter might be
+more profoundly influenced by mid-life experiences than the former in
+this regard. In trying to understand cultural differences, sociologists
+should pay more attention to the heterogeneous effects that various
+factors including social events and encounters can have.
 
@@
 the two is therefore crucial for understanding cultural differences and
-cannot be done with cross-sectional data. Instead we should extend
-panels beyond two waves to gain more leverage to understand when and in
-what form personal culture changes.
-
-A key insight from our meta-analysis is that an increase in response
-resolution (the range of response options survey respondents are given)
-correlates with an increased share of systematic variance explained by
-intrapersonal change. This may reflect specific issues being assessed by
-survey items using different scales. Alternatively, it might imply that
-when intrapersonal changes occurs, they are often subtle with answers
-moving from ``agree'' to ``strongly agree,'' rather than from ``strongly
-agree'' to ``strongly disagree.'' Such subtle changes require finer
-response options to be detected.
+cannot be done with cross-sectional data. Instead, we should extend
+panels beyond two waves and longer time periods to gain more leverage to
+understand when and in what form personal culture changes.
 
@@
 might not be able to help explain major cultural differences in adult
-populations.
+populations, and the reasons why people believe in some things but not
+others.
 
@@
 in on a broader range of theoretical debates. We hope researchers find
-our approach useful as they investigate these questions.
+our approach useful as they investigate these questions across a variety
+of problems.
 
@@
 
-\bibitem[\citeproctext]{ref-hout2016}
-Hout, Michael, and Orestes P. Hastings. 2016. {``Reliability of the
-{Core} {Items} in the {General} {Social} {Survey}: {Estimates} from the
-{Three}-{Wave} {Panels}, 2006--2014.''} \emph{Sociological Science}
-3:971--1002. doi:
-\href{https://doi.org/10.15195/v3.a43}{10.15195/v3.a43}.
-
 \bibitem[\citeproctext]{ref-data_psid}
@@
 
-\bibitem[\citeproctext]{ref-visser2004}
-Visser, Penny S., and Robert R. Mirabile. 2004. {``Attitudes in the
-Social Context: {The} Impact of Social Network Composition on
-Individual-Level Attitude Strength.''} \emph{Journal of Personality and
-Social Psychology} 779--95.
-
 \bibitem[\citeproctext]{ref-data_shp}

### 2024-09-13 drawer/tk_code/simulation.R

@@ -1,5 +1,5 @@
 
-# - quantifying change ------------------------------------------------------- #
+# - quantifying the importance of change ------------------------------------- #
 # - simulations -------------------------------------------------------------- #
 
 ### note: this script
@@ -97,7 +97,7 @@ d1 <- d1 |>
 # --- STEP 4: DRAW THE RESULTS
 
 png(
-  "./figures/simulation01.png",
+  "./figures/simulation.png",
   w = 8,
   h = 4,
   units = "in",

### 2024-09-13 scripts/03_educ.R

@@ -1,190 +1,0 @@
-
-# - quantifying change ------------------------------------------------------- #
-# - college example ---------------------------------------------------------- #
-
-### note: this script
-###       (a) prepares 8 political items from the General Social Survey,
-###       (b) calculates omega values across college and no-college subgroups,
-###       (c) compares the differences of omega values in booted samples.
-
-# ---------------------------------------------------------------------------- #
-
-rm(list = ls())
-pacman::p_load(
-  purrr,
-  furrr, 
-  hrbrthemes,
-  conflicted,
-  gssr,
-  tidyverse)
-conflicts_prefer(dplyr::select())
-conflicts_prefer(dplyr::filter())
-
-# --- helpers
-my_oka <-
-  c("#eebd64",
-    "#8b3b36",
-    "#35455d",
-    "#144525")
-theme_set(theme_ipsum_rc())
-`%nin%` = Negate(`%in%`)
-
-# --- scripts
-suppressMessages(
-  source("./scripts/helper_scripts/omega.R"))
-
-# --- data
-data("gss_panel06_long")
-data("gss_panel08_long")
-data("gss_panel10_long")
-
-# --- parallel sessions
-plan(multisession); 
-set.seed(11235) # set seed for simulation protocol
-options(future.globals.maxSize=5000*1024^2)
-
-# ---------------------------------------------------------------------------- #
-# PART 1: DATA PREPS --------------------------------------------------------- #
-# ---------------------------------------------------------------------------- #
-
-# --- get the data
-gss <- bind_rows(
-  gss_panel06_long |>
-    select(
-      firstid,
-      wave,
-      age,
-      degree,
-      partyid,
-      polviews,
-      eqwlth,
-      helpsick,
-      helpnot,
-      helpblk,
-      helppoor,
-      affrmact
-    ) |>
-    mutate(pid = paste("06", firstid, sep = "-")),
-  gss_panel08_long |>
-    select(
-      firstid,
-      wave,
-      age,
-      degree,
-      partyid,
-      polviews,
-      eqwlth,
-      helpsick,
-      helpnot,
-      helpblk,
-      helppoor,
-      affrmact
-    ) |>
-    mutate(pid = paste("08", firstid, sep = "-")),
-  gss_panel10_long |>
-    select(
-      firstid,
-      wave,
-      age,
-      degree,
-      partyid,
-      polviews,
-      eqwlth,
-      helpsick,
-      helpnot,
[…]

### 2024-09-13 scripts/cross_items.R

@@ -1,5 +1,5 @@
 
-# - quantifying change ------------------------------------------------------- #
+# - quantifying the importance of change ------------------------------------- #
 # - omega calculations ------------------------------------------------------- #
 
 ### note: this script
@@ -30,14 +30,13 @@ theme_set(theme_ipsum_rc())
 
 # --- scripts
 suppressMessages(
-  source("./scripts/helper_scripts/omega.R"))
+  source("./scripts/omega.R"))
 
 # --- data
 d <- readRDS("./data/data.rds") # nested survey x item panels
 
 # --- parallel sessions
-plan(multisession); 
-set.seed(11235) # set seed for simulation protocol
+plan(multisession)
 options(future.globals.maxSize=5000*1024^2)
 
 # ---------------------------------------------------------------------------- #
